### 2023-01-04 docs/cogsci/PNTKN.Rmd
cogsci submission

---
title: "Using pretense behavior to explore counterfactual self-simulation"
bibliography: r-references.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of london \\ London, WC1E 7HX UK
    \AND {\large \bf Chaz Firestone (chaz@jhu.edu)}
    \AND {\large \bf Ian Phillips (ianphillips@jhu.edu)} \\ Johns Hopkins University \\ Baltimore, MD 21218, USA}

abstract: >
   What we do depends on what we know. But sometimes we try to decouple our behavior from our knowledge so that we appear not to know what we really do know. Such pretense behavior requires understanding how we would behave with different knowledge — and so provides a window into counterfactual self-simulation. However, little research has characterized and evaluated pretense relative to non-pretense behavior. Here, in a large-scale, pre-registered experiment, subjects played normal games of Battleships (trying to sink ships hidden in a grid), as well as ‘pretend’ games, where they were told all the ships’ locations but had to pretend they were playing normally.  Relative to normal games, ‘pretend’ games demonstrated similar but exaggerated behavioral patterns. Furthermore, pretenders played rationally, but less so than non-pretenders. Despite these differences, ‘judge’ participants could not detect ‘pretend’ games. We discuss implications of these findings for accounts of theory of mind and metacognition.

keywords: >
    metacognition; theory of mind; pretense

output: cogsci2016::cogsci_paper

final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries, echo=FALSE, warning=FALSE}

library(groundhog)

groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret' #for cross validation
), '2022-12-01')

```

# Introduction

The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory]. This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation]. This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own.

Moreover, deception often involves pretense behavior, which in turn relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state. For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") — you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first? What would I say? How long would it take me to recover from the surprise?"). This reliance of pretense behavior on self-simulation makes it an ideal opportunity to examine metacognitive knowledge about one's own mental states, and the potential reliance of this knowledge on a self-simulation. By comparing non-pretend and pretend behavior, we can ask which aspects of their own cognitive processes subjects can and cannot effectively simulate, and which aspects are not well-represented in their mental models of their own cognition.

To this end, here we examine pretense in a game setting. Using an online version of the game Battleships, participants played a 'non-pretend' (normal) version of the game (trying to sink ships whose locations are unknown), as well as a 'pretend' version where they were given secret information about the location of hidden ships, but were instructed to behave as if they didn't have this information.

# Methods

A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb)

```{r load-data, warning=FALSE}
E2.df <- read.csv('../../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,4),sep='')) %>%
  mutate(subj_id = factor(subj_id)) %>%
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
  mutate(genuine_first=genuine_first=='True');

E2.export <- read.csv('../../experiments/Battleships2/data/batch1/prolific_export_batch1.csv',na.strings=c(""," ","NA"));

E2.bonus <- read.csv('../../experiments/Battleships2/data/batch1/bonus.csv',na.strings=c(""," ","NA"))

```

## Participants

The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University. 500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation. They were selected based on their acceptance rate (>95%) and for being native English speakers. The entire experiment took approximately 20 minutes to complete. Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).

## Procedure

Participants were first instructed that the experiment, based on the game Battleships, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment. They were then presented with a leaderboard of previous players, and given the rules of the game:

> "In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
What makes the game difficult is that you can’t see the ships; all you can see is a grid of squares, and you have to guess where the ships are. To sink a ship, you need to click on all of the squares it is located in. If you hit part of a ship, the square will turn red. If there is no ship in the square, it will turn blue."

We further explained that in this version of the game, ships can touch corners, but their sides can't touch. This explanation was accompanied by a visual presentation of legal and illegal ship configurations.

After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half game (see below for details). The order of pretend and non-pretend blocks was counterbalanced between participants. The allocation of boards (spatial configurations of ships; see Fig. 1A) to conditions was randomized between participants such that exactly one board was played in both pretend and non-pretend conditions, and this common board was different for different participants. The order of boards within a block was fully randomized, with the exception that half-games were always played last.

### Non-pretend (normal) games

In non-pretend games (Fig. 1B), participants aimed to sink two 2-square patrol boats and one 3-square submarine with as few clicks as possible. An online counter of the number of clicks was displayed on the screen. After each game, feedback was given about the number of clicks and resulting number of points obtained.

### Pretend games

Participants in pretend games were given the same explanation of Battleships, and played a practice round. However, they were then given an additional twist:

> "This time your goal is different. In this round, we’re going to tell you where the ships are, but **we want you to act like you don’t know this information**. We’ve marked the ships’ locations with a cross, so you’ll know where they are the whole time; but your job is to play the game as if these hints aren’t there. To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are. We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you'll get bonus points. But if your games look different, you won't get these bonus points. Your number of clicks in this part will not affect your bonus. Only your ability to play like you had no hints."

We informed participants that both the location and timing of their cell clicks will be measured. After one practice round and one comprehension question, participants played five pretend games (Fig. 1C), followed by one pretend half-game. Each game was followed by a short message, reminding them that a game that looks similar to the games of participants who had no hints would be awarded 10 bonus points.

### Half games

In order to directly compare participants' pretend and non-pretend games for identical belief states (genuine or pretended ignorance about where the ships are hidden), participants completed one pretend and one non-pretend game given a partly finished board with the content of 7 cells already revealed (see Fig. 1D). We designed our half games to produce a strong expectation to find a ship in specific cells, but not in others. The assignment of half-completed boards to pretend and non-pretend conditions was randomized between participants.

### Judge trials

In the final part of the experiment, participants observed the games of previous players and tried to determine who had hints and who didn't. On each trial, two empty grids were presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid (Fig. 1F). The two grids corresponded to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders. Only games shorter than one minute were chosen for presentation in this part. For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part. Judge participants observed a real time replay of the two grids, showing not only where participants clicked, but also when. After making a decision, participants were informed whether they would receive the 10 bonus points, or alternatively, whether the pretender would receive them in the event the pretender managed to trick them.

<!-- 500 players played 12 games of the game Battleships in two conditions presented in two blocks. The order of conditions was randomized between players. the standard, *non-pretend* condition, players started the game not knowing where the ships were and tried to sink all ships with as few clicks as possible. In the *pretend* condition, ship locations were marked with a cross such that players knew where they were hidden, and their task was to play the game as if they did not have this information. The sixth game in each block was a 'half-game', where participants completed the game with the content of some cells revealed by a previous player. Lastly, in judge trials, players observed a replay of the games of two previous players – one pretender and one non-pretender, and their task was to identify which was which. We created nine different 5x5 boards, each containing two 2-square patrol boats and one 3-square submarine. The assignment of boards to conditions was randomized across participants such that exactly one board was presented in both pretend and non-pretend blocks.  (see Fig. 1). -->

A more detailed description of the study procedure is provided in the study pre-registration document (LINK HIDDEN). Readers are also invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/NervzpM0Y0z).

```{r design, echo=FALSE, fig.cap="Experimental Design. See Methods for details.", out.width = '100%'}
knitr::include_graphics("../figures/Battleships_design.png")
```

# Results

```{r subject-comments, echo=FALSE}
E2.comments <- E2.df %>%
  filter(!is.na(responses) & is.na(test_part)) %>%
  dplyr::select(subj_id,responses) %>%
  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
  unnest(comments)%>%
  dplyr::select(subj_id,worker_comments)%>%
  filter(!is.na(worker_comments) & worker_comments!="") %>%
  mutate(worker_comments = gsub("[\r\n]", "",worker_comments))

E2.debrief <- E2.df %>%
  filter(!is.na(responses) & is.na(test_part)) %>%
  dplyr::select(subj_id,responses) %>%
  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
  unnest(comments)%>%
  dplyr::select(subj_id,debrief)%>%
  filter(!is.na(debrief) & debrief != "")%>%
  mutate(debrief = gsub("[\r\n]", "",debrief))

```

We designed our analyses to explore subjects' capacity for self-simulation under a counterfactual knowledge state, and the limits of this capacity. We focused on where subjects clicked and when, and asked whether this differed between pretend and non-pretend games. All analyses were pre-registered unless otherwise specified. In our pre-registration document, we committed to separately analyzing participants according to whether they pretended before or after completing a non-pretend block. Due to space limitations, and since the results of the two groups mostly agreed, we report here the pooled results from both groups of participants, and mention whenever we found different results depending on block order. When directly comparing pretend and non-pretend blocks, we perform a between-subjects comparison using data from the first block only, i.e., pretend games from participants who pretended in the first block and non-pretend games from participants who played normally in the first block. We do so to ensure that any successful pretending is not due to memory of one's own behavior in a previous block, and that non-pretend games are not biased by experience with the pretend block.

```{r num_clicks, echo=FALSE, cache=TRUE}

E2.click_df <- E2.df %>%
  dplyr::select(subj_id, 
         test_part, 
         grid_number, 
         num_clicks,
         click_log,
         genuine_first) %>%
  mutate(click_log = gsub("\'","\"", click_log)) %>%
  filter(test_part=='pretend' | test_part=='nonpretend');

E2.median_clicks <- E2.click_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(test_part,num_clicks);

E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();

E2.median_clicks_filtered <- E2.click_df %>%
  filter(!(subj_id %in% E2.lucky_pretenders))%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(test_part,num_clicks);

p<- E2.click_df %>%
    group_by(subj_id,test_part, genuine_first) %>%
    summarise(num_clicks=median(num_clicks)) %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=num_clicks,fill=test_part))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()
    # geom_vline(xintercept= median_simulated)+
    # facet_wrap(~genuine_first, nrow=2);

ggsave('../docs/figures/E2_num_clicks.png',p,width=5,height=2, dpi=300)

```

```{r useful_functions, echo=FALSE, cache=TRUE}

all_possible_boards <- read.csv('../../analysis/flat_boards.csv',header=FALSE);

create_board_states <- function(flat_positions,hit) {

  board_state = rep(NA,25)
  board_states = c();
  for (p in seq_along(flat_positions)) {
    board_states = c(board_states, paste(board_state,collapse=','));
    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
  }

  return(board_states)
};

get_likelihood <- function(board_state) {

  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);

  revealed_positions <- which(!is.na(board_state));

  boards_to_remove = c();

  for (position in revealed_positions) {
    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
  }

  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
  likelihood[revealed_positions]=NA;

  return(paste(likelihood,collapse=','))

};

get_posterior <- function(likelihood) {
    likelihood = scan(text= likelihood, what = numeric(), sep="," , quiet = TRUE);
    posterior = likelihood/sum(likelihood,na.rm=TRUE); # assuming a uniform prior
    return(paste(posterior,collapse=','))
}

get_p_click <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(posterior[flat_position])
}

get_p_click_rank <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  rank_posterior = rank(-posterior)
  return(rank_posterior[flat_position])
}

get_posterior_entropy <- function(posterior) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  posterior[which(posterior==0)]=NA; #to avoid infinity*0
  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
  return(entropy);
}

```

```{r simulate_nclicks, echo=FALSE, cache=TRUE}

## simulate number of clicks for random play

num_clicks_sim = c();
nsim = 10000
boards = rep(NaN,25)

for (i in 1:nsim) {
  board = rep(c(0),25);
  while (sum(board)<7) {
    board[sample(1:25,1)]=1
  }
  nc = which(board==1)%>%tail(n=1);
  num_clicks_sim = c(num_clicks_sim, nc);
  if (nc<25) {
    board[(nc+1):length(board)]=NaN;
  }
  boards = rbind(boards,board)
}

simulated_hits <- data.frame(click_number=1:25,mean_hit=colMeans(boards,na.rm=T))

## OPTIMAL PLAYER

simulate_optimal_game <- function(ship_positions) {
  #initialize
  board_state <- rep(NA,25); 
  nhits <- 0;
  nclicks <- 0;
  nships <- sum(ship_positions);

  while (nships>nhits) {

    posterior <- get_likelihood(paste(board_state,collapse=',')) %>%
      get_posterior() %>%
      scan(text = ., what= numeric(), sep="," , quiet = TRUE);

    best_shot = which.max(posterior);

    #click on best shot
    nclicks = nclicks+1;
    result = ship_positions[best_shot]
    if (result==1) {
      nhits = nhits+1;
    }
    board_state[best_shot]=result;

  }

  return(nclicks)

}

num_clicks_optimal_sim = c();
nsim = 100;
nboards = all_possible_boards%>%nrow();

for (i in 1:nsim) {
  cur_board = all_possible_boards[sample(1:nboards,1),];
  num_clicks_optimal_sim = c(num_clicks_optimal_sim,simulate_optimal_game(cur_board));
}

```

## Total number of clicks

To sink all ships, players had to click on at least 7 and at most 25 squares. A simulated player that clicks randomly had a median total click number of `r num_clicks_sim%>%median()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a median total click number of `r num_clicks_optimal_sim%>%median()`. Among our players, the median number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games (see Fig. 2A).

We observed no significant difference in the number of clicks between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend),E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`). However, a significant interaction with block order suggested that participants made fewer clicks in the second block compared to the first, regardless of which block was the pretend block (`r E2.median_clicks%>%gather('condition','nclicks',c('nonpretend','pretend'))%>%aov(nclicks ~ genuine_first * condition,data=.)%>%apa_print()%>%'$'(statistic)%>%'$'(genuine_first_condition)`).

In `r E2.click_df%>%filter(num_clicks==7)%>%nrow()` pretend games from `r E2.lucky_pretenders%>%length()` players, games were completed after 7 clicks only, without ever missing a ship. This never happened in non-pretend games. We assumed that these participants did not follow task instructions, and excluded them from all following analyses.

## Game optimality

Pretend games were similar to non pretend games in the total number of clicks — but were they also similar in *where* participants clicked? More specifically, did cell selections in pretend games make sense given the limited information those participants pretended to have? To ask this, we approximated optimal behavior by calculating the probability that a ship is hidden in each cell given available information, $p(ship(x_{i}))$, and the posterior probability that one should click on a square, assuming a uniform prior over cells $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$. Critically, in modeling pretend games we did not treat hints as part of this available information for extracting $p(ship(x_{i}))$, because an optimal player should ignore hints in choosing where to click next. Given this posterior map, a rational player should choose cells where $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3]. To quantify optimality, before each cell selection we computed the posterior probability map for all 'unknown' cells. Then, we ranked cells from high to low according to their posterior probability and recorded the rank of the chosen cell: a lower rank indicating more optimal behavior.

```{r parse-click-log, echo-FALSE, cache=TRUE}

E2.click_log <- readRDS('../../experiments/Battleships2/data/batch1/click_log.Rda');

# E2.click_log <- data.frame(matrix(ncol=10,nrow=0, 
#                                dimnames=list(NULL, 
#                                              c("subj_id",
#                                               "test_part", 
#                                               "grid_number", 
#                                               "num_clicks",
#                                               "i",
#                                               "j",
#                                               "hit",
#                                               "t",
#                                               "click_number",
#                                               "genuine_first"))))
# 
# 
# for (row in 1:nrow(E2.click_df)) {
#   
#     subject_click_log <- data.frame(fromJSON(E2.click_df[row, ]$click_log)) %>%
#     mutate(
#       click_number = 1:n(),
#       subj_id = E2.click_df[row, ]$subj_id,
#       test_part = E2.click_df[row, ]$test_part,
#       grid_number = E2.click_df[row, ]$grid_number,
#       num_clicks = E2.click_df[row, ]$num_clicks,
#       genuine_first = E2.click_df[row, ]$genuine_first
#     )
#     
#     E2.click_log <- rbind(E2.click_log, subject_click_log);
# }
#     
# 
# E2.click_log <- E2.click_log %>%
#   relocate(subj_id, .before = i) %>%
#   relocate(genuine_first, .before = i) %>% 
#   relocate(test_part, .before=i) %>% 
#   relocate(grid_number, .before=i) %>% 
#   relocate(click_number, .before=i) %>%
#   group_by(subj_id,test_part,grid_number) %>%
#   mutate(RT=t-lag(t,default=0)) %>%
#   filter(!(subj_id %in% E2.lucky_pretenders))%>%
#  mutate(hit_bin = hit !=0,
#         lag3 = lag(hit,3)!=0,
#         lag2 = lag(hit,2)!=0,
#         lag1 = lag(hit,1)!=0,
#         lead1 = lead(hit,1)!=0,
#         lead2 = lead(hit,2)!=0,
#         lead3 = lead(hit,3)!=0)

```

```{r simualte_data, echo=FALSE, cache=TRUE}

simulate_optimal_game <- function(ship_positions) {
  #initialize
  cur_board_state <- rep(NA,25); 
  nhits <- 0;
  nclicks <- 0;
  nships <- sum(ship_positions);
  board_state = c()
  flat_position = c();
  click_number = c();
  hit_bin = c();

  while (nships>nhits) {

    board_state = c(board_state,paste(cur_board_state,collapse=','));

    posterior <- get_likelihood(paste(cur_board_state,collapse=',')) %>%
      get_posterior() %>%
      scan(text = ., what= numeric(), sep="," , quiet = TRUE);

    best_shot = which.max(posterior);
    flat_position = c(flat_position,best_shot);

    #click on best shot
    nclicks = nclicks+1;
    click_number = c(click_number,nclicks)

    result = ship_positions[best_shot]

    hit_bin = c(hit_bin,ifelse(result==1,1,0));

    if (result==1) {
      nhits = nhits+1;
    }
    cur_board_state[best_shot]=result;

  }
  return(data.frame(board_state,flat_position,hit_bin,click_number))

}

simulate_random_game <- function(ship_positions) {
  #initialize
  cur_board_state <- rep(NA,25); 
  nhits <- 0;
  nclicks <- 0;
  nships <- sum(ship_positions);
  board_state = c()
  flat_position = c();
  click_number = c();
  hit_bin = c();

  while (nships>nhits) {

    board_state = c(board_state,paste(cur_board_state,collapse=','));

    random_guess = which(is.na(cur_board_state))%>%sample(1);
    flat_position = c(flat_position,random_guess);

    #click on best shot
    nclicks = nclicks+1;
    click_number = c(click_number,nclicks)

    result = ship_positions[random_guess]

    hit_bin = c(hit_bin,ifelse(result==1,1,0));

    if (result==1) {
      nhits = nhits+1;
    }
    cur_board_state[random_guess]=result;

  }
  return(data.frame(board_state,flat_position,hit_bin,click_number))

}

nsim = 100;
nboards = all_possible_boards%>%nrow();

for (i in 1:nsim) {
  cur_board = all_possible_boards[sample(1:nboards,1),];
  if (i==1) {
    simulated_optimal = simulate_optimal_game(cur_board)%>%mutate(game=i)
  } else {
    simulated_optimal = bind_rows(simulated_optimal, simulate_optimal_game(cur_board)%>%mutate(game=i))
  }
}

for (i in 1:nsim) {
  cur_board = all_possible_boards[sample(1:nboards,1),];
  if (i==1) {
    simulated_random = simulate_random_game(cur_board)%>%mutate(game=i)
  } else {
    simulated_random = bind_rows(simulated_random, simulate_random_game(cur_board)%>%mutate(game=i))
  }
}

```

```{r p_click, echo = FALSE, cache=TRUE}

E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')

get_random_flat_position <- function(board_state) {

  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
  return(which(is.na(board_state))%>%sample(1))

}

get_optimal_flat_position <- function(posterior) {

  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(which.max(posterior))

}

E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
  mutate(flat_position = get_optimal_flat_position(posterior),
         p_click = get_p_click(posterior,flat_position),
         p_click_rank = get_p_click_rank(posterior,flat_position),
         test_part = 'optimal')

E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
  mutate(flat_position = get_random_flat_position(board_state),
         p_click = get_p_click(posterior,flat_position),
         p_click_rank = get_p_click_rank(posterior,flat_position),
         test_part = 'random')

# E2.click_log <-E2.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>% 
#   mutate(flat_position = 1+i*5+j) %>%
#   group_by(subj_id,genuine_first,test_part,grid_number) %>%
#   arrange(click_number) %>%
#   summarise(click_number = click_number,
#             board_state=create_board_states(flat_position,hit_bin),
#             flat_position=flat_position,
#             i=i,
#             j=j,
#             RT=RT,
#             hit_bin=hit_bin) %>%
#   rowwise()%>%
#   mutate(likelihood = get_likelihood(board_state),
#          posterior = get_posterior(likelihood),
#          p_click = get_p_click(posterior,flat_position),
#          entropy = get_posterior_entropy(posterior),
#          p_click_rank = get_p_click_rank(posterior,flat_position));

E2.mean_P_click_rank <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal) %>%
  rbind(E2.click_log_with_boards_random) %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank));

E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal) %>%
  rbind(E2.click_log_with_boards_random) %>%
  rowwise() %>%
  filter(!hit_bin) %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

p<- E2.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('../figures/E2_pclick_rank.pdf',p)
ggsave('../figures/E2_pclick_rank.png',p,width=5,height=2,dpi=300)

p<- E2.mean_P_click_rank_misses_only %>%
    filter(test_part != 'optimal') %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    scale_x_continuous(limits=c(0,15))+
    theme(legend.position=c(0.10,0.85)) +
    # facet_wrap(~genuine_first, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('../figures/E2_pclick_rank_misses_only.pdf',p)
ggsave('../figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)

E2.num_irrational_clicks <- E2.click_log_with_boards %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(nic=sum(p_click==0)/5);

p<- E2.num_irrational_clicks %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=nic,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='mean number of irrational clicks per game', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2)+
    theme_classic();

ggsave('figures/../E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)

 E2.RT_p_click_cor <- E2.click_log_with_boards %>%
     filter(click_number<9 & click_number>1) %>%
     group_by(subj_id,test_part, genuine_first) %>%
     summarise(r=cor(RT,p_click,method='spearman'));

 E2.RT_entropy_cor <- E2.click_log_with_boards %>%
     filter(click_number<9 & click_number>1) %>%
     group_by(subj_id,test_part, genuine_first) %>%
     summarise(r=cor(entropy,p_click,method='spearman'))

```

The mean posterior rank of non-pretend cell selections was `r E2.mean_P_click_rank%>%filter(test_part=='nonpretend')%>%pull(p_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (see Fig. 2B; `r E2.mean_P_click_rank%>%filter(test_part=='random')%>%pull(p_click_rank)%>%mean()%>%printnum()`, `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). Pretend games were significantly less optimal than non-pretend games (`r E2.mean_P_click_rank%>%filter(test_part=='pretend')%>%pull(p_click_rank)%>%mean()%>%printnum()`; `r apa_print(t.test(E2.mean_P_click_rank%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`), but still more optimal than those of a random agent (`r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). In other words, the optimality of pretend games relative to random cell selection was not merely due to the fact that pretenders clicked on ships more than expected by chance. Even when missing a ship, their cell selections made sense given the limited information they pretended to have.

```{r numclick_optimality, echo=FALSE, fig.cap="Experimental Design", out.width = '100%', fig.cap = "A: The total number of clicks in pretend and non-pretend games was highly similar. B: Non-pretend games were significantly more optimal than pretend games, but both were more optimal than what is expected by random chance. Box: An illustration of the difference between success and optimality. A given click (highlighted with a black border) can reveal a ship (upper row) or not (lower row) regardless of whether the click makes sense given available information (left column) or not (right column). "}
knitr::include_graphics("../figures/numclicks_optimality.png")
```

## Response time analysis

```{r total_time, echo=FALSE}

E2.total_time_df <- E2.click_log %>%
 filter(click_number==num_clicks)%>%
  mutate(t=t/1000)

E2.median_time <- E2.total_time_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(t=median(t)) %>%
  spread(test_part,t);

# anova
E2.time_anva <- aov(t ~ genuine_first * condition, 
                         data = E2.median_time %>%
                           gather('condition','t',nonpretend,pretend) %>%
                           mutate(condition=as.factor(condition)));

E2.median_time_per_click <- E2.click_log %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(RT=median(RT)) %>%
  spread(test_part,RT) %>%
  mutate(diff=pretend-nonpretend);

# anova
E2.click_time_anva <- aov(RT ~ genuine_first * condition, 
                         data = E2.median_time_per_click %>%
                           gather('condition','RT',nonpretend,pretend) %>%
                           mutate(condition=as.factor(condition)))
```

In general, participants gave slower responses when pretending compared to when playing normally. Despite the similar number of clicks per game, the median game duration was `r E2.total_time_df%>%filter(test_part=='nonpretend')%>%pull(t)%>%median()%>%printnum()` seconds in the non-pretend condition and `r E2.total_time_df%>%filter(test_part=='pretend')%>%pull(t)%>%median()%>%printnum()` seconds in the pretend condition, which was significantly longer (`r t.test(E2.median_time$pretend-E2.median_time$nonpretend)%>%apa_print()%>%'$'(statistic)`). This difference showed a significant interaction with the order of conditions, such that completing games in the first block took longer (`r apa_print(E2.time_anva%>%summary())$full_result$genuine_first_condition`).

In our instructions to pretenders, we asked them to pretend not to know where the ships were, not only in *where* they chose to click, but also in *when* they clicked. In the following set of analyses, we identified patterns in click latency in non-pretend games, and asked whether the same patterns are also observed in pretend games. We focused on the effects of hitting versus missing a ship, decision uncertainty, and ship completion.

```{r RT_by_accuracy}

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
  };
  return(sd(medians))
}

E2.RT_by_lag3 <- E2.click_log %>%
  filter(click_number>3) %>%
  group_by(subj_id,test_part,lag3,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag3,RT,sep='_') %>%
  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);

E2.RT_by_lag2 <- E2.click_log %>%
  filter(click_number>2) %>%
  group_by(subj_id,test_part,lag2,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag2,RT,sep='_') %>%
  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);

E2.RT_by_lag1 <- E2.click_log %>%
  filter(click_number>1) %>%
  group_by(subj_id,test_part,lag1,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

E2.RT_by_hit <- E2.click_log %>%
  filter(click_number>0) %>%
  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)

E2.RT_by_lead1 <- E2.click_log %>%
  filter(click_number<num_clicks) %>%
  group_by(subj_id,test_part,lead1,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

E2.RT_by_lead2 <- E2.click_log %>%
  filter(click_number<num_clicks-1) %>%
  group_by(subj_id,test_part,lead2,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead2,RT,sep='_') %>%
  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)

E2.RT_by_lead3 <- E2.click_log %>%
  filter(click_number<num_clicks-2) %>%
  group_by(subj_id,test_part,lead3,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead3,RT,sep='_') %>%
  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)

E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
  merge(E2.RT_by_lag2) %>%
  merge(E2.RT_by_lag1) %>%
  merge(E2.RT_by_hit) %>%
  merge(E2.RT_by_lead1) %>%
  merge(E2.RT_by_lead2) %>%
  merge(E2.RT_by_lead3);

E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
  mutate(position = ifelse(position=="lag3_diff", 3,
                           ifelse(position=="lag2_diff",2,
                                  ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,
                                                ifelse(position=="lead1_diff",-1,
                                                       ifelse(position=="lead2_diff",-2,-3
                                                              ))))))) %>%
  dplyr::select(subj_id,test_part,genuine_first,position,difference);

p<- E2.RT_by_hit_lags_long %>% 
  filter(position<2 & position > -1)%>%
  group_by(test_part,position) %>%
  summarise(median_difference=median(difference),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=0:1,labels=c('hit on current trial', 'hit on previous trial'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
  theme_classic()

ggsave('../figures/E2_RT_by_lag_three_only_collapse.pdf',p,width=7)
ggsave('../figures/E2_RT_by_lag_three_only_collapse.png',p,width=4, height=3.5, dpi=300)

```

### Effects of hitting versus missing a ship

In non-pretend games, players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='nonpretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) when hitting compared to missing a ship (see Fig. 3A; `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). The same effect was observed in pretend games: players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='pretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) in hits compared to misses (`r E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). However, this difference between hits and misses was significantly more pronounced in pretend games (`r apa_print(t.test(E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff), E2.RT_by_hit_lags%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)))$statistic`).

The opposite effect was observed for a hit on the previous click: players were slower by `r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) after hitting a ship (`r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). This effect was also observed in pretend games, where players were slower by `r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='pretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) following a hit on the previous click (`r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). Again, the absolute effect in ms was exaggerated in the pretend condition (`r t.test(E2.RT_by_lag1%>%filter(!genuine_first & test_part=='pretend')%>%pull(lag1_diff),E2.RT_by_lag1%>%filter(genuine_first & test_part=='nonpretend')%>%pull(lag1_diff))%>%apa_print()%>%'$'(statistic)`).

### Effects of decision uncertainty

When playing Battleships, it is sometimes clear what the next cell selection should be, and sometimes more difficult to decide where to click next. To capture this notion of decision uncertainty, we calculated the entropy of the posterior distribution over cell selections $H(P)=-\sum_{i=1}^{k}P(x_i) \log P(x_{i})$, where $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$, and asked how this measure relates to decision latency, or the time taken to click on the next cell. $H(P)$ is high when players need to decide between multiple cells with a similar probability of hiding a ship, and low when there are only a few candidates with a high probability of hiding a ship. For every player and condition separately, we fitted a multiple linear regression to predict decision latency based on $H(P)$ and $H(P)^2$. The resulting coefficients were then subjected to a group-level inference. The first cell selection of each game was excluded from this analysis, because entropy was constant for the first click.

```{r entropy-RT, echo=FALSE, cache=TRUE}

E2.RT_by_entropy <- E2.click_log_with_boards %>%
  filter(click_number>1) %>%
  mutate(entropy_bin=round(entropy*2)/2) %>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT)) %>%
  group_by(test_part,subj_id,entropy_bin) %>%
  summarise(RT=median(RT),
            corrected_RT = median(corrected_RT)) %>%
  group_by(test_part,entropy_bin) %>%
  summarise(median_RT=median(RT),
            sem_RT=bootstrap_error(RT,N_perm),
            median_corrected_RT=median(corrected_RT),
            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))

p <- E2.RT_by_entropy %>%
  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
  theme_classic()

ggsave('../figures/E2_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
# ggsave('../figures/E2_RT_by_entropy_no_split.svg',p,width=5,height=2,dpi=300)

p <- E2.RT_by_entropy %>%
  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
  labs(fill=' ',x='Entropy of posterior', y='median residual RT (ms)')+
  theme_classic()

ggsave('../figures/E2_corrected_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)

E2.RT_by_entropy_coefs <- E2.click_log_with_boards %>%
  filter(click_number>1) %>%
  group_by(genuine_first,subj_id,test_part,grid_number) %>%
  mutate(entropy_centered = entropy-mean(entropy),
         entropy_squared = entropy_centered^2) %>%
  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
  group_by(genuine_first,subj_id,test_part,grid_number) %>%
  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
  mutate(tidys=list(broom::tidy(model))) %>%
  unnest(tidys) %>%
  dplyr::select(!c(std.error,statistic,p.value)) %>%
  group_by(genuine_first,subj_id,test_part,term) %>%
  summarise(estimate = mean(estimate))

E2.RT_by_entropy_coefs_summary <- E2.RT_by_entropy_coefs %>%
  group_by(test_part,term) %>%
  summarise(mean_estimate = mean(estimate),
            se_estimate = se(estimate))

p <- E2.RT_by_entropy_coefs_summary %>%
  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
  filter(term!='(Intercept)') %>%
  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='term', y='coefficient')+
  theme_classic()

ggsave('../figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)

E2.corrected_RT_by_entropy_coefs <- E2.click_log_with_boards %>%
  filter(click_number>1) %>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT)) %>%
  group_by(genuine_first,subj_id,test_part) %>%
  mutate(entropy_centered = entropy-mean(entropy),
         entropy_squared = entropy_centered^2) %>%
  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
  group_by(genuine_first,subj_id,test_part) %>%
  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
  mutate(tidys=list(broom::tidy(model))) %>%
  unnest(tidys) %>%
  dplyr::select(!c(std.error,statistic,p.value))

E2.corrected_RT_by_entropy_coefs_summary <- E2.corrected_RT_by_entropy_coefs %>%
  group_by(test_part,term) %>%
  summarise(mean_estimate = mean(estimate),
            se_estimate = se(estimate))

p <- E2.corrected_RT_by_entropy_coefs_summary %>%
  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
  filter(term!='(Intercept)') %>%
  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='term', y='coefficient')+
  theme_classic()

ggsave('../figures/E2_corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)

```

In non-pretend games, we found no evidence for a linear relation between decision entropy and decision latency (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`). We did however observe a block order effect, with a significantly negative linear modulation only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).  In pretend games this negative relation between $H(P)$ and click latency was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).

A negative linear modulation of $H(P)$ on decision latency is surprising, as it suggests that participants were quicker to decide when uncertainty was high. However, visual inspection of the RT/entropy curves (Fig. 3B) reveals a pronounced quadratic modulation, such that in both pretend and non-pretend games slower decision times are observed for intermediate $H(P)$ values of around 2. Similar to the linear effect of $H(P)$ on decision latency, here too a quadratic modulation in non-pretend games was observed only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`). Again, in pretend games this negative quadratic relation was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared')%>%pull(estimate)%>%t.test())$statistic`). A between-subjects comparison focusing on games from the first block revealed that this negative quadratic effect was stronger in pretend, compared to non-pretend games (`r apa_print(t.test(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate),E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)))$statistic`).

```{r latency, echo=FALSE, out.width = '100%', fig.cap = "A: A subtraction of decision latencies in misses versus hits in the current and previous cell selections, in pretend and non-pretend games. B. Upper panel: Effect of the entropy of the posterior distribution over cell selections (a measure of decision uncertainty) on decision latency in pretend and non-pretend games. Lower panel: mean beta coefficients for a multiple regression model, predicting decision latency from entropy and entropy squared. C: Decision latency after a hit, as a function of hit number."}
knitr::include_graphics("../figures/decision_latencies.png")
```

<!-- $H(P)$ is highest before the first click in a game and lowest just before the last one. To make sure any results are not driven by a general effect of click number on decision latency, and in deviation from our pre-registered analysis plan, we first mean-centered all decision latencies within individual subjects, conditions, and click numbers. This way, any effect of $H(P)$ on latency cannot reflect effects of click order.  -->

## Effects of ship completion

```{r}

get_hit_number <- function(hit) {

  num_A <- 0;
  num_B <- 0;
  num_C <- 0;

  hit_number <- c();

  for (i in seq(length(hit))) {
    if (hit[i]==0) {
      hit_number = c(hit_number,NaN)
    } else if (hit[i]=='A') {
      num_A = num_A + 1;
      hit_number = c(hit_number,num_A)
    } else if (hit[i]=='B') {
      num_B = num_B + 1;
      hit_number = c(hit_number,num_B)
    } else if (hit[i]=='C') {
      num_C = num_C + 1;
      hit_number = c(hit_number,num_C)
    }
  }

  return(hit_number)
}

revealed_three <- function(hit_number) {

  revealed_three <- c();
  three = FALSE;

  for (i in seq(length(hit_number))) {
    revealed_three = c(revealed_three, three);
     if (!(is.na(hit_number[i])) & hit_number[i]==3) {
      three = TRUE
    } 
  }

  return(revealed_three)
}

E2.hit_order <- E2.click_log %>%
  group_by(genuine_first,subj_id,test_part,grid_number) %>%
  summarise(click_number = click_number,
            hit = hit,
            RT = RT,
            hit_number = get_hit_number(hit),
            revealed_three = revealed_three(hit_number),
            hit_type = ifelse(hit_number==1, 'first',
                              ifelse(hit_number==2 & !revealed_three, 'second',
                                     ifelse(hit_number==2 & revealed_three, 'secondlast',
                                            ifelse(hit_number==3,'thirdlast','sea'))))) %>%
  mutate(prev_hit_type = lag(hit_type),
         prev_hit = lag(hit))

E2.RT_by_hit_order <- E2.hit_order %>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(RT=median(RT))

p <- E2.RT_by_hit_order %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(RT = median(RT),
            sem_RT = bootstrap_error(RT,N_perm)) %>%
  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT')+
  theme_classic()

ggsave('../figures/E2_RT_by_previous_hit_number_collapse.png',p,width=5,height=2,dpi=300)

##correct for trial order

E2.corrected_RT_by_hit_order <- E2.hit_order %>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT))%>%
  rowwise() %>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(corrected_RT=median(corrected_RT))

p <- E2.corrected_RT_by_hit_order %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
            corrected_RT = median(corrected_RT)) %>%
  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
  theme_classic()

ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)

## correct for true ship size

E2.RT_by_hit_order_boats_only <- E2.hit_order %>%
  filter(prev_hit!='A')%>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(RT=median(RT))

p <- E2.RT_by_hit_order_boats_only %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(RT = median(RT),
            sem_RT = bootstrap_error(RT,N_perm)) %>%
  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT')+
  theme_classic()

ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)

E2.corrected_RT_by_hit_order_boats_only <- E2.hit_order %>%
  filter(prev_hit!='A')%>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT))%>%
  rowwise() %>%
  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
  summarise(corrected_RT=median(corrected_RT))

p <- E2.corrected_RT_by_hit_order_boats_only %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
            corrected_RT = median(corrected_RT)) %>%
  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
  theme_classic()

ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed_boats_only.png',p,width=5,height=2,dpi=300)

```

The previous analyses reveal robust associations between game state and decision latency, that are highly similar, and even exaggerated, in pretend compared to non-pretend games. To get at subtler dynamics in a more direct way, in a follow-up exploratory analysis we focused on decision latencies following a hit. We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn’t been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine (see Fig. 3C). In the first case, players know the ship must continue in one of the neighboring cells. In the second case, there is a good chance the ship continues (if this ship turns out to be a submarine). In the third and fourth cases, it is clear that the ship is fully sunk.

In non-pretend games, participants were significantly slower to select the next cell when they knew they had just completed a ship (categories 3 and 4) compared to when they just hit a ship, but were not sure (category 2) or knew they had not completely sunk it (category 1; `r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`). Specifically, we find that participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).

Strikingly, we found the exact same pattern in pretend games. Players were faster to make the next cell selection when they pretended to think that the current ship might not be fully sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`). This was not merely a difference between the first, second and third hits: participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`). Further analysis confirmed that this effect remained significant when controlling for click number (`r apa_print(E2.corrected_RT_by_hit_order%>%spread(prev_hit_type,corrected_RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`), and when restricting the analysis to the second hit of a ship that is in fact of size-two (`r apa_print(E2.RT_by_hit_order_boats_only%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).

This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their their response latency in the same way it would have been affected had they been in a non-pretend game.

## Half-games

Our optimality analysis showed that pretenders’ click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be. However,  due to the high number of possible board configurations, data from full games provide limited opportunity to compare cell selections for specific game states. In addition to asking, “What guides cell selections in pretend and non-pretend games?”, we also wanted to ask, “Where exactly would pretenders and non-pretenders click, given a specific board configuration?”.

To achieve this, the sixth game in each block started not with an empty grid, but with the contents of some cells already revealed by a previous player. As before, pretenders also knew where the remaining ships were hidden, but tried to play as if they only knew what was known to this previous player. Having cell selections from 250 players for each board configuration and condition allowed us to plot and compare the distribution of clicks under a genuine, or pretend, knowledge state.

```{r, child=c('../half_games.Rmd')}
```

For example, look at Board A in Fig. 4 (upper left corner). Imagine that you need to select your next move, knowing that exactly two size-two patrol boats and one size-three submarine are hiding in the water. Where would you click next? Perhaps one of the two revealed ships is the size-three submarine. In that case you want to click on the first cell in the third row, or on the third cell in the fifth row. Or maybe these are the two size-two patrol boats. This would mean the submarine is hiding in some other place, and for it not to touch sides with the boats this must be in the first row. The hit probability map in the second column visualizes this reasoning, by plotting the normalized probability that a ship is hidden in different cells. As can be seen, the highest probabilities are to the sides of the revealed ships, and in the first row.

In the third column we plot the distribution of clicks for non-pretend players. This distribution is in agreement with the hit probability map (board A: `r cor_A_nonpretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_nonpretend%>%apa_print()%>%'$'(full_result)`). Finally, in the fourth column we plot the distribution of cell selections for pretend players. Although noisier, this distribution is also in agreement with the hit probability map (board A: `r cor_A_pretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_pretend%>%apa_print()%>%'$'(full_result)`), and more importantly, with the hit distribution of non-pretenders for the same board configuration (board A: `r cor_A_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`,  board B: `r cor_B_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`).

```{r half-games, echo=FALSE, out.width = '100%', fig.cap = "Half games. First column: the two boards, as they appeared to pretenders. Second column: objective hit probability, given game state. Third and fourth columns: empirical click probabilities in non-pretend and pretend games, respectively."}
knitr::include_graphics("../figures/Half_games.png")
```

<!-- Furthermore, compared to non-pretend games, in pretend games participants were more likely to make irrational clicks on cells where a ship could not in fact be hiding (a median of `r E2.num_irrational_clicks%>%filter(test_part=='pretend')%>%pull(nic)%>%median()` per game in pretend versus `r E2.num_irrational_clicks%>%filter(test_part=='nonpretend')%>%pull(nic)%>%median()` in non-pretend games.  `r E2.num_irrational_clicks%>%spread(test_part,nic)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).  -->
## Judge trials

```{r condition-classifier, echo=FALSE, eval=FALSE}

set.seed(42);

gamewise_data <- E2.click_log_with_boards %>% 
  filter(!genuine_first | test_part=='pretend') %>% 
  group_by(subj_id,test_part,grid_number) %>%
  filter(sum(RT)<=60000) %>%
  summarise(median_RT=median(RT),
            mean_p_click_rank = mean(p_click_rank),
            first_click_latency=RT[click_number==1],
            num_irrational_clicks = sum(p_click==0)) %>%
  ungroup()%>%
  select(test_part,mean_p_click_rank,median_RT,num_irrational_clicks);

gamewise_data_nonpretend = gamewise_data%>%filter(test_part=='nonpretend') %>%mutate(nrow=1:n())
gamewise_data_pretend = gamewise_data%>%filter(test_part=='pretend')
rows <- sample(nrow(gamewise_data_nonpretend))
gamewise_data_pretend <- gamewise_data_pretend[rows,]%>%mutate(nrow=1:n())

pair_gamewise_data <- full_join(gamewise_data_nonpretend,
                                gamewise_data_pretend,
                                by='nrow',
                                suffix=c('.nonpretend','.pretend')) %>%
  mutate(which_side = sample(c(-1,1),nrow(gamewise_data_pretend),replace=TRUE),
         mean_p_click = which_side*(mean_p_click_rank.nonpretend-mean_p_click_rank.pretend),
         median_RT = which_side*(median_RT.nonpretend-median_RT.pretend),
         num_irrational_clicks = which_side*(num_irrational_clicks.nonpretend-num_irrational_clicks.pretend),
         which_side=factor(which_side)) %>%
  select(which_side,mean_p_click,median_RT,num_irrational_clicks)

train_control <- trainControl(method = "CV", number=10, sampling='down')

# train the model on training set
model <- train(test_part ~ .,
               data = gamewise_data,
               trControl = train_control,
               method = "svmLinear",
               preProcess = c("center","scale"))

pair_model <- train(which_side ~ .,
               data = pair_gamewise_data,
               trControl = train_control,
               method = "svmLinear",
               preProcess = c("center","scale"))

# print cv scores

```
Despite the fact that pretenders were overall successful in mimicking patterns of cell selection and decision latency, they also left many clues for an independent observer to pick up on: they were slower, displayed exaggerated effects of decision uncertainty and hit success on response time, and were less optimal in choosing where to click next. To illustrate, a support vector machine (SVM) algorithm reached an accuracy level of 72% in linearly classifying condition (pretend / non-pretend) based on just three basic summary features: median decision latency, game optimality score, and number of irrational clicks per game.

```{r judge}

get_common_board = function(grid_number,test_part) {
  pretend_grids = grid_number[which(test_part=='pretend')];
  nonpretend_grids = grid_number[which(test_part=='nonpretend')];
  common_board = intersect(pretend_grids,nonpretend_grids);
  return(common_board)
};

common_boards <- E2.df %>%
  filter(test_part=='pretend' | test_part=='nonpretend') %>%
  dplyr::select(subj_id,grid_number,test_part) %>%
  group_by(subj_id) %>%
  summarise(common_board = get_common_board(grid_number,test_part))

used_for_pretend <- function(s, g) {

  return (E2.df%>%filter(test_part=='pretend' &
                           subj_id==s & 
                           grid_number == g) %>%
            nrow()==1)
}

used_for_nonpretend <- function(s, g) {

  return (E2.df%>%filter(test_part=='nonpretend' &
                           subj_id==s & 
                           grid_number == g) %>%
            nrow()==1)
}

E2.judge_df <- E2.df %>%
  filter(test_part=='judge' & !(cheat_player %in% E2.lucky_pretenders))%>%
  dplyr::select(subj_id, 
         grid_number, 
         cheat_player,
         genuine_player,
         cheater,
         noncheater,
         decision,
         replay_log,
         choose_pretender,
         correct)%>%
  rowwise()%>%
  mutate(choose_pretender=choose_pretender=='True',
         used_for_pretend = used_for_pretend(subj_id,grid_number),
         used_for_nonpretend = used_for_nonpretend(subj_id,grid_number),
         common_board = used_for_pretend & used_for_nonpretend,
         board_type = ifelse(common_board, 'common',
                             ifelse(used_for_pretend, 'pretend',
                                    ifelse(used_for_nonpretend,'nonpretend','new')))
)

p <- E2.judge_df %>%
  group_by(subj_id) %>%
  summarise(ncorrect=sum(correct)) %>%
  ggplot(aes(x=ncorrect)) +
  geom_bar(fill='black') +
  theme_classic() +
  labs(x='number of correct guesses (out of 5)',y='number of players') +
  scale_x_continuous(breaks=c(0,1,2,3,4,5));

ggsave('../figures/E2_correct_guesses.png',p,width=5,height=3,dpi=300)

p <- E2.judge_df %>%
  mutate(choose_pretender = ifelse(choose_pretender,
                                   'who had hints?', 
                                   'who had no hints?')) %>% 
  group_by(choose_pretender,subj_id) %>%
  summarise(ncorrect=sum(correct)) %>%
  ggplot(aes(x=ncorrect)) + 
  geom_bar(fill='black') +
  theme_classic() +
  labs(x='number of correct guesses (out of 5)',y='number of players') + 
  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
  facet_wrap(~choose_pretender,nrow=2);

ggsave('../figures/E2_correct_guesses_by_choose_pretender.png',p,width=5,height=3,dpi=300);

p <- E2.judge_df %>% 
  group_by(board_type,subj_id) %>%
  summarise(ncorrect=sum(correct)) %>%
  ggplot(aes(x=ncorrect)) + 
  geom_bar(fill='black') +
  theme_classic() +
  labs(x='number of correct guesses (out of 5)',y='number of players') + 
  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
  facet_wrap(~board_type,nrow=1);

ggsave('../figures/E2_correct_guesses_by_board_type.png',p,width=5,height=3,dpi=300)
```

Despite this, participants were never significantly above chance in judging which of the two presented games came from a pretender (mean accuracy: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%; t test against 50: `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%t.test(mu=50))$statistic`).

For a subset of `r E2.judge_df%>%filter(common_board)%>%pull(correct)%>%length()` players, one of the five boards presented in the judging block was already presented both in the pretend and in the non-pretend blocks. Still, even this subset of participants who had just experienced a board both as pretenders and as non-pretenders were not significantly above chance in telling which of two other players had hints for this same board (mean accuracy: 55%, $p=.141$).

## Are good pretenders also better judges?

```{r judge-pretend-cor, echo=FALSE}

unidentifiable_ss <- E2.judge_df %>%
  mutate(subj_num=substr(as.character(subj_id),1,3))%>%
  group_by(subj_num)%>%
  summarise(subj_id=subj_id, n=n())%>%
  filter(n>5)%>%
  pull(subj_id)%>%
  unique();

E2.judge_performance <- E2.judge_df %>%
  filter(!(subj_id %in% unidentifiable_ss)) %>%
  mutate(subj_num=as.numeric(substr(as.character(subj_id),1,3)))%>%
  group_by(subj_num)%>%
  summarise(judge_rate=sum(correct)/5);

E2.pretend_performance <- E2.judge_df %>%
  filter(!(subj_id %in% unidentifiable_ss)) %>%
  mutate(subj_num=cheat_player)%>%
  group_by(subj_num)%>%
  summarise(ntrials=n(),
            successful_pretend=ntrials-sum(correct),
            pretend_rate = successful_pretend/ntrials);

E2.performance_df <- merge(E2.judge_performance,E2.pretend_performance);

p <- E2.performance_df %>%
  ggplot(aes(x=judge_rate,y=pretend_rate)) +
  geom_jitter() + 
  labs(x='judge accuracy',y='pretense success rate') +
  theme_classic()

ggsave('../figures/E2_judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)

min_njudge = seq(0,10);
cors =c()
ns = c()

for (minn in min_njudge) {

  temp_df = E2.performance_df %>%
    filter(ntrials>minn)

  cur_cor = cor(temp_df$pretend_rate,temp_df$judge_rate);
  cur_n = temp_df%>%nrow()
  cors = c(cors,cur_cor)
  ns = c(ns, cur_n)
}

judge_pretend_correlation_df <- data.frame(min_njudge,cors,ns);

```

```{r judge-optimality-cor, echo=FALSE}
E2.judge_summary <- E2.judge_df %>%
  group_by(subj_id) %>%
  summarise(judge = sum(correct));

E2.optimality_summary <- E2.click_log_with_boards %>%
  group_by(subj_id,test_part) %>%
  summarise(p_click_rank = mean(p_click_rank)) %>%
  spread(test_part,p_click_rank)

E2.judge_summary <- E2.judge_summary %>%
  merge(E2.optimality_summary)
```

The abilities to pretend and to detect pretense in others both rely on some form of theory of mind: reasoning about counterfactual knowledge states of oneself in the first case, and of others in the second. For example, in scrub jays, individual birds that pilfered another bird's caches were also more careful in hiding their own food, suggesting an ability to project from their own experience of stealing to the behavior of others [@emery2001effects]. Along similar lines, we reasoned that those participants whose pretend games better resembled the games of genuine players would also be more accurate in detecting pretense in others. To our surprise, this was not at all the case (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate))$full_result`). Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend))$full_result`).

Despite our considerable sample size of 500 participants, we are cautious in making strong claims based on the absence of a correlation between pretense and pretense detection abilities in our sample. We note, however, that previous developmental research has identified weaker correlations with theory of mind measures for a tendency to intentionally deceive another player in a game setting, compared to an ability to detect and understand deception, potentially indicating partly different mechanisms [@shultz1981development].

# Discussion

Many tasks (including writing papers for conferences) require us to overcome the ‘curse of knowledge’ — the challenge of putting ourselves in the shoes of someone who doesn’t know what we know. This is just an example of a more general fact: the ability to simulate counterfactual knowledge states is necessary for efficient and flexible communication. To do this, we must have an internal self-model that can generate thoughts and behavior for hypothetical beliefs, desires, and perceptions.

Research on Bayesian Theory of Mind provides some support for the existence of such a model by measuring subjects' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational], or implicitly [@onishi2005fifteen; @liu2017ten]. Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state - in this case, a counterfactual knowledge state in which a known piece of information is unknown. Instead of relying on model inversion (e.g., “Which belief states would give rise to this behavior?”), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output. Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].

Our subjects rose to the challenge. Their pretense behavior, although not perfect, closely resembled that of non-pretenders — not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency. Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: “What would I have done if I didn’t have this piece of knowledge?”

An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game. This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something. While we cannot fully rule out this interpretation, we think it is unlikely to explain participants’ pretense ability in our data, for two indirect reasons. First, suppressing thoughts on demand is notoriously difficult  [@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible. Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with a self-simulation account. In future work, we intend to more directly rule out this knowledge-suppression hypothesis, and well as to generalize our results to other games.

In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects’ models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense. While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2023-01-04 docs/cogsci/PNTKN.tex
cogsci submission

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission
\cogscifinalcopy

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Using pretense behavior to explore counterfactual
self-simulation}

\author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of london \\ London, WC1E 7HX UK \AND {\large \bf Chaz Firestone (chaz@jhu.edu)} \AND {\large \bf Ian Phillips (ianphillips@jhu.edu)} \\ Johns Hopkins University \\ Baltimore, MD 21218, USA}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

\begin{document}

\maketitle

\begin{abstract}
What we do depends on what we know. But sometimes we try to decouple our
behavior from our knowledge so that we appear not to know what we really
do know. Such pretense behavior requires understanding how we would
behave with different knowledge --- and so provides a window into
counterfactual self-simulation. However, little research has
characterized and evaluated pretense relative to non-pretense behavior.
Here, in a large-scale, pre-registered experiment, subjects played
normal games of Battleships (trying to sink ships hidden in a grid), as
well as `pretend' games, where they were told all the ships' locations
but had to pretend they were playing normally. Relative to normal games,
`pretend' games demonstrated similar but exaggerated behavioral
patterns. Furthermore, pretenders played rationally, but less so than
non-pretenders. Despite these differences, `judge' participants could
not detect `pretend' games. We discuss implications of these findings
for accounts of theory of mind and metacognition.

\textbf{Keywords:}
metacognition; theory of mind; pretense
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The ability to intentionally deceive others relies on a capacity to
reason about mental states (Frith \& Frith, 2005). This is evident in a
similar developmental trajectory for the acquisition of theory of mind
and the ability to deceive and detect deception (Shultz \& Cloghesy,
1981; Sodian, Taylor, Harris, \& Perner, 1991; Wimmer \& Perner, 1983),
and a similar distribution of deception and theory of mind in the animal
kingdom (e.g., Emery \& Clayton, 2004; Hall \& Brosnan, 2017). This link
makes conceptual sense: to deceive others, one needs to understand that
others can have different knowledge and beliefs than one's own.

Moreover, deception often involves pretense behavior, which in turn
relies on an ability to simulate and mimic one's own behavior under a
hypothetical belief state. For example, in order to successfully deceive
your friends into thinking that you were surprised by the birthday party
they threw for you, it is not sufficient that you are able to reason
about their mental states (``I know that they are planning a surprise
party, but they don't know that I know that.'') --- you also need to
convincingly simulate and mimic your hypothetical behavior had you not
known about the party (``Where would I look first? What would I say? How
long would it take me to recover from the surprise?''). This reliance of
pretense behavior on self-simulation makes it an ideal opportunity to
examine metacognitive knowledge about one's own mental states, and the
potential reliance of this knowledge on a self-simulation. By comparing
non-pretend and pretend behavior, we can ask which aspects of their own
cognitive processes subjects can and cannot effectively simulate, and
which aspects are not well-represented in their mental models of their
own cognition.

To this end, here we examine pretense in a game setting. Using an online
version of the game Battleships, participants played a `non-pretend'
(normal) version of the game (trying to sink ships whose locations are
unknown), as well as a `pretend' version where they were given secret
information about the location of hidden ships, but were instructed to
behave as if they didn't have this information.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

A detailed pre-registration can be accessed at
\href{https://osf.io/v9zsb}{osf.io/v9zsb}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

The research complied with all relevant ethical regulations and was
approved by the Research Ethics Committee of Johns Hopkins University.
500 Participants were recruited via Prolific (prolific.co) and gave
their informed consent prior to their participation. They were selected
based on their acceptance rate (\textgreater95\%) and for being native
English speakers. The entire experiment took approximately 20 minutes to
complete. Participants' pay was equivalent to an hourly wage of 9.50
USD, in addition to a bonus payment (0.2 - 2 USD, mean = 0.90).

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

Participants were first instructed that the experiment, based on the
game Battleships, had three parts, and that they could accumulate
`points' that would later translate to a monetary bonus payment. They
were then presented with a leaderboard of previous players, and given
the rules of the game:

\begin{quote}
``In the game Battleships, your task is to sink all ships located in a
grid with as few clicks as possible. What makes the game difficult is
that you can't see the ships; all you can see is a grid of squares, and
you have to guess where the ships are. To sink a ship, you need to click
on all of the squares it is located in. If you hit part of a ship, the
square will turn red. If there is no ship in the square, it will turn
blue.''
\end{quote}

We further explained that in this version of the game, ships can touch
corners, but their sides can't touch. This explanation was accompanied
by a visual presentation of legal and illegal ship configurations.

After completing a comprehension question and a practice round,
participants completed one `pretend' and one `non-pretend' block, each
comprising five full games and one half game (see below for details).
The order of pretend and non-pretend blocks was counterbalanced between
participants. The allocation of boards (spatial configurations of ships;
see Fig. 1A) to conditions was randomized between participants such that
exactly one board was played in both pretend and non-pretend conditions,
and this common board was different for different participants. The
order of boards within a block was fully randomized, with the exception
that half-games were always played last.

\hypertarget{non-pretend-normal-games}{%
\subsubsection{Non-pretend (normal)
games}\label{non-pretend-normal-games}}

In non-pretend games (Fig. 1B), participants aimed to sink two 2-square
patrol boats and one 3-square submarine with as few clicks as possible.
An online counter of the number of clicks was displayed on the screen.
After each game, feedback was given about the number of clicks and
resulting number of points obtained.

\hypertarget{pretend-games}{%
\subsubsection{Pretend games}\label{pretend-games}}

Participants in pretend games were given the same explanation of
Battleships, and played a practice round. However, they were then given
an additional twist:

\begin{quote}
``This time your goal is different. In this round, we're going to tell
you where the ships are, but \textbf{we want you to act like you don't
know this information}. We've marked the ships' locations with a cross,
so you'll know where they are the whole time; but your job is to play
the game as if these hints aren't there. To see how good you are at
this, we're going to compare your games to the games of people who
actually had no hints, and see how similar they are. We will measure
where and when you clicked; if your clicks look similar to people who
played like normal (trying to reveal all ships with as few clicks as
possible, but without any hints), you'll get bonus points. But if your
games look different, you won't get these bonus points. Your number of
clicks in this part will not affect your bonus. Only your ability to
play like you had no hints.''
\end{quote}

We informed participants that both the location and timing of their cell
clicks will be measured. After one practice round and one comprehension
question, participants played five pretend games (Fig. 1C), followed by
one pretend half-game. Each game was followed by a short message,
reminding them that a game that looks similar to the games of
participants who had no hints would be awarded 10 bonus points.

\hypertarget{half-games}{%
\subsubsection{Half games}\label{half-games}}

In order to directly compare participants' pretend and non-pretend games
for identical belief states (genuine or pretended ignorance about where
the ships are hidden), participants completed one pretend and one
non-pretend game given a partly finished board with the content of 7
cells already revealed (see Fig. 1D). We designed our half games to
produce a strong expectation to find a ship in specific cells, but not
in others. The assignment of half-completed boards to pretend and
non-pretend conditions was randomized between participants.

\hypertarget{judge-trials}{%
\subsubsection{Judge trials}\label{judge-trials}}

In the final part of the experiment, participants observed the games of
previous players and tried to determine who had hints and who didn't. On
each trial, two empty grids were presented side by side, with a smaller
grid on top, displaying the hidden positions of ships on the grid (Fig.
1F). The two grids corresponded to the true games of two previous
players who played a version of the top grid either as pretenders or as
non-pretenders. Only games shorter than one minute were chosen for
presentation in this part. For non-pretend games, only games from the
group of participants that pretended in the second block (and played
normally in the first block) were chosen for presentation in this part.
Judge participants observed a real time replay of the two grids, showing
not only where participants clicked, but also when. After making a
decision, participants were informed whether they would receive the 10
bonus points, or alternatively, whether the pretender would receive them
in the event the pretender managed to trick them.

A more detailed description of the study procedure is provided in the
study pre-registration document (LINK HIDDEN). Readers are also invited
to try a \href{https://jatos.mindprobe.eu/publix/NervzpM0Y0z}{demo of
the experiment}.

\begin{CodeChunk}
\begin{figure}[tb]
\includegraphics[width=1\linewidth]{../figures/Battleships_design} \caption[Experimental Design]{Experimental Design. See Methods for details.}\label{fig:design}
\end{figure}
\end{CodeChunk}

\hypertarget{results}{%
\section{Results}\label{results}}

We designed our analyses to explore subjects' capacity for
self-simulation under a counterfactual knowledge state, and the limits
of this capacity. We focused on where subjects clicked and when, and
asked whether this differed between pretend and non-pretend games. All
analyses were pre-registered unless otherwise specified. In our
pre-registration document, we committed to separately analyzing
participants according to whether they pretended before or after
completing a non-pretend block. Due to space limitations, and since the
results of the two groups mostly agreed, we report here the pooled
results from both groups of participants, and mention whenever we found
different results depending on block order. When directly comparing
pretend and non-pretend blocks, we perform a between-subjects comparison
using data from the first block only, i.e., pretend games from
participants who pretended in the first block and non-pretend games from
participants who played normally in the first block. We do so to ensure
that any successful pretending is not due to memory of one's own
behavior in a previous block, and that non-pretend games are not biased
by experience with the pretend block.

\hypertarget{total-number-of-clicks}{%
\subsection{Total number of clicks}\label{total-number-of-clicks}}

To sink all ships, players had to click on at least 7 and at most 25
squares. A simulated player that clicks randomly had a median total
click number of 23, and a near-optimal greedy player that consistently
selected the square with the highest objective probability of containing
a ship had a median total click number of 14. Among our players, the
median number of clicks was 16 in both pretend and non-pretend games
(see Fig. 2A).

We observed no significant difference in the number of clicks between
the two conditions (\(t(440.10) = 0.41\), \(p = .682\)). However, a
significant interaction with block order suggested that participants
made fewer clicks in the second block compared to the first, regardless
of which block was the pretend block (\(F(1, 996) = 10.59\),
\(\mathit{MSE} = 6.55\), \(p = .001\)).

In 62 pretend games from 20 players, games were completed after 7 clicks
only, without ever missing a ship. This never happened in non-pretend
games. We assumed that these participants did not follow task
instructions, and excluded them from all following analyses.

\hypertarget{game-optimality}{%
\subsection{Game optimality}\label{game-optimality}}

Pretend games were similar to non pretend games in the total number of
clicks --- but were they also similar in \emph{where} participants
clicked? More specifically, did cell selections in pretend games make
sense given the limited information those participants pretended to
have? To ask this, we approximated optimal behavior by calculating the
probability that a ship is hidden in each cell given available
information, \(p(ship(x_{i}))\), and the posterior probability that one
should click on a square, assuming a uniform prior over cells
\(P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}\).
Critically, in modeling pretend games we did not treat hints as part of
this available information for extracting \(p(ship(x_{i}))\), because an
optimal player should ignore hints in choosing where to click next.
Given this posterior map, a rational player should choose cells where
\(P(x_{i})\) is high {[}this behavior is not strictly optimal, but
approximates optimal behavior in most cases; Audinot, Bonnet, \& Viennot
(2014), Section 3.3{]}. To quantify optimality, before each cell
selection we computed the posterior probability map for all `unknown'
cells. Then, we ranked cells from high to low according to their
posterior probability and recorded the rank of the chosen cell: a lower
rank indicating more optimal behavior.

The mean posterior rank of non-pretend cell selections was 6.44 and
significantly lower (more optimal) than that of a simulated random agent
(see Fig. 2B; 9.18, \(t(479) = -48.34\), \(p < .001\)). Pretend games
were significantly less optimal than non-pretend games (6.93;
\(t(474.06) = -6.07\), \(p < .001\)), but still more optimal than those
of a random agent (\(t(479) = -38.09\), \(p < .001\)). Critically, the
same pattern was observed when restricting analysis to cell selections
that resulted in a miss (non-pretend - pretend: \(t(469.71) = -6.30\),
\(p < .001\); pretend - random: \(t(479) = -10.26\), \(p < .001\)). In
other words, the optimality of pretend games relative to random cell
selection was not merely due to the fact that pretenders clicked on
ships more than expected by chance. Even when missing a ship, their cell
selections made sense given the limited information they pretended to
have.

\begin{CodeChunk}
\begin{figure}[tb]
\includegraphics[width=1\linewidth]{../figures/numclicks_optimality} \caption[A]{A: The total number of clicks in pretend and non-pretend games was highly similar. B: Non-pretend games were significantly more optimal than pretend games, but both were more optimal than what is expected by random chance. Box: An illustration of the difference between success and optimality. A given click (highlighted with a black border) can reveal a ship (upper row) or not (lower row) regardless of whether the click makes sense given available information (left column) or not (right column). }\label{fig:numclick_optimality}
\end{figure}
\end{CodeChunk}

\hypertarget{response-time-analysis}{%
\subsection{Response time analysis}\label{response-time-analysis}}

In general, participants gave slower responses when pretending compared
to when playing normally. Despite the similar number of clicks per game,
the median game duration was 22.31 seconds in the non-pretend condition
and 29.23 seconds in the pretend condition, which was significantly
longer (\(t(479) = 12.68\), \(p < .001\)). This difference showed a
significant interaction with the order of conditions, such that
completing games in the first block took longer (\(F(1, 956) = 21.24\),
\(\mathit{MSE} = 243.45\), \(p < .001\), \(\hat{\eta}^2_G = .022\)).

In our instructions to pretenders, we asked them to pretend not to know
where the ships were, not only in \emph{where} they chose to click, but
also in \emph{when} they clicked. In the following set of analyses, we
identified patterns in click latency in non-pretend games, and asked
whether the same patterns are also observed in pretend games. We focused
on the effects of hitting versus missing a ship, decision uncertainty,
and ship completion.

\hypertarget{effects-of-hitting-versus-missing-a-ship}{%
\subsubsection{Effects of hitting versus missing a
ship}\label{effects-of-hitting-versus-missing-a-ship}}

In non-pretend games, players were faster by 109 ms (12\%) when hitting
compared to missing a ship (see Fig. 3A; \(t(479) = -12.37\),
\(p < .001\)). The same effect was observed in pretend games: players
were faster by 293 ms (31\%) in hits compared to misses
(\(t(233) = -11.83\), \(p < .001\); pretend first: \(t(479) = -16.23\),
\(p < .001\)). However, this difference between hits and misses was
significantly more pronounced in pretend games (\(t(314.67) = -7.45\),
\(p < .001\)).

The opposite effect was observed for a hit on the previous click:
players were slower by 182 ms (15\%) after hitting a ship
(\(t(479) = 13.39\), \(p < .001\)). This effect was also observed in
pretend games, where players were slower by 236 ms (16\%) following a
hit on the previous click (\(t(479) = 12.02\), \(p < .001\)). Again, the
absolute effect in ms was exaggerated in the pretend condition
(\(t(377.12) = 2.75\), \(p = .006\)).

\hypertarget{effects-of-decision-uncertainty}{%
\subsubsection{Effects of decision
uncertainty}\label{effects-of-decision-uncertainty}}

When playing Battleships, it is sometimes clear what the next cell
selection should be, and sometimes more difficult to decide where to
click next. To capture this notion of decision uncertainty, we
calculated the entropy of the posterior distribution over cell
selections \(H(P)=-\sum_{i=1}^{k}P(x_i) \log P(x_{i})\), where
\(P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}\), and asked
how this measure relates to decision latency, or the time taken to click
on the next cell. \(H(P)\) is high when players need to decide between
multiple cells with a similar probability of hiding a ship, and low when
there are only a few candidates with a high probability of hiding a
ship. For every player and condition separately, we fitted a multiple
linear regression to predict decision latency based on \(H(P)\) and
\(H(P)^2\). The resulting coefficients were then subjected to a
group-level inference. The first cell selection of each game was
excluded from this analysis, because entropy was constant for the first
click.

In non-pretend games, we found no evidence for a linear relation between
decision entropy and decision latency (\(t(479) = -1.67\),
\(p = .096\)). We did however observe a block order effect, with a
significantly negative linear modulation only in the group that
pretended in the first (\(t(233) = -10.24\), \(p < .001\)), but not in
the second block (\(t(245) = -0.06\), \(p = .950\)). In pretend games
this negative relation between \(H(P)\) and click latency was
significant in both groups, as well as overall (\(t(479) = -7.42\),
\(p < .001\)).

A negative linear modulation of \(H(P)\) on decision latency is
surprising, as it suggests that participants were quicker to decide when
uncertainty was high. However, visual inspection of the RT/entropy
curves (Fig. 3B) reveals a pronounced quadratic modulation, such that in
both pretend and non-pretend games slower decision times are observed
for intermediate \(H(P)\) values of around 2. Similar to the linear
effect of \(H(P)\) on decision latency, here too a quadratic modulation
in non-pretend games was observed only in the group that pretended in
the first (\(t(233) = -13.26\), \(p < .001\)), but not in the second
block (\(t(245) = -1.33\), \(p = .184\)). Again, in pretend games this
negative quadratic relation was significant in both groups, as well as
overall (\(t(479) = -15.65\), \(p < .001\)). A between-subjects
comparison focusing on games from the first block revealed that this
negative quadratic effect was stronger in pretend, compared to
non-pretend games (\(t(363.42) = 3.46\), \(p = .001\)).

\begin{CodeChunk}
\begin{figure}[tb]
\includegraphics[width=1\linewidth]{../figures/decision_latencies} \caption[A]{A: A subtraction of decision latencies in misses versus hits in the current and previous cell selections, in pretend and non-pretend games. B. Upper panel: Effect of the entropy of the posterior distribution over cell selections (a measure of decision uncertainty) on decision latency in pretend and non-pretend games. Lower panel: mean beta coefficients for a multiple regression model, predicting decision latency from entropy and entropy squared. C: Decision latency after a hit, as a function of hit number.}\label{fig:latency}
\end{figure}
\end{CodeChunk}

\hypertarget{effects-of-ship-completion}{%
\subsection{Effects of ship
completion}\label{effects-of-ship-completion}}

The previous analyses reveal robust associations between game state and
decision latency, that are highly similar, and even exaggerated, in
pretend compared to non-pretend games. To get at subtler dynamics in a
more direct way, in a follow-up exploratory analysis we focused on
decision latencies following a hit. We categorized hits into four types:
first hit on a ship, second hit on a ship when the size-three submarine
hasn't been sunk yet, second hit on a ship when the size-three submarine
has already been sunk, and third hit on a submarine (see Fig. 3C). In
the first case, players know the ship must continue in one of the
neighboring cells. In the second case, there is a good chance the ship
continues (if this ship turns out to be a submarine). In the third and
fourth cases, it is clear that the ship is fully sunk.

In non-pretend games, participants were significantly slower to select
the next cell when they knew they had just completed a ship (categories
3 and 4) compared to when they just hit a ship, but were not sure
(category 2) or knew they had not completely sunk it (category 1;
\(t(404) = -16.82\), \(p < .001\)). Specifically, we find that
participants were slower by 728 ms to make the next cell selection after
hitting the second cell of a ship if the size-three submarine had
already been sunk (\(t(404) = 13.45\), \(p < .001\)).

Strikingly, we found the exact same pattern in pretend games. Players
were faster to make the next cell selection when they pretended to think
that the current ship might not be fully sunk (\(t(369) = -14.47\),
\(p < .001\)). This was not merely a difference between the first,
second and third hits: participants were slower by 754 ms to make the
next cell selection after hitting the second cell of a ship if the
size-three submarine had already been sunk (\(t(404) = 13.45\),
\(p < .001\)). Further analysis confirmed that this effect remained
significant when controlling for click number (\(t(404) = 11.28\),
\(p < .001\)), and when restricting the analysis to the second hit of a
ship that is in fact of size-two (\(t(366) = 3.91\), \(p < .001\)).

This last finding bears emphasizing: In both classes two and three,
pretenders \emph{knew} that they had just sunk a size-two ship, but in
the second case they \emph{pretended not to know} this fact, and this
affected their their response latency in the same way it would have been
affected had they been in a non-pretend game.

\hypertarget{half-games-1}{%
\subsection{Half-games}\label{half-games-1}}

Our optimality analysis showed that pretenders' click selections closely
resemble those of non-pretenders, at least in that they are not random,
but rather guided by where a ship might be. However, due to the high
number of possible board configurations, data from full games provide
limited opportunity to compare cell selections for specific game states.
In addition to asking, ``What guides cell selections in pretend and
non-pretend games?'', we also wanted to ask, ``Where exactly would
pretenders and non-pretenders click, given a specific board
configuration?''.

To achieve this, the sixth game in each block started not with an empty
grid, but with the contents of some cells already revealed by a previous
player. As before, pretenders also knew where the remaining ships were
hidden, but tried to play as if they only knew what was known to this
previous player. Having cell selections from 250 players for each board
configuration and condition allowed us to plot and compare the
distribution of clicks under a genuine, or pretend, knowledge state.

For example, look at Board A in Fig. 4 (upper left corner). Imagine that
you need to select your next move, knowing that exactly two size-two
patrol boats and one size-three submarine are hiding in the water. Where
would you click next? Perhaps one of the two revealed ships is the
size-three submarine. In that case you want to click on the first cell
in the third row, or on the third cell in the fifth row. Or maybe these
are the two size-two patrol boats. This would mean the submarine is
hiding in some other place, and for it not to touch sides with the boats
this must be in the first row. The hit probability map in the second
column visualizes this reasoning, by plotting the normalized probability
that a ship is hidden in different cells. As can be seen, the highest
probabilities are to the sides of the revealed ships, and in the first
row.

In the third column we plot the distribution of clicks for non-pretend
players. This distribution is in agreement with the hit probability map
(board A: \(r = .81\), 95\% CI \([.55, .93]\), \(t(16) = 5.48\),
\(p < .001\); board B: \(r = .87\), 95\% CI \([.68, .95]\),
\(t(16) = 7.01\), \(p < .001\)). Finally, in the fourth column we plot
the distribution of cell selections for pretend players. Although
noisier, this distribution is also in agreement with the hit probability
map (board A: \(r = .49\), 95\% CI \([.03, .78]\), \(t(16) = 2.27\),
\(p = .037\); board B: \(r = .73\), 95\% CI \([.40, .89]\),
\(t(16) = 4.28\), \(p = .001\)), and more importantly, with the hit
distribution of non-pretenders for the same board configuration (board
A: \(r = .63\), 95\% CI \([.31, .82]\), \(t(23) = 3.89\), \(p = .001\),
board B: \(r = .87\), 95\% CI \([.73, .94]\), \(t(23) = 8.60\),
\(p < .001\)).

\begin{CodeChunk}
\begin{figure}[tb]
\includegraphics[width=1\linewidth]{../figures/Half_games} \caption[Half games]{Half games. First column: the two boards, as they appeared to pretenders. Second column: objective hit probability, given game state. Third and fourth columns: empirical click probabilities in non-pretend and pretend games, respectively.}\label{fig:half-games}
\end{figure}
\end{CodeChunk}

\hypertarget{judge-trials-1}{%
\subsection{Judge trials}\label{judge-trials-1}}

Despite the fact that pretenders were overall successful in mimicking
patterns of cell selection and decision latency, they also left many
clues for an independent observer to pick up on: they were slower,
displayed exaggerated effects of decision uncertainty and hit success on
response time, and were less optimal in choosing where to click next. To
illustrate, a support vector machine (SVM) algorithm reached an accuracy
level of 72\% in linearly classifying condition (pretend / non-pretend)
based on just three basic summary features: median decision latency,
game optimality score, and number of irrational clicks per game.

Despite this, participants were never significantly above chance in
judging which of the two presented games came from a pretender (mean
accuracy: 51\%; t test against 50: \(t(499) = 1.45\), \(p = .147\)).

For a subset of 288 players, one of the five boards presented in the
judging block was already presented both in the pretend and in the
non-pretend blocks. Still, even this subset of participants who had just
experienced a board both as pretenders and as non-pretenders were not
significantly above chance in telling which of two other players had
hints for this same board (mean accuracy: 55\%, \(p=.141\)).

\hypertarget{are-good-pretenders-also-better-judges}{%
\subsection{Are good pretenders also better
judges?}\label{are-good-pretenders-also-better-judges}}

The abilities to pretend and to detect pretense in others both rely on
some form of theory of mind: reasoning about counterfactual knowledge
states of oneself in the first case, and of others in the second. For
example, in scrub jays, individual birds that pilfered another bird's
caches were also more careful in hiding their own food, suggesting an
ability to project from their own experience of stealing to the behavior
of others (Emery \& Clayton, 2001). Along similar lines, we reasoned
that those participants whose pretend games better resembled the games
of genuine players would also be more accurate in detecting pretense in
others. To our surprise, this was not at all the case (\(r = -.03\),
95\% CI \([-.13, .07]\), \(t(352) = -0.56\), \(p = .573\)). Furthermore,
we find no significant correlation between participants' accuracy in
detecting the pretender and their decision optimality in pretend games
(measured as the mean rank posterior probability of their cell
selections; \(r = .01\), 95\% CI \([-.08, .10]\), \(t(478) = 0.18\),
\(p = .854\)), nor with the cost to optimality relative to non-pretend
games (\(r = .02\), 95\% CI \([-.07, .11]\), \(t(478) = 0.47\),
\(p = .641\)).

Despite our considerable sample size of 500 participants, we are
cautious in making strong claims based on the absence of a correlation
between pretense and pretense detection abilities in our sample. We
note, however, that previous developmental research has identified
weaker correlations with theory of mind measures for a tendency to
intentionally deceive another player in a game setting, compared to an
ability to detect and understand deception, potentially indicating
partly different mechanisms (Shultz \& Cloghesy, 1981).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Many tasks (including writing papers for conferences) require us to
overcome the `curse of knowledge' --- the challenge of putting ourselves
in the shoes of someone who doesn't know what we know. This is just an
example of a more general fact: the ability to simulate counterfactual
knowledge states is necessary for efficient and flexible communication.
To do this, we must have an internal self-model that can generate
thoughts and behavior for hypothetical beliefs, desires, and
perceptions.

Research on Bayesian Theory of Mind provides some support for the
existence of such a model by measuring subjects' ability to infer
beliefs and desires from observed behavior, either explicitly (Baker,
Jara-Ettinger, Saxe, \& Tenenbaum, 2017; Baker, Saxe, \& Tenenbaum,
2009), or implicitly (Liu, Ullman, Tenenbaum, \& Spelke, 2017; Onishi \&
Baillargeon, 2005). Here we have put forward a complementary approach:
Asking participants to generate behavior based on a counterfactual
mental state - in this case, a counterfactual knowledge state in which a
known piece of information is unknown. Instead of relying on model
inversion (e.g., ``Which belief states would give rise to this
behavior?''), we ask participants to run the model forward, taking
beliefs and desires as input and producing behavior as output. Due to
the unconstrained space of possible behaviors in our task (cell
selections x decision latencies), successfully pretending not to know
demands a rich model of cognition, and is much harder to achieve based
on a quasi-scientific theory of mental states (Gopnik \& Wellman, 1994).

Our subjects rose to the challenge. Their pretense behavior, although
not perfect, closely resembled that of non-pretenders --- not only in
gross measures such as total number of clicks, but also in subtler
patterns of cell selection and decision latency. Our interpretation of
this finding is that participants engaged in rich and relatively
accurate counterfactual self-simulation: ``What would I have done if I
didn't have this piece of knowledge?''

An alternative interpretation is that instead of simulating a
counterfactual knowledge state, participants actively suppressed or
ignored unwanted knowledge, such that their entire cognitive machinery
was available to play the game. This does not require any self-modeling
or self-simulation, beyond the knowledge that suppressing knowledge can
be beneficial for pretending not to know something. While we cannot
fully rule out this interpretation, we think it is unlikely to explain
participants' pretense ability in our data, for two indirect reasons.
First, suppressing thoughts on demand is notoriously difficult (Wegner,
Schneider, Carter, \& White, 1987), and our first-person experience of
performing the task is that suppressing knowledge of ship locations is
hardly possible. Second, when asked how they had performed the task in a
debrief question, the responses of most participants were aligned with a
self-simulation account. In future work, we intend to more directly rule
out this knowledge-suppression hypothesis, and well as to generalize our
results to other games.

In conclusion, we propose that pretense behavior provides a unique
opportunity to directly test the richness and accuracy of subjects'
models of their own cognitive processes, and find evidence for accurate
pretending that surprisingly does not translate to an ability to detect
pretense. While more work is needed to exclude alternative explanations,
we interpret our findings as providing preliminary evidence for a
remarkable capacity for counterfactual self-simulation.

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}

\noindent

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-audinot2014optimal}{}}%
Audinot, M., Bonnet, F., \& Viennot, S. (2014). Optimal strategies
against a random opponent in battleship. \emph{The 19th Game Programming
Workshop}.

\leavevmode\vadjust pre{\hypertarget{ref-baker2017rational}{}}%
Baker, C. L., Jara-Ettinger, J., Saxe, R., \& Tenenbaum, J. B. (2017).
Rational quantitative attribution of beliefs, desires and percepts in
human mentalizing. \emph{Nature Human Behaviour}, \emph{1}(4), 1--10.

\leavevmode\vadjust pre{\hypertarget{ref-baker2009action}{}}%
Baker, C. L., Saxe, R., \& Tenenbaum, J. B. (2009). Action understanding
as inverse planning. \emph{Cognition}, \emph{113}(3), 329--349.

\leavevmode\vadjust pre{\hypertarget{ref-emery2001effects}{}}%
Emery, N. J., \& Clayton, N. S. (2001). Effects of experience and social
context on prospective caching strategies by scrub jays. \emph{Nature},
\emph{414}(6862), 443--446.

\leavevmode\vadjust pre{\hypertarget{ref-emery2004mentality}{}}%
Emery, N. J., \& Clayton, N. S. (2004). The mentality of crows:
Convergent evolution of intelligence in corvids and apes.
\emph{Science}, \emph{306}(5703), 1903--1907.

\leavevmode\vadjust pre{\hypertarget{ref-frith2005theory}{}}%
Frith, C., \& Frith, U. (2005). Theory of mind. \emph{Current Biology},
\emph{15}(17), R644--R645.

\leavevmode\vadjust pre{\hypertarget{ref-gopnik1994theory}{}}%
Gopnik, A., \& Wellman, H. M. (1994). The theory theory. In \emph{An
earlier version of this chapter was presented at the society for
research in child development meeting, 1991.} Cambridge University
Press.

\leavevmode\vadjust pre{\hypertarget{ref-hall2017cooperation}{}}%
Hall, K., \& Brosnan, S. F. (2017). Cooperation and deception in
primates. \emph{Infant Behavior and Development}, \emph{48}, 38--44.

\leavevmode\vadjust pre{\hypertarget{ref-liu2017ten}{}}%
Liu, S., Ullman, T. D., Tenenbaum, J. B., \& Spelke, E. S. (2017).
Ten-month-old infants infer the value of goals from the costs of
actions. \emph{Science}, \emph{358}(6366), 1038--1041.

\leavevmode\vadjust pre{\hypertarget{ref-onishi2005fifteen}{}}%
Onishi, K. H., \& Baillargeon, R. (2005). Do 15-month-old infants
understand false beliefs? \emph{Science}, \emph{308}(5719), 255--258.

\leavevmode\vadjust pre{\hypertarget{ref-shultz1981development}{}}%
Shultz, T. R., \& Cloghesy, K. (1981). Development of recursive
awareness of intention. \emph{Developmental Psychology}, \emph{17}(4),
465.

\leavevmode\vadjust pre{\hypertarget{ref-sodian1991early}{}}%
Sodian, B., Taylor, C., Harris, P. L., \& Perner, J. (1991). Early
deception and the child's theory of mind: False trails and genuine
markers. \emph{Child Development}, \emph{62}(3), 468--483.

\leavevmode\vadjust pre{\hypertarget{ref-wegner1987paradoxical}{}}%
Wegner, D. M., Schneider, D. J., Carter, S. R., \& White, T. L. (1987).
Paradoxical effects of thought suppression. \emph{Journal of Personality
and Social Psychology}, \emph{53}(1), 5.

\leavevmode\vadjust pre{\hypertarget{ref-wimmer1983beliefs}{}}%
Wimmer, H., \& Perner, J. (1983). Beliefs about beliefs: Representation
and constraining function of wrong beliefs in young children's
understanding of deception. \emph{Cognition}, \emph{13}(1), 103--128.

\end{CSLReferences}

\bibliographystyle{apacite}

\end{document}

### 2023-01-04 docs/pre-registrations/hangman_PR.Rmd
pre-registrations

---
title             : "Pretending not to know, Hangman: pre-registration document"
shorttitle        : "PNTN HANGMAN PRE-REGISTRATION"

author: 
  - name          : "Matan Mazor"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "12 Queen Square, WC1N 3BG London"
    email         : "mtnmzor@gmail.com"
  - name          : "Ian Phillips"
    affiliation   : "2"
  - name          : "Chaz Firestone"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "University College London"
  - id            : "2"
    institution   : "Johns Hopkins University"

authornote: |
  The authors declare no competing interests

abstract: |
  Generally, what we do depends on what we know. But sometimes we try to appear not to know something that we really do know. Such pretense behavior relies on counterfactual self-simulation — an understanding of how we would behave if our knowledge were different — and so provides an opportunity to investigate how well people can emulate a hypothetical knowledge state. Surprisingly, despite its immediate relevance to metacognition and theory of mind, little research has focused on quantifying pretense accuracy, relative to non-pretense behaviour. In previous experiments, we examined the ability to both produce and detect pretense behaviour using the game "Battleship" — normally played by searching for ships hidden behind cells in a grid. Here, we test the generalizability of our findings to a game where players reveal a hidden word by guessing letters ("Hangman"). In 'pretend' blocks, subjects will know what the hidden word is, but will try to play as if they lack this knowledge. Similar to our analysis of Battleship games, we will quantify subjects' self-simulation ability by comparing their behaviour in pretend and non-pretend games. We will further test players' ability to detect pretense, and examine the association between pretense ability and pretense-detection ability.

keywords          : "Pretense, self-simulation, metacognition, theory of mind"
wordcount         : "3634"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib");
library('reticulate')
library('tidyverse')
library('broom')
library('cowplot')
library('MESS') # for AUCs
library('lsr') # for effect sizes
library('pwr') # for power calculations
library('brms') # for mixed effects modeling
library('BayesFactor') # for Bayesian t test

N <- 500;
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r load-data, warning=FALSE}
E3.df <- read.csv('..\\experiments\\Hangman\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
  rename(subj_id = participant_number) %>%
  mutate(subj_id = factor(subj_id)) %>%
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part))

E3.export <-read.csv('..\\experiments\\Hangman\\data\\prolific_export_batch1.csv') %>%
  rbind(read.csv('..\\experiments\\Hangman\\data\\prolific_export_batch2.csv'));
```

# Motivation

<!-- The ability to reason about hypothetical mental states is critical for inferring goals and beliefs from observed behaviour [@baker2009action; @baker2011bayesian], predicting behaviour based on goals and belief states [@wimmer1983beliefs; @wellman2001meta], pre-commitment in action planning [@moran2021model], inference about the absence of memories [@dodson2002false], [...].  -->

The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory]. This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation]. This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own. Moreover, deception often involves pretense behaviour, which in turn relies on an ability to simulate and mimic one's own behaviour under a counterfactual belief state. For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") — you also need to convincingly simulate and mimic your hypothetical behaviour had you not known about the party ("Where would I look first? What would I say? How long would it take me to recover from the surprise?"). This reliance of pretense behaviour on self-simulation makes it an ideal opportunity to examine metacognitive knowledge about one's own mental states, and the potential reliance of this knowledge on a self-simulation. By comparing non-pretend and pretend behaviour, we can ask which aspects of their cognitive processes subjects can simulate, and which aspects are not represented in their mental models of their own cognition.

In Exp. 1 (N=100, exploratory) and 2 (N=500, pre-registered), we examined pretense behaviour using an online version of the game "Battleship". We quantified pretense quality by comparing patterns of cell selections in pretend and non-pretend games. We observed a striking ability to simulate ignorance: pretenders made rational cell selections given the limited knowledge state they simulated, and the timing of their selections qualitatively matched click-time patterns found in standard, non-pretend games. In a final 'judge' block, subjects were unable to identify pretend games in a two-alternative forced choice discrimination task, and the accuracy of their judgments was uncorrelated with the quality of their pretense.

In Exp. 3 (N=100, exploratory), we tested how general this pretense capacity is by comparing pretend and non-pretend games in an online version of the game "Hangman". In Exp. 4, we will replicate and extend our results in a larger sample (N=500).

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

The research complies with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University. Participants will be recruited via Prolific and will give informed consent prior to their participation. They will be selected based on their acceptance rate (>95%), being native English speakers, and being located in the USA. We will collect data until we reach `r N` included participants. The entire experiment will take 20 minutes to complete (median completion time in our pilot data: `r round(E3.export%>%pull(time_taken)%>%median()/60)` minutes). Participants will be paid 3.15 USD for their participation, equivalent to an hourly wage of approximately 9.50 USD, in addition to a bonus payment.

## Procedure

```{r design, echo=FALSE, fig.cap="Experimental Design: players will play 12 games of the game Hangman in two conditions presented in two blocks of 6 games. The order of blocks will be randomized between players. A: In non-pretend games, players will start the game not knowing the identity of the hidden word, and try to reveal the word with as few letter guesses as possible. B: In pretend games, players will be told the identity of the word, and their task will be to play the game as if they do not have this information. Before pretending, players will verify they had read the hidden word by typing it. C: Lastly, in 5 judge trials players will observe a replay of the game of a previous player, and try to tell whether the player was pretending or not.", out.width = '100%'}
knitr::include_graphics("figures/Hangman_design.png")
```

The first instructions screen will inform participants that the experiment, based on the game Hangman, has three parts, and that the points they will accumulate will translate to a monetary bonus payment. They will then be presented with a leaderboard of previous players. Then, the rules of the game will be presented:

> In the following game, your task is to reveal a hidden word or phrase by guessing letters. What makes the game difficult is that you can’t see the word; all you can see is a row of squares - a square for each letter. Use your mouse to make letter guesses. We will have five types of words: body parts, numbers, US states, fruit, and famous people. You will start each game with 15 points and lose one point for every guess of a letter that is not in the word.

We will then explain that "the words in this game are the kind of words that will be familiar to most English-speaking fifth-graders. We didn't pick any strange or particularly difficult words".

Once they respond correctly to a multi-choice comprehension question ("the goal of the game is to...": "reveal the word with as few letter guesses as possible"), participants will play a practice round, revealing the word PIGEON (see Fig. \@ref(fig:design)A).

After the main instructions, comprehension question and practice round, participants will complete one pretend and one non-pretend block, each followed by one half-game (see below for details). The order of pretend and non-pretend blocks will be counterbalanced between participants. Each block will comprise five games played with five out of ten different words, and one half-game. The allocation of words to conditions will be randomized between participants, with the constraint that both pretend and non-pretend blocks will include exactly one word from each category. The order of words within a block will be randomized, except for the half-game, which will always be delivered at the end.

The ten words will include two number words (ELEVEN, NINETY SIX), two famous people (DALAI LAMA, TAYLOR SWIFT), two fruits (STRAWBERRY, LEMON), two body parts (TOOTH, HEAD), and two US states (MONTANA, IOWA).

### Non-pretend games

In non-pretend games, participants will reveal a hidden word with as few letter guesses as possible. An online counter of the number of points will be displayed on the screen, deducting one point for every guess of a letter that is not in the target word. After each game, feedback will be given about the number of points obtained.

After completing the five games, participants will perform one half-game (see below for details).

### Pretend games

Participants will be given the following instructions:

> In the next part of the experiment, you'll play 6 games where you reveal a hidden word by guessing letters.

> However, this time your goal is different.

> In this round, we’re going to tell you the word in advance, but **we want you to act like you don’t know this information**.

> To see how good you are at this, we're going to compare your games to the games of people who played normally, without knowing what the word was, and see how similar they are. We will measure which letters you click and the timing of your guesses; if your clicks look similar to people who played like normal (trying to reveal the word with as few guesses as possible, but without any hints), you'll get bonus points. But if your games look different, you won't get these bonus points. Your number of clicks in this part will not affect your bonus. Only your ability to play like you didn't see the word in advance.

After one practice round, pretending not to know that the hidden word is PIGEON, and one comprehension question ("In this part of the experiment my goal is to...": "play the game as if I don't know what the word is so that I look like someone who had no hints"), participants will play five pretend games. Each game will be preceded by a short message informing subjects about the identity of the target word. To start pretending, players will be asked to type in the target word on their keyboard. After pretending, we will remind players that a game that looks similar to the games of participants who had no hints will be awarded 10 bonus points.

After completing the five games, participants will perform one half-game (see below for details).

### Half games

In order to directly compare participants' pretend and non-pretend games for identical belief states (true or pretended knowledge about the identity of the word), we will ask participants to also complete one pretend and one non-pretend game, given a partly finished game with some letters already guessed (they will be told that the computer made these guesses). The two half-game words will be one fruit: PAPAYA or BANANA, with guessed letters [A, E, I, O, M, T], and one body part: HAND, or HAIR with guessed letters (A, E, O, M, T, H, P). The assignment of category (fruit or body part) to condition (pretend and non-pretend), as well as the identity of the target word within each category (e.g., PAPAYA or BANANA), will be randomized between participants.

Instructions for the non-pretend half-game will be:

> For the next game, the computer chose the first letters for you; you can take over from where it left off. Your challenge is to  complete the game. Just like in the previous games, here also you will lose one point for each letter that you guess and is not in the word.

Instructions for the pretend half-game will be:

> For the next game, the computer chose the first letters for you; you can take over from where it left off. Just like in the previous games, here also you will know what the word is, but your bonus points will depend on your ability to play as if you didn’t know the word.

### Judge trials

In the final part of the experiment, participants will observe five games of previous players and determine who had hints and who didn't. Instructions for this part are:

>  In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of the players were shown the word (but acted like they weren't). We will show you 5 replays of the games of previous players. Your task is to decide whether they played normally or pretended. For each game that you get right, you will receive 10 points. Good luck!

Then, on each judge trial, one game of a previous player will be replayed in real time, with the target word presented above. For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) will be chosen for presentation in this part. For both pretend and non-pretend games, only games shorter than 1.5 minutes will be presented. Judge participants will indicate their decision by pressing the P and N keys on their keyboard. After making a decision, participants will be informed whether they will receive the 10 points. Whenever a pretend game is classified as a non-pretend game, they will be informed that the pretender will receive these 10 points instead of them.

Lastly, participants will be asked the following debrief questions:

> Did you have a strategy that you used for pretending you did not see the word? What was most difficult about pretending? How about telling between players who pretenders and who played for real - did you have a strategy for that?

And:

> We would appreciate it if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.

### Randomization

The order and timing of experimental events will be determined pseudo-randomly by the Mersenne Twister pseudorandom number generator, initialized in a way that ensures registration time-locking [@mazor2019novel].

## Data analysis

### Rejection criteria

Participants will be excluded if all five of their pretend games were all hits. Unlike Battleship, here we decided not to exclude pretend games that were all hits. In Hangman, getting a perfect score is possible even when not pretending, as was the case in 27% of all non-pretend games in exploratory Exp. 3.

### Data preprocessing

### Hypotheses and analysis plan

This study is designed to explore subjects' capacity for self-simulation under a counterfactual knowledge state, and the limits of this capacity. In particular, we are interested in comparing non-pretend and pretend games and identifying similarities and differences. We will also assess subjects' capacity to discriminate between pretend and non-pretend games, and how this capacity relates to their ability to pretend.

Unless otherwise specified, all hypotheses will be tested using a repeated measures t-test, with a significance level of 0.05.

All hypotheses will be tested separately for the group of participants that pretended in the first and in the second block. If results for the two groups diverge, analysis will be repeated on the first block of all participants (pretend games of pretend-first participants and non-pretend games of pretend-second participants), in a between-subjects manner, and results will be interpreted in light of this analysis.

*Hypothesis 1 (game duration)*: We will test the null hypothesis that pretend games take as long as non-pretend games. For each participant, the difference between median game duration in pretend and non-pretend games will be extracted. Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 2 (first click latency)*: We will test the null hypothesis that the first click in pretend games takes as long to execute as in non-pretend games. For each participant, a difference between median first click latency in pretend and non-pretend games will be extracted. Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 3 (number of misses)*: We will test the null hypothesis that pretend games are as long, in terms of total number of letter misses, as non-pretend games. For each participant, we will extract the difference between the median number of letter selections that do not appear in the target word in pretend and non-pretend games. Subject-wise differences will then be carried over to a t-test at the group level.

*Hypotheses 4-6 (click latency by click outcome)*: Letter selections will be classified based on whether they resulted in revealing a letter that appears in the target word (*hit*) or not (*miss*). We will then contrast median click latency between hits and misses separately for pretend and non-pretend games. We will test for a difference in selection latency as a function of outcome on pretend games, on non-pretend games, and for an interaction effect between click outcome (hit or miss) and condition (pretend or non-pretend) on click latency. The same analysis will then be repeated by classifying letter selections on the basis of whether the previous selection resulted in a hit or a miss (*Hypothesis 5*), and whether the next selection will result in a hit or a miss (*Hypothesis 6*).

#### Model based analysis

The following analyses are designed to test for differences in game optimality between pretend and non-pretend games, and for a relationship between decision difficulty and click latency. To do so, we approximate optimal behaviour by approximating the posterior probability that a letter appears in the word, given available information. Critically, in modelling pretend games we do not treat hints as part of this available information, because an optimal player should ignore this information in choosing where to click next. Given this posterior, a rational player should choose letters that have a high posterior probability of appearing in the word.

To approximate the posterior probability of letters given a game state, we will follow the following procedure:

1. We will use the category information (e.g., 'a fruit'), to obtain a probability-weighted list of category-compatible words (or names, in the case of famous people). The lists were obtained in the following way: for US states and number names, we used an exhaustive list (in the case of numbers, of numbers of 1-2 words), fruit names were taken from Wikipedia ([simple.wikipedia.org/wiki/List_of_fruits](https://simple.wikipedia.org/wiki/List_of_fruits)), famous people names from a [crowdsourced document](https://docs.google.com/spreadsheets/d/1mWez4KH0wueV_SSzdN16MMUG4qCa_nFFHToJcUvtJIA/edit#gid=4), and body-part words from existing prototypicality norms [@uyeda1980prototypicality]. We extended Uyeda and Mandler's list of body part words by adding all body part words commonly appearing in lists on the internet. In the case of famous people, names were given prior probabilities in proportion to the number of visits their Wikipedia entries received in 2021. The top 100 most popular entries were given a prior probability twice that of the next 100, three times that of the next 100, and so forth. All entries from 600 and on were given the same prior probability, seven times smaller than that of entries at the top 100 positions. In the cases of fruits and body parts, we used prototypicality norms from Uyeda and Mandler (1980) to assign higher prior probability to more prototypical words (mapped to lower numbers on the 1-7 scale used by Uyeda and Mandler). Words that were not included in the norm were given the maximum score (that is, the lowest perceived prototypicality), of 7. Body parts and fruits were included in both singular and plural forms. The plural forms were assigned a prototypicality score of 100, due to the fact that category names were presented in the singular form ('a fruit' and 'a body part'). We then took the reciprocal for each score, and normalized it by the total sum to get a probability distribution over words $p(w)$. The full prior distributions for each category are included in this pre-registration.

2. The likelihood of a game state given a target word $p(s|w)$ equals 0 when the word is inconsistent with the information available to the player (this includes the word length, the identity of letters that do not appear in the word based on previous guesses, and the identity and position of letters that do appear in the word based on previous guesses). When consistent, the likelihood is a non-zero quantity that is equal for all consistent words. Bayes rule will be used to extract the posterior over words given game state $p(w|s)=\frac{p(s|w)p(w)}{\Sigma_{w'}{p(s|w')p(w')}}$. The full non-zero $p(w|s)$ distributions for the initial states of all games are included in this pre-registration.

3. The probability that an individual letter appears in the target word $p(l|s)$ is the sum of posterior probabilities of words that contain this letter $p(l|s)=\Sigma_{w}p(w|s)\times includes(w,l)$ where $includes(w,l)$ returns 1 if $l$ appears in $w$ and 0 otherwise.

4. A posterior probability over letter selections will be obtained by dividing the probabilities of individual letters by their total sum $\frac{p(l|s)}{\Sigma_{l}p(l|s)}$.

To quantify this notion of optimality, before each cell selection we will compute the posterior probability that each of the unclicked letters appears in the target word, given the game state. Then, we will rank letters from highest to lowest according to their posterior probability and record the rank of the chosen cell.

*Hypothesis 7 (game optimality):* We will test the null hypothesis that pretend games are as optimal as non-pretend games. Mean click optimality will be contrasted between the pretend and non-pretend games of each player.  Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 7B (letter frequency heuristic):* We will repeat the game optimality analysis, but this time using the ranked frequency of the selected letter in the English language (from high to low: E, T, A, O, I, N, S, H, R, D, L, C, U, M, W, F, G, Y, P, B, V, K, J, X, Q, Z, obtained from Wikipedia). We will test the null hypothesis that letters selected in pretend games are of similar frequency to those selected in non-pretend games. Mean ranked probability will be contrasted between the pretend and non-pretend games of each player.  Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 8 (entropy-RT relation):* The entropy of the posterior map will be extracted before each cell selection using the following formula: $H(p)=\sum_{i=1}^{N} p_{i}\times log p_{i}$. We will test for within-subject associations between click latency and the linear and quadratic expansions of posterior entropy in pretend and in non-pretend games by fitting a multiple regression model to the data of each subject. We will contrast subject-wise beta values for pretend and non-pretend games to test the null hypothesis that the association between posterior entropy and click latency is the same in both conditions.

#### Judge accuracy

*Hypothesis 9 (judge accuracy):* Judge accuracy will be quantified as the number of correct guesses in the judge block, divided by 5. This quantity will be extracted separately for each subject and contrasted against chance (50%) at the group level.

*Hypothesis 9B (judge bias):* Judge bias will be quantified as the number of 'pretend' guesses in the judge block, divided by 5. This quantity will be extracted separately for each subject and contrasted against chance (50%) at the group level.

*Hypothesis 10 (judge-pretend relation):* Pretense quality will be quantified as the number of pretend games that were misclassified by other participants as non-pretend games, divided by the number of pretend games that were presented to other participants. Due to randomization, this denominator is different for different participants. We will test for a correlation between pretense quality and judge accuracy at the group level. To maximize statistical power, this hypothesis will be tested on the pooled data from both groups of participants.

#### Half games

Our main focus in analyzing half-games will be the identity of the first letter selection. We will ask whether the distribution over letter selections is different in pretend and non-pretend games, and within pretend games, between the two words of each category.

*Hypothesis 12 (irrational letter selections):* Irrational letter selections are ones that are inconsistent with the current game state, such that $p(l|s)=0$. We will use a binomial test to test the null hypothesis that the group-level proportion of irrational letter selections is similar in pretend and non-pretend games.

*Hypothesis 13 (target word effect):* For this analysis, we will focus on half-games where the first letter selection was not irrational, and ask whether knowledge of the target word had an attraction or repulsion effect (or no effect at all) on letter selection. Within each word pair (HAIR and HAND, PAPAYA and BANANA), we will contrast the proportion of letter selections that are consistent with the first target word (I and R for HAIR, P and Y for PAPAYA) or the second target word (N and D for HAND, B and A for BANANA) in pretend against non-pretend games, and within pretend games, as a function of the target word. For example, we will compare the proportion of non-pretenders who clicked on B or N among those who clicked on B,N,P or Y when encountering _A_A_A, and test whether this proportion is similar in pretend games for the same game state _A_A_A, and within pretend-games, whether it is different as a function of the target word (PAPAYA or BANANA).

## Power calculations and sample size justification

With approximately `r N/2` participants in each group (allocation to group will be determined pseudorandomly), we will have 95% statistical power to detect an effect size of `r pwr.t.test(n=N/2,sig.level=0.05,power=0.95,type='paired')$d%>%printnum()` standard deviations in a repeated measures t-test (Hypotheses 1-9). In case of a conflict between the results of the two groups, a between-subjects comparison will be performed, using data from the first block only (pretend games from the pretend-first group and non-pretend games from the pretend-second group). For this between-subjects t-test, we will have 95% statistical power to detect an effect size of `r pwr.t2n.test(n1=N/2,n2=N/2,sig.level=0.05,power=0.95)$d%>%printnum()` standard deviations.

With `r N` participants in both groups combined, we will have 95% statistical power to detect a Pearson correlation of `r pwr.r.test(n=N,sig.level=0.05,power=0.95)$r%>%printnum()` (Hypothesis 10).

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2023-01-17 docs/hangman.Rmd
working version

---
title             : "Pretending not to know"
shorttitle        : "PNTN EXP2 results"

author: 
  - name          : "Matan Mazor"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "12 Queen Square, WC1N 3BG London"
    email         : "mtnmzor@gmail.com"
  - name          : "Ian Phillips"
    affiliation   : "2"
  - name          : "Chaz Firestone"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "University College London"
  - id            : "2"
    institution   : "Johns Hopkins University"

authornote: |
  Enter author note here.

abstract: |
  Generally, what we do depends on what we know. But sometimes we try to appear not to know something that we really do know. Such pretense behavior relies on counterfactual self-simulation — an understanding of how we would behave if our knowledge were different — and so provides an opportunity to investigate how well can people emulate a hypothetical knowledge state. Surprisingly, despite its immediate relevance to metacognition and theory of mind, little research has focused on quantifying pretense accuracy, relative to non-pretense behaviour. In a first exploratory experiment, we examine the ability to both produce and detect pretense behaviour using the game "Battleships" — normally played by searching for ships hidden behind cells in a grid. In 'pretend' blocks, subjects were told where all the ships are but asked to play as if they lacked this knowledge. Analyzing non-pretend (regular) games of the same subjects, we identify robust behavioural patterns in the location and timing of cell selections, including serial dependencies in hit probability and decision time, and an association between decision time and the entropy of the posterior distribution over candidate actions. Critically, pretend games demonstrate similar, but exaggerated, behavioural patterns. Furthermore, relative to the modeled behaviour of a near-optimal Bayesian agent, pretend behaviour is markedly less optimal than non-pretend behaviour. This is because pretend play often doesn't make sense given the limited knowledge players pretend to have. However, despite these striking differences, independent "judge" participants were completely unable to discriminate the games of pretenders from non-pretenders.

keywords          : "Pretense, self-simulation, metacognition, theory of mind"

wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library(groundhog)

groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'reticulate',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel', #for ggplot words
  'caret' #for cross validation
  'kernlab' #for SVM
), '2022-12-01')

r_refs("r-references.bib")
use_condaenv("py3.8", required = TRUE)

```

```{python, echo=FALSE, eval=FALSE}
import json
import pandas as pd
import numpy as np
from os import path as path

def to_csv(filename):
   dfs=[]
   with open(filename+'.txt', encoding="utf8") as json_file:
      for i,line in enumerate(json_file):
         dfs.append(pd.read_json(line))
   group_df = pd.concat(dfs);
   cols = list(group_df.columns.values);
   cols.remove('hover_log');
   group_df = group_df[cols]
   group_df.to_csv(filename+'.csv', index=False)
   return(group_df)

to_csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1')
to_csv('..\\experiments\\Hangman2\\data\\jatos_results_batch2')

```

```{r load-data, warning=FALSE}
e4.df <- read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch2.csv',na.strings=c(""," ","NA")) %>%
  rbind(read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
  rename(subj_id = participant_number) %>%
  mutate(subj_id = factor(subj_id)) %>%
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
  filter(PROLIFIC_PID != '5ec4a156bc5aac3819ac52f2' & #reported some of the letters not showing up on their screen
        PROLIFIC_PID != '58211fc787f6b90001f13f9' & #reported some of the letters not showing up on their screen
        PROLIFIC_PID != '6149256f6335b06ade3723e0' &  # due a PROLFIIC bug, participated twice
        PROLIFIC_PID != '615ddab1e4f013092538b6c5')# due a PROLFIIC bug, participated twice

e4.export1 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export1.csv');
e4.export2 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export2.csv');
e4.export3 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export3.csv');
e4.export4 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export4.csv');
e4.export5 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export5.csv');
e4.export6 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export6.csv');

```

```{r subject-comments, eval=FALSE}
e4.comments <- e4.df %>%
  filter(is.na(test_part) & substr(responses,1,1)=='{') %>%
  dplyr::select(subj_id,PROLIFIC_PID,responses) %>%
  mutate(comments = map(responses, ~fromJSON(.) %>%
                          as.data.frame()))%>%
  unnest(comments)%>%
  dplyr::select(subj_id,PROLIFIC_PID,worker_comments)%>%
  filter(!is.na(worker_comments) & worker_comments!="") %>%
  mutate(worker_comments = gsub("[\r\n]", "",worker_comments))

e4.debrief <- e4.df %>%
  filter(is.na(test_part) & substr(responses,1,1)=='{') %>%
  dplyr::select(subj_id,PROLIFIC_PID,responses) %>%
  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
  unnest(comments)%>%
  dplyr::select(subj_id,PROLIFIC_PID,debrief)%>%
  filter(!is.na(debrief) & debrief != "")%>%
  mutate(debrief = gsub("[\r\n]", "",debrief))

```

<!-- ## Strategy use: -->

<!-- After completing the judgment blocks, participants were asked  -->

<!-- > "Did you have a strategy that you used for pretending you had no hints? How about telling between players with and without hints - did you have a strategy for that?" -->

```{r results='asis'}
# string.list <- e4.debrief%>%pull(debrief)
# cat(paste('-', string.list), sep = '\n')
# cat(paste('-', string.list), sep = '\n',file='strategy_hangman2.txt')
```

## General comments

Then, we asked participants for more general feedback about the task:

> "That's it! Before we thank you, we would appreciate if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data."

```{r results='asis'}
# string.list <- e4.comments%>%pull(worker_comments)
# cat(paste('-', string.list), sep = '\n')
```

```{r comprehension-checks}
e4.num_instructions <- e4.df %>%
  group_by(subj_id)%>%
  summarise(pretend_instructions = mean(pretend_instructions),
            nonpretend_instructions = mean(nonpretend_instructions))

```

```{r format-data}

e4.click_df <- e4.df %>%
  dplyr::select(subj_id, 
         test_part, 
         category,
         word,
         num_clicks,
         click_log,
         genuine_first,
         trial_type) %>%
  rowwise()%>%
  mutate(num_hits = strsplit(gsub(' ','',word),split='')[[1]]%>%unique()%>%length(),
         genuine_first= genuine_first=='True',
         num_misses = num_clicks-num_hits,
         click_log = gsub("\'","\"", click_log),
         click_log = gsub("None","null", click_log),
         word = factor(word,levels=c(
           'eleven',
           'ninety six',
           'dalai lama',
           'taylor swift',
           'strawberry',
           'lemon',
           'tooth',
           'head',
           'iowa',
           'montana'
         ))) %>%
  filter(test_part=='pretend' | test_part=='nonpretend') %>%
  filter(trial_type=='Hangman')

e4.median_clicks <- e4.click_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(test_part,num_clicks);

e4.median_misses <- e4.click_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(num_misses=median(num_misses)) %>%
  spread(test_part,num_misses);

e4.lucky_pretenders <- e4.median_misses%>%filter(pretend==0)%>%pull(subj_id)%>%unique();
e4.lucky_nonpretenders <- e4.median_misses%>%filter(nonpretend==0)%>%pull(subj_id)%>%unique()

# 
# median_simulated = num_clicks%>%median()
p<- e4.click_df %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=num_misses,fill=test_part))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of misses', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()+
    # geom_vline(xintercept= median_simulated)+
    facet_wrap(~genuine_first, nrow=2);

ggsave('figures/Hangman2/num_misses.png',p,width=5,height=3,dpi=300)

e4.misses_per_word <- e4.click_df %>%
  group_by(test_part,word) %>%
  summarise(num_misses=median(num_misses)) %>%
  spread(test_part,num_misses)

p <- e4.misses_per_word %>%
  ggplot(aes(x=nonpretend,y=pretend, label=word)) +
  geom_abline(intercept=0,slope=1) +
  geom_point()+
  geom_text_repel()+
  coord_fixed(ratio=1)+
  scale_x_continuous(limits=c(0,5.5), breaks=seq(0,5))+
  scale_y_continuous(limits=c(0,5.5), breaks=seq(0,5)) +
  theme(legend.position = 'none')+
  theme_bw() +
  labs(title='number of misses')

ggsave('figures/hangman2/num_misses_per_word.png',p,width=5,height=5,dpi=300)

```

```{r parse-click-log, echo-FALSE}

e4.click_log <- data.frame(matrix(ncol=11,nrow=0, 
                               dimnames=list(NULL, 
                                             c("subj_id",
                                              "test_part", 
                                              "word",
                                              "category",
                                              "num_clicks",
                                              "num_misses",
                                              "letter",
                                              "hit",
                                              "t",
                                              "click_number",
                                              "genuine_first"))))

for (row in 1:nrow(e4.click_df)) {

    subject_click_log <- data.frame(fromJSON(e4.click_df[row, ]$click_log)) %>%
    mutate(
      letter = lead(letter,1),
      t = lead(t,1),
      hit = lead(hit,1),
      click_number = 1:n(),
      subj_id = e4.click_df[row, ]$subj_id,
      test_part = e4.click_df[row, ]$test_part,
      word = e4.click_df[row, ]$word,
      category = e4.click_df[row, ]$category,
      num_clicks = e4.click_df[row, ]$num_clicks,
      num_misses = e4.click_df[row, ]$num_misses,
      genuine_first = e4.click_df[row, ]$genuine_first
    )%>%
      filter(click_number<=num_clicks)

    e4.click_log <- rbind(e4.click_log, subject_click_log);
}

e4.click_log <- e4.click_log %>%
  relocate(subj_id, .before = letter) %>%
  relocate(genuine_first, .before = letter) %>% 
  relocate(test_part, .before=letter) %>% 
  relocate(word, .before=letter) %>% 
  relocate(category, .before=letter) %>% 
  relocate(click_number, .before=letter) %>%
  group_by(subj_id,test_part,word) %>%
  mutate(RT=t-lag(t,default=0)) %>%
  group_by(subj_id,word) %>%
  rowwise()
# %>%
#   filter(!(subj_id %in% E1.lucky_pretenders))

```

```{r total_time, echo=FALSE}

e4.total_time_df <- e4.click_log %>%
 filter(click_number==num_clicks)%>%
  mutate(t=t/1000)

p <- e4.total_time_df %>%
  ggplot(aes(x=word,y=t)) +
  geom_boxplot() +
  scale_y_continuous(limits=c(0,150))

ggsave('figures/hangman_nonpretend_pilot_time_per_word.png',p)

e4.median_time <- e4.total_time_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(t=median(t)) %>%
  spread(test_part,t) %>%
  mutate(diff=pretend-nonpretend);

# anova
e4.time_anva <- aov(t ~ genuine_first * condition,
                         data = e4.median_time %>%
                           gather('condition','t',nonpretend,pretend) %>%
                           mutate(condition=as.factor(condition)))

p<- e4.total_time_df %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=t,fill=test_part))+
    geom_histogram( alpha=0.5, position = 'identity',bins=100) +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    scale_x_continuous(limits = c(0,200))+
    labs(x='median game duration', y='number of players')+
    theme(legend.position=c(0.90,0.85)) +
    facet_wrap(~genuine_first, nrow=2) +
    theme_classic();

ggsave('figures/Hangman2/total_time.pdf',p)
ggsave('figures/Hangman2/total_time.png',width=5,height=3,dpi=300)

e4.time_per_word <- e4.total_time_df %>%
  group_by(test_part,word) %>%
  summarise(t=median(t)) %>%
  spread(test_part,t)

p <- e4.time_per_word %>%
  ggplot(aes(x=pretend,y=nonpretend, label=word)) +
  geom_abline(intercept=0,slope=1) +
  geom_point()+
  geom_text(position=position_jitter(), size=3.5)+
  coord_fixed(ratio=1)+
  scale_x_continuous(limits=c(15,60))+
  scale_y_continuous(limits=c(15,60)) +
  theme(legend.position = 'none')+
  theme_classic()

ggsave('figures/Hangman2/time_per_word.png',width=3,height=3,dpi=300)

```

## Hypothesis 1: Game duration

The median game duration in non-pretend games was `r round(e4.median_time$nonpretend%>%median())` seconds in non-pretend games and `r round(e4.median_time$nonpretend%>%median())` seconds in pretend games. Pretend games were significantly shorter among subjects that pretended in the first block (`r e4.median_time%>%filter(!genuine_first)%>%pull(diff)%>%t.test()`) and in the second block (`r e4.median_time%>%filter(genuine_first)%>%pull(diff)%>%t.test()`).

```{r time_per_click, echo=FALSE}

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median(.,na.rm=T))
  };
  return(sd(medians))
}

e4.time_per_click_number <- e4.click_log %>%
  group_by(subj_id,test_part,genuine_first,click_number)%>%
  summarise(RT=median(RT)) %>%
  group_by(genuine_first,test_part,click_number) %>%
  summarise(median_RT=median(RT),
            sem_RT = bootstrap_error(RT,1000))

p <- e4.time_per_click_number %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  filter(click_number>=0 & click_number<=10) %>%
  ggplot(aes(x=click_number,y=median_RT, color=test_part, fill=test_part)) +
  geom_line() +
  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT), alpha=0.5)+
  scale_y_continuous(limits=c(0,7500))+
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  facet_wrap(~genuine_first,nrow=2) +
  theme_classic()

ggsave('figures/Hangman2/time_per_click_number.png',width=5,height=3,dpi=300)

e4.first_click_time_per_word <- e4.click_log %>%
  filter(click_number==1) %>%
  group_by(test_part,word)%>%
  summarise(t=median(t))%>%
  spread(test_part,t)

# p <- e4.time_per_first_click %>%
#   ggplot(aes(x=word,y=RT)) +
#   geom_boxplot() +
#   scale_y_continuous(limits=c(0,15000))
# 
# ggsave('figures/hangman_nonpretend_pilot_time_per_first_click.png',p)
```

## Hypothesis 2: First click latency

## Hypothesis 3: number of misses

Among our players, the median number of misses was `r e4.median_misses%>%pull(nonpretend)%>%median()` in non-pretend games and `r e4.median_misses%>%pull(pretend)%>%median()` in pretend games. This effect was significant among players that pretended in the first part (`r t.test(e4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-e4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(e4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-e4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).

Across the 10 words, the number of misses in pretend games closely matched the number of misses in non-pretend games (`r apa_print(cor.test(e4.misses_per_word$pretend,e4.misses_per_word$nonpretend,method='spearman'))$full_result`).

## Reaction times

```{r RT_by_accuracy}

e4.click_log <- e4.click_log %>%
  group_by(subj_id,test_part,word) %>%
 mutate(hit_bin = hit !=0,
        lag3 = lag(hit,3)!=0,
        lag2 = lag(hit,2)!=0,
        lag1 = lag(hit,1)!=0,
        lead1 = lead(hit,1)!=0,
        lead2 = lead(hit,2)!=0,
        lead3 = lead(hit,3)!=0)

# E1.RT_by_lag3 <- E1.click_log %>%
#   filter(click_number>3) %>%
#   group_by(subj_id,test_part,lag3,genuine_first)%>%
#   summarise(RT=median(RT))  %>%
#   spread(lag3,RT,sep='_') %>%
#   mutate(lag3_diff = lag3_TRUE-lag3_FALSE);
# 
# E1.RT_by_lag2 <- E1.click_log %>%
#   filter(click_number>2) %>%
#   group_by(subj_id,test_part,lag2,genuine_first)%>%
#   summarise(RT=median(RT))  %>%
#   spread(lag2,RT,sep='_') %>%
#   mutate(lag2_diff = lag2_TRUE-lag2_FALSE);

e4.RT_by_lag1 <- e4.click_log %>%
  filter(click_number>1 & click_number<num_clicks-1) %>%
  group_by(genuine_first,subj_id,test_part,lag1)%>%
  summarise(RT=median(RT, na.rm=T))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

e4.RT_by_hit <- e4.click_log %>%
  filter(click_number>1 & click_number<num_clicks-1) %>%
  group_by(genuine_first,subj_id,test_part,hit_bin)%>%
  summarise(RT=median(RT, na.rm=T))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)

e4.RT_by_lead1 <- e4.click_log %>%
  filter(click_number>1 & click_number<num_clicks-1) %>%
  group_by(genuine_first,subj_id,test_part,lead1)%>%
  summarise(RT=median(RT, na.rm=T))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

# E1.RT_by_lead2 <- E1.click_log %>%
#   filter(click_number<num_clicks-1) %>%
#   group_by(subj_id,test_part,lead2,genuine_first)%>%
#   summarise(RT=median(RT))  %>%
#   spread(lead2,RT,sep='_') %>%
#   mutate(lead2_diff = lead2_TRUE-lead2_FALSE)
# 
# E1.RT_by_lead3 <- E1.click_log %>%
#   filter(click_number<num_clicks-2) %>%
#   group_by(subj_id,test_part,lead3,genuine_first)%>%
#   summarise(RT=median(RT))  %>%
#   spread(lead3,RT,sep='_') %>%
#   mutate(lead3_diff = lead3_TRUE-lead3_FALSE)
# 
e4.RT_by_hit_lags <- e4.RT_by_lag1 %>%
  merge(e4.RT_by_hit) %>%
  merge(e4.RT_by_lead1)

e4.RT_by_hit_lags_long <- e4.RT_by_hit_lags %>%
  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
  mutate(position = ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,-1))) %>%
  dplyr::select(genuine_first,subj_id,test_part,position,difference);

p<- e4.RT_by_hit_lags_long %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  group_by(genuine_first,test_part,position) %>%
  summarise(median_difference=median(difference, na.rm=T),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  # mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_vline(xintercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t')+
  theme_classic()

ggsave('figures/Hangman2/RT_by_lag_acc.pdf',p,width=7)
ggsave('figures/Hangman2/RT_by_lag_acc.png',p,width=5, height=3, dpi=300)

e4.RT_by_hit_lags_by_condition <- e4.RT_by_hit_lags_long %>%
  spread(test_part,difference) %>%
  mutate(interaction = pretend-nonpretend)

p<- e4.RT_by_hit_lags_long %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  group_by(genuine_first,test_part,position) %>%
  summarise(mean_difference=mean(difference, na.rm=T),
            se_difference=se(difference)) %>%
  # mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  ggplot(aes(x=position,color=test_part,fill=test_part,y=mean_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_vline(xintercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=mean_difference-se_difference,ymax=mean_difference+se_difference),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t')+
  theme_classic()

ggsave('figures/Hangman2/RT_by_lag_acc_with_mean_instead_of_median_at_the_group_level.png',p,width=5, height=3, dpi=300)

# p<- E1.RT_by_hit %>% 
#   group_by(test_part,hit_bin,genuine_first) %>%
#   summarise(median_RT=median(RT)/1000,
#             sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
#   mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
#   ggplot(aes(x=test_part,fill=hit_bin,y=median_RT)) +
#   geom_bar(stat="identity",position="dodge", color='black')+
#   geom_errorbar(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),linetype="solid", color="black", position="dodge")+
#   facet_wrap(~genuine_first,nrow=2)+
#   labs(fill='hit',x='test part', y='median RT (seconds)')
# 
# ggsave('figures/RT_by_acc.pdf',p)
# 
# 
# E1.RT_by_prev_trial <- E1.click_log %>%
#   filter(click_number>1) %>%
#   group_by(subj_id,test_part,prev_hit, genuine_first)%>%
#   summarise(RT=median(RT))
# 
# E1.RT_by_prev_trial_long<- E1.RT_by_prev_trial %>%
#   spread(prev_hit,RT,sep='_') %>%
#   mutate(hit_diff = prev_hit_FALSE-prev_hit_TRUE)
# 
# p<- E1.RT_by_prev_trial %>% 
#   group_by(test_part,prev_hit,genuine_first) %>%
#   summarise(median_RT=median(RT)/1000,
#             sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
#   mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
#   ggplot(aes(x=test_part,fill=prev_hit,y=median_RT)) +
#   geom_bar(stat="identity",position="dodge", color='black')+
#   geom_errorbar(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),linetype="solid", color="black", position="dodge")+
#   facet_wrap(~genuine_first,nrow=2)+
#   labs(fill='hit on previous click',x='test part', y='median RT (seconds)')
# 
# ggsave('figures/RT_by_prev_acc.pdf',p)
# 
# E1.RT_by_next_trial <- E1.click_log %>%
#   filter(click_number<num_clicks) %>%
#   group_by(subj_id,test_part,next_hit, genuine_first)%>%
#   summarise(RT=median(RT))
# 
# E1.RT_by_next_trial_long<- E1.RT_by_next_trial %>%
#   spread(next_hit,RT,sep='_') %>%
#   mutate(hit_diff = next_hit_FALSE-next_hit_TRUE)
# 
# p<- E1.RT_by_next_trial %>% 
#   group_by(test_part,next_hit,genuine_first) %>%
#   summarise(median_RT=median(RT)/1000,
#             sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
#   mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
#   ggplot(aes(x=test_part,fill=next_hit,y=median_RT)) +
#   geom_bar(stat="identity",position="dodge", color='black')+
#   geom_errorbar(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),linetype="solid", color="black", position="dodge")+
#   facet_wrap(~genuine_first,nrow=2)+
#   labs(fill='hit on next click',x='test part', y='median RT (seconds)')

```

## Hypotheses 4-6: effect of hits/misses on RT

Hits were faster than misses by `r round(abs(e4.RT_by_hit_lags_by_condition%>%filter(position==0)%>%pull(nonpretend)%>%median(na.rm=T)))` ms in non-pretend games (non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==0 & genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`; pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==0 & !genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`) and by `r round(abs(e4.RT_by_hit_lags_by_condition%>%filter(position==0)%>%pull(pretend)%>%median(na.rm=T)))` ms in pretend games (non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==0 & genuine_first)%>%pull(pretend)%>%t.test())$full_result`; pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==0 & !genuine_first)%>%pull(pretend)%>%t.test())$full_result`). We find no significant interaction between these effects (non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==0 & genuine_first)%>%pull(interaction)%>%t.test())$full_result`; pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==0 & !genuine_first)%>%pull(interaction)%>%t.test())$full_result`).

Letter selections following a hit were slower than following a miss by `r round(abs(e4.RT_by_hit_lags_by_condition%>%filter(position==1)%>%pull(nonpretend)%>%median(na.rm=T)))` ms in non-pretend games (non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==1 & genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`; pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==1 & !genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`) and by `r round(abs(e4.RT_by_hit_lags_by_condition%>%filter(position==1)%>%pull(pretend)%>%median(na.rm=T)))` ms in pretend games (non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==1 & genuine_first)%>%pull(pretend)%>%t.test())$full_result`; pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==1 & !genuine_first)%>%pull(pretend)%>%t.test())$full_result`). We find no significant interaction between these effects (non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==1 & genuine_first)%>%pull(interaction)%>%t.test())$full_result`; pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==1 & !genuine_first)%>%pull(interaction)%>%t.test())$full_result`).

Letter selections preceding a hit were faster than those preceding a miss in non-pretend games only in the group that pretended first (`r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`) but not in the group that played normally first (`r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`). This effect was not observed in pretend games (pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)%>%t.test())$full_result`; non-pretend first: `r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(pretend)%>%t.test())$full_result`). We find a significant interaction between these effects in the group that pretended first (`r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(interaction)%>%t.test())$full_result`) but not in the other group (`r apa_print(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(interaction)%>%t.test())$full_result`). The difference was not significant when analyzing first blocks only (`r apa_print(t.test(e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend),e4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)))$statistic`).

```{r optimality_analysis, echo=FALSE}

number_words_df <- read.csv('../analysis/word_lists/numbers.csv')
states_df <- read.csv('../analysis/word_lists/USstates.csv')
fruit_df <- read.csv('../analysis/word_lists/fruit.csv')
body_df <-  read.csv('../analysis/word_lists/bodypart.csv')
celebrity_df <- read.csv('../analysis/word_lists/celebrities.csv')
letter_df <- read.csv('../analysis/word_lists/letters.csv')

is_consistent <- function(word_state, asked,  word) {

  word_state = strsplit(toupper(word_state),'')[[1]];
  word = strsplit(toupper(word),'')[[1]];
  asked = strsplit(toupper(asked),'')[[1]];
  not_in_word = asked[!(asked %in% word_state)]

  # check if number of letters is consistent
  if (length(word_state) != length(word)) {
    return(FALSE)
  }

  # check if the letters in the masked word are consistent
  for (i in seq_along(word)) {
    if (word_state[i] != word[i]) {
      if (word_state[i] != '_' | word[i] == ' ' | word[i] == '.') {
      return(FALSE)
      }
    }
  }

  # check if any of the misses appear in the candidate word
  if (length(Reduce(intersect,list(word,not_in_word)))>0) {
    return(FALSE)
  }

  #if the function hasn't returned FALSE, return TRUE
  return(TRUE)

}

get_word_prior <- function(category) {

  return(
    switch(category,
         "a number" = number_words_df,
         "a US state" = states_df,
         "a body part" = body_df,
         "a famous person" = celebrity_df,
         "a fruit" = fruit_df)
  )
}

get_word_probabilities <- function(word_state, asked, category) {

  word_probabilities <- get_word_prior(category) %>%
    rowwise() %>%
    mutate(likelihood = ifelse(is_consistent(word_state, asked,word),1, 0),
           pword = prior * likelihood) %>%
    ungroup() %>%
    mutate(posterior = pword/sum(pword))%>%
    filter(posterior>0) %>%
    arrange(desc(posterior))

  return(word_probabilities)

};

get_posterior <- function(word_state, asked, category) {

  word_probabilities <- get_word_probabilities(word_state, asked, category);

  letter_probabilities = c();
  for (letter in toupper(letters)) {
    if (grepl(letter,asked)) {
      letter_probabilities = c(letter_probabilities,
                               NA)
    } else {
      letter_probabilities = c(letter_probabilities,
                             word_probabilities %>%
                               filter(grepl(letter,toupper(word))) %>%
                               pull(posterior) %>%
                               sum()
    )
    }
  }
  posterior = letter_probabilities/sum(letter_probabilities, na.rm=T)

  return(paste(posterior,collapse=','))

};

get_letter_prob_posterior <- function(asked) {

  letter_probabilities <- letter_df %>%
    rowwise() %>%
    mutate(prob = ifelse(
      grepl(toupper(letter),toupper(asked)),
      NA,
      frequency)
      ) %>%
    ungroup()%>%
    mutate(
      posterior = prob/sum(prob,na.rm=T)
    )

  return(paste(letter_probabilities$posterior,collapse=','))

};

get_p_click <- function(posterior, letter) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(posterior[which(toupper(letters)==letter)])
}

get_p_click_rank <- function(posterior, letter) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  rank_posterior = rank(-posterior)
  return(rank_posterior[which(toupper(letters)==letter)])
}

get_posterior_entropy <- function(posterior) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  posterior[which(posterior==0)]=NA; #to avoid infinity*0
  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
  return(entropy);
}

e4.click_log_with_boards <- e4.click_log %>%
  dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>% 
  rowwise()%>%
  mutate(posterior = get_posterior(word_state,asked,category),
         p_click = get_p_click(posterior,letter),
         entropy = get_posterior_entropy(posterior),
         p_click_rank = get_p_click_rank(posterior,letter));

get_random_letter <- function(asked) {

  asked = strsplit(tolower(asked),'')[[1]];
  not_asked = setdiff(letters,asked)
  return(toupper(not_asked)%>%sample(1))

}

get_most_frequent_letter <- function(asked) {

   lp <- letter_df %>%
    rowwise() %>%
    mutate(prob = ifelse(
      grepl(toupper(letter),toupper(asked)),
      NA,
      frequency)
      ) %>%
    ungroup()

     return(lp[which.max(lp$prob),]$letter%>%toupper())

}

e4.click_log_with_boards_random <- e4.click_log_with_boards %>%
  mutate(letter = get_random_letter(asked),
         p_click = get_p_click(posterior,letter),
         p_click_rank = get_p_click_rank(posterior,letter),
         hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
         test_part='random');

e4.click_log_with_boards_most_frequent <- e4.click_log_with_boards %>%
  mutate(letter = get_most_frequent_letter(asked),
         p_click = get_p_click(posterior,letter),
         p_click_rank = get_p_click_rank(posterior,letter),
         hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
         test_part='most_frequent');

e4.letter_prob_click_log_with_boards <- e4.click_log %>%
  dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>% 
  rowwise()%>%
  mutate(posterior = get_letter_prob_posterior(asked),
         p_click = get_p_click(posterior,letter),
         entropy = get_posterior_entropy(posterior),
         p_click_rank = get_p_click_rank(posterior,letter));

e4.letter_prob_click_log_with_boards_random <- e4.letter_prob_click_log_with_boards %>%
  mutate(letter = get_random_letter(asked),
         p_click = get_p_click(posterior,letter),
         p_click_rank = get_p_click_rank(posterior,letter),
         hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
         test_part='random');
```

```{r word_probabilities}

words <-e4.click_log %>%
  group_by(word) %>% summarise(word_state=word_state[1], category=category[1],asked=asked[1]) %>%
  filter(word != 'vladimir putin') %>%
  mutate(word=as.character(word), category=as.character(category))

words[nrow(words)+1,] = list('taylor swift','______ _____','a famous person','')
words[nrow(words)+1,] = list('papaya','_A_A_A','a fruit','AEIOMT')
words[nrow(words)+1,] = list('banana','_A_A_A','a fruit','AEIOMT')
words[nrow(words)+1,] = list('hand','HA__','a body part','AEOMTHP')
words[nrow(words)+1,] = list('hair','HA__','a body part','AEOMTHP')

save_word_probs <- function(word,word_state, asked, category) {
  probs <- get_word_probabilities(word_state, asked, category);
  probs %>%
      write.csv(paste('../analysis/word_lists/per_game/',word,'.csv',sep=''))
  return('')
}

words %>%
  rowwise() %>%
  mutate(a=save_word_probs(word,word_state,asked,category))
```

```{r p_click}

e4.mean_P_click_rank <- e4.click_log_with_boards %>%
  rbind(e4.click_log_with_boards_random)%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

e4.mean_P_click_rank_wide <- e4.mean_P_click_rank %>%
  spread(test_part,p_click_rank) %>%
  mutate(p_vs_r = pretend-random,
         np_vs_r = nonpretend-random,
         p_vs_np = pretend-nonpretend)

e4.mean_letter_prob_P_click_rank <- e4.letter_prob_click_log_with_boards %>%
  rbind(e4.click_log_with_boards_random)%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

e4.mean_letter_prob_P_click_rank_wide <- e4.mean_letter_prob_P_click_rank %>%
  spread(test_part,p_click_rank) %>%
  mutate(p_vs_r = pretend-random,
         np_vs_r = nonpretend-random,
         p_vs_np = pretend-nonpretend)

p<- e4.mean_P_click_rank %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725")) +
    labs(x='p(letter in word) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2)+
    theme_classic();

ggsave('figures/Hangman2/pclick_rank.pdf',p)
ggsave('figures/Hangman2/pclick_rank.png',p,width=5,height=3,dpi=300);

p<- e4.mean_letter_prob_P_click_rank %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725")) +
    labs(x='p(letter) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2)+
    theme_classic();

ggsave('figures/Hangman2/letter_prob_pclick_rank.pdf',p)
ggsave('figures/Hangman2/letter_prob_pclick_rank.png',p,width=5,height=3,dpi=300)

e4.mean_P_click_rank_misses <- e4.click_log_with_boards %>%
  filter(!hit_bin) %>%
  rbind(e4.click_log_with_boards_random %>%filter(!hit_bin))%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

e4.mean_P_click_rank_misses_wide <- e4.mean_P_click_rank_misses %>%
  spread(test_part,p_click_rank) %>%
  mutate(p_vs_r = pretend-random,
         np_vs_r = nonpretend-random,
         p_vs_np = pretend-nonpretend)

e4.mean_letter_prob_P_click_rank_misses <- e4.letter_prob_click_log_with_boards %>%
  filter(!hit_bin) %>%
  rbind(e4.click_log_with_boards_random %>%filter(!hit_bin))%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

e4.mean_letter_prob_P_click_rank_misses_wide <- e4.mean_letter_prob_P_click_rank_misses %>%
  spread(test_part,p_click_rank) %>%
  mutate(p_vs_r = pretend-random,
         np_vs_r = nonpretend-random,
         p_vs_np = pretend-nonpretend)

p<- e4.mean_P_click_rank_misses %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(letter in word) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2)+
    theme_classic();

ggsave('figures/Hangman2/pclick_rank_misses.pdf',p)
ggsave('figures/Hangman2/pclick_rank_misses.png',p,width=5,height=3,dpi=300)

p<- e4.mean_letter_prob_P_click_rank_misses %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(letter) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2)+
    theme_classic();

ggsave('figures/Hangman2/letter_prob_pclick_rank_misses.pdf',p)
ggsave('figures/Hangman2/letter_prob_pclick_rank_misses.png',p,width=5,height=3,dpi=300)

e4.num_irrational_clicks <- e4.click_log_with_boards %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(nic=sum(p_click==0)/5) %>%
  spread(test_part,nic)

e4.irrational_clicks_misses_ratio <- e4.click_log_with_boards %>%
  filter(!hit_bin)%>%
  rbind(e4.click_log_with_boards_random %>%
           filter(!hit_bin))%>%
    rbind(e4.click_log_with_boards_most_frequent %>%
           filter(!hit_bin))%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(nic=sum(p_click==0)/n()) %>%
  spread(test_part,nic) %>%
  mutate(diff=pretend-nonpretend)

e4.rank_misses_by_category <- e4.click_log_with_boards %>%
  filter(!hit_bin) %>%
  rbind(e4.click_log_with_boards_random %>%filter(!hit_bin))%>%
  group_by(subj_id,test_part, category, word) %>%
  summarise(p_click_rank=mean(p_click_rank)) %>%
  group_by(test_part,category,word)%>%
  summarise(p_rank = mean(p_click_rank,na.rm=T),
            se = se(p_click_rank,na.rm=T))

p <- e4.rank_misses_by_category %>% ggplot(
  aes(x=as.factor(word),y=p_rank,fill=test_part)
) + 
  geom_bar(position=position_dodge(),
           stat='identity', color='black')+
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
  geom_errorbar(aes(ymin=p_rank-se, ymax=p_rank+se), width=.2,position=position_dodge(.9)) +
  labs(x='word',
       y='posterior rank')

ggsave('figures/Hangman2/prank_by_word.png',p,width=12,height=3,dpi=300)

```

## Hypothesis 7: optimality

Letter selections in pretend games had a higher posterior probability to be included in the word than in non-pretend games (pretend first: `r e4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`). The same was true for letter selections in pretend games (pretend first: `r e4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`). Non-pretend letter selections had a higher posterior probability among players who pretended first (`r e4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`) but not in the group that played normally first (`r e4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`). Comparing data from the first condition only in a between-subject manner revealed a significant difference (`r t.test(e4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(nonpretend),e4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).

When focusing on misses, non-pretend letter selections were more optimal than random (pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`), as well as pretend games (pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`). In both cases, non-pretend games were more optimal than pretend games (pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`)

## Hypothesis 7b: Letter frequency

Non-pretenders selected high-frequency letters more than expected by chance (pretend first: `r e4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`). The same was true for letter selections in pretend games (pretend first: `r e4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`). Letters selected by pretenders had a higher overall frequency compared to letters selected by non-pretenders (pretend first: `r e4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`). but not in the group that played normally first (`r e4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`). Comparing data from the first condition only in a between-subject manner revealed a significant difference (`r t.test(e4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(nonpretend),e4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).

This was the case also when focusing on misses only ( pretend first: `r e4.mean_letter_prob_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r e4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`)

```{r}

e4.RT_by_entropy <- e4.click_log_with_boards %>%
  filter(click_number>1) %>%
  mutate(entropy_bin=round(entropy*10)/10) %>%
  group_by(test_part,subj_id,genuine_first,entropy_bin) %>% 
  summarise(RT=median(RT)) %>%
  group_by(test_part,genuine_first,entropy_bin) %>%
  summarise(median_RT=median(RT),
            sem_RT=bootstrap_error(RT,N_perm))

p <- e4.RT_by_entropy %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
  theme_classic()

ggsave('figures/Hangman2/RT_by_entropy.png',p,width=5,height=3,dpi=300)

e4.RT_by_entropy_trimmed <- e4.click_log_with_boards %>%
  # filter(click_number>1) %>%
  filter(RT<5000 & RT>100) %>%
  mutate(entropy_bin=round(entropy*5)/5) %>%
  group_by(test_part,subj_id,genuine_first,entropy_bin) %>% 
  summarise(RT=median(RT)) %>%
  group_by(test_part,genuine_first,entropy_bin) %>%
  summarise(median_RT=median(RT),
            sem_RT=bootstrap_error(RT,N_perm))

p <- e4.RT_by_entropy_trimmed %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
  theme_classic()

ggsave('figures/hangman2/RT_by_entropy_trimmed.png',p,width=5,height=3,dpi=300)

e4.RT_by_entropy_coefs <- e4.click_log_with_boards %>%
  # filter(click_number>1) %>%
  group_by(genuine_first,subj_id,test_part,word) %>%
  mutate(entropy_centered = entropy-mean(entropy),
         entropy_squared = entropy_centered^2) %>%
  select(genuine_first,subj_id,test_part,word,click_number,entropy_centered,entropy_squared,RT) %>%
  group_by(genuine_first,subj_id,test_part,word) %>%
  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
  mutate(tidys=list(broom::tidy(model))) %>%
  unnest(tidys) %>%
  dplyr::select(!c(std.error,statistic,p.value)) %>%
  group_by(genuine_first,subj_id,test_part,term) %>%
  summarise(estimate = mean(estimate))

e4.RT_by_entropy_coefs_wide <- e4.RT_by_entropy_coefs %>% 
  spread(test_part,estimate) %>%
  mutate(diff=pretend-nonpretend)

e4.RT_by_entropy_coefs_summary <- e4.RT_by_entropy_coefs %>%
  group_by(test_part,term, genuine_first) %>%
  summarise(mean_estimate = mean(estimate),
            se_estimate = se(estimate))

p <- e4.RT_by_entropy_coefs_summary %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
  filter(term!='(Intercept)') %>%
  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='term', y='coefficient')+
  theme_classic() +
  facet_grid(rows=vars(genuine_first))

ggsave('figures/Hangman2/RT_by_entropy_coefficientst.png',p,width=5,height=5,dpi=300)

e4.corrected_RT_by_entropy_coefs <- e4.click_log_with_boards %>%
  group_by(subj_id,test_part,click_number) %>%
  mutate(corrected_RT = RT-mean(RT)) %>%
  group_by(genuine_first,subj_id,test_part) %>%
  mutate(entropy_centered = entropy-mean(entropy),
         entropy_squared = entropy_centered^2) %>%
  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
  group_by(genuine_first,subj_id,test_part) %>%
  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
  mutate(tidys=list(broom::tidy(model))) %>%
  unnest(tidys) %>%
  dplyr::select(!c(std.error,statistic,p.value))

e4.corrected_RT_by_entropy_coefs_summary <- e4.corrected_RT_by_entropy_coefs %>%
  group_by(test_part,term,genuine_first) %>%
  summarise(mean_estimate = mean(estimate),
            se_estimate = se(estimate))

p <- e4.corrected_RT_by_entropy_coefs_summary %>%
      mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
  filter(term!='(Intercept)') %>%
  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='term', y='coefficient')+
  theme_classic()+
  facet_grid(rows=vars(genuine_first))

ggsave('figures/Hangman2/corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=5,dpi=300)

```

## Hypothesis 8: Entropy/RT relationship

In non-pretend games, high uncertainty was associated with longer decision times (non-pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`). The same was true for pretend games (non-pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`). The effects were significantly weaker in pretend games (non-pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).

In non-pretend games, the relationship between uncertainty and RT was negatively quadratic, such that the longest responses were the one associated with intermediate uncertainty levels (non-pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`). The same was true for pretend games (non-pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`). The effects were significantly weaker in pretend games (non-pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r e4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).

```{r}

e4.judge_df <- e4.df %>%
  filter(test_part=='judge' & trial_type=='html-keyboard-response')%>%
  dplyr::select(subj_id, 
         word, 
         player,
         correct_response,
         key_press,
         correct)%>%
  rowwise()%>%
  mutate(decision = ifelse(key_press==80,1,0),
         pretend = ifelse(correct_response=='P',1,0)
)

e4.judge_df$correct%>%mean();

e4.judge_summary_df <- e4.judge_df %>%
  group_by(subj_id)%>%
  summarise(accuracy=mean(correct),
            bias=mean(decision))

```

```{r condition-classifier, echo=FALSE, eval=FALSE}

set.seed(42);

gamewise_data <- e4.click_log_with_boards %>% 
  filter(genuine_first != 'True' | test_part=='pretend') %>% 
  mutate(word=factor(word)) %>%
  group_by(subj_id,test_part,word) %>%
  filter(sum(RT)<=90000) %>%
  summarise(median_RT=median(RT),
            mean_p_click_rank = mean(p_click_rank),
            first_click_latency=RT[click_number==1],
            num_irrational_clicks = sum(p_click==0)) %>%
  ungroup()%>%
  select(test_part,mean_p_click_rank,median_RT,num_irrational_clicks,word) 
# 
# model.matrix(~ 0 + word, gamewise_data) %>%
#     as.data.frame() %>%
#     bind_cols(gamewise_data) %>%
#     select(x, day, everything());

# gamewise_data_nonpretend = gamewise_data%>%filter(test_part=='nonpretend') %>%mutate(nrow=1:n())
# gamewise_data_pretend = gamewise_data%>%filter(test_part=='pretend')
# rows <- sample(nrow(gamewise_data_nonpretend))
# gamewise_data_pretend <- gamewise_data_pretend[rows,]%>%mutate(nrow=1:n())
# 
# pair_gamewise_data <- full_join(gamewise_data_nonpretend,
#                                 gamewise_data_pretend,
#                                 by='nrow',
#                                 suffix=c('.nonpretend','.pretend')) %>%
#   mutate(which_side = sample(c(-1,1),nrow(gamewise_data_pretend),replace=TRUE),
#          mean_p_click = which_side*(mean_p_click_rank.nonpretend-mean_p_click_rank.pretend),
#          median_RT = which_side*(median_RT.nonpretend-median_RT.pretend),
#          num_irrational_clicks = which_side*(num_irrational_clicks.nonpretend-num_irrational_clicks.pretend),
#          which_side=factor(which_side)) %>%
#   select(which_side,mean_p_click,median_RT,num_irrational_clicks)

train_control <- trainControl(method = "CV", number=10, sampling='down')

# train the model on training set
model <- train(test_part ~ .,
               data = gamewise_data,
               trControl = train_control,
               method = "svmLinear",
               preProcess = c("center","scale"))

# print cv scores
# model
```

```{r}

unidentifiable_ss <- e4.judge_df %>%
  group_by(subj_id)%>%
  summarise(n=n())%>%
  filter(n>5)%>%
  pull(subj_id)%>%
  unique();

e4.judge_performance <- e4.judge_df %>%
  filter(!(subj_id %in% unidentifiable_ss)) %>%
  group_by(subj_id)%>%
  summarise(judge_rate=sum(correct)/5);

e4.pretend_performance <- e4.judge_df %>%
  filter(!(player %in% unidentifiable_ss & pretend==1)) %>%
  group_by(player)%>%
  summarise(ntrials=n(),
            successful=ntrials-sum(correct),
            pretend_rate = successful/ntrials) %>%
  mutate(subj_id=player)

e4.optimality_cost <- e4.click_log_with_boards %>%
  group_by(subj_id,test_part) %>%
  summarise(p_click_rank=mean(p_click_rank)) %>%
  spread(test_part,p_click_rank) %>%
  mutate(optimality_cost=nonpretend-pretend)

e4.performance_df <- merge(e4.judge_performance,e4.pretend_performance)%>%
  merge(e4.optimality_cost);

p <- e4.performance_df %>%
  ggplot(aes(x=judge_rate,y=pretend_rate)) +
  geom_jitter() + 
  labs(x='judge accuracy',y='pretense success rate') +
  theme_classic()

ggsave('figures/Hangman2/judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)

p <- e4.performance_df %>%
  ggplot(aes(x=judge_rate,y=optimality_cost)) +
  geom_jitter() + 
  labs(x='judge accuracy',y='optimality cost') +
  theme_classic()

ggsave('figures/Hangman2/optimality_judge_accuracy.png',p,width=5,height=3,dpi=300)

```

## Hypothesis 9 (judge accuracy)

Subjects were not better than chance in detecting pretense games (`r e4.judge_summary_df$accuracy%>%t.test(mu=0.5)%>%apa_print()%>%'$'(full_result)`).

## Hypothesis 9B (judge bias)

Subjects were not slightly biased to classify games as pretend games (`r e4.judge_summary_df$bias%>%t.test(mu=0.5)%>%apa_print()%>%'$'(full_result)`).

## Hypothesis 10 (judge-pretend relation)

No significant correlation between judge accuracy and ability to trick others (`r cor.test(e4.performance_df$pretend_rate,e4.performance_df$judge_rate,method='spearman')%>%apa_print()%>%'$'(full_result)`), nor with the optimality cost to pretending (`r cor.test(e4.performance_df$optimality_cost,e4.performance_df$judge_rate,method='spearman')%>%apa_print()%>%'$'(full_result)`).

### Half games

```{r}

e4.half_game_click_df <- e4.df %>%
  dplyr::select(subj_id, 
         test_part, 
         word,
         num_clicks,
         click_log,
         genuine_first) %>%
  rowwise()%>%
  mutate(num_hits = strsplit(gsub(' ','',word),split='')[[1]]%>%unique()%>%length(),
         num_misses = num_clicks-num_hits,
         click_log = gsub("\'","\"", click_log),
         click_log = gsub("None","null", click_log),
         word = factor(word,levels=c(
           'papaya',
           'banana',
           'hair',
           'hand'
         )),
         category = ifelse(word=='papaya'|word=='banana','a fruit','a body part')) %>%
  filter(test_part=='pretend_half_game' | test_part=='nonpretend_half_game')

e4.half_game_click_log <- data.frame(matrix(ncol=11,nrow=0, 
                               dimnames=list(NULL, 
                                             c("subj_id",
                                              "test_part", 
                                              "word",
                                              "category",
                                              "num_clicks",
                                              "num_misses",
                                              "letter",
                                              "hit",
                                              "t",
                                              "click_number",
                                              "genuine_first"))))

for (row in 1:nrow(e4.half_game_click_df)) {

    subject_click_log <- data.frame(fromJSON(e4.half_game_click_df[row, ]$click_log)) %>%
    mutate(
      letter = lead(letter,1),
      t = lead(t,1),
      hit = lead(hit,1),
      click_number = 1:n(),
      subj_id = e4.half_game_click_df[row, ]$subj_id,
      test_part = e4.half_game_click_df[row, ]$test_part,
      word = e4.half_game_click_df[row, ]$word,
      category = e4.half_game_click_df[row, ]$category,
      num_clicks = e4.half_game_click_df[row, ]$num_clicks,
      num_misses = e4.half_game_click_df[row, ]$num_misses,
      genuine_first = e4.half_game_click_df[row, ]$genuine_first
    )%>%
      filter(click_number<=num_clicks)

    e4.half_game_click_log <- rbind(e4.half_game_click_log, subject_click_log);
}

e4.half_game_click_log <- e4.half_game_click_log %>%
  relocate(subj_id, .before = letter) %>%
  relocate(genuine_first, .before = letter) %>% 
  relocate(test_part, .before=letter) %>% 
  relocate(word, .before=letter) %>% 
  relocate(category, .before=letter) %>% 
  relocate(click_number, .before=letter) %>%
  group_by(subj_id,test_part,word) %>%
  mutate(RT=t-lag(t,default=0)) %>%
  group_by(subj_id,word) %>%
  rowwise()

# %>%
#   mutate(posterior = get_posterior(word_state,asked,category),
#          p_click = get_p_click(posterior,letter),
#          entropy = get_posterior_entropy(posterior),
#          p_click_rank = get_p_click_rank(posterior,letter))

e4.letter_choice_frequency <- e4.half_game_click_log %>%
  filter(click_number==1) %>%
  dplyr::select(subj_id,test_part,word_state,letter) %>%
  mutate(letter=factor(letter,levels=toupper(letters))) %>%
  group_by(letter,test_part,word) %>%
  summarise(count = n()) %>%
  group_by(test_part,word) %>%
  mutate(p=count/sum(count)) %>%
  complete(letter,fill=list(count=0,p=0)) %>%
  mutate(test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend'))

e4.first_click_half_games <- e4.half_game_click_log %>%
  filter(click_number==1) %>%
  dplyr::select(subj_id,genuine_first,word,letter,t,hit,test_part) %>%
  mutate(
    makes_sense = 
    (word %in% c('hair','hand') & letter %in% c('N','D','I','R')) | 
      (word %in% c('banana','papaya') & letter %in% c('P','Y','B','N')),
    bananahand = letter %in% c('B','N','D')
  )

p <- e4.first_click_half_games %>%
  group_by(test_part,genuine_first) %>%
  summarise(makes_sense=mean(makes_sense)) %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'),
         test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend')) %>%
  ggplot(aes(x=test_part,fill=test_part,y=makes_sense))+
  geom_bar(position=position_dodge(),
           stat='identity', color='black')+
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
  facet_grid(~genuine_first)+
  theme_classic()+
  labs(y='proportion of clicks that make sense')

ggsave('figures/Hangman2/halfgames_make_sense.png',p,width=6,height=3,dpi=300)

p <- e4.first_click_half_games %>%
  filter(makes_sense)%>%
  mutate(word_class = ifelse(word %in% c('banana','hand'),'BANANA or HAND','PAPAYA or HAIR'))%>%
  group_by(test_part,genuine_first,word_class) %>%
  summarise(bananahand = mean(bananahand)) %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'),
         test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend')) %>%
  ggplot(aes(x=test_part,fill=test_part,y=bananahand))+
  geom_bar(position=position_dodge(),
           stat='identity', color='black')+
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
  facet_grid(rows=vars(genuine_first),col=vars(word_class))+
  theme_classic()+
  labs(y='(consistent with BANANA/HAND)/(cons. with something)')

ggsave('figures/Hangman2/halfgames_attraction.png',p,width=6,height=4,dpi=300)

e4.optimal_letter_choice <- e4.half_game_click_log %>% 
  filter(click_number==1) %>% 
  group_by(word) %>% 
  summarize(posterior = get_posterior(word_state[1],asked[1],category[1]),
            word_state=word_state[1])

e4.optimal_by_letter <- data.frame(matrix(ncol=5,nrow=0, 
                               dimnames=list(NULL, 
                                             c("test_part", 
                                               "word",
                                              "word_state",
                                              "letter",
                                              "p"))))

for (row in 1:nrow(e4.optimal_letter_choice)) {

   p <- scan(text= e4.optimal_letter_choice[row,]$posterior, what = numeric(), sep="," , quiet = TRUE);
   letter = LETTERS;

   this.word <- tibble(letter,p) %>%
     mutate(
       word_state = e4.optimal_letter_choice[row,]$word_state,
       word = e4.optimal_letter_choice[row,]$word,
       test_part = 'optimal'
     )

    e4.optimal_by_letter <- rbind(e4.optimal_by_letter, this.word);

}

e4.letter_choice_frequency <-
  e4.letter_choice_frequency %>%
  rbind(e4.optimal_by_letter)

p <- e4.letter_choice_frequency %>% 
  mutate(test_part = factor(test_part, levels=c('non-pretend','pretend','optimal'))) %>%
  ggplot(aes(x=letter,y=p, fill=test_part)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(rows=vars(word)) +
  theme_minimal() +
  scale_fill_manual(values=c("#69b3a2", "#404080", "#49B240"))

ggsave('figures/hangmanpilot_halfgames.png',p,width=6,height=3,dpi=300)

p <- e4.letter_choice_frequency %>%
  filter((word=='papaya' | word=='banana') & test_part != 'optimal') %>%
  mutate(letter = factor(
    ifelse(letter %in% c('B','N','P','Y'), letter, 'other'),
    levels=rev(c('B','N','P','Y','other'))),
    test_part = factor(test_part, levels=c('non-pretend','pretend'))) %>%
  group_by(test_part,word,letter) %>%
  summarise(p=sum(p)) %>%
  ggplot(aes(x=letter,y=p, fill=test_part)) +
  geom_bar(stat='identity', position='dodge') +
  coord_flip()+
  facet_grid(cols=vars(word)) +
  theme_minimal() +
  scale_fill_manual(values=c("#69b3a2", "#404080", "#49B240"))

irrational_test_p_first <- binom.test(
  e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='False') %>%
    pull(makes_sense) %>%
    sum(), 
  e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='False') %>%
    nrow(),
  e4.first_click_half_games %>% 
    filter(test_part=='nonpretend_half_game' & genuine_first=='False') %>%
    pull(makes_sense) %>%
    mean())

irrational_test_np_first <- binom.test(
  e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='True') %>%
    pull(makes_sense) %>%
    sum(), 
  e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='True') %>%
    nrow(),
  e4.first_click_half_games %>% 
    filter(test_part=='nonpretend_half_game' & genuine_first=='True') %>%
    pull(makes_sense) %>%
    mean())

attraction_test_np_first <- prop.test(
  x=c(
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('banana','hand')) %>%
    pull(bananahand) %>%
    sum(), 
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('papaya','hair')) %>%
    pull(bananahand) %>%
    sum()), 
  n= c(
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('banana','hand')) %>%
    nrow(),
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('papaya','hair')) %>%
    nrow()))

attraction_test_p_first <- prop.test(
  x=c(
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('banana','hand')) %>%
    pull(bananahand) %>%
    sum(), 
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('papaya','hair')) %>%
    pull(bananahand) %>%
    sum()), 
  n= c(
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('banana','hand')) %>%
    nrow(),
    e4.first_click_half_games %>% 
    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('papaya','hair')) %>%
    nrow()))

```

## Hypothesis 12 (irrational letter selections)

In non-pretend games, subjects almost always chose letters that made sense (non-pretend first: `r e4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`; pretend first: `r e4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`). In pretend games, these figures were significantly lower (non-pretend first: `r e4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`, $p<.001$; pretend first: `r e4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`, $p<.001$).

## Hypothesis 13 (target word effect)

```{r compute-bonus-4}

nonpretend_points <- e4.df %>%
  group_by(PROLIFIC_PID) %>%
  filter(test_part=='nonpretend') %>%
  summarise(nonpretend = sum(points,na.rm=T),
            subj_id=subj_id[1]);

judge_points <- e4.judge_df %>%
  group_by(subj_id) %>%
  summarise(judge = 10*sum(correct));

pretend_points <- e4.judge_df %>%
  filter(pretend==1) %>%
  group_by(player)%>%
  summarise(pretend=sum(correct)*10) %>%
  mutate(subj_id=player) %>%
  dplyr::select(subj_id,pretend)

bonus.df <- nonpretend_points %>%
  merge(judge_points,by='subj_id',all=T) %>%
  merge(pretend_points, by="subj_id", all=T) %>%
  mutate(points = pretend+nonpretend+judge,
         bonus=floor(points/100));

export_2_ss <- e4.export2$participant_id %>%
  unique()

bonus.df %>%
  filter(PROLIFIC_PID %in% export_2_ss)%>%
  filter(bonus>0)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\prolific_export2.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

export_3_ss <- e4.export3$participant_id %>%
  unique()

bonus.df %>%
  filter(PROLIFIC_PID %in% export_3_ss)%>%
  filter(bonus>0)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\prolific_export3.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

export_4_ss <- e4.export4$participant_id %>%
  unique()

bonus.df %>%
  filter(PROLIFIC_PID %in% export_4_ss)%>%
  filter(bonus>0)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\prolific_export4.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

export_5_ss <- e4.export5$participant_id %>%
  unique()

bonus.df %>%
  filter(PROLIFIC_PID %in% export_5_ss)%>%
  filter(bonus>0)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\prolific_export5.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

export_6_ss <- e4.export6$participant_id %>%
  unique()

bonus.df %>%
  filter(PROLIFIC_PID %in% export_6_ss)%>%
  # filter(bonus>0)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\prolific_export6.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

# this ignores participatns whose games were not presented to later participants!
nonincluded_subs <-setdiff(nonpretend_points$subj_id%>%unique(),
                           pretend_points$subj_id%>%unique())
nonincluded.pretend_points <- data.frame(subj_id=nonincluded_subs,pretend=0)

nonincluded_bonus.df <- nonpretend_points %>%
  merge(judge_points,by='subj_id',all=T) %>%
  merge(nonincluded.pretend_points, by="subj_id", all=T) %>%
  mutate(points = pretend+nonpretend+judge) %>%
  filter(points>85);

nonincluded_bonus.df %>%
  filter(PROLIFIC_PID %in% export_2_ss)%>%
  mutate(bonus=1)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\nonincluded_prolific_export2.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

nonincluded_bonus.df %>%
  filter(PROLIFIC_PID %in% export_3_ss)%>%
  mutate(bonus=1)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\nonincluded_prolific_export3.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

nonincluded_bonus.df %>%
  filter(PROLIFIC_PID %in% export_4_ss)%>%
  mutate(bonus=1)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\nonincluded_prolific_export4.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

nonincluded_bonus.df %>%
  filter(PROLIFIC_PID %in% export_5_ss)%>%
  mutate(bonus=1)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\nonincluded_prolific_export5.csv', 
              sep=',',col.names = F, row.names=F, quote=F)

nonincluded_bonus.df %>%
  filter(PROLIFIC_PID %in% export_6_ss)%>%
  mutate(bonus=1)%>%
  dplyr::select(PROLIFIC_PID,bonus)%>%
  write.table('..\\experiments\\Hangman2\\data\\bonuses\\nonincluded_prolific_export6.csv', 
              sep=',',col.names = F, row.names=F, quote=F)
```

# Discussion

\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```

### 2023-01-18 docs/cogsci/PNTKN.Rmd
fixed naming of E2.click_log_with_boards (it was unclear how this structure was created)

@@
 author-information: > 
-    \author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of london \\ London, WC1E 7HX UK
+    \author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of London \\ London, WC1E 7HX UK
     \AND {\large \bf Chaz Firestone (chaz@jhu.edu)}
@@
 Moreover, deception often involves pretense behavior, which in turn relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state.
-For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") — you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
+For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
 What would I say?
@@
 500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
-They were selected based on their acceptance rate (>95%) and for being native English speakers.
+They were selected based on their acceptance rate (\>95%) and for being native English speakers.
 The entire experiment took approximately 20 minutes to complete.
@@
 > "In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
-What makes the game difficult is that you can’t see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
+What makes the game difficult is that you can't see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
 To sink a ship, you need to click on all of the squares it is located in.
@@
 > "This time your goal is different.
-In this round, we’re going to tell you where the ships are, but **we want you to act like you don’t know this information**.
-We’ve marked the ships’ locations with a cross, so you’ll know where they are the whole time; but your job is to play the game as if these hints aren’t there.
+In this round, we're going to tell you where the ships are, but **we want you to act like you don't know this information**.
+We've marked the ships' locations with a cross, so you'll know where they are the whole time; but your job is to play the game as if these hints aren't there.
 To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are.
@@
 
-Pretend games were similar to non pretend games in the total number of clicks — but were they also similar in *where* participants clicked?
+Pretend games were similar to non pretend games in the total number of clicks --- but were they also similar in *where* participants clicked?
 More specifically, did cell selections in pretend games make sense given the limited information those participants pretended to have?
@@
 
-E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')
-
-get_random_flat_position <- function(board_state) {
-
-  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
-  return(which(is.na(board_state))%>%sample(1))
-
-}
-
-get_optimal_flat_position <- function(posterior) {
-
-  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  return(which.max(posterior))
-
-}
-
-E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_optimal_flat_position(posterior),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'optimal')
-
-E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_random_flat_position(board_state),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'random')
-
-# E2.click_log <-E2.click_log %>%
+# a <-E2.click_log %>%
 #   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>%
@@
 
+E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')
+
+get_random_flat_position <- function(board_state) {
+
+  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
+  return(which(is.na(board_state))%>%sample(1))
+
+}
+
+get_optimal_flat_position <- function(posterior) {
+
+  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  return(which.max(posterior))
+
+}
+
+E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
+  mutate(flat_position = get_optimal_flat_position(posterior),
+         p_click = get_p_click(posterior,flat_position),
+         p_click_rank = get_p_click_rank(posterior,flat_position),
+         test_part = 'optimal')
+
+E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
+  mutate(flat_position = get_random_flat_position(board_state),
+         p_click = get_p_click(posterior,flat_position),
+         p_click_rank = get_p_click_rank(posterior,flat_position),
+         test_part = 'random')
+
 E2.mean_P_click_rank <- E2.click_log_with_boards %>%
@@
 To get at subtler dynamics in a more direct way, in a follow-up exploratory analysis we focused on decision latencies following a hit.
-We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn’t been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine (see Fig.
+We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn't been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine (see Fig.
 3C).
@@
 
-Our optimality analysis showed that pretenders’ click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be.
+Our optimality analysis showed that pretenders' click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be.
 However, due to the high number of possible board configurations, data from full games provide limited opportunity to compare cell selections for specific game states.
-In addition to asking, “What guides cell selections in pretend and non-pretend games?”, we also wanted to ask, “Where exactly would pretenders and non-pretenders click, given a specific board configuration?”.
+In addition to asking, "What guides cell selections in pretend and non-pretend games?", we also wanted to ask, "Where exactly would pretenders and non-pretenders click, given a specific board configuration?".
 
@@
 -->
+
 ## Judge trials
@@
 ```
+
 Despite the fact that pretenders were overall successful in mimicking patterns of cell selection and decision latency, they also left many clues for an independent observer to pick up on: they were slower, displayed exaggerated effects of decision uncertainty and hit success on response time, and were less optimal in choosing where to click next.
@@
 
-Many tasks (including writing papers for conferences) require us to overcome the ‘curse of knowledge’ — the challenge of putting ourselves in the shoes of someone who doesn’t know what we know.
+Many tasks (including writing papers for conferences) require us to overcome the 'curse of knowledge' --- the challenge of putting ourselves in the shoes of someone who doesn't know what we know.
 This is just an example of a more general fact: the ability to simulate counterfactual knowledge states is necessary for efficient and flexible communication.
@@
 Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state - in this case, a counterfactual knowledge state in which a known piece of information is unknown.
-Instead of relying on model inversion (e.g., “Which belief states would give rise to this behavior?”), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
+Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
 Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
@@
 Our subjects rose to the challenge.
-Their pretense behavior, although not perfect, closely resembled that of non-pretenders — not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
-Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: “What would I have done if I didn’t have this piece of knowledge?”
+Their pretense behavior, although not perfect, closely resembled that of non-pretenders --- not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
+Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: "What would I have done if I didn't have this piece of knowledge?"
 
@@
 This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain participants’ pretense ability in our data, for two indirect reasons.
+While we cannot fully rule out this interpretation, we think it is unlikely to explain participants' pretense ability in our data, for two indirect reasons.
 First, suppressing thoughts on demand is notoriously difficult [@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible.
@@
 
-In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects’ models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
+In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects' models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
 While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.
@@
 
+```{=tex}
 \setlength{\parindent}{-0.1in} 
 \setlength{\leftskip}{0.125in}
+```
 \noindent

### 2023-01-18 docs/cogsci/PNTKN.Rmd
fixd again

@@
 
-# a <-E2.click_log %>%
+# E2.click_log_with_boards <-E2.click_log %>%
 #   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>%

### 2023-01-18 analysis/decisionEntropy.Rmd
Measuring a tendency to make decisions according to strict rules

--- 2023-01-18 docs/cogsci/PNTKN.Rmd
@@
 ---
-title: "Using pretense behavior to explore counterfactual self-simulation"
-bibliography: r-references.bib
-csl: apa6.csl
-document-params: "10pt, letterpaper"
-
-author-information: > 
-    \author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of London \\ London, WC1E 7HX UK
-    \AND {\large \bf Chaz Firestone (chaz@jhu.edu)}
-    \AND {\large \bf Ian Phillips (ianphillips@jhu.edu)} \\ Johns Hopkins University \\ Baltimore, MD 21218, USA}
-
-abstract: >
-   What we do depends on what we know.
-But sometimes we try to decouple our behavior from our knowledge so that we appear not to know what we really do know.
-Such pretense behavior requires understanding how we would behave with different knowledge — and so provides a window into counterfactual self-simulation.
-However, little research has characterized and evaluated pretense relative to non-pretense behavior.
-Here, in a large-scale, pre-registered experiment, subjects played normal games of Battleships (trying to sink ships hidden in a grid), as well as ‘pretend’ games, where they were told all the ships’ locations but had to pretend they were playing normally.
-Relative to normal games, ‘pretend’ games demonstrated similar but exaggerated behavioral patterns.
-Furthermore, pretenders played rationally, but less so than non-pretenders.
-Despite these differences, ‘judge’ participants could not detect ‘pretend’ games.
-We discuss implications of these findings for accounts of theory of mind and metacognition.
-
-keywords: >
-    metacognition; theory of mind; pretense
-
-output: cogsci2016::cogsci_paper
-
-final-submission: \cogscifinalcopy
+title: "Decision entropy"
+author: "Matan Mazor"
+date: "2023-01-18"
+output: html_document
 ---
 
-```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
-                      fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, 
-                      message=F, sanitize = T)
-```
-
-```{r, libraries, echo=FALSE, warning=FALSE}
-
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
 library(groundhog)
@@
 
+load('../.Rdata')
 ```
 
-# Introduction
+## Decision entropy
 
-The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory].
-This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation].
-This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own.
+Unlike the main analysis, where we look at the likelihood of hitting a ship, here we look at the prior of clicking on a cell, regardless of possible ships' locations.
 
-Moreover, deception often involves pretense behavior, which in turn relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state.
-For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
-What would I say?
-How long would it take me to recover from the surprise?").
-This reliance of pretense behavior on self-simulation makes it an ideal opportunity to examine metacognitive knowledge about one's own mental states, and the potential reliance of this knowledge on a self-simulation.
-By comparing non-pretend and pretend behavior, we can ask which aspects of their own cognitive processes subjects can and cannot effectively simulate, and which aspects are not well-represented in their mental models of their own cognition.
+```{r get-priors, echo=FALSE, cache=TRUE}
 
-To this end, here we examine pretense in a game setting.
-Using an online version of the game Battleships, participants played a 'non-pretend' (normal) version of the game (trying to sink ships whose locations are unknown), as well as a 'pretend' version where they were given secret information about the location of hidden ships, but were instructed to behave as if they didn't have this information.
-
-# Methods
-
-A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb)
-
-```{r load-data, warning=FALSE}
-E2.df <- read.csv('../../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
-  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,4),sep='')) %>%
-  mutate(subj_id = factor(subj_id)) %>%
-  # the dash is breaking r
-  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
-  mutate(genuine_first=genuine_first=='True');
-
-E2.export <- read.csv('../../experiments/Battleships2/data/batch1/prolific_export_batch1.csv',na.strings=c(""," ","NA"));
-
-E2.bonus <- read.csv('../../experiments/Battleships2/data/batch1/bonus.csv',na.strings=c(""," ","NA"))
+E2.prior_from_group <- E2.click_log_with_boards %>% 
+  filter(test_part=='nonpretend')%>%
+  mutate(flat_position = factor(flat_position, levels=seq(25))) %>%
+  group_by(grid_number,click_number, flat_position) %>%
+  summarise(n=n())%>%
+  group_by(click_number, flat_position)%>%
+  do( data.frame(.,other_n=sapply(.$grid_number, function(i) mean(.$n[!.$grid_number %in% i])))) %>%
+  group_by(grid_number, click_number) %>%
+  mutate(p=other_n/sum(other_n)) %>%
+  dplyr::select(grid_number,click_number,flat_position,p) %>%
+  spread(flat_position,p, fill=0, sep='_') %>%
+  ungroup();
 
@@
 
-## Participants
+```{r decision-entropy-functions, echo=FALSE, cache=TRUE}
 
-The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
-500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
-They were selected based on their acceptance rate (\>95%) and for being native English speakers.
-The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
-
-## Procedure
-
-Participants were first instructed that the experiment, based on the game Battleships, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment.
-They were then presented with a leaderboard of previous players, and given the rules of the game:
-
-> "In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
-What makes the game difficult is that you can't see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
-To sink a ship, you need to click on all of the squares it is located in.
-If you hit part of a ship, the square will turn red.
-If there is no ship in the square, it will turn blue."
-
-We further explained that in this version of the game, ships can touch corners, but their sides can't touch.
-This explanation was accompanied by a visual presentation of legal and illegal ship configurations.
-
-After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half game (see below for details).
-The order of pretend and non-pretend blocks was counterbalanced between participants.
-The allocation of boards (spatial configurations of ships; see Fig.
-1A) to conditions was randomized between participants such that exactly one board was played in both pretend and non-pretend conditions, and this common board was different for different participants.
-The order of boards within a block was fully randomized, with the exception that half-games were always played last.
-
-### Non-pretend (normal) games
-
-In non-pretend games (Fig.
-1B), participants aimed to sink two 2-square patrol boats and one 3-square submarine with as few clicks as possible.
-An online counter of the number of clicks was displayed on the screen.
-After each game, feedback was given about the number of clicks and resulting number of points obtained.
-
-### Pretend games
-
-Participants in pretend games were given the same explanation of Battleships, and played a practice round. However, they were then given an additional twist:
-
-> "This time your goal is different.
-In this round, we're going to tell you where the ships are, but **we want you to act like you don't know this information**.
-We've marked the ships' locations with a cross, so you'll know where they are the whole time; but your job is to play the game as if these hints aren't there.
-To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are.
-We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you'll get bonus points.
-But if your games look different, you won't get these bonus points.
-Your number of clicks in this part will not affect your bonus.
-Only your ability to play like you had no hints."
-
-We informed participants that both the location and timing of their cell clicks will be measured.
-After one practice round and one comprehension question, participants played five pretend games (Fig.
-1C), followed by one pretend half-game.
-Each game was followed by a short message, reminding them that a game that looks similar to the games of participants who had no hints would be awarded 10 bonus points.
-
-### Half games
-
-In order to directly compare participants' pretend and non-pretend games for identical belief states (genuine or pretended ignorance about where the ships are hidden), participants completed one pretend and one non-pretend game given a partly finished board with the content of 7 cells already revealed (see Fig.
-1D).
-We designed our half games to produce a strong expectation to find a ship in specific cells, but not in others.
-The assignment of half-completed boards to pretend and non-pretend conditions was randomized between participants.
-
-### Judge trials
-
-In the final part of the experiment, participants observed the games of previous players and tried to determine who had hints and who didn't.
-On each trial, two empty grids were presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid (Fig.
-1F).
-The two grids corresponded to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders.
-Only games shorter than one minute were chosen for presentation in this part.
-For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
-Judge participants observed a real time replay of the two grids, showing not only where participants clicked, but also when.
-After making a decision, participants were informed whether they would receive the 10 bonus points, or alternatively, whether the pretender would receive them in the event the pretender managed to trick them.
-
-<!-- 500 players played 12 games of the game Battleships in two conditions presented in two blocks.
-The order of conditions was randomized between players.
-the standard, *non-pretend* condition, players started the game not knowing where the ships were and tried to sink all ships with as few clicks as possible.
-In the *pretend* condition, ship locations were marked with a cross such that players knew where they were hidden, and their task was to play the game as if they did not have this information.
-The sixth game in each block was a 'half-game', where participants completed the game with the content of some cells revealed by a previous player.
-Lastly, in judge trials, players observed a replay of the games of two previous players – one pretender and one non-pretender, and their task was to identify which was which.
-We created nine different 5x5 boards, each containing two 2-square patrol boats and one 3-square submarine.
-The assignment of boards to conditions was randomized across participants such that exactly one board was presented in both pretend and non-pretend blocks.
-(see Fig.
-1).
--->
-
-A more detailed description of the study procedure is provided in the study pre-registration document (LINK HIDDEN).
-Readers are also invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/NervzpM0Y0z).
-
-```{r design, echo=FALSE, fig.cap="Experimental Design. See Methods for details.", out.width = '100%'}
-knitr::include_graphics("../figures/Battleships_design.png")
-```
-
-# Results
-
-```{r subject-comments, echo=FALSE}
-E2.comments <- E2.df %>%
-  filter(!is.na(responses) & is.na(test_part)) %>%
-  dplyr::select(subj_id,responses) %>%
-  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
-  unnest(comments)%>%
-  dplyr::select(subj_id,worker_comments)%>%
-  filter(!is.na(worker_comments) & worker_comments!="") %>%
-  mutate(worker_comments = gsub("[\r\n]", "",worker_comments))
-
-E2.debrief <- E2.df %>%
-  filter(!is.na(responses) & is.na(test_part)) %>%
-  dplyr::select(subj_id,responses) %>%
-  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
-  unnest(comments)%>%
-  dplyr::select(subj_id,debrief)%>%
-  filter(!is.na(debrief) & debrief != "")%>%
-  mutate(debrief = gsub("[\r\n]", "",debrief))
-
-```
-
-We designed our analyses to explore subjects' capacity for self-simulation under a counterfactual knowledge state, and the limits of this capacity.
-We focused on where subjects clicked and when, and asked whether this differed between pretend and non-pretend games.
-All analyses were pre-registered unless otherwise specified.
-In our pre-registration document, we committed to separately analyzing participants according to whether they pretended before or after completing a non-pretend block.
-Due to space limitations, and since the results of the two groups mostly agreed, we report here the pooled results from both groups of participants, and mention whenever we found different results depending on block order.
-When directly comparing pretend and non-pretend blocks, we perform a between-subjects comparison using data from the first block only, i.e., pretend games from participants who pretended in the first block and non-pretend games from participants who played normally in the first block.
-We do so to ensure that any successful pretending is not due to memory of one's own behavior in a previous block, and that non-pretend games are not biased by experience with the pretend block.
-
-```{r num_clicks, echo=FALSE, cache=TRUE}
-
-E2.click_df <- E2.df %>%
-  dplyr::select(subj_id, 
-         test_part, 
-         grid_number, 
-         num_clicks,
-         click_log,
-         genuine_first) %>%
-  mutate(click_log = gsub("\'","\"", click_log)) %>%
-  filter(test_part=='pretend' | test_part=='nonpretend');
-
-E2.median_clicks <- E2.click_df %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(test_part,num_clicks);
-
-E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();
-
-E2.median_clicks_filtered <- E2.click_df %>%
-  filter(!(subj_id %in% E2.lucky_pretenders))%>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(test_part,num_clicks);
-
-p<- E2.click_df %>%
-    group_by(subj_id,test_part, genuine_first) %>%
-    summarise(num_clicks=median(num_clicks)) %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=num_clicks,fill=test_part))+
-    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='median number of clicks', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    theme_classic()
-    # geom_vline(xintercept= median_simulated)+
-    # facet_wrap(~genuine_first, nrow=2);
-
-ggsave('../docs/figures/E2_num_clicks.png',p,width=5,height=2, dpi=300)
-
-```
-
-```{r useful_functions, echo=FALSE, cache=TRUE}
-
-all_possible_boards <- read.csv('../../analysis/flat_boards.csv',header=FALSE);
-
-create_board_states <- function(flat_positions,hit) {
-
-  board_state = rep(NA,25)
-  board_states = c();
-  for (p in seq_along(flat_positions)) {
-    board_states = c(board_states, paste(board_state,collapse=','));
-    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
-  }
-
-  return(board_states)
-};
-
-get_likelihood <- function(board_state) {
+get_group_prior <- function(grid_i, click_i, board_state) {
 
@@
 
-  revealed_positions <- which(!is.na(board_state));
+  hidden_cells = ifelse(is.na(board_state),1,0);
 
-  boards_to_remove = c();
+  group_prior = E2.prior_from_group %>%
+    filter(grid_number==grid_i & click_number==click_i) %>%
+    select(flat_position_1:flat_position_25) %>%
+    as.numeric()
 
-  for (position in revealed_positions) {
-    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
-  }
+  group_prior_for_hidden_cells = group_prior*hidden_cells/sum(group_prior*hidden_cells)
 
-  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
-  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
-  likelihood[revealed_positions]=NA;
-
-  return(paste(likelihood,collapse=','))
+  return(paste(group_prior_for_hidden_cells,collapse=','))
 
@@
 
-get_posterior <- function(likelihood) {
-    likelihood = scan(text= likelihood, what = numeric(), sep="," , quiet = TRUE);
-    posterior = likelihood/sum(likelihood,na.rm=TRUE); # assuming a uniform prior
-    return(paste(posterior,collapse=','))
+get_prior_click <- function(prior, flat_position) {
+  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
+  return(prior[flat_position])
 }
 
-get_p_click <- function(posterior, flat_position) {
-  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  return(posterior[flat_position])
+get_prior_click_rank <- function(prior, flat_position) {
+  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
+  rank_prior = rank(-prior)
+  return(rank_prior[flat_position])
 }
 
-get_p_click_rank <- function(posterior, flat_position) {
-  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  rank_posterior = rank(-posterior)
-  return(rank_posterior[flat_position])
-}
-
-get_posterior_entropy <- function(posterior) {
-  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  posterior[which(posterior==0)]=NA; #to avoid infinity*0
-  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
+get_prior_entropy <- function(prior) {
+  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
+  prior[which(prior==0)]=NA; #to avoid infinity*0
+  entropy <- -sum(unlist(prior)*(unlist(log(prior))),na.rm=TRUE);
   return(entropy);
-}
-
-```
-
-```{r simulate_nclicks, echo=FALSE, cache=TRUE}
-
-## simulate number of clicks for random play
-
-num_clicks_sim = c();
-nsim = 10000
-boards = rep(NaN,25)
-
-for (i in 1:nsim) {
-  board = rep(c(0),25);
-  while (sum(board)<7) {
-    board[sample(1:25,1)]=1
-  }
-  nc = which(board==1)%>%tail(n=1);
-  num_clicks_sim = c(num_clicks_sim, nc);
-  if (nc<25) {
-    board[(nc+1):length(board)]=NaN;
-  }
-  boards = rbind(boards,board)
-}
-
-simulated_hits <- data.frame(click_number=1:25,mean_hit=colMeans(boards,na.rm=T))
-
-## OPTIMAL PLAYER
-
-simulate_optimal_game <- function(ship_positions) {
-  #initialize
-  board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-
-  while (nships>nhits) {
-
-    posterior <- get_likelihood(paste(board_state,collapse=',')) %>%
-      get_posterior() %>%
-      scan(text = ., what= numeric(), sep="," , quiet = TRUE);
-
-    best_shot = which.max(posterior);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    result = ship_positions[best_shot]
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    board_state[best_shot]=result;
-
-  }
-
-  return(nclicks)
-
-}
-
-num_clicks_optimal_sim = c();
-nsim = 100;
-nboards = all_possible_boards%>%nrow();
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  num_clicks_optimal_sim = c(num_clicks_optimal_sim,simulate_optimal_game(cur_board));
-}
-
-```
-
-## Total number of clicks
-
-To sink all ships, players had to click on at least 7 and at most 25 squares.
-A simulated player that clicks randomly had a median total click number of `r num_clicks_sim%>%median()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a median total click number of `r num_clicks_optimal_sim%>%median()`.
-Among our players, the median number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games (see Fig.
-2A).
-
-We observed no significant difference in the number of clicks between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend),E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
-However, a significant interaction with block order suggested that participants made fewer clicks in the second block compared to the first, regardless of which block was the pretend block (`r E2.median_clicks%>%gather('condition','nclicks',c('nonpretend','pretend'))%>%aov(nclicks ~ genuine_first * condition,data=.)%>%apa_print()%>%'$'(statistic)%>%'$'(genuine_first_condition)`).
-
-In `r E2.click_df%>%filter(num_clicks==7)%>%nrow()` pretend games from `r E2.lucky_pretenders%>%length()` players, games were completed after 7 clicks only, without ever missing a ship.
-This never happened in non-pretend games.
-We assumed that these participants did not follow task instructions, and excluded them from all following analyses.
-
-## Game optimality
-
-Pretend games were similar to non pretend games in the total number of clicks --- but were they also similar in *where* participants clicked?
-More specifically, did cell selections in pretend games make sense given the limited information those participants pretended to have?
-To ask this, we approximated optimal behavior by calculating the probability that a ship is hidden in each cell given available information, $p(ship(x_{i}))$, and the posterior probability that one should click on a square, assuming a uniform prior over cells $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$.
-Critically, in modeling pretend games we did not treat hints as part of this available information for extracting $p(ship(x_{i}))$, because an optimal player should ignore hints in choosing where to click next.
-Given this posterior map, a rational player should choose cells where $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3].
-To quantify optimality, before each cell selection we computed the posterior probability map for all 'unknown' cells.
-Then, we ranked cells from high to low according to their posterior probability and recorded the rank of the chosen cell: a lower rank indicating more optimal behavior.
-
-```{r parse-click-log, echo-FALSE, cache=TRUE}
-
-E2.click_log <- readRDS('../../experiments/Battleships2/data/batch1/click_log.Rda');
-
-# E2.click_log <- data.frame(matrix(ncol=10,nrow=0, 
-#                                dimnames=list(NULL, 
-#                                              c("subj_id",
-#                                               "test_part", 
-#                                               "grid_number", 
-#                                               "num_clicks",
-#                                               "i",
-#                                               "j",
-#                                               "hit",
-#                                               "t",
-#                                               "click_number",
-#                                               "genuine_first"))))
-# 
-# 
-# for (row in 1:nrow(E2.click_df)) {
-#   
-#     subject_click_log <- data.frame(fromJSON(E2.click_df[row, ]$click_log)) %>%
-#     mutate(
-#       click_number = 1:n(),
-#       subj_id = E2.click_df[row, ]$subj_id,
-#       test_part = E2.click_df[row, ]$test_part,
-#       grid_number = E2.click_df[row, ]$grid_number,
-#       num_clicks = E2.click_df[row, ]$num_clicks,
-#       genuine_first = E2.click_df[row, ]$genuine_first
-#     )
-#     
-#     E2.click_log <- rbind(E2.click_log, subject_click_log);
-# }
-#     
-# 
-# E2.click_log <- E2.click_log %>%
-#   relocate(subj_id, .before = i) %>%
-#   relocate(genuine_first, .before = i) %>% 
-#   relocate(test_part, .before=i) %>% 
-#   relocate(grid_number, .before=i) %>% 
-#   relocate(click_number, .before=i) %>%
-#   group_by(subj_id,test_part,grid_number) %>%
-#   mutate(RT=t-lag(t,default=0)) %>%
-#   filter(!(subj_id %in% E2.lucky_pretenders))%>%
-#  mutate(hit_bin = hit !=0,
-#         lag3 = lag(hit,3)!=0,
-#         lag2 = lag(hit,2)!=0,
-#         lag1 = lag(hit,1)!=0,
-#         lead1 = lead(hit,1)!=0,
-#         lead2 = lead(hit,2)!=0,
-#         lead3 = lead(hit,3)!=0)
-
-```
-
-```{r simualte_data, echo=FALSE, cache=TRUE}
-
-simulate_optimal_game <- function(ship_positions) {
-  #initialize
-  cur_board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-  board_state = c()
-  flat_position = c();
-  click_number = c();
-  hit_bin = c();
-
-  while (nships>nhits) {
-
-    board_state = c(board_state,paste(cur_board_state,collapse=','));
-
-    posterior <- get_likelihood(paste(cur_board_state,collapse=',')) %>%
-      get_posterior() %>%
-      scan(text = ., what= numeric(), sep="," , quiet = TRUE);
-
-    best_shot = which.max(posterior);
-    flat_position = c(flat_position,best_shot);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    click_number = c(click_number,nclicks)
-
-    result = ship_positions[best_shot]
-
-    hit_bin = c(hit_bin,ifelse(result==1,1,0));
-
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    cur_board_state[best_shot]=result;
-
-  }
-  return(data.frame(board_state,flat_position,hit_bin,click_number))
-
-}
-
-simulate_random_game <- function(ship_positions) {
-  #initialize
-  cur_board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-  board_state = c()
-  flat_position = c();
-  click_number = c();
-  hit_bin = c();
-
-  while (nships>nhits) {
-
-    board_state = c(board_state,paste(cur_board_state,collapse=','));
-
-    random_guess = which(is.na(cur_board_state))%>%sample(1);
-    flat_position = c(flat_position,random_guess);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    click_number = c(click_number,nclicks)
-
-    result = ship_positions[random_guess]
-
-    hit_bin = c(hit_bin,ifelse(result==1,1,0));
-
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    cur_board_state[random_guess]=result;
-
-  }
-  return(data.frame(board_state,flat_position,hit_bin,click_number))
-
-}
-
-nsim = 100;
-nboards = all_possible_boards%>%nrow();
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  if (i==1) {
-    simulated_optimal = simulate_optimal_game(cur_board)%>%mutate(game=i)
-  } else {
-    simulated_optimal = bind_rows(simulated_optimal, simulate_optimal_game(cur_board)%>%mutate(game=i))
-  }
-}
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  if (i==1) {
-    simulated_random = simulate_random_game(cur_board)%>%mutate(game=i)
-  } else {
-    simulated_random = bind_rows(simulated_random, simulate_random_game(cur_board)%>%mutate(game=i))
-  }
 }
@@
 ```
+## Including Plots
 
-The mean posterior rank of non-pretend cell selections was `r E2.mean_P_click_rank%>%filter(test_part=='nonpretend')%>%pull(p_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (see Fig.
-2B; `r E2.mean_P_click_rank%>%filter(test_part=='random')%>%pull(p_click_rank)%>%mean()%>%printnum()`, `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Pretend games were significantly less optimal than non-pretend games (`r E2.mean_P_click_rank%>%filter(test_part=='pretend')%>%pull(p_click_rank)%>%mean()%>%printnum()`; `r apa_print(t.test(E2.mean_P_click_rank%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`), but still more optimal than those of a random agent (`r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-In other words, the optimality of pretend games relative to random cell selection was not merely due to the fact that pretenders clicked on ships more than expected by chance.
-Even when missing a ship, their cell selections made sense given the limited information they pretended to have.
+You can also embed plots, for example:
 
-```{r numclick_optimality, echo=FALSE, fig.cap="Experimental Design", out.width = '100%', fig.cap = "A: The total number of clicks in pretend and non-pretend games was highly similar.
-B: Non-pretend games were significantly more optimal than pretend games, but both were more optimal than what is expected by random chance.
-Box: An illustration of the difference between success and optimality.
-A given click (highlighted with a black border) can reveal a ship (upper row) or not (lower row) regardless of whether the click makes sense given available information (left column) or not (right column).
-"}
-knitr::include_graphics("../figures/numclicks_optimality.png")
+```{r pressure, echo=FALSE}
+plot(pressure)
 ```
 
-## Response time analysis
-
-```{r total_time, echo=FALSE}
-
-E2.total_time_df <- E2.click_log %>%
- filter(click_number==num_clicks)%>%
-  mutate(t=t/1000)
-
-E2.median_time <- E2.total_time_df %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(t=median(t)) %>%
-  spread(test_part,t);
-
-# anova
-E2.time_anva <- aov(t ~ genuine_first * condition, 
-                         data = E2.median_time %>%
-                           gather('condition','t',nonpretend,pretend) %>%
-                           mutate(condition=as.factor(condition)));
-
-E2.median_time_per_click <- E2.click_log %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(RT=median(RT)) %>%
-  spread(test_part,RT) %>%
-  mutate(diff=pretend-nonpretend);
-
-# anova
-E2.click_time_anva <- aov(RT ~ genuine_first * condition, 
-                         data = E2.median_time_per_click %>%
-                           gather('condition','RT',nonpretend,pretend) %>%
-                           mutate(condition=as.factor(condition)))
-```
-
-In general, participants gave slower responses when pretending compared to when playing normally.
-Despite the similar number of clicks per game, the median game duration was `r E2.total_time_df%>%filter(test_part=='nonpretend')%>%pull(t)%>%median()%>%printnum()` seconds in the non-pretend condition and `r E2.total_time_df%>%filter(test_part=='pretend')%>%pull(t)%>%median()%>%printnum()` seconds in the pretend condition, which was significantly longer (`r t.test(E2.median_time$pretend-E2.median_time$nonpretend)%>%apa_print()%>%'$'(statistic)`).
-This difference showed a significant interaction with the order of conditions, such that completing games in the first block took longer (`r apa_print(E2.time_anva%>%summary())$full_result$genuine_first_condition`).
-
-In our instructions to pretenders, we asked them to pretend not to know where the ships were, not only in *where* they chose to click, but also in *when* they clicked.
-In the following set of analyses, we identified patterns in click latency in non-pretend games, and asked whether the same patterns are also observed in pretend games.
-We focused on the effects of hitting versus missing a ship, decision uncertainty, and ship completion.
-
-```{r RT_by_accuracy}
-
-N_perm <- 1000;
-bootstrap_error <- function(x, N_perm) {
-  N <- length(x)
-  medians = c();
-  for (i in 1:N_perm) {
-    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
-  };
-  return(sd(medians))
-}
-
-E2.RT_by_lag3 <- E2.click_log %>%
-  filter(click_number>3) %>%
-  group_by(subj_id,test_part,lag3,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag3,RT,sep='_') %>%
-  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);
-
-E2.RT_by_lag2 <- E2.click_log %>%
-  filter(click_number>2) %>%
-  group_by(subj_id,test_part,lag2,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag2,RT,sep='_') %>%
-  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);
-
-E2.RT_by_lag1 <- E2.click_log %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,lag1,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag1,RT,sep='_') %>%
-  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
-
-E2.RT_by_hit <- E2.click_log %>%
-  filter(click_number>0) %>%
-  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(hit_bin,RT,sep='_') %>%
-  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
-
-E2.RT_by_lead1 <- E2.click_log %>%
-  filter(click_number<num_clicks) %>%
-  group_by(subj_id,test_part,lead1,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead1,RT,sep='_') %>%
-  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
-
-E2.RT_by_lead2 <- E2.click_log %>%
-  filter(click_number<num_clicks-1) %>%
-  group_by(subj_id,test_part,lead2,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead2,RT,sep='_') %>%
-  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)
-
-E2.RT_by_lead3 <- E2.click_log %>%
-  filter(click_number<num_clicks-2) %>%
-  group_by(subj_id,test_part,lead3,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead3,RT,sep='_') %>%
-  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)
-
-E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
-  merge(E2.RT_by_lag2) %>%
-  merge(E2.RT_by_lag1) %>%
-  merge(E2.RT_by_hit) %>%
-  merge(E2.RT_by_lead1) %>%
-  merge(E2.RT_by_lead2) %>%
-  merge(E2.RT_by_lead3);
-
-E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
-  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
-  mutate(position = ifelse(position=="lag3_diff", 3,
-                           ifelse(position=="lag2_diff",2,
-                                  ifelse(position=="lag1_diff",1,
-                                         ifelse(position=="hit_diff",0,
-                                                ifelse(position=="lead1_diff",-1,
-                                                       ifelse(position=="lead2_diff",-2,-3
-                                                              ))))))) %>%
-  dplyr::select(subj_id,test_part,genuine_first,position,difference);
-
-p<- E2.RT_by_hit_lags_long %>% 
-  filter(position<2 & position > -1)%>%
-  group_by(test_part,position) %>%
-  summarise(median_difference=median(difference),
-            sem_difference=bootstrap_error(difference,N_perm)) %>%
-  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  scale_x_continuous(breaks=0:1,labels=c('hit on current trial', 'hit on previous trial'))+
-  geom_abline(slope=0,intercept=0,linetype=3)+
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.pdf',p,width=7)
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.png',p,width=4, height=3.5, dpi=300)
-
-```
-
-### Effects of hitting versus missing a ship
-
-In non-pretend games, players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='nonpretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) when hitting compared to missing a ship (see Fig.
-3A; `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-The same effect was observed in pretend games: players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='pretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) in hits compared to misses (`r E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-However, this difference between hits and misses was significantly more pronounced in pretend games (`r apa_print(t.test(E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff), E2.RT_by_hit_lags%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)))$statistic`).
-
-The opposite effect was observed for a hit on the previous click: players were slower by `r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) after hitting a ship (`r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-This effect was also observed in pretend games, where players were slower by `r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='pretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) following a hit on the previous click (`r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Again, the absolute effect in ms was exaggerated in the pretend condition (`r t.test(E2.RT_by_lag1%>%filter(!genuine_first & test_part=='pretend')%>%pull(lag1_diff),E2.RT_by_lag1%>%filter(genuine_first & test_part=='nonpretend')%>%pull(lag1_diff))%>%apa_print()%>%'$'(statistic)`).
-
-### Effects of decision uncertainty
-
-When playing Battleships, it is sometimes clear what the next cell selection should be, and sometimes more difficult to decide where to click next.
-To capture this notion of decision uncertainty, we calculated the entropy of the posterior distribution over cell selections $H(P)=-\sum_{i=1}^{k}P(x_i) \log P(x_{i})$, where $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$, and asked how this measure relates to decision latency, or the time taken to click on the next cell.
-$H(P)$ is high when players need to decide between multiple cells with a similar probability of hiding a ship, and low when there are only a few candidates with a high probability of hiding a ship.
-For every player and condition separately, we fitted a multiple linear regression to predict decision latency based on $H(P)$ and $H(P)^2$.
-The resulting coefficients were then subjected to a group-level inference.
-The first cell selection of each game was excluded from this analysis, because entropy was constant for the first click.
-
-```{r entropy-RT, echo=FALSE, cache=TRUE}
-
-E2.RT_by_entropy <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  mutate(entropy_bin=round(entropy*2)/2) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(test_part,subj_id,entropy_bin) %>%
-  summarise(RT=median(RT),
-            corrected_RT = median(corrected_RT)) %>%
-  group_by(test_part,entropy_bin) %>%
-  summarise(median_RT=median(RT),
-            sem_RT=bootstrap_error(RT,N_perm),
-            median_corrected_RT=median(corrected_RT),
-            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))
-
-p <- E2.RT_by_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
-# ggsave('../figures/E2_RT_by_entropy_no_split.svg',p,width=5,height=2,dpi=300)
-
-p <- E2.RT_by_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of posterior', y='median residual RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.RT_by_entropy_coefs <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  mutate(entropy_centered = entropy-mean(entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value)) %>%
-  group_by(genuine_first,subj_id,test_part,term) %>%
-  summarise(estimate = mean(estimate))
-
-E2.RT_by_entropy_coefs_summary <- E2.RT_by_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.RT_by_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.corrected_RT_by_entropy_coefs <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  mutate(entropy_centered = entropy-mean(entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value))
-
-E2.corrected_RT_by_entropy_coefs_summary <- E2.corrected_RT_by_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.corrected_RT_by_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-```
-
-In non-pretend games, we found no evidence for a linear relation between decision entropy and decision latency (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).
-We did however observe a block order effect, with a significantly negative linear modulation only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
-In pretend games this negative relation between $H(P)$ and click latency was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).
-
-A negative linear modulation of $H(P)$ on decision latency is surprising, as it suggests that participants were quicker to decide when uncertainty was high.
-However, visual inspection of the RT/entropy curves (Fig.
-3B) reveals a pronounced quadratic modulation, such that in both pretend and non-pretend games slower decision times are observed for intermediate $H(P)$ values of around 2.
-Similar to the linear effect of $H(P)$ on decision latency, here too a quadratic modulation in non-pretend games was observed only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
-Again, in pretend games this negative quadratic relation was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared')%>%pull(estimate)%>%t.test())$statistic`).
-A between-subjects comparison focusing on games from the first block revealed that this negative quadratic effect was stronger in pretend, compared to non-pretend games (`r apa_print(t.test(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate),E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)))$statistic`).
-
-```{r latency, echo=FALSE, out.width = '100%', fig.cap = "A: A subtraction of decision latencies in misses versus hits in the current and previous cell selections, in pretend and non-pretend games.
-B.
-Upper panel: Effect of the entropy of the posterior distribution over cell selections (a measure of decision uncertainty) on decision latency in pretend and non-pretend games.
-Lower panel: mean beta coefficients for a multiple regression model, predicting decision latency from entropy and entropy squared.
-C: Decision latency after a hit, as a function of hit number."}
-knitr::include_graphics("../figures/decision_latencies.png")
-```
-
-<!-- $H(P)$ is highest before the first click in a game and lowest just before the last one.
-To make sure any results are not driven by a general effect of click number on decision latency, and in deviation from our pre-registered analysis plan, we first mean-centered all decision latencies within individual subjects, conditions, and click numbers.
-This way, any effect of $H(P)$ on latency cannot reflect effects of click order.
--->
-
-## Effects of ship completion
-
-```{r}
-
-get_hit_number <- function(hit) {
-
-  num_A <- 0;
-  num_B <- 0;
-  num_C <- 0;
-
-  hit_number <- c();
-
-  for (i in seq(length(hit))) {
-    if (hit[i]==0) {
-      hit_number = c(hit_number,NaN)
-    } else if (hit[i]=='A') {
-      num_A = num_A + 1;
-      hit_number = c(hit_number,num_A)
-    } else if (hit[i]=='B') {
-      num_B = num_B + 1;
-      hit_number = c(hit_number,num_B)
-    } else if (hit[i]=='C') {
-      num_C = num_C + 1;
-      hit_number = c(hit_number,num_C)
-    }
-  }
-
-  return(hit_number)
-}
-
-revealed_three <- function(hit_number) {
-
-  revealed_three <- c();
-  three = FALSE;
-
-  for (i in seq(length(hit_number))) {
-    revealed_three = c(revealed_three, three);
-     if (!(is.na(hit_number[i])) & hit_number[i]==3) {
-      three = TRUE
-    } 
-  }
-
-  return(revealed_three)
-}
-
-E2.hit_order <- E2.click_log %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  summarise(click_number = click_number,
-            hit = hit,
-            RT = RT,
-            hit_number = get_hit_number(hit),
-            revealed_three = revealed_three(hit_number),
-            hit_type = ifelse(hit_number==1, 'first',
-                              ifelse(hit_number==2 & !revealed_three, 'second',
-                                     ifelse(hit_number==2 & revealed_three, 'secondlast',
-                                            ifelse(hit_number==3,'thirdlast','sea'))))) %>%
-  mutate(prev_hit_type = lag(hit_type),
-         prev_hit = lag(hit))
-
-E2.RT_by_hit_order <- E2.hit_order %>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(RT=median(RT))
-
-p <- E2.RT_by_hit_order %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(RT = median(RT),
-            sem_RT = bootstrap_error(RT,N_perm)) %>%
-  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_previous_hit_number_collapse.png',p,width=5,height=2,dpi=300)
-
-##correct for trial order
-
-E2.corrected_RT_by_hit_order <- E2.hit_order %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT))%>%
-  rowwise() %>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(corrected_RT=median(corrected_RT))
-
-p <- E2.corrected_RT_by_hit_order %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
-            corrected_RT = median(corrected_RT)) %>%
-  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)
-
-## correct for true ship size
-
-E2.RT_by_hit_order_boats_only <- E2.hit_order %>%
-  filter(prev_hit!='A')%>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(RT=median(RT))
-
-p <- E2.RT_by_hit_order_boats_only %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(RT = median(RT),
-            sem_RT = bootstrap_error(RT,N_perm)) %>%
-  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)
-
-E2.corrected_RT_by_hit_order_boats_only <- E2.hit_order %>%
-  filter(prev_hit!='A')%>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT))%>%
-  rowwise() %>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(corrected_RT=median(corrected_RT))
-
-p <- E2.corrected_RT_by_hit_order_boats_only %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
-            corrected_RT = median(corrected_RT)) %>%
-  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed_boats_only.png',p,width=5,height=2,dpi=300)
-
-```
-
-The previous analyses reveal robust associations between game state and decision latency, that are highly similar, and even exaggerated, in pretend compared to non-pretend games.
-To get at subtler dynamics in a more direct way, in a follow-up exploratory analysis we focused on decision latencies following a hit.
-We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn't been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine (see Fig.
-3C).
-In the first case, players know the ship must continue in one of the neighboring cells.
-In the second case, there is a good chance the ship continues (if this ship turns out to be a submarine).
-In the third and fourth cases, it is clear that the ship is fully sunk.
-
-In non-pretend games, participants were significantly slower to select the next cell when they knew they had just completed a ship (categories 3 and 4) compared to when they just hit a ship, but were not sure (category 2) or knew they had not completely sunk it (category 1; `r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-Specifically, we find that participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-
-Strikingly, we found the exact same pattern in pretend games.
-Players were faster to make the next cell selection when they pretended to think that the current ship might not be fully sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).
-This was not merely a difference between the first, second and third hits: participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-Further analysis confirmed that this effect remained significant when controlling for click number (`r apa_print(E2.corrected_RT_by_hit_order%>%spread(prev_hit_type,corrected_RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`), and when restricting the analysis to the second hit of a ship that is in fact of size-two (`r apa_print(E2.RT_by_hit_order_boats_only%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).
-
-This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their their response latency in the same way it would have been affected had they been in a non-pretend game.
-
-## Half-games
-
-Our optimality analysis showed that pretenders' click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be.
-However, due to the high number of possible board configurations, data from full games provide limited opportunity to compare cell selections for specific game states.
-In addition to asking, "What guides cell selections in pretend and non-pretend games?", we also wanted to ask, "Where exactly would pretenders and non-pretenders click, given a specific board configuration?".
-
-To achieve this, the sixth game in each block started not with an empty grid, but with the contents of some cells already revealed by a previous player.
-As before, pretenders also knew where the remaining ships were hidden, but tried to play as if they only knew what was known to this previous player.
-Having cell selections from 250 players for each board configuration and condition allowed us to plot and compare the distribution of clicks under a genuine, or pretend, knowledge state.
-
-```{r, child=c('../half_games.Rmd')}
-```
-
-For example, look at Board A in Fig.
-4 (upper left corner).
-Imagine that you need to select your next move, knowing that exactly two size-two patrol boats and one size-three submarine are hiding in the water.
-Where would you click next?
-Perhaps one of the two revealed ships is the size-three submarine.
-In that case you want to click on the first cell in the third row, or on the third cell in the fifth row.
-Or maybe these are the two size-two patrol boats.
-This would mean the submarine is hiding in some other place, and for it not to touch sides with the boats this must be in the first row.
-The hit probability map in the second column visualizes this reasoning, by plotting the normalized probability that a ship is hidden in different cells.
-As can be seen, the highest probabilities are to the sides of the revealed ships, and in the first row.
-
-In the third column we plot the distribution of clicks for non-pretend players.
-This distribution is in agreement with the hit probability map (board A: `r cor_A_nonpretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_nonpretend%>%apa_print()%>%'$'(full_result)`).
-Finally, in the fourth column we plot the distribution of cell selections for pretend players.
-Although noisier, this distribution is also in agreement with the hit probability map (board A: `r cor_A_pretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_pretend%>%apa_print()%>%'$'(full_result)`), and more importantly, with the hit distribution of non-pretenders for the same board configuration (board A: `r cor_A_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`, board B: `r cor_B_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`).
-
-```{r half-games, echo=FALSE, out.width = '100%', fig.cap = "Half games.
-First column: the two boards, as they appeared to pretenders.
-Second column: objective hit probability, given game state.
-Third and fourth columns: empirical click probabilities in non-pretend and pretend games, respectively."}
-knitr::include_graphics("../figures/Half_games.png")
-```
-
-<!-- Furthermore, compared to non-pretend games, in pretend games participants were more likely to make irrational clicks on cells where a ship could not in fact be hiding (a median of `r E2.num_irrational_clicks%>%filter(test_part=='pretend')%>%pull(nic)%>%median()` per game in pretend versus `r E2.num_irrational_clicks%>%filter(test_part=='nonpretend')%>%pull(nic)%>%median()` in non-pretend games.
-`r E2.num_irrational_clicks%>%spread(test_part,nic)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
--->
-
-## Judge trials
-
-```{r condition-classifier, echo=FALSE, eval=FALSE}
-
-set.seed(42);
-
-gamewise_data <- E2.click_log_with_boards %>% 
-  filter(!genuine_first | test_part=='pretend') %>% 
-  group_by(subj_id,test_part,grid_number) %>%
-  filter(sum(RT)<=60000) %>%
-  summarise(median_RT=median(RT),
-            mean_p_click_rank = mean(p_click_rank),
-            first_click_latency=RT[click_number==1],
-            num_irrational_clicks = sum(p_click==0)) %>%
-  ungroup()%>%
-  select(test_part,mean_p_click_rank,median_RT,num_irrational_clicks);
-
-gamewise_data_nonpretend = gamewise_data%>%filter(test_part=='nonpretend') %>%mutate(nrow=1:n())
-gamewise_data_pretend = gamewise_data%>%filter(test_part=='pretend')
-rows <- sample(nrow(gamewise_data_nonpretend))
-gamewise_data_pretend <- gamewise_data_pretend[rows,]%>%mutate(nrow=1:n())
-
-pair_gamewise_data <- full_join(gamewise_data_nonpretend,
-                                gamewise_data_pretend,
-                                by='nrow',
-                                suffix=c('.nonpretend','.pretend')) %>%
-  mutate(which_side = sample(c(-1,1),nrow(gamewise_data_pretend),replace=TRUE),
-         mean_p_click = which_side*(mean_p_click_rank.nonpretend-mean_p_click_rank.pretend),
-         median_RT = which_side*(median_RT.nonpretend-median_RT.pretend),
-         num_irrational_clicks = which_side*(num_irrational_clicks.nonpretend-num_irrational_clicks.pretend),
-         which_side=factor(which_side)) %>%
-  select(which_side,mean_p_click,median_RT,num_irrational_clicks)
-
-train_control <- trainControl(method = "CV", number=10, sampling='down')
-
-# train the model on training set
-model <- train(test_part ~ .,
-               data = gamewise_data,
-               trControl = train_control,
-               method = "svmLinear",
-               preProcess = c("center","scale"))
-
-pair_model <- train(which_side ~ .,
-               data = pair_gamewise_data,
-               trControl = train_control,
-               method = "svmLinear",
-               preProcess = c("center","scale"))
-
-# print cv scores
-
-```
-
-Despite the fact that pretenders were overall successful in mimicking patterns of cell selection and decision latency, they also left many clues for an independent observer to pick up on: they were slower, displayed exaggerated effects of decision uncertainty and hit success on response time, and were less optimal in choosing where to click next.
-To illustrate, a support vector machine (SVM) algorithm reached an accuracy level of 72% in linearly classifying condition (pretend / non-pretend) based on just three basic summary features: median decision latency, game optimality score, and number of irrational clicks per game.
-
-```{r judge}
-
-get_common_board = function(grid_number,test_part) {
-  pretend_grids = grid_number[which(test_part=='pretend')];
-  nonpretend_grids = grid_number[which(test_part=='nonpretend')];
-  common_board = intersect(pretend_grids,nonpretend_grids);
-  return(common_board)
-};
-
-common_boards <- E2.df %>%
-  filter(test_part=='pretend' | test_part=='nonpretend') %>%
-  dplyr::select(subj_id,grid_number,test_part) %>%
-  group_by(subj_id) %>%
-  summarise(common_board = get_common_board(grid_number,test_part))
-
-used_for_pretend <- function(s, g) {
-
-  return (E2.df%>%filter(test_part=='pretend' &
-                           subj_id==s & 
-                           grid_number == g) %>%
-            nrow()==1)
-}
-
-used_for_nonpretend <- function(s, g) {
-
-  return (E2.df%>%filter(test_part=='nonpretend' &
-                           subj_id==s & 
-                           grid_number == g) %>%
-            nrow()==1)
-}
-
-E2.judge_df <- E2.df %>%
-  filter(test_part=='judge' & !(cheat_player %in% E2.lucky_pretenders))%>%
-  dplyr::select(subj_id, 
-         grid_number, 
-         cheat_player,
-         genuine_player,
-         cheater,
-         noncheater,
-         decision,
-         replay_log,
-         choose_pretender,
-         correct)%>%
-  rowwise()%>%
-  mutate(choose_pretender=choose_pretender=='True',
-         used_for_pretend = used_for_pretend(subj_id,grid_number),
-         used_for_nonpretend = used_for_nonpretend(subj_id,grid_number),
-         common_board = used_for_pretend & used_for_nonpretend,
-         board_type = ifelse(common_board, 'common',
-                             ifelse(used_for_pretend, 'pretend',
-                                    ifelse(used_for_nonpretend,'nonpretend','new')))
-)
-
-p <- E2.judge_df %>%
-  group_by(subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) +
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') +
-  scale_x_continuous(breaks=c(0,1,2,3,4,5));
-
-ggsave('../figures/E2_correct_guesses.png',p,width=5,height=3,dpi=300)
-
-p <- E2.judge_df %>%
-  mutate(choose_pretender = ifelse(choose_pretender,
-                                   'who had hints?', 
-                                   'who had no hints?')) %>% 
-  group_by(choose_pretender,subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) + 
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') + 
-  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
-  facet_wrap(~choose_pretender,nrow=2);
-
-ggsave('../figures/E2_correct_guesses_by_choose_pretender.png',p,width=5,height=3,dpi=300);
-
-p <- E2.judge_df %>% 
-  group_by(board_type,subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) + 
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') + 
-  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
-  facet_wrap(~board_type,nrow=1);
-
-ggsave('../figures/E2_correct_guesses_by_board_type.png',p,width=5,height=3,dpi=300)
-```
-
-Despite this, participants were never significantly above chance in judging which of the two presented games came from a pretender (mean accuracy: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%; t test against 50: `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%t.test(mu=50))$statistic`).
-
-For a subset of `r E2.judge_df%>%filter(common_board)%>%pull(correct)%>%length()` players, one of the five boards presented in the judging block was already presented both in the pretend and in the non-pretend blocks.
-Still, even this subset of participants who had just experienced a board both as pretenders and as non-pretenders were not significantly above chance in telling which of two other players had hints for this same board (mean accuracy: 55%, $p=.141$).
-
-## Are good pretenders also better judges?
-
-```{r judge-pretend-cor, echo=FALSE}
-
-unidentifiable_ss <- E2.judge_df %>%
-  mutate(subj_num=substr(as.character(subj_id),1,3))%>%
-  group_by(subj_num)%>%
-  summarise(subj_id=subj_id, n=n())%>%
-  filter(n>5)%>%
-  pull(subj_id)%>%
-  unique();
-
-E2.judge_performance <- E2.judge_df %>%
-  filter(!(subj_id %in% unidentifiable_ss)) %>%
-  mutate(subj_num=as.numeric(substr(as.character(subj_id),1,3)))%>%
-  group_by(subj_num)%>%
-  summarise(judge_rate=sum(correct)/5);
-
-E2.pretend_performance <- E2.judge_df %>%
-  filter(!(subj_id %in% unidentifiable_ss)) %>%
-  mutate(subj_num=cheat_player)%>%
-  group_by(subj_num)%>%
-  summarise(ntrials=n(),
-            successful_pretend=ntrials-sum(correct),
-            pretend_rate = successful_pretend/ntrials);
-
-E2.performance_df <- merge(E2.judge_performance,E2.pretend_performance);
-
-p <- E2.performance_df %>%
-  ggplot(aes(x=judge_rate,y=pretend_rate)) +
-  geom_jitter() + 
-  labs(x='judge accuracy',y='pretense success rate') +
-  theme_classic()
-
-ggsave('../figures/E2_judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)
-
-min_njudge = seq(0,10);
-cors =c()
-ns = c()
-
-for (minn in min_njudge) {
-
-  temp_df = E2.performance_df %>%
-    filter(ntrials>minn)
-
-  cur_cor = cor(temp_df$pretend_rate,temp_df$judge_rate);
-  cur_n = temp_df%>%nrow()
-  cors = c(cors,cur_cor)
-  ns = c(ns, cur_n)
-}
-
-judge_pretend_correlation_df <- data.frame(min_njudge,cors,ns);
-
-```
-
-```{r judge-optimality-cor, echo=FALSE}
-E2.judge_summary <- E2.judge_df %>%
-  group_by(subj_id) %>%
-  summarise(judge = sum(correct));
-
-E2.optimality_summary <- E2.click_log_with_boards %>%
-  group_by(subj_id,test_part) %>%
-  summarise(p_click_rank = mean(p_click_rank)) %>%
-  spread(test_part,p_click_rank)
-
-E2.judge_summary <- E2.judge_summary %>%
-  merge(E2.optimality_summary)
-```
-
-The abilities to pretend and to detect pretense in others both rely on some form of theory of mind: reasoning about counterfactual knowledge states of oneself in the first case, and of others in the second.
-For example, in scrub jays, individual birds that pilfered another bird's caches were also more careful in hiding their own food, suggesting an ability to project from their own experience of stealing to the behavior of others [@emery2001effects].
-Along similar lines, we reasoned that those participants whose pretend games better resembled the games of genuine players would also be more accurate in detecting pretense in others.
-To our surprise, this was not at all the case (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate))$full_result`).
-Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend))$full_result`).
-
-Despite our considerable sample size of 500 participants, we are cautious in making strong claims based on the absence of a correlation between pretense and pretense detection abilities in our sample.
-We note, however, that previous developmental research has identified weaker correlations with theory of mind measures for a tendency to intentionally deceive another player in a game setting, compared to an ability to detect and understand deception, potentially indicating partly different mechanisms [@shultz1981development].
-
-# Discussion
-
-Many tasks (including writing papers for conferences) require us to overcome the 'curse of knowledge' --- the challenge of putting ourselves in the shoes of someone who doesn't know what we know.
-This is just an example of a more general fact: the ability to simulate counterfactual knowledge states is necessary for efficient and flexible communication.
-To do this, we must have an internal self-model that can generate thoughts and behavior for hypothetical beliefs, desires, and perceptions.
-
-Research on Bayesian Theory of Mind provides some support for the existence of such a model by measuring subjects' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational], or implicitly [@onishi2005fifteen; @liu2017ten].
-Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state - in this case, a counterfactual knowledge state in which a known piece of information is unknown.
-Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
-Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
-
-Our subjects rose to the challenge.
-Their pretense behavior, although not perfect, closely resembled that of non-pretenders --- not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
-Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: "What would I have done if I didn't have this piece of knowledge?"
-
-An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game.
-This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain participants' pretense ability in our data, for two indirect reasons.
-First, suppressing thoughts on demand is notoriously difficult [@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible.
-Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with a self-simulation account.
-In future work, we intend to more directly rule out this knowledge-suppression hypothesis, and well as to generalize our results to other games.
-
-In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects' models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
-While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.
-
-# References
-
-```{r}
-# References will be generated automatically by Pandoc and included here.
-# The following code is some latex to format the bibliography. Do not remove it.
-```
-
-```{=tex}
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
-```
-\noindent
+Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-01-18 analysis/decisionEntropy.Rmd
Analysis of decision conformity/stereotipicity (not sure that's a word)

@@
 ---
+
+These scripts can be run after running PTKN.Rmd from docs, because they use objects and functions that are defined there.
+
+Here I compare subjects' tendency to make 'conventional' clicks (that is, clicks that are often made by participants at this stage of the game), in pretend and non-pretend games.
 
@@
 
+# What  I refer to as 'prior' here is the overall tendency of participants to
+# click on some cells. To make sure this quantity is not driven by the content
+# of the board itself, for each board I compute this quantity across all other
+# boards, not including itself.
 E2.prior_from_group <- E2.click_log_with_boards %>% 
-  filter(test_part=='nonpretend')%>%
+  filter(test_part=='nonpretend') %>% # we are interested in true games
   mutate(flat_position = factor(flat_position, levels=seq(25))) %>%
   group_by(grid_number,click_number, flat_position) %>%
-  summarise(n=n())%>%
+  summarise(n=n())%>% # count the number of clicks per cell per board per timepoint
   group_by(click_number, flat_position)%>%
+  ## for each grid, average across all other grids.
   do( data.frame(.,other_n=sapply(.$grid_number, function(i) mean(.$n[!.$grid_number %in% i])))) %>%
   group_by(grid_number, click_number) %>%
-  mutate(p=other_n/sum(other_n)) %>%
+  mutate(p=other_n/sum(other_n)) %>% # normalize to get a probability function
   dplyr::select(grid_number,click_number,flat_position,p) %>%
-  spread(flat_position,p, fill=0, sep='_') %>%
+  spread(flat_position,p, fill=0, sep='_') %>% # pivot to wider format
   ungroup();
@@
 
-  hidden_cells = ifelse(is.na(board_state),1,0);
+  hidden_cells = ifelse(is.na(board_state),1,0); #only hidden cells can be clicked on
 
@@
 };
+
+## the next three functions are exactly the same as the corresponding functions for
+## the posterior, with the variable name posterior replaced with prior. I did 
+## this just to avoid confusion later.
 
@@
 
-# E2.click_log_with_boards <-E2.click_log %>%
-#   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>%
-#   mutate(flat_position = 1+i*5+j) %>%
-#   group_by(subj_id,genuine_first,test_part,grid_number) %>%
-#   arrange(click_number) %>%
-#   summarise(click_number = click_number,
-#             board_state=create_board_states(flat_position,hit_bin),
-#             flat_position=flat_position,
-#             i=i,
-#             j=j,
-#             RT=RT,
-#             hit_bin=hit_bin) %>%
-#   rowwise()%>%
-#   mutate(likelihood = get_likelihood(board_state),
-#          posterior = get_posterior(likelihood),
-#          p_click = get_p_click(posterior,flat_position),
-#          entropy = get_posterior_entropy(posterior),
-#          p_click_rank = get_p_click_rank(posterior,flat_position));
+E2.click_log_with_boards_and_prior <-E2.click_log_with_boards %>%
+  rowwise()%>%
+  mutate(prior = get_group_prior(grid_number, click_number, board_state),
+         prior_click = get_p_click(prior,flat_position),
+         prior_entropy = get_prior_entropy(prior),
+         prior_click_rank = get_prior_click_rank(prior,flat_position));
 
-E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')
+E2.click_log_with_prior_optimal <- E2.click_log_with_boards_optimal %>%
+  mutate(prior = get_group_prior(grid_number, click_number, board_state),
+         prior_click = get_p_click(prior,flat_position),
+         prior_entropy = get_prior_entropy(prior),
+         prior_click_rank = get_prior_click_rank(prior,flat_position),
+         test_part=='optimal');
 
-get_random_flat_position <- function(board_state) {
-
-  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
-  return(which(is.na(board_state))%>%sample(1))
-
-}
-
-get_optimal_flat_position <- function(posterior) {
-
-  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  return(which.max(posterior))
-
-}
-
-E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_optimal_flat_position(posterior),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'optimal')
-
-E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_random_flat_position(board_state),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
+E2.click_log_with_prior_random <- E2.click_log_with_boards_random %>%
+  mutate(prior = get_group_prior(grid_number, click_number, board_state),
+         prior_click = get_p_click(prior,flat_position),
+         prior_entropy = get_prior_entropy(prior),
+         prior_click_rank = get_prior_click_rank(prior,flat_position),
          test_part = 'random')
 
-E2.mean_P_click_rank <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal) %>%
-  rbind(E2.click_log_with_boards_random) %>%
+E2.mean_prior_click_rank <- E2.click_log_with_boards_and_prior %>%
+  dplyr::select(subj_id,test_part,genuine_first,prior_click_rank)%>%
+  rbind(E2.click_log_with_prior_optimal %>% ungroup() %>%
+          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank)) %>%
+  rbind(E2.click_log_with_prior_random %>% ungroup() %>%
+          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank)) %>%
   group_by(subj_id,test_part, genuine_first) %>%
-  summarise(p_click_rank=mean(p_click_rank));
+  summarise(prior_click_rank=mean(prior_click_rank));
 
-E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal) %>%
-  rbind(E2.click_log_with_boards_random) %>%
+E2.mean_prior_click_rank_misses_only <- E2.click_log_with_boards_and_prior %>%
+  dplyr::select(subj_id,test_part,genuine_first,prior_click_rank, hit_bin)%>%
+  rbind(E2.click_log_with_prior_optimal %>% ungroup() %>%
+          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank, hit_bin)) %>%
+  rbind(E2.click_log_with_prior_random %>% ungroup() %>%
+          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank, hit_bin)) %>%
   rowwise() %>%
@@
   group_by(subj_id,test_part, genuine_first) %>%
-  summarise(p_click_rank=mean(p_click_rank))
+  summarise(prior_click_rank=mean(prior_click_rank));
 
-p<- E2.mean_P_click_rank %>%
+p<- E2.mean_prior_click_rank %>%
     filter(test_part != 'optimal') %>%
     mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=p_click_rank,fill=test_part))+
+    ggplot(aes(x=prior_click_rank,fill=test_part))+
     geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
     scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    labs(x='p(click here) - rank', y='number of players')+
     theme(legend.position=c(0.10,0.85)) +
@@
 
-ggsave('../figures/E2_pclick_rank.pdf',p)
-ggsave('../figures/E2_pclick_rank.png',p,width=5,height=2,dpi=300)
+ggsave('../docs/figures/E2_prior_click_rank.pdf',p)
+ggsave('../docs/figures/E2_prior_click_rank.png',p,width=5,height=2,dpi=300)
 
-p<- E2.mean_P_click_rank_misses_only %>%
+p<- E2.mean_prior_click_rank_misses_only %>%
     filter(test_part != 'optimal') %>%
     mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=p_click_rank,fill=test_part))+
+    ggplot(aes(x=prior_click_rank,fill=test_part))+
     geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
@@
 
-E2.num_irrational_clicks <- E2.click_log_with_boards %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(nic=sum(p_click==0)/5);
-
-p<- E2.num_irrational_clicks %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=nic,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='mean number of irrational clicks per game', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    facet_wrap(~genuine_first, nrow=2)+
-    theme_classic();
-
-ggsave('figures/../E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)
-
- E2.RT_p_click_cor <- E2.click_log_with_boards %>%
+ E2.RT_prior_click_cor <- E2.click_log_with_boards_and_prior %>%
      filter(click_number<9 & click_number>1) %>%
      group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(RT,p_click,method='spearman'));
+     summarise(r=cor(RT,prior_click,method='spearman'));
 
- E2.RT_entropy_cor <- E2.click_log_with_boards %>%
+ E2.RT_prior_entropy_cor <- E2.click_log_with_boards_and_prior %>%
      filter(click_number<9 & click_number>1) %>%
      group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(entropy,p_click,method='spearman'))
+     summarise(r=cor(prior_entropy,RT,method='spearman'))
 
 ```
-## Including Plots
+The mean prior rank of non-pretend cell selections was `r E2.mean_prior_click_rank%>%filter(test_part=='nonpretend')%>%pull(prior_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (see Fig.
+2B; `r E2.mean_prior_click_rank%>%filter(test_part=='random')%>%pull(prior_click_rank)%>%mean()%>%printnum()`, `r E2.mean_prior_click_rank%>%spread(test_part,prior_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Pretend games were significantly less stereotypical than non-pretend games (`r E2.mean_prior_click_rank%>%filter(test_part=='pretend')%>%pull(prior_click_rank)%>%mean()%>%printnum()`; `r apa_print(t.test(E2.mean_prior_click_rank%>%filter(genuine_first&test_part=='nonpretend')%>%pull(prior_click_rank), E2.mean_prior_click_rank%>%filter(!genuine_first&test_part=='pretend')%>%pull(prior_click_rank)))$statistic`), but still more optimal than those of a random agent (`r E2.mean_prior_click_rank%>%spread(test_part,prior_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_prior_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(prior_click_rank), E2.mean_prior_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(prior_click_rank)))$statistic`; pretend - random: `r E2.mean_prior_click_rank_misses_only%>%spread(test_part,prior_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
 
-You can also embed plots, for example:
+```{r prior-entropy-RT, echo=FALSE, cache=TRUE}
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+E2.RT_by_prior_entropy <- E2.click_log_with_boards_and_prior %>%
+  filter(click_number>1) %>%
+  mutate(entropy_bin=round(prior_entropy*2)/2) %>%
+  group_by(subj_id,test_part,click_number) %>%
+  mutate(corrected_RT = RT-mean(RT)) %>%
+  group_by(test_part,subj_id,entropy_bin) %>%
+  summarise(RT=median(RT),
+            corrected_RT = median(corrected_RT)) %>%
+  group_by(test_part,entropy_bin) %>%
+  summarise(median_RT=median(RT),
+            sem_RT=bootstrap_error(RT,N_perm),
+            median_corrected_RT=median(corrected_RT),
+            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))
+
+p <- E2.RT_by_prior_entropy %>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Entropy of prior', y='median RT (ms)')+
+  theme_classic()
+
+ggsave('../figures/E2_RT_by_prior_entropy_no_split.png',p,width=5,height=2,dpi=300)
+# ggsave('../figures/E2_RT_by_prior_entropy_no_split.svg',p,width=5,height=2,dpi=300)
+
+p <- E2.RT_by_prior_entropy %>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Entropy of prior', y='median residual RT (ms)')+
+  theme_classic()
+
+ggsave('../figures/E2_corrected_RT_by_prior_entropy_no_split.png',p,width=5,height=2,dpi=300)
+
+E2.RT_by_prior_entropy_coefs <- E2.click_log_with_boards_and_prior %>%
+  filter(click_number>1) %>%
+  group_by(genuine_first,subj_id,test_part,grid_number) %>%
+  mutate(entropy_centered = prior_entropy-mean(prior_entropy),
+         entropy_squared = entropy_centered^2) %>%
+  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
+  group_by(genuine_first,subj_id,test_part,grid_number) %>%
+  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
+  mutate(tidys=list(broom::tidy(model))) %>%
+  unnest(tidys) %>%
+  dplyr::select(!c(std.error,statistic,p.value)) %>%
+  group_by(genuine_first,subj_id,test_part,term) %>%
+  summarise(estimate = mean(estimate))
+
+E2.RT_by_prior_entropy_coefs_summary <- E2.RT_by_prior_entropy_coefs %>%
+  group_by(test_part,term) %>%
+  summarise(mean_estimate = mean(estimate),
+            se_estimate = se(estimate))
+
+p <- E2.RT_by_prior_entropy_coefs_summary %>%
+  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
+  filter(term!='(Intercept)') %>%
+  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_bar(stat='identity',color='black',position=position_dodge()) +
+  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
+  labs(fill=' ',x='term', y='coefficient')+
+  theme_classic()
+
+ggsave('../figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
+
+E2.corrected_RT_by_prior_entropy_coefs <- E2.click_log_with_boards_and_prior %>%
+  filter(click_number>1) %>%
+  group_by(subj_id,test_part,click_number) %>%
+  mutate(corrected_RT = RT-mean(RT)) %>%
+  group_by(genuine_first,subj_id,test_part) %>%
+  mutate(entropy_centered = prior_entropy-mean(prior_entropy),
+         entropy_squared = entropy_centered^2) %>%
+  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
+  group_by(genuine_first,subj_id,test_part) %>%
+  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
+  mutate(tidys=list(broom::tidy(model))) %>%
+  unnest(tidys) %>%
+  dplyr::select(!c(std.error,statistic,p.value))
+
+E2.corrected_RT_by_prior_entropy_coefs_summary <- E2.corrected_RT_by_prior_entropy_coefs %>%
+  group_by(test_part,term) %>%
+  summarise(mean_estimate = mean(estimate),
+            se_estimate = se(estimate))
+
+p <- E2.corrected_RT_by_prior_entropy_coefs_summary %>%
+  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
+  filter(term!='(Intercept)') %>%
+  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_bar(stat='identity',color='black',position=position_dodge()) +
+  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
+  labs(fill=' ',x='term', y='coefficient')+
+  theme_classic()
+
+ggsave('../figures/E2_corrected_RT_by_prior_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
+
 ```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-01-18 analysis/decisionEntropy.Rmd
click entropy analysis

@@
   group_by(click_number, flat_position)%>%
-  ## for each grid, average across all other grids.
-  do( data.frame(.,other_n=sapply(.$grid_number, function(i) mean(.$n[!.$grid_number %in% i])))) %>%
+  ## for each grid, sum across all other grids.
+  do( data.frame(.,other_n=sapply(.$grid_number, function(i) sum(.$n[!.$grid_number %in% i])))) %>%
   group_by(grid_number, click_number) %>%
@@
 ```
+
+```{r entropy-by-erial-position, echo=FALSE, cache=TRUE}
+
+#compute Shannon entropy
+entropy <- function(target) {
+  freq <- table(target)/length(target)
+  # vectorize
+  vec <- as.data.frame(freq)[,2]
+  #drop 0 to avoid NaN resulting from log2
+  vec<-vec[vec>0]
+  #compute entropy
+  -sum(vec * log2(vec))
+}
+
+E2.click_entropy_by_serial_position <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_optimal)%>%
+  rbind(E2.click_log_with_boards_random)%>%
+  # filter(click_number<15) %>%
+  group_by(subj_id,test_part,click_number)%>%
+  summarise(entropy=entropy(flat_position))
+
+E2.click_entropy_by_serial_position_summary <- E2.click_entropy_by_serial_position %>%
+  group_by(test_part,click_number) %>%
+  summarise(mean_entropy=mean(entropy),
+            se_entropy=se(entropy)) %>%
+  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));
+
+p <- E2.click_entropy_by_serial_position_summary %>%
+  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
+  geom_line() +
+  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
+  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
+  labs(x='click number', y='entropy') +
+  theme_classic()
+
+ggsave('../figures/E2_entropy_by_click_nnumber.png',p,width=5,height=5,dpi=300)
+
+```

### 2023-01-18 researchLog.Rmd
click entropy analysis

---
title: "Research log"
author: "Matan Mazor"
date: "2023-01-18"
output: github_document
bibliography: references.bib
---

## Analysis of decision conformity (2023-01-18)

Instead of computing p(click) relative to places where a ship could be hiding, I now computed it relative to the behaviour of other players in the same click number in non-pretend games. My reasoning was that pretenders may be following what people usually do in a rigid and inflexible manner, instead of adapting it flexibly to the game state.

Turns out that's not the case. Pretenders and non-pretenders followed similarly stereotypical click sequences, with slightly less stereotypical sequences in pretend games. The relation between entropy of this new measure and RT was negatively quadratic, but not very strongly so, and only in pretend games.

The entropy of clicks across games seem generally similar in pretend and non-pretend games, but for some reason it decreases at a round the 10th click. I'm not sure why this happens.

### 2023-01-19 analysis/decisionConformityBattleship.Rmd
decision conformity analysis

--- 2023-01-18 analysis/decisionEntropy.Rmd
@@
 ---
-title: "Decision entropy"
+title: "Decision conformity: Battleship"
 author: "Matan Mazor"
@@
 
-#compute Shannon entropy
+#compute Shannon entropy, but only until the first timepoint at which some gamed ended
 entropy <- function(target) {
+  if (length(target)<5) {
+    return(NaN)
+  } else {
     freq <- table(target)/length(target)
@@
     #compute entropy
-  -sum(vec * log2(vec))
+    return(-sum(vec * log2(vec)))
+  }
 }
@@
   group_by(test_part,click_number) %>%
-  summarise(mean_entropy=mean(entropy),
-            se_entropy=se(entropy)) %>%
+  summarise(mean_entropy=mean(entropy, na.rm=T),
+            se_entropy=se(entropy, na.rm=T)) %>%
   mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));
@@
 
-ggsave('../figures/E2_entropy_by_click_nnumber.png',p,width=5,height=5,dpi=300)
+ggsave('../docs/figures/E2_entropy_by_click_nnumber.png',p,width=7,height=5,dpi=300)
 

### 2023-01-19 researchLog.Rmd
decision conformity analysis

@@
 output: github_document
-bibliography: references.bib
 ---
@@
 The entropy of clicks across games seem generally similar in pretend and non-pretend games, but for some reason it decreases at a round the 10th click. I'm not sure why this happens.
+
+## Analysis of decision conformity (2023-01-19)
+
+Part of the solution to the problem is that entropy decreases when there are fewer possible events, and that happens when the click number is high enough such that some games have already ended.
+When restricting the analysis only to cases in which the games haven't ended yet, entropy remains high.
+But for some reason it does decrease for the random and optimal agents:
+
+![](docs/figures/decision_entropy_with_stars.png){width="615"}
+
+It might be subjects that go over the board systematically that decrease this measure.
+
+The important thing is that early on, entropy seems to be higher for nonpretend games.
+
+Oddly, the difference is already apparent in the first click 🤔

### 2023-01-20 analysis/decisionConformityBattleship.Rmd
correlation between entropy effect and pclick cost

@@
   mutate(prior = get_group_prior(grid_number, click_number, board_state),
-         prior_click = get_p_click(prior,flat_position),
+         prior_click = get_prior_click(prior,flat_position),
          prior_entropy = get_prior_entropy(prior),
@@
 
-ggsave('../figures/E2_pclick_rank_misses_only.pdf',p)
-ggsave('../figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
+ggsave('../docs/figures/E2_pclick_rank_misses_only.pdf',p)
+ggsave('../docs/figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
 
@@
 ```
+
+```{r entropy-pclick-correlation, echo=FALSE, cache=TRUE}
+
+E2.entropy_summary <- E2.click_entropy_by_serial_position %>%
+  filter(click_number>8) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(entropy=mean(entropy,na.rm=T))
+
+E2.p_click_summary <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_random) %>%
+  ungroup() %>%
+  filter(click_number<8) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click=mean(p_click))
+
+E2.p_click_summary %>%
+  merge(E2.entropy_summary) %>%
+  ggplot(aes(x=entropy,y=p_click, color=test_part))+
+  geom_point(size=2,alpha=0.8)
+
+```

### 2023-01-24 docs/Methods and results.Rmd
putting everything in one document

--- 2023-01-18 docs/cogsci/PNTKN.Rmd
@@
 ---
-title: "Using pretense behavior to explore counterfactual self-simulation"
-bibliography: r-references.bib
-csl: apa6.csl
-document-params: "10pt, letterpaper"
+title             : "Pretending not to know"
+shorttitle        : "Pretending not to know"
 
-author-information: > 
-    \author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of London \\ London, WC1E 7HX UK
-    \AND {\large \bf Chaz Firestone (chaz@jhu.edu)}
-    \AND {\large \bf Ian Phillips (ianphillips@jhu.edu)} \\ Johns Hopkins University \\ Baltimore, MD 21218, USA}
+author: 
+  - name          : "Matan Mazor"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Malet St., London WC1E 7HX"
+    email         : "mtnmzor@gmail.com"
+  - name          : "Chaz Firestone"
+    affiliation   : "2"
+  - name          : "Ian Phillips"
+    affiliation   : "2"
 
-abstract: >
-   What we do depends on what we know.
-But sometimes we try to decouple our behavior from our knowledge so that we appear not to know what we really do know.
-Such pretense behavior requires understanding how we would behave with different knowledge — and so provides a window into counterfactual self-simulation.
-However, little research has characterized and evaluated pretense relative to non-pretense behavior.
-Here, in a large-scale, pre-registered experiment, subjects played normal games of Battleships (trying to sink ships hidden in a grid), as well as ‘pretend’ games, where they were told all the ships’ locations but had to pretend they were playing normally.
-Relative to normal games, ‘pretend’ games demonstrated similar but exaggerated behavioral patterns.
-Furthermore, pretenders played rationally, but less so than non-pretenders.
-Despite these differences, ‘judge’ participants could not detect ‘pretend’ games.
-We discuss implications of these findings for accounts of theory of mind and metacognition.
+affiliation:
+  - id            : "1"
+    institution   : "Birkbeck, University of London"
+  - id            : "2"
+    institution   : "Johns Hopkins University"
 
-keywords: >
-    metacognition; theory of mind; pretense
+authornote: |
+  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
 
-output: cogsci2016::cogsci_paper
+  Enter author note here.
 
-final-submission: \cogscifinalcopy
+abstract: |
+  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
+
+  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
+
+  One sentence clearly stating the **general problem** being addressed by  this particular study.
+
+  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
+
+  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
+
+  One or two sentences to put the results into a more **general context**.
+
+  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
+
+  <!-- https://tinyurl.com/ybremelq -->
+
+keywords          : "pretense, theory of mind, metacognition, self-model, self-simulation"
+wordcount         : "X"
+
+bibliography      : "r-references.bib"
+
+floatsintext      : no
+linenumbers       : yes
+draft             : no
+mask              : no
+
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+
+classoption       : "man"
+output            : papaja::apa6_word
 ---
 
+```{r setup, include = FALSE}
+library("papaja")
+r_refs("r-references.bib")
+```
+
 ```{r global_options, include=FALSE}
+set.seed(42)
 knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
@@
 
-# Introduction
+# Exp. 1: Battleship
 
-The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory].
-This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation].
-This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own.
-
-Moreover, deception often involves pretense behavior, which in turn relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state.
-For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
-What would I say?
-How long would it take me to recover from the surprise?").
-This reliance of pretense behavior on self-simulation makes it an ideal opportunity to examine metacognitive knowledge about one's own mental states, and the potential reliance of this knowledge on a self-simulation.
-By comparing non-pretend and pretend behavior, we can ask which aspects of their own cognitive processes subjects can and cannot effectively simulate, and which aspects are not well-represented in their mental models of their own cognition.
-
-To this end, here we examine pretense in a game setting.
-Using an online version of the game Battleships, participants played a 'non-pretend' (normal) version of the game (trying to sink ships whose locations are unknown), as well as a 'pretend' version where they were given secret information about the location of hidden ships, but were instructed to behave as if they didn't have this information.
-
-# Methods
+## Methods
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
 
@@
 
-```{r load-data, warning=FALSE}
-E2.df <- read.csv('../../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
+```{r battleship-load-data, warning=FALSE}
+
+E2.df <- read.csv('../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
   mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,4),sep='')) %>%
@@
 
-E2.export <- read.csv('../../experiments/Battleships2/data/batch1/prolific_export_batch1.csv',na.strings=c(""," ","NA"));
+E2.export <- read.csv('../experiments/Battleships2/data/batch1/prolific_export_batch1.csv',na.strings=c(""," ","NA"));
 
-E2.bonus <- read.csv('../../experiments/Battleships2/data/batch1/bonus.csv',na.strings=c(""," ","NA"))
+E2.bonus <- read.csv('../experiments/Battleships2/data/batch1/bonus.csv',na.strings=c(""," ","NA"))
 
@@
 
-## Participants
+### Participants
 
@@
 
-## Procedure
+### Procedure
 
@@
 
-### Non-pretend (normal) games
+#### Non-pretend (normal) games
 
@@
 
-### Pretend games
+#### Pretend games
 
@@
 
-### Half games
+```{r battleship-design, echo=FALSE, fig.cap="Experimental Design. See Methods for details.", out.width = '100%'}
+knitr::include_graphics("figures/Battleships_design.png")
+```
+
+#### Half games
 
@@
 
-### Judge trials
+#### Judge trials
 
@@
 
-<!-- 500 players played 12 games of the game Battleships in two conditions presented in two blocks.
-The order of conditions was randomized between players.
-the standard, *non-pretend* condition, players started the game not knowing where the ships were and tried to sink all ships with as few clicks as possible.
-In the *pretend* condition, ship locations were marked with a cross such that players knew where they were hidden, and their task was to play the game as if they did not have this information.
-The sixth game in each block was a 'half-game', where participants completed the game with the content of some cells revealed by a previous player.
-Lastly, in judge trials, players observed a replay of the games of two previous players – one pretender and one non-pretender, and their task was to identify which was which.
-We created nine different 5x5 boards, each containing two 2-square patrol boats and one 3-square submarine.
-The assignment of boards to conditions was randomized across participants such that exactly one board was presented in both pretend and non-pretend blocks.
-(see Fig.
-1).
--->
-
-A more detailed description of the study procedure is provided in the study pre-registration document (LINK HIDDEN).
+A more detailed description of the study procedure is provided in the study [pre-registration document](https://osf.io/v9zsb).
 Readers are also invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/NervzpM0Y0z).
 
-```{r design, echo=FALSE, fig.cap="Experimental Design. See Methods for details.", out.width = '100%'}
-knitr::include_graphics("../figures/Battleships_design.png")
-```
+### Data analysis
+<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->
 
-# Results
+## Results
 
-```{r subject-comments, echo=FALSE}
+```{r battleship-subject-comments, echo=FALSE}
+
 E2.comments <- E2.df %>%
@@
 In our pre-registration document, we committed to separately analyzing participants according to whether they pretended before or after completing a non-pretend block.
-Due to space limitations, and since the results of the two groups mostly agreed, we report here the pooled results from both groups of participants, and mention whenever we found different results depending on block order.
-When directly comparing pretend and non-pretend blocks, we perform a between-subjects comparison using data from the first block only, i.e., pretend games from participants who pretended in the first block and non-pretend games from participants who played normally in the first block.
+Whenever there is a mismatch between the two groups, we focus on the first block of each participant (only pretend games from participants who pretended first, and only standard games from participants who played normally first).
 We do so to ensure that any successful pretending is not due to memory of one's own behavior in a previous block, and that non-pretend games are not biased by experience with the pretend block.
 
-```{r num_clicks, echo=FALSE, cache=TRUE}
+```{r battleship-num_clicks, echo=FALSE, cache=TRUE}
 
@@
 
-```{r useful_functions, echo=FALSE, cache=TRUE}
+```{r battleship-useful_functions, echo=FALSE, cache=TRUE}
 
@@
 
-## Total number of clicks
+*Hypothesis 1 (game duration)*: We will test the null hypothesis that pretend games take as long as non-pretend games.
+For each participant, a difference between median game duration in pretend and non-pretend games will be extracted.
+Subject-wise differences will then be carried over to a t-test at the group level.
 
-To sink all ships, players had to click on at least 7 and at most 25 squares.
-A simulated player that clicks randomly had a median total click number of `r num_clicks_sim%>%median()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a median total click number of `r num_clicks_optimal_sim%>%median()`.
-Among our players, the median number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games (see Fig.
-2A).
+*Hypothesis 2 (first click latency)*: We will test the null hypothesis that the first click in pretend games takes as long to execute as in non-pretend games.
+For each participant, a difference between median first click latency in pretend and non-pretend games will be extracted.
+Subject-wise differences will then be carried over to a t-test at the group level.
 
-We observed no significant difference in the number of clicks between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend),E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
-However, a significant interaction with block order suggested that participants made fewer clicks in the second block compared to the first, regardless of which block was the pretend block (`r E2.median_clicks%>%gather('condition','nclicks',c('nonpretend','pretend'))%>%aov(nclicks ~ genuine_first * condition,data=.)%>%apa_print()%>%'$'(statistic)%>%'$'(genuine_first_condition)`).
+*Hypothesis 3 (number of clicks)*: We will test the null hypothesis that pretend games are as long, in terms of total number of clicks, as non-pretend games.
+For each participant, a difference between median number of clicks in pretend and non-pretend games will be extracted.
+Subject-wise differences will then be carried over to a t-test at the group level.
 
-In `r E2.click_df%>%filter(num_clicks==7)%>%nrow()` pretend games from `r E2.lucky_pretenders%>%length()` players, games were completed after 7 clicks only, without ever missing a ship.
-This never happened in non-pretend games.
-We assumed that these participants did not follow task instructions, and excluded them from all following analyses.
+*Hypotheses 4-6 (click latency by click outcome)*: Trials will be classified based on whether they resulted in hitting a ship (*hit*) or not (*miss*).
+We will then contrast median click latency between hits and misses separately for pretend and non-pretend games.
+We will test for a difference in trial latency as a function of outcome on pretend games, on non-pretend games, and for an interaction effect between click outcome (hit or miss) and condition (pretend or non-pretend) on click latency.
+The same analysis will then be repeated by classifying trials on the basis of whether the previous trial resulted in a hit or a miss (*Hypothesis 5*), and whether the next trial will result in a hit or a miss (*Hypothesis 6*).
 
-## Game optimality
+# Exp. 2: Hangman
 
-Pretend games were similar to non pretend games in the total number of clicks --- but were they also similar in *where* participants clicked?
-More specifically, did cell selections in pretend games make sense given the limited information those participants pretended to have?
-To ask this, we approximated optimal behavior by calculating the probability that a ship is hidden in each cell given available information, $p(ship(x_{i}))$, and the posterior probability that one should click on a square, assuming a uniform prior over cells $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$.
-Critically, in modeling pretend games we did not treat hints as part of this available information for extracting $p(ship(x_{i}))$, because an optimal player should ignore hints in choosing where to click next.
-Given this posterior map, a rational player should choose cells where $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3].
-To quantify optimality, before each cell selection we computed the posterior probability map for all 'unknown' cells.
-Then, we ranked cells from high to low according to their posterior probability and recorded the rank of the chosen cell: a lower rank indicating more optimal behavior.
+## Methods
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
 
-```{r parse-click-log, echo-FALSE, cache=TRUE}
+### Participants
 
-E2.click_log <- readRDS('../../experiments/Battleships2/data/batch1/click_log.Rda');
+### Material
 
-# E2.click_log <- data.frame(matrix(ncol=10,nrow=0, 
-#                                dimnames=list(NULL, 
-#                                              c("subj_id",
-#                                               "test_part", 
-#                                               "grid_number", 
-#                                               "num_clicks",
-#                                               "i",
-#                                               "j",
-#                                               "hit",
-#                                               "t",
-#                                               "click_number",
-#                                               "genuine_first"))))
-# 
-# 
-# for (row in 1:nrow(E2.click_df)) {
-#   
-#     subject_click_log <- data.frame(fromJSON(E2.click_df[row, ]$click_log)) %>%
-#     mutate(
-#       click_number = 1:n(),
-#       subj_id = E2.click_df[row, ]$subj_id,
-#       test_part = E2.click_df[row, ]$test_part,
-#       grid_number = E2.click_df[row, ]$grid_number,
-#       num_clicks = E2.click_df[row, ]$num_clicks,
-#       genuine_first = E2.click_df[row, ]$genuine_first
-#     )
-#     
-#     E2.click_log <- rbind(E2.click_log, subject_click_log);
-# }
-#     
-# 
-# E2.click_log <- E2.click_log %>%
-#   relocate(subj_id, .before = i) %>%
-#   relocate(genuine_first, .before = i) %>% 
-#   relocate(test_part, .before=i) %>% 
-#   relocate(grid_number, .before=i) %>% 
-#   relocate(click_number, .before=i) %>%
-#   group_by(subj_id,test_part,grid_number) %>%
-#   mutate(RT=t-lag(t,default=0)) %>%
-#   filter(!(subj_id %in% E2.lucky_pretenders))%>%
-#  mutate(hit_bin = hit !=0,
-#         lag3 = lag(hit,3)!=0,
-#         lag2 = lag(hit,2)!=0,
-#         lag1 = lag(hit,1)!=0,
-#         lead1 = lead(hit,1)!=0,
-#         lead2 = lead(hit,2)!=0,
-#         lead3 = lead(hit,3)!=0)
+### Procedure
 
-```
+### Data analysis
+We used `r cite_r("r-references.bib")` for all our analyses.
 
-```{r simualte_data, echo=FALSE, cache=TRUE}
-
-simulate_optimal_game <- function(ship_positions) {
-  #initialize
-  cur_board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-  board_state = c()
-  flat_position = c();
-  click_number = c();
-  hit_bin = c();
-
-  while (nships>nhits) {
-
-    board_state = c(board_state,paste(cur_board_state,collapse=','));
-
-    posterior <- get_likelihood(paste(cur_board_state,collapse=',')) %>%
-      get_posterior() %>%
-      scan(text = ., what= numeric(), sep="," , quiet = TRUE);
-
-    best_shot = which.max(posterior);
-    flat_position = c(flat_position,best_shot);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    click_number = c(click_number,nclicks)
-
-    result = ship_positions[best_shot]
-
-    hit_bin = c(hit_bin,ifelse(result==1,1,0));
-
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    cur_board_state[best_shot]=result;
-
-  }
-  return(data.frame(board_state,flat_position,hit_bin,click_number))
-
-}
-
-simulate_random_game <- function(ship_positions) {
-  #initialize
-  cur_board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-  board_state = c()
-  flat_position = c();
-  click_number = c();
-  hit_bin = c();
-
-  while (nships>nhits) {
-
-    board_state = c(board_state,paste(cur_board_state,collapse=','));
-
-    random_guess = which(is.na(cur_board_state))%>%sample(1);
-    flat_position = c(flat_position,random_guess);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    click_number = c(click_number,nclicks)
-
-    result = ship_positions[random_guess]
-
-    hit_bin = c(hit_bin,ifelse(result==1,1,0));
-
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    cur_board_state[random_guess]=result;
-
-  }
-  return(data.frame(board_state,flat_position,hit_bin,click_number))
-
-}
-
-nsim = 100;
-nboards = all_possible_boards%>%nrow();
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  if (i==1) {
-    simulated_optimal = simulate_optimal_game(cur_board)%>%mutate(game=i)
-  } else {
-    simulated_optimal = bind_rows(simulated_optimal, simulate_optimal_game(cur_board)%>%mutate(game=i))
-  }
-}
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  if (i==1) {
-    simulated_random = simulate_random_game(cur_board)%>%mutate(game=i)
-  } else {
-    simulated_random = bind_rows(simulated_random, simulate_random_game(cur_board)%>%mutate(game=i))
-  }
-}
-
-```
-
-```{r p_click, echo = FALSE, cache=TRUE}
-
-# a <-E2.click_log %>%
-#   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>%
-#   mutate(flat_position = 1+i*5+j) %>%
-#   group_by(subj_id,genuine_first,test_part,grid_number) %>%
-#   arrange(click_number) %>%
-#   summarise(click_number = click_number,
-#             board_state=create_board_states(flat_position,hit_bin),
-#             flat_position=flat_position,
-#             i=i,
-#             j=j,
-#             RT=RT,
-#             hit_bin=hit_bin) %>%
-#   rowwise()%>%
-#   mutate(likelihood = get_likelihood(board_state),
-#          posterior = get_posterior(likelihood),
-#          p_click = get_p_click(posterior,flat_position),
-#          entropy = get_posterior_entropy(posterior),
-#          p_click_rank = get_p_click_rank(posterior,flat_position));
-
-E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')
-
-get_random_flat_position <- function(board_state) {
-
-  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
-  return(which(is.na(board_state))%>%sample(1))
-
-}
-
-get_optimal_flat_position <- function(posterior) {
-
-  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  return(which.max(posterior))
-
-}
-
-E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_optimal_flat_position(posterior),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'optimal')
-
-E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_random_flat_position(board_state),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'random')
-
-E2.mean_P_click_rank <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal) %>%
-  rbind(E2.click_log_with_boards_random) %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(p_click_rank=mean(p_click_rank));
-
-E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal) %>%
-  rbind(E2.click_log_with_boards_random) %>%
-  rowwise() %>%
-  filter(!hit_bin) %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(p_click_rank=mean(p_click_rank))
-
-p<- E2.mean_P_click_rank %>%
-    filter(test_part != 'optimal') %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=p_click_rank,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    # facet_wrap(~genuine_first, nrow=2)+
-    # scale_x_reverse() +
-    theme_classic();
-
-ggsave('../figures/E2_pclick_rank.pdf',p)
-ggsave('../figures/E2_pclick_rank.png',p,width=5,height=2,dpi=300)
-
-p<- E2.mean_P_click_rank_misses_only %>%
-    filter(test_part != 'optimal') %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=p_click_rank,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
-    scale_x_continuous(limits=c(0,15))+
-    theme(legend.position=c(0.10,0.85)) +
-    # facet_wrap(~genuine_first, nrow=2)+
-    # scale_x_reverse() +
-    theme_classic();
-
-ggsave('../figures/E2_pclick_rank_misses_only.pdf',p)
-ggsave('../figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
-
-E2.num_irrational_clicks <- E2.click_log_with_boards %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(nic=sum(p_click==0)/5);
-
-p<- E2.num_irrational_clicks %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=nic,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='mean number of irrational clicks per game', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    facet_wrap(~genuine_first, nrow=2)+
-    theme_classic();
-
-ggsave('figures/../E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)
-
- E2.RT_p_click_cor <- E2.click_log_with_boards %>%
-     filter(click_number<9 & click_number>1) %>%
-     group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(RT,p_click,method='spearman'));
-
- E2.RT_entropy_cor <- E2.click_log_with_boards %>%
-     filter(click_number<9 & click_number>1) %>%
-     group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(entropy,p_click,method='spearman'))
-
-```
-
-The mean posterior rank of non-pretend cell selections was `r E2.mean_P_click_rank%>%filter(test_part=='nonpretend')%>%pull(p_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (see Fig.
-2B; `r E2.mean_P_click_rank%>%filter(test_part=='random')%>%pull(p_click_rank)%>%mean()%>%printnum()`, `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Pretend games were significantly less optimal than non-pretend games (`r E2.mean_P_click_rank%>%filter(test_part=='pretend')%>%pull(p_click_rank)%>%mean()%>%printnum()`; `r apa_print(t.test(E2.mean_P_click_rank%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`), but still more optimal than those of a random agent (`r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-In other words, the optimality of pretend games relative to random cell selection was not merely due to the fact that pretenders clicked on ships more than expected by chance.
-Even when missing a ship, their cell selections made sense given the limited information they pretended to have.
-
-```{r numclick_optimality, echo=FALSE, fig.cap="Experimental Design", out.width = '100%', fig.cap = "A: The total number of clicks in pretend and non-pretend games was highly similar.
-B: Non-pretend games were significantly more optimal than pretend games, but both were more optimal than what is expected by random chance.
-Box: An illustration of the difference between success and optimality.
-A given click (highlighted with a black border) can reveal a ship (upper row) or not (lower row) regardless of whether the click makes sense given available information (left column) or not (right column).
-"}
-knitr::include_graphics("../figures/numclicks_optimality.png")
-```
-
-## Response time analysis
-
-```{r total_time, echo=FALSE}
-
-E2.total_time_df <- E2.click_log %>%
- filter(click_number==num_clicks)%>%
-  mutate(t=t/1000)
-
-E2.median_time <- E2.total_time_df %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(t=median(t)) %>%
-  spread(test_part,t);
-
-# anova
-E2.time_anva <- aov(t ~ genuine_first * condition, 
-                         data = E2.median_time %>%
-                           gather('condition','t',nonpretend,pretend) %>%
-                           mutate(condition=as.factor(condition)));
-
-E2.median_time_per_click <- E2.click_log %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(RT=median(RT)) %>%
-  spread(test_part,RT) %>%
-  mutate(diff=pretend-nonpretend);
-
-# anova
-E2.click_time_anva <- aov(RT ~ genuine_first * condition, 
-                         data = E2.median_time_per_click %>%
-                           gather('condition','RT',nonpretend,pretend) %>%
-                           mutate(condition=as.factor(condition)))
-```
-
-In general, participants gave slower responses when pretending compared to when playing normally.
-Despite the similar number of clicks per game, the median game duration was `r E2.total_time_df%>%filter(test_part=='nonpretend')%>%pull(t)%>%median()%>%printnum()` seconds in the non-pretend condition and `r E2.total_time_df%>%filter(test_part=='pretend')%>%pull(t)%>%median()%>%printnum()` seconds in the pretend condition, which was significantly longer (`r t.test(E2.median_time$pretend-E2.median_time$nonpretend)%>%apa_print()%>%'$'(statistic)`).
-This difference showed a significant interaction with the order of conditions, such that completing games in the first block took longer (`r apa_print(E2.time_anva%>%summary())$full_result$genuine_first_condition`).
-
-In our instructions to pretenders, we asked them to pretend not to know where the ships were, not only in *where* they chose to click, but also in *when* they clicked.
-In the following set of analyses, we identified patterns in click latency in non-pretend games, and asked whether the same patterns are also observed in pretend games.
-We focused on the effects of hitting versus missing a ship, decision uncertainty, and ship completion.
-
-```{r RT_by_accuracy}
-
-N_perm <- 1000;
-bootstrap_error <- function(x, N_perm) {
-  N <- length(x)
-  medians = c();
-  for (i in 1:N_perm) {
-    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
-  };
-  return(sd(medians))
-}
-
-E2.RT_by_lag3 <- E2.click_log %>%
-  filter(click_number>3) %>%
-  group_by(subj_id,test_part,lag3,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag3,RT,sep='_') %>%
-  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);
-
-E2.RT_by_lag2 <- E2.click_log %>%
-  filter(click_number>2) %>%
-  group_by(subj_id,test_part,lag2,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag2,RT,sep='_') %>%
-  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);
-
-E2.RT_by_lag1 <- E2.click_log %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,lag1,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag1,RT,sep='_') %>%
-  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
-
-E2.RT_by_hit <- E2.click_log %>%
-  filter(click_number>0) %>%
-  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(hit_bin,RT,sep='_') %>%
-  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
-
-E2.RT_by_lead1 <- E2.click_log %>%
-  filter(click_number<num_clicks) %>%
-  group_by(subj_id,test_part,lead1,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead1,RT,sep='_') %>%
-  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
-
-E2.RT_by_lead2 <- E2.click_log %>%
-  filter(click_number<num_clicks-1) %>%
-  group_by(subj_id,test_part,lead2,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead2,RT,sep='_') %>%
-  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)
-
-E2.RT_by_lead3 <- E2.click_log %>%
-  filter(click_number<num_clicks-2) %>%
-  group_by(subj_id,test_part,lead3,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead3,RT,sep='_') %>%
-  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)
-
-E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
-  merge(E2.RT_by_lag2) %>%
-  merge(E2.RT_by_lag1) %>%
-  merge(E2.RT_by_hit) %>%
-  merge(E2.RT_by_lead1) %>%
-  merge(E2.RT_by_lead2) %>%
-  merge(E2.RT_by_lead3);
-
-E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
-  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
-  mutate(position = ifelse(position=="lag3_diff", 3,
-                           ifelse(position=="lag2_diff",2,
-                                  ifelse(position=="lag1_diff",1,
-                                         ifelse(position=="hit_diff",0,
-                                                ifelse(position=="lead1_diff",-1,
-                                                       ifelse(position=="lead2_diff",-2,-3
-                                                              ))))))) %>%
-  dplyr::select(subj_id,test_part,genuine_first,position,difference);
-
-p<- E2.RT_by_hit_lags_long %>% 
-  filter(position<2 & position > -1)%>%
-  group_by(test_part,position) %>%
-  summarise(median_difference=median(difference),
-            sem_difference=bootstrap_error(difference,N_perm)) %>%
-  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  scale_x_continuous(breaks=0:1,labels=c('hit on current trial', 'hit on previous trial'))+
-  geom_abline(slope=0,intercept=0,linetype=3)+
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.pdf',p,width=7)
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.png',p,width=4, height=3.5, dpi=300)
-
-```
-
-### Effects of hitting versus missing a ship
-
-In non-pretend games, players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='nonpretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) when hitting compared to missing a ship (see Fig.
-3A; `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-The same effect was observed in pretend games: players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='pretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) in hits compared to misses (`r E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-However, this difference between hits and misses was significantly more pronounced in pretend games (`r apa_print(t.test(E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff), E2.RT_by_hit_lags%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)))$statistic`).
-
-The opposite effect was observed for a hit on the previous click: players were slower by `r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) after hitting a ship (`r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-This effect was also observed in pretend games, where players were slower by `r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='pretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) following a hit on the previous click (`r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Again, the absolute effect in ms was exaggerated in the pretend condition (`r t.test(E2.RT_by_lag1%>%filter(!genuine_first & test_part=='pretend')%>%pull(lag1_diff),E2.RT_by_lag1%>%filter(genuine_first & test_part=='nonpretend')%>%pull(lag1_diff))%>%apa_print()%>%'$'(statistic)`).
-
-### Effects of decision uncertainty
-
-When playing Battleships, it is sometimes clear what the next cell selection should be, and sometimes more difficult to decide where to click next.
-To capture this notion of decision uncertainty, we calculated the entropy of the posterior distribution over cell selections $H(P)=-\sum_{i=1}^{k}P(x_i) \log P(x_{i})$, where $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$, and asked how this measure relates to decision latency, or the time taken to click on the next cell.
-$H(P)$ is high when players need to decide between multiple cells with a similar probability of hiding a ship, and low when there are only a few candidates with a high probability of hiding a ship.
-For every player and condition separately, we fitted a multiple linear regression to predict decision latency based on $H(P)$ and $H(P)^2$.
-The resulting coefficients were then subjected to a group-level inference.
-The first cell selection of each game was excluded from this analysis, because entropy was constant for the first click.
-
-```{r entropy-RT, echo=FALSE, cache=TRUE}
-
-E2.RT_by_entropy <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  mutate(entropy_bin=round(entropy*2)/2) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(test_part,subj_id,entropy_bin) %>%
-  summarise(RT=median(RT),
-            corrected_RT = median(corrected_RT)) %>%
-  group_by(test_part,entropy_bin) %>%
-  summarise(median_RT=median(RT),
-            sem_RT=bootstrap_error(RT,N_perm),
-            median_corrected_RT=median(corrected_RT),
-            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))
-
-p <- E2.RT_by_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
-# ggsave('../figures/E2_RT_by_entropy_no_split.svg',p,width=5,height=2,dpi=300)
-
-p <- E2.RT_by_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of posterior', y='median residual RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.RT_by_entropy_coefs <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  mutate(entropy_centered = entropy-mean(entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value)) %>%
-  group_by(genuine_first,subj_id,test_part,term) %>%
-  summarise(estimate = mean(estimate))
-
-E2.RT_by_entropy_coefs_summary <- E2.RT_by_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.RT_by_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.corrected_RT_by_entropy_coefs <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  mutate(entropy_centered = entropy-mean(entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value))
-
-E2.corrected_RT_by_entropy_coefs_summary <- E2.corrected_RT_by_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.corrected_RT_by_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-```
-
-In non-pretend games, we found no evidence for a linear relation between decision entropy and decision latency (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).
-We did however observe a block order effect, with a significantly negative linear modulation only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
-In pretend games this negative relation between $H(P)$ and click latency was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).
-
-A negative linear modulation of $H(P)$ on decision latency is surprising, as it suggests that participants were quicker to decide when uncertainty was high.
-However, visual inspection of the RT/entropy curves (Fig.
-3B) reveals a pronounced quadratic modulation, such that in both pretend and non-pretend games slower decision times are observed for intermediate $H(P)$ values of around 2.
-Similar to the linear effect of $H(P)$ on decision latency, here too a quadratic modulation in non-pretend games was observed only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
-Again, in pretend games this negative quadratic relation was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared')%>%pull(estimate)%>%t.test())$statistic`).
-A between-subjects comparison focusing on games from the first block revealed that this negative quadratic effect was stronger in pretend, compared to non-pretend games (`r apa_print(t.test(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate),E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)))$statistic`).
-
-```{r latency, echo=FALSE, out.width = '100%', fig.cap = "A: A subtraction of decision latencies in misses versus hits in the current and previous cell selections, in pretend and non-pretend games.
-B.
-Upper panel: Effect of the entropy of the posterior distribution over cell selections (a measure of decision uncertainty) on decision latency in pretend and non-pretend games.
-Lower panel: mean beta coefficients for a multiple regression model, predicting decision latency from entropy and entropy squared.
-C: Decision latency after a hit, as a function of hit number."}
-knitr::include_graphics("../figures/decision_latencies.png")
-```
-
-<!-- $H(P)$ is highest before the first click in a game and lowest just before the last one.
-To make sure any results are not driven by a general effect of click number on decision latency, and in deviation from our pre-registered analysis plan, we first mean-centered all decision latencies within individual subjects, conditions, and click numbers.
-This way, any effect of $H(P)$ on latency cannot reflect effects of click order.
--->
-
-## Effects of ship completion
-
-```{r}
-
-get_hit_number <- function(hit) {
-
-  num_A <- 0;
-  num_B <- 0;
-  num_C <- 0;
-
-  hit_number <- c();
-
-  for (i in seq(length(hit))) {
-    if (hit[i]==0) {
-      hit_number = c(hit_number,NaN)
-    } else if (hit[i]=='A') {
-      num_A = num_A + 1;
-      hit_number = c(hit_number,num_A)
-    } else if (hit[i]=='B') {
-      num_B = num_B + 1;
-      hit_number = c(hit_number,num_B)
-    } else if (hit[i]=='C') {
-      num_C = num_C + 1;
-      hit_number = c(hit_number,num_C)
-    }
-  }
-
-  return(hit_number)
-}
-
-revealed_three <- function(hit_number) {
-
-  revealed_three <- c();
-  three = FALSE;
-
-  for (i in seq(length(hit_number))) {
-    revealed_three = c(revealed_three, three);
-     if (!(is.na(hit_number[i])) & hit_number[i]==3) {
-      three = TRUE
-    } 
-  }
-
-  return(revealed_three)
-}
-
-E2.hit_order <- E2.click_log %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  summarise(click_number = click_number,
-            hit = hit,
-            RT = RT,
-            hit_number = get_hit_number(hit),
-            revealed_three = revealed_three(hit_number),
-            hit_type = ifelse(hit_number==1, 'first',
-                              ifelse(hit_number==2 & !revealed_three, 'second',
-                                     ifelse(hit_number==2 & revealed_three, 'secondlast',
-                                            ifelse(hit_number==3,'thirdlast','sea'))))) %>%
-  mutate(prev_hit_type = lag(hit_type),
-         prev_hit = lag(hit))
-
-E2.RT_by_hit_order <- E2.hit_order %>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(RT=median(RT))
-
-p <- E2.RT_by_hit_order %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(RT = median(RT),
-            sem_RT = bootstrap_error(RT,N_perm)) %>%
-  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_previous_hit_number_collapse.png',p,width=5,height=2,dpi=300)
-
-##correct for trial order
-
-E2.corrected_RT_by_hit_order <- E2.hit_order %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT))%>%
-  rowwise() %>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(corrected_RT=median(corrected_RT))
-
-p <- E2.corrected_RT_by_hit_order %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
-            corrected_RT = median(corrected_RT)) %>%
-  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)
-
-## correct for true ship size
-
-E2.RT_by_hit_order_boats_only <- E2.hit_order %>%
-  filter(prev_hit!='A')%>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(RT=median(RT))
-
-p <- E2.RT_by_hit_order_boats_only %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(RT = median(RT),
-            sem_RT = bootstrap_error(RT,N_perm)) %>%
-  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)
-
-E2.corrected_RT_by_hit_order_boats_only <- E2.hit_order %>%
-  filter(prev_hit!='A')%>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT))%>%
-  rowwise() %>%
-  group_by(genuine_first,subj_id,test_part,prev_hit_type) %>%
-  summarise(corrected_RT=median(corrected_RT))
-
-p <- E2.corrected_RT_by_hit_order_boats_only %>%
-  filter(!is.na(prev_hit_type))%>%
-  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
-  group_by(test_part,prev_hit_type) %>%
-  summarise(sem_corrected_RT = se(corrected_RT,N_perm),
-            corrected_RT = median(corrected_RT)) %>%
-  ggplot(aes(x=prev_hit_type,y=corrected_RT,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=corrected_RT-sem_corrected_RT,ymax=corrected_RT+sem_corrected_RT),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='previous hit', y='median RT, corrected for serial order')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed_boats_only.png',p,width=5,height=2,dpi=300)
-
-```
-
-The previous analyses reveal robust associations between game state and decision latency, that are highly similar, and even exaggerated, in pretend compared to non-pretend games.
-To get at subtler dynamics in a more direct way, in a follow-up exploratory analysis we focused on decision latencies following a hit.
-We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn't been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine (see Fig.
-3C).
-In the first case, players know the ship must continue in one of the neighboring cells.
-In the second case, there is a good chance the ship continues (if this ship turns out to be a submarine).
-In the third and fourth cases, it is clear that the ship is fully sunk.
-
-In non-pretend games, participants were significantly slower to select the next cell when they knew they had just completed a ship (categories 3 and 4) compared to when they just hit a ship, but were not sure (category 2) or knew they had not completely sunk it (category 1; `r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-Specifically, we find that participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-
-Strikingly, we found the exact same pattern in pretend games.
-Players were faster to make the next cell selection when they pretended to think that the current ship might not be fully sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).
-This was not merely a difference between the first, second and third hits: participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-Further analysis confirmed that this effect remained significant when controlling for click number (`r apa_print(E2.corrected_RT_by_hit_order%>%spread(prev_hit_type,corrected_RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`), and when restricting the analysis to the second hit of a ship that is in fact of size-two (`r apa_print(E2.RT_by_hit_order_boats_only%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).
-
-This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their their response latency in the same way it would have been affected had they been in a non-pretend game.
-
-## Half-games
-
-Our optimality analysis showed that pretenders' click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be.
-However, due to the high number of possible board configurations, data from full games provide limited opportunity to compare cell selections for specific game states.
-In addition to asking, "What guides cell selections in pretend and non-pretend games?", we also wanted to ask, "Where exactly would pretenders and non-pretenders click, given a specific board configuration?".
-
-To achieve this, the sixth game in each block started not with an empty grid, but with the contents of some cells already revealed by a previous player.
-As before, pretenders also knew where the remaining ships were hidden, but tried to play as if they only knew what was known to this previous player.
-Having cell selections from 250 players for each board configuration and condition allowed us to plot and compare the distribution of clicks under a genuine, or pretend, knowledge state.
-
-```{r, child=c('../half_games.Rmd')}
-```
-
-For example, look at Board A in Fig.
-4 (upper left corner).
-Imagine that you need to select your next move, knowing that exactly two size-two patrol boats and one size-three submarine are hiding in the water.
-Where would you click next?
-Perhaps one of the two revealed ships is the size-three submarine.
-In that case you want to click on the first cell in the third row, or on the third cell in the fifth row.
-Or maybe these are the two size-two patrol boats.
-This would mean the submarine is hiding in some other place, and for it not to touch sides with the boats this must be in the first row.
-The hit probability map in the second column visualizes this reasoning, by plotting the normalized probability that a ship is hidden in different cells.
-As can be seen, the highest probabilities are to the sides of the revealed ships, and in the first row.
-
-In the third column we plot the distribution of clicks for non-pretend players.
-This distribution is in agreement with the hit probability map (board A: `r cor_A_nonpretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_nonpretend%>%apa_print()%>%'$'(full_result)`).
-Finally, in the fourth column we plot the distribution of cell selections for pretend players.
-Although noisier, this distribution is also in agreement with the hit probability map (board A: `r cor_A_pretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_pretend%>%apa_print()%>%'$'(full_result)`), and more importantly, with the hit distribution of non-pretenders for the same board configuration (board A: `r cor_A_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`, board B: `r cor_B_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`).
-
-```{r half-games, echo=FALSE, out.width = '100%', fig.cap = "Half games.
-First column: the two boards, as they appeared to pretenders.
-Second column: objective hit probability, given game state.
-Third and fourth columns: empirical click probabilities in non-pretend and pretend games, respectively."}
-knitr::include_graphics("../figures/Half_games.png")
-```
-
-<!-- Furthermore, compared to non-pretend games, in pretend games participants were more likely to make irrational clicks on cells where a ship could not in fact be hiding (a median of `r E2.num_irrational_clicks%>%filter(test_part=='pretend')%>%pull(nic)%>%median()` per game in pretend versus `r E2.num_irrational_clicks%>%filter(test_part=='nonpretend')%>%pull(nic)%>%median()` in non-pretend games.
-`r E2.num_irrational_clicks%>%spread(test_part,nic)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
--->
-
-## Judge trials
-
-```{r condition-classifier, echo=FALSE, eval=FALSE}
-
-set.seed(42);
-
-gamewise_data <- E2.click_log_with_boards %>% 
-  filter(!genuine_first | test_part=='pretend') %>% 
-  group_by(subj_id,test_part,grid_number) %>%
-  filter(sum(RT)<=60000) %>%
-  summarise(median_RT=median(RT),
-            mean_p_click_rank = mean(p_click_rank),
-            first_click_latency=RT[click_number==1],
-            num_irrational_clicks = sum(p_click==0)) %>%
-  ungroup()%>%
-  select(test_part,mean_p_click_rank,median_RT,num_irrational_clicks);
-
-gamewise_data_nonpretend = gamewise_data%>%filter(test_part=='nonpretend') %>%mutate(nrow=1:n())
-gamewise_data_pretend = gamewise_data%>%filter(test_part=='pretend')
-rows <- sample(nrow(gamewise_data_nonpretend))
-gamewise_data_pretend <- gamewise_data_pretend[rows,]%>%mutate(nrow=1:n())
-
-pair_gamewise_data <- full_join(gamewise_data_nonpretend,
-                                gamewise_data_pretend,
-                                by='nrow',
-                                suffix=c('.nonpretend','.pretend')) %>%
-  mutate(which_side = sample(c(-1,1),nrow(gamewise_data_pretend),replace=TRUE),
-         mean_p_click = which_side*(mean_p_click_rank.nonpretend-mean_p_click_rank.pretend),
-         median_RT = which_side*(median_RT.nonpretend-median_RT.pretend),
-         num_irrational_clicks = which_side*(num_irrational_clicks.nonpretend-num_irrational_clicks.pretend),
-         which_side=factor(which_side)) %>%
-  select(which_side,mean_p_click,median_RT,num_irrational_clicks)
-
-train_control <- trainControl(method = "CV", number=10, sampling='down')
-
-# train the model on training set
-model <- train(test_part ~ .,
-               data = gamewise_data,
-               trControl = train_control,
-               method = "svmLinear",
-               preProcess = c("center","scale"))
-
-pair_model <- train(which_side ~ .,
-               data = pair_gamewise_data,
-               trControl = train_control,
-               method = "svmLinear",
-               preProcess = c("center","scale"))
-
-# print cv scores
-
-```
-
-Despite the fact that pretenders were overall successful in mimicking patterns of cell selection and decision latency, they also left many clues for an independent observer to pick up on: they were slower, displayed exaggerated effects of decision uncertainty and hit success on response time, and were less optimal in choosing where to click next.
-To illustrate, a support vector machine (SVM) algorithm reached an accuracy level of 72% in linearly classifying condition (pretend / non-pretend) based on just three basic summary features: median decision latency, game optimality score, and number of irrational clicks per game.
-
-```{r judge}
-
-get_common_board = function(grid_number,test_part) {
-  pretend_grids = grid_number[which(test_part=='pretend')];
-  nonpretend_grids = grid_number[which(test_part=='nonpretend')];
-  common_board = intersect(pretend_grids,nonpretend_grids);
-  return(common_board)
-};
-
-common_boards <- E2.df %>%
-  filter(test_part=='pretend' | test_part=='nonpretend') %>%
-  dplyr::select(subj_id,grid_number,test_part) %>%
-  group_by(subj_id) %>%
-  summarise(common_board = get_common_board(grid_number,test_part))
-
-used_for_pretend <- function(s, g) {
-
-  return (E2.df%>%filter(test_part=='pretend' &
-                           subj_id==s & 
-                           grid_number == g) %>%
-            nrow()==1)
-}
-
-used_for_nonpretend <- function(s, g) {
-
-  return (E2.df%>%filter(test_part=='nonpretend' &
-                           subj_id==s & 
-                           grid_number == g) %>%
-            nrow()==1)
-}
-
-E2.judge_df <- E2.df %>%
-  filter(test_part=='judge' & !(cheat_player %in% E2.lucky_pretenders))%>%
-  dplyr::select(subj_id, 
-         grid_number, 
-         cheat_player,
-         genuine_player,
-         cheater,
-         noncheater,
-         decision,
-         replay_log,
-         choose_pretender,
-         correct)%>%
-  rowwise()%>%
-  mutate(choose_pretender=choose_pretender=='True',
-         used_for_pretend = used_for_pretend(subj_id,grid_number),
-         used_for_nonpretend = used_for_nonpretend(subj_id,grid_number),
-         common_board = used_for_pretend & used_for_nonpretend,
-         board_type = ifelse(common_board, 'common',
-                             ifelse(used_for_pretend, 'pretend',
-                                    ifelse(used_for_nonpretend,'nonpretend','new')))
-)
-
-p <- E2.judge_df %>%
-  group_by(subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) +
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') +
-  scale_x_continuous(breaks=c(0,1,2,3,4,5));
-
-ggsave('../figures/E2_correct_guesses.png',p,width=5,height=3,dpi=300)
-
-p <- E2.judge_df %>%
-  mutate(choose_pretender = ifelse(choose_pretender,
-                                   'who had hints?', 
-                                   'who had no hints?')) %>% 
-  group_by(choose_pretender,subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) + 
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') + 
-  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
-  facet_wrap(~choose_pretender,nrow=2);
-
-ggsave('../figures/E2_correct_guesses_by_choose_pretender.png',p,width=5,height=3,dpi=300);
-
-p <- E2.judge_df %>% 
-  group_by(board_type,subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) + 
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') + 
-  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
-  facet_wrap(~board_type,nrow=1);
-
-ggsave('../figures/E2_correct_guesses_by_board_type.png',p,width=5,height=3,dpi=300)
-```
-
-Despite this, participants were never significantly above chance in judging which of the two presented games came from a pretender (mean accuracy: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%; t test against 50: `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%t.test(mu=50))$statistic`).
-
-For a subset of `r E2.judge_df%>%filter(common_board)%>%pull(correct)%>%length()` players, one of the five boards presented in the judging block was already presented both in the pretend and in the non-pretend blocks.
-Still, even this subset of participants who had just experienced a board both as pretenders and as non-pretenders were not significantly above chance in telling which of two other players had hints for this same board (mean accuracy: 55%, $p=.141$).
-
-## Are good pretenders also better judges?
-
-```{r judge-pretend-cor, echo=FALSE}
-
-unidentifiable_ss <- E2.judge_df %>%
-  mutate(subj_num=substr(as.character(subj_id),1,3))%>%
-  group_by(subj_num)%>%
-  summarise(subj_id=subj_id, n=n())%>%
-  filter(n>5)%>%
-  pull(subj_id)%>%
-  unique();
-
-E2.judge_performance <- E2.judge_df %>%
-  filter(!(subj_id %in% unidentifiable_ss)) %>%
-  mutate(subj_num=as.numeric(substr(as.character(subj_id),1,3)))%>%
-  group_by(subj_num)%>%
-  summarise(judge_rate=sum(correct)/5);
-
-E2.pretend_performance <- E2.judge_df %>%
-  filter(!(subj_id %in% unidentifiable_ss)) %>%
-  mutate(subj_num=cheat_player)%>%
-  group_by(subj_num)%>%
-  summarise(ntrials=n(),
-            successful_pretend=ntrials-sum(correct),
-            pretend_rate = successful_pretend/ntrials);
-
-E2.performance_df <- merge(E2.judge_performance,E2.pretend_performance);
-
-p <- E2.performance_df %>%
-  ggplot(aes(x=judge_rate,y=pretend_rate)) +
-  geom_jitter() + 
-  labs(x='judge accuracy',y='pretense success rate') +
-  theme_classic()
-
-ggsave('../figures/E2_judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)
-
-min_njudge = seq(0,10);
-cors =c()
-ns = c()
-
-for (minn in min_njudge) {
-
-  temp_df = E2.performance_df %>%
-    filter(ntrials>minn)
-
-  cur_cor = cor(temp_df$pretend_rate,temp_df$judge_rate);
-  cur_n = temp_df%>%nrow()
-  cors = c(cors,cur_cor)
-  ns = c(ns, cur_n)
-}
-
-judge_pretend_correlation_df <- data.frame(min_njudge,cors,ns);
-
-```
-
-```{r judge-optimality-cor, echo=FALSE}
-E2.judge_summary <- E2.judge_df %>%
-  group_by(subj_id) %>%
-  summarise(judge = sum(correct));
-
-E2.optimality_summary <- E2.click_log_with_boards %>%
-  group_by(subj_id,test_part) %>%
-  summarise(p_click_rank = mean(p_click_rank)) %>%
-  spread(test_part,p_click_rank)
-
-E2.judge_summary <- E2.judge_summary %>%
-  merge(E2.optimality_summary)
-```
-
-The abilities to pretend and to detect pretense in others both rely on some form of theory of mind: reasoning about counterfactual knowledge states of oneself in the first case, and of others in the second.
-For example, in scrub jays, individual birds that pilfered another bird's caches were also more careful in hiding their own food, suggesting an ability to project from their own experience of stealing to the behavior of others [@emery2001effects].
-Along similar lines, we reasoned that those participants whose pretend games better resembled the games of genuine players would also be more accurate in detecting pretense in others.
-To our surprise, this was not at all the case (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate))$full_result`).
-Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend))$full_result`).
-
-Despite our considerable sample size of 500 participants, we are cautious in making strong claims based on the absence of a correlation between pretense and pretense detection abilities in our sample.
-We note, however, that previous developmental research has identified weaker correlations with theory of mind measures for a tendency to intentionally deceive another player in a game setting, compared to an ability to detect and understand deception, potentially indicating partly different mechanisms [@shultz1981development].
+## Results
 
@@
 
-Many tasks (including writing papers for conferences) require us to overcome the 'curse of knowledge' --- the challenge of putting ourselves in the shoes of someone who doesn't know what we know.
-This is just an example of a more general fact: the ability to simulate counterfactual knowledge states is necessary for efficient and flexible communication.
-To do this, we must have an internal self-model that can generate thoughts and behavior for hypothetical beliefs, desires, and perceptions.
-
-Research on Bayesian Theory of Mind provides some support for the existence of such a model by measuring subjects' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational], or implicitly [@onishi2005fifteen; @liu2017ten].
-Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state - in this case, a counterfactual knowledge state in which a known piece of information is unknown.
-Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
-Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
-
-Our subjects rose to the challenge.
-Their pretense behavior, although not perfect, closely resembled that of non-pretenders --- not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
-Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: "What would I have done if I didn't have this piece of knowledge?"
-
-An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game.
-This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain participants' pretense ability in our data, for two indirect reasons.
-First, suppressing thoughts on demand is notoriously difficult [@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible.
-Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with a self-simulation account.
-In future work, we intend to more directly rule out this knowledge-suppression hypothesis, and well as to generalize our results to other games.
-
-In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects' models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
-While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.
+\newpage
 
@@
 
-```{r}
-# References will be generated automatically by Pandoc and included here.
-# The following code is some latex to format the bibliography. Do not remove it.
-```
-
-```{=tex}
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
-```
-\noindent
+::: {#refs custom-style="Bibliography"}
+:::

### 2023-01-24 docs/Methods and results.Rmd
first three hypotheses in battleships

@@
 
-all_possible_boards <- read.csv('../../analysis/flat_boards.csv',header=FALSE);
+all_possible_boards <- read.csv('../analysis/flat_boards.csv',header=FALSE);
 
@@
 }
+
+```
+
+```{r battleship-parse-click-log, echo-FALSE, cache=TRUE}
+
+E2.click_log <- readRDS('../experiments/Battleships2/data/batch1/click_log.Rda');
+
+# E2.click_log <- data.frame(matrix(ncol=10,nrow=0, 
+#                                dimnames=list(NULL, 
+#                                              c("subj_id",
+#                                               "test_part", 
+#                                               "grid_number", 
+#                                               "num_clicks",
+#                                               "i",
+#                                               "j",
+#                                               "hit",
+#                                               "t",
+#                                               "click_number",
+#                                               "genuine_first"))))
+# 
+# 
+# for (row in 1:nrow(E2.click_df)) {
+#   
+#     subject_click_log <- data.frame(fromJSON(E2.click_df[row, ]$click_log)) %>%
+#     mutate(
+#       click_number = 1:n(),
+#       subj_id = E2.click_df[row, ]$subj_id,
+#       test_part = E2.click_df[row, ]$test_part,
+#       grid_number = E2.click_df[row, ]$grid_number,
+#       num_clicks = E2.click_df[row, ]$num_clicks,
+#       genuine_first = E2.click_df[row, ]$genuine_first
+#     )
+#     
+#     E2.click_log <- rbind(E2.click_log, subject_click_log);
+# }
+#     
+# 
+# E2.click_log <- E2.click_log %>%
+#   relocate(subj_id, .before = i) %>%
+#   relocate(genuine_first, .before = i) %>% 
+#   relocate(test_part, .before=i) %>% 
+#   relocate(grid_number, .before=i) %>% 
+#   relocate(click_number, .before=i) %>%
+#   group_by(subj_id,test_part,grid_number) %>%
+#   mutate(RT=t-lag(t,default=0)) %>%
+#   filter(!(subj_id %in% E2.lucky_pretenders))%>%
+#  mutate(hit_bin = hit !=0,
+#         lag3 = lag(hit,3)!=0,
+#         lag2 = lag(hit,2)!=0,
+#         lag1 = lag(hit,1)!=0,
+#         lead1 = lead(hit,1)!=0,
+#         lead2 = lead(hit,2)!=0,
+#         lead3 = lead(hit,3)!=0)
+
+```
+
+```{r battleship-duration, echo-FALSE, cache=TRUE}
+
+E2.duration_log <- E2.click_log %>% 
+  group_by(subj_id,test_part,grid_number, genuine_first) %>%
+  summarise(duration=sum(RT)) %>% 
+  group_by(subj_id,test_part,genuine_first) %>% 
+  summarise(duration=median(duration)) %>% 
+  spread(test_part,duration) %>%
+  mutate(diff=pretend-nonpretend)
+
+```
+
+*Hypothesis 1 (game duration)*: Pretend games were longer than non-pretend games by `r (E2.duration_log %>%pull(diff)%>%mean()/1000)%>%round()` seconds on average.
+This was true in the pretend-first group (`r E2.duration_log %>% filter(!genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), and in the pretend-second groups (`r E2.duration_log %>% filter(genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+```{r battleship-first-click, echo-FALSE, cache=TRUE}
+
+E2.first_click_log <- E2.click_log %>% 
+  group_by(subj_id,test_part,grid_number, genuine_first) %>%
+  summarise(RT=RT[click_number==1]) %>% 
+  group_by(subj_id,test_part,genuine_first) %>% 
+  summarise(RT=median(RT)) %>% 
+  spread(test_part,RT) %>%
+  mutate(diff=pretend-nonpretend)
+
+```
+
+*Hypothesis 2 (first click latency)*: The first click in pretend games took longer to execute by `r E2.first_click_log %>%pull(diff)%>%mean()%>%round()` ms.
+This was true in the pretend-first group (`r E2.first_click_log %>% filter(!genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), but not in the pretend-second groups (`r E2.first_click_log %>% filter(genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+As per our pre-registration document, we focused our analysis on the first blocks only, using a between-subject t-test.
+This test revealed a significant difference in the first click latency (`r t.test(E2.first_click_log%>%filter(!genuine_first)%>%pull(pretend), E2.first_click_log%>%filter(genuine_first)%>%pull(nonpretend))%>%apa_print()%>%'$'(full_result)`).
+
+```{r battleship-num-clicks, echo=FALSE, cache=TRUE}
+
+E2.click_df <- E2.df %>%
+  dplyr::select(subj_id, 
+         test_part, 
+         grid_number, 
+         num_clicks,
+         click_log,
+         genuine_first) %>%
+  mutate(click_log = gsub("\'","\"", click_log)) %>%
+  filter(test_part=='pretend' | test_part=='nonpretend');
+
+E2.median_clicks <- E2.click_df %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(num_clicks=median(num_clicks)) %>%
+  spread(test_part,num_clicks) %>%
+  mutate(diff=pretend-nonpretend);
+
+E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();
+
+E2.median_clicks_filtered <- E2.click_df %>%
+  filter(!(subj_id %in% E2.lucky_pretenders))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(num_clicks=median(num_clicks)) %>%
+  spread(test_part,num_clicks);
+
+p<- E2.click_df %>%
+    group_by(subj_id,test_part, genuine_first) %>%
+    summarise(num_clicks=median(num_clicks)) %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=num_clicks,fill=test_part))+
+    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='median number of clicks', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic()
+    # geom_vline(xintercept= median_simulated)+
+    # facet_wrap(~genuine_first, nrow=2);
+
+ggsave('../docs/figures/E2_num_clicks.png',p,width=5,height=2, dpi=300)
 
@@
 
-*Hypothesis 1 (game duration)*: We will test the null hypothesis that pretend games take as long as non-pretend games.
-For each participant, a difference between median game duration in pretend and non-pretend games will be extracted.
-Subject-wise differences will then be carried over to a t-test at the group level.
+*Hypothesis 3 (number of clicks)*:
 
-*Hypothesis 2 (first click latency)*: We will test the null hypothesis that the first click in pretend games takes as long to execute as in non-pretend games.
-For each participant, a difference between median first click latency in pretend and non-pretend games will be extracted.
-Subject-wise differences will then be carried over to a t-test at the group level.
+To sink all ships, players had to click on at least 7 and at most 25 squares.
+A simulated player that clicks randomly had a median total click number of `r num_clicks_sim%>%median()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a median total click number of `r num_clicks_optimal_sim%>%median()`.
+Among our players, the median number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games.
 
-*Hypothesis 3 (number of clicks)*: We will test the null hypothesis that pretend games are as long, in terms of total number of clicks, as non-pretend games.
-For each participant, a difference between median number of clicks in pretend and non-pretend games will be extracted.
-Subject-wise differences will then be carried over to a t-test at the group level.
+No significant difference in the total number of clicks was observed in the pretend-first group (`r E2.median_clicks%>%filter(!genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), but a significant difference was observed in the pretend-second group (`r E2.median_clicks%>%filter(genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Note that the two effects point in opposite directions, reflecting a tendency to make fewer clicks in the second block of the game.
+Crucially, when focusing on the first block only, We observed no significant difference in the number of clicks between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend),E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
+
+In `r E2.click_df%>%filter(num_clicks==7)%>%nrow()` pretend games from `r E2.lucky_pretenders%>%length()` players, games were completed after 7 clicks only, without ever missing a ship.
+This never happened in non-pretend games.
+We assumed that these participants did not follow the task instructions, and excluded them from all analyses.
 

### 2023-01-25 docs/preregisteredMethodsAndResults.Rmd
pre-registered methods and results

--- 2023-01-24 docs/Methods and results.Rmd
@@
 
-*Hypotheses 4-6 (click latency by click outcome)*: Trials will be classified based on whether they resulted in hitting a ship (*hit*) or not (*miss*).
+In our instructions to pretenders, we asked them to pretend not to know where the ships were, not only in *where* they chose to click, but also in *when* they clicked.
+In the following set of analyses, we identified patterns in click latency in non-pretend games, and asked whether the same patterns are also observed in pretend games.
+We focused on the effects of hitting versus missing a ship, decision uncertainty, and ship completion.
+
+```{r RT_by_accuracy}
+
+N_perm <- 1000;
+bootstrap_error <- function(x, N_perm) {
+  N <- length(x)
+  medians = c();
+  for (i in 1:N_perm) {
+    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
+  };
+  return(sd(medians))
+}
+
+E2.RT_by_lag3 <- E2.click_log %>%
+  filter(click_number>3) %>%
+  group_by(subj_id,test_part,lag3,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag3,RT,sep='_') %>%
+  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);
+
+E2.RT_by_lag2 <- E2.click_log %>%
+  filter(click_number>2) %>%
+  group_by(subj_id,test_part,lag2,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag2,RT,sep='_') %>%
+  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);
+
+E2.RT_by_lag1 <- E2.click_log %>%
+  filter(click_number>1) %>%
+  group_by(subj_id,test_part,lag1,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag1,RT,sep='_') %>%
+  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
+
+E2.RT_by_hit <- E2.click_log %>%
+  filter(click_number>0) %>%
+  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(hit_bin,RT,sep='_') %>%
+  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
+
+E2.RT_by_lead1 <- E2.click_log %>%
+  filter(click_number<num_clicks) %>%
+  group_by(subj_id,test_part,lead1,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead1,RT,sep='_') %>%
+  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
+
+E2.RT_by_lead2 <- E2.click_log %>%
+  filter(click_number<num_clicks-1) %>%
+  group_by(subj_id,test_part,lead2,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead2,RT,sep='_') %>%
+  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)
+
+E2.RT_by_lead3 <- E2.click_log %>%
+  filter(click_number<num_clicks-2) %>%
+  group_by(subj_id,test_part,lead3,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead3,RT,sep='_') %>%
+  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)
+
+E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
+  merge(E2.RT_by_lag2) %>%
+  merge(E2.RT_by_lag1) %>%
+  merge(E2.RT_by_hit) %>%
+  merge(E2.RT_by_lead1) %>%
+  merge(E2.RT_by_lead2) %>%
+  merge(E2.RT_by_lead3);
+
+E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
+  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
+  mutate(position = ifelse(position=="lag3_diff", 3,
+                           ifelse(position=="lag2_diff",2,
+                                  ifelse(position=="lag1_diff",1,
+                                         ifelse(position=="hit_diff",0,
+                                                ifelse(position=="lead1_diff",-1,
+                                                       ifelse(position=="lead2_diff",-2,-3
+                                                              ))))))) %>%
+  dplyr::select(subj_id,test_part,genuine_first,position,difference);
+
+p<- E2.RT_by_hit_lags_long %>% 
+  filter(position<2 & position > -1)%>%
+  group_by(test_part,position) %>%
+  summarise(median_difference=median(difference),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=0:1,labels=c('hit on current trial', 'hit on previous trial'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
+  theme_classic()
+
+ggsave('../figures/E2_RT_by_lag_three_only_collapse.pdf',p,width=7)
+ggsave('../figures/E2_RT_by_lag_three_only_collapse.png',p,width=4, height=3.5, dpi=300)
+
+```
+
+*Hypothesis 4 (click latency by outcome)*: In non-pretend games, trials that resulted in a hit were faster by `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%mean()%>%round()%>%abs()` than those that resulted in a miss (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`) will be classified based on whether they resulted in hitting a ship (*hit*) or not (*miss*).
 We will then contrast median click latency between hits and misses separately for pretend and non-pretend games.

### 2023-01-26 experiments/Battleships2/protocolFolder/docs/exp2_PR.Rmd
experiment code: Battleship

---
title             : "Pretending not to know, Exp.2: pre-registration document"
shorttitle        : "PNTN EXP2 PRE-REGISTRATION"

author: 
  - name          : "Matan Mazor"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "12 Queen Square, WC1N 3BG London"
    email         : "mtnmzor@gmail.com"
  - name          : "Ian Phillips"
    affiliation   : "2"
  - name          : "Chaz Firestone"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "University College London"
  - id            : "2"
    institution   : "Johns Hopkins University"

authornote: |
  The authors declare no competing interests

abstract: |
  Generally, what we do depends on what we know. But sometimes we try to appear not to know something that we really do know. Such pretense behavior relies on counterfactual self-simulation — an understanding of how we would behave if our knowledge were different — and so provides an opportunity to investigate how well people can emulate a hypothetical knowledge state. Surprisingly, despite its immediate relevance to metacognition and theory of mind, little research has focused on quantifying pretense accuracy, relative to non-pretense behaviour. In a replication of a first exploratory experiment, we examine the ability to both produce and detect pretense behaviour using the game "Battleships" — normally played by searching for ships hidden behind cells in a grid. In 'pretend' blocks, subjects will be told where all the ships are but be asked to play as if they lacked this knowledge. We will quantify subjects' self-simulation ability by comparing their behaviour in pretend and non-pretend games. We will further test players' ability to detect pretense, and examine the association between pretense ability and pretense-detection ability.

keywords          : "Pretense, self-simulation, metacognition, theory of mind"
wordcount         : "3634"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib");
library('reticulate')
library('tidyverse')
library('broom')
library('cowplot')
library('MESS') # for AUCs
library('lsr') # for effect sizes
library('pwr') # for power calculations
library('brms') # for mixed effects modeling
library('BayesFactor') # for Bayesian t test

N <- 500;
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r load-data, warning=FALSE}
E1.df <- read.csv('..\\experiments\\Battleships\\data\\batch1\\jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
  rename(subj_id = participant_number) %>%
   # we accidentally have two subjects 251: 5c360ede943bda0001db4f1b and 6092f8b1c80be84bfe9fc57d.
  # to fix this, I change 6092f8b1c80be84bfe9fc57d to be 2511
  mutate(subj_id = ifelse(PROLIFIC_PID=='6092f8b1c80be84bfe9fc57d',2511,subj_id))%>%
  mutate(subj_id = factor(subj_id)) %>%
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
  mutate(genuine_first=genuine_first=='True')

E1.export <-read.csv('..\\experiments\\Battleships\\data\\batch1\\prolific_export_batch1.csv');
```

# Motivation

<!-- The ability to reason about hypothetical mental states is critical for inferring goals and beliefs from observed behaviour [@baker2009action; @baker2011bayesian], predicting behaviour based on goals and belief states [@wimmer1983beliefs; @wellman2001meta], pre-commitment in action planning [@moran2021model], inference about the absence of memories [@dodson2002false], [...].  -->

The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory]. This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation]. This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own. Moreover, deception often involves pretense behaviour, which in turn relies on an ability to simulate and mimic one's own behaviour under a hypothetical belief state. For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") — you also need to convincingly simulate and mimic your hypothetical behaviour had you not known about the party ("Where would I look first? What would I say? How long would it take me to recover from the surprise?"). This reliance of pretense behaviour on self-simulation makes it an ideal opportunity to examine metacognitive knowledge about one's own mental states, and the potential reliance of this knowledge on a self-simulation. By comparing non-pretend and pretend behaviour, we can ask which aspects of their cognitive processes subjects can simulate, and which aspects are not represented in their mental models of their own cognition.

To this end, here we examine pretense in a game setting. Using an online version of the game "Battleships", participants will play a non-pretend (normal) version of the game, as well as a pretend version where they will be given secret information about the location of hidden ships, but will try to behave as if they don't have this information. Examining different aspects of pretend and non-pretend games (such as decision latency, sequential dependencies and game optimality), we will ask to what extent can people simulate their own play behaviour under a different knowledge state. We will further ask whether participants can discriminate between pretend- and non-pretend games of other participants, and whether this ability predicts their own pretense quality.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

The research complies with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University. Participants will be recruited via Prolific and will give informed consent prior to their participation. They will be selected based on their acceptance rate (>95%) and for being native English speakers. We will collect data until we reach `r N` included participants. The entire experiment will take 20 minutes to complete (median completion time in our pilot data: `r printnum(E1.export%>%pull(time_taken)%>%median()/60)` minutes). Participants will be paid 3.15 USD for their participation, equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment.

## Procedure

```{r design, echo=FALSE, fig.cap="Experimental Design: players will play 10 games of the game Battleships (on a 5x5 board) in two conditions presented in two blocks of 5 games. The order of blocks will be randomized between players. In non-pretend games, players will start the game not knowing where the ships are and will try to sink all ships with as few clicks as possible. In pretend games, ship locations will be marked with a cross such that players will know where they are hiding, and their task will be to play the game as if they do not have this information. Lastly, in judge trials, players will observe a replay of the games of two previous players – one pretender and one non-pretender, and their task will be to identify which was which. Lower panel: we created nine different boards, each containing two 2-square patrol boats and one 3-square submarine. The assignment of boards to conditions will be randomized across participants in the following way: first, the order of boards will be permuted randomly for each participant. Then, the first 5 boards in this permuted order will be used for non-pretend games, the last 5 boards will be used for pretend games, and boards 1,3,5,7 and 9 will be used for judge trials. This means that one board (“the common board”) will be used in all three parts of the experiment. The identity of the common board will be different for each participant.", out.width = '100%'}
knitr::include_graphics("figures/Battleships_design.png")
```

The first instructions screen will inform participants that the experiment, based on the game Battleships, has three parts, and that the points they will accumulate will translate to a monetary bonus payment. They will then be presented with a leaderboard of previous players. Then, the rules of the game will be presented:

> In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
What makes the game difficult is that you can’t see the ships; all you can see is a grid of squares, and you have to guess where the ships are. To sink a ship, you need to click on all of the squares it is located in. If you hit part of a ship, the square will turn red. If there is no ship in the square, it will turn blue.

We will further explain that in this version of the game, ships can touch corners, but their sides can't touch. This explanation will be accompanied by a visual presentation of legal and illegal ship configurations.

Once they respond correctly to a multi-choice comprehension question ("the goal of the game is to...": "sink all ships with as few clicks as possible"), participants will play a practice round on a 3x3 board.

After the main instructions, comprehension question and practice round, participants will complete one pretend and one non-pretend block, each followed by one half-game (see below for details). The order of pretend and non-pretend blocks will be counterbalanced between participants. Each block will comprise five games played on five out of nine different boards. The allocation of boards to conditions will be randomized between participants in the following way: for each participant, we will randomly determine a random permutation of the nine boards. Based on this permutation, boards 1-5 will be used in the non-pretend block, boards 5-9 will be used in the pretend block, and boards 1,3,5,7 and 9 will be used in the final judge block (see below for details). The order of boards within a block will be randomized.

### Non-pretend games

In non-pretend games, participants will sink one 3-square submarine and two 2-square patrol boats with as few clicks as possible. An online counter of the number of clicks will be displayed on the screen. After each game, feedback will be given about the number of clicks and resulting number of points obtained.

After completing the five games, participants will perform one half-game (see below for details).

### Pretend games

Participants will be given the following instructions:

> In the next part of the experiment, you'll play 5 games where you sink one 3-square submarine and two 2-square patrol boats.

> However, this time your goal is different.

> In this round, we’re going to tell you where the ships are, but **we want you to act like you don’t know this information**. We’ve marked the ships’ locations with a cross, so you’ll know where they are the whole time; but your job is to play the game as if these hints aren’t there.

> To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are. We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you'll get bonus points. But if your games look different, you won't get these bonus points. Your number of clicks in this part will not affect your bonus. Only your ability to play like you had no hints.

After one practice round, pretending not to know the locations of two 2-square patrol boats on a 3x3 board, and one comprehension question ("In this part of the experiment my goal is to...": "play the game as if the cross hints were not there so that I look like someone who had no hints"), participants will play five pretend games. Each game will be followed by a short message, reminding them that a game that looks similar to the game of participants who had no hints will be awarded with 10 bonus points.

After completing the five games, participants will perform one half-game (see below for details).

### Half games

In order to directly compare participants' pretend and non-pretend games for identical belief states (true or pretended knowledge about which cells hide water and which hide ships), we will ask participants to also complete one pretend and one non-pretend game, given a partly finished board with the content of 7 cells already revealed (see Fig. \@ref(fig:halfgames)). We designed our half games to produce a strong expectation to find a ship in specific cells, but not in others. In one of the two games (A), these expectations are not fulfilled. In the other (B), they are. The assignment of half-completed boards to pretend and non-pretend conditions will be randomized between participants.

Instructions for the non-pretend half-game will be:

> For the next game, you will continue the game of another player from where he or she stopped. Just like in the previous games, here also there are two 2-square patrol boats and one 3-square submarine hiding in the water, and you will get a bonus for revealing all the ships with as few clicks as possible.

Instructions for the pretend half-game will be:

> For the next game, you will continue the game of another player from where he or she stopped. Just like in the previous games, here also there are two 2-square patrol boats and one 3-square submarine hiding in the water. We marked for you the true locations of the ships with a cross, but your bonus points will depend on your ability to play as if you didn't have these hints.

```{r game_board_analysis, echo=FALSE}

all_possible_boards <- read.csv('../analysis/flat_boards.csv',header=FALSE);

create_board_states <- function(flat_positions,hit) {

  board_state = rep(NA,25)
  board_states = c();
  for (p in seq_along(flat_positions)) {
    board_states = c(board_states, paste(board_state,collapse=','));
    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
  }

  return(board_states)
};

get_likelihood <- function(board_state) {

  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);

  revealed_positions <- which(!is.na(board_state));

  boards_to_remove = c();

  for (position in revealed_positions) {
    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
  }

  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
  likelihood[revealed_positions]=NA;

  return(paste(likelihood,collapse=','))

};

get_posterior <- function(likelihood) {
    likelihood = scan(text= likelihood, what = numeric(), sep="," , quiet = TRUE);
    posterior = likelihood/sum(likelihood,na.rm=TRUE); # assuming a uniform prior
    return(paste(posterior,collapse=','))
}

get_p_click <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(posterior[flat_position])
}

get_p_click_rank <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  rank_posterior = rank(-posterior)
  return(rank_posterior[flat_position])
}

get_posterior_entropy <- function(posterior) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  posterior[which(posterior==0)]=NA; #to avoid infinity*0
  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
  return(entropy);
}

plot_posterior <- function(posterior,file_name) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  board_df <- data.frame(i=rep(c(1,2,3,4,5),each=5),j=rep(c(1,2,3,4,5),5),posterior=posterior);
  p <- ggplot(board_df,aes(x=j,y=i,fill=posterior)) +
    geom_tile() +
    coord_equal() +
    scale_x_continuous(breaks=c()) +
    scale_y_continuous(breaks=c(),trans='reverse') +
    scale_fill_gradient(low='#649bc8',high='#c83237',na.value = "white",limits = c(0,0.25))+
    labs(x='',y='') +
    theme_classic();

  ggsave(paste('figures/',file_name,'.pdf',sep=''),width=5,dpi=300);
}

board_state <- "NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,0,1,1,0,NA,NA,NA,0,NA,NA,1,1,NA,NA,NA"
posterior <- board_state%>%get_likelihood()%>%get_posterior();
plot_posterior(posterior,'posterior_boardA');

board_state <- "NA,NA,NA,NA,0,NA,NA,0,NA,NA,NA,NA,NA,NA,0,NA,NA,1,0,NA,NA,NA,1,NA,1"
posterior <- board_state%>%get_likelihood()%>%get_posterior();
plot_posterior(posterior,'posterior_boardB');

```

```{r halfgames, echo=FALSE, fig.cap="The two half games. Left column: True location of ships is marked with a cross, but this will only be visible to participants in the pretend version. Right column: objective probability of hitting a ship, given board configuration.", out.width = '100%'}
knitr::include_graphics("figures/Half_games.png")
```

### Judge trials

In the final part of the experiment, participants will observe the games of previous players and determine who had hints and who didn't. Instructions for this part are:

>  In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of two players was given hints (and acted like they didn't). We will do 5 trials, each featuring two new players, only one of which had hints.

For a random half of the participant, the instructions continue:

> We will show you replays of their games side by side. Your task is to decide who had hints (and acted like they didn't). After you've watched the games of both players, click on the board of the player who had hints. For each game that you get right, you will receive 10 points. Good luck!

For the other half, the instructions continue:

> We will show you replays of their games side by side. Your task is to decide who played the game normally (without hints). After you've watched the games of both players, click on the board of the player who played normally. For each game that you get right, you will receive 10 points. Good luck!

Then, on each judge trial, two empty grids will be presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid. The two grids will correspond to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders. Only games shorter than one minute will be chosen for presentation in this part. For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) will be chosen for presentation in this part. Judge participants will press one of two buttons to start a replay of one of the two grids. Replays will be presented in "real time", mimicking not only where participants clicked, but also when. Only after watching both replays will it be possible to click on one of the two boards to indicate a decision. Alternatively, participants will be allowed re-watch the two games until they make up their minds. After making a decision, participants will be informed whether they will receive the 10 points, or alternatively, whether the pretender will receive them in case the pretender managed to trick them.

Lastly, participants will be asked the following debrief questions:

> Did you have a strategy that you used for pretending you had no hints? How about telling between players with and without hints - did you have a strategy for that?

And:

> We would appreciate if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.

### Randomization

The order and timing of experimental events will be determined pseudo-randomly by the Mersenne Twister pseudorandom number generator, initialized in a way that ensures registration time-locking [@mazor2019novel].

## Data analysis

### Rejection criteria

Pretend games in which all clicks resulted in hits will not be included for analysis. Similarly, participants will be excluded if two or more of their pretend games were all hits.

### Data preprocessing

Data will not be preprocessed in any way.

### Hypotheses and analysis plan

This study is designed to explore subjects' capacity for self-simulation under a counterfactual knowledge state, and the limits of this capacity. In particular, we are interested in comparing non-pretend and pretend games and identifying similarities and differences. We will also assess subjects' capacity to discriminate pretend and non-pretend games, and how this capacity relates to their ability to pretend.

Unless otherwise specified, all hypotheses will be tested using a repeated measures t-test, with a significance level of 0.05.

All hypotheses will be tested separately for the group of participants that pretended in the first and in the second block. If results for the two groups diverge, analysis will be repeated on the first block of all participants (pretend games of pretend-first participants and non-pretend games of pretend-second participants), in a between-subjects manner, and results will be interpreted in light of this analysis.

*Hypothesis 1 (game duration)*: We will test the null hypothesis that pretend games take as long as non-pretend games. For each participant, a difference between median game duration in pretend and non-pretend games will be extracted. Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 2 (first click latency)*: We will test the null hypothesis that the first click in pretend games takes as long to execute as in non-pretend games. For each participant, a difference between median first click latency in pretend and non-pretend games will be extracted. Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 3 (number of clicks)*: We will test the null hypothesis that pretend games are as long, in terms of total number of clicks, as non-pretend games. For each participant, a difference between median number of clicks in pretend and non-pretend games will be extracted. Subject-wise differences will then be carried over to a t-test at the group level.

*Hypotheses 4-6 (click latency by click outcome)*: Trials will be classified based on whether they resulted in hitting a ship (*hit*) or not (*miss*). We will then contrast median click latency between hits and misses separately for pretend and non-pretend games. We will test for a difference in trial latency as a function of outcome on pretend games, on non-pretend games, and for an interaction effect between click outcome (hit or miss) and condition (pretend or non-pretend) on click latency. The same analysis will then be repeated by classifying trials on the basis of whether the previous trial resulted in a hit or a miss (*Hypothesis 5*), and whether the next trial will result in a hit or a miss (*Hypothesis 6*).

#### Model based analysis

The following analyses are designed to test for differences in game optimality between pretend and non-pretend games, and for a relationship between decision difficulty and click latency. To do so, we approximate optimal behaviour by calculating the posterior probability that a ship is hiding in each cell, given available information. Critically, in modelling pretend games we do not treat hints as part of this available information, because an optimal player should ignore this information in choosing where to click next. Given this posterior map, a rational player should choose cells that have a high posterior probability of hiding a ship. This behaviour is not strictly optimal, but approximates optimal behaviour in most cases [@audinot2014optimal, Section 3.3].

To quantify optimality, before each cell selection we will compute the posterior probability that a ship is hidden behind each of the ‘unknown’ cells. Then, we will rank cells from highest to lowest according to their posterior probability and record the rank of the chosen cell.

*Hypothesis 7 (game optimality):* We will test the null hypothesis that pretend games are as optimal as non-pretend games. Mean click optimality will be contrasted between the pretend and non-pretend games of each player.  Subject-wise differences will then be carried over to a t-test at the group level.

*Hypothesis 8 (entropy-RT relation):* The entropy of the posterior map will be extracted before each cell selection using the following formula: $H(p)=\sum_{i=1}^{N} p_{i}\times log p_{i}$. We will test for within-subject associations between click latency and the linear and quadratic expansions of posterior entropy in pretend and in non-pretend games by fitting a multiple regression model to the data of each subject. We will contrast subject-wise beta values for pretend and non-pretend games to test the null hypothesis that the association between posterior entropy and click latency is the same in both conditions.

#### Judge accuracy

*Hypothesis 9 (judge accuracy):* Judge accuracy will be quantified as the number of correct guesses in the judge block, divided by 5. This quantity will be extracted separately for each subject and contrasted against chance (50%) at the group level.

*Hypothesis 10 (judge-pretend relation):* Pretense quality will be quantified as the number of pretend games that were misclassified by other participants as non-pretend games, divided by the number of pretend games that were presented to other participants. Due to randomization, this denominator is different for different participants. We will test for a correlation between pretense quality and judge accuracy at the group level. To maximize statistical power, this hypothesis will be tested on the pooled data from both groups of participants.

#### Half games

Data from half games will be subjected to exploratory analysis, focusing on the spatial distribution of clicks in pretend and non-pretend games. For each board (A or B) and condition (pretend or non-pretend), we will produce a heat map of click probabilities. This heat map will be compared against the true posterior map (see Fig. \@ref(fig:halfgames)). Depending on the consistency in cell selections among different players, we will perform this analysis not only for the first click, but also for subsequent clicks.

## Power calculations and sample size justification

With approximately `r N/2` participants in each group (allocation to group will be determined pseudorandomly), we will have 95% statistical power to detect an effect size of `r pwr.t.test(n=N/2,sig.level=0.05,power=0.95,type='paired')$d%>%printnum()` standard deviations in a repeated measures t-test (Hypotheses 1-9). In case of a conflict between the results of the two groups, a between-subjects comparison will be performed, using data from the first block only (pretend games from the pretend-first group and non-pretend games from the pretend-second group). For this between-subjects t-test, we will have 95% statistical power to detect an effect size of `r pwr.t2n.test(n1=N/2,n2=N/2,sig.level=0.05,power=0.95)$d%>%printnum()` standard deviations.

With `r N` participants in both groups combined, we will have 95% statistical power to detect a Pearson correlation of `r pwr.r.test(n=N,sig.level=0.05,power=0.95)$r%>%printnum()` (Hypothesis 10).

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2023-01-26 experiments/Battleships2/code/jspsych-6.1.0/plugins/jspsych-battleships.js

@@ -1,0 +1,298 @@
+jsPsych.plugins["Battleships"] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'Battleships',
+
+		parameters: {
+			grid: {
+				type: jsPsych.plugins.parameterType.INT,
+				pretty_name: "Grid",
+				default: [
+					['0','0','0','0','0'],
+					['A','A','A','0','0'],
+					['0','0','0','0','0'],
+					['0','0','0','B','0'],
+					['C','C','0','B','0']
+				],
+				description: "Grid. 0 means water, everything else is a ship."
+			},
+			cheat: {
+				type: jsPsych.plugins.parameterType.BOOL,
+				pretty_name:'cheat',
+				default: false,
+				description: 'When true, participants see the position of the ships.'
+			},
+			end_screen_time: {
+				type: jsPsych.plugins.parameterType.INT,
+				pretty_name:'end screen time',
+				default: 2000,
+				description: 'For how many milliseconds is the end screen presented?'
+			},
+		 text: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			 pretty_name:'text',
+			 default:
+`To win, you need to sink one 3-square submarine
+and two 2-square patrol boats.`,
+			 description: 'Text to display on top of grid.'
+		 },
+		 draw_attention_to_instructions_time: {
+				 type: jsPsych.plugins.parameterType.INT,
+				 pretty_name: 'draw attention to instructions time',
+				 default: 0,
+				 description: 'For how many milliseconds is will the instructions be flashing?'
+		 },
+		 already_clicked: {
+ 				type: jsPsych.plugins.parameterType.INT,
+ 				pretty_name: "Already clicked cells",
+ 				default: [],
+ 				description: "For half games. List of tuples of i,j that have been clicked. [] for fresh start."
+		 }
+		}
+	}
+
+	plugin.trial = function(display_element, trial) {
+
+		display_element.innerHTML = '';
+
+		//open a p5 sketch
+		let sketch = function(p) {
+
+		const du = p.min([window.innerWidth, window.innerHeight, 500])*7/10 //drawing unit
+		const left_margin = p.round((window.innerWidth-du)/2);
[…]

### 2023-01-26 experiments/Battleships2/protocolFolder/docs/board_state.R

@@ -1,0 +1,49 @@
+all_possible_boards <- read.csv('../analysis/flat_boards.csv',header=FALSE);
+
+create_board_states <- function(flat_positions,hit) {
+  
+  board_state = rep(NA,25)
+  board_states = c();
+  for (p in seq_along(flat_positions)) {
+    board_states = c(board_states, paste(board_state,collapse=','));
+    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
+  }
+  
+  return(board_states)
+};
+
+get_likelihood <- function(board_state) {
+  
+  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
+  
+  revealed_positions <- which(!is.na(board_state));
+  
+  boards_to_remove = c();
+  
+  for (position in revealed_positions) {
+    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
+  }
+  
+  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
+  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
+  likelihood[revealed_positions]=NA;
+  
+  return(paste(likelihood,collapse=','))
+  
+};
+
+
+
+E1.click_log_with_boards <- E1.click_log %>%
+  select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>% 
+  mutate(flat_position = 1+i*5+j) %>%
+  group_by(subj_id,genuine_first,test_part,grid_number) %>%
+  arrange(click_number) %>%
+  summarise(board_state=create_board_states(flat_position,hit_bin),
+            flat_position=flat_position,
+            i=i,
+            j=j,
+            RT=RT,
+            hit_bin=hit_bin) %>%
+  rowwise()%>%
+  mutate(likelihood = get_likelihood(board_state));

### 2023-01-26 experiments/Battleships2/protocolFolder/experiment/jspsych-6.1.0/plugins/jspsych-battleships-replay-2players.js

@@ -1,0 +1,412 @@
+jsPsych.plugins["Battleships_replay_2players"] = (function() {
+
+    var plugin = {};
+
+    plugin.info = {
+        name: 'Battleships_replay_2players',
+
+        parameters: {
+            grid: {
+                type: jsPsych.plugins.parameterType.INT,
+                pretty_name: "Grid",
+                default: [
+                    ['0', '0', '0', '0', '0'],
+                    ['A', 'A', 'A', '0', '0'],
+                    ['0', '0', '0', '0', '0'],
+                    ['0', '0', '0', 'B', '0'],
+                    ['C', 'C', '0', 'B', '0']
+                ],
+                description: "Grid. 0 means water, everything else is a ship."
+            },
+            end_screen_time: {
+                type: jsPsych.plugins.parameterType.INT,
+                pretty_name: 'end screen time',
+                default: 2000,
+                description: 'For how many milliseconds is the end screen presented?'
+            },
+            text: {
+                type: jsPsych.plugins.parameterType.STRING,
+                pretty_name: 'text',
+                default: `To win, you need to sink one 3-square submarine
+and two 2-square patrol boats.`,
+                description: 'Text to display on top of grid.'
+            },
+            description: {
+                type: jsPsych.plugins.parameterType.STRING,
+                pretty_name: 'description',
+                default: {
+                    right: `Replaying player 2's game:`,
+                    left: `Replaying Player 1's game`
+                },
+                description: 'Description'
+            },
+            click_log: {
+                type: jsPsych.plugins.parameterType.OBJECT,
+                pretty_name: 'click_log',
+                default: {
+                    'right': {
+                        "i": [1, 2, 3, 1, 1, 2, 2, 3, 2, 3, 3, 4, 4, 4],
+                        "j": [0, 0, 0, 1, 2, 3, 4, 4, 2, 3, 2, 3, 1, 0],
+                        "t": [3559, 3906, 4284, 5152, 6111, 7086, 7696, 7993, 8532, 9145, 9580, 10129, 11173.5, 11552],
+                        "hit": ["A", "0", "0", "A", "A", "0", "0", "0", "0", "B", "0", "B", "C", "C"]
+                    },
+                    "left": {
+                        "i": [1, 2, 1, 2, 1, 0, 1, 3, 3, 4, 4, 4, 4],
[…]

### 2023-01-26 experiments/Battleships2/protocolFolder/experiment/jspsych-6.1.0/plugins/jspsych-battleships-replay.js

@@ -1,0 +1,238 @@
+jsPsych.plugins["Battleships_replay"] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'Battleships_replay',
+
+		parameters: {
+			grid: {
+				type: jsPsych.plugins.parameterType.INT,
+				pretty_name: "Grid",
+				default: [
+					['0','0','0','0','0'],
+					['A','A','A','0','0'],
+					['0','0','0','0','0'],
+					['0','0','0','B','0'],
+					['C','C','0','B','0']
+				],
+				description: "Grid. 0 means water, everything else is a ship."
+			},
+			cheat: {
+				type: jsPsych.plugins.parameterType.BOOL,
+				pretty_name:'cheat',
+				default: false,
+				description: 'When true, participants see the position of the ships.'
+			},
+			end_screen_time: {
+				type: jsPsych.plugins.parameterType.INT,
+				pretty_name:'end screen time',
+				default: 2000,
+				description: 'For how many milliseconds is the end screen presented?'
+			},
+		 text: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			 pretty_name:'text',
+			 default:
+`To win, you need to sink one 3-square submarine
+and two 2-square patrol boats.`,
+			 description: 'Text to display on top of grid.'
+		 },
+		 description: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			 pretty_name:'description',
+			 default:
+ `Replaying Matan's game:`,
+			 description: 'Description'
+		 },
+		 click_log: {
+			 type: jsPsych.plugins.parameterType.OBJECT,
+			 pretty_name:'click_log',
+			 default: {"i":[1,2,3,1,1,2,2,3,2,3,3,4,4,4],
+			 "j":[0,0,0,1,2,3,4,4,2,3,2,3,1,0],
+			 "t":[3559,3906,4284,5152,6111,7086,7696,7993,8532,9145,9580,10129,11173.5,11552],
+			 "hit":["A","0","0","A","A","0","0","0","0","B","0","B","C","C"]},
+			 description: 'Click log'
+		 }
+		}
+	}
+
+	plugin.trial = function(display_element, trial) {
+
+		display_element.innerHTML = '';
+
+		//open a p5 sketch
+		let sketch = function(p) {
+
+		const du = p.min([window.innerWidth, window.innerHeight, 500])*7/10 //drawing unit
[…]

### 2023-01-26 experiments/Hangman2/code/jspsych-6.1.0/plugins/jspsych-hangman.js
hangman experiment

@@ -1,0 +1,506 @@
+jsPsych.plugins["Hangman"] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'Hangman',
+
+		parameters: {
+
+			word: {
+				type: jsPsych.plugins.parameterType.STRING,
+				pretty_name: "Word",
+				default: "ZEBRA",
+				description: "The target word"
+			},
+
+			cheat: {
+				type: jsPsych.plugins.parameterType.BOOL,
+				pretty_name:'cheat',
+				default: true,
+				description: 'When true, participants see the position of the ships.'
+			},
+
+			end_screen_time: {
+				type: jsPsych.plugins.parameterType.INT,
+				pretty_name:'end screen time',
+				default: 2000,
+				description: 'For how many milliseconds is the end screen presented?'
+			},
+
+		 text: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			 pretty_name:'text',
+			 default:
+`Use your keyboard to reveal the letter with as few clicks as possible.`,
+			 description: 'Text to display on top of grid.'
+		 },
+
+		 category: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			pretty_name:'category',
+			default: `an animal`,
+			description: 'Category of word (as hint)'
+		},
+
+		 draw_attention_to_instructions_time: {
+				 type: jsPsych.plugins.parameterType.INT,
+				 pretty_name: 'draw attention to instructions time',
+				 default: 0,
+				 description: 'For how many milliseconds is will the instructions be flashing?'
+		 },
+
+		 asked: {
+ 				type: jsPsych.plugins.parameterType.STRING,
+ 				pretty_name: "Already asked_letters",
+ 				default: [],
+ 				description: "For half games. List of letters that are already revealed"
+		 },
+
+		 draw_attention_to_instructions_time: {
+				 type: jsPsych.plugins.parameterType.INT,
+				 pretty_name: 'draw attention to instructions time',
+				 default: 0,
+				 description: 'For how many milliseconds is will the instructions be flashing?'
+		 },
+		}
+	}
+
+	plugin.trial = function(display_element, trial) {
+
+		trial.word = trial.word.toUpperCase();
+
+		display_element.innerHTML = '';
+
+		//open a p5 sketch
+		let sketch = function(p) {
+
[…]

### 2023-01-26 experiments/Hangman2/protocolFolder2/docs/word_probabilities/makeBodyDF.R

@@ -1,0 +1,22 @@
+library(tidyverse)
+
+five_letter_bodyparts <- strsplit('belly, spine, skull, scalp, thigh, torso, thumb, wrist, blood, bones, cheek, elbow, spine, skull, scalp, thigh, torso, thumb, wrist, waist', 'penis',', ')[[1]] %>%tolower()%>%unique()
+five_letter_bodyparts_plural <- c('veins')
+four_letter_bodyparts <- strsplit('Neck, Back, Head, Foot, Hand, Nose, Bone, Iris, Brow, Face, Chin, Skin, Lung, Hair, Vein, Butt, Anus, Lung, Shin, Palm, Calf, Nail, Vein, Lash, Brow', ', ')[[1]]%>%tolower()%>%unique()
+
+word = c(five_letter_bodyparts,four_letter_bodyparts,five_letter_bodyparts_plural)
+
+basic_list <- data.frame(word) %>% mutate(score=ifelse(word=='veins', 100,7))
+
+word_df <- read.csv('word_lists/bodypart_proto.csv')%>%
+  rbind(basic_list)%>%
+  group_by(word) %>%
+  summarise(score=min(score)) %>%
+  mutate(invscore=1/score) %>%
+  ungroup()%>%
+  mutate(prior=invscore/sum(invscore)) %>%
+  dplyr::select(word, score, prior)
+
+
+word_df %>%
+  write.csv('word_lists/bodypart.csv')

### 2023-01-26 experiments/Hangman2/protocolFolder2/docs/word_probabilities/makeCelebDF.R

@@ -1,0 +1,43 @@
+library(tidyverse)
+library(httr)
+library(jsonlite)
+library(stringr)
+
+celebs <- read.csv('word_lists/celebs.csv')%>%
+  pull(name) %>%
+  unique()
+
+# using the Wikimedia API to get page view counts
+getWikiPageViewCounts <- function(name,first_date,last_date) {
+
+  #convert the name to the Right_Format
+  formatted_name <- str_to_title(name) %>%
+    str_replace_all(' ', '_')
+
+  query <- paste('https://wikimedia.org/api/rest_v1/metrics/pageviews/per-article/en.wikipedia.org/all-access/all-agents/',
+                formatted_name,
+                '/monthly/',
+                as.character(first_date),
+                '/',
+                as.character(last_date),
+                sep='');
+
+  res <- httr::GET(query);
+
+  data <- fromJSON(rawToChar(res$content))
+  return(data$items$views%>%sum())
+
+}
+
+word_df <- data.frame(word = tolower(celebs)) %>%
+  rowwise()%>%
+  mutate(views= getWikiPageViewCounts(word,20210101,20211231))%>%
+  ungroup() %>%
+  mutate(prior=views/sum(views)) %>%
+  arrange(desc(prior)) %>%
+  filter(prior>0)
+
+
+word_df %>%
+  write.csv('word_lists/celebrities.csv')
+

### 2023-01-26 experiments/Hangman2/protocolFolder2/docs/word_probabilities/makeFruitDF.R

@@ -1,0 +1,343 @@
+library(tidyverse)
+
+fruit_from_wiki <- "*[[Abiu]]
+*[[Acai]]
+*[[Acerola]]
+*[[Akebi]]
+*[[Ackee]]
+*[[African Cherry Orange]]
+*[[American Mayapple]]
+*[[Apple]]
+*[[Apricot]]
+*[[Araza]]
+*[[Avocado]]
+*[[Banana]]
+*[[Bilberry]]
+*[[Blackberry]]
+*[[Blackcurrant]]
+*[[Black sapote]]
+*[[Blueberry]]
+*[[Boysenberry]]
+*[[Breadfruit]]
+*[[Buddhas hand]]
+*[[Cactus pear]]
+*[[Canistel]]
+*[[Cashew]]
+*[[Cempedak]]
+*[[Cherimoya]]
+*[[Cherry]]
+*[[Chico fruit]]
+*[[Cloudberry]]
+*[[Coco De Mer]]
+*[[Coconut]]
+*[[Crab apple]]
+*[[Cranberry]]
+*[[Currant]]
+*[[Damson]]
+*[[Date palm|Date]]
+*[[Dragonfruit]]
+*[[Pitaya]]
+*[[Durian]]
+*[[Egg Fruit]]
+*[[Elderberry]]
+*[[Feijoa]]
+*[[Fig]]
+*[[Finger Lime]]
+*[[Caviar Lime]]
+*[[Goji berry]]
+*[[Gooseberry]]
+*[[Grape]]
+*[[Raisin]]
+*[[Grapefruit]]
+*[[Grewia asiatica (phalsa or falsa)]]
+*[[Guava]]
+*[[Hala Fruit]]
+*[[Honeyberry]]
+*[[Huckleberry]]
+*[[Jabuticaba]]
+*[[Jackfruit]]
+*[[Jambul]]
+*[[Japanese plum]]
+*[[Jostaberry]]
+*[[Jujube]]
+*[[Juniper berry]]
+*[[Kaffir Lime]]
+*[[Kiwano]]
+*[[Kiwifruit]]
+*[[Kumquat]]
+*[[Lemon]]
+*[[Lime]]
+*[[Loganberry]]
+*[[Longan]]
+*[[Loquat]]
+*[[Lulo]]
+*[[Lychee]]
+*[[Magellan Barberry]]
+*[[Mamey Apple]]
+*[[Mamey Sapote]]
+*[[Mango]]
[…]

### 2023-01-26 experiments/Hangman2/protocolFolder2/docs/word_probabilities/makeNumberWordsDF.R

@@ -1,0 +1,52 @@
+library(tidyverse)
+
+numbers = c('zero')
+
+one_to_nine = c('one',
+                'two',
+                'three',
+                'four',
+                'five',
+                'six',
+                'seven',
+                'eight',
+                'nine');
+
+numbers = c(numbers, one_to_nine);
+
+numbers = c(numbers, c(
+  'ten',
+  'eleven',
+  'twelve',
+  'thirteen',
+  'fourteen',
+  'fifteen',
+  'sixteen',
+  'seventeen',
+  'eighteen',
+  'nineteen'
+))
+
+tens = c('twenty',
+         'thirty',
+         'forty',
+         'fifty',
+         'sixty',
+         'seventy',
+         'eighty',
+         'ninety')
+
+for (ten in tens) {
+  numbers = c(numbers, ten)
+  for (digit in one_to_nine) {
+    numbers = c(numbers, paste(ten,digit, sep= ' '))
+  }
+}
+
+numbers = c(numbers,'hundred','thousand', 'million', 'billion', 'dozen', 'googol')
+
+word_df <- data.frame(word = numbers) %>%
+  mutate(prior=1/length(numbers))
+
+word_df %>%
+  write.csv('word_lists/numbers.csv')

### 2023-01-26 experiments/Hangman2/protocolFolder2/docs/word_probabilities/makeStateDF.R

@@ -1,0 +1,11 @@
+library(tidyverse)
+
+state_list <- read.csv('word_lists/states.csv')%>%
+  pull(state) %>%
+  unique()
+
+word_df <- data.frame(word = tolower(state_list)) %>%
+  mutate(prior=1/length(state_list))
+
+word_df %>%
+  write.csv('word_lists/USstates.csv')

### 2023-01-26 experiments/Hangman2/protocolFolder2/experiment/jspsych-6.1.0/plugins/jspsych-hangman-replay.js

@@ -1,0 +1,548 @@
+jsPsych.plugins["Hangman_replay"] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'Hangman_replay',
+
+		parameters: {
+
+			word: {
+				type: jsPsych.plugins.parameterType.STRING,
+				pretty_name: "Word",
+				default: "ZEBRA",
+				description: "The target word"
+			},
+
+			cheat: {
+				type: jsPsych.plugins.parameterType.BOOL,
+				pretty_name:'cheat',
+				default: false,
+				description: 'When true, participants see the position of the ships.'
+			},
+
+			end_screen_time: {
+				type: jsPsych.plugins.parameterType.INT,
+				pretty_name:'end screen time',
+				default: 2000,
+				description: 'For how many milliseconds is the end screen presented?'
+			},
+
+		 text: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			 pretty_name:'text',
+			 default:
+`Use your keyboard to reveal the letter with as few clicks as possible.`,
+			 description: 'Text to display on top of grid.'
+		 },
+
+		 category: {
+			 type: jsPsych.plugins.parameterType.STRING,
+			pretty_name:'category',
+			default: `an animal`,
+			description: 'Category of word (as hint)'
+		},
+
+		 draw_attention_to_instructions_time: {
+				 type: jsPsych.plugins.parameterType.INT,
+				 pretty_name: 'draw attention to instructions time',
+				 default: 0,
+				 description: 'For how many milliseconds is will the instructions be flashing?'
+		 },
+
+		 asked: {
+ 				type: jsPsych.plugins.parameterType.STRING,
+ 				pretty_name: "Already asked_letters",
+ 				default: [],
+ 				description: "For half games. List of letters that are already revealed"
+		 },
+		 click_log: {
+			 type: jsPsych.plugins.parameterType.OBJECT,
+			 pretty_name:'click_log',
[…]

### 2023-01-26 experiments/Hangman2/protocolFolder2/experiment/jspsych-6.1.0/plugins/jspsych-p5text.js

@@ -1,0 +1,106 @@
+jsPsych.plugins["p5Text"] = (function() {
+
+	var plugin = {};
+
+	plugin.info = {
+		name: 'p5Text',
+		parameters: {
+			title: {
+				type: jsPsych.plugins.parameterType.STRING,
+				default: '',
+			},
+			text: {
+				type: jsPsych.plugins.parameterType.STRING,
+				default: 'text',
+			},
+			footer: {
+				type: jsPsych.plugins.parameterType.STRING,
+				default: 'press Space to continue',
+			},
+			draw_function: {
+				type: jsPsych.plugins.parameterType.FUNCTION,
+				default: function(p, trial) {return}
+			},
+			key_function: {
+				type: jsPsych.plugins.parameterType.FUNCTION,
+				default: function(p, trial) {return}
+			},
+			tite: {
+				type: jsPsych.plugins.parameterType.STRING,
+				default: ''
+			}
+		}
+	}
+
+	plugin.trial = function(display_element, trial) {
+
+		display_element.innerHTML = ''
+
+		//open a p5 sketch
+		let sketch = function(p) {
+
+		//sketch setup
+		p.setup = function() {
+			p.createCanvas(window.innerWidth, window.innerHeight);
+			p.fill(255); //white
+			p.strokeWeight(0)
+			p.background(128); //gray
+			p.textFont('Quicksand');
+      p.rectMode(p.CENTER)
+		}
+
+		//organize everything in one sequence
+		p.draw = function() {
+
+      p.textAlign(p.LEFT,p.CENTER)
+			p.background(128); //gray
+			p.textSize(24)
+			p.text(trial.text, p.width/2, p.height*(5/12), 600, 300);
+
+			p.push()
+			p.textAlign(p.CENTER, p.CENTER)
+			// p.textStyle(p.BOLD)
+			p.textSize(50)
+			p.text(trial.title,p.width/2, p.height/6, 600, 100)
+			// p.pop()
+			//
+			// p.push()
+			p.textSize(18)
+			p.textStyle(p.NORMAL)
+			p.text(trial.footer, p.width/2, p.height*(5/6), 600, 60);
+			p.strokeWeight(1)
+			p.stroke(255)
+			p.line(p.width/2-150,p.height*(3/4),p.width/2+150,p.height*(3/4))
+			p.pop()
+
+			trial.draw_function(p, trial)
+		}
+
+		p.keyPressed = function() {
+
[…]

### 2023-01-26 experiments/Hangman2/protocolFolder2/experiment/jspsych-6.1.0/plugins/jspsych-type-this.js

@@ -1,0 +1,80 @@
+/**
+ * jspsych-type-this
+ * a jspsych plugin to get a subject to type a word, for the hangman games
+ * based on the jspsych survey text plugin. Since I don't fully understand
+ * JQuery, I'm sure there are many redundancies and inefficient code lines here.
+ * It's good enough for my purpose though!
+ *
+ * Matan Mazor
+ *
+ * documentation: docs.jspsych.org
+ *
+ */
+
+
+jsPsych.plugins['type-this'] = (function() {
+
+  var plugin = {};
+
+  plugin.info = {
+    name: 'type-this',
+    description: '',
+    parameters: {
+          word: {
+            type: jsPsych.plugins.parameterType.STRING,
+            pretty_name: 'Word',
+            default: 'banana',
+            description: 'The word the subject is asked to type'
+          }
+        }
+      }
+
+  plugin.trial = function(display_element, trial) {
+
+    var responses = [];
+    var html = `<div id="jspsych-survey-text-preamble" class="jspsych-survey-text-preamble"><p>The next word is ${trial.word.toUpperCase()}, but your task is to pretend you don't know that.</p>
+      <p>Type ${trial.word.toUpperCase()} to confirm:</p></div>
+      <form id="jspsych-survey-text-form">
+      <div id="jspsych-survey-text-question" class="jspsych-survey-text-question" style="margin: 2em 0em;">
+      <input type="text" id="input-0"  name="#jspsych-survey-text-response-0" size="${trial.word.length}" autofocus required placeholder=""></input>
+      </div>
+      <input type="submit" id="jspsych-survey-text-next" class="jspsych-btn jspsych-survey-text" value="pretend!"></input>
+      </form>`
+
+    display_element.innerHTML = html;
+
+    // backup in case autofocus doesn't work
+    display_element.querySelector('#input-0').focus();
+
+    display_element.querySelector('#jspsych-survey-text-form').addEventListener('submit', function(e) {
+      e.preventDefault();
+      // measure response time
+      var endTime = performance.now();
+      var response_time = endTime - startTime;
+      var val = document.querySelector('#jspsych-survey-text-question').querySelector('textarea, input').value;
+      responses.push(val);
+
+      // if the subject types in the correct word:
+      if (val.toUpperCase()==trial.word.toUpperCase()) {
+
+        // save data
+        var trialdata = {
+          "rt": response_time,
+          "response": val,
+          "responses": responses
+        };
+
[…]

### 2023-01-26 docs/preregisteredMethodsAndResults.Rmd
preregistered analysis

@@
 
-bibliography      : "r-references.bib"
-
 floatsintext      : no
@@
 classoption       : "man"
-output            : papaja::apa6_word
+output            : bookdown::html_book
+bibliography: references.bib
 ---
-
-```{r setup, include = FALSE}
-library("papaja")
-r_refs("r-references.bib")
-```
 
@@
   'jsonlite', # parsing data from sort_trial
-  'caret' #for cross validation
+  'caret', #for cross validation
+  'ggrepel' #for word scatterplots
 ), '2022-12-01')
@@
 ## Methods
+
 We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
@@
 
-A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb)
+A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb).
+The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019] (protocol sum: 60c270410375e8a192468fc1a0e9c93da60d5e203eb2760b621a8631a26f4c5c; [link to relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/44ef964c0d7d4a8f3b76ec52316a8e624c41a047/experiments/Battleships2/code/index.html#L865-L876)).
 
@@
 ### Data analysis
+
 <!-- We used `r cite_r("r-references.bib")` for all our analyses. -->
@@
 
-*Hypothesis 1 (game duration)*: Pretend games were longer than non-pretend games by `r (E2.duration_log %>%pull(diff)%>%mean()/1000)%>%round()` seconds on average.
+### Hypothesis 1 (game duration)
+
+Pretend games were longer than non-pretend games by `r (E2.duration_log %>%pull(diff)%>%median()/1000)%>%round()` seconds on average.
 This was true in the pretend-first group (`r E2.duration_log %>% filter(!genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), and in the pretend-second groups (`r E2.duration_log %>% filter(genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
@@
 
-*Hypothesis 2 (first click latency)*: The first click in pretend games took longer to execute by `r E2.first_click_log %>%pull(diff)%>%mean()%>%round()` ms.
+### Hypothesis 2 (first click latency)
+
+The first click in pretend games took longer to execute by `r E2.first_click_log %>%pull(diff)%>%median()%>%round()` ms.
 This was true in the pretend-first group (`r E2.first_click_log %>% filter(!genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), but not in the pretend-second groups (`r E2.first_click_log %>% filter(genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
@@
 
-*Hypothesis 3 (number of clicks)*:
+### Hypothesis 3 (number of clicks)
 
@@
 We assumed that these participants did not follow the task instructions, and excluded them from all analyses.
-
-In our instructions to pretenders, we asked them to pretend not to know where the ships were, not only in *where* they chose to click, but also in *when* they clicked.
-In the following set of analyses, we identified patterns in click latency in non-pretend games, and asked whether the same patterns are also observed in pretend games.
-We focused on the effects of hitting versus missing a ship, decision uncertainty, and ship completion.
 
@@
 p<- E2.RT_by_hit_lags_long %>% 
-  filter(position<2 & position > -1)%>%
+  filter(position<2 & position > -2)%>%
   group_by(test_part,position) %>%
@@
   scale_color_manual(values=c("#69b3a2", "#404080")) +
-  scale_x_continuous(breaks=0:1,labels=c('hit on current trial', 'hit on previous trial'))+
+  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
   geom_abline(slope=0,intercept=0,linetype=3)+
@@
 
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.pdf',p,width=7)
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.png',p,width=4, height=3.5, dpi=300)
-
-```
-
-*Hypothesis 4 (click latency by outcome)*: In non-pretend games, trials that resulted in a hit were faster by `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%mean()%>%round()%>%abs()` than those that resulted in a miss (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`) will be classified based on whether they resulted in hitting a ship (*hit*) or not (*miss*).
-We will then contrast median click latency between hits and misses separately for pretend and non-pretend games.
-We will test for a difference in trial latency as a function of outcome on pretend games, on non-pretend games, and for an interaction effect between click outcome (hit or miss) and condition (pretend or non-pretend) on click latency.
-The same analysis will then be repeated by classifying trials on the basis of whether the previous trial resulted in a hit or a miss (*Hypothesis 5*), and whether the next trial will result in a hit or a miss (*Hypothesis 6*).
+ggsave('figures/E2_RT_by_lag.pdf',p,width=7)
+ggsave('figures/E2_RT_by_lag.png',p,width=4, height=3.5, dpi=300)
+
+```
+
+### Hypothesis 4 (click latency by outcome)
+
+In non-pretend games, cell selections that resulted in a hit were faster by `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median()%>%round()%>%abs()` ms.
+than those that resulted in a miss (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+In pretend games, cell selections that resulted in a hit were faster by `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median()%>%round()%>%abs()` ms.
+than those that resulted in a miss (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='pretend' & genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effect of hitting versus missing a ship was significantly stronger in pretend games (pretend-first group: `r E2.RT_by_hit_lags_long %>% filter(position==0 & !genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags_long %>% filter(position==0 & genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+### Hypothesis 5 (click latency by previous outcome)
+
+In non-pretend games, cell selections that followed a hit were slower by `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()%>%abs()` ms.
+than those that followed a miss (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+In pretend games, cell selections that followed in a hit were slower by `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()%>%abs()` ms.
+than those that followed a miss (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='pretend' & genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effect of following a hit versus a miss was significantly stronger in pretend games only in the pretend-second group (`r E2.RT_by_hit_lags_long %>% filter(position==1 & genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-first group: `r E2.RT_by_hit_lags_long %>% filter(position==1 & !genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Crucially, the effect was significantly stronger when focusing on the first blocks only (`r t.test(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(lag1_diff), E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(lag1_diff))%>%apa_print()%>%'$'(statistic)`).
+
+### Hypothesis 6 (click latency by next outcome)
+
+In non-pretend games, cell selections that preceded a hit were slower than those that preceded a miss only in the group that played normally first (pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(lead1_diff)%>%t.test()%>%apa_print()%>%'$'(full_result)`), but not in the group that pretended first (`r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(lead1_diff)%>%t.test()%>%apa_print()%>%'$'(full_result)`).
+
+In pretend games, the timing of cell selections did not covary with the outcomes of future cell selections (pretend-first group: `r E2.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(lead1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_hit_lags %>% filter(test_part=='pretend' & genuine_first)%>%pull(lead1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effect of the next outcome was significantly weaker in pretend games only in the pretend-second group (`r E2.RT_by_hit_lags_long %>% filter(position==-1 & genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`) but not in the pretend-first group (`r E2.RT_by_hit_lags_long %>% filter(position==-1 & !genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Crucially, the effect was significantly weaker when focusing on the first blocks only (`r t.test(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(lead1_diff), E2.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(lead1_diff))%>%apa_print()%>%'$'(statistic)`).
+
+### Model based analysis
+
+Pretend games were similar to non pretend games in the total number of clicks --- but were they also similar in *where* participants clicked?
+More specifically, did cell selections in pretend games make sense given the limited information those participants pretended to have?
+To ask this, we approximated optimal behavior by calculating the probability that a ship is hidden in each cell given available information, $p(ship(x_{i}))$, and the posterior probability that one should click on a square, assuming a uniform prior over cells $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$.
+Critically, in modeling pretend games we did not treat hints as part of this available information for extracting $p(ship(x_{i}))$, because an optimal player should ignore hints in choosing where to click next.
+Given this posterior map, a rational player should choose cells where $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3].
+To quantify optimality, before each cell selection we computed the posterior probability map for all 'unknown' cells.
+Then, we ranked cells from high to low according to their posterior probability and recorded the rank of the chosen cell: a lower rank indicating more optimal behavior.
+
+```{r battleship-p_click, echo = FALSE, cache=TRUE}
+
+# E2.click_log_with_boards <-E2.click_log %>%
+#   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>%
+#   mutate(flat_position = 1+i*5+j) %>%
+#   group_by(subj_id,genuine_first,test_part,grid_number) %>%
+#   arrange(click_number) %>%
+#   summarise(click_number = click_number,
+#             board_state=create_board_states(flat_position,hit_bin),
+#             flat_position=flat_position,
+#             i=i,
+#             j=j,
+#             RT=RT,
+#             hit_bin=hit_bin) %>%
+#   rowwise()%>%
+#   mutate(likelihood = get_likelihood(board_state),
+#          posterior = get_posterior(likelihood),
+#          p_click = get_p_click(posterior,flat_position),
+#          entropy = get_posterior_entropy(posterior),
+#          p_click_rank = get_p_click_rank(posterior,flat_position));
+
+E2.click_log_with_boards <- readRDS('../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')
+
+get_random_flat_position <- function(board_state) {
+
+  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
+  return(which(is.na(board_state))%>%sample(1))
+
+}
+
+get_optimal_flat_position <- function(posterior) {
+
+  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  return(which.max(posterior))
+
+}
+
+E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
+  mutate(flat_position = get_optimal_flat_position(posterior),
+         p_click = get_p_click(posterior,flat_position),
+         p_click_rank = get_p_click_rank(posterior,flat_position),
+         test_part = 'optimal')
+
+E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
+  mutate(flat_position = get_random_flat_position(board_state),
+         p_click = get_p_click(posterior,flat_position),
+         p_click_rank = get_p_click_rank(posterior,flat_position),
+         test_part = 'random')
+
+E2.mean_P_click_rank <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_optimal) %>%
+  rbind(E2.click_log_with_boards_random) %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(p_click_rank=mean(p_click_rank));
+
+E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_optimal) %>%
+  rbind(E2.click_log_with_boards_random) %>%
+  rowwise() %>%
+  filter(!hit_bin) %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+p<- E2.mean_P_click_rank %>%
+    filter(test_part != 'optimal') %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+ggsave('figures/E2_pclick_rank.pdf',p)
+ggsave('figures/E2_pclick_rank.png',p,width=5,height=2,dpi=300)
+
+p<- E2.mean_P_click_rank_misses_only %>%
+    filter(test_part != 'optimal') %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    scale_x_continuous(limits=c(0,15))+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+ggsave('figures/E2_pclick_rank_misses_only.pdf',p)
+ggsave('figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
+
+E2.num_irrational_clicks <- E2.click_log_with_boards %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(nic=sum(p_click==0)/5);
+
+p<- E2.num_irrational_clicks %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=nic,fill=test_part))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='mean number of irrational clicks per game', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    facet_wrap(~genuine_first, nrow=2)+
+    theme_classic();
+
+ggsave('figures/E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)
+
+ E2.RT_p_click_cor <- E2.click_log_with_boards %>%
+     filter(click_number<9 & click_number>1) %>%
+     group_by(subj_id,test_part, genuine_first) %>%
+     summarise(r=cor(RT,p_click,method='spearman'));
+
+ E2.RT_entropy_cor <- E2.click_log_with_boards %>%
+     filter(click_number<9 & click_number>1) %>%
+     group_by(subj_id,test_part, genuine_first) %>%
+     summarise(r=cor(entropy,RT,method='spearman'))
+
+```
+
+### Hypothesis 7 (game optimality)
+
+The mean posterior rank of non-pretend cell selections was `r E2.mean_P_click_rank%>%filter(test_part=='nonpretend')%>%pull(p_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (`r E2.mean_P_click_rank%>%filter(test_part=='random')%>%pull(p_click_rank)%>%mean()%>%printnum()`, `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Pretend games were significantly less optimal than non-pretend games (`r E2.mean_P_click_rank%>%filter(test_part=='pretend')%>%pull(p_click_rank)%>%mean()%>%printnum()`; pretend-first group: `r E2.mean_P_click_rank%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.mean_P_click_rank%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), but still more optimal than those of a random agent (`r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+In other words, the optimality of pretend games relative to random cell selection was not merely due to the fact that pretenders clicked on ships more than expected by chance.
+Even when missing a ship, their cell selections made sense given the limited information they pretended to have.
+
+### Hypothesis 8 (entropy-RT relation)
+
+When playing Battleships, it is sometimes clear what the next cell selection should be, and sometimes more difficult to decide where to click next.
+To capture this notion of decision uncertainty, we calculated the entropy of the posterior distribution over cell selections $H(P)=-\sum_{i=1}^{k}P(x_i) \log P(x_{i})$, where $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$, and asked how this measure relates to decision latency, or the time taken to click on the next cell.
+$H(P)$ is high when players need to decide between multiple cells with a similar probability of hiding a ship, and low when there are only a few candidates with a high probability of hiding a ship.
+For every player and condition separately, we fitted a multiple linear regression to predict decision latency based on $H(P)$ and $H(P)^2$.
+The resulting coefficients were then subjected to a group-level inference.
+The first cell selection of each game was excluded from this analysis, because entropy was constant for the first click.
+
+```{r entropy-RT, echo=FALSE, cache=TRUE}
+
+E2.RT_by_entropy <- E2.click_log_with_boards %>%
+  filter(click_number>1) %>%
+  mutate(entropy_bin=round(entropy*2)/2) %>%
+  group_by(subj_id,test_part,click_number) %>%
+  mutate(corrected_RT = RT-mean(RT)) %>%
+  group_by(test_part,subj_id,entropy_bin) %>%
+  summarise(RT=median(RT),
+            corrected_RT = median(corrected_RT)) %>%
+  group_by(test_part,entropy_bin) %>%
+  summarise(median_RT=median(RT),
+            sem_RT=bootstrap_error(RT,N_perm),
+            median_corrected_RT=median(corrected_RT),
+            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))
+
+p <- E2.RT_by_entropy %>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
+  theme_classic()
+
+ggsave('figures/E2_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
+# ggsave('../figures/E2_RT_by_entropy_no_split.svg',p,width=5,height=2,dpi=300)
+
+p <- E2.RT_by_entropy %>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Entropy of posterior', y='median residual RT (ms)')+
+  theme_classic()
+
+ggsave('figures/E2_corrected_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
+
+E2.RT_by_entropy_coefs <- E2.click_log_with_boards %>%
+  filter(click_number>1) %>%
+  group_by(genuine_first,subj_id,test_part,grid_number) %>%
+  mutate(entropy_centered = entropy-mean(entropy),
+         entropy_squared = entropy_centered^2) %>%
+  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
+  group_by(genuine_first,subj_id,test_part,grid_number) %>%
+  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
+  mutate(tidys=list(broom::tidy(model))) %>%
+  unnest(tidys) %>%
+  dplyr::select(!c(std.error,statistic,p.value)) %>%
+  group_by(genuine_first,subj_id,test_part,term) %>%
+  summarise(estimate = mean(estimate))
+
+E2.RT_by_entropy_coefs_summary <- E2.RT_by_entropy_coefs %>%
+  group_by(test_part,term) %>%
+  summarise(mean_estimate = mean(estimate),
+            se_estimate = se(estimate))
+
+p <- E2.RT_by_entropy_coefs_summary %>%
+  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
+  filter(term!='(Intercept)') %>%
+  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_bar(stat='identity',color='black',position=position_dodge()) +
+  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
+  labs(fill=' ',x='term', y='coefficient')+
+  theme_classic()
+
+ggsave('figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
+
+E2.corrected_RT_by_entropy_coefs <- E2.click_log_with_boards %>%
+  filter(click_number>1) %>%
+  group_by(subj_id,test_part,click_number) %>%
+  mutate(corrected_RT = RT-mean(RT)) %>%
+  group_by(genuine_first,subj_id,test_part) %>%
+  mutate(entropy_centered = entropy-mean(entropy),
+         entropy_squared = entropy_centered^2) %>%
+  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
+  group_by(genuine_first,subj_id,test_part) %>%
+  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
+  mutate(tidys=list(broom::tidy(model))) %>%
+  unnest(tidys) %>%
+  dplyr::select(!c(std.error,statistic,p.value))
+
+E2.corrected_RT_by_entropy_coefs_summary <- E2.corrected_RT_by_entropy_coefs %>%
+  group_by(test_part,term) %>%
+  summarise(mean_estimate = mean(estimate),
+            se_estimate = se(estimate))
+
+p <- E2.corrected_RT_by_entropy_coefs_summary %>%
+  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
+  filter(term!='(Intercept)') %>%
+  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_bar(stat='identity',color='black',position=position_dodge()) +
+  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
+  labs(fill=' ',x='term', y='coefficient')+
+  theme_classic()
+
+ggsave('figures/E2_corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
+
+```
+
+In non-pretend games, we found no evidence for a linear relation between decision entropy and decision latency among participants who played normally in the first block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
+In contrast, a significantly negative linear modulation was observed in the group that pretended in the first and played normally in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
+
+In pretend games, this negative relation between $H(P)$ and click latency was significant in both groups (pretend-first group: `r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_centered' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`; pretend-second group: `r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_centered' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
+We found no significant difference between the magnitude of the linear effect in pretend and non-pretend games (pretend-first group: `r E2.RT_by_entropy_coefs%>%filter(!genuine_first & term=='entropy_centered')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_entropy_coefs%>%filter(genuine_first & term=='entropy_centered')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+Similarly, we found no evidence for a quadratic relation between decision entropy and decision latency in non-pretend games of participants who played normally in the first block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
+Again, a significantly negative quadratic modulation was observed in the non-pretend games of players who pretended in the first and played normally in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
+Like the linear effect, this negative quadratic relation was significant in the pretend games of both groups (pretend-first group: `r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`; pretend-second group: `r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
+This negative quadratic effect was stronger in pretend compared to non-pretend games (pretend-first group: `r E2.RT_by_entropy_coefs%>%filter(!genuine_first & term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_entropy_coefs%>%filter(genuine_first & term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+```{r judge}
+
+get_common_board = function(grid_number,test_part) {
+  pretend_grids = grid_number[which(test_part=='pretend')];
+  nonpretend_grids = grid_number[which(test_part=='nonpretend')];
+  common_board = intersect(pretend_grids,nonpretend_grids);
+  return(common_board)
+};
+
+common_boards <- E2.df %>%
+  filter(test_part=='pretend' | test_part=='nonpretend') %>%
+  dplyr::select(subj_id,grid_number,test_part) %>%
+  group_by(subj_id) %>%
+  summarise(common_board = get_common_board(grid_number,test_part))
+
+used_for_pretend <- function(s, g) {
+
+  return (E2.df%>%filter(test_part=='pretend' &
+                           subj_id==s & 
+                           grid_number == g) %>%
+            nrow()==1)
+}
+
+used_for_nonpretend <- function(s, g) {
+
+  return (E2.df%>%filter(test_part=='nonpretend' &
+                           subj_id==s & 
+                           grid_number == g) %>%
+            nrow()==1)
+}
+
+E2.judge_df <- E2.df %>%
+  filter(test_part=='judge' & !(cheat_player %in% E2.lucky_pretenders))%>%
+  dplyr::select(subj_id, 
+         grid_number, 
+         cheat_player,
+         genuine_player,
+         cheater,
+         noncheater,
+         decision,
+         replay_log,
+         choose_pretender,
+         correct)%>%
+  rowwise()%>%
+  mutate(choose_pretender=choose_pretender=='True',
+         used_for_pretend = used_for_pretend(subj_id,grid_number),
+         used_for_nonpretend = used_for_nonpretend(subj_id,grid_number),
+         common_board = used_for_pretend & used_for_nonpretend,
+         board_type = ifelse(common_board, 'common',
+                             ifelse(used_for_pretend, 'pretend',
+                                    ifelse(used_for_nonpretend,'nonpretend','new')))
+)
+
+p <- E2.judge_df %>%
+  group_by(subj_id) %>%
+  summarise(ncorrect=sum(correct)) %>%
+  ggplot(aes(x=ncorrect)) +
+  geom_bar(fill='black') +
+  theme_classic() +
+  labs(x='number of correct guesses (out of 5)',y='number of players') +
+  scale_x_continuous(breaks=c(0,1,2,3,4,5));
+
+ggsave('figures/E2_correct_guesses.png',p,width=5,height=3,dpi=300)
+
+p <- E2.judge_df %>%
+  mutate(choose_pretender = ifelse(choose_pretender,
+                                   'who had hints?', 
+                                   'who had no hints?')) %>% 
+  group_by(choose_pretender,subj_id) %>%
+  summarise(ncorrect=sum(correct)) %>%
+  ggplot(aes(x=ncorrect)) + 
+  geom_bar(fill='black') +
+  theme_classic() +
+  labs(x='number of correct guesses (out of 5)',y='number of players') + 
+  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
+  facet_wrap(~choose_pretender,nrow=2);
+
+ggsave('figures/E2_correct_guesses_by_choose_pretender.png',p,width=5,height=3,dpi=300);
+
+p <- E2.judge_df %>% 
+  group_by(board_type,subj_id) %>%
+  summarise(acc=mean(correct)) %>%
+  ggplot(aes(x=acc)) + 
+  geom_bar(fill='black') +
+  theme_classic() +
+  labs(x='Accuracy',y='number of players') + 
+  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
+  facet_wrap(~board_type,nrow=1);
+
+ggsave('figures/E2_acc_by_board_type.png',p,width=5,height=3,dpi=300)
+```
+
+### Hypothesis 9 (judge accuracy)
+
+Participants were not significantly above chance in judging which of the two presented games came from a pretender (mean accuracy: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%; t test against 50: `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%t.test(mu=50))$statistic`).
+
+For a subset of `r E2.judge_df%>%filter(common_board)%>%pull(correct)%>%length()` players, one of the five boards presented in the judging block was already presented both in the pretend and in the non-pretend blocks.
+Still, even this subset of participants who had just experienced a board both as pretenders and as non-pretenders were not significantly above chance in telling which of two other players had hints for this same board (mean accuracy: 55%, $p=.141$).
+
+```{r judge-pretend-cor, echo=FALSE}
+
+unidentifiable_ss <- E2.judge_df %>%
+  mutate(subj_num=substr(as.character(subj_id),1,3))%>%
+  group_by(subj_num)%>%
+  summarise(subj_id=subj_id, n=n())%>%
+  filter(n>5)%>%
+  pull(subj_id)%>%
+  unique();
+
+E2.judge_performance <- E2.judge_df %>%
+  filter(!(subj_id %in% unidentifiable_ss)) %>%
+  mutate(subj_num=as.numeric(substr(as.character(subj_id),1,3)))%>%
+  group_by(subj_num)%>%
+  summarise(judge_rate=sum(correct)/5);
+
+E2.pretend_performance <- E2.judge_df %>%
+  filter(!(subj_id %in% unidentifiable_ss)) %>%
+  mutate(subj_num=cheat_player)%>%
+  group_by(subj_num)%>%
+  summarise(ntrials=n(),
+            successful_pretend=ntrials-sum(correct),
+            pretend_rate = successful_pretend/ntrials);
+
+E2.performance_df <- merge(E2.judge_performance,E2.pretend_performance);
+
+p <- E2.performance_df %>%
+  ggplot(aes(x=judge_rate,y=pretend_rate)) +
+  geom_jitter() + 
+  labs(x='judge accuracy',y='pretense success rate') +
+  theme_classic()
+
+ggsave('../figures/E2_judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)
+
+min_njudge = seq(0,10);
+cors =c()
+ns = c()
+
+for (minn in min_njudge) {
+
+  temp_df = E2.performance_df %>%
+    filter(ntrials>minn)
+
+  cur_cor = cor(temp_df$pretend_rate,temp_df$judge_rate);
+  cur_n = temp_df%>%nrow()
+  cors = c(cors,cur_cor)
+  ns = c(ns, cur_n)
+}
+
+judge_pretend_correlation_df <- data.frame(min_njudge,cors,ns);
+
+```
+
+```{r judge-optimality-cor, echo=FALSE}
+E2.judge_summary <- E2.judge_df %>%
+  group_by(subj_id) %>%
+  summarise(judge = sum(correct));
+
+E2.optimality_summary <- E2.click_log_with_boards %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click_rank = mean(p_click_rank)) %>%
+  spread(test_part,p_click_rank)
+
+E2.judge_summary <- E2.judge_summary %>%
+  merge(E2.optimality_summary)
+```
+
+### Hypothesis 10 (judge-pretend relation)
+
+Pretense quality was quantified as the number of pretend games that were misclassified by other participants as non-pretend games, divided by the number of pretend games that were presented to other participants.
+Due to randomization, this denominator is different for different participants.
+We tested for a correlation between pretense quality and judge accuracy at the group level.
+To maximize statistical power, this hypothesis was tested on the pooled data from both groups of participants.
+We find no evidence for such a correlation (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate))$full_result`).
+Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend))$full_result`).
 
@@
 
+```{r hangman-load-data, warning=FALSE}
+
+E4.df <- read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch2.csv',na.strings=c(""," ","NA")) %>%
+  rbind(read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
+  rename(subj_id = participant_number) %>%
+  mutate(subj_id = factor(subj_id)) %>%
+  # the dash is breaking r
+  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
+  filter(PROLIFIC_PID != '5ec4a156bc5aac3819ac52f2' & #reported some of the letters not showing up on their screen
+        PROLIFIC_PID != '58211fc787f6b90001f13f9' & #reported some of the letters not showing up on their screen
+        PROLIFIC_PID != '6149256f6335b06ade3723e0' &  # due a PROLFIIC bug, participated twice
+        PROLIFIC_PID != '615ddab1e4f013092538b6c5')# due a PROLFIIC bug, participated twice
+
+E4.export1 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export1.csv');
+E4.export2 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export2.csv');
+E4.export3 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export3.csv');
+E4.export4 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export4.csv');
+E4.export5 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export5.csv');
+E4.export6 <- read.csv('..\\experiments\\Hangman2\\data\\prolific_export6.csv');
+
+E4.export <- E4.export1 %>%
+  rbind(E4.export2) %>%
+  rbind(E4.export3) %>%
+  rbind(E4.export4) %>%
+  rbind(E4.export5) %>%
+  rbind(E4.export6)
+
+E4.bonus <- read.csv('..\\experiments\\Hangman2\\data\\bonuses\\all_bonuses.csv')
+
+```
+
 ## Methods
+
 We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
@@
 
+A detailed pre-registration can be accessed at [osf.io/3thry](https://osf.io/3thry).
+The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019].
+Due to an error in the experiment code, time-locking took effect only from player number 221 (batch 3) and on (protocol sum: c4929c7fe33df1b7b52f15c789d98eab30a9cee09a8121807a3c59e28e7430a4; [relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/f98903f3f2d0e21800497dc65f8d5f823c910c88/experiments/Hangman2/code/index.html#L772-L783)).
+
 ### Participants
 
-### Material
+The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
+500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
+They were selected based on their acceptance rate (\>95%) and for being native English speakers.
+The entire experiment took approximately 20 minutes to complete.
+Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment of 1 USD that was awarded to `r E4.bonus$bonus%>%sum(na.rm=T)` players who earned 100 points or more.
 
@@
 
+```{r hangman-design, echo=FALSE, fig.cap="Experimental Design: players will play 12 games of the game Hangman in two conditions presented in two blocks of 6 games.
+The order of blocks will be randomized between players.
+A: In non-pretend games, players will start the game not knowing the identity of the hidden word, and try to reveal the word with as few letter guesses as possible.
+B: In pretend games, players will be told the identity of the word, and their task will be to play the game as if they do not have this information.
+Before pretending, players will verify they had read the hidden word by typing it.
+C: Lastly, in 5 judge trials players will observe a replay of the game of a previous player, and try to tell whether the player was pretending or not.", out.width = '100%'}
+knitr::include_graphics("figures/Hangman_design.png")
+```
+
+The first instructions screen informed participants that the experiment, based on the game Hangman, had three parts, and that the points they accumulate translate to a monetary bonus payment.
+They were then be presented with a leaderboard of previous players.
+Then, the rules of the game were presented:
+
+> In the following game, your task is to reveal a hidden word or phrase by guessing letters.
+What makes the game difficult is that you can't see the word; all you can see is a row of squares - a square for each letter.
+Use your mouse to make letter guesses.
+We will have five types of words: body parts, numbers, US states, fruit, and famous people.
+You will start each game with 15 points and lose one point for every guess of a letter that is not in the word.
+
+We then explained that "the words in this game are the kind of words that will be familiar to most English-speaking fifth-graders. We didn't pick any strange or particularly difficult words".
+
+Once they responded correctly to a multi-choice comprehension question ("the goal of the game is to...": "reveal the word with as few letter guesses as possible"), participants played a practice round, revealing the word PIGEON (see Fig.
+\@ref(fig:hangman-design)A).
+
+After the main instructions, comprehension question and practice round, participants completed one pretend and one non-pretend block, each followed by one half-game (see below for details).
+The order of pretend and non-pretend blocks was counterbalanced between participants.
+Each block comprised five games played with five out of ten different words, and one half-game.
+The allocation of words to conditions was randomized between participants, with the constraint that both pretend and non-pretend blocks included exactly one word from each category.
+The order of words within a block was randomized, except for the half-game, which was always delivered at the end.
+
+The ten words included two number words (ELEVEN, NINETY SIX), two famous people (DALAI LAMA, TAYLOR SWIFT), two fruits (STRAWBERRY, LEMON), two body parts (TOOTH, HEAD), and two US states (MONTANA, IOWA).
+
+### Non-pretend games
+
+In non-pretend games, participants revealed a hidden word with as few letter guesses as possible.
+An online counter of the number of points was displayed on the screen, deducting one point for every guess of a letter that is not in the target word.
+After each game, feedback was given about the number of points obtained.
+
+After completing the five games, participants performed one half-game (see below for details).
+
+### Pretend games
+
+Participants were given the following instructions:
+
+> In the next part of the experiment, you'll play 6 games where you reveal a hidden word by guessing letters.
+
+> However, this time your goal is different.
+
+> In this round, we're going to tell you the word in advance, but **we want you to act like you don't know this information**.
+
+> To see how good you are at this, we're going to compare your games to the games of people who played normally, without knowing what the word was, and see how similar they are.
+We will measure which letters you click and the timing of your guesses; if your clicks look similar to people who played like normal (trying to reveal the word with as few guesses as possible, but without any hints), you'll get bonus points.
+But if your games look different, you won't get these bonus points.
+Your number of clicks in this part will not affect your bonus.
+Only your ability to play like you didn't see the word in advance.
+
+After one practice round, pretending not to know that the hidden word is PIGEON, and one comprehension question ("In this part of the experiment my goal is to...": "play the game as if I don't know what the word is so that I look like someone who had no hints"), participants played five pretend games.
+Each game was preceded by a short message informing subjects about the identity of the target word.
+To start pretending, players were asked to type in the target word on their keyboard.
+After pretending, we reminded players that a game that looks similar to the games of participants who had no hints will be awarded 10 bonus points.
+
+After completing the five games, participants performed one half-game (see below for details).
+
+### Half games
+
+In order to directly compare participants' pretend and non-pretend games for identical belief states (true or pretended knowledge about the identity of the word), we asked participants to also complete one pretend and one non-pretend game, given a partly finished game with some letters already guessed (they were told that the computer made these guesses).
+The two half-game words were one fruit: PAPAYA or BANANA, with guessed letters [A, E, I, O, M, T], and one body part: HAND, or HAIR with guessed letters (A, E, O, M, T, H, P).
+The assignment of category (fruit or body part) to condition (pretend and non-pretend), as well as the identity of the target word within each category (e.g., PAPAYA or BANANA), was randomized between participants.
+
+Instructions for the non-pretend half-game were:
+
+> For the next game, the computer chose the first letters for you; you can take over from where it left off.
+Your challenge is to complete the game.
+Just like in the previous games, here also you will lose one point for each letter that you guess and is not in the word.
+
+Instructions for the pretend half-game were:
+
+> For the next game, the computer chose the first letters for you; you can take over from where it left off.
+Just like in the previous games, here also you will know what the word is, but your bonus points will depend on your ability to play as if you didn't know the word.
+
+### Judge trials
+
+In the final part of the experiment, participants observed five games of previous players and determined who had hints and who didn't. Instructions for this part were:
+
+> In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of the players were shown the word (but acted like they weren't).
+We will show you 5 replays of the games of previous players.
+Your task is to decide whether they played normally or pretended.
+For each game that you get right, you will receive 10 points.
+Good luck!
+
+Then, on each judge trial, one game of a previous player was replayed in real time, with the target word presented above.
+For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
+For both pretend and non-pretend games, only games shorter than 1.5 minutes were presented.
+Judge participants indicated their decision by pressing the P and N keys on their keyboard.
+After making a decision, participants were informed whether they received the 10 points.
+Whenever a pretend game was classified as a non-pretend game, they were informed that the pretender received these 10 points instead of them.
+
+Lastly, participants were asked the following debrief questions:
+
+> Did you have a strategy that you used for pretending you did not see the word?
+What was most difficult about pretending?
+How about telling between players who pretenders and who played for real - did you have a strategy for that?
+
+And:
+
+> We would appreciate it if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.
+
 ### Data analysis
+
 We used `r cite_r("r-references.bib")` for all our analyses.
@@
 ## Results
+
+```{r hangman-format-data}
+
+E4.click_df <- E4.df %>%
+  dplyr::select(subj_id, 
+         test_part, 
+         category,
+         word,
+         num_clicks,
+         click_log,
+         genuine_first,
+         trial_type) %>%
+  rowwise()%>%
+  mutate(num_hits = strsplit(gsub(' ','',word),split='')[[1]]%>%unique()%>%length(),
+         genuine_first= genuine_first=='True',
+         num_misses = num_clicks-num_hits,
+         click_log = gsub("\'","\"", click_log),
+         click_log = gsub("None","null", click_log),
+         word = factor(word,levels=c(
+           'eleven',
+           'ninety six',
+           'dalai lama',
+           'taylor swift',
+           'strawberry',
+           'lemon',
+           'tooth',
+           'head',
+           'iowa',
+           'montana'
+         ))) %>%
+  filter(test_part=='pretend' | test_part=='nonpretend') %>%
+  filter(trial_type=='Hangman')
+
+E4.median_clicks <- E4.click_df %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(num_clicks=median(num_clicks)) %>%
+  spread(test_part,num_clicks);
+
+E4.median_misses <- E4.click_df %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(num_misses=median(num_misses)) %>%
+  spread(test_part,num_misses);
+
+E4.lucky_pretenders <- E4.median_misses%>%filter(pretend==0)%>%pull(subj_id)%>%unique();
+E4.lucky_nonpretenders <- E4.median_misses%>%filter(nonpretend==0)%>%pull(subj_id)%>%unique()
+
+# 
+# median_simulated = num_clicks%>%median()
+p<- E4.click_df %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=num_misses,fill=test_part))+
+    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='median number of misses', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic()+
+    # geom_vline(xintercept= median_simulated)+
+    facet_wrap(~genuine_first, nrow=2);
+
+ggsave('figures/Hangman2/num_misses.png',p,width=5,height=3,dpi=300)
+
+E4.misses_per_word <- E4.click_df %>%
+  group_by(test_part,word) %>%
+  summarise(num_misses=median(num_misses)) %>%
+  spread(test_part,num_misses)
+
+p <- E4.misses_per_word %>%
+  ggplot(aes(x=nonpretend,y=pretend, label=word)) +
+  geom_abline(intercept=0,slope=1) +
+  geom_point()+
+  geom_text_repel()+
+  coord_fixed(ratio=1)+
+  scale_x_continuous(limits=c(0,5.5), breaks=seq(0,5))+
+  scale_y_continuous(limits=c(0,5.5), breaks=seq(0,5)) +
+  theme(legend.position = 'none')+
+  theme_bw() +
+  labs(title='number of misses')
+
+ggsave('figures/hangman2/num_misses_per_word.png',p,width=5,height=5,dpi=300)
+
+```
+
+```{r hangman-parse-click-log, echo-FALSE}
+
+E4.click_log <- data.frame(matrix(ncol=11,nrow=0, 
+                               dimnames=list(NULL, 
+                                             c("subj_id",
+                                              "test_part", 
+                                              "word",
+                                              "category",
+                                              "num_clicks",
+                                              "num_misses",
+                                              "letter",
+                                              "hit",
+                                              "t",
+                                              "click_number",
+                                              "genuine_first"))))
+
+for (row in 1:nrow(E4.click_df)) {
+
+    subject_click_log <- data.frame(fromJSON(E4.click_df[row, ]$click_log)) %>%
+    mutate(
+      letter = lead(letter,1),
+      t = lead(t,1),
+      hit = lead(hit,1),
+      click_number = 1:n(),
+      subj_id = E4.click_df[row, ]$subj_id,
+      test_part = E4.click_df[row, ]$test_part,
+      word = E4.click_df[row, ]$word,
+      category = E4.click_df[row, ]$category,
+      num_clicks = E4.click_df[row, ]$num_clicks,
+      num_misses = E4.click_df[row, ]$num_misses,
+      genuine_first = E4.click_df[row, ]$genuine_first
+    )%>%
+      filter(click_number<=num_clicks)
+
+    E4.click_log <- rbind(E4.click_log, subject_click_log);
+}
+
+E4.click_log <- E4.click_log %>%
+  relocate(subj_id, .before = letter) %>%
+  relocate(genuine_first, .before = letter) %>% 
+  relocate(test_part, .before=letter) %>% 
+  relocate(word, .before=letter) %>% 
+  relocate(category, .before=letter) %>% 
+  relocate(click_number, .before=letter) %>%
+  group_by(subj_id,test_part,word) %>%
+  mutate(RT=t-lag(t,default=0)) %>%
+  group_by(subj_id,word) %>%
+  rowwise()
+# %>%
+#   filter(!(subj_id %in% E1.lucky_pretenders))
+
+```
+
+```{r hangman-total_time, echo=FALSE}
+
+E4.total_time_df <- E4.click_log %>%
+ filter(click_number==num_clicks)%>%
+  mutate(t=t/1000)
+
+p <- E4.total_time_df %>%
+  ggplot(aes(x=word,y=t)) +
+  geom_boxplot() +
+  scale_y_continuous(limits=c(0,150))
+
+ggsave('figures/hangman_nonpretend_pilot_time_per_word.png',p)
+
+E4.median_time <- E4.total_time_df %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(t=median(t)) %>%
+  spread(test_part,t) %>%
+  mutate(diff=pretend-nonpretend);
+
+# anova
+E4.time_anva <- aov(t ~ genuine_first * condition,
+                         data = E4.median_time %>%
+                           gather('condition','t',nonpretend,pretend) %>%
+                           mutate(condition=as.factor(condition)))
+
+p<- E4.total_time_df %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=t,fill=test_part))+
+    geom_histogram( alpha=0.5, position = 'identity',bins=100) +
+    scale_fill_manual(values=c("#69b3a2", "#404080")) +
+    scale_x_continuous(limits = c(0,200))+
+    labs(x='median game duration', y='number of players')+
+    theme(legend.position=c(0.90,0.85)) +
+    facet_wrap(~genuine_first, nrow=2) +
+    theme_classic();
+
+ggsave('figures/Hangman2/total_time.pdf',p)
+ggsave('figures/Hangman2/total_time.png',width=5,height=3,dpi=300)
+
+E4.time_per_word <- E4.total_time_df %>%
+  group_by(test_part,word) %>%
+  summarise(t=median(t)) %>%
+  spread(test_part,t)
+
+p <- E4.time_per_word %>%
+  ggplot(aes(x=pretend,y=nonpretend, label=word)) +
+  geom_abline(intercept=0,slope=1) +
+  geom_point()+
+  geom_text(position=position_jitter(), size=3.5)+
+  coord_fixed(ratio=1)+
+  scale_x_continuous(limits=c(15,60))+
+  scale_y_continuous(limits=c(15,60)) +
+  theme(legend.position = 'none')+
+  theme_classic()
+
+ggsave('figures/Hangman2/time_per_word.png',width=3,height=3,dpi=300)
+
+```
+
+### Hypothesis 1 (Game duration)
+
+The median game duration in non-pretend games was `r round(E4.median_time$nonpretend%>%median())` seconds in non-pretend games and `r round(E4.median_time$nonpretend%>%median())` seconds in pretend games.
+Pretend games were significantly shorter among subjects that pretended in the first block (`r E4.median_time%>%filter(!genuine_first)%>%pull(diff)%>%t.test()`) and in the second block (`r E4.median_time%>%filter(genuine_first)%>%pull(diff)%>%t.test()`).
+
+```{r hangman-time_per_click, echo=FALSE}
+
+N_perm <- 1000;
+bootstrap_error <- function(x, N_perm) {
+  N <- length(x)
+  medians = c();
+  for (i in 1:N_perm) {
+    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median(.,na.rm=T))
+  };
+  return(sd(medians))
+}
+
+E4.time_per_click_number <- E4.click_log %>%
+  group_by(subj_id,test_part,genuine_first,click_number)%>%
+  summarise(RT=median(RT))
+
+p <- E4.time_per_click_number %>%
+  group_by(genuine_first,test_part,click_number) %>%
+  summarise(median_RT=median(RT),
+            sem_RT = bootstrap_error(RT,1000)) %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  filter(click_number>=0 & click_number<=10) %>%
+  ggplot(aes(x=click_number,y=median_RT, color=test_part, fill=test_part)) +
+  geom_line() +
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT), alpha=0.5)+
+  scale_y_continuous(limits=c(0,7500))+
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  facet_wrap(~genuine_first,nrow=2) +
+  theme_classic()
+
+ggsave('figures/Hangman2/time_per_click_number.png',width=5,height=3,dpi=300)
+
+E4.first_click_time_per_word <- E4.click_log %>%
+  filter(click_number==1) %>%
+  group_by(test_part,word)%>%
+  summarise(t=median(t))%>%
+  spread(test_part,t)
+
+# p <- E4.time_per_first_click %>%
+#   ggplot(aes(x=word,y=RT)) +
+#   geom_boxplot() +
+#   scale_y_continuous(limits=c(0,15000))
+# 
+# ggsave('figures/hangman_nonpretend_pilot_time_per_first_click.png',p)
+```
+
+### Hypothesis 2 (first click latency)
+
+The median time taken to make the first letter selection was `r printnum(E4.time_per_click_number %>%filter(test_part=='pretend' & click_number==1)%>%pull(RT)%>%median()/1000)` seconds in pretend games and `r printnum(E4.time_per_click_number %>%filter(test_part=='nonpretend' & click_number==1)%>%pull(RT)%>%median()/1000)` in non-pretend games.
+First letter selections were slower in non-pretend games in both groups (pretend-first group: `r E4.time_per_click_number %>%filter(test_part=='nonpretend' & click_number==1 & !genuine_first)%>%pull(RT)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.time_per_click_number %>%filter(test_part=='nonpretend' & click_number==1 & genuine_first)%>%pull(RT)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+### Hypothesis 3: number of misses
+
+The median number of misses was `r E4.median_misses%>%pull(nonpretend)%>%median()` in non-pretend games and `r E4.median_misses%>%pull(pretend)%>%median()` in pretend games.
+The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-e4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
+
+```{r hangman-RT-by-accuracy}
+
+E4.click_log <- E4.click_log %>%
+  group_by(subj_id,test_part,word) %>%
+ mutate(hit_bin = hit !=0,
+        lag3 = lag(hit,3)!=0,
+        lag2 = lag(hit,2)!=0,
+        lag1 = lag(hit,1)!=0,
+        lead1 = lead(hit,1)!=0,
+        lead2 = lead(hit,2)!=0,
+        lead3 = lead(hit,3)!=0)
+
+E4.RT_by_lag1 <- E4.click_log %>%
+  filter(click_number>1 & click_number<num_clicks-1) %>%
+  group_by(genuine_first,subj_id,test_part,lag1)%>%
+  summarise(RT=median(RT, na.rm=T))  %>%
+  spread(lag1,RT,sep='_') %>%
+  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
+
+E4.RT_by_hit <- E4.click_log %>%
+  filter(click_number>1 & click_number<num_clicks-1) %>%
+  group_by(genuine_first,subj_id,test_part,hit_bin)%>%
+  summarise(RT=median(RT, na.rm=T))  %>%
+  spread(hit_bin,RT,sep='_') %>%
+  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
+
+E4.RT_by_lead1 <- E4.click_log %>%
+  filter(click_number>1 & click_number<num_clicks-1) %>%
+  group_by(genuine_first,subj_id,test_part,lead1)%>%
+  summarise(RT=median(RT, na.rm=T))  %>%
+  spread(lead1,RT,sep='_') %>%
+  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
+
+E4.RT_by_hit_lags <- E4.RT_by_lag1 %>%
+  merge(E4.RT_by_hit) %>%
+  merge(E4.RT_by_lead1)
+
+E4.RT_by_hit_lags_long <- E4.RT_by_hit_lags %>%
+  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
+  mutate(position = ifelse(position=="lag1_diff",1,
+                                         ifelse(position=="hit_diff",0,-1))) %>%
+  dplyr::select(genuine_first,subj_id,test_part,position,difference);
+
+p<- E4.RT_by_hit_lags_long %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  group_by(genuine_first,test_part,position) %>%
+  summarise(median_difference=median(difference, na.rm=T),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  # mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_vline(xintercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  facet_wrap(~genuine_first,nrow=2)+
+  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t')+
+  theme_classic()
+
+ggsave('figures/Hangman2/RT_by_lag_acc.pdf',p,width=7)
+ggsave('figures/Hangman2/RT_by_lag_acc.png',p,width=5, height=3, dpi=300)
+
+E4.RT_by_hit_lags_by_condition <- E4.RT_by_hit_lags_long %>%
+  spread(test_part,difference) %>%
+  mutate(interaction = pretend-nonpretend)
+
+p<- E4.RT_by_hit_lags_long %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  group_by(genuine_first,test_part,position) %>%
+  summarise(mean_difference=mean(difference, na.rm=T),
+            se_difference=se(difference)) %>%
+  # mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=mean_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_vline(xintercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=mean_difference-se_difference,ymax=mean_difference+se_difference),alpha=0.3,show.legend=F)+
+  facet_wrap(~genuine_first,nrow=2)+
+  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t')+
+  theme_classic()
+
+ggsave('figures/Hangman2/RT_by_lag_acc_with_mean_instead_of_median_at_the_group_level.png',p,width=5, height=3, dpi=300)
+
+```
+
+### Hypothesis 4 (click latency by outcome)
+
+In non-pretend games, cell selections that resulted in a hit were faster by `r round(abs(e4.RT_by_hit_lags_by_condition%>%filter(position==0)%>%pull(nonpretend)%>%median(na.rm=T)))` ms than those that resulted in a miss (pretend-first group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+In pretend games, cell selections that resulted in a hit were faster by `r round(abs(E4.RT_by_hit_lags_by_condition%>%filter(position==0)%>%pull(pretend)%>%median(na.rm=T)))` ms than those that resulted in a miss (pretend-first group: `r E4.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags %>% filter(test_part=='pretend' & genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effect of hitting versus missing a ship was marginally weaker in pretend games in the group that pretended first(`r E4.RT_by_hit_lags_long %>% filter(position==0 & !genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`) but not in the group that played normally first (`r E4.RT_by_hit_lags_long %>% filter(position==0 & genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Crucially, pretend and nonpretend games were not different according to this measure when focusing on the first block only (`r t.test(E4.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff), E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff))%>%apa_print()%>%'$'(statistic)`).
+
+### Hypothesis 5 (click latency by previous outcome)
+
+In non-pretend games, cell selections that followed a hit were not faster or slower than those that followed a hit (pretend-first group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+In contrast, pretend cell selections that followed in a hit were slower by `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()%>%abs()` ms.
+than those that followed a miss (pretend-first group: `r E4.RT_by_hit_lags %>% filter(test_part=='pretend' & !genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags %>% filter(test_part=='pretend' & genuine_first)%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effect of following a hit versus a miss was not significantly stronger in pretend games (pretend-first group: `r E4.RT_by_hit_lags_long %>% filter(position==1 & !genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags_long %>% filter(position==1 & genuine_first)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+### Hypothesis 6 (click latency by next outcome)
+
+Letter selections preceding a hit were faster than those preceding a miss in non-pretend games only in the group that pretended first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`) but not in the group that played normally first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`).
+This effect was not observed in pretend games (pretend first: `r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)%>%t.test())$full_result`; non-pretend first: `r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(pretend)%>%t.test())$full_result`).
+We find a significant interaction between these effects in the group that pretended first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(interaction)%>%t.test())$full_result`) but not in the other group (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(interaction)%>%t.test())$full_result`).
+The difference was not significant when analyzing first blocks only (`r apa_print(t.test(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend),E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)))$statistic`).
+
+### Model based analysis
+
+The next analyses were designed to test for differences in game optimality between pretend and non-pretend games, and for a relationship between decision difficulty and click latency.
+To do so, we approximated optimal behaviour by approximating the posterior probability that a letter appears in the word, given available information.
+Critically, in modelling pretend games we do not treat hints as part of this available information, because an optimal player should ignore this information in choosing where to click next.
+Given this posterior, a rational player should choose letters that have a high posterior probability of appearing in the word.
+
+To approximate the posterior probability of letters given a game state, we followed the following procedure:
+
+1.
+We used the category information (e.g., 'a fruit'), to obtain a probability-weighted list of category-compatible words (or names, in the case of famous people).
+The lists were obtained in the following way: for US states and number names, we used an exhaustive list (in the case of numbers, of numbers of 1-2 words), fruit names were taken from Wikipedia ([simple.wikipedia.org/wiki/List_of_fruits](https://simple.wikipedia.org/wiki/List_of_fruits)), famous people names from a [crowdsourced document](https://docs.google.com/spreadsheets/d/1mWez4KH0wueV_SSzdN16MMUG4qCa_nFFHToJcUvtJIA/edit#gid=4), and body-part words from existing prototypicality norms [@uyeda1980prototypicality].
+We extended Uyeda and Mandler's list of body part words by adding all body part words commonly appearing in lists on the internet.
+In the case of famous people, names were given prior probabilities in proportion to the number of visits their Wikipedia entries received in 2021.
+The top 100 most popular entries were given a prior probability twice that of the next 100, three times that of the next 100, and so forth.
+All entries from 600 and on were given the same prior probability, seven times smaller than that of entries at the top 100 positions.
+In the cases of fruits and body parts, we used prototypicality norms from Uyeda and Mandler (1980) to assign higher prior probability to more prototypical words (mapped to lower numbers on the 1-7 scale used by Uyeda and Mandler).
+Words that were not included in the norm were given the maximum score (that is, the lowest perceived prototypicality), of 7.
+Body parts and fruits were included in both singular and plural forms.
+The plural forms were assigned a prototypicality score of 100, due to the fact that category names were presented in the singular form ('a fruit' and 'a body part').
+We then took the reciprocal for each score, and normalized it by the total sum to get a probability distribution over words $p(w)$.
+The full prior distributions for each category are included in this pre-registration.
+
+2.
+The likelihood of a game state given a target word $p(s|w)$ equals 0 when the word is inconsistent with the information available to the player (this includes the word length, the identity of letters that do not appear in the word based on previous guesses, and the identity and position of letters that do appear in the word based on previous guesses).
+When consistent, the likelihood is a non-zero quantity that is equal for all consistent words.
+Bayes rule was used to extract the posterior over words given game state $p(w|s)=\frac{p(s|w)p(w)}{\Sigma_{w'}{p(s|w')p(w')}}$.
+The full non-zero $p(w|s)$ distributions for the initial states of all games are included in this pre-registration.
+
+3.
+The probability that an individual letter appears in the target word $p(l|s)$ is the sum of posterior probabilities of words that contain this letter $p(l|s)=\Sigma_{w}p(w|s)\times includes(w,l)$ where $includes(w,l)$ returns 1 if $l$ appears in $w$ and 0 otherwise.
+
+4.  A posterior probability over letter selections was obtained by dividing the probabilities of individual letters by their total sum $\frac{p(l|s)}{\Sigma_{l}p(l|s)}$.
+
+To quantify this notion of optimality, before each cell selection we computed the posterior probability that each of the unclicked letters appeared in the target word, given the game state.
+Then, we ranked letters from highest to lowest according to their posterior probability and recorded the rank of the chosen cell.
+
+```{r hangman-optimality-analysis, echo=FALSE}
+
+number_words_df <- read.csv('../analysis/word_lists/numbers.csv')
+states_df <- read.csv('../analysis/word_lists/USstates.csv')
+fruit_df <- read.csv('../analysis/word_lists/fruit.csv')
+body_df <-  read.csv('../analysis/word_lists/bodypart.csv')
+celebrity_df <- read.csv('../analysis/word_lists/celebrities.csv')
+letter_df <- read.csv('../analysis/word_lists/letters.csv')
+
+is_consistent <- function(word_state, asked,  word) {
+
+  word_state = strsplit(toupper(word_state),'')[[1]];
+  word = strsplit(toupper(word),'')[[1]];
+  asked = strsplit(toupper(asked),'')[[1]];
+  not_in_word = asked[!(asked %in% word_state)]
+
+  # check if number of letters is consistent
+  if (length(word_state) != length(word)) {
+    return(FALSE)
+  }
+
+  # check if the letters in the masked word are consistent
+  for (i in seq_along(word)) {
+    if (word_state[i] != word[i]) {
+      if (word_state[i] != '_' | word[i] == ' ' | word[i] == '.') {
+      return(FALSE)
+      }
+    }
+  }
+
+  # check if any of the misses appear in the candidate word
+  if (length(Reduce(intersect,list(word,not_in_word)))>0) {
+    return(FALSE)
+  }
+
+  #if the function hasn't returned FALSE, return TRUE
+  return(TRUE)
+
+}
+
+get_word_prior <- function(category) {
+
+  return(
+    switch(category,
+         "a number" = number_words_df,
+         "a US state" = states_df,
+         "a body part" = body_df,
+         "a famous person" = celebrity_df,
+         "a fruit" = fruit_df)
+  )
+}
+
+get_word_probabilities <- function(word_state, asked, category) {
+
+  word_probabilities <- get_word_prior(category) %>%
+    rowwise() %>%
+    mutate(likelihood = ifelse(is_consistent(word_state, asked,word),1, 0),
+           pword = prior * likelihood) %>%
+    ungroup() %>%
+    mutate(posterior = pword/sum(pword))%>%
+    filter(posterior>0) %>%
+    arrange(desc(posterior))
+
+  return(word_probabilities)
+
+};
+
+get_posterior <- function(word_state, asked, category) {
+
+  word_probabilities <- get_word_probabilities(word_state, asked, category);
+
+  letter_probabilities = c();
+  for (letter in toupper(letters)) {
+    if (grepl(letter,asked)) {
+      letter_probabilities = c(letter_probabilities,
+                               NA)
+    } else {
+      letter_probabilities = c(letter_probabilities,
+                             word_probabilities %>%
+                               filter(grepl(letter,toupper(word))) %>%
+                               pull(posterior) %>%
+                               sum()
+    )
+    }
+  }
+  posterior = letter_probabilities/sum(letter_probabilities, na.rm=T)
+
+  return(paste(posterior,collapse=','))
+
+};
+
+get_letter_prob_posterior <- function(asked) {
+
+  letter_probabilities <- letter_df %>%
+    rowwise() %>%
+    mutate(prob = ifelse(
+      grepl(toupper(letter),toupper(asked)),
+      NA,
+      frequency)
+      ) %>%
+    ungroup()%>%
+    mutate(
+      posterior = prob/sum(prob,na.rm=T)
+    )
+
+  return(paste(letter_probabilities$posterior,collapse=','))
+
+};
+
+get_p_click <- function(posterior, letter) {
+  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  return(posterior[which(toupper(letters)==letter)])
+}
+
+get_p_click_rank <- function(posterior, letter) {
+  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  rank_posterior = rank(-posterior)
+  return(rank_posterior[which(toupper(letters)==letter)])
+}
+
+get_posterior_entropy <- function(posterior) {
+  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  posterior[which(posterior==0)]=NA; #to avoid infinity*0
+  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
+  return(entropy);
+}
+
+# E4.click_log_with_boards <- E4.click_log %>%
+#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>% 
+#   rowwise()%>%
+#   mutate(posterior = get_posterior(word_state,asked,category),
+#          p_click = get_p_click(posterior,letter),
+#          entropy = get_posterior_entropy(posterior),
+#          p_click_rank = get_p_click_rank(posterior,letter));
+
+get_random_letter <- function(asked) {
+
+  asked = strsplit(tolower(asked),'')[[1]];
+  not_asked = setdiff(letters,asked)
+  return(toupper(not_asked)%>%sample(1))
+
+}
+
+get_most_frequent_letter <- function(asked) {
+
+   lp <- letter_df %>%
+    rowwise() %>%
+    mutate(prob = ifelse(
+      grepl(toupper(letter),toupper(asked)),
+      NA,
+      frequency)
+      ) %>%
+    ungroup()
+
+     return(lp[which.max(lp$prob),]$letter%>%toupper())
+
+}
+
+# E4.click_log_with_boards_random <- E4.click_log_with_boards %>%
+#   mutate(letter = get_random_letter(asked),
+#          p_click = get_p_click(posterior,letter),
+#          p_click_rank = get_p_click_rank(posterior,letter),
+#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
+#          test_part='random');
+
+# E4.click_log_with_boards_most_frequent <- E4.click_log_with_boards %>%
+#   mutate(letter = get_most_frequent_letter(asked),
+#          p_click = get_p_click(posterior,letter),
+#          p_click_rank = get_p_click_rank(posterior,letter),
+#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
+#          test_part='most_frequent');
+
+# E4.letter_prob_click_log_with_boards <- E4.click_log %>%
+#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>% 
+#   rowwise()%>%
+#   mutate(posterior = get_letter_prob_posterior(asked),
+#          p_click = get_p_click(posterior,letter),
+#          entropy = get_posterior_entropy(posterior),
+#          p_click_rank = get_p_click_rank(posterior,letter));
+
+# E4.letter_prob_click_log_with_boards_random <- E4.letter_prob_click_log_with_boards %>%
+#   mutate(letter = get_random_letter(asked),
+#          p_click = get_p_click(posterior,letter),
+#          p_click_rank = get_p_click_rank(posterior,letter),
+#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
+#          test_part='random');
+
+load('E4.click_log_with_boards.RData') # this file includes all the above dataframes.
+
+```
+
+```{r hangman-word-probabilities, echo=F}
+
+words <-E4.click_log %>%
+  group_by(word) %>% summarise(word_state=word_state[1], category=category[1],asked=asked[1]) %>%
+  filter(word != 'vladimir putin') %>%
+  mutate(word=as.character(word), category=as.character(category))
+
+words[nrow(words)+1,] = list('taylor swift','______ _____','a famous person','')
+words[nrow(words)+1,] = list('papaya','_A_A_A','a fruit','AEIOMT')
+words[nrow(words)+1,] = list('banana','_A_A_A','a fruit','AEIOMT')
+words[nrow(words)+1,] = list('hand','HA__','a body part','AEOMTHP')
+words[nrow(words)+1,] = list('hair','HA__','a body part','AEOMTHP')
+
+save_word_probs <- function(word,word_state, asked, category) {
+  probs <- get_word_probabilities(word_state, asked, category);
+  probs %>%
+      write.csv(paste('../analysis/word_lists/per_game/',word,'.csv',sep=''))
+  return('')
+}
+
+words %>%
+  rowwise() %>%
+  mutate(a=save_word_probs(word,word_state,asked,category))
+```
+
+```{r hangman-p-click, echo=F}
+
+E4.mean_P_click_rank <- E4.click_log_with_boards %>%
+  rbind(E4.click_log_with_boards_random)%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+E4.mean_P_click_rank_wide <- E4.mean_P_click_rank %>%
+  spread(test_part,p_click_rank) %>%
+  mutate(p_vs_r = pretend-random,
+         np_vs_r = nonpretend-random,
+         p_vs_np = pretend-nonpretend)
+
+E4.mean_letter_prob_P_click_rank <- E4.letter_prob_click_log_with_boards %>%
+  rbind(E4.click_log_with_boards_random)%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+E4.mean_letter_prob_P_click_rank_wide <- E4.mean_letter_prob_P_click_rank %>%
+  spread(test_part,p_click_rank) %>%
+  mutate(p_vs_r = pretend-random,
+         np_vs_r = nonpretend-random,
+         p_vs_np = pretend-nonpretend)
+
+p<- E4.mean_P_click_rank %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
+    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725")) +
+    labs(x='p(letter in word) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    facet_wrap(~genuine_first, nrow=2)+
+    theme_classic();
+
+ggsave('figures/Hangman2/pclick_rank.pdf',p)
+ggsave('figures/Hangman2/pclick_rank.png',p,width=5,height=3,dpi=300);
+
+p<- E4.mean_letter_prob_P_click_rank %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
+    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725")) +
+    labs(x='p(letter) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    facet_wrap(~genuine_first, nrow=2)+
+    theme_classic();
+
+ggsave('figures/Hangman2/letter_prob_pclick_rank.pdf',p)
+ggsave('figures/Hangman2/letter_prob_pclick_rank.png',p,width=5,height=3,dpi=300)
+
+E4.mean_P_click_rank_misses <- E4.click_log_with_boards %>%
+  filter(!hit_bin) %>%
+  rbind(E4.click_log_with_boards_random %>%filter(!hit_bin))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+E4.mean_P_click_rank_misses_wide <- E4.mean_P_click_rank_misses %>%
+  spread(test_part,p_click_rank) %>%
+  mutate(p_vs_r = pretend-random,
+         np_vs_r = nonpretend-random,
+         p_vs_np = pretend-nonpretend)
+
+E4.mean_letter_prob_P_click_rank_misses <- E4.letter_prob_click_log_with_boards %>%
+  filter(!hit_bin) %>%
+  rbind(E4.click_log_with_boards_random %>%filter(!hit_bin))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+E4.mean_letter_prob_P_click_rank_misses_wide <- E4.mean_letter_prob_P_click_rank_misses %>%
+  spread(test_part,p_click_rank) %>%
+  mutate(p_vs_r = pretend-random,
+         np_vs_r = nonpretend-random,
+         p_vs_np = pretend-nonpretend)
+
+p<- E4.mean_P_click_rank_misses %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(letter in word) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    facet_wrap(~genuine_first, nrow=2)+
+    theme_classic();
+
+ggsave('figures/Hangman2/pclick_rank_misses.pdf',p)
+ggsave('figures/Hangman2/pclick_rank_misses.png',p,width=5,height=3,dpi=300)
+
+p<- E4.mean_letter_prob_P_click_rank_misses %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(letter) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    facet_wrap(~genuine_first, nrow=2)+
+    theme_classic();
+
+ggsave('figures/Hangman2/letter_prob_pclick_rank_misses.pdf',p)
+ggsave('figures/Hangman2/letter_prob_pclick_rank_misses.png',p,width=5,height=3,dpi=300)
+
+E4.num_irrational_clicks <- E4.click_log_with_boards %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(nic=sum(p_click==0)/5) %>%
+  spread(test_part,nic)
+
+E4.irrational_clicks_misses_ratio <- E4.click_log_with_boards %>%
+  filter(!hit_bin)%>%
+  rbind(E4.click_log_with_boards_random %>%
+           filter(!hit_bin))%>%
+    rbind(E4.click_log_with_boards_most_frequent %>%
+           filter(!hit_bin))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(nic=sum(p_click==0)/n()) %>%
+  spread(test_part,nic) %>%
+  mutate(diff=pretend-nonpretend)
+
+E4.rank_misses_by_category <- E4.click_log_with_boards %>%
+  filter(!hit_bin) %>%
+  rbind(E4.click_log_with_boards_random %>%filter(!hit_bin))%>%
+  group_by(subj_id,test_part, category, word) %>%
+  summarise(p_click_rank=mean(p_click_rank)) %>%
+  group_by(test_part,category,word)%>%
+  summarise(p_rank = mean(p_click_rank,na.rm=T),
+            se = se(p_click_rank,na.rm=T))
+
+p <- E4.rank_misses_by_category %>% ggplot(
+  aes(x=as.factor(word),y=p_rank,fill=test_part)
+) + 
+  geom_bar(position=position_dodge(),
+           stat='identity', color='black')+
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+  geom_errorbar(aes(ymin=p_rank-se, ymax=p_rank+se), width=.2,position=position_dodge(.9)) +
+  labs(x='word',
+       y='posterior rank')
+
+ggsave('figures/Hangman2/prank_by_word.png',p,width=12,height=3,dpi=300)
+
+```
+
+### Hypothesis 7: optimality
+
+Letter selections in pretend games had a higher posterior probability to be included in the word than in random games (pretend first: `r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+The same was true for letter selections in pretend games (pretend first: `r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Non-pretend letter selections had a higher posterior probability among players who pretended first (`r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`) but not in the group that played normally first (`r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Comparing data from the first condition only in a between-subject manner revealed a significant difference (`r t.test(E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
+
+When focusing on misses, non-pretend letter selections were more optimal than random (pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`), as well as pretend games (pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+In both cases, non-pretend games were more optimal than pretend games (pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`)
+
+### Hypothesis 7b: Letter frequency
+
+Non-pretenders selected high-frequency letters more than expected by chance (pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+The same was true for letter selections in pretend games (pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Letters selected by pretenders had a higher overall frequency compared to letters selected by non-pretenders (pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+but not in the group that played normally first (`r E4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Comparing data from the first condition only in a between-subject manner revealed a significant difference (`r t.test(E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
+
+This was the case also when focusing on misses only ( pretend first: `r E4.mean_letter_prob_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; first block only: `r t.test(E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
 

### 2023-01-26 docs/preregisteredMethodsAndResults.Rmd
full pre-registered results

@@
   - id            : "2"
-    institution   : "Johns Hopkins University"
-
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
-
-abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
+    institution   : "Johns Hopkins University"s
 
@@
 classoption       : "man"
-output            : bookdown::html_book
+output            : bookdown::gitbook
 bibliography: references.bib
@@
 
-E4.click_log <- data.frame(matrix(ncol=11,nrow=0, 
-                               dimnames=list(NULL, 
-                                             c("subj_id",
-                                              "test_part", 
-                                              "word",
-                                              "category",
-                                              "num_clicks",
-                                              "num_misses",
-                                              "letter",
-                                              "hit",
-                                              "t",
-                                              "click_number",
-                                              "genuine_first"))))
-
-for (row in 1:nrow(E4.click_df)) {
-
-    subject_click_log <- data.frame(fromJSON(E4.click_df[row, ]$click_log)) %>%
-    mutate(
-      letter = lead(letter,1),
-      t = lead(t,1),
-      hit = lead(hit,1),
-      click_number = 1:n(),
-      subj_id = E4.click_df[row, ]$subj_id,
-      test_part = E4.click_df[row, ]$test_part,
-      word = E4.click_df[row, ]$word,
-      category = E4.click_df[row, ]$category,
-      num_clicks = E4.click_df[row, ]$num_clicks,
-      num_misses = E4.click_df[row, ]$num_misses,
-      genuine_first = E4.click_df[row, ]$genuine_first
-    )%>%
-      filter(click_number<=num_clicks)
-
-    E4.click_log <- rbind(E4.click_log, subject_click_log);
-}
-
-E4.click_log <- E4.click_log %>%
-  relocate(subj_id, .before = letter) %>%
-  relocate(genuine_first, .before = letter) %>% 
-  relocate(test_part, .before=letter) %>% 
-  relocate(word, .before=letter) %>% 
-  relocate(category, .before=letter) %>% 
-  relocate(click_number, .before=letter) %>%
-  group_by(subj_id,test_part,word) %>%
-  mutate(RT=t-lag(t,default=0)) %>%
-  group_by(subj_id,word) %>%
-  rowwise()
-# %>%
-#   filter(!(subj_id %in% E1.lucky_pretenders))
-
+# E4.click_log <- data.frame(matrix(ncol=11,nrow=0, 
+#                                dimnames=list(NULL, 
+#                                              c("subj_id",
+#                                               "test_part", 
+#                                               "word",
+#                                               "category",
+#                                               "num_clicks",
+#                                               "num_misses",
+#                                               "letter",
+#                                               "hit",
+#                                               "t",
+#                                               "click_number",
+#                                               "genuine_first"))))
+# 
+# 
+# for (row in 1:nrow(E4.click_df)) {
+#   
+#     subject_click_log <- data.frame(fromJSON(E4.click_df[row, ]$click_log)) %>%
+#     mutate(
+#       letter = lead(letter,1),
+#       t = lead(t,1),
+#       hit = lead(hit,1),
+#       click_number = 1:n(),
+#       subj_id = E4.click_df[row, ]$subj_id,
+#       test_part = E4.click_df[row, ]$test_part,
+#       word = E4.click_df[row, ]$word,
+#       category = E4.click_df[row, ]$category,
+#       num_clicks = E4.click_df[row, ]$num_clicks,
+#       num_misses = E4.click_df[row, ]$num_misses,
+#       genuine_first = E4.click_df[row, ]$genuine_first
+#     )%>%
+#       filter(click_number<=num_clicks)
+#     
+#     E4.click_log <- rbind(E4.click_log, subject_click_log);
+# }
+#     
+# 
+# E4.click_log <- E4.click_log %>%
+#   relocate(subj_id, .before = letter) %>%
+#   relocate(genuine_first, .before = letter) %>% 
+#   relocate(test_part, .before=letter) %>% 
+#   relocate(word, .before=letter) %>% 
+#   relocate(category, .before=letter) %>% 
+#   relocate(click_number, .before=letter) %>%
+#   group_by(subj_id,test_part,word) %>%
+#   mutate(RT=t-lag(t,default=0)) %>%
+#   group_by(subj_id,word) %>%
+#   rowwise()
+
+load('E4.click_log.RData')
 ```
@@
 The median game duration in non-pretend games was `r round(E4.median_time$nonpretend%>%median())` seconds in non-pretend games and `r round(E4.median_time$nonpretend%>%median())` seconds in pretend games.
-Pretend games were significantly shorter among subjects that pretended in the first block (`r E4.median_time%>%filter(!genuine_first)%>%pull(diff)%>%t.test()`) and in the second block (`r E4.median_time%>%filter(genuine_first)%>%pull(diff)%>%t.test()`).
+Pretend games were significantly shorter among subjects that pretended in the first block (`r apa_print(E4.median_time%>%filter(!genuine_first)%>%pull(diff)%>%t.test())$statistic`) and in the second block (`r apa_print(E4.median_time%>%filter(genuine_first)%>%pull(diff)%>%t.test())$statistic`).
 
@@
 
-### Hypothesis 3: number of misses
+### Hypothesis 3 (number of misses)
 
 The median number of misses was `r E4.median_misses%>%pull(nonpretend)%>%median()` in non-pretend games and `r E4.median_misses%>%pull(pretend)%>%median()` in pretend games.
-The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-e4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
+The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
 
@@
 
-In non-pretend games, cell selections that resulted in a hit were faster by `r round(abs(e4.RT_by_hit_lags_by_condition%>%filter(position==0)%>%pull(nonpretend)%>%median(na.rm=T)))` ms than those that resulted in a miss (pretend-first group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+In non-pretend games, cell selections that resulted in a hit were faster by `r round(abs(E4.RT_by_hit_lags_by_condition%>%filter(position==0)%>%pull(nonpretend)%>%median(na.rm=T)))` ms than those that resulted in a miss (pretend-first group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
 
@@
 
-words %>%
-  rowwise() %>%
-  mutate(a=save_word_probs(word,word_state,asked,category))
 ```
@@
 
-### Hypothesis 7: optimality
+### Hypothesis 7 (optimality)
 
@@
 
-### Hypothesis 7b: Letter frequency
+### Hypothesis 7b (Letter frequency)
 
@@
 
+```{r hangman-RT-entropy, echo=F}
+
+E4.RT_by_entropy <- E4.click_log_with_boards %>%
+  filter(click_number>1) %>%
+  mutate(entropy_bin=round(entropy*10)/10) %>%
+  group_by(test_part,subj_id,genuine_first,entropy_bin) %>% 
+  summarise(RT=median(RT)) %>%
+  group_by(test_part,genuine_first,entropy_bin) %>%
+  summarise(median_RT=median(RT),
+            sem_RT=bootstrap_error(RT,N_perm))
+
+p <- E4.RT_by_entropy %>%
+  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
+  facet_wrap(~genuine_first,nrow=2)+
+  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
+  theme_classic()
+
+ggsave('figures/Hangman2/RT_by_entropy.png',p,width=5,height=3,dpi=300)
+
+E4.RT_by_entropy_trimmed <- E4.click_log_with_boards %>%
+  # filter(click_number>1) %>%
+  filter(RT<5000 & RT>100) %>%
+  mutate(entropy_bin=round(entropy*5)/5) %>%
+  group_by(test_part,subj_id,genuine_first,entropy_bin) %>% 
+  summarise(RT=median(RT)) %>%
+  group_by(test_part,genuine_first,entropy_bin) %>%
+  summarise(median_RT=median(RT),
+            sem_RT=bootstrap_error(RT,N_perm))
+
+p <- E4.RT_by_entropy_trimmed %>%
+  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
+  facet_wrap(~genuine_first,nrow=2)+
+  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
+  theme_classic()
+
+ggsave('figures/hangman2/RT_by_entropy_trimmed.png',p,width=5,height=3,dpi=300)
+
+E4.RT_by_entropy_coefs <- E4.click_log_with_boards %>%
+  # filter(click_number>1) %>%
+  group_by(genuine_first,subj_id,test_part,word) %>%
+  mutate(entropy_centered = entropy-mean(entropy),
+         entropy_squared = entropy_centered^2) %>%
+  select(genuine_first,subj_id,test_part,word,click_number,entropy_centered,entropy_squared,RT) %>%
+  group_by(genuine_first,subj_id,test_part,word) %>%
+  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
+  mutate(tidys=list(broom::tidy(model))) %>%
+  unnest(tidys) %>%
+  dplyr::select(!c(std.error,statistic,p.value)) %>%
+  group_by(genuine_first,subj_id,test_part,term) %>%
+  summarise(estimate = mean(estimate))
+
+E4.RT_by_entropy_coefs_wide <- E4.RT_by_entropy_coefs %>% 
+  spread(test_part,estimate) %>%
+  mutate(diff=pretend-nonpretend)
+
+E4.RT_by_entropy_coefs_summary <- E4.RT_by_entropy_coefs %>%
+  group_by(test_part,term, genuine_first) %>%
+  summarise(mean_estimate = mean(estimate),
+            se_estimate = se(estimate))
+
+p <- E4.RT_by_entropy_coefs_summary %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
+  filter(term!='(Intercept)') %>%
+  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_bar(stat='identity',color='black',position=position_dodge()) +
+  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
+  labs(fill=' ',x='term', y='coefficient')+
+  theme_classic() +
+  facet_grid(rows=vars(genuine_first))
+
+ggsave('figures/Hangman2/RT_by_entropy_coefficientst.png',p,width=5,height=5,dpi=300)
+
+E4.corrected_RT_by_entropy_coefs <- E4.click_log_with_boards %>%
+  group_by(subj_id,test_part,click_number) %>%
+  mutate(corrected_RT = RT-mean(RT)) %>%
+  group_by(genuine_first,subj_id,test_part) %>%
+  mutate(entropy_centered = entropy-mean(entropy),
+         entropy_squared = entropy_centered^2) %>%
+  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
+  group_by(genuine_first,subj_id,test_part) %>%
+  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
+  mutate(tidys=list(broom::tidy(model))) %>%
+  unnest(tidys) %>%
+  dplyr::select(!c(std.error,statistic,p.value))
+
+E4.corrected_RT_by_entropy_coefs_summary <- E4.corrected_RT_by_entropy_coefs %>%
+  group_by(test_part,term,genuine_first) %>%
+  summarise(mean_estimate = mean(estimate),
+            se_estimate = se(estimate))
+
+p <- E4.corrected_RT_by_entropy_coefs_summary %>%
+      mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
+  filter(term!='(Intercept)') %>%
+  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_bar(stat='identity',color='black',position=position_dodge()) +
+  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
+  labs(fill=' ',x='term', y='coefficient')+
+  theme_classic()+
+  facet_grid(rows=vars(genuine_first))
+
+ggsave('figures/Hangman2/corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=5,dpi=300)
+
+```
+
+### Hypothesis 8(Entropy/RT relationship)
+
+In non-pretend games, high uncertainty was associated with longer decision times (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The same was true for pretend games (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effects were significantly weaker in pretend games (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_centered' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+In non-pretend games, the relationship between uncertainty and RT was negatively quadratic, such that the longest responses were the one associated with intermediate uncertainty levels (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(nonpretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The same was true for pretend games (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & genuine_first)%>%pull(pretend)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+The effects were significantly weaker in pretend games (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+```{r hangman-judge, echo=F}
+
+word_used_for_pretend <- function(s, w) {
+
+  return (E4.df%>%filter(test_part=='pretend' &
+                           subj_id==s & 
+                           word == w) %>%
+            nrow()>0)
+}
+
+word_used_for_nonpretend <- function(s, w) {
+
+  return (E4.df%>%filter(test_part=='nonpretend' &
+                           subj_id==s & 
+                           word == w) %>%
+            nrow()>0)
+}
+
+# E4.judge_df <- E4.df %>%
+#   filter(test_part=='judge' & trial_type=='html-keyboard-response')%>%
+#   dplyr::select(subj_id, 
+#          word, 
+#          player,
+#          correct_response,
+#          key_press,
+#          correct)%>%
+#   rowwise()%>%
+#   mutate(used_for_pretend = word_used_for_pretend(subj_id,word),
+#          used_for_nonpretend = word_used_for_nonpretend(subj_id,word),
+#          common_board = used_for_pretend & used_for_nonpretend,
+#          word_type = ifelse(common_board, 'common',
+#                              ifelse(used_for_pretend, 'pretend',
+#                                     ifelse(used_for_nonpretend,'nonpretend','new')))
+# )
+
+load('E4.judge_df.RData')
+
+E4.judge_summary_df <- E4.judge_df %>%
+  group_by(subj_id) %>%
+  summarise(accuracy=mean(correct),
+            accuracy_pretend_words = mean(correct[word_type=='pretend']),
+            accuracy_nonpretend_words = mean(correct[word_type=='nonpretend']),
+            bias = mean(key_press==80),
+            bias_pretend_words = mean(key_press[word_type=='pretend']==80),
+            bias_nonpretend_words = mean(key_press[word_type=='nonpretend']==80)) %>%
+  mutate(bias_experience_effect = bias_pretend_words-bias_nonpretend_words)
+
+```
+
+### Hypothesis 9 (judge accuracy):
+
+Participants were not significantly above chance in judging which of the two presented games came from a pretender (`r apa_print(t.test(E4.judge_summary_df$accuracy, mu=0.5))$full_result`).
+This was the case also when focusing on words that judges encountered when pretending (`r apa_print(t.test(E4.judge_summary_df$accuracy_pretend_words, mu=0.5))$full_result`), and when focusing on words that judges encountered when playing normally (`r apa_print(t.test(E4.judge_summary_df$accuracy_nonpretend_words, mu=0.5))$full_result`).
+
+*Hypothesis 9B (judge bias):* Participants were slightly biased to classify games as pretend games (`r apa_print(E4.judge_summary_df$bias%>%t.test(mu=0.5))$full_result`).
+
+```{r hangman-judge-pretend-correlation}
+
+unidentifiable_ss <- E4.judge_df %>%
+  group_by(subj_id)%>%
+  summarise(n=n())%>%
+  filter(n>5)%>%
+  pull(subj_id)%>%
+  unique();
+
+E4.judge_performance <- E4.judge_df %>%
+  filter(!(subj_id %in% unidentifiable_ss)) %>%
+  group_by(subj_id)%>%
+  summarise(judge_rate=sum(correct)/5);
+
+E4.pretend_performance <- E4.judge_df %>%
+  filter(!(player %in% unidentifiable_ss & pretend==1)) %>%
+  group_by(player)%>%
+  summarise(ntrials=n(),
+            successful=ntrials-sum(correct),
+            pretend_rate = successful/ntrials) %>%
+  mutate(subj_id=player)
+
+E4.optimality_cost <- E4.click_log_with_boards %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click_rank=mean(p_click_rank)) %>%
+  spread(test_part,p_click_rank) %>%
+  mutate(optimality_cost=nonpretend-pretend)
+
+E4.performance_df <- merge(E4.judge_performance,
+                           E4.pretend_performance) %>%
+  merge(E4.optimality_cost);
+
+p <- E4.performance_df %>%
+  ggplot(aes(x=judge_rate,y=pretend_rate)) +
+  geom_jitter() + 
+  labs(x='judge accuracy',y='pretense success rate') +
+  theme_classic()
+
+ggsave('figures/Hangman2/judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)
+
+p <- E4.performance_df %>%
+  ggplot(aes(x=judge_rate,y=optimality_cost)) +
+  geom_jitter() + 
+  labs(x='judge accuracy',y='optimality cost') +
+  theme_classic()
+
+ggsave('figures/Hangman2/optimality_judge_accuracy.png',p,width=5,height=3,dpi=300)
+
+```
+
+## Hypothesis 10 (judge-pretend relation)
+
+No significant correlation between judge accuracy and ability to trick others (`r cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman')%>%apa_print()%>%'$'(full_result)`), nor with the optimality cost to pretending (`r cor.test(E4.performance_df$optimality_cost,E4.performance_df$judge_rate,method='spearman')%>%apa_print()%>%'$'(full_result)`).
+
+```{r hangman-half-games}
+
+E4.half_game_click_df <- E4.df %>%
+  dplyr::select(subj_id, 
+         test_part, 
+         word,
+         num_clicks,
+         click_log,
+         genuine_first) %>%
+  rowwise()%>%
+  mutate(num_hits = strsplit(gsub(' ','',word),split='')[[1]]%>%unique()%>%length(),
+         num_misses = num_clicks-num_hits,
+         click_log = gsub("\'","\"", click_log),
+         click_log = gsub("None","null", click_log),
+         word = factor(word,levels=c(
+           'papaya',
+           'banana',
+           'hair',
+           'hand'
+         )),
+         category = ifelse(word=='papaya'|word=='banana','a fruit','a body part')) %>%
+  filter(test_part=='pretend_half_game' | test_part=='nonpretend_half_game')
+
+E4.half_game_click_log <- data.frame(matrix(ncol=11,nrow=0, 
+                               dimnames=list(NULL, 
+                                             c("subj_id",
+                                              "test_part", 
+                                              "word",
+                                              "category",
+                                              "num_clicks",
+                                              "num_misses",
+                                              "letter",
+                                              "hit",
+                                              "t",
+                                              "click_number",
+                                              "genuine_first"))))
+
+for (row in 1:nrow(E4.half_game_click_df)) {
+
+    subject_click_log <- data.frame(fromJSON(E4.half_game_click_df[row, ]$click_log)) %>%
+    mutate(
+      letter = lead(letter,1),
+      t = lead(t,1),
+      hit = lead(hit,1),
+      click_number = 1:n(),
+      subj_id = E4.half_game_click_df[row, ]$subj_id,
+      test_part = E4.half_game_click_df[row, ]$test_part,
+      word = E4.half_game_click_df[row, ]$word,
+      category = E4.half_game_click_df[row, ]$category,
+      num_clicks = E4.half_game_click_df[row, ]$num_clicks,
+      num_misses = E4.half_game_click_df[row, ]$num_misses,
+      genuine_first = E4.half_game_click_df[row, ]$genuine_first
+    )%>%
+      filter(click_number<=num_clicks)
+
+    E4.half_game_click_log <- rbind(E4.half_game_click_log, subject_click_log);
+}
+
+E4.half_game_click_log <- E4.half_game_click_log %>%
+  relocate(subj_id, .before = letter) %>%
+  relocate(genuine_first, .before = letter) %>% 
+  relocate(test_part, .before=letter) %>% 
+  relocate(word, .before=letter) %>% 
+  relocate(category, .before=letter) %>% 
+  relocate(click_number, .before=letter) %>%
+  group_by(subj_id,test_part,word) %>%
+  mutate(RT=t-lag(t,default=0)) %>%
+  group_by(subj_id,word) %>%
+  rowwise()
+
+E4.letter_choice_frequency <- E4.half_game_click_log %>%
+  filter(click_number==1) %>%
+  dplyr::select(subj_id,test_part,word_state,letter) %>%
+  mutate(letter=factor(letter,levels=toupper(letters))) %>%
+  group_by(letter,test_part,word) %>%
+  summarise(count = n()) %>%
+  group_by(test_part,word) %>%
+  mutate(p=count/sum(count)) %>%
+  complete(letter,fill=list(count=0,p=0)) %>%
+  mutate(test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend'))
+
+E4.first_click_half_games <- E4.half_game_click_log %>%
+  filter(click_number==1) %>%
+  dplyr::select(subj_id,genuine_first,word,letter,t,hit,test_part) %>%
+  mutate(
+    makes_sense = 
+    (word %in% c('hair','hand') & letter %in% c('N','D','I','R')) | 
+      (word %in% c('banana','papaya') & letter %in% c('P','Y','B','N')),
+    bananahand = letter %in% c('B','N','D')
+  )
+
+p <- E4.first_click_half_games %>%
+  group_by(test_part,genuine_first) %>%
+  summarise(makes_sense=mean(makes_sense)) %>%
+  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'),
+         test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend')) %>%
+  ggplot(aes(x=test_part,fill=test_part,y=makes_sense))+
+  geom_bar(position=position_dodge(),
+           stat='identity', color='black')+
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+  facet_grid(~genuine_first)+
+  theme_classic()+
+  labs(y='proportion of clicks that make sense')
+
+ggsave('figures/Hangman2/halfgames_make_sense.png',p,width=6,height=3,dpi=300)
+
+p <- E4.first_click_half_games %>%
+  filter(makes_sense)%>%
+  mutate(word_class = ifelse(word %in% c('banana','hand'),'BANANA or HAND','PAPAYA or HAIR'))%>%
+  group_by(test_part,genuine_first,word_class) %>%
+  summarise(bananahand = mean(bananahand)) %>%
+  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'),
+         test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend')) %>%
+  ggplot(aes(x=test_part,fill=test_part,y=bananahand))+
+  geom_bar(position=position_dodge(),
+           stat='identity', color='black')+
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+  facet_grid(rows=vars(genuine_first),col=vars(word_class))+
+  theme_classic()+
+  labs(y='(consistent with BANANA/HAND)/(cons. with something)')
+
+ggsave('figures/Hangman2/halfgames_attraction.png',p,width=6,height=4,dpi=300)
+
+E4.optimal_letter_choice <- E4.half_game_click_log %>% 
+  filter(click_number==1) %>% 
+  group_by(word) %>% 
+  summarize(posterior = get_posterior(word_state[1],asked[1],category[1]),
+            word_state=word_state[1])
+
+E4.optimal_by_letter <- data.frame(matrix(ncol=5,nrow=0, 
+                               dimnames=list(NULL, 
+                                             c("test_part", 
+                                               "word",
+                                              "word_state",
+                                              "letter",
+                                              "p"))))
+
+for (row in 1:nrow(E4.optimal_letter_choice)) {
+
+   p <- scan(text= E4.optimal_letter_choice[row,]$posterior, what = numeric(), sep="," , quiet = TRUE);
+   letter = LETTERS;
+
+   this.word <- tibble(letter,p) %>%
+     mutate(
+       word_state = E4.optimal_letter_choice[row,]$word_state,
+       word = E4.optimal_letter_choice[row,]$word,
+       test_part = 'optimal'
+     )
+
+    E4.optimal_by_letter <- rbind(E4.optimal_by_letter, this.word);
+
+}
+
+E4.letter_choice_frequency <-
+  E4.letter_choice_frequency %>%
+  rbind(E4.optimal_by_letter)
+
+p <- E4.letter_choice_frequency %>% 
+  mutate(test_part = factor(test_part, levels=c('non-pretend','pretend','optimal'))) %>%
+  ggplot(aes(x=letter,y=p, fill=test_part)) +
+  geom_bar(stat='identity', position='dodge') +
+  facet_grid(rows=vars(word)) +
+  theme_minimal() +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#49B240"))
+
+ggsave('figures/hangmanpilot_halfgames.png',p,width=6,height=3,dpi=300)
+
+p <- E4.letter_choice_frequency %>%
+  filter((word=='papaya' | word=='banana') & test_part != 'optimal') %>%
+  mutate(letter = factor(
+    ifelse(letter %in% c('B','N','P','Y'), letter, 'other'),
+    levels=rev(c('B','N','P','Y','other'))),
+    test_part = factor(test_part, levels=c('non-pretend','pretend'))) %>%
+  group_by(test_part,word,letter) %>%
+  summarise(p=sum(p)) %>%
+  ggplot(aes(x=letter,y=p, fill=test_part)) +
+  geom_bar(stat='identity', position='dodge') +
+  coord_flip()+
+  facet_grid(cols=vars(word)) +
+  theme_minimal() +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#49B240"))
+
+irrational_test_p_first <- binom.test(
+  E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='False') %>%
+    pull(makes_sense) %>%
+    sum(), 
+  E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='False') %>%
+    nrow(),
+  E4.first_click_half_games %>% 
+    filter(test_part=='nonpretend_half_game' & genuine_first=='False') %>%
+    pull(makes_sense) %>%
+    mean())
+
+irrational_test_np_first <- binom.test(
+  E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='True') %>%
+    pull(makes_sense) %>%
+    sum(), 
+  E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='True') %>%
+    nrow(),
+  E4.first_click_half_games %>% 
+    filter(test_part=='nonpretend_half_game' & genuine_first=='True') %>%
+    pull(makes_sense) %>%
+    mean())
+
+attraction_test_np_first <- prop.test(
+  x=c(
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('banana','hand')) %>%
+    pull(bananahand) %>%
+    sum(), 
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('papaya','hair')) %>%
+    pull(bananahand) %>%
+    sum()), 
+  n= c(
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('banana','hand')) %>%
+    nrow(),
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('papaya','hair')) %>%
+    nrow()))
+
+attraction_test_p_first <- prop.test(
+  x=c(
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('banana','hand')) %>%
+    pull(bananahand) %>%
+    sum(), 
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('papaya','hair')) %>%
+    pull(bananahand) %>%
+    sum()), 
+  n= c(
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('banana','hand')) %>%
+    nrow(),
+    E4.first_click_half_games %>% 
+    filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('papaya','hair')) %>%
+    nrow()))
+
+```
+
+## Hypothesis 12 (irrational letter selections)
+
+In non-pretend games, subjects almost always chose letters that made sense (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`).
+In pretend games, these figures were significantly lower (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`, $p<.001$; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`, $p<.001$).
+
+## Hypothesis 13 (target word effect)
+
+Excluding irrational clicks, non-pretenders had an overall preference to select letters that are consistent with the words 'BANANA' and 'HAND' (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='False' & makes_sense)%>%pull(bananahand)%>%mean()%>%printnum()`; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='True' & makes_sense)%>%pull(bananahand)%>%mean()%>%printnum()`).
+Pretenders' preference for the same letters was weaker (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense)%>%pull(bananahand)%>%mean()%>%printnum()`, $p<.001$; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense)%>%pull(bananahand)%>%mean()%>%printnum()`, $p<.001$).
+
+Within pretend games, preference for letters that are consistent with HAND or BANANA was significantly stronger when the word itself was BANANA or HAND, compared to when it was PAPAYA or HAIR (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('hand','banana'))%>%pull(bananahand)%>%mean()%>%printnum()` vs.
+`r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='False' & makes_sense & word %in% c('hair','papaya'))%>%pull(bananahand)%>%mean()%>%printnum()`, $p<.001$; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('hand','banana'))%>%pull(bananahand)%>%mean()%>%printnum()` vs.
+`r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='True' & makes_sense & word %in% c('hair','papaya'))%>%pull(bananahand)%>%mean()%>%printnum()`, $p<.001$).
+
 # Discussion

### 2023-01-27 docs/preregisteredMethodsAndResults.Rmd
all pre-registered analysis

@@
   - id            : "2"
-    institution   : "Johns Hopkins University"s
+    institution   : "Johns Hopkins University"
+
+authornote: |
+
+abstract: |
 
@@
 set.seed(42)
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
-                      fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, 
-                      message=F, sanitize = T)
-```
-
-```{r, libraries, echo=FALSE, warning=FALSE}
+knitr::opts_chunk$set(fig.crop = F, fig.pos = "tb", fig.path='figs/',
+                      echo=F, warning=F, cache=T, 
+                      message=F, sanitize = T, 
+                      fig.width = 10, fig.height = 5)
+```
+
+```{r, libraries, echo=FALSE, warning=FALSE, cache=F}
 
@@
 
-<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->
+We used `r cite_r("r-references.bib")` for all our analyses.
 
@@
 
+```{r battleship-h1-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E2.click_log %>% 
+  group_by(subj_id,test_part,grid_number, genuine_first) %>%
+  summarise(duration=sum(RT)/1000) %>% 
+  group_by(subj_id,test_part,genuine_first) %>% 
+  summarise(duration=median(duration)) %>%
+  ggplot(aes(x=duration,color=test_part, fill=test_part)) +
+  geom_histogram(position='identity',alpha=0.5)  +  
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='game duration (seconds)')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic()
+
+```
+
 ```{r battleship-first-click, echo-FALSE, cache=TRUE}
@@
 This test revealed a significant difference in the first click latency (`r t.test(E2.first_click_log%>%filter(!genuine_first)%>%pull(pretend), E2.first_click_log%>%filter(genuine_first)%>%pull(nonpretend))%>%apa_print()%>%'$'(full_result)`).
+
+```{r battleship-h2-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E2.click_log %>% 
+  group_by(subj_id,test_part,grid_number, genuine_first) %>%
+  summarise(RT=RT[click_number==1]) %>% 
+  group_by(subj_id,test_part,genuine_first) %>% 
+  summarise(RT=median(RT)) %>%
+  filter(RT<20000)%>%
+  ggplot(aes(x=RT,color=test_part, fill=test_part)) +
+  geom_histogram(position='identity',alpha=0.5)  +  
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='time to first cell selection (ms)')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic()
+
+```
 
@@
 num_clicks_sim = c();
-nsim = 10000
+nsim = 2500
 boards = rep(NaN,25)
@@
 }
-
-simulated_hits <- data.frame(click_number=1:25,mean_hit=colMeans(boards,na.rm=T))
 
@@
 
-num_clicks_optimal_sim = c();
-nsim = 100;
-nboards = all_possible_boards%>%nrow();
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  num_clicks_optimal_sim = c(num_clicks_optimal_sim,simulate_optimal_game(cur_board));
-}
+# nsim = 2500;
+# nboards = all_possible_boards%>%nrow();
+
+# for (i in 1:nsim) {
+#   cur_board = all_possible_boards[sample(1:nboards,1),];
+#   num_clicks_optimal_sim = c(num_clicks_optimal_sim,simulate_optimal_game(cur_board));
+# }
+# save(num_clicks_sim, num_clicks_optimal_sim, file='num_clicks_sim.RData')
+
+load('num_clicks_sim.RData')
+simulated_hits <- data.frame(click_number=1:25,mean_hit=colMeans(boards,na.rm=T))
 
@@
 We assumed that these participants did not follow the task instructions, and excluded them from all analyses.
+
+```{r battleship-h3-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E2.random_click_df <- data.frame(num_clicks_sim) %>% 
+  rename(num_clicks = num_clicks_sim) %>% 
+  mutate(test_part='random', 
+         subj_id=rep(seq(500),each=5));
+
+num_clicks_optimal_sim_2500 = num_clicks_optimal_sim[1:2500]
+
+E2.optimal_click_df <- data.frame(num_clicks_optimal_sim_2500) %>% 
+  rename(num_clicks = num_clicks_optimal_sim_2500) %>% 
+  mutate(test_part='optimal', 
+         subj_id=rep(seq(500),each=5))
+
+E2.click_df %>%
+  dplyr::select(subj_id,test_part, num_clicks) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  rbind(E2.optimal_click_df) %>%
+  rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend','optimal','random'
+  ))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_clicks=median(num_clicks)) %>%
+  ggplot(aes(x=num_clicks,fill=test_part))+
+  geom_bar( color="black", alpha=0.6, position = 'identity') +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
+  labs(x='median number of clicks', y='number of players')+
+  theme(legend.position=c(0.10,0.85)) +
+  theme_classic()
+
+```
 
@@
 
+```{r battleship-h4-6-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E2.RT_by_hit_lags_long %>% 
+  filter(position<2 & position > -2)%>%
+  group_by(test_part,position) %>%
+  summarise(median_difference=median(difference),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
+  theme(legend.position=c(0.10,0.85)) +
+  theme_classic()
+
+```
 ### Model based analysis
@@
 
+```{r battleship-h7-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E2.mean_P_click_rank %>%
+    filter(test_part != 'optimal') %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="black", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+```
+
 ### Hypothesis 8 (entropy-RT relation)
@@
 This negative quadratic effect was stronger in pretend compared to non-pretend games (pretend-first group: `r E2.RT_by_entropy_coefs%>%filter(!genuine_first & term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.RT_by_entropy_coefs%>%filter(genuine_first & term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+```{r battleship-h8-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E2.RT_by_entropy %>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
+  theme_classic()
+
+```
 
@@
 
+```{r battleship-h10-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+# E2.performance_df %>%
+#   group_by(judge_rate) %>%
+#   summarise(mean_pretend_rate = mean(pretend_rate),
+#             se_pretend_rate=se(pretend_rate)) %>%
+#   ggplot(aes(x=judge_rate, y=mean_pretend_rate, 
+#          ymin=mean_pretend_rate-se_pretend_rate, 
+#          ymax=mean_pretend_rate+se_pretend_rate)) +
+#   geom_point(size=3) +
+#   geom_errorbar()
+
+E2.performance_df %>%
+  ggplot(aes(x=pretend_rate, y=judge_rate)) +
+  geom_point() +
+  labs(x= 'successful pretense rate', y='judge accuracy') +
+  theme_classic()
+
+```
+
 # Exp. 2: Hangman
@@
   rbind(read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
-  rename(subj_id = participant_number) %>%
+  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,5),sep='')) %>%
   mutate(subj_id = factor(subj_id)) %>%
@@
 
-### Data analysis
-
-We used `r cite_r("r-references.bib")` for all our analyses.
-
 ## Results
@@
   dplyr::select(subj_id, 
+         PROLIFIC_PID,
          test_part, 
@@
 #                                               "genuine_first"))))
-# 
-# 
+
 # for (row in 1:nrow(E4.click_df)) {
@@
 # }
-#     
-# 
+
 # E4.click_log <- E4.click_log %>%
@@
 
+# save(E4.click_log, file='E4.click_log.RData')
+
 load('E4.click_log.RData')
@@
 Pretend games were significantly shorter among subjects that pretended in the first block (`r apa_print(E4.median_time%>%filter(!genuine_first)%>%pull(diff)%>%t.test())$statistic`) and in the second block (`r apa_print(E4.median_time%>%filter(genuine_first)%>%pull(diff)%>%t.test())$statistic`).
+
+```{r hangman-duration-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.click_log %>% 
+  group_by(subj_id,test_part,word, genuine_first) %>%
+  summarise(duration=sum(RT)/1000) %>% 
+  group_by(subj_id,test_part,genuine_first) %>% 
+  summarise(duration=median(duration)) %>%
+  ggplot(aes(x=duration,color=test_part, fill=test_part)) +
+  geom_histogram(position='identity',alpha=0.5)  +  
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='game duration (seconds)')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic()
+
+```
 
@@
 
+```{r hangman-h2-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.time_per_click_number %>% 
+  filter(click_number==1 & subj_id != '5975df4') %>%
+  mutate(RT=RT/1000)%>%
+  filter(RT<40) %>%
+  ggplot(aes(x=RT,color=test_part, fill=test_part)) +
+  geom_histogram(position='identity',alpha=0.5)  +  
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='time to first cell selection (seconds)')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic()
+
+```
+
 ### Hypothesis 3 (number of misses)
@@
 The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
+
+```{r hangman-h3-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.click_df %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_misses)) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend','optimal','random'
+  ))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_clicks=median(num_misses)) %>%
+  ggplot(aes(x=num_clicks,fill=test_part))+
+  geom_bar( color="black", alpha=0.6, position = 'identity') +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position=c(0.10,0.85)) +
+  theme_classic()
+
+```
 
@@
 
+```{r hangman-h4-6-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.RT_by_hit_lags_long %>% 
+  filter(position<2 & position > -2)%>%
+  group_by(test_part,position) %>%
+  summarise(median_difference=median(difference, na.rm=T),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
+  theme(legend.position=c(0.10,0.85)) +
+  theme_classic()
+
+```
+
 ### Model based analysis
@@
 
-is_consistent <- function(word_state, asked,  word) {
+word_is_consistent <- function(word_state, asked,  word) {
 
@@
     rowwise() %>%
-    mutate(likelihood = ifelse(is_consistent(word_state, asked,word),1, 0),
+    mutate(likelihood = ifelse(word_is_consistent(word_state, asked,word),1, 0),
            pword = prior * likelihood) %>%
@@
 
-get_posterior <- function(word_state, asked, category) {
+get_letter_posterior <- function(word_state, asked, category) {
 
@@
 
-get_p_click <- function(posterior, letter) {
+get_p_letter <- function(posterior, letter) {
   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
@@
 
-get_p_click_rank <- function(posterior, letter) {
+get_p_letter_rank <- function(posterior, letter) {
   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
@@
 
-get_posterior_entropy <- function(posterior) {
+get_letter_posterior_entropy <- function(posterior) {
   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
@@
 #   rowwise()%>%
-#   mutate(posterior = get_posterior(word_state,asked,category),
-#          p_click = get_p_click(posterior,letter),
-#          entropy = get_posterior_entropy(posterior),
-#          p_click_rank = get_p_click_rank(posterior,letter));
+#   mutate(posterior = get_letter_posterior(word_state,asked,category),
+#          p_click = get_p_letter(posterior,letter),
+#          entropy = get_letter_posterior_entropy(posterior),
+#          p_click_rank = get_p_letter_rank(posterior,letter));
 
@@
 #   mutate(letter = get_random_letter(asked),
-#          p_click = get_p_click(posterior,letter),
-#          p_click_rank = get_p_click_rank(posterior,letter),
+#          p_click = get_p_letter(posterior,letter),
+#          p_click_rank = get_p_letter_rank(posterior,letter),
 #          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
@@
 #   mutate(letter = get_most_frequent_letter(asked),
-#          p_click = get_p_click(posterior,letter),
-#          p_click_rank = get_p_click_rank(posterior,letter),
+#          p_click = get_p_letter(posterior,letter),
+#          p_click_rank = get_p_letter_rank(posterior,letter),
 #          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
@@
 #   mutate(posterior = get_letter_prob_posterior(asked),
-#          p_click = get_p_click(posterior,letter),
-#          entropy = get_posterior_entropy(posterior),
-#          p_click_rank = get_p_click_rank(posterior,letter));
+#          p_click = get_p_letter(posterior,letter),
+#          entropy = get_letter_posterior_entropy(posterior),
+#          p_click_rank = get_p_letter_rank(posterior,letter));
 
@@
 #   mutate(letter = get_random_letter(asked),
-#          p_click = get_p_click(posterior,letter),
-#          p_click_rank = get_p_click_rank(posterior,letter),
+#          p_click = get_p_letter(posterior,letter),
+#          p_click_rank = get_p_letter_rank(posterior,letter),
 #          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
 #          test_part='random');
+
+# save(E4.click_log_with_boards,
+#      E4.click_log_with_boards_random,
+#      E4.click_log_with_boards_most_frequent,
+#      E4.letter_prob_click_log_with_boards,
+#      E4.letter_prob_click_log_with_boards_random,
+#      file='E4.click_log_with_boards.RData')
 
@@
 
+```{r hangman-h7-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.mean_P_click_rank %>%
+    filter(test_part != 'optimal') %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="black", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+```
 ### Hypothesis 7b (Letter frequency)
@@
 This was the case also when focusing on misses only ( pretend first: `r E4.mean_letter_prob_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; first block only: `r t.test(E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
+
+```{r hangman-h7b-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.mean_letter_prob_P_click_rank %>%
+    filter(test_part != 'optimal') %>%
+    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="black", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+```
 
@@
 
-### Hypothesis 8(Entropy/RT relationship)
+### Hypothesis 8 (Entropy/RT relationship)
 
@@
 The effects were significantly weaker in pretend games (pretend-first group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & !genuine_first)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+
+```{r hangman-h8-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.click_log_with_boards %>%
+  filter(click_number>1) %>%
+  mutate(entropy_bin=round(entropy*3)/3) %>%
+  group_by(test_part,subj_id,entropy_bin) %>% 
+  summarise(RT=median(RT)) %>%
+  group_by(test_part,entropy_bin) %>%
+  summarise(median_RT=median(RT),
+            sem_RT=bootstrap_error(RT,N_perm)) %>%
+  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
+  theme_classic()
+
+```
 
@@
 #          used_for_nonpretend = word_used_for_nonpretend(subj_id,word),
+#          subj_id = as.numeric(substr(subj_id,1,3)),
 #          common_board = used_for_pretend & used_for_nonpretend,
@@
 E4.pretend_performance <- E4.judge_df %>%
-  filter(!(player %in% unidentifiable_ss & pretend==1)) %>%
+  mutate(pretend = correct_response=='P')%>%
+  filter(!(player %in% unidentifiable_ss) & pretend) %>%
   group_by(player)%>%
@@
 E4.optimality_cost <- E4.click_log_with_boards %>%
+  mutate(subj_id = as.numeric(substr(subj_id,1,3))) %>%
+  filter(!(subj_id %in% unidentifiable_ss)) %>%
   group_by(subj_id,test_part) %>%
@@
 E4.performance_df <- merge(E4.judge_performance,
-                           E4.pretend_performance) %>%
-  merge(E4.optimality_cost);
+                           E4.pretend_performance %>%
+                             dplyr::select(subj_id,pretend_rate)) %>%
+  merge(E4.optimality_cost, by='subj_id');
 
@@
 
-## Hypothesis 10 (judge-pretend relation)
+### Hypothesis 10 (judge-pretend relation)
 
 No significant correlation between judge accuracy and ability to trick others (`r cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman')%>%apa_print()%>%'$'(full_result)`), nor with the optimality cost to pretending (`r cor.test(E4.performance_df$optimality_cost,E4.performance_df$judge_rate,method='spearman')%>%apa_print()%>%'$'(full_result)`).
+
+```{r hangman-h10-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+# E4.performance_df %>%
+#   group_by(judge_rate) %>%
+#   summarise(mean_pretend_rate = mean(pretend_rate),
+#             se_pretend_rate=se(pretend_rate)) %>%
+#   ggplot(aes(x=judge_rate, y=mean_pretend_rate,
+#          ymin=mean_pretend_rate-se_pretend_rate,
+#          ymax=mean_pretend_rate+se_pretend_rate)) +
+#   geom_point(size=3) +
+#   geom_errorbar()
+
+E4.performance_df %>%
+  ggplot(aes(x=pretend_rate, y=judge_rate)) +
+  geom_point() +
+  labs(x= 'successful pretense rate', y='judge accuracy') +
+  theme_classic()
+
+```
 
@@
   group_by(word) %>% 
-  summarize(posterior = get_posterior(word_state[1],asked[1],category[1]),
+  summarize(posterior = get_letter_posterior(word_state[1],asked[1],category[1]),
             word_state=word_state[1])
@@
 
-## Hypothesis 12 (irrational letter selections)
+### Hypothesis 12 (irrational letter selections)
 
@@
 
-## Hypothesis 13 (target word effect)
+```{r hangman-h12-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+N_perm <- 1000;
+bootstrap_error_binomial <- function(x, N_perm) {
+  N <- length(x)
+  means = c();
+  for (i in 1:N_perm) {
+    means = c(means,sample(x,replace=TRUE,size=N)%>%mean())
+  };
+  return(sd(means))
+}
+
+E4.first_click_half_games %>%
+  group_by(test_part) %>%
+  summarise(se = bootstrap_error_binomial(makes_sense,N_perm),
+            makes_sense=mean(makes_sense)) %>%
+  mutate(test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend')) %>%
+  ggplot(aes(x=test_part,fill=test_part,y=makes_sense))+
+  geom_bar(position=position_dodge(),
+           stat='identity', color='black')+
+  geom_errorbar(aes(ymax = makes_sense+se, ymin=makes_sense-se))+
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+  theme_classic()+
+  labs(y='proportion of clicks that make sense')
+
+```
+
+### Hypothesis 13 (target word effect)
 
@@
 
-# Discussion
+```{r hangman-h13-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.first_click_half_games %>%
+  filter(makes_sense)%>%
+  mutate(word_class = ifelse(word %in% c('banana','hand'),'BANANA or HAND','PAPAYA or HAIR'))%>%
+  group_by(test_part,word_class) %>%
+  summarise(se = bootstrap_error_binomial(bananahand,N_perm),
+            bananahand= mean(bananahand)) %>%
+  mutate(test_part=ifelse(test_part=='pretend_half_game','pretend','non-pretend')) %>%
+  ggplot(aes(x=test_part,fill=test_part,y=bananahand))+
+  geom_bar(position=position_dodge(),
+           stat='identity', color='black')+
+  geom_errorbar(aes(ymax = bananahand+se, ymin=bananahand-se))+
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+  theme_classic()+
+  facet_grid(col=vars(word_class))+
+  labs(y='(consistent with BANANA/HAND)/(cons. with something)')
+
+```
 

### 2023-01-27 analysis/clickEntropyBattleship.Rmd
pre-registered analysis and some exploratory analysis

--- 2023-01-19 analysis/decisionConformityBattleship.Rmd
@@
----
-title: "Decision conformity: Battleship"
-author: "Matan Mazor"
-date: "2023-01-18"
-output: html_document
----
-
-These scripts can be run after running PTKN.Rmd from docs, because they use objects and functions that are defined there.
-
-Here I compare subjects' tendency to make 'conventional' clicks (that is, clicks that are often made by participants at this stage of the game), in pretend and non-pretend games.
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
-library(groundhog)
-
-groundhog.library(c(
-  'png',
-  'grid',
-  'ggplot2',
-  'svglite',
-  'xtable',
-  'papaja',
-  'tidyverse',
-  'broom',
-  'cowplot',
-  'MESS', # for AUCs
-  'lsr', # for effect sizes
-  'pwr', # for power calculations
-  'brms', # for mixed effects modeling
-  'BayesFactor', # for Bayesian t test
-  'jsonlite', # parsing data from sort_trial
-  'caret' #for cross validation
-), '2022-12-01')
-
-load('../.Rdata')
-```
-
-## Decision entropy
-
-Unlike the main analysis, where we look at the likelihood of hitting a ship, here we look at the prior of clicking on a cell, regardless of possible ships' locations.
-
-```{r get-priors, echo=FALSE, cache=TRUE}
-
-# What  I refer to as 'prior' here is the overall tendency of participants to
-# click on some cells. To make sure this quantity is not driven by the content
-# of the board itself, for each board I compute this quantity across all other
-# boards, not including itself.
-E2.prior_from_group <- E2.click_log_with_boards %>% 
-  filter(test_part=='nonpretend') %>% # we are interested in true games
-  mutate(flat_position = factor(flat_position, levels=seq(25))) %>%
-  group_by(grid_number,click_number, flat_position) %>%
-  summarise(n=n())%>% # count the number of clicks per cell per board per timepoint
-  group_by(click_number, flat_position)%>%
-  ## for each grid, sum across all other grids.
-  do( data.frame(.,other_n=sapply(.$grid_number, function(i) sum(.$n[!.$grid_number %in% i])))) %>%
-  group_by(grid_number, click_number) %>%
-  mutate(p=other_n/sum(other_n)) %>% # normalize to get a probability function
-  dplyr::select(grid_number,click_number,flat_position,p) %>%
-  spread(flat_position,p, fill=0, sep='_') %>% # pivot to wider format
-  ungroup();
-
-```
-
-```{r decision-entropy-functions, echo=FALSE, cache=TRUE}
-
-get_group_prior <- function(grid_i, click_i, board_state) {
-
-  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
-
-  hidden_cells = ifelse(is.na(board_state),1,0); #only hidden cells can be clicked on
-
-  group_prior = E2.prior_from_group %>%
-    filter(grid_number==grid_i & click_number==click_i) %>%
-    select(flat_position_1:flat_position_25) %>%
-    as.numeric()
-
-  group_prior_for_hidden_cells = group_prior*hidden_cells/sum(group_prior*hidden_cells)
-
-  return(paste(group_prior_for_hidden_cells,collapse=','))
-
-};
-
-## the next three functions are exactly the same as the corresponding functions for
-## the posterior, with the variable name posterior replaced with prior. I did 
-## this just to avoid confusion later.
-
-get_prior_click <- function(prior, flat_position) {
-  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
-  return(prior[flat_position])
-}
-
-get_prior_click_rank <- function(prior, flat_position) {
-  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
-  rank_prior = rank(-prior)
-  return(rank_prior[flat_position])
-}
-
-get_prior_entropy <- function(prior) {
-  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
-  prior[which(prior==0)]=NA; #to avoid infinity*0
-  entropy <- -sum(unlist(prior)*(unlist(log(prior))),na.rm=TRUE);
-  return(entropy);
-}
-
-```
-
-```{r p_click, echo = FALSE, cache=TRUE}
-
-E2.click_log_with_boards_and_prior <-E2.click_log_with_boards %>%
-  rowwise()%>%
-  mutate(prior = get_group_prior(grid_number, click_number, board_state),
-         prior_click = get_p_click(prior,flat_position),
-         prior_entropy = get_prior_entropy(prior),
-         prior_click_rank = get_prior_click_rank(prior,flat_position));
-
-E2.click_log_with_prior_optimal <- E2.click_log_with_boards_optimal %>%
-  mutate(prior = get_group_prior(grid_number, click_number, board_state),
-         prior_click = get_p_click(prior,flat_position),
-         prior_entropy = get_prior_entropy(prior),
-         prior_click_rank = get_prior_click_rank(prior,flat_position),
-         test_part=='optimal');
-
-E2.click_log_with_prior_random <- E2.click_log_with_boards_random %>%
-  mutate(prior = get_group_prior(grid_number, click_number, board_state),
-         prior_click = get_p_click(prior,flat_position),
-         prior_entropy = get_prior_entropy(prior),
-         prior_click_rank = get_prior_click_rank(prior,flat_position),
-         test_part = 'random')
-
-E2.mean_prior_click_rank <- E2.click_log_with_boards_and_prior %>%
-  dplyr::select(subj_id,test_part,genuine_first,prior_click_rank)%>%
-  rbind(E2.click_log_with_prior_optimal %>% ungroup() %>%
-          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank)) %>%
-  rbind(E2.click_log_with_prior_random %>% ungroup() %>%
-          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank)) %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(prior_click_rank=mean(prior_click_rank));
-
-E2.mean_prior_click_rank_misses_only <- E2.click_log_with_boards_and_prior %>%
-  dplyr::select(subj_id,test_part,genuine_first,prior_click_rank, hit_bin)%>%
-  rbind(E2.click_log_with_prior_optimal %>% ungroup() %>%
-          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank, hit_bin)) %>%
-  rbind(E2.click_log_with_prior_random %>% ungroup() %>%
-          dplyr::select(subj_id,test_part,genuine_first,prior_click_rank, hit_bin)) %>%
-  rowwise() %>%
-  filter(!hit_bin) %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(prior_click_rank=mean(prior_click_rank));
-
-p<- E2.mean_prior_click_rank %>%
-    filter(test_part != 'optimal') %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=prior_click_rank,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(click here) - rank', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    # facet_wrap(~genuine_first, nrow=2)+
-    # scale_x_reverse() +
-    theme_classic();
-
-ggsave('../docs/figures/E2_prior_click_rank.pdf',p)
-ggsave('../docs/figures/E2_prior_click_rank.png',p,width=5,height=2,dpi=300)
-
-p<- E2.mean_prior_click_rank_misses_only %>%
-    filter(test_part != 'optimal') %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=prior_click_rank,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
-    scale_x_continuous(limits=c(0,15))+
-    theme(legend.position=c(0.10,0.85)) +
-    # facet_wrap(~genuine_first, nrow=2)+
-    # scale_x_reverse() +
-    theme_classic();
-
-ggsave('../figures/E2_pclick_rank_misses_only.pdf',p)
-ggsave('../figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
-
- E2.RT_prior_click_cor <- E2.click_log_with_boards_and_prior %>%
-     filter(click_number<9 & click_number>1) %>%
-     group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(RT,prior_click,method='spearman'));
-
- E2.RT_prior_entropy_cor <- E2.click_log_with_boards_and_prior %>%
-     filter(click_number<9 & click_number>1) %>%
-     group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(prior_entropy,RT,method='spearman'))
-
-```
-The mean prior rank of non-pretend cell selections was `r E2.mean_prior_click_rank%>%filter(test_part=='nonpretend')%>%pull(prior_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (see Fig.
-2B; `r E2.mean_prior_click_rank%>%filter(test_part=='random')%>%pull(prior_click_rank)%>%mean()%>%printnum()`, `r E2.mean_prior_click_rank%>%spread(test_part,prior_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Pretend games were significantly less stereotypical than non-pretend games (`r E2.mean_prior_click_rank%>%filter(test_part=='pretend')%>%pull(prior_click_rank)%>%mean()%>%printnum()`; `r apa_print(t.test(E2.mean_prior_click_rank%>%filter(genuine_first&test_part=='nonpretend')%>%pull(prior_click_rank), E2.mean_prior_click_rank%>%filter(!genuine_first&test_part=='pretend')%>%pull(prior_click_rank)))$statistic`), but still more optimal than those of a random agent (`r E2.mean_prior_click_rank%>%spread(test_part,prior_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_prior_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(prior_click_rank), E2.mean_prior_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(prior_click_rank)))$statistic`; pretend - random: `r E2.mean_prior_click_rank_misses_only%>%spread(test_part,prior_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-
-```{r prior-entropy-RT, echo=FALSE, cache=TRUE}
-
-E2.RT_by_prior_entropy <- E2.click_log_with_boards_and_prior %>%
-  filter(click_number>1) %>%
-  mutate(entropy_bin=round(prior_entropy*2)/2) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(test_part,subj_id,entropy_bin) %>%
-  summarise(RT=median(RT),
-            corrected_RT = median(corrected_RT)) %>%
-  group_by(test_part,entropy_bin) %>%
-  summarise(median_RT=median(RT),
-            sem_RT=bootstrap_error(RT,N_perm),
-            median_corrected_RT=median(corrected_RT),
-            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))
-
-p <- E2.RT_by_prior_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of prior', y='median RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_prior_entropy_no_split.png',p,width=5,height=2,dpi=300)
-# ggsave('../figures/E2_RT_by_prior_entropy_no_split.svg',p,width=5,height=2,dpi=300)
-
-p <- E2.RT_by_prior_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of prior', y='median residual RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_prior_entropy_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.RT_by_prior_entropy_coefs <- E2.click_log_with_boards_and_prior %>%
-  filter(click_number>1) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  mutate(entropy_centered = prior_entropy-mean(prior_entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value)) %>%
-  group_by(genuine_first,subj_id,test_part,term) %>%
-  summarise(estimate = mean(estimate))
-
-E2.RT_by_prior_entropy_coefs_summary <- E2.RT_by_prior_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.RT_by_prior_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.corrected_RT_by_prior_entropy_coefs <- E2.click_log_with_boards_and_prior %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  mutate(entropy_centered = prior_entropy-mean(prior_entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value))
-
-E2.corrected_RT_by_prior_entropy_coefs_summary <- E2.corrected_RT_by_prior_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.corrected_RT_by_prior_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_prior_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-```
-
-```{r entropy-by-erial-position, echo=FALSE, cache=TRUE}
+```{r entropy-by-serial-position, echo=FALSE, cache=TRUE}
 
@@
 
+E2.click_entropy_by_serial_position_wide <- E2.click_entropy_by_serial_position %>%
+  spread(test_part,entropy) %>% 
+  mutate(diff=pretend-nonpretend)
+
 E2.click_entropy_by_serial_position_summary <- E2.click_entropy_by_serial_position %>%
@@
 ```
+
+```{r entropy-pclick-correlation, echo=FALSE, cache=TRUE}
+
+E2.entropy_summary <- E2.click_entropy_by_serial_position %>%
+  filter(click_number<8) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(entropy=mean(entropy,na.rm=T))
+
+E2.p_click_summary <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_random) %>%
+  ungroup() %>%
+  filter(click_number<8) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+E2.entropy_cost <- E2.p_click_summary %>%
+  merge(E2.entropy_summary)%>%
+  group_by(subj_id)%>%
+  summarise(p_click_cost = p_click_rank[test_part=='pretend']-
+              p_click_rank[test_part=='nonpretend'],
+            entropy_cost = entropy[test_part=='pretend']-
+              entropy[test_part=='nonpretend'])
+
+```

### 2023-01-27 analysis/clickEntropyHangman.Rmd
pre-registered analysis and some exploratory analysis

```{r entropy-by-serial-position, echo=FALSE, cache=TRUE}

#compute Shannon entropy, but only until the first timepoint at which some gamed ended
entropy <- function(target) {
  if (length(target)<5) {
    return(NaN)
  } else {
    freq <- table(target)/length(target)
    # vectorize
    vec <- as.data.frame(freq)[,2]
    #drop 0 to avoid NaN resulting from log2
    vec<-vec[vec>0]
    #compute entropy
    return(-sum(vec * log2(vec)))
  }
}

e4.click_entropy_by_serial_position <- e4.click_log_with_boards %>%
  mutate(flat_position = which(LETTERS==letter)) %>%
  # filter(click_number<15) %>%
  group_by(subj_id,test_part,click_number)%>%
  summarise(entropy=entropy(flat_position))

e4.click_entropy_by_serial_position_summary <- e4.click_entropy_by_serial_position %>%
  group_by(test_part,click_number) %>%
  summarise(mean_entropy=mean(entropy, na.rm=T),
            se_entropy=se(entropy, na.rm=T)) %>%
  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));

p <- e4.click_entropy_by_serial_position_summary %>%
  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
  geom_line() +
  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
  labs(x='click number', y='entropy') +
  theme_classic()

ggsave('../docs/figures/e4_entropy_by_click_nnumber.png',p,width=7,height=5,dpi=300)

```

### 2023-01-27 analysis/decisionConformityBattleship.Rmd
pre-registered analysis and some exploratory analysis

@@
 ```
-
-```{r entropy-by-erial-position, echo=FALSE, cache=TRUE}
-
-#compute Shannon entropy, but only until the first timepoint at which some gamed ended
-entropy <- function(target) {
-  if (length(target)<5) {
-    return(NaN)
-  } else {
-    freq <- table(target)/length(target)
-    # vectorize
-    vec <- as.data.frame(freq)[,2]
-    #drop 0 to avoid NaN resulting from log2
-    vec<-vec[vec>0]
-    #compute entropy
-    return(-sum(vec * log2(vec)))
-  }
-}
-
-E2.click_entropy_by_serial_position <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal)%>%
-  rbind(E2.click_log_with_boards_random)%>%
-  # filter(click_number<15) %>%
-  group_by(subj_id,test_part,click_number)%>%
-  summarise(entropy=entropy(flat_position))
-
-E2.click_entropy_by_serial_position_summary <- E2.click_entropy_by_serial_position %>%
-  group_by(test_part,click_number) %>%
-  summarise(mean_entropy=mean(entropy, na.rm=T),
-            se_entropy=se(entropy, na.rm=T)) %>%
-  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));
-
-p <- E2.click_entropy_by_serial_position_summary %>%
-  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
-  geom_line() +
-  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
-  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
-  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
-  labs(x='click number', y='entropy') +
-  theme_classic()
-
-ggsave('../docs/figures/E2_entropy_by_click_nnumber.png',p,width=7,height=5,dpi=300)
-
-```
-
-```{r entropy-pclick-correlation, echo=FALSE, cache=TRUE}
-
-E2.entropy_summary <- E2.click_entropy_by_serial_position %>%
-  filter(click_number>8) %>%
-  group_by(subj_id,test_part) %>%
-  summarise(entropy=mean(entropy,na.rm=T))
-
-E2.p_click_summary <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_random) %>%
-  ungroup() %>%
-  filter(click_number<8) %>%
-  group_by(subj_id,test_part) %>%
-  summarise(p_click=mean(p_click))
-
-E2.p_click_summary %>%
-  merge(E2.entropy_summary) %>%
-  ggplot(aes(x=entropy,y=p_click, color=test_part))+
-  geom_point(size=2,alpha=0.8)
-
-```

### 2023-01-27 analysis/shipCompletion.Rmd
pre-registered analysis and some exploratory analysis

--- 2023-01-04 docs/cogsci/PNTKN.Rmd
@@
----
-title: "Using pretense behavior to explore counterfactual self-simulation"
-bibliography: r-references.bib
-csl: apa6.csl
-document-params: "10pt, letterpaper"
-
-author-information: > 
-    \author{{\large \bf Matan Mazor (mtnmzor@gmail.com)} \\ Birkbeck, University of london \\ London, WC1E 7HX UK
-    \AND {\large \bf Chaz Firestone (chaz@jhu.edu)}
-    \AND {\large \bf Ian Phillips (ianphillips@jhu.edu)} \\ Johns Hopkins University \\ Baltimore, MD 21218, USA}
-
-abstract: >
-   What we do depends on what we know.
-But sometimes we try to decouple our behavior from our knowledge so that we appear not to know what we really do know.
-Such pretense behavior requires understanding how we would behave with different knowledge — and so provides a window into counterfactual self-simulation.
-However, little research has characterized and evaluated pretense relative to non-pretense behavior.
-Here, in a large-scale, pre-registered experiment, subjects played normal games of Battleships (trying to sink ships hidden in a grid), as well as ‘pretend’ games, where they were told all the ships’ locations but had to pretend they were playing normally.
-Relative to normal games, ‘pretend’ games demonstrated similar but exaggerated behavioral patterns.
-Furthermore, pretenders played rationally, but less so than non-pretenders.
-Despite these differences, ‘judge’ participants could not detect ‘pretend’ games.
-We discuss implications of these findings for accounts of theory of mind and metacognition.
-
-keywords: >
-    metacognition; theory of mind; pretense
-
-output: cogsci2016::cogsci_paper
-
-final-submission: \cogscifinalcopy
----
-
-```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
-                      fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, 
-                      message=F, sanitize = T)
-```
-
-```{r, libraries, echo=FALSE, warning=FALSE}
-
-library(groundhog)
-
-groundhog.library(c(
-  'png',
-  'grid',
-  'ggplot2',
-  'svglite',
-  'xtable',
-  'papaja',
-  'tidyverse',
-  'broom',
-  'cowplot',
-  'MESS', # for AUCs
-  'lsr', # for effect sizes
-  'pwr', # for power calculations
-  'brms', # for mixed effects modeling
-  'BayesFactor', # for Bayesian t test
-  'jsonlite', # parsing data from sort_trial
-  'caret' #for cross validation
-), '2022-12-01')
-
-```
-
-# Introduction
-
-The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory].
-This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation].
-This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own.
-
-Moreover, deception often involves pretense behavior, which in turn relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state.
-For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") — you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
-What would I say?
-How long would it take me to recover from the surprise?").
-This reliance of pretense behavior on self-simulation makes it an ideal opportunity to examine metacognitive knowledge about one's own mental states, and the potential reliance of this knowledge on a self-simulation.
-By comparing non-pretend and pretend behavior, we can ask which aspects of their own cognitive processes subjects can and cannot effectively simulate, and which aspects are not well-represented in their mental models of their own cognition.
-
-To this end, here we examine pretense in a game setting.
-Using an online version of the game Battleships, participants played a 'non-pretend' (normal) version of the game (trying to sink ships whose locations are unknown), as well as a 'pretend' version where they were given secret information about the location of hidden ships, but were instructed to behave as if they didn't have this information.
-
-# Methods
-
-A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb)
-
-```{r load-data, warning=FALSE}
-E2.df <- read.csv('../../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
-  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,4),sep='')) %>%
-  mutate(subj_id = factor(subj_id)) %>%
-  # the dash is breaking r
-  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
-  mutate(genuine_first=genuine_first=='True');
-
-E2.export <- read.csv('../../experiments/Battleships2/data/batch1/prolific_export_batch1.csv',na.strings=c(""," ","NA"));
-
-E2.bonus <- read.csv('../../experiments/Battleships2/data/batch1/bonus.csv',na.strings=c(""," ","NA"))
-
-```
-
-## Participants
-
-The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
-500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
-They were selected based on their acceptance rate (>95%) and for being native English speakers.
-The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
-
-## Procedure
-
-Participants were first instructed that the experiment, based on the game Battleships, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment.
-They were then presented with a leaderboard of previous players, and given the rules of the game:
-
-> "In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
-What makes the game difficult is that you can’t see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
-To sink a ship, you need to click on all of the squares it is located in.
-If you hit part of a ship, the square will turn red.
-If there is no ship in the square, it will turn blue."
-
-We further explained that in this version of the game, ships can touch corners, but their sides can't touch.
-This explanation was accompanied by a visual presentation of legal and illegal ship configurations.
-
-After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half game (see below for details).
-The order of pretend and non-pretend blocks was counterbalanced between participants.
-The allocation of boards (spatial configurations of ships; see Fig.
-1A) to conditions was randomized between participants such that exactly one board was played in both pretend and non-pretend conditions, and this common board was different for different participants.
-The order of boards within a block was fully randomized, with the exception that half-games were always played last.
-
-### Non-pretend (normal) games
-
-In non-pretend games (Fig.
-1B), participants aimed to sink two 2-square patrol boats and one 3-square submarine with as few clicks as possible.
-An online counter of the number of clicks was displayed on the screen.
-After each game, feedback was given about the number of clicks and resulting number of points obtained.
-
-### Pretend games
-
-Participants in pretend games were given the same explanation of Battleships, and played a practice round. However, they were then given an additional twist:
-
-> "This time your goal is different.
-In this round, we’re going to tell you where the ships are, but **we want you to act like you don’t know this information**.
-We’ve marked the ships’ locations with a cross, so you’ll know where they are the whole time; but your job is to play the game as if these hints aren’t there.
-To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are.
-We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you'll get bonus points.
-But if your games look different, you won't get these bonus points.
-Your number of clicks in this part will not affect your bonus.
-Only your ability to play like you had no hints."
-
-We informed participants that both the location and timing of their cell clicks will be measured.
-After one practice round and one comprehension question, participants played five pretend games (Fig.
-1C), followed by one pretend half-game.
-Each game was followed by a short message, reminding them that a game that looks similar to the games of participants who had no hints would be awarded 10 bonus points.
-
-### Half games
-
-In order to directly compare participants' pretend and non-pretend games for identical belief states (genuine or pretended ignorance about where the ships are hidden), participants completed one pretend and one non-pretend game given a partly finished board with the content of 7 cells already revealed (see Fig.
-1D).
-We designed our half games to produce a strong expectation to find a ship in specific cells, but not in others.
-The assignment of half-completed boards to pretend and non-pretend conditions was randomized between participants.
-
-### Judge trials
-
-In the final part of the experiment, participants observed the games of previous players and tried to determine who had hints and who didn't.
-On each trial, two empty grids were presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid (Fig.
-1F).
-The two grids corresponded to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders.
-Only games shorter than one minute were chosen for presentation in this part.
-For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
-Judge participants observed a real time replay of the two grids, showing not only where participants clicked, but also when.
-After making a decision, participants were informed whether they would receive the 10 bonus points, or alternatively, whether the pretender would receive them in the event the pretender managed to trick them.
-
-<!-- 500 players played 12 games of the game Battleships in two conditions presented in two blocks.
-The order of conditions was randomized between players.
-the standard, *non-pretend* condition, players started the game not knowing where the ships were and tried to sink all ships with as few clicks as possible.
-In the *pretend* condition, ship locations were marked with a cross such that players knew where they were hidden, and their task was to play the game as if they did not have this information.
-The sixth game in each block was a 'half-game', where participants completed the game with the content of some cells revealed by a previous player.
-Lastly, in judge trials, players observed a replay of the games of two previous players – one pretender and one non-pretender, and their task was to identify which was which.
-We created nine different 5x5 boards, each containing two 2-square patrol boats and one 3-square submarine.
-The assignment of boards to conditions was randomized across participants such that exactly one board was presented in both pretend and non-pretend blocks.
-(see Fig.
-1).
--->
-
-A more detailed description of the study procedure is provided in the study pre-registration document (LINK HIDDEN).
-Readers are also invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/NervzpM0Y0z).
-
-```{r design, echo=FALSE, fig.cap="Experimental Design. See Methods for details.", out.width = '100%'}
-knitr::include_graphics("../figures/Battleships_design.png")
-```
-
-# Results
-
-```{r subject-comments, echo=FALSE}
-E2.comments <- E2.df %>%
-  filter(!is.na(responses) & is.na(test_part)) %>%
-  dplyr::select(subj_id,responses) %>%
-  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
-  unnest(comments)%>%
-  dplyr::select(subj_id,worker_comments)%>%
-  filter(!is.na(worker_comments) & worker_comments!="") %>%
-  mutate(worker_comments = gsub("[\r\n]", "",worker_comments))
-
-E2.debrief <- E2.df %>%
-  filter(!is.na(responses) & is.na(test_part)) %>%
-  dplyr::select(subj_id,responses) %>%
-  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
-  unnest(comments)%>%
-  dplyr::select(subj_id,debrief)%>%
-  filter(!is.na(debrief) & debrief != "")%>%
-  mutate(debrief = gsub("[\r\n]", "",debrief))
-
-```
-
-We designed our analyses to explore subjects' capacity for self-simulation under a counterfactual knowledge state, and the limits of this capacity.
-We focused on where subjects clicked and when, and asked whether this differed between pretend and non-pretend games.
-All analyses were pre-registered unless otherwise specified.
-In our pre-registration document, we committed to separately analyzing participants according to whether they pretended before or after completing a non-pretend block.
-Due to space limitations, and since the results of the two groups mostly agreed, we report here the pooled results from both groups of participants, and mention whenever we found different results depending on block order.
-When directly comparing pretend and non-pretend blocks, we perform a between-subjects comparison using data from the first block only, i.e., pretend games from participants who pretended in the first block and non-pretend games from participants who played normally in the first block.
-We do so to ensure that any successful pretending is not due to memory of one's own behavior in a previous block, and that non-pretend games are not biased by experience with the pretend block.
-
-```{r num_clicks, echo=FALSE, cache=TRUE}
-
-E2.click_df <- E2.df %>%
-  dplyr::select(subj_id, 
-         test_part, 
-         grid_number, 
-         num_clicks,
-         click_log,
-         genuine_first) %>%
-  mutate(click_log = gsub("\'","\"", click_log)) %>%
-  filter(test_part=='pretend' | test_part=='nonpretend');
-
-E2.median_clicks <- E2.click_df %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(test_part,num_clicks);
-
-E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();
-
-E2.median_clicks_filtered <- E2.click_df %>%
-  filter(!(subj_id %in% E2.lucky_pretenders))%>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(test_part,num_clicks);
-
-p<- E2.click_df %>%
-    group_by(subj_id,test_part, genuine_first) %>%
-    summarise(num_clicks=median(num_clicks)) %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=num_clicks,fill=test_part))+
-    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='median number of clicks', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    theme_classic()
-    # geom_vline(xintercept= median_simulated)+
-    # facet_wrap(~genuine_first, nrow=2);
-
-ggsave('../docs/figures/E2_num_clicks.png',p,width=5,height=2, dpi=300)
-
-```
-
-```{r useful_functions, echo=FALSE, cache=TRUE}
-
-all_possible_boards <- read.csv('../../analysis/flat_boards.csv',header=FALSE);
-
-create_board_states <- function(flat_positions,hit) {
-
-  board_state = rep(NA,25)
-  board_states = c();
-  for (p in seq_along(flat_positions)) {
-    board_states = c(board_states, paste(board_state,collapse=','));
-    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
-  }
-
-  return(board_states)
-};
-
-get_likelihood <- function(board_state) {
-
-  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
-
-  revealed_positions <- which(!is.na(board_state));
-
-  boards_to_remove = c();
-
-  for (position in revealed_positions) {
-    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
-  }
-
-  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
-  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
-  likelihood[revealed_positions]=NA;
-
-  return(paste(likelihood,collapse=','))
-
-};
-
-get_posterior <- function(likelihood) {
-    likelihood = scan(text= likelihood, what = numeric(), sep="," , quiet = TRUE);
-    posterior = likelihood/sum(likelihood,na.rm=TRUE); # assuming a uniform prior
-    return(paste(posterior,collapse=','))
-}
-
-get_p_click <- function(posterior, flat_position) {
-  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  return(posterior[flat_position])
-}
-
-get_p_click_rank <- function(posterior, flat_position) {
-  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  rank_posterior = rank(-posterior)
-  return(rank_posterior[flat_position])
-}
-
-get_posterior_entropy <- function(posterior) {
-  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  posterior[which(posterior==0)]=NA; #to avoid infinity*0
-  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
-  return(entropy);
-}
-
-```
-
-```{r simulate_nclicks, echo=FALSE, cache=TRUE}
-
-## simulate number of clicks for random play
-
-num_clicks_sim = c();
-nsim = 10000
-boards = rep(NaN,25)
-
-for (i in 1:nsim) {
-  board = rep(c(0),25);
-  while (sum(board)<7) {
-    board[sample(1:25,1)]=1
-  }
-  nc = which(board==1)%>%tail(n=1);
-  num_clicks_sim = c(num_clicks_sim, nc);
-  if (nc<25) {
-    board[(nc+1):length(board)]=NaN;
-  }
-  boards = rbind(boards,board)
-}
-
-simulated_hits <- data.frame(click_number=1:25,mean_hit=colMeans(boards,na.rm=T))
-
-## OPTIMAL PLAYER
-
-simulate_optimal_game <- function(ship_positions) {
-  #initialize
-  board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-
-  while (nships>nhits) {
-
-    posterior <- get_likelihood(paste(board_state,collapse=',')) %>%
-      get_posterior() %>%
-      scan(text = ., what= numeric(), sep="," , quiet = TRUE);
-
-    best_shot = which.max(posterior);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    result = ship_positions[best_shot]
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    board_state[best_shot]=result;
-
-  }
-
-  return(nclicks)
-
-}
-
-num_clicks_optimal_sim = c();
-nsim = 100;
-nboards = all_possible_boards%>%nrow();
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  num_clicks_optimal_sim = c(num_clicks_optimal_sim,simulate_optimal_game(cur_board));
-}
-
-```
-
-## Total number of clicks
-
-To sink all ships, players had to click on at least 7 and at most 25 squares.
-A simulated player that clicks randomly had a median total click number of `r num_clicks_sim%>%median()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a median total click number of `r num_clicks_optimal_sim%>%median()`.
-Among our players, the median number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games (see Fig.
-2A).
-
-We observed no significant difference in the number of clicks between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend),E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
-However, a significant interaction with block order suggested that participants made fewer clicks in the second block compared to the first, regardless of which block was the pretend block (`r E2.median_clicks%>%gather('condition','nclicks',c('nonpretend','pretend'))%>%aov(nclicks ~ genuine_first * condition,data=.)%>%apa_print()%>%'$'(statistic)%>%'$'(genuine_first_condition)`).
-
-In `r E2.click_df%>%filter(num_clicks==7)%>%nrow()` pretend games from `r E2.lucky_pretenders%>%length()` players, games were completed after 7 clicks only, without ever missing a ship.
-This never happened in non-pretend games.
-We assumed that these participants did not follow task instructions, and excluded them from all following analyses.
-
-## Game optimality
-
-Pretend games were similar to non pretend games in the total number of clicks — but were they also similar in *where* participants clicked?
-More specifically, did cell selections in pretend games make sense given the limited information those participants pretended to have?
-To ask this, we approximated optimal behavior by calculating the probability that a ship is hidden in each cell given available information, $p(ship(x_{i}))$, and the posterior probability that one should click on a square, assuming a uniform prior over cells $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$.
-Critically, in modeling pretend games we did not treat hints as part of this available information for extracting $p(ship(x_{i}))$, because an optimal player should ignore hints in choosing where to click next.
-Given this posterior map, a rational player should choose cells where $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3].
-To quantify optimality, before each cell selection we computed the posterior probability map for all 'unknown' cells.
-Then, we ranked cells from high to low according to their posterior probability and recorded the rank of the chosen cell: a lower rank indicating more optimal behavior.
-
-```{r parse-click-log, echo-FALSE, cache=TRUE}
-
-E2.click_log <- readRDS('../../experiments/Battleships2/data/batch1/click_log.Rda');
-
-# E2.click_log <- data.frame(matrix(ncol=10,nrow=0, 
-#                                dimnames=list(NULL, 
-#                                              c("subj_id",
-#                                               "test_part", 
-#                                               "grid_number", 
-#                                               "num_clicks",
-#                                               "i",
-#                                               "j",
-#                                               "hit",
-#                                               "t",
-#                                               "click_number",
-#                                               "genuine_first"))))
-# 
-# 
-# for (row in 1:nrow(E2.click_df)) {
-#   
-#     subject_click_log <- data.frame(fromJSON(E2.click_df[row, ]$click_log)) %>%
-#     mutate(
-#       click_number = 1:n(),
-#       subj_id = E2.click_df[row, ]$subj_id,
-#       test_part = E2.click_df[row, ]$test_part,
-#       grid_number = E2.click_df[row, ]$grid_number,
-#       num_clicks = E2.click_df[row, ]$num_clicks,
-#       genuine_first = E2.click_df[row, ]$genuine_first
-#     )
-#     
-#     E2.click_log <- rbind(E2.click_log, subject_click_log);
-# }
-#     
-# 
-# E2.click_log <- E2.click_log %>%
-#   relocate(subj_id, .before = i) %>%
-#   relocate(genuine_first, .before = i) %>% 
-#   relocate(test_part, .before=i) %>% 
-#   relocate(grid_number, .before=i) %>% 
-#   relocate(click_number, .before=i) %>%
-#   group_by(subj_id,test_part,grid_number) %>%
-#   mutate(RT=t-lag(t,default=0)) %>%
-#   filter(!(subj_id %in% E2.lucky_pretenders))%>%
-#  mutate(hit_bin = hit !=0,
-#         lag3 = lag(hit,3)!=0,
-#         lag2 = lag(hit,2)!=0,
-#         lag1 = lag(hit,1)!=0,
-#         lead1 = lead(hit,1)!=0,
-#         lead2 = lead(hit,2)!=0,
-#         lead3 = lead(hit,3)!=0)
-
-```
-
-```{r simualte_data, echo=FALSE, cache=TRUE}
-
-simulate_optimal_game <- function(ship_positions) {
-  #initialize
-  cur_board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-  board_state = c()
-  flat_position = c();
-  click_number = c();
-  hit_bin = c();
-
-  while (nships>nhits) {
-
-    board_state = c(board_state,paste(cur_board_state,collapse=','));
-
-    posterior <- get_likelihood(paste(cur_board_state,collapse=',')) %>%
-      get_posterior() %>%
-      scan(text = ., what= numeric(), sep="," , quiet = TRUE);
-
-    best_shot = which.max(posterior);
-    flat_position = c(flat_position,best_shot);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    click_number = c(click_number,nclicks)
-
-    result = ship_positions[best_shot]
-
-    hit_bin = c(hit_bin,ifelse(result==1,1,0));
-
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    cur_board_state[best_shot]=result;
-
-  }
-  return(data.frame(board_state,flat_position,hit_bin,click_number))
-
-}
-
-simulate_random_game <- function(ship_positions) {
-  #initialize
-  cur_board_state <- rep(NA,25); 
-  nhits <- 0;
-  nclicks <- 0;
-  nships <- sum(ship_positions);
-  board_state = c()
-  flat_position = c();
-  click_number = c();
-  hit_bin = c();
-
-  while (nships>nhits) {
-
-    board_state = c(board_state,paste(cur_board_state,collapse=','));
-
-    random_guess = which(is.na(cur_board_state))%>%sample(1);
-    flat_position = c(flat_position,random_guess);
-
-    #click on best shot
-    nclicks = nclicks+1;
-    click_number = c(click_number,nclicks)
-
-    result = ship_positions[random_guess]
-
-    hit_bin = c(hit_bin,ifelse(result==1,1,0));
-
-    if (result==1) {
-      nhits = nhits+1;
-    }
-    cur_board_state[random_guess]=result;
-
-  }
-  return(data.frame(board_state,flat_position,hit_bin,click_number))
-
-}
-
-nsim = 100;
-nboards = all_possible_boards%>%nrow();
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  if (i==1) {
-    simulated_optimal = simulate_optimal_game(cur_board)%>%mutate(game=i)
-  } else {
-    simulated_optimal = bind_rows(simulated_optimal, simulate_optimal_game(cur_board)%>%mutate(game=i))
-  }
-}
-
-for (i in 1:nsim) {
-  cur_board = all_possible_boards[sample(1:nboards,1),];
-  if (i==1) {
-    simulated_random = simulate_random_game(cur_board)%>%mutate(game=i)
-  } else {
-    simulated_random = bind_rows(simulated_random, simulate_random_game(cur_board)%>%mutate(game=i))
-  }
-}
-
-```
-
-```{r p_click, echo = FALSE, cache=TRUE}
-
-E2.click_log_with_boards <- readRDS('../../experiments/Battleships2/data/batch1/click_log_with_boards.Rda')
-
-get_random_flat_position <- function(board_state) {
-
-  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
-  return(which(is.na(board_state))%>%sample(1))
-
-}
-
-get_optimal_flat_position <- function(posterior) {
-
-  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
-  return(which.max(posterior))
-
-}
-
-E2.click_log_with_boards_optimal <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_optimal_flat_position(posterior),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'optimal')
-
-E2.click_log_with_boards_random <- E2.click_log_with_boards %>%
-  mutate(flat_position = get_random_flat_position(board_state),
-         p_click = get_p_click(posterior,flat_position),
-         p_click_rank = get_p_click_rank(posterior,flat_position),
-         test_part = 'random')
-
-# E2.click_log <-E2.click_log %>%
-#   dplyr::select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>% 
-#   mutate(flat_position = 1+i*5+j) %>%
-#   group_by(subj_id,genuine_first,test_part,grid_number) %>%
-#   arrange(click_number) %>%
-#   summarise(click_number = click_number,
-#             board_state=create_board_states(flat_position,hit_bin),
-#             flat_position=flat_position,
-#             i=i,
-#             j=j,
-#             RT=RT,
-#             hit_bin=hit_bin) %>%
-#   rowwise()%>%
-#   mutate(likelihood = get_likelihood(board_state),
-#          posterior = get_posterior(likelihood),
-#          p_click = get_p_click(posterior,flat_position),
-#          entropy = get_posterior_entropy(posterior),
-#          p_click_rank = get_p_click_rank(posterior,flat_position));
-
-E2.mean_P_click_rank <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal) %>%
-  rbind(E2.click_log_with_boards_random) %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(p_click_rank=mean(p_click_rank));
-
-E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
-  rbind(E2.click_log_with_boards_optimal) %>%
-  rbind(E2.click_log_with_boards_random) %>%
-  rowwise() %>%
-  filter(!hit_bin) %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(p_click_rank=mean(p_click_rank))
-
-p<- E2.mean_P_click_rank %>%
-    filter(test_part != 'optimal') %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=p_click_rank,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    # facet_wrap(~genuine_first, nrow=2)+
-    # scale_x_reverse() +
-    theme_classic();
-
-ggsave('../figures/E2_pclick_rank.pdf',p)
-ggsave('../figures/E2_pclick_rank.png',p,width=5,height=2,dpi=300)
-
-p<- E2.mean_P_click_rank_misses_only %>%
-    filter(test_part != 'optimal') %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=p_click_rank,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
-    scale_x_continuous(limits=c(0,15))+
-    theme(legend.position=c(0.10,0.85)) +
-    # facet_wrap(~genuine_first, nrow=2)+
-    # scale_x_reverse() +
-    theme_classic();
-
-ggsave('../figures/E2_pclick_rank_misses_only.pdf',p)
-ggsave('../figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
-
-E2.num_irrational_clicks <- E2.click_log_with_boards %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(nic=sum(p_click==0)/5);
-
-p<- E2.num_irrational_clicks %>%
-    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
-    ggplot(aes(x=nic,fill=test_part))+
-    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='mean number of irrational clicks per game', y='number of players')+
-    theme(legend.position=c(0.10,0.85)) +
-    facet_wrap(~genuine_first, nrow=2)+
-    theme_classic();
-
-ggsave('figures/../E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)
-
- E2.RT_p_click_cor <- E2.click_log_with_boards %>%
-     filter(click_number<9 & click_number>1) %>%
-     group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(RT,p_click,method='spearman'));
-
- E2.RT_entropy_cor <- E2.click_log_with_boards %>%
-     filter(click_number<9 & click_number>1) %>%
-     group_by(subj_id,test_part, genuine_first) %>%
-     summarise(r=cor(entropy,p_click,method='spearman'))
-
-```
-
-The mean posterior rank of non-pretend cell selections was `r E2.mean_P_click_rank%>%filter(test_part=='nonpretend')%>%pull(p_click_rank)%>%mean()%>%printnum()` and significantly lower (more optimal) than that of a simulated random agent (see Fig.
-2B; `r E2.mean_P_click_rank%>%filter(test_part=='random')%>%pull(p_click_rank)%>%mean()%>%printnum()`, `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Pretend games were significantly less optimal than non-pretend games (`r E2.mean_P_click_rank%>%filter(test_part=='pretend')%>%pull(p_click_rank)%>%mean()%>%printnum()`; `r apa_print(t.test(E2.mean_P_click_rank%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`), but still more optimal than those of a random agent (`r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-In other words, the optimality of pretend games relative to random cell selection was not merely due to the fact that pretenders clicked on ships more than expected by chance.
-Even when missing a ship, their cell selections made sense given the limited information they pretended to have.
-
-```{r numclick_optimality, echo=FALSE, fig.cap="Experimental Design", out.width = '100%', fig.cap = "A: The total number of clicks in pretend and non-pretend games was highly similar.
-B: Non-pretend games were significantly more optimal than pretend games, but both were more optimal than what is expected by random chance.
-Box: An illustration of the difference between success and optimality.
-A given click (highlighted with a black border) can reveal a ship (upper row) or not (lower row) regardless of whether the click makes sense given available information (left column) or not (right column).
-"}
-knitr::include_graphics("../figures/numclicks_optimality.png")
-```
-
-## Response time analysis
-
-```{r total_time, echo=FALSE}
-
-E2.total_time_df <- E2.click_log %>%
- filter(click_number==num_clicks)%>%
-  mutate(t=t/1000)
-
-E2.median_time <- E2.total_time_df %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(t=median(t)) %>%
-  spread(test_part,t);
-
-# anova
-E2.time_anva <- aov(t ~ genuine_first * condition, 
-                         data = E2.median_time %>%
-                           gather('condition','t',nonpretend,pretend) %>%
-                           mutate(condition=as.factor(condition)));
-
-E2.median_time_per_click <- E2.click_log %>%
-  group_by(subj_id,test_part, genuine_first) %>%
-  summarise(RT=median(RT)) %>%
-  spread(test_part,RT) %>%
-  mutate(diff=pretend-nonpretend);
-
-# anova
-E2.click_time_anva <- aov(RT ~ genuine_first * condition, 
-                         data = E2.median_time_per_click %>%
-                           gather('condition','RT',nonpretend,pretend) %>%
-                           mutate(condition=as.factor(condition)))
-```
-
-In general, participants gave slower responses when pretending compared to when playing normally.
-Despite the similar number of clicks per game, the median game duration was `r E2.total_time_df%>%filter(test_part=='nonpretend')%>%pull(t)%>%median()%>%printnum()` seconds in the non-pretend condition and `r E2.total_time_df%>%filter(test_part=='pretend')%>%pull(t)%>%median()%>%printnum()` seconds in the pretend condition, which was significantly longer (`r t.test(E2.median_time$pretend-E2.median_time$nonpretend)%>%apa_print()%>%'$'(statistic)`).
-This difference showed a significant interaction with the order of conditions, such that completing games in the first block took longer (`r apa_print(E2.time_anva%>%summary())$full_result$genuine_first_condition`).
-
-In our instructions to pretenders, we asked them to pretend not to know where the ships were, not only in *where* they chose to click, but also in *when* they clicked.
-In the following set of analyses, we identified patterns in click latency in non-pretend games, and asked whether the same patterns are also observed in pretend games.
-We focused on the effects of hitting versus missing a ship, decision uncertainty, and ship completion.
-
-```{r RT_by_accuracy}
-
-N_perm <- 1000;
-bootstrap_error <- function(x, N_perm) {
-  N <- length(x)
-  medians = c();
-  for (i in 1:N_perm) {
-    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
-  };
-  return(sd(medians))
-}
-
-E2.RT_by_lag3 <- E2.click_log %>%
-  filter(click_number>3) %>%
-  group_by(subj_id,test_part,lag3,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag3,RT,sep='_') %>%
-  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);
-
-E2.RT_by_lag2 <- E2.click_log %>%
-  filter(click_number>2) %>%
-  group_by(subj_id,test_part,lag2,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag2,RT,sep='_') %>%
-  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);
-
-E2.RT_by_lag1 <- E2.click_log %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,lag1,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lag1,RT,sep='_') %>%
-  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
-
-E2.RT_by_hit <- E2.click_log %>%
-  filter(click_number>0) %>%
-  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(hit_bin,RT,sep='_') %>%
-  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
-
-E2.RT_by_lead1 <- E2.click_log %>%
-  filter(click_number<num_clicks) %>%
-  group_by(subj_id,test_part,lead1,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead1,RT,sep='_') %>%
-  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
-
-E2.RT_by_lead2 <- E2.click_log %>%
-  filter(click_number<num_clicks-1) %>%
-  group_by(subj_id,test_part,lead2,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead2,RT,sep='_') %>%
-  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)
-
-E2.RT_by_lead3 <- E2.click_log %>%
-  filter(click_number<num_clicks-2) %>%
-  group_by(subj_id,test_part,lead3,genuine_first)%>%
-  summarise(RT=median(RT))  %>%
-  spread(lead3,RT,sep='_') %>%
-  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)
-
-E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
-  merge(E2.RT_by_lag2) %>%
-  merge(E2.RT_by_lag1) %>%
-  merge(E2.RT_by_hit) %>%
-  merge(E2.RT_by_lead1) %>%
-  merge(E2.RT_by_lead2) %>%
-  merge(E2.RT_by_lead3);
-
-E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
-  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
-  mutate(position = ifelse(position=="lag3_diff", 3,
-                           ifelse(position=="lag2_diff",2,
-                                  ifelse(position=="lag1_diff",1,
-                                         ifelse(position=="hit_diff",0,
-                                                ifelse(position=="lead1_diff",-1,
-                                                       ifelse(position=="lead2_diff",-2,-3
-                                                              ))))))) %>%
-  dplyr::select(subj_id,test_part,genuine_first,position,difference);
-
-p<- E2.RT_by_hit_lags_long %>% 
-  filter(position<2 & position > -1)%>%
-  group_by(test_part,position) %>%
-  summarise(median_difference=median(difference),
-            sem_difference=bootstrap_error(difference,N_perm)) %>%
-  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  scale_x_continuous(breaks=0:1,labels=c('hit on current trial', 'hit on previous trial'))+
-  geom_abline(slope=0,intercept=0,linetype=3)+
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.pdf',p,width=7)
-ggsave('../figures/E2_RT_by_lag_three_only_collapse.png',p,width=4, height=3.5, dpi=300)
-
-```
-
-### Effects of hitting versus missing a ship
-
-In non-pretend games, players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='nonpretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) when hitting compared to missing a ship (see Fig.
-3A; `r E2.RT_by_hit_lags%>%filter(test_part=='nonpretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-The same effect was observed in pretend games: players were faster by `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms (`r round(E2.RT_by_hit%>%filter(test_part=='pretend')%>%mutate(relative=hit_diff/hit_bin_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) in hits compared to misses (`r E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend first: `r E2.RT_by_hit_lags%>%filter(test_part=='pretend')%>%pull(hit_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-However, this difference between hits and misses was significantly more pronounced in pretend games (`r apa_print(t.test(E2.RT_by_hit_lags%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff), E2.RT_by_hit_lags%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)))$statistic`).
-
-The opposite effect was observed for a hit on the previous click: players were slower by `r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) after hitting a ship (`r E2.RT_by_lag1%>%filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-This effect was also observed in pretend games, where players were slower by `r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms (`r round(E2.RT_by_lag1%>%filter(test_part=='pretend')%>%mutate(relative=lag1_diff/lag1_TRUE)%>%pull(relative)%>%mean()%>%abs()*100)`%) following a hit on the previous click (`r E2.RT_by_lag1%>%filter(test_part=='pretend')%>%pull(lag1_diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Again, the absolute effect in ms was exaggerated in the pretend condition (`r t.test(E2.RT_by_lag1%>%filter(!genuine_first & test_part=='pretend')%>%pull(lag1_diff),E2.RT_by_lag1%>%filter(genuine_first & test_part=='nonpretend')%>%pull(lag1_diff))%>%apa_print()%>%'$'(statistic)`).
-
-### Effects of decision uncertainty
-
-When playing Battleships, it is sometimes clear what the next cell selection should be, and sometimes more difficult to decide where to click next.
-To capture this notion of decision uncertainty, we calculated the entropy of the posterior distribution over cell selections $H(P)=-\sum_{i=1}^{k}P(x_i) \log P(x_{i})$, where $P(x_{i})=\frac{p(Ship(x_i))}{\sum_{j=1}^{k}p(Ship(x_j))}$, and asked how this measure relates to decision latency, or the time taken to click on the next cell.
-$H(P)$ is high when players need to decide between multiple cells with a similar probability of hiding a ship, and low when there are only a few candidates with a high probability of hiding a ship.
-For every player and condition separately, we fitted a multiple linear regression to predict decision latency based on $H(P)$ and $H(P)^2$.
-The resulting coefficients were then subjected to a group-level inference.
-The first cell selection of each game was excluded from this analysis, because entropy was constant for the first click.
-
-```{r entropy-RT, echo=FALSE, cache=TRUE}
-
-E2.RT_by_entropy <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  mutate(entropy_bin=round(entropy*2)/2) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(test_part,subj_id,entropy_bin) %>%
-  summarise(RT=median(RT),
-            corrected_RT = median(corrected_RT)) %>%
-  group_by(test_part,entropy_bin) %>%
-  summarise(median_RT=median(RT),
-            sem_RT=bootstrap_error(RT,N_perm),
-            median_corrected_RT=median(corrected_RT),
-            sem_corrected_RT=bootstrap_error(corrected_RT,N_perm))
-
-p <- E2.RT_by_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of posterior', y='median RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
-# ggsave('../figures/E2_RT_by_entropy_no_split.svg',p,width=5,height=2,dpi=300)
-
-p <- E2.RT_by_entropy %>%
-  ggplot(aes(x=entropy_bin,color=test_part,fill=test_part,y=median_corrected_RT)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_path(size=1)+
-  geom_ribbon(aes(ymin=median_corrected_RT-sem_corrected_RT,ymax=median_corrected_RT+sem_corrected_RT),alpha=0.3,show.legend=F)+
-  labs(fill=' ',x='Entropy of posterior', y='median residual RT (ms)')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_entropy_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.RT_by_entropy_coefs <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  mutate(entropy_centered = entropy-mean(entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,grid_number,click_number,entropy_centered,entropy_squared,RT) %>%
-  group_by(genuine_first,subj_id,test_part,grid_number) %>%
-  do(model=lm(RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value)) %>%
-  group_by(genuine_first,subj_id,test_part,term) %>%
-  summarise(estimate = mean(estimate))
-
-E2.RT_by_entropy_coefs_summary <- E2.RT_by_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.RT_by_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-E2.corrected_RT_by_entropy_coefs <- E2.click_log_with_boards %>%
-  filter(click_number>1) %>%
-  group_by(subj_id,test_part,click_number) %>%
-  mutate(corrected_RT = RT-mean(RT)) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  mutate(entropy_centered = entropy-mean(entropy),
-         entropy_squared = entropy_centered^2) %>%
-  select(genuine_first,subj_id,test_part,entropy_centered,entropy_squared,corrected_RT) %>%
-  group_by(genuine_first,subj_id,test_part) %>%
-  do(model=lm(corrected_RT~entropy_centered+entropy_squared,data=.)) %>%
-  mutate(tidys=list(broom::tidy(model))) %>%
-  unnest(tidys) %>%
-  dplyr::select(!c(std.error,statistic,p.value))
-
-E2.corrected_RT_by_entropy_coefs_summary <- E2.corrected_RT_by_entropy_coefs %>%
-  group_by(test_part,term) %>%
-  summarise(mean_estimate = mean(estimate),
-            se_estimate = se(estimate))
-
-p <- E2.corrected_RT_by_entropy_coefs_summary %>%
-  mutate(term = factor(term,levels=c('entropy_centered','entropy_squared'), labels = c('entropy','entropy squared')))%>%
-  filter(term!='(Intercept)') %>%
-  ggplot(aes(x=term,y=mean_estimate,fill=test_part)) +
-  scale_fill_manual(values=c("#69b3a2", "#404080")) +
-  scale_color_manual(values=c("#69b3a2", "#404080")) +
-  geom_bar(stat='identity',color='black',position=position_dodge()) +
-  geom_errorbar(aes(ymin=mean_estimate-se_estimate,ymax=mean_estimate+se_estimate),width=.2,position=position_dodge(.9))+
-  labs(fill=' ',x='term', y='coefficient')+
-  theme_classic()
-
-ggsave('../figures/E2_corrected_RT_by_entropy_coefficients_no_split.png',p,width=5,height=2,dpi=300)
-
-```
-
-In non-pretend games, we found no evidence for a linear relation between decision entropy and decision latency (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).
-We did however observe a block order effect, with a significantly negative linear modulation only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_centered' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
-In pretend games this negative relation between $H(P)$ and click latency was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_centered')%>%pull(estimate)%>%t.test())$statistic`).
-
-A negative linear modulation of $H(P)$ on decision latency is surprising, as it suggests that participants were quicker to decide when uncertainty was high.
-However, visual inspection of the RT/entropy curves (Fig.
-3B) reveals a pronounced quadratic modulation, such that in both pretend and non-pretend games slower decision times are observed for intermediate $H(P)$ values of around 2.
-Similar to the linear effect of $H(P)$ on decision latency, here too a quadratic modulation in non-pretend games was observed only in the group that pretended in the first (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)%>%t.test())$statistic`), but not in the second block (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate)%>%t.test())$statistic`).
-Again, in pretend games this negative quadratic relation was significant in both groups, as well as overall (`r apa_print(E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared')%>%pull(estimate)%>%t.test())$statistic`).
-A between-subjects comparison focusing on games from the first block revealed that this negative quadratic effect was stronger in pretend, compared to non-pretend games (`r apa_print(t.test(E2.RT_by_entropy_coefs%>%filter(test_part=='nonpretend' &term=='entropy_squared' & genuine_first)%>%pull(estimate),E2.RT_by_entropy_coefs%>%filter(test_part=='pretend' &term=='entropy_squared' & !genuine_first)%>%pull(estimate)))$statistic`).
-
-```{r latency, echo=FALSE, out.width = '100%', fig.cap = "A: A subtraction of decision latencies in misses versus hits in the current and previous cell selections, in pretend and non-pretend games.
-B.
-Upper panel: Effect of the entropy of the posterior distribution over cell selections (a measure of decision uncertainty) on decision latency in pretend and non-pretend games.
-Lower panel: mean beta coefficients for a multiple regression model, predicting decision latency from entropy and entropy squared.
-C: Decision latency after a hit, as a function of hit number."}
-knitr::include_graphics("../figures/decision_latencies.png")
-```
-
-<!-- $H(P)$ is highest before the first click in a game and lowest just before the last one.
-To make sure any results are not driven by a general effect of click number on decision latency, and in deviation from our pre-registered analysis plan, we first mean-centered all decision latencies within individual subjects, conditions, and click numbers.
-This way, any effect of $H(P)$ on latency cannot reflect effects of click order.
--->
-
-## Effects of ship completion
-
-```{r}
+```{r ship-completion}
 
@@
 
-ggsave('../figures/E2_RT_by_previous_hit_number_collapse.png',p,width=5,height=2,dpi=300)
+ggsave('../docs/figures/E2_RT_by_previous_hit_number_collapse.png',p,width=5,height=2,dpi=300)
 
@@
 
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)
+ggsave('../docs/figures/E2_corrected_RT_by_previous_hit_number_collapsed.png',p,width=5,height=2,dpi=300)
 
@@
 
-ggsave('../figures/E2_corrected_RT_by_previous_hit_number_collapsed_boats_only.png',p,width=5,height=2,dpi=300)
+ggsave('../docs/figures/E2_corrected_RT_by_previous_hit_number_collapsed_boats_only.png',p,width=5,height=2,dpi=300)
 
 ```
-
-The previous analyses reveal robust associations between game state and decision latency, that are highly similar, and even exaggerated, in pretend compared to non-pretend games.
-To get at subtler dynamics in a more direct way, in a follow-up exploratory analysis we focused on decision latencies following a hit.
-We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn’t been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine (see Fig.
-3C).
-In the first case, players know the ship must continue in one of the neighboring cells.
-In the second case, there is a good chance the ship continues (if this ship turns out to be a submarine).
-In the third and fourth cases, it is clear that the ship is fully sunk.
-
-In non-pretend games, participants were significantly slower to select the next cell when they knew they had just completed a ship (categories 3 and 4) compared to when they just hit a ship, but were not sure (category 2) or knew they had not completely sunk it (category 1; `r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-Specifically, we find that participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-
-Strikingly, we found the exact same pattern in pretend games.
-Players were faster to make the next cell selection when they pretended to think that the current ship might not be fully sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).
-This was not merely a difference between the first, second and third hits: participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).
-Further analysis confirmed that this effect remained significant when controlling for click number (`r apa_print(E2.corrected_RT_by_hit_order%>%spread(prev_hit_type,corrected_RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`), and when restricting the analysis to the second hit of a ship that is in fact of size-two (`r apa_print(E2.RT_by_hit_order_boats_only%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).
-
-This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their their response latency in the same way it would have been affected had they been in a non-pretend game.
-
-## Half-games
-
-Our optimality analysis showed that pretenders’ click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be.
-However,  due to the high number of possible board configurations, data from full games provide limited opportunity to compare cell selections for specific game states.
-In addition to asking, “What guides cell selections in pretend and non-pretend games?”, we also wanted to ask, “Where exactly would pretenders and non-pretenders click, given a specific board configuration?”.
-
-To achieve this, the sixth game in each block started not with an empty grid, but with the contents of some cells already revealed by a previous player.
-As before, pretenders also knew where the remaining ships were hidden, but tried to play as if they only knew what was known to this previous player.
-Having cell selections from 250 players for each board configuration and condition allowed us to plot and compare the distribution of clicks under a genuine, or pretend, knowledge state.
-
-```{r, child=c('../half_games.Rmd')}
-```
-
-For example, look at Board A in Fig.
-4 (upper left corner).
-Imagine that you need to select your next move, knowing that exactly two size-two patrol boats and one size-three submarine are hiding in the water.
-Where would you click next?
-Perhaps one of the two revealed ships is the size-three submarine.
-In that case you want to click on the first cell in the third row, or on the third cell in the fifth row.
-Or maybe these are the two size-two patrol boats.
-This would mean the submarine is hiding in some other place, and for it not to touch sides with the boats this must be in the first row.
-The hit probability map in the second column visualizes this reasoning, by plotting the normalized probability that a ship is hidden in different cells.
-As can be seen, the highest probabilities are to the sides of the revealed ships, and in the first row.
-
-In the third column we plot the distribution of clicks for non-pretend players.
-This distribution is in agreement with the hit probability map (board A: `r cor_A_nonpretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_nonpretend%>%apa_print()%>%'$'(full_result)`).
-Finally, in the fourth column we plot the distribution of cell selections for pretend players.
-Although noisier, this distribution is also in agreement with the hit probability map (board A: `r cor_A_pretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_pretend%>%apa_print()%>%'$'(full_result)`), and more importantly, with the hit distribution of non-pretenders for the same board configuration (board A: `r cor_A_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`,  board B: `r cor_B_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`).
-
-```{r half-games, echo=FALSE, out.width = '100%', fig.cap = "Half games.
-First column: the two boards, as they appeared to pretenders.
-Second column: objective hit probability, given game state.
-Third and fourth columns: empirical click probabilities in non-pretend and pretend games, respectively."}
-knitr::include_graphics("../figures/Half_games.png")
-```
-
-<!-- Furthermore, compared to non-pretend games, in pretend games participants were more likely to make irrational clicks on cells where a ship could not in fact be hiding (a median of `r E2.num_irrational_clicks%>%filter(test_part=='pretend')%>%pull(nic)%>%median()` per game in pretend versus `r E2.num_irrational_clicks%>%filter(test_part=='nonpretend')%>%pull(nic)%>%median()` in non-pretend games.
-`r E2.num_irrational_clicks%>%spread(test_part,nic)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
--->
-## Judge trials
-
-```{r condition-classifier, echo=FALSE, eval=FALSE}
-
-set.seed(42);
-
-gamewise_data <- E2.click_log_with_boards %>% 
-  filter(!genuine_first | test_part=='pretend') %>% 
-  group_by(subj_id,test_part,grid_number) %>%
-  filter(sum(RT)<=60000) %>%
-  summarise(median_RT=median(RT),
-            mean_p_click_rank = mean(p_click_rank),
-            first_click_latency=RT[click_number==1],
-            num_irrational_clicks = sum(p_click==0)) %>%
-  ungroup()%>%
-  select(test_part,mean_p_click_rank,median_RT,num_irrational_clicks);
-
-gamewise_data_nonpretend = gamewise_data%>%filter(test_part=='nonpretend') %>%mutate(nrow=1:n())
-gamewise_data_pretend = gamewise_data%>%filter(test_part=='pretend')
-rows <- sample(nrow(gamewise_data_nonpretend))
-gamewise_data_pretend <- gamewise_data_pretend[rows,]%>%mutate(nrow=1:n())
-
-pair_gamewise_data <- full_join(gamewise_data_nonpretend,
-                                gamewise_data_pretend,
-                                by='nrow',
-                                suffix=c('.nonpretend','.pretend')) %>%
-  mutate(which_side = sample(c(-1,1),nrow(gamewise_data_pretend),replace=TRUE),
-         mean_p_click = which_side*(mean_p_click_rank.nonpretend-mean_p_click_rank.pretend),
-         median_RT = which_side*(median_RT.nonpretend-median_RT.pretend),
-         num_irrational_clicks = which_side*(num_irrational_clicks.nonpretend-num_irrational_clicks.pretend),
-         which_side=factor(which_side)) %>%
-  select(which_side,mean_p_click,median_RT,num_irrational_clicks)
-
-train_control <- trainControl(method = "CV", number=10, sampling='down')
-
-# train the model on training set
-model <- train(test_part ~ .,
-               data = gamewise_data,
-               trControl = train_control,
-               method = "svmLinear",
-               preProcess = c("center","scale"))
-
-pair_model <- train(which_side ~ .,
-               data = pair_gamewise_data,
-               trControl = train_control,
-               method = "svmLinear",
-               preProcess = c("center","scale"))
-
-# print cv scores
-
-```
-Despite the fact that pretenders were overall successful in mimicking patterns of cell selection and decision latency, they also left many clues for an independent observer to pick up on: they were slower, displayed exaggerated effects of decision uncertainty and hit success on response time, and were less optimal in choosing where to click next.
-To illustrate, a support vector machine (SVM) algorithm reached an accuracy level of 72% in linearly classifying condition (pretend / non-pretend) based on just three basic summary features: median decision latency, game optimality score, and number of irrational clicks per game.
-
-```{r judge}
-
-get_common_board = function(grid_number,test_part) {
-  pretend_grids = grid_number[which(test_part=='pretend')];
-  nonpretend_grids = grid_number[which(test_part=='nonpretend')];
-  common_board = intersect(pretend_grids,nonpretend_grids);
-  return(common_board)
-};
-
-common_boards <- E2.df %>%
-  filter(test_part=='pretend' | test_part=='nonpretend') %>%
-  dplyr::select(subj_id,grid_number,test_part) %>%
-  group_by(subj_id) %>%
-  summarise(common_board = get_common_board(grid_number,test_part))
-
-used_for_pretend <- function(s, g) {
-
-  return (E2.df%>%filter(test_part=='pretend' &
-                           subj_id==s & 
-                           grid_number == g) %>%
-            nrow()==1)
-}
-
-used_for_nonpretend <- function(s, g) {
-
-  return (E2.df%>%filter(test_part=='nonpretend' &
-                           subj_id==s & 
-                           grid_number == g) %>%
-            nrow()==1)
-}
-
-E2.judge_df <- E2.df %>%
-  filter(test_part=='judge' & !(cheat_player %in% E2.lucky_pretenders))%>%
-  dplyr::select(subj_id, 
-         grid_number, 
-         cheat_player,
-         genuine_player,
-         cheater,
-         noncheater,
-         decision,
-         replay_log,
-         choose_pretender,
-         correct)%>%
-  rowwise()%>%
-  mutate(choose_pretender=choose_pretender=='True',
-         used_for_pretend = used_for_pretend(subj_id,grid_number),
-         used_for_nonpretend = used_for_nonpretend(subj_id,grid_number),
-         common_board = used_for_pretend & used_for_nonpretend,
-         board_type = ifelse(common_board, 'common',
-                             ifelse(used_for_pretend, 'pretend',
-                                    ifelse(used_for_nonpretend,'nonpretend','new')))
-)
-
-p <- E2.judge_df %>%
-  group_by(subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) +
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') +
-  scale_x_continuous(breaks=c(0,1,2,3,4,5));
-
-ggsave('../figures/E2_correct_guesses.png',p,width=5,height=3,dpi=300)
-
-p <- E2.judge_df %>%
-  mutate(choose_pretender = ifelse(choose_pretender,
-                                   'who had hints?', 
-                                   'who had no hints?')) %>% 
-  group_by(choose_pretender,subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) + 
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') + 
-  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
-  facet_wrap(~choose_pretender,nrow=2);
-
-ggsave('../figures/E2_correct_guesses_by_choose_pretender.png',p,width=5,height=3,dpi=300);
-
-p <- E2.judge_df %>% 
-  group_by(board_type,subj_id) %>%
-  summarise(ncorrect=sum(correct)) %>%
-  ggplot(aes(x=ncorrect)) + 
-  geom_bar(fill='black') +
-  theme_classic() +
-  labs(x='number of correct guesses (out of 5)',y='number of players') + 
-  scale_x_continuous(breaks=c(0,1,2,3,4,5)) + 
-  facet_wrap(~board_type,nrow=1);
-
-ggsave('../figures/E2_correct_guesses_by_board_type.png',p,width=5,height=3,dpi=300)
-```
-
-Despite this, participants were never significantly above chance in judging which of the two presented games came from a pretender (mean accuracy: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%; t test against 50: `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%t.test(mu=50))$statistic`).
-
-For a subset of `r E2.judge_df%>%filter(common_board)%>%pull(correct)%>%length()` players, one of the five boards presented in the judging block was already presented both in the pretend and in the non-pretend blocks.
-Still, even this subset of participants who had just experienced a board both as pretenders and as non-pretenders were not significantly above chance in telling which of two other players had hints for this same board (mean accuracy: 55%, $p=.141$).
-
-## Are good pretenders also better judges?
-
-```{r judge-pretend-cor, echo=FALSE}
-
-unidentifiable_ss <- E2.judge_df %>%
-  mutate(subj_num=substr(as.character(subj_id),1,3))%>%
-  group_by(subj_num)%>%
-  summarise(subj_id=subj_id, n=n())%>%
-  filter(n>5)%>%
-  pull(subj_id)%>%
-  unique();
-
-E2.judge_performance <- E2.judge_df %>%
-  filter(!(subj_id %in% unidentifiable_ss)) %>%
-  mutate(subj_num=as.numeric(substr(as.character(subj_id),1,3)))%>%
-  group_by(subj_num)%>%
-  summarise(judge_rate=sum(correct)/5);
-
-E2.pretend_performance <- E2.judge_df %>%
-  filter(!(subj_id %in% unidentifiable_ss)) %>%
-  mutate(subj_num=cheat_player)%>%
-  group_by(subj_num)%>%
-  summarise(ntrials=n(),
-            successful_pretend=ntrials-sum(correct),
-            pretend_rate = successful_pretend/ntrials);
-
-E2.performance_df <- merge(E2.judge_performance,E2.pretend_performance);
-
-p <- E2.performance_df %>%
-  ggplot(aes(x=judge_rate,y=pretend_rate)) +
-  geom_jitter() + 
-  labs(x='judge accuracy',y='pretense success rate') +
-  theme_classic()
-
-ggsave('../figures/E2_judge_pretend_accuracy.png',p,width=5,height=3,dpi=300)
-
-min_njudge = seq(0,10);
-cors =c()
-ns = c()
-
-for (minn in min_njudge) {
-
-  temp_df = E2.performance_df %>%
-    filter(ntrials>minn)
-
-  cur_cor = cor(temp_df$pretend_rate,temp_df$judge_rate);
-  cur_n = temp_df%>%nrow()
-  cors = c(cors,cur_cor)
-  ns = c(ns, cur_n)
-}
-
-judge_pretend_correlation_df <- data.frame(min_njudge,cors,ns);
-
-```
-
-```{r judge-optimality-cor, echo=FALSE}
-E2.judge_summary <- E2.judge_df %>%
-  group_by(subj_id) %>%
-  summarise(judge = sum(correct));
-
-E2.optimality_summary <- E2.click_log_with_boards %>%
-  group_by(subj_id,test_part) %>%
-  summarise(p_click_rank = mean(p_click_rank)) %>%
-  spread(test_part,p_click_rank)
-
-E2.judge_summary <- E2.judge_summary %>%
-  merge(E2.optimality_summary)
-```
-
-The abilities to pretend and to detect pretense in others both rely on some form of theory of mind: reasoning about counterfactual knowledge states of oneself in the first case, and of others in the second.
-For example, in scrub jays, individual birds that pilfered another bird's caches were also more careful in hiding their own food, suggesting an ability to project from their own experience of stealing to the behavior of others [@emery2001effects].
-Along similar lines, we reasoned that those participants whose pretend games better resembled the games of genuine players would also be more accurate in detecting pretense in others.
-To our surprise, this was not at all the case (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate))$full_result`).
-Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend))$full_result`).
-
-Despite our considerable sample size of 500 participants, we are cautious in making strong claims based on the absence of a correlation between pretense and pretense detection abilities in our sample.
-We note, however, that previous developmental research has identified weaker correlations with theory of mind measures for a tendency to intentionally deceive another player in a game setting, compared to an ability to detect and understand deception, potentially indicating partly different mechanisms [@shultz1981development].
-
-# Discussion
-
-Many tasks (including writing papers for conferences) require us to overcome the ‘curse of knowledge’ — the challenge of putting ourselves in the shoes of someone who doesn’t know what we know.
-This is just an example of a more general fact: the ability to simulate counterfactual knowledge states is necessary for efficient and flexible communication.
-To do this, we must have an internal self-model that can generate thoughts and behavior for hypothetical beliefs, desires, and perceptions.
-
-Research on Bayesian Theory of Mind provides some support for the existence of such a model by measuring subjects' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational], or implicitly [@onishi2005fifteen; @liu2017ten].
-Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state - in this case, a counterfactual knowledge state in which a known piece of information is unknown.
-Instead of relying on model inversion (e.g., “Which belief states would give rise to this behavior?”), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
-Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
-
-Our subjects rose to the challenge.
-Their pretense behavior, although not perfect, closely resembled that of non-pretenders — not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
-Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: “What would I have done if I didn’t have this piece of knowledge?”
-
-An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game.
-This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain participants’ pretense ability in our data, for two indirect reasons.
-First, suppressing thoughts on demand is notoriously difficult  [@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible.
-Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with a self-simulation account.
-In future work, we intend to more directly rule out this knowledge-suppression hypothesis, and well as to generalize our results to other games.
-
-In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects’ models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
-While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.
-
-# References
-
-```{r}
-# References will be generated automatically by Pandoc and included here.
-# The following code is some latex to format the bibliography. Do not remove it.
-```
-
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
-\noindent

### 2023-01-27 docs/half_games.rmd
pre-registered analysis and some exploratory analysis

```{r}

plot_posterior <- function(posterior,file_name) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  board_df <- data.frame(i=rep(c(1,2,3,4,5),each=5),j=rep(c(1,2,3,4,5),5),posterior=posterior);
  p <- ggplot(board_df,aes(x=j,y=i,fill=posterior)) +
    geom_tile() +
    coord_equal() +
    scale_x_continuous(breaks=c()) +
    scale_y_continuous(breaks=c(),trans='reverse') +
    scale_fill_gradient(low='#649bc8',high='#c83237', na.value ='white')+
    labs(x='',y='') +
    theme_classic();

  ggsave(paste('figures/',file_name,'.png',sep=''),width=5,dpi=300);
    ggsave(paste('figures/',file_name,'.svg',sep=''),width=5,dpi=300);

}

boardA <- "NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,1,1,0,0,NA,NA,0,NA,NA,1,1,NA,NA,NA"
posteriorA <- boardA%>%get_likelihood()%>%get_posterior();
plot_posterior(posteriorA,'posteriorA');

boardB <- "NA,NA,NA,NA,0,NA,NA,0,NA,NA,NA,NA,NA,NA,0,NA,NA,1,0,NA,NA,NA,1,NA,1"
posteriorB <- boardB%>%get_likelihood()%>%get_posterior();
plot_posterior(posteriorB,'posteriorB');

```

```{r}

E2.half_game_click_df <- E2.df %>%
  dplyr::select(subj_id, 
         test_part, 
         grid_number, 
         num_clicks,
         click_log,
         genuine_first) %>%
  mutate(click_log = gsub("\'","\"", click_log)) %>%
  filter(test_part=='pretend_half_game' | test_part=='nonpretend_half_game');

E2.half_game_click_log <- data.frame(matrix(ncol=10,nrow=0, 
                               dimnames=list(NULL, 
                                             c("subj_id",
                                              "test_part", 
                                              "grid_number", 
                                              "num_clicks",
                                              "i",
                                              "j",
                                              "hit",
                                              "t",
                                              "click_number",
                                              "genuine_first"))))

for (row in 1:nrow(E2.half_game_click_df)) {

    subject_click_log <- data.frame(fromJSON(E2.half_game_click_df[row, ]$click_log)) %>%
    mutate(
      click_number = 1:n(),
      subj_id = E2.half_game_click_df[row, ]$subj_id,
      test_part = E2.half_game_click_df[row, ]$test_part,
      grid_number = E2.half_game_click_df[row, ]$grid_number,
      num_clicks = E2.half_game_click_df[row, ]$num_clicks,
      genuine_first = E2.half_game_click_df[row, ]$genuine_first
    )

    E2.half_game_click_log <- rbind(E2.half_game_click_log, subject_click_log);
}

E2.half_game_click_log <- E2.half_game_click_log %>%
  relocate(subj_id, .before = i) %>%
  relocate(genuine_first, .before = i) %>% 
  relocate(test_part, .before=i) %>% 
  relocate(grid_number, .before=i) %>% 
  relocate(click_number, .before=i) %>%
  group_by(subj_id,test_part,grid_number) %>%
  mutate(RT=t-lag(t,default=0)) %>%
  filter(!(subj_id %in% E2.lucky_pretenders))

```

```{r}

E2.half_game_first_clicks <- E2.half_game_click_log %>%
  filter(click_number==1) %>%
  group_by(i,j, grid_number,test_part) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  complete(i,j,grid_number,test_part,fill=list(n=0)) %>%
  group_by(grid_number,test_part) %>%
  mutate(n=n/sum(n), x=i*5+j);

p<- E2.half_game_first_clicks %>%
  ggplot(aes(x=j,y=i,fill=n)) +
  geom_tile(color='black') +
  # geom_text(aes(label=printnum(n)))+
  coord_equal() +
  scale_x_continuous(breaks=c()) +
  scale_y_continuous(breaks=c(),trans='reverse') +
  scale_fill_gradient(low='#649bc8',high='#c83237', na.value ='white')+
  labs(x='',y='') +
  theme_classic() + 
  facet_wrap(~grid_number+test_part, nrow=2);

ggsave('figures/E2_half_game_first_clicks.png',width=10,height=10,dpi=300)
ggsave('figures/E2_half_game_first_clicks.svg',width=10,height=10,dpi=300)

p<- E2.half_game_first_clicks %>%
  group_by(grid_number,test_part) %>%
  mutate(p=n/sum(n)) %>%
  ggplot(aes(x=j,y=i)) +
  geom_tile(color='black',fill = NA) +
  # geom_text(aes(label=printnum(n)))+
  coord_equal() +
  # Area fills, convert factors to numeric
  geom_rect(
    aes(xmin = as.numeric(j) - 0.5, 
        xmax = as.numeric(j) + 0.5,
        ymin = as.numeric(i) + 0.5-p,
        ymax = as.numeric(i) + 0.5),
    fill='black',alpha=1

  )+
  scale_x_continuous(breaks=c()) +
  scale_y_continuous(breaks=c(),trans='reverse') +
  labs(x='',y='') +
  theme_classic() + 
  facet_wrap(~grid_number+test_part, nrow=2);

ggsave('figures/E2_half_game_first_clicks_height.svg',width=10,height=10,dpi=300)

## check correlation with posterior

cor_A_nonpretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='nonpretend_half_game' & grid_number=='A') %>% 
                               pull(n),scan(text= posteriorA, what = numeric(), sep="," , quiet = TRUE));

cor_A_pretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='pretend_half_game' & grid_number=='A') %>% 
                               pull(n),scan(text= posteriorA, what = numeric(), sep="," , quiet = TRUE));

cor_A_pretend_nonpretend <- cor.test(E2.half_game_first_clicks %>%
             filter(test_part=='nonpretend_half_game' & grid_number=='A') %>% 
             pull(n), 
             E2.half_game_first_clicks %>%
             filter(test_part=='pretend_half_game' & grid_number=='A') %>% 
             pull(n))

cor_B_nonpretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='nonpretend_half_game' & grid_number=='B') %>% 
                               pull(n),scan(text= posteriorB, what = numeric(), sep="," , quiet = TRUE));

cor_B_pretend <- cor.test(E2.half_game_first_clicks %>%
                               filter(test_part=='pretend_half_game' & grid_number=='B') %>% 
                               pull(n),scan(text= posteriorB, what = numeric(), sep="," , quiet = TRUE));

cor_B_pretend_nonpretend <- cor.test(E2.half_game_first_clicks %>%
             filter(test_part=='nonpretend_half_game' & grid_number=='B') %>% 
             pull(n), 
             E2.half_game_first_clicks %>%
             filter(test_part=='pretend_half_game' & grid_number=='B') %>% 
             pull(n))

```

### 2023-01-27 docs/ExploratoryResults.Rmd
suppression or simulation for batteship (optimality only)

---
title             : "Pretending not to know: exploratory analyses"
shorttitle        : "Pretending not to know"

author: 
  - name          : "Matan Mazor"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Malet St., London WC1E 7HX"
    email         : "mtnmzor@gmail.com"
  - name          : "Chaz Firestone"
    affiliation   : "2"
  - name          : "Ian Phillips"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Birkbeck, University of London"
  - id            : "2"
    institution   : "Johns Hopkins University"

authornote: |

abstract: |

keywords          : "pretense, theory of mind, metacognition, self-model, self-simulation"
wordcount         : "X"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : bookdown::gitbook
bibliography: references.bib
---

```{r global_options, include=FALSE}
set.seed(42)
knitr::opts_chunk$set(fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T, 
                      fig.width = 10, fig.height = 5)
```

```{r, libraries, echo=FALSE, warning=FALSE, cache=F}

library(groundhog)

groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel' #for word scatterplots
), '2022-12-01')

```

# Exp. 1: Battleship; Exploratory results

```{r battleship-load, warning=FALSE}

E2.df <- read.csv('../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,4),sep='')) %>%
  mutate(subj_id = factor(subj_id)) %>%
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
  mutate(genuine_first=genuine_first=='True');

E2.click_df <- E2.df %>%
  dplyr::select(subj_id, 
         test_part, 
         grid_number, 
         num_clicks,
         click_log,
         genuine_first) %>%
  mutate(click_log = gsub("\'","\"", click_log)) %>%
  filter(test_part=='pretend' | test_part=='nonpretend');

E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();

E2.click_log <- readRDS('../experiments/Battleships2/data/batch1/click_log.Rda');

load('E2.click_log_with_boards.RData')

```

```{r functions, warning=FALSE}

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
  };
  return(sd(medians))
}

all_possible_boards <- read.csv('../analysis/flat_boards.csv',header=FALSE);

create_board_states <- function(flat_positions,hit) {

  board_state = rep(NA,25)
  board_states = c();
  for (p in seq_along(flat_positions)) {
    board_states = c(board_states, paste(board_state,collapse=','));
    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
  }

  return(board_states)
};

get_likelihood <- function(board_state) {

  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);

  revealed_positions <- which(!is.na(board_state));

  boards_to_remove = c();

  for (position in revealed_positions) {
    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
  }

  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
  likelihood[revealed_positions]=NA;

  return(paste(likelihood,collapse=','))

};

get_posterior <- function(likelihood) {
    likelihood = scan(text= likelihood, what = numeric(), sep="," , quiet = TRUE);
    posterior = likelihood/sum(likelihood,na.rm=TRUE); # assuming a uniform prior
    return(paste(posterior,collapse=','))
}

get_p_click <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(posterior[flat_position])
}

get_p_click_rank <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  rank_posterior = rank(-posterior)
  return(rank_posterior[flat_position])
}

get_posterior_entropy <- function(posterior) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  posterior[which(posterior==0)]=NA; #to avoid infinity*0
  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
  return(entropy);
}

```

## Ship completion

```{r, child=c('../analysis/shipCompletion.Rmd')}
```

To get at subtler dynamics in a more direct way, in a follow-up exploratory analysis we focused on decision latencies following a hit. We categorized hits into four types: first hit on a ship, second hit on a ship when the size-three submarine hasn't been sunk yet, second hit on a ship when the size-three submarine has already been sunk, and third hit on a submarine. In the first case, players know the ship must continue in one of the neighboring cells. In the second case, there is a good chance the ship continues (if this ship turns out to be a submarine). In the third and fourth cases, it is clear that the ship is fully sunk.

In non-pretend games, participants were significantly slower to select the next cell when they knew they had just completed a ship (categories 3 and 4) compared to when they just hit a ship, but were not sure (category 2) or knew they had not completely sunk it (category 1; `r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`). Specifically, we find that participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`).

Strikingly, we found the exact same pattern in pretend games. Players were faster to make the next cell selection when they pretended to think that the current ship might not be fully sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=first+second-secondlast-thirdlast)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`). This was not merely a difference between the first, second and third hits: participants were slower by `r E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%median(na.rm=TRUE)%>%round()` ms to make the next cell selection after hitting the second cell of a ship if the size-three submarine had already been sunk (`r apa_print(E2.RT_by_hit_order%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`). Further analysis confirmed that this effect remained significant when controlling for click number (`r apa_print(E2.corrected_RT_by_hit_order%>%spread(prev_hit_type,corrected_RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='nonpretend')%>%pull(diff)%>%t.test())$statistic`), and when restricting the analysis to the second hit of a ship that is in fact of size-two (`r apa_print(E2.RT_by_hit_order_boats_only%>%spread(prev_hit_type,RT)%>%mutate(diff=secondlast-second)%>%filter(test_part=='pretend')%>%pull(diff)%>%t.test())$statistic`).

This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their their response latency in the same way it would have been affected had they been in a non-pretend game.

```{r battleship-ship-completion-plot}

E2.RT_by_hit_order %>%
  filter(!is.na(prev_hit_type))%>%
  mutate(prev_hit_type = factor(prev_hit_type,levels=c('first','second','secondlast','thirdlast'), labels = c('first','second','second and last','third')))%>%
  group_by(test_part,prev_hit_type) %>%
  summarise(RT = median(RT),
            sem_RT = bootstrap_error(RT,N_perm)) %>%
  ggplot(aes(x=prev_hit_type,y=RT,fill=test_part)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  geom_bar(stat='identity',color='black',position=position_dodge()) +
  geom_errorbar(aes(ymin=RT-sem_RT,ymax=RT+sem_RT),width=.2,position=position_dodge(.9))+
  labs(fill=' ',x='previous hit', y='median RT')+
  theme_classic()

```

## Half games

Our optimality analysis showed that pretenders' click selections closely resemble those of non-pretenders, at least in that they are not random, but rather guided by where a ship might be. However, due to the high number of possible board configurations, data from full games provide limited opportunity to compare cell selections for specific game states. In addition to asking, "What guides cell selections in pretend and non-pretend games?", we also wanted to ask, "Where exactly would pretenders and non-pretenders click, given a specific board configuration?".

To achieve this, the sixth game in each block started not with an empty grid, but with the contents of some cells already revealed by a previous player. As before, pretenders also knew where the remaining ships were hidden, but tried to play as if they only knew what was known to this previous player. Having cell selections from 250 players for each board configuration and condition allowed us to plot and compare the distribution of clicks under a genuine, or pretend, knowledge state.

```{r, child=c('half_games.Rmd')}
```

In the third column we plot the distribution of clicks for non-pretend players. This distribution is in agreement with the hit probability map (board A: `r cor_A_nonpretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_nonpretend%>%apa_print()%>%'$'(full_result)`). Finally, in the fourth column we plot the distribution of cell selections for pretend players. Although noisier, this distribution is also in agreement with the hit probability map (board A: `r cor_A_pretend%>%apa_print()%>%'$'(full_result)`; board B: `r cor_B_pretend%>%apa_print()%>%'$'(full_result)`), and more importantly, with the hit distribution of non-pretenders for the same board configuration (board A: `r cor_A_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`, board B: `r cor_B_pretend_nonpretend%>%apa_print()%>%'$'(full_result)`).

```{r half-games, echo=FALSE, out.width = '100%', fig.cap = "Half games. First column: the two boards, as they appeared to pretenders. Second column: objective hit probability, given game state. Third and fourth columns: empirical click probabilities in non-pretend and pretend games, respectively."}
knitr::include_graphics("figures/Half_games.png")
```

## Within-participant click entropy

```{r, child=c('../analysis/clickEntropyBattleship.Rmd')}
```

In this analysis, we asked how flexible players were in their click sequences, across games. We quantified flexibility as the Shannon entropy of cell selections across games, for a given click number. For example, if a player always started their games by clicking in the top left corner, their flexibility score for the first click will be $H([1,1,1,1,1])=0$. Flexibility was quantified separately for pretend and non-pretend games.

```{r battleship-flexibility-plot}

E2.click_entropy_by_serial_position_summary %>%
  filter(test_part %in% c('pretend','nonpretend') & click_number<=20) %>%
  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
  geom_line() +
  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
  labs(x='click number', y='flexibility score (Shannon entropy)') +
  theme_classic() +
  scale_x_continuous(breaks=seq(20))

```

Flexibility scores increased as a function of click number for both pretend and non-pretend games, as expected if players adjust their behaviour based on the outcomes of previous clicks. This process reached a plateau of around $H=2.1$ at click number 6.

Importantly, before that point, flexibility was persistently lower in pretend games (click #1: `r apa_print(E2.click_entropy_by_serial_position_wide %>% filter(click_number==1)%>%pull(diff)%>%t.test())$statistic`; click #2: `r apa_print(E2.click_entropy_by_serial_position_wide %>% filter(click_number==2)%>%pull(diff)%>%t.test())$statistic`; click #3: `r apa_print(E2.click_entropy_by_serial_position_wide %>% filter(click_number==3)%>%pull(diff)%>%t.test())$statistic`; click #4: `r apa_print(E2.click_entropy_by_serial_position_wide %>% filter(click_number==4)%>%pull(diff)%>%t.test())$statistic`; click #5: `r apa_print(E2.click_entropy_by_serial_position_wide %>% filter(click_number==5)%>%pull(diff)%>%t.test())$statistic`).

We reasoned that this tendency to rigidly follow a pre-defined plan may underlie at least some of the difference in game optimality between pretend and non-pretend games. Interestingly, however, we find no significant correlation between the pretend/non-pretend differences in flexibility and optimality in the first 5 clicks (`r apa_print(cor.test(E2.entropy_cost$entropy_cost, E2.entropy_cost$p_click_cost))$full_result`).

## Suppression or simulation?

```{r suppression-or-simulation}

E2.debrief_labeled <- read.csv('..\\experiments\\Battleships2\\data\\debrief.csv',na.strings=c(""," ","NA")) %>%
  dplyr::select(-subj_id) %>%
  merge(E2.debrief <- E2.df %>%
          filter(!is.na(responses) & is.na(test_part)) %>%
          dplyr::select(subj_id,responses) %>%
          mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
          unnest(comments)%>%
          dplyr::select(subj_id,debrief)%>%
          filter(!is.na(debrief) & debrief != "")%>%
          mutate(debrief = gsub("[\r\n]", "",debrief)))

E4.debrief_labeled <- read.csv('..\\experiments\\Hangman2\\data\\debrief.csv',na.strings=c(""," ","NA"))

E2.simulators <- E2.debrief_labeled %>%
  filter(simulation==1 & suppression==0) %>%
  pull(subj_id)

E2.suppressors <- E2.debrief_labeled %>%
  filter(simulation==0 & suppression==1) %>%
  pull(subj_id)

E4.simulators <- E4.debrief_labeled %>%
  filter(simulation==1 & suppression==0) %>%
  pull(subj_id)

E4.suppressors <- E4.debrief_labeled %>%
  filter(simulation==0 & suppression==1) %>%
  pull(subj_id)

```

At the end of the experiment, players were asked whether they had a strategy for pretending or for detecting pretense. I (MM) manually labeled their responses according to three criteria: mentions of suppression (e.g., "Trying to ignore the hints"), simulation (e.g., "Trying to imagine how I would react"), and rules (e.g., "Checking the corners and center before random selection of tiles"). Some answers got more than one label and some got none. Overall, `r E2.debrief_labeled$suppression%>%sum()` mentioned using suppression, `r E2.debrief_labeled$simulation%>%sum()` mentioned using simulation, and `r E2.debrief_labeled$rules%>%sum()` mentioned following rules.

In the following analysis, we focus on the first two labels and carve out two sets of players: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but no suppression.

### Click optimality

```{r battleship-p_click-sup-sim, echo = FALSE, cache=TRUE}

E2.sup_sim.mean_P_click_rank <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal) %>%
  rbind(E2.click_log_with_boards_random) %>%
  ungroup() %>%
  filter(subj_id %in% c(E2.simulators, E2.suppressors)) %>%
  mutate(group = ifelse(subj_id %in% E2.simulators, 'simulators','suppressors')) %>%
  group_by(subj_id,test_part, group) %>%
  summarise(p_click_rank=mean(p_click_rank));

E2.sup_sim.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal) %>%
  rbind(E2.click_log_with_boards_random) %>%
  ungroup() %>%
  filter(subj_id %in% c(E2.simulators, E2.suppressors) & !hit_bin) %>%
  mutate(group = ifelse(subj_id %in% E2.simulators, 'simulators','suppressors')) %>%
  group_by(subj_id,test_part, group) %>%
  summarise(p_click_rank=mean(p_click_rank));

E2.sup_sim.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~group, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('figures/E2_pclick_rank_by_heuristic.pdf')
ggsave('figures/E2_pclick_rank_by_heuristic.png',width=5,height=2,dpi=300)

p<- E2.sup_sim.mean_P_click_rank %>%
    filter(test_part != 'optimal') %>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    scale_x_continuous(limits=c(0,15))+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~group, nrow=2)+
    # scale_x_reverse() +
    theme_classic();

ggsave('figures/E2_pclick_rank_by_heuristic_misses_only.pdf',p)
ggsave('figures/E2_pclick_rank_by_heuristic_misses_only.png',p,width=5,height=2,dpi=300)

```
The mean negative optimality score of non-pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)))$statistic`). The mean negative optimality score of pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)))$statistic`). To the very least, simulators did as well as suppressors in pretending.

# Exp. 2: Hangman; Exploratory results

```{r hangman-load, warning=FALSE}

E4.df <- read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch2.csv',na.strings=c(""," ","NA")) %>%
  rbind(read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,5),sep='')) %>%
  mutate(subj_id = factor(subj_id)) %>%
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
  filter(PROLIFIC_PID != '5ec4a156bc5aac3819ac52f2' & #reported some of the letters not showing up on their screen
        PROLIFIC_PID != '58211fc787f6b90001f13f9' & #reported some of the letters not showing up on their screen
        PROLIFIC_PID != '6149256f6335b06ade3723e0' &  # due a PROLFIIC bug, participated twice
        PROLIFIC_PID != '615ddab1e4f013092538b6c5')# due a PROLFIIC bug, participated twice

E4.click_df <- E4.df %>%
  dplyr::select(subj_id, 
         PROLIFIC_PID,
         test_part, 
         category,
         word,
         num_clicks,
         click_log,
         genuine_first,
         trial_type) %>%
  rowwise()%>%
  mutate(num_hits = strsplit(gsub(' ','',word),split='')[[1]]%>%unique()%>%length(),
         genuine_first= genuine_first=='True',
         num_misses = num_clicks-num_hits,
         click_log = gsub("\'","\"", click_log),
         click_log = gsub("None","null", click_log),
         word = factor(word,levels=c(
           'eleven',
           'ninety six',
           'dalai lama',
           'taylor swift',
           'strawberry',
           'lemon',
           'tooth',
           'head',
           'iowa',
           'montana'
         ))) %>%
  filter(test_part=='pretend' | test_part=='nonpretend') %>%
  filter(trial_type=='Hangman')

load('E4.click_log.RData')

load('E4.click_log_with_boards.RData')

load('E4.judge_df.RData')

```

## Within-participant click entropy

```{r, child=c('../analysis/clickEntropyHangman.Rmd')}
```

```{r hangman-flexibility-plot}

E4.click_entropy_by_serial_position_summary %>%
  filter(test_part %in% c('pretend','nonpretend') & click_number<=10) %>%
  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
  geom_line() +
  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
  labs(x='click number', y='flexibility score (Shannon entropy)') +
  theme_classic() +
  scale_x_continuous(breaks=seq(20))

```

Like for Battleship, flexibility scores increased as a function of click number for both pretend and non-pretend games, as expected if players adjust their behaviour based on the outcomes of previous clicks. This process reached a plateau of around $H=2$ and click number 6.

Importantly, before that point, flexibility was persistently lower in pretend games (click #1: `r apa_print(E4.click_entropy_by_serial_position_wide %>% filter(click_number==1)%>%pull(diff)%>%t.test())$statistic`; click #2: `r apa_print(E4.click_entropy_by_serial_position_wide %>% filter(click_number==2)%>%pull(diff)%>%t.test())$statistic`; click #3: `r apa_print(E4.click_entropy_by_serial_position_wide %>% filter(click_number==3)%>%pull(diff)%>%t.test())$statistic`; click #4: `r apa_print(E4.click_entropy_by_serial_position_wide %>% filter(click_number==4)%>%pull(diff)%>%t.test())$statistic`; click #5: `r apa_print(E4.click_entropy_by_serial_position_wide %>% filter(click_number==5)%>%pull(diff)%>%t.test())$statistic`).

We reasoned that this tendency to rigidly follow a pre-defined plan may underlie at least some of the difference in game optimality between pretend and non-pretend games. We find evidence for such a correlation (`r apa_print(cor.test(E4.entropy_cost$entropy_cost, E4.entropy_cost$p_click_cost))$full_result`). Note that the direction is negative because high optimality is mapped to low values, but high flexibility is mapped to high values.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2023-01-28 docs/ExploratoryResults.Rmd
simulation or pretense in battleship and hangman (RT and optimality analysis)

@@
   for (i in 1:N_perm) {
-    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
+    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median(na.rm=T))
   };
-  return(sd(medians))
+  return(sd(medians,na.rm=T))
 }
@@
 
-```{r suppression-or-simulation}
+```{r battleship-suppression-or-simulation}
 
@@
 
-E4.debrief_labeled <- read.csv('..\\experiments\\Hangman2\\data\\debrief.csv',na.strings=c(""," ","NA"))
-
 E2.simulators <- E2.debrief_labeled %>%
@@
 E2.suppressors <- E2.debrief_labeled %>%
-  filter(simulation==0 & suppression==1) %>%
-  pull(subj_id)
-
-E4.simulators <- E4.debrief_labeled %>%
-  filter(simulation==1 & suppression==0) %>%
-  pull(subj_id)
-
-E4.suppressors <- E4.debrief_labeled %>%
   filter(simulation==0 & suppression==1) %>%
@@
 
-In the following analysis, we focus on the first two labels and carve out two sets of players: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but no suppression.
+In the following analysis, we focus on the first two labels and carve out two sets of players: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but not suppression.
 
@@
 ```
+
 The mean negative optimality score of non-pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)))$statistic`).
@@
 To the very least, simulators did as well as suppressors in pretending.
+
+### RT by outcome
+
+```{r RT_by_accuracy-sup-sim}
+
+E2.RT_by_lag3 <- E2.click_log %>%
+  filter(click_number>3) %>%
+  group_by(subj_id,test_part,lag3,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag3,RT,sep='_') %>%
+  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);
+
+E2.RT_by_lag2 <- E2.click_log %>%
+  filter(click_number>2) %>%
+  group_by(subj_id,test_part,lag2,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag2,RT,sep='_') %>%
+  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);
+
+E2.RT_by_lag1 <- E2.click_log %>%
+  filter(click_number>1) %>%
+  group_by(subj_id,test_part,lag1,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag1,RT,sep='_') %>%
+  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
+
+E2.RT_by_hit <- E2.click_log %>%
+  filter(click_number>0) %>%
+  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(hit_bin,RT,sep='_') %>%
+  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
+
+E2.RT_by_lead1 <- E2.click_log %>%
+  filter(click_number<num_clicks) %>%
+  group_by(subj_id,test_part,lead1,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead1,RT,sep='_') %>%
+  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
+
+E2.RT_by_lead2 <- E2.click_log %>%
+  filter(click_number<num_clicks-1) %>%
+  group_by(subj_id,test_part,lead2,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead2,RT,sep='_') %>%
+  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)
+
+E2.RT_by_lead3 <- E2.click_log %>%
+  filter(click_number<num_clicks-2) %>%
+  group_by(subj_id,test_part,lead3,genuine_first)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead3,RT,sep='_') %>%
+  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)
+
+E2.RT_by_hit_lags <- E2.RT_by_lag3 %>%
+  merge(E2.RT_by_lag2) %>%
+  merge(E2.RT_by_lag1) %>%
+  merge(E2.RT_by_hit) %>%
+  merge(E2.RT_by_lead1) %>%
+  merge(E2.RT_by_lead2) %>%
+  merge(E2.RT_by_lead3);
+
+E2.RT_by_hit_lags_long <- E2.RT_by_hit_lags %>%
+  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
+  mutate(position = ifelse(position=="lag3_diff", 3,
+                           ifelse(position=="lag2_diff",2,
+                                  ifelse(position=="lag1_diff",1,
+                                         ifelse(position=="hit_diff",0,
+                                                ifelse(position=="lead1_diff",-1,
+                                                       ifelse(position=="lead2_diff",-2,-3
+                                                              ))))))) %>%
+  dplyr::select(subj_id,test_part,genuine_first,position,difference);
+
+E2.RT_by_hit_lags_long %>% 
+  filter(position<2 & position > -2)%>%
+  filter(subj_id %in% E2.simulators | subj_id %in% E2.suppressors) %>%
+  mutate(group = ifelse(subj_id %in% E2.simulators, 
+                        'simulators','suppressors')) %>%
+  group_by(test_part,position, group) %>%
+  summarise(median_difference=median(difference),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
+  theme_classic() +
+  facet_wrap(~group, nrow=2)
+
+ggsave('figures/E2_RT_by_lag_sim_sup.pdf',width=7)
+ggsave('figures/E2_RT_by_lag.png_sim_sup.png',width=4, height=3.5, dpi=300)
+
+```
+
+Visual inspection reveals similar temporal profiles in both groups (the higher standard error among suppressors is due to the lower number of subjects in this group).
+Specifically, in pretend games, cell selections that resulted in a hit were faster than those that resulted in a miss in both groups (suppressors: `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.suppressors)%>%pull(hit_diff)%>%t.test())$full_result`; simulators:  `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.simulators)%>%pull(hit_diff)%>%t.test())$full_result`), and cell selections that followed a hit were slower than those that followed a miss in both groups (suppressors: `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.suppressors)%>%pull(lag1_diff)%>%t.test())$full_result`; simulators:  `r apa_print(E2.RT_by_hit_lags %>% filter(test_part=='pretend' & subj_id %in% E2.simulators)%>%pull(lag1_diff)%>%t.test())$full_result`).
 
@@
 
+## Suppression or simulation?
+
+```{r hangman-suppression-or-simulation}
+
+E4.debrief_labeled <- read.csv('..\\experiments\\Hangman2\\data\\debrief.csv',na.strings=c(""," ","NA"))
+
+E4.simulators <- E4.debrief_labeled %>%
+  filter(simulation==1 & suppression==0) %>%
+  pull(subj_id)
+
+E4.suppressors <- E4.debrief_labeled %>%
+  filter(simulation==0 & suppression==1) %>%
+  pull(subj_id)
+
+```
+
+At the end of the experiment, players were asked whether they had a strategy for pretending or for detecting pretense.
+I (MM) manually labeled their responses according to three criteria: mentions of suppression (e.g., "Trying to ignore the hints"), simulation (e.g., "Trying to imagine how I would react"), and rules (e.g., "Checking the corners and center before random selection of tiles").
+Some answers got more than one label and some got none.
+Overall, `r E4.debrief_labeled$suppression%>%sum()` mentioned using suppression, `r E4.debrief_labeled$simulation%>%sum()` mentioned using simulation, and `r E4.debrief_labeled$heuristic%>%sum()` mentioned following a heuristic.
+
+Since there were only `r E4.suppressors%>%length` players who reported using suppression but not simulation, we focus our analysis only on the `r E4.simulators%>%length` who reported using simulation but no suppression.
+
+### Click optimality
+
+```{r battleship-p_click-sim, echo = FALSE, cache=TRUE}
+
+E4.sim.mean_P_click_rank <- E4.click_log_with_boards %>%
+  rbind(E4.click_log_with_boards_random) %>%
+  ungroup() %>%
+  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators)) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click_rank=mean(p_click_rank));
+
+E4.sim.mean_P_click_rank_misses_only <- E4.click_log_with_boards %>%
+  rbind(E4.click_log_with_boards_random) %>%
+  ungroup() %>%
+  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators) & !hit_bin) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click_rank=mean(p_click_rank));
+
+E4.sim.mean_P_click_rank %>%
+    filter(test_part != 'optimal') %>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(letter in word) - rank', y='number of players', title='Simulators only')+
+    theme(legend.position=c(0.10,0.85)) +
+    # scale_x_reverse() +
+    theme_classic();
+
+ggsave('figures/E4_pclick_rank_sim_only.pdf')
+ggsave('figures/E4_pclick_rank_sim_only.png',width=5,height=2,dpi=300)
+
+p<- E4.sim.mean_P_click_rank_misses_only %>%
+    filter(test_part != 'optimal') %>%
+    ggplot(aes(x=p_click_rank,fill=test_part))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(letter in word) - rank', y='number of players', title='Simulators only: misses')+
+    theme(legend.position=c(0.10,0.85)) +
+    # scale_x_reverse() +
+    theme_classic();
+
+ggsave('figures/E4_pclick_rank_sim_only_misses_only.pdf',p)
+ggsave('figures/E4_pclick_rank_sim_only_misses_only.png',p,width=5,height=2,dpi=300)
+
+```
+
+Among simulators, pretend games were significantly more optimal than non-pretend games (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`) even when restricting the analysis to letter selections that resulted in a miss (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`).
+Like in the full sapmle, simulators were more optimal when playing normally than when pretending (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`), also when focusing on misses (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+
+### RT by outcome
+
+```{r hangman-RT-by-accuracy-sim}
+
+E4.click_log <- E4.click_log %>%
+  group_by(subj_id,test_part,word) %>%
+ mutate(hit_bin = hit !=0,
+        lag3 = lag(hit,3)!=0,
+        lag2 = lag(hit,2)!=0,
+        lag1 = lag(hit,1)!=0,
+        lead1 = lead(hit,1)!=0,
+        lead2 = lead(hit,2)!=0,
+        lead3 = lead(hit,3)!=0)
+
+E4.RT_by_lag1 <- E4.click_log %>%
+  filter(click_number>1 & click_number<num_clicks-1) %>%
+  group_by(genuine_first,subj_id,test_part,lag1)%>%
+  summarise(RT=median(RT, na.rm=T))  %>%
+  spread(lag1,RT,sep='_') %>%
+  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
+
+E4.RT_by_hit <- E4.click_log %>%
+  filter(click_number>1 & click_number<num_clicks-1) %>%
+  group_by(genuine_first,subj_id,test_part,hit_bin)%>%
+  summarise(RT=median(RT, na.rm=T))  %>%
+  spread(hit_bin,RT,sep='_') %>%
+  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
+
+E4.RT_by_lead1 <- E4.click_log %>%
+  filter(click_number>1 & click_number<num_clicks-1) %>%
+  group_by(genuine_first,subj_id,test_part,lead1)%>%
+  summarise(RT=median(RT, na.rm=T))  %>%
+  spread(lead1,RT,sep='_') %>%
+  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
+
+E4.RT_by_hit_lags <- E4.RT_by_lag1 %>%
+  merge(E4.RT_by_hit) %>%
+  merge(E4.RT_by_lead1)
+
+E4.RT_by_hit_lags_long <- E4.RT_by_hit_lags %>%
+  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
+  mutate(position = ifelse(position=="lag1_diff",1,
+                                         ifelse(position=="hit_diff",0,-1))) %>%
+  dplyr::select(genuine_first,subj_id,test_part,position,difference);
+
+E4.RT_by_hit_lags_long %>%
+  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators)) %>%
+  group_by(test_part,position) %>%
+  summarise(median_difference=median(difference, na.rm=T),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_vline(xintercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t', title='Simulators only')+
+  theme_classic()
+
+ggsave('figures/Hangman2/RT_by_lag_acc_sim.pdf',width=7)
+ggsave('figures/Hangman2/RT_by_lag_acc_sim.png',width=5, height=3, dpi=300)
+
+E4.sim.RT_by_hit_lags_by_condition <- E4.RT_by_hit_lags_long %>%
+  filter(as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators)) %>%
+  spread(test_part,difference) %>%
+  mutate(interaction = pretend-nonpretend)
+
+```
+
+The simulators group showed a similar temporal profile to that of the entire sample.
+Specifically, in pretend games, cell selections that resulted in a hit were faster than those that resulted in a miss (`r apa_print(E4.RT_by_hit_lags %>% filter(test_part=='pretend' & as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators))%>%pull(hit_diff)%>%t.test())$full_result`), and cell selections that followed a hit were slower than those that followed a miss (`r apa_print(E4.RT_by_hit_lags %>% filter(test_part=='pretend' & as.numeric(substr(subj_id,1,3)) %in% c(E4.simulators))%>%pull(lag1_diff)%>%t.test())$full_result`).
+
 \newpage

### 2023-01-28 docs/preregisteredMethodsAndResults.Rmd
simulation or pretense in battleship and hangman (RT and optimality analysis)

@@
 ---
-title             : "Pretending not to know"
+title             : "Pretending not to know: pre-registered analyses"
 shorttitle        : "Pretending not to know"
@@
     scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    labs(x='p(letter in word) - rank', y='number of players')+
     theme(legend.position=c(0.10,0.85)) +
@@
 ```
+
 ### Hypothesis 7b (Letter frequency)
@@
     scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
-    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    labs(x='p(letter in English text) - rank', y='number of players')+
     theme(legend.position=c(0.10,0.85)) +

### 2023-01-28 docs/preregisteredMethodsAndResults.Rmd
typo

@@
 
-Participants were not significantly above chance in judging which of the two presented games came from a pretender (`r apa_print(t.test(E4.judge_summary_df$accuracy, mu=0.5))$full_result`).
+Participants were not significantly above chance in judging whether the presented game came from a pretender (`r apa_print(t.test(E4.judge_summary_df$accuracy, mu=0.5))$full_result`).
 This was the case also when focusing on words that judges encountered when pretending (`r apa_print(t.test(E4.judge_summary_df$accuracy_pretend_words, mu=0.5))$full_result`), and when focusing on words that judges encountered when playing normally (`r apa_print(t.test(E4.judge_summary_df$accuracy_nonpretend_words, mu=0.5))$full_result`).

### 2023-03-10 docs/ExploratoryResults.Rmd
Hangman correlation plot

@@
 
+## Correlation in number of misses
+
+```{r hangman-misses-per-word}
+
+E4.misses_per_word <- E4.click_df %>%
+  group_by(test_part,word) %>%
+  summarise(num_misses=median(num_misses)) %>%
+  spread(test_part,num_misses)
+
+```
+
+Some words were more difficult to reveal.
+This required in players making more unsuccessful letter guesses in attempting to reveal some words compared to others.
+When pretending, players made more unsuccessful letter guesses for the exact same words `r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(full_result)`.
+
+```{r hangman-correlation-plot}
+
+E4.misses_per_word %>%
+  ggplot(aes(x=nonpretend,y=pretend, label=word)) +
+  geom_abline(intercept=0,slope=1) +
+  geom_point()+
+  geom_text_repel()+
+  coord_fixed(ratio=1)+
+  scale_x_continuous(limits=c(0,5.5), breaks=seq(0,5))+
+  scale_y_continuous(limits=c(0,5.5), breaks=seq(0,5)) +
+  theme(legend.position = 'none')+
+  theme_bw() +
+  labs(title='number of misses')
+
+```
+
 ## Within-participant click entropy

### 2023-03-14 docs/paper.Rmd
paper draft

---
title             : "Pretending not to know"
shorttitle        : "Pretending not to know"

author: 
  - name          : "Matan Mazor"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Malet Street, London WC1E 7HX"
    email         : "mtnmzor@gmail.com"
  #   role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
  #     - Conceptualization
  #     - Writing - Original Draft Preparation
  #     - Writing - Review & Editing

  - name          : "Chaz Firestone"
    affiliation   : "3"

  - name          : "Ian Phillips"
    affiliation   : "3"
    # role:
    #   - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Birkbeck, University of London"
  - id            : "2"
    institution   : "Wellcome Centre for Human Neuroimaging, UCL"
  - id            : "3"
    institution   : "Johns Hopkins University"

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "pretense; metacognition; theory of mind"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel' #for word scatterplots
), '2022-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')
```

# Introduction

# Results

In two experiments, players played online versions of two information-seeking games: Battleship and Hangman. In these games, players' goal is to reveal an unknown world state (grid configuration in Battleship, a hidden word in Hangman) with as few steps (cell or letter selections) as possible. Critically, in addition to playing five games in their standard version, players also completed five pretend games in which the true world state was known to them all along, but their task was to behave as if it was not (see Fig. \@ref(fig:design)). We measured subjects' capacity to simulate a counterfactual knowledge state by comparing patterns of decisions and decision times in pretend and non-pretend games. Our full pre-registered results are available [online](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html).

As a first gross measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games. Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()`), with no difference between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend)-7,E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend)-7)%>%apa_print()%>%'$'(statistic)`). In Hangman, pretenders tended to make slightly more letter guesses on average, controlling for word length (mean = `r E4.median_misses%>%pull(pretend)%>%mean()`) than did non-pretenders (mean = `r E4.median_misses%>%pull(nonpretend)%>%mean()`; `r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(pretend),E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend))%>%apa_print()%>%'$'(statistic)`).

Furthermore, the ten target words we used in Hangman differed in difficulty: some required more letter guesses than others, controlling for word length. Strikingly, pretenders were highly successful in making more letter guesses when attempting to reveal those words that would have been easier to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(full_result)`). This correlation provides a first clue of an impressive capacity to act in accordance with how one would have acted if they didn't know what they know.

Having established that pretenders were able to mimic the total number of guesses in standard games, we next turned to the content of pretend and non-pretend guesses. In order to directly compare pretend and non-pretend guesses for the same board state, Battleship players completed two half-games in which they were instructed to continue the game from a half-completed state. Hundreds of cell selections for the same board state revealed a strong correlation between the spatial distributions of pretend and non-pretend guesses (), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.

To further examine the decisional processes behind this strong alignment, we compared the degree to which pretend and non-pretend guesses made sense within the context of the game. We approximated guess rationality by calculating the probability that a guess will be successful (in revealing a ship or word letters) given available information, $p(hit(x_{i}))$, and the posterior probability that one should make a guess, assuming a uniform prior over cells or letters $P(x_{i})=\frac{p(hit(x_i))}{\sum_{j=1}^{k}p(hit(x_j))}$. Given this posterior map, it makes sense to guess cells or letters for which $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3]. In the non-pretend versions of both games, guesses were more rational than expected by chance (Battleship:, Hangman:). Despite being less rational than non-pretend guesses, pretend guesses were more rational than expected by chance (Battleship:, Hangman:). Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship:, Hangman:): even when missing ships or letters, pretend guesses made sense given the limited information players pretended to have.

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2023-04-08 analysis/clickEntropyBattleship.Rmd
fixed entropy to log2 instead of log

@@
-```{r entropy-by-serial-position, echo=FALSE, cache=TRUE}
+```{r battleship-entropy-by-serial-position, echo=FALSE, cache=TRUE}
 
@@
 
-```{r entropy-pclick-correlation, echo=FALSE, cache=TRUE}
+```{r battleship-entropy-pclick-correlation, echo=FALSE, cache=TRUE}
 
 E2.entropy_summary <- E2.click_entropy_by_serial_position %>%
-  filter(click_number<8) %>%
+  filter(click_number<6) %>%
   group_by(subj_id,test_part) %>%
@@
   ungroup() %>%
-  filter(click_number<8) %>%
+  filter(click_number<6) %>%
   group_by(subj_id,test_part) %>%

### 2023-04-08 analysis/clickEntropyHangman.Rmd
fixed entropy to log2 instead of log

@@
-```{r entropy-by-serial-position, echo=FALSE, cache=TRUE}
+```{r hangman-entropy-by-serial-position, echo=FALSE, cache=TRUE}
 
@@
 
-e4.click_entropy_by_serial_position <- e4.click_log_with_boards %>%
+E4.click_entropy_by_serial_position <- E4.click_log_with_boards %>%
   mutate(flat_position = which(LETTERS==letter)) %>%
@@
 
-e4.click_entropy_by_serial_position_summary <- e4.click_entropy_by_serial_position %>%
+E4.click_entropy_by_serial_position_wide <- E4.click_entropy_by_serial_position %>%
+  spread(test_part,entropy) %>% 
+  mutate(diff=pretend-nonpretend)
+
+E4.click_entropy_by_serial_position_summary <- E4.click_entropy_by_serial_position %>%
   group_by(test_part,click_number) %>%
@@
 
-p <- e4.click_entropy_by_serial_position_summary %>%
+p <- E4.click_entropy_by_serial_position_summary %>%
   ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
@@
 ```
+
+```{r hangman-entropy-pclick-correlation, echo=FALSE, cache=TRUE}
+
+E4.entropy_summary <- E4.click_entropy_by_serial_position %>%
+  filter(click_number<6) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(entropy=mean(entropy,na.rm=T))
+
+E4.p_click_summary <- E4.click_log_with_boards %>%
+  rbind(E4.click_log_with_boards_random) %>%
+  ungroup() %>%
+  filter(click_number<6) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_click_rank=mean(p_click_rank))
+
+E4.entropy_cost <- E4.p_click_summary %>%
+  merge(E4.entropy_summary)%>%
+  group_by(subj_id)%>%
+  summarise(p_click_cost = p_click_rank[test_part=='pretend']-
+              p_click_rank[test_part=='nonpretend'],
+            entropy_cost = entropy[test_part=='pretend']-
+              entropy[test_part=='nonpretend'])
+
+```

### 2023-04-08 docs/ExploratoryResults.Rmd
fixed entropy to log2 instead of log

@@
   posterior[which(posterior==0)]=NA; #to avoid infinity*0
-  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
+  entropy <- -sum(unlist(posterior)*(unlist(log2(posterior))),na.rm=TRUE);
   return(entropy);

### 2023-04-08 docs/hangman.Rmd
fixed entropy to log2 instead of log

@@
 classoption       : "man"
-output            : papaja::apa6_word
+output            : bookdown::html_book
 ---
@@
   'ggrepel', #for ggplot words
-  'caret' #for cross validation
+  'caret', #for cross validation
   'kernlab' #for SVM
@@
   write.table('..\\experiments\\Hangman2\\data\\bonuses\\prolific_export3.csv', 
-              sep=',',col.names = F, row.names=F, quote=F)
+              sep=',', row.names=F, quote=F)
 

### 2023-04-08 docs/preregisteredMethodsAndResults.Rmd
fixed entropy to log2 instead of log

@@
 
+get_word_state <- function(word, asked) {
+  word_state = ''
+  for (letter in strsplit(word,'')[[1]]) {
+    value <- ifelse(grepl(toupper(letter), paste(toupper(asked),' ', sep='')), letter, '_')
+    word_state <- paste(word_state, value, sep='')
+  }
+  return(toupper(word_state))
+}
+
 word_is_consistent <- function(word_state, asked,  word) {
-
-  word_state = strsplit(toupper(word_state),'')[[1]];
-  word = strsplit(toupper(word),'')[[1]];
-  asked = strsplit(toupper(asked),'')[[1]];
-  not_in_word = asked[!(asked %in% word_state)]
-
-  # check if number of letters is consistent
-  if (length(word_state) != length(word)) {
-    return(FALSE)
-  }
-
-  # check if the letters in the masked word are consistent
-  for (i in seq_along(word)) {
-    if (word_state[i] != word[i]) {
-      if (word_state[i] != '_' | word[i] == ' ' | word[i] == '.') {
-      return(FALSE)
-      }
-    }
-  }
-
-  # check if any of the misses appear in the candidate word
-  if (length(Reduce(intersect,list(word,not_in_word)))>0) {
-    return(FALSE)
-  }
-
-  #if the function hasn't returned FALSE, return TRUE
-  return(TRUE)
-
+  return(get_word_state(word,asked)==toupper(word_state))
 }
@@
 
-};
+  }
 
@@
     posterior[which(posterior==0)]=NA; #to avoid infinity*0
-  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
+    entropy <- -sum(unlist(posterior)*(unlist(log2(posterior))),na.rm=TRUE);
     return(entropy);
@@
 #          p_click_rank = get_p_letter_rank(posterior,letter));
+#
 
@@
   filter(click_number>1) %>%
-  mutate(entropy_bin=round(entropy*3)/3) %>%
+  mutate(entropy_bin=round(entropy*2)/2) %>%
   group_by(test_part,subj_id,entropy_bin) %>% 

### 2023-04-15 analysis/formatDataForTuring.Rmd
orthogonalize

---
title: "Format Datan For Turing"
author: "Matan Mazor"
date: "2023-04-07"
output: html_document
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel', #for word scatterplots
  'pracma' # for GramSchmidt
), '2022-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')
```
## Calculate information gain

```{r hangman-information-gain, echo=FALSE}

get_word_state <- function(word, asked) {
  word_state = ''
  for (letter in strsplit(word,'')[[1]]) {
    value <- ifelse(grepl(toupper(letter), paste(toupper(asked),' ', sep='')), letter, '_')
    word_state <- paste(word_state, value, sep='')
  }
  return(toupper(word_state))
}

word_is_consistent <- function(word_state, asked,  word) {
  return(get_word_state(word,asked)==toupper(word_state))
}

get_word_entropy <- function(word_probabilities) {

    entropy <- -sum(unlist(word_probabilities$posterior)*
                    (unlist(log2(word_probabilities$posterior))),na.rm=TRUE);
  return(entropy)
};

get_letter_eig <- function(word_state, asked, category, 
                                          entropy, word_probabilities, letter) {

  if (grepl(toupper(letter),toupper(asked))) {
    return(NA)
  } else {
    word_probabilities <- word_probabilities %>%
      rowwise() %>%
      mutate(new_state = get_word_state(word,paste(asked,letter,sep=''))) %>%
      group_by(new_state) %>%
      summarize(prob = sum(posterior),
                entropy = -sum(unlist(posterior)*
                    (unlist(log2(posterior/sum(posterior)))),na.rm=TRUE))
    expected_entropy = sum(unlist(word_probabilities$prob)*unlist(word_probabilities$entropy))
    return(entropy-expected_entropy)
  }
}

#expected information gain
get_eig <- function(word_state,asked,category) {
  print(word_state)
  word_probabilities = get_word_probabilities(word_state,asked,category);
  entropy = get_word_entropy(word_probabilities)
  #if entropy is 0 there is no more uncertainty to resolve
  if (entropy==0) {
    print('entropy is 0')
    return(paste(rep(0,26),collapse=','))
  } else {
    letter_eig = c();
    for (letter in toupper(letters)) {
      if (grepl(letter,asked)) {
        letter_eig = c(letter_eig,
                                 NA)
      } else {
        letter_eig = c(letter_eig,
                               get_letter_eig(word_state,asked,category,entropy, 
                                              word_probabilities, letter)
      )
      }
    }
  }
  return(paste(letter_eig,collapse=','))
}

# E4.click_log_with_information_gain <- E4.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>%
#   rowwise()%>%
#   mutate(eig = get_eig(word_state,asked,category))

load('E4eig.RData')

# get_letter_posterior <- function(word_state, asked, category) {
# 
#   word_probabilities <- get_word_probabilities(word_state, asked, category);
# 
#   letter_probabilities = c();
#   for (letter in toupper(letters)) {
#     if (grepl(letter,asked)) {
#       letter_probabilities = c(letter_probabilities,
#                                NA)
#     } else {
#       letter_probabilities = c(letter_probabilities,
#                              word_probabilities %>%
#                                filter(grepl(letter,toupper(word))) %>%
#                                pull(posterior) %>%
#                                sum()
#     )
#     }
#   }
#   posterior = letter_probabilities/sum(letter_probabilities, na.rm=T)
# 
#   return(paste(posterior,collapse=','))
# 
# };
# 
# get_letter_prob_posterior <- function(asked) {
# 
#   letter_probabilities <- letter_df %>%
#     rowwise() %>%
#     mutate(prob = ifelse(
#       grepl(toupper(letter),toupper(asked)),
#       NA,
#       frequency)
#       ) %>%
#     ungroup()%>%
#     mutate(
#       posterior = prob/sum(prob,na.rm=T)
#     )
# 
#   return(paste(letter_probabilities$posterior,collapse=','))
# 
# };
# 
# get_p_letter <- function(posterior, letter) {
#   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
#   return(posterior[which(toupper(letters)==letter)])
# }
# 
# get_p_letter_rank <- function(posterior, letter) {
#   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
#   rank_posterior = rank(-posterior)
#   return(rank_posterior[which(toupper(letters)==letter)])
# }
# 
# get_letter_posterior_entropy <- function(posterior) {
#   posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
#   posterior[which(posterior==0)]=NA; #to avoid infinity*0
#   entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
#   return(entropy);
# }

# E4.click_log_with_boards <- E4.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>%
#   rowwise()%>%
#   mutate(posterior = get_letter_posterior(word_state,asked,category),
#          p_click = get_p_letter(posterior,letter),
#          entropy = get_letter_posterior_entropy(posterior),
#          p_click_rank = get_p_letter_rank(posterior,letter));

# get_random_letter <- function(asked) {
# 
#   asked = strsplit(tolower(asked),'')[[1]];
#   not_asked = setdiff(letters,asked)
#   return(toupper(not_asked)%>%sample(1))
# 
# }
# 
# get_most_frequent_letter <- function(asked) {
# 
#    lp <- letter_df %>%
#     rowwise() %>%
#     mutate(prob = ifelse(
#       grepl(toupper(letter),toupper(asked)),
#       NA,
#       frequency)
#       ) %>%
#     ungroup()
# 
#      return(lp[which.max(lp$prob),]$letter%>%toupper())
# 
# }

# E4.click_log_with_boards_random <- E4.click_log_with_boards %>%
#   mutate(letter = get_random_letter(asked),
#          p_click = get_p_letter(posterior,letter),
#          p_click_rank = get_p_letter_rank(posterior,letter),
#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
#          test_part='random');

# E4.click_log_with_boards_most_frequent <- E4.click_log_with_boards %>%
#   mutate(letter = get_most_frequent_letter(asked),
#          p_click = get_p_letter(posterior,letter),
#          p_click_rank = get_p_letter_rank(posterior,letter),
#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
#          test_part='most_frequent');

# E4.letter_prob_click_log_with_boards <- E4.click_log %>%
#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,asked,click_number,letter,RT,hit_bin) %>%
#   rowwise()%>%
#   mutate(posterior = get_letter_prob_posterior(asked),
#          p_click = get_p_letter(posterior,letter),
#          entropy = get_letter_posterior_entropy(posterior),
#          p_click_rank = get_p_letter_rank(posterior,letter));

# E4.letter_prob_click_log_with_boards_random <- E4.letter_prob_click_log_with_boards %>%
#   mutate(letter = get_random_letter(asked),
#          p_click = get_p_letter(posterior,letter),
#          p_click_rank = get_p_letter_rank(posterior,letter),
#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
#          test_part='random');

# save(E4.click_log_with_boards,
#      E4.click_log_with_boards_random,
#      E4.click_log_with_boards_most_frequent,
#      E4.letter_prob_click_log_with_boards,
#      E4.letter_prob_click_log_with_boards_random,
#      file='E4.click_log_with_boards.RData')

load('../docs/E4.click_log_with_boards.RData') # this file includes all the above dataframes.

```

```{r format}

E4.letters_posterior_wide <- E4.click_log_with_boards %>% 
  dplyr::select(subj_id, test_part,word,posterior,letter, hit_bin) %>%
  separate(posterior, sep=",", into=paste0("post_",LETTERS)) %>%
  mutate_at(vars(post_A:post_Z), as.numeric) %>% 
  mutate(across(starts_with("post"), ~ifelse(is.na(.), 0, 1), .names = "{sub('post', 'mask', .col)}"),
         across(starts_with("post"), ~ifelse(is.na(.), 0, .)))

E4.letters_prior_wide <- E4.letter_prob_click_log_with_boards %>% 
  dplyr::select(subj_id, test_part,word,posterior,letter) %>%
  separate(posterior, sep=",", into=paste0("prior_",LETTERS)) %>%
  mutate_at(vars(prior_A:prior_Z), as.numeric) %>% 
  mutate(across(starts_with("prior"), ~ifelse(is.na(.), 0, 1), .names = "{sub('prior', 'mask', .col)}"),
         across(starts_with("prior"), ~ifelse(is.na(.), 0, .)))

E4.letters_eig_wide <- E4.click_log_with_information_gain %>% 
  dplyr::select(subj_id, test_part,word,eig,letter) %>%
  separate(eig, sep=",", into=paste0("eig_",LETTERS)) %>%
  mutate_at(vars(eig_A:eig_Z), as.numeric) %>% 
  mutate(across(starts_with("eig"), ~ifelse(is.na(.), 0, .)))

E4.letters_features_wide <- E4.letters_posterior_wide %>%
  merge(E4.letters_prior_wide) %>%
  merge(E4.letters_eig_wide)

E4.letters_features_wide %>%
  write.csv('dataForTuring/E4.csv')
```

```{r normalize-features}
E4.letters_features_normalized <- E4.letters_features_wide;

post_columns <- paste0("post_",LETTERS)
prior_columns <- paste0("prior_",LETTERS)
eig_columns <- paste0("eig_",LETTERS)

for (subj in E4.letters_features_normalized$subj_id%>%unique()) {

  subj_rows <- which(E4.letters_features_normalized$subj_id==subj)
  post_subset <- E4.letters_features_normalized[subj_rows,post_columns];
  # Calculate the global mean and standard deviation for the subset
  post_mean <- mean(as.vector(unlist((post_subset))), na.rm = TRUE)
  post_sd <- sd(as.vector(unlist((post_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_post_subset <- post_subset %>% mutate(across(everything(), ~ (. - post_mean) / post_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_normalized[subj_rows, post_columns] <- normalized_post_subset

  prior_subset <- E4.letters_features_normalized[subj_rows,prior_columns];
  # Calculate the global mean and standard deviation for the subset
  prior_mean <- mean(as.vector(unlist((prior_subset))), na.rm = TRUE)
  prior_sd <- sd(as.vector(unlist((prior_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_prior_subset <- prior_subset %>% mutate(across(everything(), ~ (. - prior_mean) / prior_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_normalized[subj_rows, prior_columns] <- normalized_prior_subset

  eig_subset <- E4.letters_features_normalized[subj_rows,eig_columns];
  # Calculate the global mean and standard deviation for the subset
  eig_mean <- mean(as.vector(unlist((eig_subset))), na.rm = TRUE)
  eig_sd <- sd(as.vector(unlist((eig_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_eig_subset <- eig_subset %>% mutate(across(everything(), ~ (. - eig_mean) / eig_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_normalized[subj_rows, eig_columns] <- normalized_eig_subset;
}

E4.letters_features_normalized %>%
  write.csv('dataForTuring/E4norm.csv')

```

```{r orthogonalize-eig}
E4.letters_features_orth <- E4.letters_features_normalized;

post_columns <- paste0("post_",LETTERS)
prior_columns <- paste0("prior_",LETTERS)
eig_columns <- paste0("eig_",LETTERS)
mask_columns <- paste0("mask_", LETTERS)

num_rows <- nrow(E4.letters_features_orth)

pb = txtProgressBar(min = 0, max = num_rows, initial = 0)

for (i_row in 1:num_rows) {

  mask_subset <- unlist(E4.letters_features_orth[i_row,mask_columns], use.names=F);
  mask_indices <- which(mask_subset==1);
  post_subset <- unlist(E4.letters_features_orth[i_row,post_columns[mask_indices]], use.names=F);
  prior_subset <- unlist(E4.letters_features_orth[i_row,prior_columns[mask_indices]], use.names=F);
  eig_subset <- unlist(E4.letters_features_orth[i_row,eig_columns[mask_indices]], use.names=F);

    matrix <- matrix(c(post_subset,prior_subset,eig_subset),length(mask_indices),3)
    if (Rank(mask_indices)==3) {
    orthogonalized_matrix <- gramSchmidt(matrix)$Q

    orth_post_subset <- orthogonalized_matrix[,1];
    orth_prior_subset <- orthogonalized_matrix[,2];
    orth_eig_subset <- orthogonalized_matrix[,3];
  } else {
    orth_post_subset <- post_subset;
    orth_prior_subset <- prior_subset;
    orth_eig_subset <- eig_subset;
  }

 # Replace the original columns with the orthogonalized ones
  E4.letters_features_orth[i_row, post_columns[mask_indices]] <- orth_post_subset
  E4.letters_features_orth[i_row, prior_columns[mask_indices]] <- orth_prior_subset
  E4.letters_features_orth[i_row, eig_columns[mask_indices]] <- orth_eig_subset

  setTxtProgressBar(pb, i_row)

}

close(pb)

# NORMALIZE!

num_subj <- E4.letters_features_orth$subj_id%>%unique()%>%length()
pb = txtProgressBar(min = 0, max = num_subj, initial = 0)

i_subj=0
for (subj in E4.letters_features_orth$subj_id%>%unique()) {
  i_subj = i_subj+1
  subj_rows <- which(E4.letters_features_orth$subj_id==subj)
  post_subset <- E4.letters_features_orth[subj_rows,post_columns];
  # Calculate the global mean and standard deviation for the subset
  post_mean <- mean(as.vector(unlist((post_subset))), na.rm = TRUE)
  post_sd <- sd(as.vector(unlist((post_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_post_subset <- post_subset %>% mutate(across(everything(), ~ (. - post_mean) / post_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_orth[subj_rows, post_columns] <- normalized_post_subset

  prior_subset <- E4.letters_features_orth[subj_rows,prior_columns];
  # Calculate the global mean and standard deviation for the subset
  prior_mean <- mean(as.vector(unlist((prior_subset))), na.rm = TRUE)
  prior_sd <- sd(as.vector(unlist((prior_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_prior_subset <- prior_subset %>% mutate(across(everything(), ~ (. - prior_mean) / prior_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_orth[subj_rows, prior_columns] <- normalized_prior_subset

  eig_subset <- E4.letters_features_orth[subj_rows,eig_columns];
  # Calculate the global mean and standard deviation for the subset
  eig_mean <- mean(as.vector(unlist((eig_subset))), na.rm = TRUE)
  eig_sd <- sd(as.vector(unlist((eig_subset))), na.rm = TRUE)

  # Normalize the selected columns
  normalized_eig_subset <- eig_subset %>% mutate(across(everything(), ~ (. - eig_mean) / eig_sd))

  # Replace the original columns with the normalized ones
  E4.letters_features_orth[subj_rows, eig_columns] <- normalized_eig_subset;

  setTxtProgressBar(pb, i_subj)

}

E4.letters_features_orth %>%
  write.csv('dataForTuring/E4orth.csv')

```

```{r}

E4.letters_features_wide_misses <- E4.letters_features_wide %>%
  filter(!hit_bin)

E4.subjects_with_enough_misses <- E4.letters_features_wide_misses %>%
  group_by(subj_id,test_part) %>%
  summarise(n=n()) %>%
  group_by(subj_id) %>%
  summarise(n=min(n)) %>%
  filter(n>5) %>%
  pull(subj_id)

E4.letters_features_wide_misses %>%
  filter(subj_id %in% E4.subjects_with_enough_misses) %>%
  write.csv('dataForTuring/E4misses.csv')

```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2023-04-16 analysis/formatDataForTuring.Rmd
model-based analysis with orthogonalized regressors

@@
     matrix <- matrix(c(post_subset,prior_subset,eig_subset),length(mask_indices),3)
-    if (Rank(mask_indices)==3) {
+    if (Rank(matrix)==3) {
     orthogonalized_matrix <- gramSchmidt(matrix)$Q

### 2023-04-16 docs/ExploratoryResults.Rmd
model-based analysis with orthogonalized regressors

@@
 
+## Model-based analysis
+
+We considered three factors that contribute to letter selections in Hangman:
+
+1. The prior probability of a letter in the English language $p(X)$.
+2. The posterior probability that a letter is present in the target word, given the game state $S$ and the category $p(X|S)$.
+3.
+The expected information gain this letter provides regarding the identity of the target word $H(S)-\sum p(x_v)H(S,x_v)$ (where $x_v$ is one potential outcome of guessing the letter $X$, for example, finding it in the second and fifth positions).
+
+To quantify the contribution of each of these factors to letter selection, we fitted a multinomial logistic regression model to the data of each player.
+We compared the ensuing coefficients in 'pretend' and 'non-pretend' conditions.
+
+```{r hangman-load-coefficients, echo=FALSE}
+
+#get list of subjects
+E4.data_for_turing <- read.csv('../analysis/dataForTuring/E4norm.csv')
+E4.subjects <- E4.data_for_turing$subj_id%>%unique()
+
+#initialize
+E4.model_coefficients_raw <- data.frame()
+
+#fill the array
+for (i_subj in 1:length(E4.subjects)) {
+  E4.model_coefficients_raw <- E4.model_coefficients_raw %>%
+    rbind(read.csv(sprintf('../analysis/model_fits/E4/subj%d_normalized.csv',i_subj)) %>%
+            mutate(subj_id=E4.subjects[i_subj]) %>%
+            dplyr::select(subj_id,term,estimate))
+}
+
+# Organize in a more convenient form
+# non-pretend dataframe
+
+E4.np_coefficients <- E4.model_coefficients_raw %>% 
+  group_by(subj_id) %>% 
+  summarize(test_part = 'nonpretend',
+            prior = estimate[term == 'coef_prior'],
+            post = estimate[term == 'coef_post'],
+            eig = estimate[term == 'coef_eig'])
+
+E4.p_coefficients <- E4.model_coefficients_raw %>% 
+  group_by(subj_id) %>% 
+  summarize(test_part='pretend',
+            prior = estimate[term == 'coef_prior'] +
+              estimate[term == 'coef_prior_delta'],
+            post = estimate[term == 'coef_post'] +
+              estimate[term == 'coef_post_delta'],
+            eig = estimate[term == 'coef_eig'] +
+              estimate[term == 'coef_eig_delta'])
+
+E4.model_coefficients <- E4.np_coefficients %>%
+  rbind(E4.p_coefficients)
+```
+
+```{r hangman-coefficients_plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.model_coefficients %>%
+  gather('term','coefficient',prior,post,eig) %>%
+  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend'),
+                            labels=c('non-pretend','pretend')),
+         term = factor(term, levels = c('prior','post','eig'), labels = c('prior','posterior','expected information gain'))) %>%
+  ggplot(aes(x=coefficient,fill=test_part, color=test_part))+
+  geom_histogram( alpha=0.6, position = 'identity', bins=100) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#146464", "#404080")) +
+  labs(x='coefficient (a.u.)', y='number of players')+
+  theme(legend.position=c(0.10,0.85)) +
+  scale_x_continuous(limits=c(-2,5)) +
+  facet_wrap(~term, nrow=3)+
+  # scale_x_reverse() +
+  theme_classic();
+
+ggsave('../docs/figures/E4coefs.png',width=5,height=4.5)
+```
+
+All three terms had a positive effect in both conditions.
+Prior probability (base rate) had a stronger effect in pretend games (`r apa_print(E4.model_coefficients %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+Hit posterior probability showed no such difference, and had a similar effect on players' decisions in both conditions (`r apa_print(E4.model_coefficients %>% dplyr::select(subj_id,test_part,post)%>%spread(test_part,post)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+Finally, expected information gain had a weaker effect in pretend games (`r apa_print(E4.model_coefficients %>% dplyr::select(subj_id,test_part,eig)%>%spread(test_part,eig)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+
+```{r hangman-load-coefficients-orth, echo=FALSE}
+
+#get list of subjects
+
+#initialize
+E4.model_coefficients_orth_raw <- data.frame()
+
+#fill the array
+for (i_subj in 1:length(E4.subjects)) {
+  E4.model_coefficients_orth_raw <- E4.model_coefficients_orth_raw %>%
+    rbind(read.csv(sprintf('../analysis/model_fits/E4/subj%d_orth.csv',i_subj)) %>%
+            mutate(subj_id=E4.subjects[i_subj]) %>%
+            dplyr::select(subj_id,term,estimate))
+}
+
+# Organize in a more convenient form
+# non-pretend dataframe
+
+E4.np_orth_coefficients <- E4.model_coefficients_orth_raw %>% 
+  group_by(subj_id) %>% 
+  summarize(test_part = 'nonpretend',
+            prior = estimate[term == 'coef_prior'],
+            post = estimate[term == 'coef_post'],
+            eig = estimate[term == 'coef_eig'])
+
+E4.p_orth_coefficients <- E4.model_coefficients_orth_raw %>% 
+  group_by(subj_id) %>% 
+  summarize(test_part='pretend',
+            prior = estimate[term == 'coef_prior'] +
+              estimate[term == 'coef_prior_delta'],
+            post = estimate[term == 'coef_post'] +
+              estimate[term == 'coef_post_delta'],
+            eig = estimate[term == 'coef_eig'] +
+              estimate[term == 'coef_eig_delta'])
+
+E4.model_coefficients_orth <- E4.np_orth_coefficients %>%
+  rbind(E4.p_orth_coefficients)
+```
+
+Given colinearity between model predictors, we suspected that the effects of expected information gain may be driven by posterior probability effects.
+To isolate unique contributions of prior probability and expected information gain, we repeated the same analysis, but this time serially orthogonalizing the features (in the order [posterior, prior, expected information gain], and using the Gram Schmidt algorithm) prior to fitting the multinomial regression model.
+This time, expected information gain had no significant effect on decision in non-pretend (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='nonpretend')%>%pull(eig)%>%t.test())$statistic`) and pretend games (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='pretend')%>%pull(eig)%>%t.test())$statistic`), rendering it possible that the previously observed effects were confounded with prior and posterior probability.
+In contrast, both prior and posterior probability had a significant positive effect, with prior probability having a stronger effect in pretend games (`r apa_print(E4.model_coefficients_orth %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+
+```{r hangman-coefficients_plot_orth, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
+
+E4.model_coefficients_orth %>%
+  gather('term','coefficient',prior,post,eig) %>%
+  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend'),
+                            labels=c('non-pretend','pretend')),
+         term = factor(term, levels = c('prior','post','eig'), labels = c('prior','posterior','expected information gain'))) %>%
+  ggplot(aes(x=coefficient,fill=test_part, color=test_part))+
+  geom_histogram( alpha=0.6, position = 'identity', bins=100) +
+  scale_fill_manual(values=c("#69b3a2", "#404080")) +
+  scale_color_manual(values=c("#146464", "#404080")) +
+  labs(x='coefficient (a.u.)', y='number of players', title='Orthogonalized vectors')+
+  theme(legend.position=c(0.10,0.85)) +
+  scale_x_continuous(limits=c(-2,5)) +
+  facet_wrap(~term, nrow=3)+
+  # scale_x_reverse() +
+  theme_classic();
+
+ggsave('../docs/figures/E4coefs_orth.png',width=5,height=4.5)
+```
+
 ## Suppression or simulation?

### 2023-04-24 docs/paper.Rmd
simulated hangman games and n misses analysis

@@
 load('../.RData')
+
+#color scheme
+colors <-  list()
+colors$p <- "#404080"
+colors$np <- "#69b3a2"
+colors$random <- 'gray'
+  colors$greedy <- 'black'
 ```
@@
 
-In two experiments, players played online versions of two information-seeking games: Battleship and Hangman.
-In these games, players' goal is to reveal an unknown world state (grid configuration in Battleship, a hidden word in Hangman) with as few steps (cell or letter selections) as possible.
-Critically, in addition to playing five games in their standard version, players also completed five pretend games in which the true world state was known to them all along, but their task was to behave as if it was not (see Fig.
+In two experiments, players played online versions of two information-seeking games: Battleship (N=`r E2.df$subj_id%>%unique()%>%length()`) and Hangman (N=`r E4.df$subj_id%>%unique()%>%length()`).
+These games typically start in a state of ignorance, where players' goal is to reveal an unknown world state (grid configuration in Battleship, a hidden word in Hangman) with as few steps (cell or letter selections) as possible.
+Critically, in addition to playing five games in their standard version, players also completed five pretend games in which the true world state was known to them from the start, but their task was to behave as if it was not (see Fig.
 \@ref(fig:design)).
-We measured subjects' capacity to simulate a counterfactual knowledge state by comparing patterns of decisions and decision times in pretend and non-pretend games.
+We measured their capacity to simulate a counterfactual knowledge state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games.
 Our full pre-registered results are available [online](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html).
 
+```{r battleship-design, echo=FALSE, fig.cap="Experimental Design in Exp.
+1 (left) and 2 (right).
+In non-pretend games, players revealed ships by guessing cells in a grid (A) or revealed a word by guessing letters (E).
+In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (F), but asked players to play as if they didn't have access to this information.
+In half games (C and G), players continued the game from a given state, with the content of some cells and the position of some letters already revealed.
+Lastly, players watched replays of the games of previous players and guessed were pretend games (D and H).", out.width = '100%'}
+knitr::include_graphics("figures/methods.png")
+```
+
 As a first gross measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games.
-Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()`), with no difference between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend)-7,E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend)-7)%>%apa_print()%>%'$'(statistic)`).
-In Hangman, pretenders tended to make slightly more letter guesses on average, controlling for word length (mean = `r E4.median_misses%>%pull(pretend)%>%mean()`) than did non-pretenders (mean = `r E4.median_misses%>%pull(nonpretend)%>%mean()`; `r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(pretend),E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend))%>%apa_print()%>%'$'(statistic)`).
+Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()`).
+In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()`; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()`).
+Despite an overall bias in the number of guesses, pretend Hangman games provided evidence for a more impressive, item-specific alignment: pretenders were successful in making more letter guesses when attempting to reveal words that would have been easier to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`).
+This correlation provides a first clue of a human capacity to act in accordance with a counterfactual knowledge state.
 
-Furthermore, the ten target words we used in Hangman differed in difficulty: some required more letter guesses than others, controlling for word length.
-Strikingly, pretenders were highly successful in making more letter guesses when attempting to reveal those words that would have been easier to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(full_result)`).
-This correlation provides a first clue of an impressive capacity to act in accordance with how one would have acted if they didn't know what they know.
+```{r num-misses, echo=FALSE, fig.cap="Panels A and B: median number of misses Battleship and Hangman games, in non-pretend (green) and pretend (purple) games.
+For reference, the expected number of misses is also indicated for a fully random agent, and for a myopic agent that maximizes the probability of a hit.
+C: The median number of misses in Hangman for pretend and non-pretend games, as a function of the target word.
+", out.width = '100%'}
 
-Having established that pretenders were able to mimic the total number of guesses in standard games, we next turned to the content of pretend and non-pretend guesses.
+load('../analysis/simulated_hangman_games.RData')
+
+E2.random_mean_misses <- E2.random_click_df %>%
+           group_by(subj_id) %>%
+           summarise(num_misses=median(num_clicks)-7) %>%
+           pull(num_misses) %>%
+                  mean()
+
+E2.greedy_mean_misses <- E2.optimal_click_df %>%
+           group_by(subj_id) %>%
+           summarise(num_misses=median(num_clicks)-7) %>%
+           pull(num_misses) %>%
+                  mean()
+
+E4.random_mean_misses <- simulated_random_hangman_games %>% 
+  group_by(subj_id) %>%
+  summarise(num_misses=median(nmisses)) %>%
+  pull(num_misses) %>%
+  mean()
+
+E4.greedy_mean_misses <- simulated_optimal_hangman_games %>% 
+  pull(nmisses) %>%
+  mean()
+
+E2.click_df %>%
+  dplyr::select(subj_id,test_part, num_clicks) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_clicks)-7) %>%
+  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
+  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
+  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position='none') +
+  theme_classic() +
+  geom_vline(xintercept = E2.greedy_mean_misses, color='black', linetype=2) + 
+  annotate("text", x = E2.greedy_mean_misses, y = 75, label = "Greedy", angle = 90, vjust = -1, size = 5)+
+  geom_vline(xintercept = E2.random_mean_misses, color='black', linetype=2) +
+  annotate("text", x = E2.random_mean_misses, y = 75, label = "Random", angle = 90, vjust = -1, size = 5)+
+  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
+  theme(legend.pos='na')
+
+ggsave('figures/num_misses_E2.png',width=4.5, height=2.5)
+
+E4.click_df %>%
+  dplyr::select(subj_id,test_part, num_misses) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_misses)) %>%
+  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
+  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
+  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position='none') +
+  theme_classic() +
+  geom_vline(xintercept = E4.greedy_mean_misses, color='black', linetype=2) + 
+  annotate("text", x = E4.greedy_mean_misses, y = 95, label = "Greedy", angle = 90, vjust = -1, size = 5)+
+  geom_vline(xintercept = E4.random_mean_misses, color='black', linetype=2) +
+  annotate("text", x = E4.random_mean_misses, y = 95, label = "Random", angle = 90, vjust = -1, size = 5)+
+  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
+  theme(legend.pos='na')
+
+ggsave('figures/num_misses_E4.png',width=4.5, height=2.5)
+knitr::include_graphics("figures/num_misses.png")
+
+```
+Having established an alignment in the total number of guesses, we next turned to the content of pretend and non-pretend guesses.
 In order to directly compare pretend and non-pretend guesses for the same board state, Battleship players completed two half-games in which they were instructed to continue the game from a half-completed state.
-Hundreds of cell selections for the same board state revealed a strong correlation between the spatial distributions of pretend and non-pretend guesses (), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.
+Hundreds of cell selections for the same board state revealed a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_nonpretend)$table$estimate`, $p$ `r apa_print(cor_A_nonpretend)$table$p.value`; board B: $r=$ `r apa_print(cor_B_nonpretend)$table$estimate`, $p$ `r apa_print(cor_B_nonpretend)$table$p.value`), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.
 
 To further examine the decisional processes behind this strong alignment, we compared the degree to which pretend and non-pretend guesses made sense within the context of the game.
-We approximated guess rationality by calculating the probability that a guess will be successful (in revealing a ship or word letters) given available information, $p(hit(x_{i}))$, and the posterior probability that one should make a guess, assuming a uniform prior over cells or letters $P(x_{i})=\frac{p(hit(x_i))}{\sum_{j=1}^{k}p(hit(x_j))}$.
-Given this posterior map, it makes sense to guess cells or letters for which $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3].
-In the non-pretend versions of both games, guesses were more rational than expected by chance (Battleship:, Hangman:).
-Despite being less rational than non-pretend guesses, pretend guesses were more rational than expected by chance (Battleship:, Hangman:).
-Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship:, Hangman:): even when missing ships or letters, pretend guesses made sense given the limited information players pretended to have.
+When playing Battleship and Hangman, it makes sense to guess cells or letters for which the probability of hitting a ship or revealing a letter is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases, @audinot2014optimal, Section 3.3].
+In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank_wide%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`): even when missing ships or letters, pretend guesses made sense given the limited information players pretended to have.
+
+Good pretense is a function not only of the number and content of players' decisions, but also of their timing.
+Here too, pretend games showed the same qualitative patterns as non-pretend games.
+Like non-pretenders, pretenders were faster in their successful guesses (Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median()%>%round()%>%abs()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median()%>%round()%>%abs()` ms;  Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()%>%abs()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()%>%abs()` ms) and slowed down immediately after a hit (Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()%>%abs()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()%>%abs()` ms;  Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()%>%abs()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()%>%abs()` ms).
+Furthermore, decision uncertainty (quantified as the posterior entropy over cell or letter options) had a similar quadratic effect on decision time in both non-pretend and pretend versions of the games, with the slowest responses associated with mid-range levels of entropy (see Fig.
+).
+
+However impressive, this capacity for simulating a state of ignorance was not perfect.
+Most markedly, in their attempt to behave as if they didn't know the true state of the game, pretenders tended to follow rigid heuristics and rules, as evidenced by higher predictability of decision sequences across different games of the same player (quantified as the within-participant sequential guess entropy for clicks 1-5, see [online supplementary materials](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#within-participant-click-entropy)).
+In Hangman, a popular heuristic is to guess frequent letters such as vowels.
+We found that pretenders guessed more high-frequency letters than did non-pretenders, suggesting they relied more on this strategy.
+This limitation was especially evident in Hangman half-games, where players completed the game from a half-completed state.
+When asked to reveal the hidden fruit "_A_A_A", `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('banana', 'papaya'))%>%pull(bananahand)%>%mean())`% of the non-pretenders guessed one of the letters 'B' or 'N'.
+Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(E4.letter_choice_frequenc_wide %>% filter(test_part=='pretend' & word=='banana') %>% pull('B') %>% mean()*100)`%).
+Importantly, half of the pretenders were given different information: they were told that the hidden word is the less prototypical fruit PAPAYA.
+Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', revealing that many pretenders were unable to predict that the first fruit that would come to their minds is BANANA, not PAPAYA.
+A similar pattern was observed for the prototypical body part word HA(ND) and its surprising counterpart HA(IR): when playing normally, `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('hand', 'hair'))%>%pull(bananahand)%>%mean())`% of the players selected letters that are consistent with the prototypical option HAND.
+Pretenders for whom the word was HAND showed the same tendency (`r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hand')%>%pull(bananahand)%>%mean())`%), in contrast to only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hair')%>%pull(bananahand)%>%mean())`% of the pretenders who saw the word HAIR.
+
+Finally, after performing pretend and non-pretend games, participants were presented with game replays of previous players and judged who was pretending and who was playing for real.
+In both games, pretense detection was at chance (Battleship: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%, Hangman: `r round(E4.judge_summary_df$accuracy%>%mean()*100)`).
+Moreover, we find no sign of a correlation between pretense quality (measured as players' ability to trick judges to think they were not pretending) and pretense detection ability (measured as proportion correct; Battleship: `r apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate, method='spearman'))$estimate`, Hangman: `r apa_print(cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman'))$estimate`), suggesting the two capacities rely on at least partly different cognitive processes.
 

### 2023-04-24 docs/preregisteredMethodsAndResults.Rmd
simulated hangman games and n misses analysis

@@
 To sink all ships, players had to click on at least 7 and at most 25 squares.
-A simulated player that clicks randomly had a median total click number of `r num_clicks_sim%>%median()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a median total click number of `r num_clicks_optimal_sim%>%median()`.
-Among our players, the median number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games.
+A simulated player that clicks randomly had a mean total click number of `r num_clicks_sim%>%mean()%>%printnum()`, and a near-optimal greedy player that consistently selected the square with the highest objective probability of containing a ship had a mean total click number of `r num_clicks_optimal_sim%>%mean()%>%printnum()`.
+Among our players, the mean number of clicks was `r E2.median_clicks%>%pull(nonpretend)%>%mean()%>%printnum()` in non-pretend games and `r E2.median_clicks%>%pull(pretend)%>%mean()%>%printnum()` in pretend games.
 
@@
 To maximize statistical power, this hypothesis was tested on the pooled data from both groups of participants.
-We find no evidence for such a correlation (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate))$full_result`).
-Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend))$full_result`).
+We find no evidence for such a correlation (`r  apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate, method='spearman'))$full_result`).
+Furthermore, we find no significant correlation between participants' accuracy in detecting the pretender and their decision optimality in pretend games (measured as the mean rank posterior probability of their cell selections; `r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend, method='spearman'))$full_result`), nor with the cost to optimality relative to non-pretend games (`r apa_print(cor.test(E2.judge_summary$judge,E2.judge_summary$pretend-E2.judge_summary$nonpretend, method='spearman'))$full_result`).
 
@@
 
-### Hypothesis 3 (number of misses)
-
-The median number of misses was `r E4.median_misses%>%pull(nonpretend)%>%median()` in non-pretend games and `r E4.median_misses%>%pull(pretend)%>%median()` in pretend games.
-The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
-
 ```{r hangman-h3-plot, echo-FALSE, cache=TRUE, fig.width=10,fig.height=5}
@@
 
+get_optimal_letter <- function(posterior) {
+
+  posterior = scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  return(LETTERS[which.max(posterior)])
+
+}
+
 # E4.click_log_with_boards_random <- E4.click_log_with_boards %>%
@@
 #   mutate(letter = get_most_frequent_letter(asked),
+#          p_click = get_p_letter(posterior,letter),
+#          p_click_rank = get_p_letter_rank(posterior,letter),
+#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
+#          test_part='most_frequent');
+
+# E4.click_log_with_boards_optimal <- E4.click_log_with_boards %>%
+#   mutate(letter = get_optimal_letter(posterior),
 #          p_click = get_p_letter(posterior,letter),
@@
 
+```
+
+### Hypothesis 3 (number of misses)
+
+The mean number of misses was `r E4.median_misses%>%pull(nonpretend)%>%mean()` in non-pretend games and `r E4.median_misses%>%pull(pretend)%>%mean()` in pretend games.
+The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
+
+```{r simulate_nclicks-hangman, echo=FALSE, cache=TRUE}
+
+## simulate number of clicks for random play
+
+words_and_categories <- E4.click_df %>%
+  group_by(word) %>%
+  summarise(category=category[1])
+
+mask_word <- function(word, asked) {
+# Replace non-space characters with underscores
+  pattern <- sprintf("[^[:space:]%s]", toupper(asked))
+  result_string <- gsub(pattern, "_", toupper(word))
+  return(result_string)
+}
+
+simulate_optimal_hangman <- function(word,category) {
+  #initialize
+  word_state <- mask_word(toupper(word),''); 
+  asked <- ''
+  nclicks <- 0;
+  nmisses <- 0;
+
+  while (grepl('_',word_state) & nclicks<26) {
+
+    posterior <- get_letter_posterior(toupper(word_state),toupper(asked),category);
+    posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE)
+
+    best_shot = LETTERS[which.max(posterior)];
+
+    #click on best shot
+    nclicks = nclicks+1;
+    asked = paste(asked,best_shot)
+    new_word_state = mask_word(word,asked)
+    if (new_word_state==word_state) {
+      nmisses=nmisses+1;
+    }
+    word_state = new_word_state
+  }
+
+  return(nmisses)
+
+}
+
+simulated_optimal_hangman_games <- words_and_categories %>%
+  rowwise() %>%
+  mutate(nmisses=simulate_optimal_hangman(as.character(word),category))
+
+simulate_random_hangman <- function(word,category) {
+  #initialize
+  word_state <- mask_word(toupper(word),''); 
+  asked <- ''
+  nclicks <- 0;
+  nmisses <- 0;
+
+  while (grepl('_',word_state) & nclicks<26) {
+
+    random_guess = sample(Filter(function(letter) !str_detect(asked, fixed(letter)), LETTERS),1)
+
+    #click on best shot
+    nclicks = nclicks+1;
+    asked = paste(asked,random_guess)
+    new_word_state = mask_word(word,asked)
+    if (new_word_state==word_state) {
+      nmisses=nmisses+1;
+    }
+    word_state = new_word_state
+  }
+
+  return(nmisses)
+
+}
+
+simulated_random_hangman_games <- data.frame()
+
+for (i in seq(500)) {
+  simulated_random_hangman_games <- rbind(simulated_random_hangman_games,
+                                          words_and_categories %>% 
+                                            sample_n(5) %>%
+                                            mutate(subj_id=i) %>%
+                                            rowwise() %>%
+                                            mutate(nmisses = simulate_random_hangman(word,category)))
+} 
 ```
@@
 
+E4.letter_choice_frequenc_wide <- E4.letter_choice_frequency %>%
+  dplyr::select(-count, -word_state) %>% 
+  spread(letter,p)
+
 E4.first_click_half_games <- E4.half_game_click_log %>%

### 2023-04-25 docs/half_games.rmd
multiple regression with prior

@@
 plot_posterior(posteriorB,'posteriorB');
+
+boardA <- "NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,1,1,0,0,NA,NA,0,NA,NA,1,1,NA,NA,NA"
+posteriorA <- boardA%>%get_likelihood();
+plot_posterior(posteriorA,'likelihoodA');
+
+boardB <- "NA,NA,NA,NA,0,NA,NA,0,NA,NA,NA,NA,NA,NA,0,NA,NA,1,0,NA,NA,NA,1,NA,1"
+posteriorB <- boardB%>%get_likelihood();
+plot_posterior(posteriorB,'likelihoodB');
 
@@
 
+E2.half_game_board_A <- c(12,13,14,15,18,21,22)
+E2.half_game_board_B <- c(5,8,15,18,19,23,25)
+
 E2.half_game_first_clicks <- E2.half_game_click_log %>%
@@
   group_by(grid_number,test_part) %>%
-  mutate(n=n/sum(n), x=i*5+j);
+  mutate(n=n/sum(n), x=i*5+j) %>%
+  mutate(n = ifelse(grid_number=='A' & (i*5+j+1) %in% E2.half_game_board_A, NaN, n),
+         n = ifelse(grid_number=='B' & (i*5+j+1) %in% E2.half_game_board_B, NaN, n));
 
@@
 ```
+
+```{r incorporating prior}
+
+E2.prior_all <- E2.click_log %>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+E2.prior_np <- E2.click_log %>%
+  filter(test_part=='nonpretend')%>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+E2.prior_first <- E2.click_log %>%
+  filter(click_number==1)%>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+E2.prior_first_np <- E2.click_log %>%
+  filter(click_number==1 & test_part=='nonpretend')%>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+E2.prior_first5 <- E2.click_log %>%
+  filter(click_number<6)%>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+E2.prior_first5np <- E2.click_log %>%
+  filter(click_number<6 & test_part=='nonpretend')%>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+plot_prior <- function(prior,file_name) {
+  prior <- scan(text= prior, what = numeric(), sep="," , quiet = TRUE);
+  board_df <- data.frame(i=rep(c(1,2,3,4,5),each=5),j=rep(c(1,2,3,4,5),5),prior=prior);
+  p <- ggplot(board_df,aes(x=j,y=i,fill=prior)) +
+    geom_tile() +
+    coord_equal() +
+    scale_x_continuous(breaks=c()) +
+    scale_y_continuous(breaks=c(),trans='reverse') +
+    scale_fill_gradient(low='#649bc8',high='#c83237', na.value ='white')+
+    labs(x='',y='') +
+    theme_classic();
+
+  ggsave(paste('figures/',file_name,'.png',sep=''),width=5,dpi=300);
+    ggsave(paste('figures/',file_name,'.svg',sep=''),width=5,dpi=300);
+
+  return(p)
+
+}
+
+ p<- plot_prior(paste(E2.prior_all$p,collapse=','),'priorall');
+ p<- plot_prior(paste(E2.prior_np$p,collapse=','),'priornp');
+ p<- plot_prior(paste(E2.prior_first$p,collapse=','),'priorfirst');
+ p<- plot_prior(paste(E2.prior_first_np$p,collapse=','),'priorfirstnp');
+ p<- plot_prior(paste(E2.prior_first5$p,collapse=','),'priorfirst5');
+ p<- plot_prior(paste(E2.prior_first5np$p,collapse=','),'priorfirst5np');
+
+```

### 2023-05-11 docs/ExploratoryResults.Rmd
small formatting tweaks

@@
 
-author: 
-  - name          : "Matan Mazor"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Malet St., London WC1E 7HX"
-    email         : "mtnmzor@gmail.com"
-  - name          : "Chaz Firestone"
-    affiliation   : "2"
-  - name          : "Ian Phillips"
-    affiliation   : "2"
-
-affiliation:
-  - id            : "1"
-    institution   : "Birkbeck, University of London"
-  - id            : "2"
-    institution   : "Johns Hopkins University"
+# author: 
+#   - name          : "Matan Mazor"
+#     affiliation   : "1"
+#     corresponding : yes    # Define only one corresponding author
+#     address       : "Malet St., London WC1E 7HX"
+#     email         : "mtnmzor@gmail.com"
+#   - name          : "Chaz Firestone"
+#     affiliation   : "2"
+#   - name          : "Ian Phillips"
+#     affiliation   : "2"
+# 
+# affiliation:
+#   - id            : "1"
+#     institution   : "Birkbeck, University of London"
+#   - id            : "2"
+#     institution   : "Johns Hopkins University"
 
@@
 ```
+## Heuristic use
 
-## Within-participant click entropy
+In Battleship, players sometimes have a fixed sequence of clicks that they follow, at least for the first moves.
+To quantify this, we plot the distribution of the first 5 clicks from all all games of a random subset of 100 players:
+
+```{r get-Battleship-prior, fig.cap="Distribution of the first 5 clicks in all games, from a subset of 100 random players"}
+
+set.seed(42)
+all_subjects <- E2.click_log$subj_id%>%unique();
+random100 <- all_subjects[sample(length(all_subjects),100)]
+
+E2.prior_first5 <- E2.click_log %>%
+  filter(click_number<6 & subj_id %in% random100)%>%
+  group_by(i,j)%>%
+  summarise(n=n()) %>%
+  mutate(p=n/sum(n),
+         flat_position = 1+i*5+j)
+
+ggplot(E2.prior_first5,aes(x=j,y=i,fill=p)) +
+  geom_tile() +
+  coord_equal() +
+  scale_x_continuous(breaks=c()) +
+  scale_y_continuous(breaks=c(),trans='reverse') +
+  scale_fill_gradient(low='#649bc8',high='#c83237', na.value ='white')+
+  labs(x='',y='') +
+  theme_classic() +
+  labs(title='click probability in Battleship')
+```
+
+As can be seen, subjects prefer to click on the diagonals, with a slight preference for the left hand side.
+We then asked how well aligned are players' clicks with this map in pretend and non-pretend games.
+For this analysis, we only used data from the remaining `r all_subjects%>%length()-random100%>%length()` players whose games were not used for generating the above map.
+
+```{r battleship-prior-comparison}
+
+battleship_prior <- E2.prior_first5$p;
+
+get_p_heuristic_rank <- function(posterior, flat_position) {
+  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
+  prior = battleship_prior*posterior/posterior #to get rid of nans
+  rank_prior = rank(-prior)
+  return(rank_prior[flat_position])
+}
+
+E2.click_log_with_heuristic <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_optimal) %>%
+  rbind(E2.click_log_with_boards_random) %>%
+  rowwise() %>%
+  mutate(p_heuristic_rank = get_p_heuristic_rank(posterior,flat_position))
+
+E2.mean_P_heuristic_rank <- E2.click_log_with_heuristic %>%
+  filter(!(subj_id %in% random100))%>%
+  group_by(subj_id,test_part) %>%
+  summarise(p_heuristic_rank=mean(p_heuristic_rank))
+
+E2.mean_P_heuristic_rank %>%
+    filter(test_part != 'optimal') %>%
+    ggplot(aes(x=p_heuristic_rank,fill=test_part))+
+    geom_histogram( color="black", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='mean rank according to the above map', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+```
+
+We find no difference in this measure between pretend and non-pretend games (`r E2.mean_P_heuristic_rank%>%spread(test_part,p_heuristic_rank)%>%mutate(diff=nonpretend-pretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+In the next analysis we use a more sensitive, player-specific measure of players' tendency to follow a fixed heuristic, versus playing flexibly.
+
+## Within-participant sequential guess entropy
 
@@
 E4.misses_per_word %>%
-  ggplot(aes(x=nonpretend,y=pretend, label=word)) +
+  ggplot(aes(x=nonpretend,y=pretend, label=toupper(word))) +
   geom_abline(intercept=0,slope=1) +
@@
   coord_fixed(ratio=1)+
-  scale_x_continuous(limits=c(0,5.5), breaks=seq(0,5))+
-  scale_y_continuous(limits=c(0,5.5), breaks=seq(0,5)) +
+  scale_x_continuous(limits=c(0,5), breaks=seq(0,5))+
+  scale_y_continuous(limits=c(0,5), breaks=seq(0,5)) +
   theme(legend.position = 'none')+
   theme_bw() +
-  labs(title='number of misses')
+  labs(x='median number of misses: non-pretend',
+       y='median number of misses: pretend')
+
+ggsave('figures/E4_scatter_misses.png',width=3.5, height=3.5)
 
@@
 
-## Within-participant click entropy
+## Within-participant sequential guess entropy
 

### 2023-05-11 docs/preregisteredMethodsAndResults.Rmd
small formatting tweaks

@@
 
-We used `r cite_r("r-references.bib")` for all our analyses.
+<!-- We used `r cite_r("r-references.bib")` for all our analyses. -->
 
@@
 ```
+
 ### Model based analysis
@@
 
-E2.mean_P_click_rank_misses_only <- E2.click_log_with_boards %>%
+E2.mean_P_click_rank_misses <- E2.click_log_with_boards %>%
   rbind(E2.click_log_with_boards_optimal) %>%
@@
 
-p<- E2.mean_P_click_rank_misses_only %>%
+p<- E2.mean_P_click_rank_misses %>%
     filter(test_part != 'optimal') %>%
@@
 
-ggsave('figures/E2_pclick_rank_misses_only.pdf',p)
-ggsave('figures/E2_pclick_rank_misses_only.png',p,width=5,height=2,dpi=300)
+ggsave('figures/E2_pclick_rank_misses.pdf',p)
+ggsave('figures/E2_pclick_rank_misses.png',p,width=5,height=2,dpi=300)
 
@@
 Pretend games were significantly less optimal than non-pretend games (`r E2.mean_P_click_rank%>%filter(test_part=='pretend')%>%pull(p_click_rank)%>%mean()%>%printnum()`; pretend-first group: `r E2.mean_P_click_rank%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`; pretend-second group: `r E2.mean_P_click_rank%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), but still more optimal than those of a random agent (`r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses_only%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Critically, the same pattern was observed when restricting analysis to cell selections that resulted in a miss (non-pretend - pretend: `r apa_print(t.test(E2.mean_P_click_rank_misses%>%filter(genuine_first&test_part=='nonpretend')%>%pull(p_click_rank), E2.mean_P_click_rank_misses%>%filter(!genuine_first&test_part=='pretend')%>%pull(p_click_rank)))$statistic`; pretend - random: `r E2.mean_P_click_rank_misses%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
 In other words, the optimality of pretend games relative to random cell selection was not merely due to the fact that pretenders clicked on ships more than expected by chance.
@@
 #          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
-#          test_part='most_frequent');
+#          test_part='optimal');
 
@@
 
-# save(E4.click_log_with_boards,
-#      E4.click_log_with_boards_random,
-#      E4.click_log_with_boards_most_frequent,
+# E4.letter_prob_click_log_with_boards_optimal <- E4.letter_prob_click_log_with_boards %>%
+#   dplyr::select(subj_id,genuine_first,test_part,word,category,word_state,click_number,posterior) %>%
+#   left_join(E4.click_log_with_boards_optimal %>%
+#               ungroup()%>%
+#               dplyr::select(subj_id,word,click_number,letter,hit_bin),
+#             by=c('subj_id','word','click_number')) %>%
+#   mutate(p_click = get_p_letter(posterior,letter),
+#          p_click_rank = get_p_letter_rank(posterior,letter),
+#          hit_bin = tolower(letter) %in% strsplit(tolower(word),'')[[1]],
+#          test_part='optimal');
+
+# save(E4.click_log_with_boards, E4.click_log_with_boards_random,
+# E4.click_log_with_boards_most_frequent, E4.click_log_with_boards_optimal,
 # E4.letter_prob_click_log_with_boards,
 # E4.letter_prob_click_log_with_boards_random,
+# E4.letter_prob_click_log_with_boards_optimal,
 # file='E4.click_log_with_boards.RData')
@@
   rbind(E4.click_log_with_boards_random)%>%
+  rbind(E4.click_log_with_boards_optimal)%>%
   group_by(subj_id,test_part, genuine_first) %>%
@@
 E4.mean_letter_prob_P_click_rank <- E4.letter_prob_click_log_with_boards %>%
-  rbind(E4.click_log_with_boards_random)%>%
+  dplyr::select(subj_id,genuine_first, test_part,word,p_click_rank)%>%
+  rbind(E4.letter_prob_click_log_with_boards_random%>%
+    dplyr::select(subj_id,genuine_first, test_part,word,p_click_rank))%>%
+  rbind(E4.letter_prob_click_log_with_boards_optimal %>%
+    dplyr::select(subj_id,genuine_first, test_part,word,p_click_rank))%>%
   group_by(subj_id,test_part, genuine_first) %>%
@@
     geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
-    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725", 'black')) +
     labs(x='p(letter in word) - rank', y='number of players')+
@@
     geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
-    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080",  "#FDE725", 'black')) +
     labs(x='p(letter) - rank', y='number of players')+
@@
   rbind(E4.click_log_with_boards_random %>%filter(!hit_bin))%>%
+  rbind(E4.click_log_with_boards_optimal %>%filter(!hit_bin))%>%
   group_by(subj_id,test_part, genuine_first) %>%
@@
     geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
     labs(x='p(letter in word) - rank', y='number of players')+
@@
     geom_histogram( alpha=0.6, position = 'identity', nbins=100) +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
     labs(x='p(letter) - rank', y='number of players')+
@@
            stat='identity', color='black')+
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
   geom_errorbar(aes(ymin=p_rank-se, ymax=p_rank+se), width=.2,position=position_dodge(.9)) +
@@
     geom_histogram( color="black", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
     labs(x='p(letter in word) - rank', y='number of players')+
@@
     geom_histogram( color="black", alpha=0.6, position = 'identity') +
-    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", 'black')) +
     labs(x='p(letter in English text) - rank', y='number of players')+
@@
   filter(click_number==1) %>%
-  dplyr::select(subj_id,test_part,word_state,letter) %>%
+  dplyr::select(subj_id,test_part,letter) %>%
   mutate(letter=factor(letter,levels=toupper(letters))) %>%
@@
 E4.letter_choice_frequenc_wide <- E4.letter_choice_frequency %>%
-  dplyr::select(-count, -word_state) %>% 
+  dplyr::select(-count) %>% 
   spread(letter,p)
@@
 
+p <- E4.letter_choice_frequency %>% 
+  filter(test_part %in% c('pretend','non-pretend') & word %in% c('papaya','banana')) %>%
+  mutate(letter = ifelse(letter %in% c('B','N','P','Y'), letter, 'other')) %>%
+  mutate(letter = factor(letter, levels=c('B', 'N', 'P', 'Y','other')))%>%
+  group_by(test_part,word,letter) %>%
+  summarise(p=sum(p)) %>%
+  mutate(test_part = factor(test_part, levels=c('non-pretend','pretend'))) %>%
+  mutate(word = factor(word, levels=c('banana','papaya'), labels=c('BANANA','PAPAYA')))%>%
+  ggplot(aes(x=reorder(letter, desc(letter)),y=p, fill=reorder(word, desc(word)))) +
+  geom_bar(stat='identity', position='dodge') +
+  facet_grid(cols=vars(test_part)) +
+  theme_classic() +
+  scale_fill_manual(values=c("#f16154","#fed601")) +
+  labs(x='',y='frequency') +
+  coord_flip() +
+  theme(legend.pos='na')
+
+ggsave('figures/Hangman2/bananapapaya.png',p,width=4,height=1.8,dpi=300)
+
+p <- E4.letter_choice_frequency %>% 
+  filter(test_part %in% c('pretend','non-pretend') & word %in% c('hair','hand')) %>%
+  mutate(letter = ifelse(letter %in% c('N','D','I','R'), letter, 'other')) %>%
+  mutate(letter = factor(letter, levels=c('N', 'D', 'I', 'R','other')))%>%
+  group_by(test_part,word,letter) %>%
+  summarise(p=sum(p)) %>%
+  mutate(test_part = factor(test_part, levels=c('non-pretend','pretend'))) %>%
+  mutate(word = factor(word, levels=c('hand','hair'), labels=c('HAND','HAIR')))%>%
+  ggplot(aes(x=reorder(letter, desc(letter)),y=p, fill=reorder(word, desc(word)))) +
+  geom_bar(stat='identity', position='dodge') +
+  facet_grid(cols=vars(test_part)) +
+  theme_classic() +
+  scale_fill_manual(values=c("#9a3300","#f1c27d")) +
+  labs(x='',y='frequency') +
+  coord_flip() +
+  theme(legend.pos='na')
+
+ggsave('figures/Hangman2/handhair.png',p,width=4,height=1.8,dpi=300)
+
 ```

### 2023-05-16 analysis/genderDifferences.Rmd
gender differences

---
title: "GenderDifferences"
author: "Matan Mazor"
date: "2023-05-16"
output: html_document
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel', #for word scatterplots
  'memoise' # for function memoization
), '2022-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')
```

## Load model coefficients

```{r create-gender-dfs, echo=FALSE}

E2.gender <- E2.export %>%
  filter(status=='APPROVED') %>%
  merge(E2.df %>%
          mutate(participant_id=PROLIFIC_PID) %>%
          dplyr::select(participant_number,participant_id,subj_id),by='participant_id') %>%
  dplyr::select(subj_id,Sex) %>%
  group_by(subj_id) %>%
  summarise(sex=Sex[1]) %>%
  rowwise()

E4.gender <- E4.export %>%
  filter(status=='APPROVED') %>%
  merge(E4.df %>%
          mutate(participant_id=PROLIFIC_PID) %>%
          dplyr::select(participant_number,participant_id,subj_id),by='participant_id') %>%
  dplyr::select(subj_id,Sex) %>%
  group_by(subj_id) %>%
  summarise(sex=Sex[1]) %>%
  rowwise()
```

```{r p_click, echo-FALSE, cache=TRUE}

E2.p_click_by_sex <- E2.p_click_summary %>%
  merge(E2.gender) %>%
  filter(sex %in% c('Female','Male'))

E2.p_click_by_sex_summary <- E2.p_click_by_sex %>% 
  group_by(test_part,sex) %>%
  summarise(p_click=mean(p_click_rank))

E4.p_click_by_sex <- E4.p_click_summary %>%
  merge(E4.gender) %>%
  filter(sex %in% c('Female','Male'))

E4.p_click_by_sex_summary <- E4.p_click_by_sex %>% 
  group_by(test_part,sex) %>%
  summarise(p_click=mean(p_click_rank))

apa_table(
  E2.p_click_by_sex_summary
  , caption = "p_click by gender in Battleship."
  , escape = TRUE
)

apa_table(
  E4.p_click_by_sex_summary
  , caption = "p_click by gender in Hangman."
  , escape = TRUE
)
```

We find no differences between the two sexes in the posterior probability of a hit, in neither pretend nor non-pretend games.

```{r hangman-halfgames, echo=F, cache=T}

E4.first_guess_by_sex <- E4.first_click_half_games %>%
  merge(E4.gender) %>%
  filter(sex %in% c('Female','Male'))

E4.papaya_by_sex <- E4.first_guess_by_sex %>%
  filter(word=='papaya') %>%
  group_by(test_part,sex) %>%
  summarise(banana=mean(bananahand & makes_sense),
            papaya = mean(!bananahand & makes_sense),
            other = mean(!makes_sense))

E4.hair_by_sex <- E4.first_guess_by_sex %>%
  filter(word=='hair') %>%
  group_by(test_part,sex) %>%
  summarise(hand=mean(bananahand & makes_sense),
            hair = mean(!bananahand & makes_sense),
            other = mean(!makes_sense))

apa_table(
  E4.papaya_by_sex
  , caption = "first click by gender in Hangman half games, when the target word is papaya. Showing proportion of guesses that are consistent with banana, papaya, or other."
  , escape = TRUE
)

apa_table(
  E4.hair_by_sex
  , caption = "first click by gender in Hangman half games, when the target word is hair. Showing proportion of guesses that are consistent with hand, hair, or other."
  , escape = TRUE
)

```

### 2023-10-18 experiments/demos/Battleships/jspsych-6.1.0/plugins/jspsych-battleships-replay-2players.js
demos

@@ -1,317 +1,0 @@
-jsPsych.plugins["Battleships_replay_2players"] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'Battleships_replay_2players',
-
-		parameters: {
-			grid: {
-				type: jsPsych.plugins.parameterType.INT,
-				pretty_name: "Grid",
-				default: [
-					['0','0','0','0','0'],
-					['A','A','A','0','0'],
-					['0','0','0','0','0'],
-					['0','0','0','B','0'],
-					['C','C','0','B','0']
-				],
-				description: "Grid. 0 means water, everything else is a ship."
-			},
-			end_screen_time: {
-				type: jsPsych.plugins.parameterType.INT,
-				pretty_name:'end screen time',
-				default: 2000,
-				description: 'For how many milliseconds is the end screen presented?'
-			},
-		 text: {
-			 type: jsPsych.plugins.parameterType.STRING,
-			 pretty_name:'text',
-			 default:
-`To win, you need to sink one 3-square submarine
-and two 2-square patrol boats.`,
-			 description: 'Text to display on top of grid.'
-		 },
-		 description: {
-			 type: jsPsych.plugins.parameterType.STRING,
-			 pretty_name:'description',
-			 default: {right: `Replaying player 2's game:`, left: `Replaying Player 1's game`},
-			 description: 'Description'
-		 },
-		 click_log: {
-			 type: jsPsych.plugins.parameterType.OBJECT,
-			 pretty_name:'click_log',
-			 default: {'right':{"i":[1,2,3,1,1,2,2,3,2,3,3,4,4,4],
-			 "j":[0,0,0,1,2,3,4,4,2,3,2,3,1,0],
-			 "t":[3559,3906,4284,5152,6111,7086,7696,7993,8532,9145,9580,10129,11173.5,11552],
-			 "hit":["A","0","0","A","A","0","0","0","0","B","0","B","C","C"]},
-			 "left":{"i":[1,2,1,2,1,0,1,3,3,4,4,4,4],
-			 "j":[3,3,2,2,1,2,0,2,3,3,2,1,0],
-			 "t":[1439,1932.5,4361,4998,5583,6664,7382,8805,9230,10242,11254.5,12390,13410],
-			 "hit":["0","0","A","0","A","0","A","0","B","B","0","C","C"]}},
-			 description: 'Click log'
-		 }
-		}
-	}
-
[…]

### 2023-10-18 experiments/demos/Battleships/jspsych-6.1.0/plugins/jspsych-battleships-replay.js

@@ -1,238 +1,0 @@
-jsPsych.plugins["Battleships_replay"] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'Battleships_replay',
-
-		parameters: {
-			grid: {
-				type: jsPsych.plugins.parameterType.INT,
-				pretty_name: "Grid",
-				default: [
-					['0','0','0','0','0'],
-					['A','A','A','0','0'],
-					['0','0','0','0','0'],
-					['0','0','0','B','0'],
-					['C','C','0','B','0']
-				],
-				description: "Grid. 0 means water, everything else is a ship."
-			},
-			cheat: {
-				type: jsPsych.plugins.parameterType.BOOL,
-				pretty_name:'cheat',
-				default: false,
-				description: 'When true, participants see the position of the ships.'
-			},
-			end_screen_time: {
-				type: jsPsych.plugins.parameterType.INT,
-				pretty_name:'end screen time',
-				default: 2000,
-				description: 'For how many milliseconds is the end screen presented?'
-			},
-		 text: {
-			 type: jsPsych.plugins.parameterType.STRING,
-			 pretty_name:'text',
-			 default:
-`To win, you need to sink one 3-square submarine
-and two 2-square patrol boats.`,
-			 description: 'Text to display on top of grid.'
-		 },
-		 description: {
-			 type: jsPsych.plugins.parameterType.STRING,
-			 pretty_name:'description',
-			 default:
- `Replaying Matan's game:`,
-			 description: 'Description'
-		 },
-		 click_log: {
-			 type: jsPsych.plugins.parameterType.OBJECT,
-			 pretty_name:'click_log',
-			 default: {"i":[1,2,3,1,1,2,2,3,2,3,3,4,4,4],
-			 "j":[0,0,0,1,2,3,4,4,2,3,2,3,1,0],
-			 "t":[3559,3906,4284,5152,6111,7086,7696,7993,8532,9145,9580,10129,11173.5,11552],
-			 "hit":["A","0","0","A","A","0","0","0","0","B","0","B","C","C"]},
-			 description: 'Click log'
-		 }
-		}
-	}
-
-	plugin.trial = function(display_element, trial) {
-
-		display_element.innerHTML = '';
-
-		//open a p5 sketch
-		let sketch = function(p) {
-
-		const du = p.min([window.innerWidth, window.innerHeight, 500])*7/10 //drawing unit
[…]

### 2023-10-18 experiments/demos/Battleships/jspsych-6.1.0/plugins/jspsych-battleships.js

@@ -1,225 +1,0 @@
-jsPsych.plugins["Battleships"] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'Battleships',
-
-		parameters: {
-			grid: {
-				type: jsPsych.plugins.parameterType.INT,
-				pretty_name: "Grid",
-				default: [
-					['0','0','0','0','0'],
-					['A','A','A','0','0'],
-					['0','0','0','0','0'],
-					['0','0','0','B','0'],
-					['C','C','0','B','0']
-				],
-				description: "Grid. 0 means water, everything else is a ship."
-			},
-			cheat: {
-				type: jsPsych.plugins.parameterType.BOOL,
-				pretty_name:'cheat',
-				default: false,
-				description: 'When true, participants see the position of the ships.'
-			},
-			end_screen_time: {
-				type: jsPsych.plugins.parameterType.INT,
-				pretty_name:'end screen time',
-				default: 2000,
-				description: 'For how many milliseconds is the end screen presented?'
-			},
-		 text: {
-			 type: jsPsych.plugins.parameterType.STRING,
-			 pretty_name:'text',
-			 default:
-`To win, you need to sink one 3-square submarine
-and two 2-square patrol boats.`,
-			 description: 'Text to display on top of grid.'
-		 }
-		}
-	}
-
-	plugin.trial = function(display_element, trial) {
-
-		display_element.innerHTML = '';
-
-		//open a p5 sketch
-		let sketch = function(p) {
-
-		const du = p.min([window.innerWidth, window.innerHeight, 500])*7/10 //drawing unit
-		const left_margin = p.round((window.innerWidth-du)/2);
-		const top_margin = p.round((window.innerHeight-du)/3);
-		const square_size = Math.floor(du/trial.grid.length);
-		const colors = {
-			'sea':p.color(100,155,200),
-			'unknown':p.color(255,230,166),
-			'ship':p.color(200,50,55)
-		}
-		var grid_state = trial.grid.map(([...rest]) => rest.map(x => 'unknown'));
-		var click_log = {i:[],j:[],t:[],hit:[]};
-		const num_nonzero = trial.grid.flat().reduce((a,b)=>a+(b=='0'? 0 : 1),0);
-		var hits = 0;;
-		var last_click_time = p.millis()
-
-
-		function grid_coordinates_to_screen_coordinates(i,j) {
[…]

### 2023-10-18 experiments/demos/Battleships/jspsych-6.1.0/plugins/jspsych-disc.js

@@ -1,338 +1,0 @@
-jsPsych.plugins["disc"] = (function() {
-
-var plugin = {};
-
-plugin.info = {
-  name: 'disc',
-  parameters: {
-    stimulus1_function: {
-        type: jsPsych.plugins.parameterType.FUNCTION,
-        pretty_name: "stimulus_function",
-        default: (p,t,du) => {p.noFill(); p.strokeWeight(2); p.stroke(255); p.rect(0,0,du/10,du/10)}
-    },
-    stimulus2_function: {
-        type: jsPsych.plugins.parameterType.FUNCTION,
-        pretty_name: "stimulus_function",
-        default: (p,t,du) => {p.noFill(); p.strokeWeight(2); p.stroke(255); p.circle(0,0,du/10)}
-    },
-    mask_function: {
-        type: jsPsych.plugins.parameterType.FUNCTION,
-        pretty_name: "mask_function",
-        default: (p,t,du) => {p.push(); p.fill(255); p.rect(0,0,du/5,du/5); p.pop()}
-    },
-    stim_duration: {
-      type: jsPsych.plugins.parameterType.FLOAT,
-      pretty_name: "stim duration",
-      default: 30,
-      description: "Duration of stimulus mask in milliseconds"
-    },
-    SOA: {
-      type: jsPsych.plugins.parameterType.FLOAT,
-      pretty_name: "SOA",
-      default: 30,
-      description: "Duration of prestimulus mask in milliseconds"
-    },
-    mask_duration: {
-      type: jsPsych.plugins.parameterType.FLOAT,
-      pretty_name: "mask duration",
-      default: 100,
-      description: "Duration of mask in milliseconds"
-    },
-    mask_2_duration: {
-      type: jsPsych.plugins.parameterType.FLOAT,
-      pretty_name: "mask2 duration",
-      default: 100,
-      description: "Duration of poststimulus mask in milliseconds"
-    },
-    which_stimulus: {
-      type: jsPsych.plugins.parameterType.INT,
-      pretty_name: "Which stimulus (1 or 2)",
-      default: 1
-    },
-    fixation_duration: {
-      type: jsPsych.plugins.parameterType.FLOAT,
-      pretty_name: "Fixation duration",
-      default: 500,
-      description: "Duration of fixation in milliseconds"
-    },
-    choices: {
-      type: jsPsych.plugins.parameterType.STRING,
-      pretty_name: "Choices",
-      description: "Choice keys",
-      default: ['f','j']
-    },
[…]

### 2023-10-18 experiments/demos/Battleships/jspsych-6.1.0/plugins/jspsych-p5text.js

@@ -1,106 +1,0 @@
-jsPsych.plugins["p5Text"] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'p5Text',
-		parameters: {
-			title: {
-				type: jsPsych.plugins.parameterType.STRING,
-				default: '',
-			},
-			text: {
-				type: jsPsych.plugins.parameterType.STRING,
-				default: 'text',
-			},
-			footer: {
-				type: jsPsych.plugins.parameterType.STRING,
-				default: 'press Space to continue',
-			},
-			draw_function: {
-				type: jsPsych.plugins.parameterType.FUNCTION,
-				default: function(p, trial) {return}
-			},
-			key_function: {
-				type: jsPsych.plugins.parameterType.FUNCTION,
-				default: function(p, trial) {return}
-			},
-			tite: {
-				type: jsPsych.plugins.parameterType.STRING,
-				default: ''
-			}
-		}
-	}
-
-	plugin.trial = function(display_element, trial) {
-
-		display_element.innerHTML = ''
-
-		//open a p5 sketch
-		let sketch = function(p) {
-
-		//sketch setup
-		p.setup = function() {
-			p.createCanvas(window.innerWidth, window.innerHeight);
-			p.fill(255); //white
-			p.strokeWeight(0)
-			p.background(128); //gray
-			p.textFont('Quicksand');
-      p.rectMode(p.CENTER)
-		}
-
-		//organize everything in one sequence
-		p.draw = function() {
-
-      p.textAlign(p.LEFT,p.CENTER)
-			p.background(128); //gray
-			p.textSize(24)
-			p.text(trial.text, p.width/2, p.height*(5/12), 600, 300);
-
-			p.push()
-			p.textAlign(p.CENTER, p.CENTER)
-			// p.textStyle(p.BOLD)
-			p.textSize(50)
-			p.text(trial.title,p.width/2, p.height/6, 600, 100)
-			// p.pop()
-			//
-			// p.push()
-			p.textSize(18)
-			p.textStyle(p.NORMAL)
-			p.text(trial.footer, p.width/2, p.height*(5/6), 600, 60);
-			p.strokeWeight(1)
-			p.stroke(255)
-			p.line(p.width/2-150,p.height*(3/4),p.width/2+150,p.height*(3/4))
-			p.pop()
-
-			trial.draw_function(p, trial)
-		}
-
-		p.keyPressed = function() {
-
[…]

### 2023-10-18 experiments/demos/Battleships/jspsych-6.1.0/plugins/jspsych-p5vs_yn.js

@@ -1,182 +1,0 @@
-jsPsych.plugins["p5vs_yn"] = (function() {
-
-	var plugin = {};
-
-	plugin.info = {
-		name: 'p5vs_yn',
-		parameters: {
-			title: {
-				type: jsPsych.plugins.parameterType.STRING,
-				default: '',
-			},
-      target_function: {
-        type: jsPsych.plugins.parameterType.FUNCTION,
-        default: (p,t,du) => {p.rect(0,0,p.sin(t/1000)*du/10,p.sin(t/1000)*du/10)}
-      },
-      distractor_function: {
-        type: jsPsych.plugins.parameterType.FUNCTION,
-        default: (p,t,du,dist_num) => {p.circle(0,0,p.sin(t/1000+dist_num/5)*du/10)}
-      },
-      set_size: {
-        type: jsPsych.plugins.parameterType.NUMBER,
-        default: 10
-      },
-      target_present: {
-        type: jsPsych.plugins.parameterType.BOOL,
-        default: false
-      },
-      instructions_duration: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        default: 1000 //ms
-      },
-      fixation_duration: {
-        type: jsPsych.plugins.parameterType.FLOAT,
-        default: 500 //ms
-      },
-      choices: {
-				type: jsPsych.plugins.parameterType.STRING,
-				pretty_name: "Choices",
-				description: "Choice keys",
-        default: ['f','j']
-			}
-		}
-	}
-
-	plugin.trial = function(display_element, trial) {
-
-		display_element.innerHTML = ''
-
-		//open a p5 sketch
-		let sketch = function(p) {
-
-    const du = p.min([window.innerWidth, window.innerHeight, 600])*7/10 //drawing unit
-    const grid_size =  p.ceil(p.sqrt(trial.set_size+1))+3
-    const ticks = Array.from(new Array(grid_size), (x,i) => i/grid_size-0.5).slice(1)
-    var positions = []
-    for (var i = 0; i < ticks.length; i++) {
-      // This is where you'll capture that last value
-      for (var j = 0; j < ticks.length; j++) {
-        positions.push([ticks[i],ticks[j]]);
-      }
-    }
-    p.shuffle(positions, true)
-    trial.target_position = positions[0]
-    trial.distractor_positions = positions.slice(1,trial.set_size+1)
-
-		trial.jitter = p.random(500)
-		trial.fixation_duration+=trial.jitter
-
-
-		//sketch setup
-		p.setup = function() {
[…]

### 2023-10-18 docs/paper.Rmd
full paper

--- 2023-03-14 docs/paper.Rmd
@@
 ---
-title             : "Pretending not to know"
-shorttitle        : "Pretending not to know"
+title             : "Pretense reveals a capacity for online simulation of counterfactual belief states"
+shorttitle        : "Pretense reveals a capacity for online simulation of counterfactual belief states"
 
@@
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
+  What we do depends on what we know.
+But sometimes we try to decouple our behavior from our knowledge so as to appear not to know what we really do know.
+Such pretense behavior requires understanding how we would behave with different knowledge — and so provides a window into counterfactual self-simulation.
+However, little research has characterized and evaluated pretense relative to non-pretense behavior.
+Here, in two large-scale, pre-registered experiments, we tested people’s ability to conceal knowledge in the context of information-seeking games (Battleship and Hangman).
+Participants played normal games, as well as ‘pretend’ games in which they had full access to the game state, but pretended they didn’t.
+Overall, we find a highly impressive capacity for pretense — so impressive that independent 'judge' participants were completely unable to discriminate between normal and ‘pretend’ games.
+Specifically, pretenders successfully reproduced subtle features of real players' decision-making and timing.
+Nonetheless, pretense was characterized by pervasive differences (undetected by judges but revealed by follow-up analyses): pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules.
+Our findings thus reveal a powerful but imperfect ability to simulate decisions and behaviour based on a counterfactual belief state, and open up a new approach to studying self-simulation.
 
@@
 wordcount         : "X"
-
-bibliography      : "r-references.bib"
 
@@
 output            : papaja::apa6_word
+bibliography: references.bib
 ---
 
-```{r setup, include = FALSE}
-library("papaja")
-r_refs("r-references.bib")
-```
-
-```{r analysis-preferences}
+```{r analysis-preferences, warning=F, message=F}
 # Seed for random number generation
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning=F, message=F)
 
@@
 load('../.RData')
+
+r_refs("r-references.bib")
+
+#color scheme
+colors <-  list()
+colors$p <- "#404080"
+colors$np <- "#69b3a2"
+colors$random <- 'gray'
+colors$greedy <- 'black'
 ```
@@
 
+<!-- The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory].
+This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation].
+This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own.
+-->
+
+Pretense relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state.
+For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
+What would I say?
+How long would it take me to recover from the surprise?").
+This is not a trivial challenge: previous research demonstrates that access to knowledge distorts our beliefs about what we would have thought [@fischhoff1975] or perceived [@harley2004] without that knowledge.
+Indeed, based on previous research, it is possible that people are completely unable to mentally simulate their own behaviour under a counterfactual state of ignorance.
+
+<!-- to what extent humans are capable of generating complex behaviour that is aligned with a counterfactual belief state, and specifically with a counterfactual state of ignorance. -->
+
+To this end, we examine pretense in a game setting.
+Using an online version of the games Battleships and Hangman, participants played a 'non-pretend' (normal) version of the game, as well as a 'pretend' version where they were given secret information about the location of hidden ships/target word, but were instructed to behave as if they didn't have this information.
+Overall, we find a highly impressive capacity for pretense, with pretenders mirroring subtle features of real players' decision-making and timing.
+At the same time, pretense was characterized by pervasive differences, with pretenders consistently playing less rationally than normal players, sometimes rigidly following heuristics and rules.
+Whilst powerful, then, our capacity for self-simulation is imperfect, displaying tell-tale biases in our internal models of our own cognition and decision making.
+
 # Results
 
-In two experiments, players played online versions of two information-seeking games: Battleship and Hangman.
-In these games, players' goal is to reveal an unknown world state (grid configuration in Battleship, a hidden word in Hangman) with as few steps (cell or letter selections) as possible.
-Critically, in addition to playing five games in their standard version, players also completed five pretend games in which the true world state was known to them all along, but their task was to behave as if it was not (see Fig.
+In two experiments, players played online versions of two information-seeking games: Battleship (N=`r E2.df$subj_id%>%unique()%>%length()`) and Hangman (N=`r E4.df$subj_id%>%unique()%>%length()`).
+These games typically start in a state of ignorance, where players' goal is to reveal an unknown world state (grid configuration in Battleship, a hidden word in Hangman) with as few steps (cell or letter selections) as possible.
+Critically, in addition to playing five games in their standard version, players also completed five pretend games in which the true world state was known to them from the start, but their task was to behave as if it was not (see Fig.
 \@ref(fig:design)).
-We measured subjects' capacity to simulate a counterfactual knowledge state by comparing patterns of decisions and decision times in pretend and non-pretend games.
+We measured their capacity to simulate a counterfactual knowledge state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games.
 Our full pre-registered results are available [online](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html).
 
+```{r design, echo=FALSE, fig.cap="Experimental Design in Exp.
+1 (left) and 2 (right).
+In non-pretend games, players revealed ships by guessing cells in a grid (A) or revealed a word by guessing letters (E).
+In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (F), but asked players to play as if they didn't have access to this information.
+In half games (C and G), players continued the game from a given state, with the content of some cells and the position of some letters already revealed.
+Lastly, players watched replays of the games of previous players and guessed which were pretend games (D and H).", out.width = '100%'}
+knitr::include_graphics("figures/methods.png")
+```
+
 As a first gross measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games.
-Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()`), with no difference between the two conditions (`r t.test(E2.median_clicks%>%filter(genuine_first)%>%pull(nonpretend)-7,E2.median_clicks%>%filter(!genuine_first)%>%pull(pretend)-7)%>%apa_print()%>%'$'(statistic)`).
-In Hangman, pretenders tended to make slightly more letter guesses on average, controlling for word length (mean = `r E4.median_misses%>%pull(pretend)%>%mean()`) than did non-pretenders (mean = `r E4.median_misses%>%pull(nonpretend)%>%mean()`; `r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(pretend),E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend))%>%apa_print()%>%'$'(statistic)`).
+Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()`; $p=$ `r apa_print(E2.median_clicks$diff%>%t.test())$table$p.value`; Fig.
+\@ref(fig:decisions)A).
+In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()`; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()`; $p$ `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+\@ref(fig:decisions)B).
+Despite an overall bias in the number of guesses, pretend Hangman games provided evidence for a more impressive, item-specific alignment: pretenders were successful in making more letter guesses when attempting to reveal words that would have been harder to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`; Fig.
+\@ref(fig:decisions)B).
+This correlation provides a first clue for a human capacity to act in accordance with a counterfactual knowledge state.
 
-Furthermore, the ten target words we used in Hangman differed in difficulty: some required more letter guesses than others, controlling for word length.
-Strikingly, pretenders were highly successful in making more letter guesses when attempting to reveal those words that would have been easier to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(full_result)`).
-This correlation provides a first clue of an impressive capacity to act in accordance with how one would have acted if they didn't know what they know.
+(ref:decisions-caption) Battleship and Hangman guesses in pretend and non-pretend games.
+A: median number of misses Battleship and Hangman games, in non-pretend (green) and pretend (purple) games.
+For reference, the expected number of misses is indicated by a reference line for a fully random agent, and for a "greedy" agent that maximizes the probability of a hit in each step.
+B: The median number of misses in Hangman for pretend and non-pretend games, as a function of the target word.
+C: in half-games, players continued the game from a half-finished state.
+Showing spatial guess distributions for pretend and non-pretend games, alongside their corresponding hit probability maps.
+D: Cell and letter selections were ranked according to their relative hit probability.
+Plotting the median rank per subject in pretend and non-pretend games, with reference lines for the expected rank probability for a random agent, and for a "greedy" agent that maximizes the probability of a hit in each step.
+E: same as panel D but discarding all guesses that resulted in a hit.
 
-Having established that pretenders were able to mimic the total number of guesses in standard games, we next turned to the content of pretend and non-pretend guesses.
+```{r decisions, echo=FALSE, message= F, warning=F, fig.cap="(ref:decisions-caption)", out.width = '100%'}
+
+num_misses_df <- E2.click_df %>%
+  dplyr::select(subj_id,test_part, num_clicks) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_clicks)-7) %>%
+  mutate(game='Battleship') %>%
+  rbind(
+    E4.click_df %>%
+  dplyr::select(subj_id,test_part, num_misses) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_misses)) %>%
+    mutate(game='Hangman')
+  )
+load('../analysis/simulated_hangman_games.RData')
+
+E2.random_mean_misses <- E2.random_click_df %>%
+           group_by(subj_id) %>%
+           summarise(num_misses=median(num_clicks)-7) %>%
+           pull(num_misses) %>%
+                  mean()
+
+E2.greedy_mean_misses <- E2.optimal_click_df %>%
+           group_by(subj_id) %>%
+           summarise(num_misses=median(num_clicks)-7) %>%
+           pull(num_misses) %>%
+                  mean()
+
+E4.random_mean_misses <- simulated_random_hangman_games %>% 
+  group_by(subj_id) %>%
+  summarise(num_misses=median(nmisses)) %>%
+  pull(num_misses) %>%
+  mean()
+
+E4.greedy_mean_misses <- simulated_optimal_hangman_games %>% 
+  pull(nmisses) %>%
+  mean()
+
+ref_lines <- data.frame(
+  game = c('Battleship','Battleship','Hangman','Hangman'),
+  xintercept = c(E2.random_mean_misses, E2.greedy_mean_misses, 
+                 E4.random_mean_misses, E4.greedy_mean_misses),
+  label = c('Random','Greedy','Random','Greedy'),
+  y = c(75,75,75,115)
+)
+
+p <- num_misses_df %>%
+  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
+  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
+  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position='none') +
+  theme_classic() +
+  geom_vline(data=ref_lines, aes(xintercept = xintercept), color='black', linetype=2) +
+  geom_text(data=ref_lines, aes(x = xintercept, y=y, label=label), angle = 90, vjust = -1, size = 3, inherit.aes = FALSE)+
+  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
+  facet_wrap(~game, nrow=2) +
+  theme(legend.position='na',strip.background = element_blank(), panel.border=element_blank())
+
+ggsave('figures/num_misses_facet.png',p, width=4, height=3)
+
+p<-  E2.click_df %>%
+  dplyr::select(subj_id,test_part, num_clicks) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_clicks)-7) %>%
+  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
+  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
+  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position='none') +
+  theme_classic() +
+  geom_vline(xintercept = E2.greedy_mean_misses, color='black', linetype=2) +
+  annotate("text", x = E2.greedy_mean_misses, y = 75, label = "Greedy", angle = 90, vjust = -1, size = 5)+
+  geom_vline(xintercept = E2.random_mean_misses, color='black', linetype=2) +
+  annotate("text", x = E2.random_mean_misses, y = 75, label = "Random", angle = 90, vjust = -1, size = 5)+
+  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
+  theme(legend.pos='na')
+
+ggsave('figures/num_misses_E2.png',p, width=4.5, height=2.5)
+
+p <- E4.click_df %>%
+  dplyr::select(subj_id,test_part, num_misses) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(num_misses=median(num_misses)) %>%
+  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
+  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
+  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position='none') +
+  theme_classic() +
+  geom_vline(xintercept = E4.greedy_mean_misses, color='black', linetype=2) +
+  annotate("text", x = E4.greedy_mean_misses, y = 95, label = "Greedy", angle = 90, vjust = -1, size = 5)+
+  geom_vline(xintercept = E4.random_mean_misses, color='black', linetype=2) +
+  annotate("text", x = E4.random_mean_misses, y = 95, label = "Random", angle = 90, vjust = -1, size = 5)+
+  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
+  theme(legend.pos='na')
+
+ggsave('figures/num_misses_E4.png', p, width=4.5, height=2.5)
+knitr::include_graphics("figures/decisions.png")
+
+```
+
+Having established an alignment in the total number of guesses, we next turned to the content of pretend and non-pretend guesses.
 In order to directly compare pretend and non-pretend guesses for the same board state, Battleship players completed two half-games in which they were instructed to continue the game from a half-completed state.
-Hundreds of cell selections for the same board state revealed a strong correlation between the spatial distributions of pretend and non-pretend guesses (), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.
+Hundreds of cell selections for the same board state revealed a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_pretend_nonpretend)$table$estimate`, $p=$ `r apa_print(cor_A_pretend_nonpretend)$table$p.value`; board B: $r=$ `r apa_print(cor_B_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_B_pretend_nonpretend)$table$p.value`; Fig.
+\@ref(fig:decisions)C), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.
 
 To further examine the decisional processes behind this strong alignment, we compared the degree to which pretend and non-pretend guesses made sense within the context of the game.
-We approximated guess rationality by calculating the probability that a guess will be successful (in revealing a ship or word letters) given available information, $p(hit(x_{i}))$, and the posterior probability that one should make a guess, assuming a uniform prior over cells or letters $P(x_{i})=\frac{p(hit(x_i))}{\sum_{j=1}^{k}p(hit(x_j))}$.
-Given this posterior map, it makes sense to guess cells or letters for which $P(x_{i})$ is high [this behavior is not strictly optimal, but approximates optimal behavior in most cases; @audinot2014optimal, Section 3.3].
-In the non-pretend versions of both games, guesses were more rational than expected by chance (Battleship:, Hangman:).
-Despite being less rational than non-pretend guesses, pretend guesses were more rational than expected by chance (Battleship:, Hangman:).
-Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship:, Hangman:): even when missing ships or letters, pretend guesses made sense given the limited information players pretended to have.
+When playing Battleship and Hangman, it makes sense to guess cells or letters for which the probability of hitting a ship or revealing a letter is high [this "greedy" behavior is not strictly optimal, but approximates optimal behavior in most cases, @audinot2014optimal, Section 3.3].
+In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank_wide%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; Fig.
+\@ref(fig:decisions)D).
+Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; Fig.
+\@ref(fig:decisions)E): that is, even when missing ships or letters, pretend guesses made sense given the limited information players pretended to have.
+
+Good pretense is a function not only of the number and content of players' decisions, but also of their timing.
+Here too, pretend games showed the same qualitative patterns as non-pretend games.
+Like non-pretenders, pretenders were faster in their successful guesses (Difference in decision time between hits and misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms) and slowed down immediately after a hit (Difference in decision time between guesses that followed hits versus misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms; Fig.
+\@ref(fig:time)A).
+Furthermore, decision uncertainty, quantified as the Shannon entropy of the posterior distribution over cell or letter options, had a similar quadratic effect on decision time in both non-pretend and pretend versions of the games, with the slowest responses associated with mid-range levels of entropy (see Fig.
+\@ref(fig:time)B).
+
+```{r time, echo=FALSE, message= F, warning=F, fig.cap="Patterns of reaction time in pretend and non-pretend guesses.
+A: median decision times in and following hits and misses.
+In both Battleship and Hangman, hits were faster on average than misses, but guesses following a hit were slower on average than those following a miss.
+This pattern was mimicked in pretend games.
+B: median decision times as a function of decision uncertainty, quantified as the entropy of the posterior over guess options.
+In both Hangman and Battleship, guesses were slowest for mid-range levels of entropy, and this pattern was mimicked in pretend games.", out.width = '100%'}
+knitr::include_graphics("figures/time.png")
+```
+
+This capacity for simulating a state of ignorance was not perfect.
+First, despite showing the same qualitative effects, decision times in Battleship (but not Hangman) pretend games were systematically more pronounced relative to non-pretend games, as if participants were acting based on an exaggerated mental image of their own counterfactual behaviour.
+Furthermore, in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result.
+The within-participant sequential guess entropy, a measure of the inverse predictability of decision sequences across different games of the same player, was significantly reduced in pretend games (Fig.
+\@ref(fig:rigidity)A; see [online supplementary materials](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#within-participant-click-entropy)).
+More specifically, Hangman pretenders were more likely to guess letters that are frequent in the English language (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig.
+\@ref(fig:rigidity)B).
+
+This limitation on incorporating evidence into the decision-making process was especially evident in Hangman half-games, where players completed the game from a half-completed state.
+When asked to reveal the hidden fruit "\_A_A_A", `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('banana', 'papaya'))%>%pull(bananahand)%>%mean())`% of the non-pretenders guessed one of the letters 'B' or 'N' (Fig.
+\@ref(fig:rigidity)C, yellow bars in left column).
+Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word== 'banana')%>%pull(bananahand)%>%mean())`%.
+Importantly, half of the pretenders were given different information: they were told that the hidden word is the less prototypical fruit PAPAYA.
+Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', revealing that many pretenders were unable to predict that the first fruit that would come to their minds is BANANA, not PAPAYA (Fig.
+\@ref(fig:rigidity)C, yellow bars in right column).
+A similar pattern was observed for the prototypical body part word HA(ND) and its surprising counterpart HA(IR): when playing normally, `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('hand', 'hair'))%>%pull(bananahand)%>%mean())`% of the players selected letters that are consistent with the prototypical option HAND.
+Pretenders for whom the word was HAND showed the same tendency (`r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hand')%>%pull(bananahand)%>%mean())`%), in contrast to only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hair')%>%pull(bananahand)%>%mean())`% of the pretenders who saw the word HAIR (Fig.
+\@ref(fig:rigidity)C, blue bars).
+
+```{r rigidity, echo=FALSE, message= F, warning=F, fig.cap="Limitations on flexible decision-making when pretending.
+A: sequential guess entropy, a measure of the (inverse) predictability of individual players' guesses as a function of click number and guess number.
+In both Battleship and Hangman, Sequential guess entropy increased with click number, and was overall lower in pretend games.
+B: letter frequency of Hangman guesses: the median rank frequency per participant in pretend and non-pretend games, with reference lines for the expected rank frequency for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step.
+C: first letter guesses in Hangman half games, as a function of pretense condition and target word.
+Letters appear in alphabetical order.
+letters that appear in the target word are marked in yellow (BANANA), red (PAPAYA), blue (HAND) and magenta (HAIR).
+For reference, the overall distribution of letter guesses across all games and conditions is given below.
+", out.width = '100%'}
+knitr::include_graphics("figures/rigidity.png")
+```
+
+Despite these limitations, pretenders were able to fool other players into believing they were playing for real.
+After performing pretend and non-pretend games, participants were presented with game replays of previous players and judged who was pretending and who was playing for real.
+In both games, pretense detection was at chance (Battleship: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%, Hangman: `r round(E4.judge_summary_df$accuracy%>%mean()*100)`%).
+This is in line with previous findings of near-chance accuracy in lie detection [@bond2006].
+Moreover, we find no sign of a correlation between pretense quality (measured as players' ability to trick judges to think they were not pretending) and pretense detection ability (measured as proportion correct; Battleship: `r apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate, method='spearman'))$estimate`, Hangman: `r apa_print(cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman'))$estimate`), indicating that pretense and pretense detection rely on at least partly different cognitive processes.
+
+# Discussion
+
+In two experiments, we examined participants' ability to mimic a state of ignorance in a game setting.
+We find that pretenders were able to successfully emulate decisions taken under a true state of ignorance, including reproducing subtle effects of guess accuracy and decision uncertainty on decision time.
+We also identify reliable signatures of pretend-ignorance on players' decisions, including a cost to decision rationality and an increased tendency to follow heuristics and rules.
+Together, our findings reveal an impressive but limited capacity for an online simulation of counterfactual belief states.
+
+Previous research has identified limitations on the human capacity to encapsulate knowledge from affecting our decisions and behaviour [@fischhoff1975; @harley2004].
+Our findings reveal that notwithstanding these limitations, humans are capable of approximating their hypothetical behaviour had they not known what they do in fact know.
+This capacity is not limited to making similar decisions to the ones we would have made had we not known; we also find that pretend decision times reproduce qualitative patterns observed under a true state of ignorance.
+This suggests that our participants were not engaging in a superficial mimicry of counterfactual behavioural patterns, but instead simulated, in real time, the cognitive processes that give rise to these patterns.
+
+An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored the revealed game state such that their entire cognitive machinery was available to play the game.
+This does not require any self-simulation, only a capacity to intentionally unsee, or forget, relevant evidence.
+While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense for three reasons.
+First, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical].
+Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with self-simulation or rule-following (see [exploratory analysis]([https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).) Finally, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig.
+\@ref(fig:time)A): a finding that is consistent with the simulation of a simplified self-model, but is hard to explain if participants were merely suppressing their current knowledge.
+
+Research on Bayesian Theory of Mind provides some support for such internal simulations of decision-making processes.
+This is often studied by measuring participants' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational], or implicitly [@onishi2005fifteen; @liu2017ten].
+Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state.
+In this case, a counterfactual knowledge state in which a known piece of information is unknown.
+Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
+Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
+
+While previous work focused on people's inability to perfectly predict their behaviour had they not known something ["hindsight bias", @fischhoff1975, @roese2012], our paradigm and analysis revealed a more nuanced picture, with an impressive but limited capacity to simulate a counterfactual state of ignorance.
+Alongside pretenders' many successes, we identify specific limitations, including a tendency to follow heuristics and rules and a scaling of response time profiles.
+These limitations are useful in showing that pretending not to know relies on more than the simulation of the same computational mechanisms that underlie genuine decision-making [@saxe2005].
+More work is now needed to trace the boundaries of the capacity to pretend not to know, and how these boundaries map onto limitations in the human capacity to simulate counterfactual mental states.
+
+<!-- including patterns of active information seeking and knowledge update in the face of new evidence. -->
+
+<!-- We consider two possible explanations to  -->
+
+<!-- Our subjects rose to the challenge.
+Their pretense behavior, although not perfect, closely resembled that of non-pretenders \-\-- not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
+Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: "What would I have done if I didn't have this piece of knowledge?".
+-->
+
+<!-- An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game.
+This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something.
+While we cannot fully rule out this interpretation, we think it is unlikely to explain participants' pretense ability in our data, for two indirect reasons.
+First, suppressing thoughts on demand is notoriously difficult [\@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible.
+Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).
+-->
+
+<!-- In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects' models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
+While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.
+-->
 
 # Methods
+
 We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
@@
 
-## Participants
+## Exp. 1: Battleship
 
-## Material
+A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb).
+The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019] (protocol sum: 60c270410375e8a192468fc1a0e9c93da60d5e203eb2760b621a8631a26f4c5c; [link to relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/44ef964c0d7d4a8f3b76ec52316a8e624c41a047/experiments/Battleships2/code/index.html#L865-L876)).
 
-## Procedure
+### Participants
 
-## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
+The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
+500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
+They were selected based on their acceptance rate (\>95%) and for being native English speakers.
+The entire experiment took approximately 20 minutes to complete.
+Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
 
-# Results
+### Procedure
 
-# Discussion
+Participants were first instructed that the experiment, based on the game Battleships, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment.
+They were then presented with a leaderboard of previous players, and given the rules of the game:
+
+> "In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
+What makes the game difficult is that you can't see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
+To sink a ship, you need to click on all of the squares it is located in.
+If you hit part of a ship, the square will turn red.
+If there is no ship in the square, it will turn blue."
+
+We further explained that in this version of the game, ships can touch corners, but their sides can't touch.
+This explanation was accompanied by a visual presentation of legal and illegal ship configurations.
+
+After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half game (see below for details).
+The order of pretend and non-pretend blocks was counterbalanced between participants.
+The allocation of boards to conditions was randomized between participants such that exactly one board was played in both pretend and non-pretend conditions, and this common board was different for different participants.
+The order of boards within a block was fully randomized, with the exception that half-games were always played last.
+
+#### Non-pretend (normal) games
+
+In non-pretend games (Fig.
+\@ref(fig:design)A), participants aimed to sink two 2-square patrol boats and one 3-square submarine with as few clicks as possible.
+An online counter of the number of clicks was displayed on the screen.
+After each game, feedback was given about the number of clicks and resulting number of points obtained.
+
+#### Pretend games
+
+Participants in pretend games were given the same explanation of Battleships, and played a practice round. However, they were then given an additional instruction:
+
+> "This time your goal is different.
+In this round, we're going to tell you where the ships are, but **we want you to act like you don't know this information**.
+We've marked the ships' locations with a cross, so you'll know where they are the whole time; but your job is to play the game as if these hints aren't there.
+To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are.
+We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you'll get bonus points.
+But if your games look different, you won't get these bonus points.
+Your number of clicks in this part will not affect your bonus.
+Only your ability to play like you had no hints."
+
+We informed participants that both the location and timing of their cell clicks will be measured.
+After one practice round and one comprehension question, participants played five pretend games (Fig.
+\@ref(fig:design)B), followed by one pretend half-game.
+Each game was followed by a short message, reminding them that a game that looks similar to the games of participants who had no hints would be awarded 10 bonus points.
+
+#### Half games
+
+In order to directly compare participants' pretend and non-pretend games for identical belief states (genuine or pretended ignorance about where the ships are hidden), participants completed one pretend and one non-pretend game given a partly finished board with the content of 7 cells already revealed (Fig.
+\@ref(fig:design)C).
+We designed our half games to produce a strong expectation to find a ship in specific cells, but not in others.
+The assignment of half-completed boards to pretend and non-pretend conditions was randomized between participants.
+
+#### Judge trials
+
+In the final part of the experiment, participants observed the games of previous players and tried to determine who had hints and who didn't.
+On each trial, two empty grids were presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid (Fig.
+\@ref(fig:design)D).
+The two grids corresponded to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders.
+Only games shorter than one minute from the group of participants that pretended in the second block were chosen for presentation in this part.
+Judge participants observed a real time replay of the two grids, showing not only where participants clicked, but also when.
+After making a decision, participants were informed whether they would receive the 10 bonus points, or alternatively, whether the pretender would receive them in the event the pretender managed to trick them.
+
+Readers are invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/NervzpM0Y0z).
+
+## Hangman
+
+A detailed pre-registration can be accessed at [osf.io/3thry](https://osf.io/3thry).
+The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019].
+Due to an error in the experiment code, time-locking took effect only from player number 221 (batch 3) and on (protocol sum: c4929c7fe33df1b7b52f15c789d98eab30a9cee09a8121807a3c59e28e7430a4; [relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/f98903f3f2d0e21800497dc65f8d5f823c910c88/experiments/Hangman2/code/index.html#L772-L783)).
+
+### Participants
+
+The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
+500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
+They were selected based on their acceptance rate (\>95%) and for being native English speakers.
+The entire experiment took approximately 20 minutes to complete.
+Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment of 1 USD that was awarded to `r E4.bonus$bonus%>%sum(na.rm=T)` players who earned 100 points or more.
+
+### Procedure
+
+The first instructions screen informed participants that the experiment, based on the game Hangman, had three parts, and that the points they accumulate translate to a monetary bonus payment.
+They were then be presented with a leaderboard of previous players.
+Then, the rules of the game were presented:
+
+> In the following game, your task is to reveal a hidden word or phrase by guessing letters.
+What makes the game difficult is that you can't see the word; all you can see is a row of squares - a square for each letter.
+Use your mouse to make letter guesses.
+We will have five types of words: body parts, numbers, US states, fruit, and famous people.
+You will start each game with 15 points and lose one point for every guess of a letter that is not in the word.
+
+We then explained that "the words in this game are the kind of words that will be familiar to most English-speaking fifth-graders. We didn't pick any strange or particularly difficult words".
+
+Once they responded correctly to a multi-choice comprehension question ("the goal of the game is to...": "reveal the word with as few letter guesses as possible"), participants played a practice round, revealing the word PIGEON (see Fig.
+\@ref(fig:design)E).
+
+After the main instructions, comprehension question and practice round, participants completed one pretend and one non-pretend block, each followed by one half-game (see below for details).
+The order of pretend and non-pretend blocks was counterbalanced between participants.
+Each block comprised five games played with five out of ten different words, and one half-game.
+The allocation of words to conditions was randomized between participants, with the constraint that both pretend and non-pretend blocks included exactly one word from each category.
+The order of words within a block was randomized, except for the half-game, which was always delivered at the end.
+
+The ten words included two number words (ELEVEN, NINETY SIX), two famous people (DALAI LAMA, TAYLOR SWIFT), two fruits (STRAWBERRY, LEMON), two body parts (TOOTH, HEAD), and two US states (MONTANA, IOWA).
+
+### Non-pretend games
+
+In non-pretend games, participants revealed a hidden word with as few letter guesses as possible.
+An online counter of the number of points was displayed on the screen, deducting one point for every guess of a letter that is not in the target word.
+After each game, feedback was given about the number of points obtained.
+
+After completing the five games, participants performed one half-game (see below for details).
+
+### Pretend games
+
+Participants were given the following instructions:
+
+> In the next part of the experiment, you'll play 6 games where you reveal a hidden word by guessing letters.
+
+> However, this time your goal is different.
+
+> In this round, we're going to tell you the word in advance, but **we want you to act like you don't know this information**.
+
+> To see how good you are at this, we're going to compare your games to the games of people who played normally, without knowing what the word was, and see how similar they are.
+We will measure which letters you click and the timing of your guesses; if your clicks look similar to people who played like normal (trying to reveal the word with as few guesses as possible, but without any hints), you'll get bonus points.
+But if your games look different, you won't get these bonus points.
+Your number of clicks in this part will not affect your bonus.
+Only your ability to play like you didn't see the word in advance.
+
+After one practice round, pretending not to know that the hidden word is PIGEON, and one comprehension question ("In this part of the experiment my goal is to...": "play the game as if I don't know what the word is so that I look like someone who had no hints"), participants played five pretend games (Fig.
+\@ref(fig:design)F).
+Each game was preceded by a short message informing subjects about the identity of the target word.
+To start pretending, players were asked to type in the target word on their keyboard.
+After pretending, we reminded players that a game that looks similar to the games of participants who had no hints will be awarded 10 bonus points.
+
+After completing the five games, participants performed one half-game (see below for details).
+
+### Half games
+
+In order to directly compare participants' pretend and non-pretend games for identical belief states (true or pretended knowledge about the identity of the word), we asked participants to also complete one pretend and one non-pretend game, given a partly finished game with some letters already guessed (they were told that the computer made these guesses; Fig.
+\@ref(fig:design)G).
+The two half-game words were one fruit: PAPAYA or BANANA, with guessed letters [A, E, I, O, M, T], and one body part: HAND, or HAIR with guessed letters (A, E, O, M, T, H, P).
+The assignment of category (fruit or body part) to condition (pretend and non-pretend), as well as the identity of the target word within each category (e.g., PAPAYA or BANANA), was randomized between participants.
+
+Instructions for the non-pretend half-game were:
+
+> For the next game, the computer chose the first letters for you; you can take over from where it left off.
+Your challenge is to complete the game.
+Just like in the previous games, here also you will lose one point for each letter that you guess and is not in the word.
+
+Instructions for the pretend half-game were:
+
+> For the next game, the computer chose the first letters for you; you can take over from where it left off.
+Just like in the previous games, here also you will know what the word is, but your bonus points will depend on your ability to play as if you didn't know the word.
+
+### Judge trials
+
+In the final part of the experiment, participants observed five games of previous players and determined who had hints and who didn't. Instructions for this part were:
+
+> In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of the players were shown the word (but acted like they weren't).
+We will show you 5 replays of the games of previous players.
+Your task is to decide whether they played normally or pretended.
+For each game that you get right, you will receive 10 points.
+Good luck!
+
+Then, on each judge trial, one game of a previous player was replayed in real time, with the target word presented above.
+For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
+For both pretend and non-pretend games, only games shorter than 1.5 minutes were presented.
+Judge participants indicated their decision by pressing the P and N keys on their keyboard.
+After making a decision, participants were informed whether they received the 10 points.
+Whenever a pretend game was classified as a non-pretend game, they were informed that the pretender received these 10 points instead of them.
+
+Lastly, participants were asked the following debrief questions:
+
+> Did you have a strategy that you used for pretending you did not see the word?
+What was most difficult about pretending?
+How about telling between players who pretenders and who played for real - did you have a strategy for that?
+
+And:
+
+> We would appreciate it if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.
+
+Readers are invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/5bpW3f50OG9).
 

### 2024-01-07 docs/ExploratoryResults.Rmd
fixed a typo

@@
   ggplot(aes(x=nonpretend,y=pretend, label=toupper(word))) +
-  geom_abline(intercept=0,slope=1) +
+  geom_abline(intercept=0,slope=1, color='lightgray') +
   geom_point()+
@@
 
-Among simulators, pretend games were significantly more optimal than non-pretend games (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`) even when restricting the analysis to letter selections that resulted in a miss (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`).
+Among simulators, pretend games were significantly more optimal than random (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`) even when restricting the analysis to letter selections that resulted in a miss (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`).
 Like in the full sapmle, simulators were more optimal when playing normally than when pretending (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`), also when focusing on misses (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).

### 2024-01-07 docs/ExploratoryResults.Rmd
typo

@@
 Among simulators, pretend games were significantly more optimal than random (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`) even when restricting the analysis to letter selections that resulted in a miss (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test())$full_result`).
-Like in the full sapmle, simulators were more optimal when playing normally than when pretending (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`), also when focusing on misses (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+Like in the full sample, simulators were more optimal when playing normally than when pretending (`r apa_print(E4.sim.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`), also when focusing on misses (`r apa_print(E4.sim.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
 

### 2024-01-07 docs/paper.Rmd
Ian's comments

@@
 ---
-title             : "Pretense reveals a capacity for online simulation of counterfactual belief states"
-shorttitle        : "Pretense reveals a capacity for online simulation of counterfactual belief states"
+title             : "Pretending not to know: Pretense behavior reveals the limits of self-simulation"
+shorttitle        : "Pretending not to know"
 
@@
   - name          : "Matan Mazor"
-    affiliation   : "1,2"
+    affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
-    address       : "Malet Street, London WC1E 7HX"
-    email         : "mtnmzor@gmail.com"
+    address       : "All Souls College, High Street, Oxford OX1 4AL"
+    email         : "matan.mazor@all-souls.ox.ac.uk"
   #   role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
   - name          : "Chaz Firestone"
-    affiliation   : "3"
+    affiliation   : "2"
 
   - name          : "Ian Phillips"
-    affiliation   : "3"
+    affiliation   : "2"
     # role:
@@
   - id            : "1"
-    institution   : "Birkbeck, University of London"
+    institution   : "University of Oxford"
   - id            : "2"
-    institution   : "Wellcome Centre for Human Neuroimaging, UCL"
-  - id            : "3"
     institution   : "Johns Hopkins University"
@@
 But sometimes we try to decouple our behavior from our knowledge so as to appear not to know what we really do know.
-Such pretense behavior requires understanding how we would behave with different knowledge — and so provides a window into counterfactual self-simulation.
+Such pretense behavior requires understanding how we would behave with different knowledge, and so provides a window into counterfactual self-simulation.
 However, little research has characterized and evaluated pretense relative to non-pretense behavior.
 Here, in two large-scale, pre-registered experiments, we tested people’s ability to conceal knowledge in the context of information-seeking games (Battleship and Hangman).
-Participants played normal games, as well as ‘pretend’ games in which they had full access to the game state, but pretended they didn’t.
-Overall, we find a highly impressive capacity for pretense — so impressive that independent 'judge' participants were completely unable to discriminate between normal and ‘pretend’ games.
-Specifically, pretenders successfully reproduced subtle features of real players' decision-making and timing.
-Nonetheless, pretense was characterized by pervasive differences (undetected by judges but revealed by follow-up analyses): pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules.
-Our findings thus reveal a powerful but imperfect ability to simulate decisions and behaviour based on a counterfactual belief state, and open up a new approach to studying self-simulation.
+Participants played normal games (in which they began with no knowledge of the ships’ locations or the word’s identity), as well as ‘pretend’ games in which they were shown the game's full solution but were instructed to play *as though* they were ignorant.
+We find a remarkable capacity for pretense, with pretenders successfully reproducing broad patterns and subtle features of real players’ decision-making and timing — so much so that 'judge' paticipants failed to distinguish real players from pretenders.
+Nonetheless, pretense was characterized by pervasive differences, undetected by judges but revealed by follow-up analyses: pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules.
+Our findings thus reveal a powerful but imperfect ability to simulate decisions and behavior based on a counterfactual belief state, and open up a new approach to studying self-simulation.
 
@@
 keywords          : "pretense; metacognition; theory of mind"
-wordcount         : "X"
+wordcount         : "5986"
 
@@
 
-Pretense relies on an ability to simulate and mimic one's own behavior under a hypothetical belief state.
-For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that.") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first?
+Pretense relies on an ability to simulate and mimic one's own behavior under a counterfactual belief state.
+For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that...") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first had I not known?
 What would I say?
 How long would it take me to recover from the surprise?").
-This is not a trivial challenge: previous research demonstrates that access to knowledge distorts our beliefs about what we would have thought [@fischhoff1975] or perceived [@harley2004] without that knowledge.
-Indeed, based on previous research, it is possible that people are completely unable to mentally simulate their own behaviour under a counterfactual state of ignorance.
+This is not a trivial challenge: previous research on "hindsight biases" suggests that knowledge about the actual state of the world can interfere with our ability to correctly judge what we would have believed [@fischhoff1975; @fischhoff1977; @wood1978; @roese2012] or perceived [@harley2004; @bernstein2007; @bernstein2012] without this knowledge.
+Such biases remain potent even when instructing participants to overcome them [@harley2004; @pohl1996].
+Moreover, even if pretenders can correctly determine what they would have believed, they must further accurately simulate how they would think and behave in this different belief state.
 
-<!-- to what extent humans are capable of generating complex behaviour that is aligned with a counterfactual belief state, and specifically with a counterfactual state of ignorance. -->
+The reliance of this kind of epistemic pretense on self-simulation makes it a promising tool for revealing the structure and content of people's internal models of their own minds.
+When directly asked, participants are able to provide relatively accurate descriptions of their own decision-making [@morris] and perception [@levin2008; @mazor2023].
+Pretending not to know opens a new window into the structure and content of this metacognitive knowledge, with two important advantages.
+First, by not relying on explicit reports, pretense has the potential to reveal implicit self-knowledge -- that is, structured knowledge about the self that is not reportable.
+And second, data obtained from pretense experiments can be analyzed and modeled using the same tools employed by cognitive scientists to study non-pretense behavior, affording a direct and finer-grained comparison between pretend and genuine decision-making.
+
+<!-- to what extent humans are capable of generating complex behavior that is aligned with a counterfactual belief state, and specifically with a counterfactual state of ignorance. -->
 
 To this end, we examine pretense in a game setting.
-Using an online version of the games Battleships and Hangman, participants played a 'non-pretend' (normal) version of the game, as well as a 'pretend' version where they were given secret information about the location of hidden ships/target word, but were instructed to behave as if they didn't have this information.
-Overall, we find a highly impressive capacity for pretense, with pretenders mirroring subtle features of real players' decision-making and timing.
-At the same time, pretense was characterized by pervasive differences, with pretenders consistently playing less rationally than normal players, sometimes rigidly following heuristics and rules.
-Whilst powerful, then, our capacity for self-simulation is imperfect, displaying tell-tale biases in our internal models of our own cognition and decision making.
+Using an online version of the games Battleship and Hangman (in which players seek to uncover the locations of enemy ships or the identity of a word), participants played a 'non-pretend' (normal) version of the game, as well as a 'pretend' version where they were given complete information about the hidden ships / target word but were instructed to behave as if they didn't have this information.
+Overall, we find a highly impressive capacity for pretending not to know, with pretenders mirroring broad patterns and subtle features of real players' decisions and decision times.
+However, we also find that pretense was characterized by pervasive biases and limitations, although these were undetected by human observers.
 
@@
 
-In two experiments, players played online versions of two information-seeking games: Battleship (N=`r E2.df$subj_id%>%unique()%>%length()`) and Hangman (N=`r E4.df$subj_id%>%unique()%>%length()`).
-These games typically start in a state of ignorance, where players' goal is to reveal an unknown world state (grid configuration in Battleship, a hidden word in Hangman) with as few steps (cell or letter selections) as possible.
-Critically, in addition to playing five games in their standard version, players also completed five pretend games in which the true world state was known to them from the start, but their task was to behave as if it was not (see Fig.
+In two experiments, players played online versions of two information-seeking games: Battleship (N=`r E2.df$subj_id%>%unique()%>%length()` players) and Hangman (N=`r E4.df$subj_id%>%unique()%>%length()` players).
+These games traditionally start in a state of ignorance, with a player's goal being to reveal an unknown world state (ship locations in Battleship, a hidden word in Hangman) in as few steps (cell or letter selections) as possible.
+Critically, in addition to playing five standard games, players in our experiments also completed five 'pretend' games in which the solution to the game was known to them from the start, but where their task was to behave as if it was not -- i.e.
+to play as though they did not have this information (see Fig.
 \@ref(fig:design)).
-We measured their capacity to simulate a counterfactual knowledge state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games.
-Our full pre-registered results are available [online](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html).
+We measured participants' capacity to simulate a counterfactual state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games.
+Our full pre-registered results are available [online](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html) together with the report-generating code.
+Readers are invited to try [demos of the experiments](https://matanmazor.github.io/ignorance/experiments/demos/pretend).
 
@@
 In non-pretend games, players revealed ships by guessing cells in a grid (A) or revealed a word by guessing letters (E).
-In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (F), but asked players to play as if they didn't have access to this information.
-In half games (C and G), players continued the game from a given state, with the content of some cells and the position of some letters already revealed.
+In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (F), but asked players to play as if they didn't have this information.
+In half-games (C and G), players continued the game from a given state, with the content of some cells and the position of some letters already revealed.
 Lastly, players watched replays of the games of previous players and guessed which were pretend games (D and H).", out.width = '100%'}
@@
 
-As a first gross measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games.
-Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()`; $p=$ `r apa_print(E2.median_clicks$diff%>%t.test())$table$p.value`; Fig.
+## Measuring pretense quality
+
+As a first measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games.
+Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()%>%printnum()`; $p=$ `r apa_print(E2.median_clicks$diff%>%t.test())$table$p.value`; Fig.
 \@ref(fig:decisions)A).
-In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()`; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()`; $p$ `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+`r E2.lucky_pretenders%>%length` pretenders who immediately discovered all ships without making errors were excluded from all further analyses, in accordance with our pre-registered plan.
+With these subjects excluded, the number of cell selections remained very similar in pretend (mean = \`r `` E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()` ``) and non-pretend games (mean = \`r `` E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()` ``; $p=$`r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+\@ref(fig:decisions)A).
+`r E2.lucky_pretenders%>%length`).
+In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()%>%printnum()` misses; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()%>%printnum()` misses; $p$ `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
 \@ref(fig:decisions)B).
-Despite an overall bias in the number of guesses, pretend Hangman games provided evidence for a more impressive, item-specific alignment: pretenders were successful in making more letter guesses when attempting to reveal words that would have been harder to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`; Fig.
+Despite an overall bias in the number of guesses, pretend Hangman games showed an impressive, item-specific alignment: pretenders were successful in making more letter guesses when attempting to reveal words that would have been harder to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`; Fig.
 \@ref(fig:decisions)B).
-This correlation provides a first clue for a human capacity to act in accordance with a counterfactual knowledge state.
+This strong correlation provides evidence for a human capacity to act in accordance with a counterfactual knowledge state.
 
 (ref:decisions-caption) Battleship and Hangman guesses in pretend and non-pretend games.
-A: median number of misses Battleship and Hangman games, in non-pretend (green) and pretend (purple) games.
+A: median number of misses in Battleship and Hangman games, in non-pretend (green) and pretend (purple) games.
 For reference, the expected number of misses is indicated by a reference line for a fully random agent, and for a "greedy" agent that maximizes the probability of a hit in each step.
 B: The median number of misses in Hangman for pretend and non-pretend games, as a function of the target word.
-C: in half-games, players continued the game from a half-finished state.
-Showing spatial guess distributions for pretend and non-pretend games, alongside their corresponding hit probability maps.
+C: Spatial guess distributions for pretend and non-pretend half-games (where players continued the game from a half-finished state) alongside their corresponding hit probability maps.
 D: Cell and letter selections were ranked according to their relative hit probability.
@@
 When playing Battleship and Hangman, it makes sense to guess cells or letters for which the probability of hitting a ship or revealing a letter is high [this "greedy" behavior is not strictly optimal, but approximates optimal behavior in most cases, @audinot2014optimal, Section 3.3].
-In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank_wide%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; Fig.
+In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
+Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(np_vs_r))%>%apa_print%>%'$'(statistic)`; Fig.
 \@ref(fig:decisions)D).
-Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; Fig.
-\@ref(fig:decisions)E): that is, even when missing ships or letters, pretend guesses made sense given the limited information players pretended to have.
+Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r t.test(-E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`; Fig.
+\@ref(fig:decisions)E): that is, even when incorrectly guessing a ship's location or a letter's identity, pretend guesses made sense given the limited information players pretended to have.
 
@@
 \@ref(fig:time)B).
+In other words, despite knowing the game's solution with full certainty, pretenders successfully feigned subtle qualitative effects of counterfactual uncertainty on their decision times.
 
-```{r time, echo=FALSE, message= F, warning=F, fig.cap="Patterns of reaction time in pretend and non-pretend guesses.
-A: median decision times in and following hits and misses.
+```{r time, echo=FALSE, message= F, warning=F, fig.cap="Patterns of decision time in pretend and non-pretend games.
+A: median decision times forhits and misses, as well as the decisions following them.
 In both Battleship and Hangman, hits were faster on average than misses, but guesses following a hit were slower on average than those following a miss.
@@
 
-This capacity for simulating a state of ignorance was not perfect.
-First, despite showing the same qualitative effects, decision times in Battleship (but not Hangman) pretend games were systematically more pronounced relative to non-pretend games, as if participants were acting based on an exaggerated mental image of their own counterfactual behaviour.
-Furthermore, in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result.
-The within-participant sequential guess entropy, a measure of the inverse predictability of decision sequences across different games of the same player, was significantly reduced in pretend games (Fig.
+## Pretense limitations
+
+Though impressive, the capacity for simulating a state of ignorance was not perfect.
+First, despite showing the same qualitative effects, decision time patterns in Battleship pretend games (but not Hangman pretend games) were systematically more pronounced relative to non-pretend games, as if participants were acting based on an exaggerated mental model of their own counterfactual behavior.
+Furthermore, the within-participant sequential guess entropy, a measure of the inverse predictability of decision sequences across different games of the same player, was significantly reduced in pretend games (Fig.
 \@ref(fig:rigidity)A; see [online supplementary materials](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#within-participant-click-entropy)).
-More specifically, Hangman pretenders were more likely to guess letters that are frequent in the English language (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig.
+More specifically, Hangman pretenders were more likely to guess letters that appear frequently in English words (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig.
 \@ref(fig:rigidity)B).
+This suggests that in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result.
 
-This limitation on incorporating evidence into the decision-making process was especially evident in Hangman half-games, where players completed the game from a half-completed state.
+This limitation on incorporating evidence into the (simulated) decision-making process was especially evident in Hangman half-games, where players completed the game from a half-completed state.
 When asked to reveal the hidden fruit "\_A_A_A", `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('banana', 'papaya'))%>%pull(bananahand)%>%mean())`% of the non-pretenders guessed one of the letters 'B' or 'N' (Fig.
@@
 Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word== 'banana')%>%pull(bananahand)%>%mean())`%.
-Importantly, half of the pretenders were given different information: they were told that the hidden word is the less prototypical fruit PAPAYA.
-Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', revealing that many pretenders were unable to predict that the first fruit that would come to their minds is BANANA, not PAPAYA (Fig.
+Importantly, half of the pretenders were given different information: they were told that the hidden word was the less prototypical fruit PAPAYA.
+Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', revealing that many pretenders were unable to predict that the first fruit that would have come to their minds was BANANA, not PAPAYA (Fig.
 \@ref(fig:rigidity)C, yellow bars in right column).
 A similar pattern was observed for the prototypical body part word HA(ND) and its surprising counterpart HA(IR): when playing normally, `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('hand', 'hair'))%>%pull(bananahand)%>%mean())`% of the players selected letters that are consistent with the prototypical option HAND.
-Pretenders for whom the word was HAND showed the same tendency (`r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hand')%>%pull(bananahand)%>%mean())`%), in contrast to only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hair')%>%pull(bananahand)%>%mean())`% of the pretenders who saw the word HAIR (Fig.
+This figure was `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hand')%>%pull(bananahand)%>%mean())`% among pretenders for whom the target word was HAND, in contrast to only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hair')%>%pull(bananahand)%>%mean())`% among pretenders for whom the target word was HAIR (Fig.
 \@ref(fig:rigidity)C, blue bars).
@@
 A: sequential guess entropy, a measure of the (inverse) predictability of individual players' guesses as a function of click number and guess number.
-In both Battleship and Hangman, Sequential guess entropy increased with click number, and was overall lower in pretend games.
+In both Battleship and Hangman, sequential guess entropy increased with click number, and was overall lower in pretend games.
 B: letter frequency of Hangman guesses: the median rank frequency per participant in pretend and non-pretend games, with reference lines for the expected rank frequency for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step.
-C: first letter guesses in Hangman half games, as a function of pretense condition and target word.
+C: first letter guesses in Hangman half-games, as a function of pretense condition and target word.
 Letters appear in alphabetical order.
@@
 
-Despite these limitations, pretenders were able to fool other players into believing they were playing for real.
-After performing pretend and non-pretend games, participants were presented with game replays of previous players and judged who was pretending and who was playing for real.
+## Failure to detect pretending
+
+Despite these limitations, pretenders' behavior resembles real players' behaviour closely enough that they were able to fool other participants into believing they were playing for real.
+After performing pretend and non-pretend games, participants were presented with game replays of previous players and took on the new role of being a 'judge' who had to determine who was pretending and who was playing for real.
 In both games, pretense detection was at chance (Battleship: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%, Hangman: `r round(E4.judge_summary_df$accuracy%>%mean()*100)`%).
 This is in line with previous findings of near-chance accuracy in lie detection [@bond2006].
-Moreover, we find no sign of a correlation between pretense quality (measured as players' ability to trick judges to think they were not pretending) and pretense detection ability (measured as proportion correct; Battleship: `r apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate, method='spearman'))$estimate`, Hangman: `r apa_print(cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman'))$estimate`), indicating that pretense and pretense detection rely on at least partly different cognitive processes.
+Moreover, we find no sign of a correlation between pretense quality (measured as players' ability to trick judges into thinking they were not pretending) and pretense detection ability (measured as proportion correct; Battleship: `r apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate, method='spearman'))$estimate`, Hangman: `r apa_print(cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman'))$estimate`), indicating that pretense and pretense detection rely on at least partly different cognitive processes.
 
@@
 
-In two experiments, we examined participants' ability to mimic a state of ignorance in a game setting.
-We find that pretenders were able to successfully emulate decisions taken under a true state of ignorance, including reproducing subtle effects of guess accuracy and decision uncertainty on decision time.
-We also identify reliable signatures of pretend-ignorance on players' decisions, including a cost to decision rationality and an increased tendency to follow heuristics and rules.
-Together, our findings reveal an impressive but limited capacity for an online simulation of counterfactual belief states.
+In two experiments, we examined participants' ability to mimic a state of ignorance in a game setting, building on the recent recognition of games as a powerful tool for studying decision making [@allen2024].
+We find that pretenders were able to successfully emulate decisions taken under a true state of ignorance, including reproducing both broad patterns and subtle effects of guess accuracy and decision uncertainty on decision time.
+We also identify reliable signatures of pretend-ignorance on players' decisions, including a cost to decision rationality and an increased tendency to follow heuristics and rules, even though these signatures went undetected by 'judges' asked to discriminate real from pretend games.
+Collectively, our findings reveal a rich and impressive, but ultimately limited, capacity for online simulation of counterfactual belief states.
 
-Previous research has identified limitations on the human capacity to encapsulate knowledge from affecting our decisions and behaviour [@fischhoff1975; @harley2004].
-Our findings reveal that notwithstanding these limitations, humans are capable of approximating their hypothetical behaviour had they not known what they do in fact know.
-This capacity is not limited to making similar decisions to the ones we would have made had we not known; we also find that pretend decision times reproduce qualitative patterns observed under a true state of ignorance.
-This suggests that our participants were not engaging in a superficial mimicry of counterfactual behavioural patterns, but instead simulated, in real time, the cognitive processes that give rise to these patterns.
+Previous research has identified limitations in our capacity to prevent knowledge from influencing our decisions and behavior [@fischhoff1975; @fischhoff1977; @wood1978; @roese2012; @harley2004].
+Our findings reveal that notwithstanding these limitations, humans are capable of approximating their hypothetical behavior had they not known what they in fact do know.
+This capacity goes beyond making similar decisions to the ones they would have made had they not known; pretenders were also able to generate decision times that reproduce subtle qualitative patterns observed under a true state of ignorance.
 
-An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored the revealed game state such that their entire cognitive machinery was available to play the game.
-This does not require any self-simulation, only a capacity to intentionally unsee, or forget, relevant evidence.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense for three reasons.
-First, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical].
-Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with self-simulation or rule-following (see [exploratory analysis]([https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).) Finally, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig.
+This alignment between pretend and non-pretend decision times may indicate that pretenders were capable of controlling their decision times at will, enacting their intuitive beliefs of how decision time should be affected by knowledge and uncertainty.
+In a recent developmental study, children aged 5-10 were capable of making inferences about knowledge from the observed response time of other agents [e.g., inferring that decision-makers who take longer to decide make higher-quality decisions, @richardson2022].
+Our findings here may show that people are also capable of using this sort of intuitive knowledge to guide their own pretense behavior.
+Alternatively, this mimicry of decision time patterns may reflect a similarity in the computations leading to pretend and non-pretend decision making.
+According to this interpretation, pretenders may have paused for longer when making some decisions because the simulated computation that led to these decisions took longer.
+
+Research on Bayesian Theory of Mind provides some support for such internal simulations of decision-making processes.
+These are often studied by measuring participants' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational; @richardson2022], or implicitly [@onishi2005fifteen; @liu2017ten].
+Here we have proposed a complementary approach: Asking participants to generate behavior based on a counterfactual mental state---In this case, a counterfactual knowledge state in which a known piece of information is unknown.
+Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking counterfactual beliefs and desires as input and producing behavior as output.
+Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
+
+An alternative interpretation of our results is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored the revealed game state such that their entire cognitive machinery was available to play the game.
+This would not require self-simulation, only a capacity to intentionally 'unsee', or forget, relevant evidence.
+While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense, for at least four reasons.
+First, we tried to make such suppression as hard as possible, by presenting the game solution on top of the game board for the entire duration of pretend games, and by having participants type the target word before pretend Hangman games.
+Second, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical].
+Third, when asked how they had performed the task in a debrief question, the responses of a significant majority of participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).
+Finally, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig.
 \@ref(fig:time)A): a finding that is consistent with the simulation of a simplified self-model, but is hard to explain if participants were merely suppressing their current knowledge.
 
-Research on Bayesian Theory of Mind provides some support for such internal simulations of decision-making processes.
-This is often studied by measuring participants' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational], or implicitly [@onishi2005fifteen; @liu2017ten].
-Here we have put forward a complementary approach: Asking participants to generate behavior based on a counterfactual mental state.
-In this case, a counterfactual knowledge state in which a known piece of information is unknown.
-Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking beliefs and desires as input and producing behavior as output.
-Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
-
-While previous work focused on people's inability to perfectly predict their behaviour had they not known something ["hindsight bias", @fischhoff1975, @roese2012], our paradigm and analysis revealed a more nuanced picture, with an impressive but limited capacity to simulate a counterfactual state of ignorance.
-Alongside pretenders' many successes, we identify specific limitations, including a tendency to follow heuristics and rules and a scaling of response time profiles.
-These limitations are useful in showing that pretending not to know relies on more than the simulation of the same computational mechanisms that underlie genuine decision-making [@saxe2005].
-More work is now needed to trace the boundaries of the capacity to pretend not to know, and how these boundaries map onto limitations in the human capacity to simulate counterfactual mental states.
+Together, our findings reveal a non-trivial capacity for pretending not to know.
+Complementing previous work on cognitive and perceptual hindsight biases, we show that people are capable of accurately simulating diverse aspects of their decision-making processes, although they exhibit systematic shortcomings.
+Further research into these and similar limitations may continue to reveal the simplifications, abstractions, and biases in people's models of their own minds.
 
@@
 The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019] (protocol sum: 60c270410375e8a192468fc1a0e9c93da60d5e203eb2760b621a8631a26f4c5c; [link to relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/44ef964c0d7d4a8f3b76ec52316a8e624c41a047/experiments/Battleships2/code/index.html#L865-L876)).
+All pre-registered analyses are available [in this link](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html).
 
@@
 
-The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
-500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
+The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.
+`E2.df$subj_id%>%unique()%>%length()` Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
 They were selected based on their acceptance rate (\>95%) and for being native English speakers.
 The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
+Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r``E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
 
@@
 
-Participants were first instructed that the experiment, based on the game Battleships, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment.
+Participants were first instructed that the experiment, based on the game Battleship, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment.
 They were then presented with a leaderboard of previous players, and given the rules of the game:
 
-> "In the game Battleships, your task is to sink all ships located in a grid with as few clicks as possible.
+> *"In the game Battleship, your task is to sink all ships located in a grid with as few clicks as possible.
 What makes the game difficult is that you can't see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
@@
 If you hit part of a ship, the square will turn red.
-If there is no ship in the square, it will turn blue."
+If there is no ship in the square, it will turn blue.*"
 
@@
 
-After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half game (see below for details).
+After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half-game (see below for details).
 The order of pretend and non-pretend blocks was counterbalanced between participants.
@@
 
-Participants in pretend games were given the same explanation of Battleships, and played a practice round. However, they were then given an additional instruction:
+Participants in pretend games were given the same explanation of Battleship, and played a practice round. However, they were then given an additional instruction:
 
-> "This time your goal is different.
+> *"This time your goal is different.
 In this round, we're going to tell you where the ships are, but **we want you to act like you don't know this information**.
@@
 Your number of clicks in this part will not affect your bonus.
-Only your ability to play like you had no hints."
+Only your ability to play like you had no hints.*"
 
-We informed participants that both the location and timing of their cell clicks will be measured.
+We informed participants that both the location and timing of their cell clicks would be measured.
 After one practice round and one comprehension question, participants played five pretend games (Fig.
@@
 
-#### Half games
+#### Half-games
 
@@
 \@ref(fig:design)C).
-We designed our half games to produce a strong expectation to find a ship in specific cells, but not in others.
+We designed our half-games to produce a strong expectation to find a ship in specific cells, but not in others.
 The assignment of half-completed boards to pretend and non-pretend conditions was randomized between participants.
@@
 
-In the final part of the experiment, participants observed the games of previous players and tried to determine who had hints and who didn't.
+```{r judge-game-proportion-Battleship, echo=FALSE, message= F, warning=F}
+E2.judge_np_game_proportion <- E2.click_log %>% 
+  filter(genuine_first & test_part=='nonpretend') %>%
+  group_by(subj_id,test_part,grid_number, genuine_first) %>% 
+  summarise(duration=sum(RT)/1000) %>% 
+  mutate(included = duration<=60) %>%
+  pull(included) %>%
+  mean()
+
+E2.judge_p_game_proportion <- E2.click_log %>% 
+  filter(test_part=='pretend') %>%
+  group_by(subj_id,test_part,grid_number, genuine_first) %>% 
+  summarise(duration=sum(RT)/1000) %>% 
+  mutate(included = duration<=60) %>%
+  pull(included) %>%
+  mean()
+```
+
+In the final part of the experiment, participants observed the games of previous players and tried to determine who were pretenders and who were non-pretenders.
 On each trial, two empty grids were presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid (Fig.
@@
 The two grids corresponded to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders.
-Only games shorter than one minute from the group of participants that pretended in the second block were chosen for presentation in this part.
+For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
+For both pretend and non-pretend games, only games shorter than one minute (`r (E2.judge_np_game_proportion*100)%>%round()`% of included non-pretend games and `r (E2.judge_p_game_proportion*100)%>%round()`% of pretend games) were presented.
 Judge participants observed a real time replay of the two grids, showing not only where participants clicked, but also when.
@@
 
-Readers are invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/NervzpM0Y0z).
+Readers are invited to try a [demo of the experiment](https://matanmazor.github.io/ignorance/experiments/demos/Battleship).
 
@@
 
-The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
-500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
+The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.
+`E4.df$subj_id%>%unique()%>%length()` Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
 They were selected based on their acceptance rate (\>95%) and for being native English speakers.
@@
 The first instructions screen informed participants that the experiment, based on the game Hangman, had three parts, and that the points they accumulate translate to a monetary bonus payment.
-They were then be presented with a leaderboard of previous players.
+They were then presented with a leaderboard of previous players.
 Then, the rules of the game were presented:
 
-> In the following game, your task is to reveal a hidden word or phrase by guessing letters.
+> *"In the following game, your task is to reveal a hidden word or phrase by guessing letters.
 What makes the game difficult is that you can't see the word; all you can see is a row of squares - a square for each letter.
@@
 We will have five types of words: body parts, numbers, US states, fruit, and famous people.
-You will start each game with 15 points and lose one point for every guess of a letter that is not in the word.
+You will start each game with 15 points and lose one point for every guess of a letter that is not in the word.*"
 
@@
 
-Once they responded correctly to a multi-choice comprehension question ("the goal of the game is to...": "reveal the word with as few letter guesses as possible"), participants played a practice round, revealing the word PIGEON (see Fig.
+Once they responded correctly to a multiple-choice comprehension question ("the goal of the game is to...": "reveal the word with as few letter guesses as possible"), participants played a practice round, revealing the word PIGEON (see Fig.
 \@ref(fig:design)E).
@@
 
-### Non-pretend games
+#### Non-pretend games
 
@@
 
-> In the next part of the experiment, you'll play 6 games where you reveal a hidden word by guessing letters.
+> *"*In the next part of the experiment, you'll play 6 games where you reveal a hidden word by guessing letters.
 
@@
 Your number of clicks in this part will not affect your bonus.
-Only your ability to play like you didn't see the word in advance.
+Only your ability to play like you didn't see the word in advance.*"*
 
@@
 To start pretending, players were asked to type in the target word on their keyboard.
+The target word remained on the screen, in green letters, until the end of the game.
 After pretending, we reminded players that a game that looks similar to the games of participants who had no hints will be awarded 10 bonus points.
@@
 
-### Half games
+#### Half-games
 
@@
 
-> For the next game, the computer chose the first letters for you; you can take over from where it left off.
+> *"For the next game, the computer chose the first letters for you; you can take over from where it left off.
 Your challenge is to complete the game.
-Just like in the previous games, here also you will lose one point for each letter that you guess and is not in the word.
+Just like in the previous games, here also you will lose one point for each letter that you guess and is not in the word.*"
 
@@
 
-> For the next game, the computer chose the first letters for you; you can take over from where it left off.
-Just like in the previous games, here also you will know what the word is, but your bonus points will depend on your ability to play as if you didn't know the word.
+> *"For the next game, the computer chose the first letters for you; you can take over from where it left off.
+Just like in the previous games, here also you will know what the word is, but your bonus points will depend on your ability to play as if you didn't know the word.*"
 
-### Judge trials
+#### Judge trials
+
+```{r judge-game-proportion-Hangman, echo=FALSE, message= F, warning=F}
+E4.judge_np_game_proportion <- E4.click_log %>% 
+  filter(genuine_first & test_part=='nonpretend') %>%
+  group_by(subj_id,test_part,word, genuine_first) %>% 
+  summarise(duration=sum(RT)/1000) %>% 
+  mutate(included = duration<=90) %>%
+  pull(included) %>%
+  mean()
+
+E4.judge_p_game_proportion <- E4.click_log %>% 
+  filter(test_part=='pretend') %>%
+  group_by(subj_id,test_part,word, genuine_first) %>% 
+  summarise(duration=sum(RT)/1000) %>% 
+  mutate(included = duration<=90) %>%
+  pull(included) %>%
+  mean()
+```
 
@@
 
-> In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of the players were shown the word (but acted like they weren't).
+> *"In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of the players were shown the word (but acted like they weren't).
 We will show you 5 replays of the games of previous players.
@@
 For each game that you get right, you will receive 10 points.
-Good luck!
+Good luck!*"
 
@@
 For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
-For both pretend and non-pretend games, only games shorter than 1.5 minutes were presented.
+For both pretend and non-pretend games, only games shorter than 1.5 minutes (`r (E4.judge_np_game_proportion*100)%>%round()`% of included non-pretend games and `r (E4.judge_p_game_proportion*100)%>%round()`% of pretend games) were presented.
 Judge participants indicated their decision by pressing the P and N keys on their keyboard.
@@
 
-> Did you have a strategy that you used for pretending you did not see the word?
+> *"Did you have a strategy that you used for pretending you did not see the word?
 What was most difficult about pretending?
-How about telling between players who pretenders and who played for real - did you have a strategy for that?
+How about telling between players who pretenders and who played for real - did you have a strategy for that?*"
 
@@
 
-> We would appreciate it if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.
+> *"We would appreciate it if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.*"
 
-Readers are invited to try a [demo of the experiment](https://jatos.mindprobe.eu/publix/5bpW3f50OG9).
+Readers are invited to try a [demo of the experiment](https://matanmazor.github.io/ignorance/experiments/demos/Hangman).
 

### 2024-01-07 docs/paper.Rmd
typo

@@
 Participants played normal games (in which they began with no knowledge of the ships’ locations or the word’s identity), as well as ‘pretend’ games in which they were shown the game's full solution but were instructed to play *as though* they were ignorant.
-We find a remarkable capacity for pretense, with pretenders successfully reproducing broad patterns and subtle features of real players’ decision-making and timing — so much so that 'judge' paticipants failed to distinguish real players from pretenders.
+We find a remarkable capacity for pretense, with pretenders successfully reproducing broad patterns and subtle features of real players’ decision-making and timing — so much so that 'judge' participants failed to distinguish real players from pretenders.
 Nonetheless, pretense was characterized by pervasive differences, undetected by judges but revealed by follow-up analyses: pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules.
@@
 The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r``E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
+Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (``` r``E2.bonus$bonus%>%min() ``` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
 

### 2024-01-07 docs/paper.Rmd
typo

@@
 `r E2.lucky_pretenders%>%length` pretenders who immediately discovered all ships without making errors were excluded from all further analyses, in accordance with our pre-registered plan.
-With these subjects excluded, the number of cell selections remained very similar in pretend (mean = \`r `` E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()` ``) and non-pretend games (mean = \`r `` E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()` ``; $p=$`r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
-\@ref(fig:decisions)A).
-`r E2.lucky_pretenders%>%length`).
+With these subjects excluded, the number of cell selections remained very similar in pretend (mean = `r  E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()`; $p=$`r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+\@ref(fig:decisions)A)).
 In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()%>%printnum()` misses; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()%>%printnum()` misses; $p$ `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
@@
 ```{r time, echo=FALSE, message= F, warning=F, fig.cap="Patterns of decision time in pretend and non-pretend games.
-A: median decision times forhits and misses, as well as the decisions following them.
+A: median decision times for hits and misses, as well as the decisions following them.
 In both Battleship and Hangman, hits were faster on average than misses, but guesses following a hit were slower on average than those following a miss.

### 2024-01-08 docs/paper.Rmd
british spelling

@@
 
-Despite these limitations, pretenders' behavior resembles real players' behaviour closely enough that they were able to fool other participants into believing they were playing for real.
+Despite these limitations, pretenders' behavior resembles real players' behavior closely enough that they were able to fool other participants into believing they were playing for real.
 After performing pretend and non-pretend games, participants were presented with game replays of previous players and took on the new role of being a 'judge' who had to determine who was pretending and who was playing for real.

### 2024-01-09 docs/ExploratoryResults.Rmd
online materials

@@
 
-This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their their response latency in the same way it would have been affected had they been in a non-pretend game.
+This last finding bears emphasizing: In both classes two and three, pretenders *knew* that they had just sunk a size-two ship, but in the second case they *pretended not to know* this fact, and this affected their response latency in the same way it would have been affected had they been in a non-pretend game.
 
@@
 ```
+
 ## Heuristic use
@@
 
-Flexibility scores increased as a function of click number for both pretend and non-pretend games, as expected if players adjust their behaviour based on the outcomes of previous clicks.
+Flexibility scores increased as a function of click number for both pretend and non-pretend games, as expected if players adjust their behavior based on the outcomes of previous clicks.
 This process reached a plateau of around $H=2.1$ at click number 6.
@@
 
-In the following analysis, we focus on the first two labels and carve out two sets of players: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but not suppression.
+In the following analysis, we focus on the first two labels and sort players into two groups: `r E2.suppressors%>%length` players who reported using suppression but not simulation, and `r E2.simulators%>%length` who reported using simulation but not suppression.
 
@@
 
-The mean negative optimality score of non-pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)))$statistic`).
-The mean negative optimality score of pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)))$statistic`).
+The mean negative optimality score of non-pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()%>%printnum()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)%>%mean()%>%printnum()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='nonpretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='nonpretend') %>%pull(p_click_rank)))$statistic`).
+The mean negative optimality score of pretend games was `r E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()%>%printnum()` among suppressors and `r E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)%>%mean()%>%printnum()` among simulators (`r apa_print(t.test(E2.sup_sim.mean_P_click_rank%>%filter(group=='suppressors' & test_part=='pretend') %>%pull(p_click_rank), E2.sup_sim.mean_P_click_rank%>%filter(group=='simulators' & test_part=='pretend') %>%pull(p_click_rank)))$statistic`).
+Note that a higher score here means lower optimality.
 To the very least, simulators did as well as suppressors in pretending.
@@
 
-Like for Battleship, flexibility scores increased as a function of click number for both pretend and non-pretend games, as expected if players adjust their behaviour based on the outcomes of previous clicks.
+Like for Battleship, flexibility scores increased as a function of click number for both pretend and non-pretend games, as expected if players adjust their behavior based on the outcomes of previous clicks.
 This process reached a plateau of around $H=2$ and click number 6.
@@
 To quantify the contribution of each of these factors to letter selection, we fitted a multinomial logistic regression model to the data of each player.
+The model was specified and fit using the Turing Julia library for Bayesian inference with probabilistic programming [@ge2018].
 We compared the ensuing coefficients in 'pretend' and 'non-pretend' conditions.
@@
 
-Given colinearity between model predictors, we suspected that the effects of expected information gain may be driven by posterior probability effects.
+Given collinearity between model predictors, we suspected that the effects of expected information gain may be driven by posterior probability effects.
 To isolate unique contributions of prior probability and expected information gain, we repeated the same analysis, but this time serially orthogonalizing the features (in the order [posterior, prior, expected information gain], and using the Gram Schmidt algorithm) prior to fitting the multinomial regression model.
-This time, expected information gain had no significant effect on decision in non-pretend (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='nonpretend')%>%pull(eig)%>%t.test())$statistic`) and pretend games (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='pretend')%>%pull(eig)%>%t.test())$statistic`), rendering it possible that the previously observed effects were confounded with prior and posterior probability.
+This time, expected information gain had no significant effect on decisions in non-pretend (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='nonpretend')%>%pull(eig)%>%t.test())$statistic`) and pretend games (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='pretend')%>%pull(eig)%>%t.test())$statistic`), rendering it possible that the previously observed effects were confounded with prior and posterior probability.
 In contrast, both prior and posterior probability had a significant positive effect, with prior probability having a stronger effect in pretend games (`r apa_print(E4.model_coefficients_orth %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
@@
 At the end of the experiment, players were asked whether they had a strategy for pretending or for detecting pretense.
-I (MM) manually labeled their responses according to three criteria: mentions of suppression (e.g., "Trying to ignore the hints"), simulation (e.g., "Trying to imagine how I would react"), and rules (e.g., "Checking the corners and center before random selection of tiles").
+I (MM) manually labeled their responses according to three criteria: mentions of suppression (e.g., "I just tried to forget what the word was"), simulation (e.g., " I tried to focus on how I would have deduced the word if I hadn't known it"), and rules (e.g., "I started with vowels in the order we receive them in school").
 Some answers got more than one label and some got none.

### 2024-01-09 docs/preregisteredMethodsAndResults.Rmd
online materials

@@
 E2.performance_df %>%
-  ggplot(aes(x=pretend_rate, y=judge_rate)) +
-  geom_point() +
-  labs(x= 'successful pretense rate', y='judge accuracy') +
-  theme_classic()
+  group_by(judge_rate,pretend_rate)%>%
+  summarise(n=n())%>%
+  ggplot(aes(x=judge_rate,y=pretend_rate,size=n)) +
+  geom_point()+
+  labs(x='judge accuracy',y='pretense success rate') +
+  theme_classic()+
+  theme(legend.pos='na')
+
+ggsave('figures/Battleship2/judge_pretend_accuracy.png',dpi=300)
 
@@
 The research complied with all relevant ethical regulations and was approved by the Research Ethics Committee of Johns Hopkins University.
-500 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
+501 Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
 They were selected based on their acceptance rate (\>95%) and for being native English speakers.
@@
 
-The median game duration in non-pretend games was `r round(E4.median_time$nonpretend%>%median())` seconds in non-pretend games and `r round(E4.median_time$nonpretend%>%median())` seconds in pretend games.
+The median game duration in non-pretend games was `r round(E4.median_time$nonpretend%>%median())` seconds in non-pretend games and `r round(E4.median_time$pretend%>%median())` seconds in pretend games.
 Pretend games were significantly shorter among subjects that pretended in the first block (`r apa_print(E4.median_time%>%filter(!genuine_first)%>%pull(diff)%>%t.test())$statistic`) and in the second block (`r apa_print(E4.median_time%>%filter(genuine_first)%>%pull(diff)%>%t.test())$statistic`).
@@
 ```
+
+### Hypothesis 3 (number of misses)
+
+The mean number of misses was `r E4.median_misses%>%pull(nonpretend)%>%mean()` in non-pretend games and `r E4.median_misses%>%pull(pretend)%>%mean()` in pretend games.
+The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
 
@@
 
-Letter selections preceding a hit were faster than those preceding a miss in non-pretend games only in the group that pretended first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`) but not in the group that played normally first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend)%>%t.test())$full_result`).
-This effect was not observed in pretend games (pretend first: `r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)%>%t.test())$full_result`; non-pretend first: `r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(pretend)%>%t.test())$full_result`).
-We find a significant interaction between these effects in the group that pretended first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(interaction)%>%t.test())$full_result`) but not in the other group (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(interaction)%>%t.test())$full_result`).
+Letter selections preceding a hit were faster than those preceding a miss in non-pretend games only in the group that pretended first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(nonpretend)%>%t.test())$statistic`) but not in the group that played normally first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend)%>%t.test())$statistic`).
+This effect was not observed in pretend games (pretend first: `r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)%>%t.test())$statistic`; non-pretend first: `r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(pretend)%>%t.test())$statistic`).
+We find a significant interaction between these effects in the group that pretended first (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(interaction)%>%t.test())$statistic`) but not in the other group (`r apa_print(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(interaction)%>%t.test())$statistic`).
 The difference was not significant when analyzing first blocks only (`r apa_print(t.test(E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & genuine_first)%>%pull(nonpretend),E4.RT_by_hit_lags_by_condition%>%filter(position==-1 & !genuine_first)%>%pull(pretend)))$statistic`).
@@
 The next analyses were designed to test for differences in game optimality between pretend and non-pretend games, and for a relationship between decision difficulty and click latency.
-To do so, we approximated optimal behaviour by approximating the posterior probability that a letter appears in the word, given available information.
+To do so, we approximated optimal behavior by approximating the posterior probability that a letter appears in the word, given available information.
 Critically, in modelling pretend games we do not treat hints as part of this available information, because an optimal player should ignore this information in choosing where to click next.
@@
 
-### Hypothesis 3 (number of misses)
-
-The mean number of misses was `r E4.median_misses%>%pull(nonpretend)%>%mean()` in non-pretend games and `r E4.median_misses%>%pull(pretend)%>%mean()` in pretend games.
-The difference between the two was significant among players that pretended in the first part (`r t.test(E4.median_misses%>%filter(!genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`), and among those that pretended in the second part (`r t.test(E4.median_misses%>%filter(genuine_first)%>%pull(nonpretend)-E4.median_misses%>%filter(genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).
-
 ```{r simulate_nclicks-hangman, echo=FALSE, cache=TRUE}
@@
 
-Letter selections in pretend games had a higher posterior probability to be included in the word than in random games (pretend first: `r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Letter selections in non-pretend games had a higher posterior probability to be included in the word than in random games (pretend first: `r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(np_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
 The same was true for letter selections in pretend games (pretend first: `r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_r)%>%t.test()%>%apa_print%>%'$'(statistic)`).
-Non-pretend letter selections had a higher posterior probability among players who pretended first (`r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`) but not in the group that played normally first (`r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
+Non-pretend letter selections had a higher posterior probability than pretend letter selections among players who pretended first (`r E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`) but not in the group that played normally first (`r E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
 Comparing data from the first condition only in a between-subject manner revealed a significant difference (`r t.test(E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
@@
 Letters selected by pretenders had a higher overall frequency compared to letters selected by non-pretenders (pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
-but not in the group that played normally first (`r E4.mean_letter_prob_P_click_rank_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`).
 Comparing data from the first condition only in a between-subject manner revealed a significant difference (`r t.test(E4.mean_P_click_rank_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
 
-This was the case also when focusing on misses only ( pretend first: `r E4.mean_letter_prob_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; first block only: `r t.test(E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
+When focusing on misses only, pretend letter selections were again more aligned with the frequency of letters in English words than were non-pretend letter selections ( pretend first: `r E4.mean_letter_prob_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; non-pretend first: `r E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(p_vs_np)%>%t.test()%>%apa_print%>%'$'(statistic)`; first block only: `r t.test(E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%pull(nonpretend),E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print%>%'$'(full_result)`).
 
@@
 
-Participants were not significantly above chance in judging whether the presented game came from a pretender (`r apa_print(t.test(E4.judge_summary_df$accuracy, mu=0.5))$full_result`).
+Participants were not significantly above chance in judging whether the presented game came from a pretender or from a genuine player (`r apa_print(t.test(E4.judge_summary_df$accuracy, mu=0.5))$full_result`).
 This was the case also when focusing on words that judges encountered when pretending (`r apa_print(t.test(E4.judge_summary_df$accuracy_pretend_words, mu=0.5))$full_result`), and when focusing on words that judges encountered when playing normally (`r apa_print(t.test(E4.judge_summary_df$accuracy_nonpretend_words, mu=0.5))$full_result`).
@@
 p <- E4.performance_df %>%
-  ggplot(aes(x=judge_rate,y=pretend_rate)) +
-  geom_jitter() + 
+  group_by(judge_rate,pretend_rate)%>%
+  summarise(n=n())%>%
+  ggplot(aes(x=judge_rate,y=pretend_rate,size=n)) +
   labs(x='judge accuracy',y='pretense success rate') +
-  theme_classic()
+  theme_classic()+
+  theme(legend.pos='na')
 
@@
 E4.performance_df %>%
-  ggplot(aes(x=pretend_rate, y=judge_rate)) +
+  group_by(pretend_rate,judge_rate) %>%
+  summarise(n=n()) %>% 
+  ggplot (aes(x=pretend_rate,y=judge_rate,size=n)) +
   geom_point() +
   labs(x= 'successful pretense rate', y='judge accuracy') +
-  theme_classic()
-
+  theme_classic()+
+  theme(legend.pos='na')
+
+ggsave('figures/Hangman2/judge_pretend_accuracy.png',p,dpi=300)
 ```
@@
 
-### Hypothesis 12 (irrational letter selections)
-
-In non-pretend games, subjects almost always chose letters that made sense (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`).
+### Hypothesis 12 (irrational letter selections in half-games)
+
+In non-pretend games, subjects almost always chose letters that made sense, that is, that were consistent with at least one word that matched the presented category and game state (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`).
 In pretend games, these figures were significantly lower (pretend-first group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='False')%>%pull(makes_sense)%>%mean()%>%printnum()`, $p<.001$; pretend-second group: `r E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & genuine_first=='True')%>%pull(makes_sense)%>%mean()%>%printnum()`, $p<.001$).
@@
 
-### Hypothesis 13 (target word effect)
+### Hypothesis 13 (target word effect in half-games)
 

### 2024-01-10 docs/ExploratoryResults.Rmd
exploratory results
updated path to Hangman data

@@
 
-load('E2.click_log_with_boards.RData')
+load('../experiments/Battleships2/data/batch1/click_log_with_boards.RData')
 
@@
 
-load('E4.click_log.RData')
+load('../experiments/Hangman2/data/E4.click_log.RData')
 
-load('E4.click_log_with_boards.RData')
+load('../experiments/Hangman2/data/E4.click_log_with_boards.RData')
 
-load('E4.judge_df.RData')
+load('../experiments/Hangman2/data/E4.judge_df.RData')
 

### 2024-01-10 docs/ExploratoryResults.Rmd
within-subject variability

@@
 
+## Variability in number of misses
+
+```{r battleship-num-misses-variability}
+
+E2.sd_num_clicks <- E2.click_df %>%
+  filter(!(subj_id %in% E2.lucky_pretenders))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_clicks=sd(num_clicks)) %>%
+  spread(test_part,sd_num_clicks) %>% 
+  mutate(diff=nonpretend-pretend);
+
+E2.sd_num_clicks_relative_to_np <- E2.click_df %>%
+  filter(!(subj_id %in% E2.lucky_pretenders))%>%
+  group_by(subj_id)%>%
+  mutate(mean_nonpretend=mean(num_clicks[test_part=='nonpretend']),
+         squared_dist_to_mean = (num_clicks-mean_nonpretend)**2)%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_clicks=sqrt(sum(squared_dist_to_mean)/(n()-1))) %>%
+  spread(test_part,sd_num_clicks) %>% 
+  mutate(diff=nonpretend-pretend);
+
+```
+
+We quantified the within-subject variability in the number of misses in pretend and non-pretend games.
+This is a function of how similar games were to each other, within condition and within individual.
+The standard deviation in non-pretend games was on average `r E2.sd_num_clicks$nonpretend%>%mean()%>%printnum()`, and significantly higher than in pretend games (mean SD = `r E2.sd_num_clicks$pretend%>%mean()%>%printnum()`, `r apa_print(E2.sd_num_clicks$diff%>%t.test())$statistic`).
+In other words, pretend games were significantly more similar to each other, in number of guesses, than non-pretend games were to each other.
+
+Next, we asked whether pretend games were not only more similar among themselves, but also to the average non-pretend game, than were non-pretend games.
+We therefore replaced the $\mu_{pretend}$ in the standard deviation formula with $\mu_{non-pretend}$: $\sigma'=\sqrt{\frac{1}{N}\Sigma_{i=1}^{N} (x_{i}-\mu_{non-pretend})^2}$.
+Strikingly, we find that a typical pretend game was more similar, in number of misses, **to the average non-pretend game**, than was a typical non-pretend game (`r apa_print(E2.sd_num_clicks_relative_to_np$diff%>%t.test())$full_result`).
+
 ## Within-participant sequential guess entropy
@@
 ```
+
+## Variability in number of misses
+
+```{r hangman-num-misses-variability}
+
+E4.sd_num_misses <- E4.click_df %>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_misses=sd(num_misses)) %>%
+  spread(test_part,sd_num_misses) %>% 
+  mutate(diff=nonpretend-pretend);
+
+E4.sd_num_misses_relative_to_np <- E4.click_df %>%
+  group_by(subj_id)%>%
+  mutate(mean_nonpretend=mean(num_misses[test_part=='nonpretend']),
+         squared_dist_to_mean = (num_misses-mean_nonpretend)**2)%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_clicks=sqrt(sum(squared_dist_to_mean)/(n()-1))) %>%
+  spread(test_part,sd_num_clicks) %>% 
+  mutate(diff=nonpretend-pretend);
+
+E4.sd_num_misses_relative_to_np_controlling_for_word_difficulty <- E4.click_df %>%
+  group_by(word)%>%
+  mutate(num_misses_corrected=num_misses-mean(num_misses[test_part=='nonpretend']))%>%
+  group_by(subj_id)%>%
+  mutate(mean_nonpretend=mean(num_misses_corrected[test_part=='nonpretend']),
+         squared_dist_to_mean = (num_misses_corrected-mean_nonpretend)**2)%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_clicks=sqrt(sum(squared_dist_to_mean)/(n()-1))) %>%
+  spread(test_part,sd_num_clicks) %>% 
+  mutate(diff=nonpretend-pretend);
+
+```
+Similar to Battleship, the within-subject standard deviation in number of misses in non-pretend games (`r E4.sd_num_misses$nonpretend%>%mean()%>%printnum()`) was on average significantly higher than in pretend games (mean SD = `r E4.sd_num_misses$pretend%>%mean()%>%printnum()`, `r apa_print(E4.sd_num_misses$diff%>%t.test())$statistic`).
+Here too, pretend games were significantly more similar to each other, in terms of number of misses, than non-pretend games.
+
+In order to test whether pretend games were more similar to the average non-pretend game than non-pretend games themselves, we first regressed out the effect of word on number of misses, since within-subject differences between the conditions can reflect the random assignment of words to conditions.
+We then replaced the $\mu_{pretend}$ in the standard deviation formula with $\mu_{non-pretend}$: $\sigma'=\sqrt{\frac{1}{N}\Sigma_{i=1}^{N} (x_{i}-\mu_{non-pretend})^2}$, as we did for Battleship.
+We again find that a typical pretend game was more similar, in number of misses (controlling for word difficulty), to the average pretend game, than was a typical non-pretend game (`r apa_print(E4.sd_num_misses_relative_to_np_controlling_for_word_difficulty$diff%>%t.test())$full_result`).
 

### 2024-01-10 docs/ExploratoryResults.Rmd
fixed error in analysis

@@
 
+p<- E2.click_df %>%
+  filter(!(subj_id %in% E2.lucky_pretenders))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_clicks=sd(num_clicks)) %>%
+  group_by(test_part)%>%
+  summarise(
+    mean=mean(sd_num_clicks),
+    se=se(sd_num_clicks)
+    ) %>%
+  ggplot(aes(x=test_part,
+             color=test_part,
+             y=mean))+
+  geom_point(size=10)+
+  geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color='black')+
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+
 ```
@@
 
-Next, we asked whether pretend games were not only more similar among themselves, but also to the average non-pretend game, than were non-pretend games.
+<!-- Next, we asked whether pretend games were not only more similar among themselves, but also to the average non-pretend game, than were non-pretend games.
 We therefore replaced the $\mu_{pretend}$ in the standard deviation formula with $\mu_{non-pretend}$: $\sigma'=\sqrt{\frac{1}{N}\Sigma_{i=1}^{N} (x_{i}-\mu_{non-pretend})^2}$.
-Strikingly, we find that a typical pretend game was more similar, in number of misses, **to the average non-pretend game**, than was a typical non-pretend game (`r apa_print(E2.sd_num_clicks_relative_to_np$diff%>%t.test())$full_result`).
+Here, a typical pretend game was less similar, in number of misses, to the average non-pretend game, than was a typical non-pretend game (`r apa_print(E2.sd_num_clicks_relative_to_np$diff%>%t.test())$full_result`).
+-->
 
@@
   scale_x_continuous(breaks=seq(20))
+
+p<- E2.click_entropy_by_serial_position_summary %>%
+    filter(test_part %in% c('pretend','nonpretend') & click_number<=8) %>%
+    ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
+    geom_line() +
+    geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
+    scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
+    labs(x='click number', y='') +
+    theme_classic() +
+    scale_x_continuous(breaks=seq(20))+
+  theme(legend.pos='na')
+
+ggsave('figures/Battleship2/sequential_guess_entropy.png',width=2.5,height=1.5,dpi=300)
 
@@
 
-In order to test whether pretend games were more similar to the average non-pretend game than non-pretend games themselves, we first regressed out the effect of word on number of misses, since within-subject differences between the conditions can reflect the random assignment of words to conditions.
+<!-- In order to test whether pretend games were more similar to the average non-pretend game than non-pretend games themselves, we first regressed out the effect of word on number of misses, since within-subject differences between the conditions can reflect the random assignment of words to conditions.
 We then replaced the $\mu_{pretend}$ in the standard deviation formula with $\mu_{non-pretend}$: $\sigma'=\sqrt{\frac{1}{N}\Sigma_{i=1}^{N} (x_{i}-\mu_{non-pretend})^2}$, as we did for Battleship.
-We again find that a typical pretend game was more similar, in number of misses (controlling for word difficulty), to the average pretend game, than was a typical non-pretend game (`r apa_print(E4.sd_num_misses_relative_to_np_controlling_for_word_difficulty$diff%>%t.test())$full_result`).
+We again find that a typical pretend game was less similar, in number of misses (controlling for word difficulty), to the average non-pretend game, than was a typical non-pretend game (`r apa_print(E4.sd_num_misses_relative_to_np_controlling_for_word_difficulty$diff%>%t.test())$full_result`).
+-->
 
@@
   scale_x_continuous(breaks=seq(20))
+
+p<- E4.click_entropy_by_serial_position_summary %>%
+    filter(test_part %in% c('pretend','nonpretend') & click_number<=8) %>%
+    ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
+    geom_line() +
+    geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
+    scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
+    labs(x='click number', y='') +
+    theme_classic() +
+    scale_x_continuous(breaks=seq(20))+
+  theme(legend.pos='na')
+
+ggsave('figures/Hangman2/sequential_guess_entropy.png',width=2.5,height=1.5,dpi=300)
 

### 2024-01-17 docs/paper.Rmd
incorporating stereotypicality analysis

@@
 ---
-title             : "Pretending not to know: Pretense behavior reveals the limits of self-simulation"
+title             : "Pretending not to know reveals a powerful capacity for self-simulation"
 shorttitle        : "Pretending not to know"
@@
 Overall, we find a highly impressive capacity for pretending not to know, with pretenders mirroring broad patterns and subtle features of real players' decisions and decision times.
-However, we also find that pretense was characterized by pervasive biases and limitations, although these were undetected by human observers.
+However, we also find that pretense was characterized by over-acting, stereotypical behavior, and suboptimal incorporation of new information, although these were undetected by human observers.
 
@@
 ```{r design, echo=FALSE, fig.cap="Experimental Design in Exp.
-1 (left) and 2 (right).
-In non-pretend games, players revealed ships by guessing cells in a grid (A) or revealed a word by guessing letters (E).
-In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (F), but asked players to play as if they didn't have this information.
-In half-games (C and G), players continued the game from a given state, with the content of some cells and the position of some letters already revealed.
-Lastly, players watched replays of the games of previous players and guessed which were pretend games (D and H).", out.width = '100%'}
-knitr::include_graphics("figures/methods.png")
+1 (upper panel) and 2 (lower panel).
+In non-pretend games, players revealed ships by guessing cells in a grid (A) or revealed a word by guessing letters (D).
+In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (E), but asked players to play as if they didn’t have this information.
+Lastly, players watched replays of the games of previous players and guessed which were pretend games (C and F).", out.width = '100%'}
+knitr::include_graphics("figures/methods2.png")
 ```
@@
 
-## Pretense limitations
+```{r sd_misses, echo=FALSE, cache=TRUE}
+
+sd_misses_df <- E2.click_df %>%
+  dplyr::select(subj_id,test_part, num_clicks) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(sd_misses=sd(num_clicks)) %>%
+  mutate(game='Battleship') %>%
+  rbind(
+    E4.click_df %>%
+  dplyr::select(subj_id,test_part, num_misses) %>%
+  mutate(subj_id=as.character(subj_id)) %>%
+  # rbind(E2.optimal_click_df) %>%
+  # rbind(E2.random_click_df) %>%
+  mutate(test_part = factor(test_part, levels=c(
+    'nonpretend','pretend'), labels = c(
+    'non-pretend','pretend'))) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(sd_misses=sd(num_misses)) %>%
+    mutate(game='Hangman')
+  )
+
+sd_misses_df_wide <- sd_misses_df %>%
+  mutate(test_part=ifelse(test_part=='pretend','p','np'))%>%
+  pivot_wider(values_from=sd_misses,
+              names_from=c(test_part,game))
+
+p <- sd_misses_df %>%
+  mutate(sd_misses = round(sd_misses*2)/2)%>%
+  ggplot(aes(x=sd_misses,fill=test_part, color=test_part))+
+  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
+  scale_x_continuous(limits=c(0,8), breaks=seq(0,10)) +
+  labs(x='median number of misses', y='number of players')+
+  theme(legend.position='none') +
+  theme_classic() +
+  # geom_vline(data=ref_lines, aes(xintercept = xintercept), color='black', linetype=2) +
+  # geom_text(data=ref_lines, aes(x = xintercept, y=y, label=label), angle = 90, vjust = -1, size = 3, inherit.aes = FALSE)+
+  geom_bar(alpha=0.6, position = 'identity', colour='black', size=0.5) +
+  facet_wrap(~game, nrow=2) +
+  theme(legend.position='na',strip.background = element_blank(), panel.border=element_blank())+
+  labs(y='number of players',x="\u03C3(number of misses)")
+
+ggsave('figures/sd_misses_facet.png',p, width=2.5, height=3)
+
+```
+
+## Stereotypical, imperfect self-simulation
 
 Though impressive, the capacity for simulating a state of ignorance was not perfect.
-First, despite showing the same qualitative effects, decision time patterns in Battleship pretend games (but not Hangman pretend games) were systematically more pronounced relative to non-pretend games, as if participants were acting based on an exaggerated mental model of their own counterfactual behavior.
-Furthermore, the within-participant sequential guess entropy, a measure of the inverse predictability of decision sequences across different games of the same player, was significantly reduced in pretend games (Fig.
-\@ref(fig:rigidity)A; see [online supplementary materials](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#within-participant-click-entropy)).
-More specifically, Hangman pretenders were more likely to guess letters that appear frequently in English words (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig.
-\@ref(fig:rigidity)B).
+Importantly, the limitations and biases we observe are consistent with the simulation of a stereotypical, “cartoon” model of decision-making, rather than leakage of concealed information into the decision-making process.
+First, despite showing the same qualitative effects, decision time patterns in Battleship pretend games (but not Hangman pretend games) were systematically more pronounced relative to non-pretend games: a form of “over-acting”.
+Furthermore, pretend games followed stereotypical patterns, and as a result were more homogeneous than non-pretend games.
+Despite a highly similar average number of misses in pretend and non-pretend games (Fig.
+\@ref(fig:decisions)A), the number of unsuccessful guesses was overwhelmingly less variable in pretend relative to non-pretend games (Battleship: sd=`r sd_misses_df_wide$p_Battleship%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Battleship%>%mean(na.rm=T)%>%printnum()` in non-pretend games, $p$ `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Battleship-np_Battleship)%>%pull(diff)%>%t.test())$table$p.value`; Hangman: sd=`r sd_misses_df_wide$p_Hangman%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Hangman%>%mean(na.rm=T)%>%printnum()` in non-pretend games, $p$ `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Hangman-np_Hangman)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+\@ref(fig:rigidity)A).
+Moreover, the within-participant sequential guess entropy, a measure of the inverse predictability of decision sequences across different games of the same player, was significantly reduced in pretend games (Fig.
+\@ref(fig:rigidity)B; see [online supplementary materials](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#within-participant-click-entropy)), consistent with an attempt to enact typical, or average, behavior in a state of ignorance.
+Finally, Hangman pretenders were more likely to guess letters that appear frequently in English words (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig.
+\@ref(fig:rigidity)C).
 This suggests that in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result.
@@
 When asked to reveal the hidden fruit "\_A_A_A", `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('banana', 'papaya'))%>%pull(bananahand)%>%mean())`% of the non-pretenders guessed one of the letters 'B' or 'N' (Fig.
-\@ref(fig:rigidity)C, yellow bars in left column).
+\@ref(fig:rigidity)D, yellow bars in left column).
 Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word== 'banana')%>%pull(bananahand)%>%mean())`%.
@@
 Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', revealing that many pretenders were unable to predict that the first fruit that would have come to their minds was BANANA, not PAPAYA (Fig.
-\@ref(fig:rigidity)C, yellow bars in right column).
+\@ref(fig:rigidity)D, yellow bars in right column).
 A similar pattern was observed for the prototypical body part word HA(ND) and its surprising counterpart HA(IR): when playing normally, `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('hand', 'hair'))%>%pull(bananahand)%>%mean())`% of the players selected letters that are consistent with the prototypical option HAND.
 This figure was `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hand')%>%pull(bananahand)%>%mean())`% among pretenders for whom the target word was HAND, in contrast to only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hair')%>%pull(bananahand)%>%mean())`% among pretenders for whom the target word was HAIR (Fig.
-\@ref(fig:rigidity)C, blue bars).
+\@ref(fig:rigidity)D, blue bars).
 
 ```{r rigidity, echo=FALSE, message= F, warning=F, fig.cap="Limitations on flexible decision-making when pretending.
-A: sequential guess entropy, a measure of the (inverse) predictability of individual players' guesses as a function of click number and guess number.
+A: variability in the number of misses (extracted individually for each player and then averaged) was lower in pretend games.
+B: sequential guess entropy, a measure of the (inverse) predictability of individual players’ guesses as a function of click number and guess number.
 In both Battleship and Hangman, sequential guess entropy increased with click number, and was overall lower in pretend games.
-B: letter frequency of Hangman guesses: the median rank frequency per participant in pretend and non-pretend games, with reference lines for the expected rank frequency for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step.
-C: first letter guesses in Hangman half-games, as a function of pretense condition and target word.
+C: letter frequency of Hangman guesses: the median rank frequency per participant in pretend and non-pretend games, with reference lines for the expected rank frequency for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step.
+D: first letter guesses in Hangman half-games, as a function of pretense condition and target word.
 Letters appear in alphabetical order.
 letters that appear in the target word are marked in yellow (BANANA), red (PAPAYA), blue (HAND) and magenta (HAIR).
-For reference, the overall distribution of letter guesses across all games and conditions is given below.
-", out.width = '100%'}
-knitr::include_graphics("figures/rigidity.png")
+For reference, the overall distribution of letter guesses across all games and conditions is given below.", out.width = '100%'}
+knitr::include_graphics("figures/stereotypical2.png")
 ```
@@
 
-Despite these limitations, pretenders' behavior resembles real players' behavior closely enough that they were able to fool other participants into believing they were playing for real.
+Despite these limitations, pretenders' behavior resembled real players' behavior closely enough that they were able to fool other participants into believing they were playing for real.
 After performing pretend and non-pretend games, participants were presented with game replays of previous players and took on the new role of being a 'judge' who had to determine who was pretending and who was playing for real.
@@
 This would not require self-simulation, only a capacity to intentionally 'unsee', or forget, relevant evidence.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense, for at least four reasons.
+While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense, for at least five reasons.
 First, we tried to make such suppression as hard as possible, by presenting the game solution on top of the game board for the entire duration of pretend games, and by having participants type the target word before pretend Hangman games.
@@
 Third, when asked how they had performed the task in a debrief question, the responses of a significant majority of participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).
-Finally, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig.
+Fourth, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig.
 \@ref(fig:time)A): a finding that is consistent with the simulation of a simplified self-model, but is hard to explain if participants were merely suppressing their current knowledge.
+Finally, pretend games were more similar to each other than were non-pretend games to themselves.
+This is again consistent with an attempt to simulate typical behavior.
 
@@
 The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (``` r``E2.bonus$bonus%>%min() ``` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
+Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
 

### 2024-01-20 docs/paper.Rmd
fixed wrong p value

@@
 `r E2.lucky_pretenders%>%length` pretenders who immediately discovered all ships without making errors were excluded from all further analyses, in accordance with our pre-registered plan.
-With these subjects excluded, the number of cell selections remained very similar in pretend (mean = `r  E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()`; $p=$`r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+With these subjects excluded, the number of cell selections remained very similar in pretend (mean = `r  E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()`; $p=$`r apa_print(E2.median_clicks_filtered%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
 \@ref(fig:decisions)A)).

### 2024-01-20 analysis/cleanDataForSharing.r

@@ -1,0 +1,37 @@
+library(dplyr)
+library(tidyverse)
+
+Battleship_df <- read.csv('../experiments/Battleships2/data/batch1/jatos_results_batch1.csv',na.strings=c(""," ","NA")) %>%
+  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,4),sep='')) %>%
+  # the dash is breaking r
+  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
+  mutate(genuine_first=genuine_first=='True') %>%
+  dplyr::select("subject_identifier", "subj_id","trial_type", "trial_index", "time_elapsed", "internal_node_id",
+                "protocol_sum", "subject_sum",
+                "genuine_first", "choose_pretender", "total_points",
+                "pretend_instructions", "nonpretend_instructions","rt", "responses","correct",
+                "test_part","grid", "click_log", "final_grid_state","num_clicks", "points",
+                "cheat", "grid_number", "replay_log","decision", "genuine_player",
+                "cheat_player", "cheater", "noncheater")
+
+write.table(Battleship_df, file="../experiments/Battleships2/data/batch1/Battleship_for_sharing.csv",quote=TRUE,sep=', ', row.names=FALSE,col.names=TRUE)
+
+Hangman_df <- read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch2.csv',na.strings=c(""," ","NA")) %>%
+  rbind(read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
+  mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,5),sep='')) %>%
+  mutate(subj_id = factor(subj_id)) %>%
+  # the dash is breaking r
+  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part)) %>%
+  filter(PROLIFIC_PID != '5ec4a156bc5aac3819ac52f2' & #reported some of the letters not showing up on their screen
+           PROLIFIC_PID != '58211fc787f6b90001f13f9' & #reported some of the letters not showing up on their screen
+           PROLIFIC_PID != '6149256f6335b06ade3723e0' &  # due a PROLFIIC bug, participated twice
+           PROLIFIC_PID != '615ddab1e4f013092538b6c5') %>%# due a PROLFIIC bug, participated twice
[…]

### 2024-01-22 docs/paper.Rmd
fixed p value typo

@@
 In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(np_vs_r))%>%apa_print%>%'$'(statistic)`; Fig.
+Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`; Fig.
 \@ref(fig:decisions)D).

### 2024-08-22 docs/preregisteredMethodsAndResults.Rmd
paths

@@
 
-# save(E4.click_log, file='E4.click_log.RData')
-
-load('E4.click_log.RData')
+# save(..experiments/Hangman2/data/E4.click_log, file='E4.click_log.RData')
+
+load('..experiments/Hangman2/data/E4.click_log.RData')
 ```
@@
 # E4.letter_prob_click_log_with_boards_optimal,
-# file='E4.click_log_with_boards.RData')
-
-load('E4.click_log_with_boards.RData') # this file includes all the above dataframes.
+# file='../experiments/Hangman2/data/E4.click_log_with_boards.RData')
+
+load('../experiments/Hangman2/data/E4.click_log_with_boards.RData') # this file includes all the above dataframes.
 
@@
 
-load('E4.judge_df.RData')
+load('../experiments/Hangman2/data/E4.judge_df.RData')
 

### 2024-08-22 analysis/JudgmentBasedAnalysis.Rmd
judgment based analysis

---
title: "Judgment Based Analysis"
output: html_document
date: "2024-08-22"
---

```{r analysis-preferences, warning=F, message=F, echo=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning=F, message=F)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel' #for word scatterplots
), '2023-12-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')

papaja::r_refs("r-references.bib")

#color scheme
colors <-  list()
colors$p <- "#404080"
colors$np <- "#69b3a2"
colors$random <- 'gray'
colors$greedy <- 'black'
```

## Construct a subj_id mapping

```{r subj_id_map, warning=F, message=F, echo=FALSE}

E2.subj_mapping <- E2.df %>% 
  group_by(subj_id) %>%
  summarise(jatos_id=subject_identifier[1])

E4.subj_mapping <- E4.df %>% 
  group_by(subj_id) %>%
  summarise(jatos_id=subject_identifier[1])
```

## Organize games according to judgments

```{r games_by_judgments, warning=F, message=F, echo=FALSE}

# Transform the dataframe
E2.judgment <- E2.judge_df %>%
  # Create rows for cheat_player
  mutate(player = cheat_player, judged_pretend = correct, judge=subj_id) %>%
  select(player, grid_number, judged_pretend, judge) %>%
  # Bind rows for genuine_player
  bind_rows(
    E2.judge_df %>%
      mutate(player = genuine_player, judged_pretend = 1 - correct, judge=subj_id) %>%
      select(player, grid_number, judged_pretend, judge)
  ) %>%
  arrange(grid_number, player, judged_pretend)%>%
  rename(jatos_id=player) %>%
  merge(E2.subj_mapping, by='jatos_id')

E4.judgment <- E4.judge_df %>%
  mutate(judged_pretend = ifelse(correct_response=='P',correct,1-correct)) %>%
  select(player,word,judged_pretend, subj_id)%>%
  rename(jatos_id=player,
         judge=subj_id) %>%
  merge(E4.subj_mapping, by='jatos_id')
```

## Merge with main dfs

```{r merge_with_df, warning=F, message=F, echo=FALSE}

E2.click_df_by_judgment <- E2.click_df %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_df_by_judgment <- E4.click_df %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E2.click_log_with_boards_by_judgment <- E2.click_log_with_boards %>% 
  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

E4.click_log_with_boards_by_judgment <- E4.click_log_with_boards %>% 
  merge(E4.judgment, by=c('subj_id','word')) %>%
  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))

```

## Number of misses

```{r num_misses, warning=F, message=F, echo=FALSE}

E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
  group_by(judge,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks)%>%
  mutate(diff=pretend-nonpretend);

p<- E2.click_df_by_judgment %>%
    group_by(judged,judge)%>%
    summarise(num_clicks=median(num_clicks))%>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of judges')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)

E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
  group_by(judge,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks) %>%
  mutate(diff=pretend-nonpretend);

p<- E4.click_df_by_judgment %>%
    group_by(judged,judge)%>%
    summarise(num_clicks=median(num_clicks))%>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='median number of clicks', y='number of judges')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
```

Games with fewer misses are more likely to be categorised as pretend games in Battleship (`r apa_print(t.test(E2.median_clicks_by_judgment%>%pull(diff)))$full_result`) but not in Hangman (`r apa_print(t.test(E4.median_clicks_by_judgment%>%pull(diff)))$full_result`).

```{r merge_with_df, warning=F, message=F, echo=FALSE}

E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
  group_by(subj_id,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks);

E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();

E2.median_clicks_by_judgment_filtered <- E2.click_df_by_judgment %>%
  filter(!(subj_id %in% E2.lucky_pretenders))%>%
  group_by(subj_id,judged, genuine_first) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks);

p<- E2.click_df_by_judgment %>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='number of clicks', y='number of games')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)

E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
  group_by(subj_id,judged) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(judged,num_clicks);

p<- E4.click_df_by_judgment %>%
    ggplot(aes(x=num_clicks,fill=judged))+
    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='number of clicks', y='number of games')+
    theme(legend.position=c(0.10,0.85)) +
    theme_classic()

ggsave('../docs/figures/judgment/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
```

### 2024-08-23 analysis/JudgmentBasedAnalysis.Rmd
RT analysis, accuracy analysis, guesses by word

@@
   # Create rows for cheat_player
-  mutate(player = cheat_player, judged_pretend = correct, judge=subj_id) %>%
-  select(player, grid_number, judged_pretend, judge) %>%
+  mutate(player = cheat_player, judged_pretend = correct, judge_id=subj_id) %>%
+  select(player, grid_number, judged_pretend, judge_id) %>%
   # Bind rows for genuine_player
@@
     E2.judge_df %>%
-      mutate(player = genuine_player, judged_pretend = 1 - correct, judge=subj_id) %>%
-      select(player, grid_number, judged_pretend, judge)
+      mutate(player = genuine_player, judged_pretend = 1 - correct, judge_id=subj_id) %>%
+      select(player, grid_number, judged_pretend, judge_id)
   ) %>%
@@
   rename(jatos_id=player,
-         judge=subj_id) %>%
+         judge_id=subj_id) %>%
   merge(E4.subj_mapping, by='jatos_id')
@@
 
+E2.click_log_by_judgment <- E2.click_log %>% 
+  merge(E2.judgment, by=c('subj_id','grid_number')) %>%
+  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))
+
+E4.click_log_by_judgment <- E4.click_log %>% 
+  merge(E4.judgment, by=c('subj_id','word')) %>%
+  mutate(judged = ifelse(judged_pretend==1,'pretend','nonpretend'))
+
 ```
@@
 E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
-  group_by(judge,judged) %>%
+  group_by(judge_id,judged) %>%
   summarise(num_clicks=median(num_clicks)) %>%
@@
 p<- E2.click_df_by_judgment %>%
-    group_by(judged,judge)%>%
+    group_by(judged,judge_id)%>%
     summarise(num_clicks=median(num_clicks))%>%
@@
 E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
-  group_by(judge,judged) %>%
+  group_by(judge_id,judged) %>%
   summarise(num_clicks=median(num_clicks)) %>%
@@
 p<- E4.click_df_by_judgment %>%
-    group_by(judged,judge)%>%
+    group_by(judged,judge_id)%>%
     summarise(num_clicks=median(num_clicks))%>%
@@
 ```
+## Number of misses, controlling for word
 
-Games with fewer misses are more likely to be categorised as pretend games in Battleship (`r apa_print(t.test(E2.median_clicks_by_judgment%>%pull(diff)))$full_result`) but not in Hangman (`r apa_print(t.test(E4.median_clicks_by_judgment%>%pull(diff)))$full_result`).
+```{r num_misses, warning=F, message=F, echo=FALSE}
 
-```{r merge_with_df, warning=F, message=F, echo=FALSE}
+E4.median_clicks_per_word <- E4.click_df %>%
+  group_by(word)%>%
+  summarise(med_clicks=median(num_clicks))
 
-E2.median_clicks_by_judgment  <- E2.click_df_by_judgment  %>%
-  group_by(subj_id,judged) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(judged,num_clicks);
+E4.corrected_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
+  merge(E4.median_clicks_per_word,by='word')%>%
+  mutate(relative_clicks=num_clicks-med_clicks)%>%
+  group_by(judge_id,judged) %>%
+  summarise(num_clicks=median(relative_clicks)) %>%
+  spread(judged,num_clicks) %>%
+  mutate(diff=pretend-nonpretend);
 
-E2.lucky_pretenders <- E2.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique();
+E4.corrected_clicks_by_judgment_mean  <- E4.click_df_by_judgment  %>%
+  merge(E4.median_clicks_per_word,by='word')%>%
+  mutate(relative_clicks=num_clicks-med_clicks)%>%
+  group_by(judge_id,judged) %>%
+  summarise(num_clicks=mean(relative_clicks)) %>%
+  spread(judged,num_clicks) %>%
+  mutate(diff=pretend-nonpretend);
 
-E2.median_clicks_by_judgment_filtered <- E2.click_df_by_judgment %>%
-  filter(!(subj_id %in% E2.lucky_pretenders))%>%
-  group_by(subj_id,judged, genuine_first) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(judged,num_clicks);
+```
+No effect in Hangman even when controlling for average number of clicks per word (`r APA_print(E4.corrected_clicks_by_judgment$diff%>%t.test())$statistic`), but a weak effect in the opposite direction (such that longer games are judged to be pretend games) when using mean instead of median (which might actually make sense; `r APA_print(E4.corrected_clicks_by_judgment$diff%>%t.test())$statistic`).
 
-p<- E2.click_df_by_judgment %>%
-    ggplot(aes(x=num_clicks,fill=judged))+
-    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
+## Misses per word
+
+```{r num_misses, warning=F, message=F, echo=FALSE}
+
+E4.clicks_by_word_by_judgment  <- E4.click_df_by_judgment  %>%
+  group_by(word,judged) %>%
+  summarise(num_clicks=mean(num_clicks)) %>%
+  spread(judged,num_clicks) %>%
+  mutate(diff=pretend-nonpretend);
+
+E4.clicks_by_word_by_judgment %>%
+  ggplot(aes(x=nonpretend,y=pretend, label=word))+
+  geom_label()+
+  geom_abline(slope=1,intercept=0)+
+  labs(x='judged non-pretend',y='judged pretend')
+```
+
+The same effect to what we see when categorising based on true pretend/non-pretend, such that people are more likely to judge unusually short or long games as nonpretend games.
+
+## Game duration
+
+```{r duration, warning=F, message=F, echo=FALSE}
+
+E2.duration_by_judgment <- E2.click_log_with_boards_by_judgment %>% 
+  group_by(judge_id,judged,grid_number) %>%
+  summarise(duration=sum(RT)) %>% 
+  group_by(judge_id,judged) %>% 
+  summarise(duration=median(duration)) %>% 
+  spread(judged,duration) %>%
+  mutate(diff=pretend-nonpretend)
+
+E4.duration_by_judgment <- E4.click_log_with_boards_by_judgment %>% 
+  group_by(judge_id,judged,word) %>%
+  summarise(duration=sum(RT)) %>% 
+  group_by(judge_id,judged) %>% 
+  summarise(duration=median(duration)) %>% 
+  spread(judged,duration) %>%
+  mutate(diff=pretend-nonpretend)
+
+```
+Long and short games were similarly likely to be categorised as pretend games in Battleship (`r apa_print(t.test(E2.duration_by_judgment%>%pull(diff)))$full_result`) and in Hangman (`r apa_print(t.test(E4.duration_by_judgment%>%pull(diff)))$full_result`).
+
+```{r RT_by_accuracy}
+
+N_perm <- 1000;
+bootstrap_error <- function(x, N_perm) {
+  N <- length(x)
+  medians = c();
+  for (i in 1:N_perm) {
+    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median(., na.rm=T))
+  };
+  return(sd(medians))
+}
+
+E2.RT_by_lag1_by_judgment <- E2.click_log_by_judgment %>%
+  filter(click_number>1) %>%
+  group_by(judge_id,judged, lag1)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag1,RT,sep='_') %>%
+  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
+
+E2.RT_by_lag1_by_judgment_wide <- E2.RT_by_lag1_by_judgment %>%
+  dplyr::select(judge_id,judged,lag1_diff) %>%
+  spread(judged,lag1_diff)%>%
+  mutate(diff=pretend-nonpretend)
+
+E2.RT_by_hit_by_judgment <- E2.click_log_by_judgment %>%
+  filter(click_number>0) %>%
+  group_by(judge_id,judged,hit_bin)%>%
+  summarise(RT=median(RT))  %>%
+  spread(hit_bin,RT,sep='_') %>%
+  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
+
+E2.RT_by_hit_by_judgment_wide <- E2.RT_by_hit_by_judgment %>%
+  dplyr::select(judge_id,judged,hit_diff) %>%
+  spread(judged,hit_diff)%>%
+  mutate(diff=pretend-nonpretend)
+
+E2.RT_by_lead1_by_judgment <- E2.click_log_by_judgment %>%
+  group_by(judge_id,judged, lead1)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead1,RT,sep='_') %>%
+  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
+
+E2.RT_by_lead1_by_judgment_wide <- E2.RT_by_lead1_by_judgment %>%
+  dplyr::select(judge_id,judged,lead1_diff) %>%
+  spread(judged,lead1_diff)%>%
+  mutate(diff=pretend-nonpretend)
+
+E2.RT_by_hit_lags_by_judgment <-E2.RT_by_lag1_by_judgment %>%
+  merge(E2.RT_by_hit_by_judgment) %>%
+  merge(E2.RT_by_lead1_by_judgment)
+
+E2.RT_by_hit_lags_long_by_judgment <- E2.RT_by_hit_lags_by_judgment %>%
+  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
+  mutate(position = ifelse(position=="lag1_diff",1,
+                                         ifelse(position=="hit_diff",0,-1))) %>%
+  dplyr::select(judge_id,judged,position,difference);
+
+p<- E2.RT_by_hit_lags_long_by_judgment %>% 
+  filter(position<2 & position > -2)%>%
+  group_by(judged,position) %>%
+  summarise(median_difference=median(difference),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=judged,fill=judged,y=median_difference)) +
   scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='number of clicks', y='number of games')+
-    theme(legend.position=c(0.10,0.85)) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
   theme_classic()
 
-ggsave('../docs/figures/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
+ggsave('figures/judgment/E2_RT_by_lag.png',p,width=4, height=3.5, dpi=300)
 
-E4.median_clicks_by_judgment  <- E4.click_df_by_judgment  %>%
-  group_by(subj_id,judged) %>%
-  summarise(num_clicks=median(num_clicks)) %>%
-  spread(judged,num_clicks);
+p
 
-p<- E4.click_df_by_judgment %>%
-    ggplot(aes(x=num_clicks,fill=judged))+
-    geom_bar( color="#e9ecef", alpha=0.6, position = 'identity') +
+E4.RT_by_lag1_by_judgment <- E4.click_log_by_judgment %>%
+  filter(click_number>1) %>%
+  group_by(judge_id,judged, lag1)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lag1,RT,sep='_') %>%
+  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)
+
+E4.RT_by_lag1_by_judgment_wide <- E4.RT_by_lag1_by_judgment %>%
+  dplyr::select(judge_id,judged,lag1_diff) %>%
+  spread(judged,lag1_diff)%>%
+  mutate(diff=pretend-nonpretend)
+
+E4.RT_by_hit_by_judgment <- E4.click_log_by_judgment %>%
+  filter(click_number>0) %>%
+  group_by(judge_id,judged,hit_bin)%>%
+  summarise(RT=median(RT))  %>%
+  spread(hit_bin,RT,sep='_') %>%
+  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)
+
+E4.RT_by_hit_by_judgment_wide <- E4.RT_by_hit_by_judgment %>%
+  dplyr::select(judge_id,judged,hit_diff) %>%
+  spread(judged,hit_diff)%>%
+  mutate(diff=pretend-nonpretend)
+
+E4.RT_by_lead1_by_judgment <- E4.click_log_by_judgment %>%
+  group_by(judge_id,judged, lead1)%>%
+  summarise(RT=median(RT))  %>%
+  spread(lead1,RT,sep='_') %>%
+  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);
+
+E4.RT_by_lead1_by_judgment_wide <- E4.RT_by_lead1_by_judgment %>%
+  dplyr::select(judge_id,judged,lead1_diff) %>%
+  spread(judged,lead1_diff)%>%
+  mutate(diff=pretend-nonpretend)
+
+E4.RT_by_hit_lags_by_judgment <-E4.RT_by_lag1_by_judgment %>%
+  merge(E4.RT_by_hit_by_judgment) %>%
+  merge(E4.RT_by_lead1_by_judgment)
+
+E4.RT_by_hit_lags_long_by_judgment <- E4.RT_by_hit_lags_by_judgment %>%
+  gather("position","difference",c("lag1_diff", "hit_diff", "lead1_diff"))%>%
+  mutate(position = ifelse(position=="lag1_diff",1,
+                                         ifelse(position=="hit_diff",0,-1))) %>%
+  dplyr::select(judge_id,judged,position,difference);
+
+p<- E4.RT_by_hit_lags_long_by_judgment %>% 
+  filter(position<2 & position > -2)%>%
+  group_by(judged,position) %>%
+  summarise(median_difference=median(difference, na.rm=T),
+            sem_difference=bootstrap_error(difference,N_perm)) %>%
+  ggplot(aes(x=position,color=judged,fill=judged,y=median_difference)) +
   scale_fill_manual(values=c("#69b3a2", "#404080")) +
-    labs(x='number of clicks', y='number of games')+
-    theme(legend.position=c(0.10,0.85)) +
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  scale_x_continuous(breaks=-1:1,labels=c('hit on next trial', 'hit on current trial', 'hit on previous trial'))+
+  geom_abline(slope=0,intercept=0,linetype=3)+
+  geom_path(size=1)+
+  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
+  labs(fill=' ',x='', y='RT effect (ms) of hit on trial t')+
   theme_classic()
 
-ggsave('../docs/figures/judgment/judgment/E2_num_clicks.png',p,width=5,height=2, dpi=300)
+ggsave('figures/judgment/E4_RT_by_lag.png',p,width=4, height=3.5, dpi=300)
+
+p
+
 ```
+
+No difference in the effect of current hit on RT between games that were judged to be pretend vs.
+non-pretend (`r apa_print(E2.RT_by_hit_by_judgment_wide$diff%>%t.test())$full_result`) in Battleship.
+Similar in Hangman (`r apa_print(E4.RT_by_hit_by_judgment_wide$diff%>%t.test())$full_result`).
+
+No difference in the effect of previous guess outcome on RT between games that were judged to be pretend vs.
+non-pretend (`r apa_print(E2.RT_by_lag1_by_judgment_wide$diff%>%t.test())$full_result`) in Battleship.
+Similar in Hangman (`r apa_print(E4.RT_by_lag1_by_judgment_wide$diff%>%t.test())$full_result`).
+
+No difference in the effect of next guess outcome on RT between games that were judged to be pretend vs.
+non-pretend (`r apa_print(E2.RT_by_lead1_by_judgment_wide$diff%>%t.test())$full_result`) in Battleship.
+Similar in Hangman (`r apa_print(E4.RT_by_lead1_by_judgment_wide$diff%>%t.test())$full_result`).
+
+### Model based analysis
+
+```{r battleship-p_click, echo = FALSE, cache=TRUE}
+
+E2.mean_P_click_rank_by_judgment <- E2.click_log_with_boards_by_judgment %>%
+  group_by(judge_id,judged) %>%
+  summarise(p_click_rank=mean(p_click_rank));
+
+E2.mean_P_click_rank_misses_by_judgment <- E2.click_log_with_boards_by_judgment %>%
+  rowwise() %>%
+  filter(!hit_bin) %>%
+  group_by(judge_id,judged) %>%
+  summarise(p_click_rank=mean(p_click_rank));
+
+p<- E2.mean_P_click_rank_by_judgment %>%
+    ggplot(aes(x=p_click_rank,fill=judged))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+ggsave('figures/judgment/E2_pclick_rank.pdf',p)
+ggsave('figures/judgement/E2_pclick_rank.png',p,width=5,height=2,dpi=300)
+
+p<- E2.mean_P_click_rank_misses_by_judgment %>%
+    ggplot(aes(x=p_click_rank,fill=judged))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725")) +
+    labs(x='p(ship in clicked square) - rank', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    # facet_wrap(~genuine_first, nrow=2)+
+    # scale_x_reverse() +
+    theme_classic();
+
+ggsave('figures/judgment/E2_pclick_rank_misses.pdf',p)
+ggsave('figures/judgement/E2_pclick_rank_misses.png',p,width=5,height=2,dpi=300)
+
+E2.num_irrational_clicks_by_judgment <- E2.click_log_with_boards_by_judgment %>%
+  group_by(judge_id,judged) %>%
+  summarise(nic=sum(p_click==0)/5);
+
+p<- E2.num_irrational_clicks_by_judgment %>%
+    ggplot(aes(x=nic,fill=judged))+
+    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
+    scale_fill_manual(values=c("#69b3a2", "#404080")) +
+    labs(x='mean number of irrational clicks per game', y='number of players')+
+    theme(legend.position=c(0.10,0.85)) +
+    theme_classic();
+
+ggsave('figures/judgement/E2_num_irrational_clicks.png',p,width=5,height=3,dpi=300)
+
+```
+
+Seems like people are judging irrational games as pretend much more.
+
+### Accuracy
+
+```{r accuracy, echo = FALSE, cache=TRUE}
+
+E2.judgment_summary <- E2.click_log_by_judgment %>% 
+  group_by(subj_id,grid_number) %>%
+  summarise(judged=as.factor(judged[1]),test_part=as.factor(test_part[1]))
+
+table(E2.judgment_summary$test_part,E2.judgment_summary$judged)
+
+E4.judgment_summary <- E4.click_log_by_judgment %>% 
+  group_by(subj_id,word) %>%
+  summarise(judged=as.factor(judged[1]),test_part=as.factor(test_part[1]))
+
+table(E4.judgment_summary$test_part,E4.judgment_summary$judged)
+
+```

### 2025-08-31 analysis/testPartConditionInteraction.Rmd
Psych Science revisions

---
title             : "Interaction between test part and condition"
shorttitle        : "Interaction between test part and condition"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : word_document
---

```{r global_options, include=FALSE}
set.seed(42)
knitr::opts_chunk$set(fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T, 
                      fig.width = 10, fig.height = 5)
```

```{r, libraries, echo=FALSE, warning=FALSE, cache=F}

library(groundhog)

groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel', #for word scatterplots
  'cocor' # differences between correlations
), '2025-06-01')

```

```{r, load data}
# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')
```

In both experiments, the order of pretend and non-pretend blocks was counterbalanced between participants. Here we ask whether differences between pretend and non-pretend effects are stronger in the first part of the experiment compared to the second part of the experiment. Overall, we find no significant interaction effects in any of our measures. Marginally significant effects appear for Hangman (a reduction of the effect of pretense on the number of misses and on click optimality between the first and the second parts), and in Battleship (a reduction of the effect of pretense on the effect of guess accuracy on decision time between the first and second parts).

## Number of clicks

### Battleship:

No interaction between test part (first/second) and condition (pretend/non-pretend) on number of clicks:

`r apa_print(t.test(E2.median_clicks_filtered %>% filter(genuine_first) %>% mutate(mean = (pretend+nonpretend)/2)%>%pull(mean),E2.median_clicks_filtered %>% filter(!genuine_first) %>% mutate(mean = (pretend+nonpretend)/2)%>%pull(mean)))$full_result`

### Hangman:

No interaction between test part (first/second) and condition (pretend/non-pretend) on number of misses:

`r apa_print(t.test(E4.median_misses %>% filter(genuine_first) %>% mutate(mean = (pretend+nonpretend)/2)%>%pull(mean),E4.median_misses %>% filter(!genuine_first) %>% mutate(mean = (pretend+nonpretend)/2)%>%pull(mean)))$full_result`

A marginally significant effect was driven by a descriptively stronger tendency to underestimate the number of misses in the first part of the experiment (a contrast between pretend and non-pretend from the first part only: `r apa_print(t.test(E4.median_misses %>% filter(genuine_first) %>% mutate(mean = (nonpretend))%>%pull(mean),E4.median_misses %>% filter(!genuine_first) %>% mutate(mean = (pretend))%>%pull(mean)))$estimate`)) compared to the second part of the experiment (a contrast between pretend and non-pretend from the second part only: `r apa_print(t.test(E4.median_misses %>% filter(!genuine_first) %>% mutate(mean = (nonpretend))%>%pull(mean),E4.median_misses %>% filter(genuine_first) %>% mutate(mean = (pretend))%>%pull(mean)))$estimate`)).

## Correlation in number of guesses: Hangman

```{r, correlations in misses}

E4.misses_per_word_gf <- E4.click_df %>%
  filter(genuine_first) %>%
  group_by(test_part,word) %>%
  summarise(num_misses=median(num_misses)) %>%
  spread(test_part,num_misses)

E4.misses_per_word_gs <- E4.click_df %>%
  filter(!genuine_first) %>%
  group_by(test_part,word) %>%
  summarise(num_misses=median(num_misses)) %>%
  spread(test_part,num_misses)
```

The Spearman correlation between the number of misses in pretend games and non-pretend games was `r cor(E4.misses_per_word_gf$pretend, E4.misses_per_word_gf$nonpretend,method='spearman')%>%printnum()` in the first part and `r cor(E4.misses_per_word_gs$pretend, E4.misses_per_word_gs$nonpretend,method='spearman')%>%printnum()` in the second part. The difference between the two correlations was not statistically significant in a Fisher z test (p>0.19).

## Click optimality

### Battleship:

No interaction between test part (first/second) and condition (pretend/non-pretend) on click optimality:

`r apa_print(t.test(E2.mean_P_click_rank%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E2.mean_P_click_rank%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

### Hangman:

No interaction between test part (first/second) and condition (pretend/non-pretend) on click optimality:

`r apa_print(t.test(E4.mean_P_click_rank%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E4.mean_P_click_rank%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

A marginally significant effect was driven by a descriptively bigger optimality cost for pretending in the first part of the experiment (a contrast between pretend and non-pretend from the first part only: `r apa_print(t.test(E4.mean_P_click_rank%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%pull(nonpretend),E4.mean_P_click_rank%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%pull(pretend)))$estimate`)) compared to the second part of the experiment (a contrast between pretend and non-pretend from the second part only: `r apa_print(t.test(E4.mean_P_click_rank%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%pull(nonpretend),E4.mean_P_click_rank%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%pull(pretend)))$estimate`)).

## Click optimality: misses only

### Battleship:

No interaction between test part (first/second) and condition (pretend/non-pretend) on click optimality when restricting the analysis to misses only:

`r apa_print(t.test(E2.mean_P_click_rank_misses_only%>%filter(genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E2.mean_P_click_rank_misses_only%>%filter(!genuine_first)%>%spread(test_part,p_click_rank)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

### Hangman:

No interaction between test part (first/second) and condition (pretend/non-pretend) on click optimality when restricting the analysis to misses only:

`r apa_print(t.test(E4.mean_P_click_rank_misses_wide%>%filter(genuine_first)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E4.mean_P_click_rank_misses_wide%>%filter(!genuine_first)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

## Effect of ship completion: Battleship

```{r pneighbour, echo=FALSE, message= F, warning=F}
source('../analysis/pneighbour.R')
```

No interaction between test part (first/second) and condition (pretend/non-pretend) on the effect of sinking the submarine on the probability of checking whether a size-2 patrol boat is a submarine:

`r apa_print(t.test(E2.pneighbour_wide%>%filter(genuine_first_nonpretend)%>%mutate(mean=(diff_pretend+diff_nonpretend)/2)%>%pull(mean),E2.pneighbour_wide%>%filter(!genuine_first_nonpretend)%>%mutate(mean=(diff_pretend+diff_nonpretend)/2)%>%pull(mean)))$full_result`

## Effect of guess accuracy on decision time

### Battleship:
#### Previous guess:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the effect of the accuracy of the previous guess on the time taken to make the next one:

`r apa_print(t.test(E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

A marginally significant effect was driven by a descriptively stronger effect in the first part of the experiment (a contrast between pretend and non-pretend from the first part only: `r apa_print(t.test(E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%pull(nonpretend),E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%pull(pretend)))$estimate`) compared to the second part of the experiment (`r apa_print(t.test(E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%pull(nonpretend),E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%pull(pretend)))$estimate`). Here, the effect in the second part of the experiment was not significant (`r apa_print(t.test(E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%pull(nonpretend),E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%pull(pretend)))$statistic`).

#### Current guess:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the difference in RT between accurate and inaccurate guesses:

`r apa_print(t.test(E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

A marginally significant effect was driven by a descriptively stronger effect in the first part of the experiment (a contrast between pretend and non-pretend from the first part only: `r apa_print(t.test(E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%pull(nonpretend),E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%pull(pretend)))$estimate`) compared to the second part of the experiment (`r apa_print(t.test(E2.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%pull(nonpretend),E2.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%pull(pretend)))$estimate`).

### Hangman:
#### Previous guess:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the effect of the accuracy of the previous guess on the time taken to make the next one:

`r apa_print(t.test(E4.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E4.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,lag1_diff,test_part)%>%spread(test_part,lag1_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

#### Current guess:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the difference in RT between accurate and inaccurate guesses:

`r apa_print(t.test(E4.RT_by_hit_lags %>% filter(genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E4.RT_by_hit_lags %>% filter(!genuine_first) %>% dplyr::select(subj_id,hit_diff,test_part)%>%spread(test_part,hit_diff)%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

## Effect of uncertainty on decision time

### Battleship:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the quadratic effect of uncertainty on decision time:

`r apa_print(t.test(E2.RT_by_entropy_coefs_wide %>% filter(genuine_first & term=='entropy_squared')%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E2.RT_by_entropy_coefs_wide %>% filter(!genuine_first & term=='entropy_squared')%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

### Hangman:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the quadratic effect of uncertainty on decision time:

`r apa_print(t.test(E4.RT_by_entropy_coefs_wide %>% filter(genuine_first & term=='entropy_squared')%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E4.RT_by_entropy_coefs_wide %>% filter(!genuine_first & term=='entropy_squared')%>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

## Standard deviation of number of misses

```{r sd_misses, echo=FALSE, cache=TRUE}

sd_misses_df <- E2.click_df %>%
  dplyr::select(genuine_first, subj_id,test_part, num_clicks) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,genuine_first, test_part) %>%
  summarise(sd_misses=sd(num_clicks)) %>%
  mutate(game='Battleship') %>%
  rbind(
    E4.click_df %>%
  dplyr::select(genuine_first, subj_id,test_part, num_misses) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,genuine_first, test_part) %>%
  summarise(sd_misses=sd(num_misses)) %>%
    mutate(game='Hangman')
  )
```

### Battleship:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the inter-game variability in the number of misses:

`r apa_print(t.test(sd_misses_df %>% filter(genuine_first&game=='Battleship')%>%group_by(subj_id)%>%summarise(mean=mean(sd_misses))%>%pull(mean),sd_misses_df %>% filter(!genuine_first&game=='Battleship')%>%group_by(subj_id)%>%summarise(mean=mean(sd_misses))%>%pull(mean)))$full_result`

### Hangman:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the inter-game variability in the number of misses:

`r apa_print(t.test(sd_misses_df %>% filter(genuine_first&game=='Hangman')%>%group_by(subj_id)%>%summarise(mean=mean(sd_misses))%>%pull(mean),sd_misses_df %>% filter(!genuine_first&game=='Hangman')%>%group_by(subj_id)%>%summarise(mean=mean(sd_misses))%>%pull(mean)))$full_result`

## Shannon entropy of decisions

### Battleship:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the Shannon entropy of participants' guesses (averaged across guesses 1-6):

`r apa_print(t.test(E2.entropy_summary %>% spread(test_part,entropy) %>% merge(E2.first_click_log%>%dplyr::select(subj_id,genuine_first)) %>% filter(genuine_first) %>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E2.entropy_summary %>% spread(test_part,entropy) %>% merge(E2.first_click_log%>%dplyr::select(subj_id,genuine_first)) %>% filter(!genuine_first) %>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

### Hangman:

No interaction between test part (first/second) and condition (pretend/non-pretend) on the Shannon entropy of participants' guesses (averaged across guesses 1-6):

`r apa_print(t.test(E4.entropy_summary %>% spread(test_part,entropy) %>% merge(E4.median_clicks%>%dplyr::select(subj_id,genuine_first)) %>% filter(genuine_first) %>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean),E4.entropy_summary %>% spread(test_part,entropy) %>% merge(E4.median_clicks%>%dplyr::select(subj_id,genuine_first)) %>% filter(!genuine_first) %>%mutate(mean=(pretend+nonpretend)/2)%>%pull(mean)))$full_result`

### 2025-08-31 docs/ExploratoryResults.Rmd
Psych Science revisions

@@
 classoption       : "man"
-output            : bookdown::gitbook
+output            : word_document
+
 bibliography: references.bib
@@
   'ggrepel' #for word scatterplots
-), '2022-12-01')
+), '2025-06-01')
 
@@
 p<- E2.click_df %>%
+  mutate(test_part=factor(test_part,labels=c('non-pretend','pretend')))%>%
   filter(!(subj_id %in% E2.lucky_pretenders))%>%
@@
              y=mean))+
-  geom_point(size=10)+
-  geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color='black')+
+  geom_point(size=6)+
+  geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color='black', width=0.3)+
   scale_color_manual(values=c("#69b3a2", "#404080")) +
+  labs(x="",y="")+
+  theme(legend.pos='na')+
+  theme_classic()+
+  scale_y_continuous(limits=c(0,2.8))+
+  theme(legend.pos='na')
+
+ggsave('figures/Battleship2/num_guess_sd.png',p,width=2.5,height=1.5,dpi=300)
 
@@
 
-E2.debrief_labeled <- read.csv('..\\experiments\\Battleships2\\data\\debrief.csv',na.strings=c(""," ","NA")) %>%
+E2.debrief_labeled <- read.csv('../experiments/Battleships2/data/debrief.csv',na.strings=c(""," ","NA")) %>%
   dplyr::select(-subj_id) %>%
@@
 
-E4.df <- read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch2.csv',na.strings=c(""," ","NA")) %>%
-  rbind(read.csv('..\\experiments\\Hangman2\\data\\jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
+E4.df <- read.csv('../experiments/Hangman2/data/jatos_results_batch2.csv',na.strings=c(""," ","NA")) %>%
+  rbind(read.csv('../experiments/Hangman2/data/jatos_results_batch1.csv',na.strings=c(""," ","NA"))) %>%
   mutate(subj_id = paste(as.character(participant_number),substr(PROLIFIC_PID,1,5),sep='')) %>%
@@
 
-```
+p<- E4.click_df %>%
+  mutate(test_part=factor(test_part,levels=c('nonpretend','pretend'),labels=c('non-pretend','pretend')))%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(sd_num_clicks=sd(num_misses)) %>%
+  group_by(test_part)%>%
+  summarise(
+    mean=mean(sd_num_clicks),
+    se=se(sd_num_clicks)
+    ) %>%
+  ggplot(aes(x=test_part,
+             color=test_part,
+             y=mean))+
+  geom_point(size=6)+
+  geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color='black', width=0.3)+
+  scale_color_manual(values=c("#69b3a2", "#404080")) +
+  labs(x="",y="")+
+  theme(legend.pos='na')+
+  theme_classic()+
+  scale_y_continuous(limits=c(0,2.8))+
+  theme(legend.pos='na')
+
+ggsave('figures/Hangman2/num_guess_sd.png',p,width=2.5,height=1.5,dpi=300)
+
+```
+
 Similar to Battleship, the within-subject standard deviation in number of misses in non-pretend games (`r E4.sd_num_misses$nonpretend%>%mean()%>%printnum()`) was on average significantly higher than in pretend games (mean SD = `r E4.sd_num_misses$pretend%>%mean()%>%printnum()`, `r apa_print(E4.sd_num_misses$diff%>%t.test())$statistic`).
@@
 
-E4.debrief_labeled <- read.csv('..\\experiments\\Hangman2\\data\\debrief.csv',na.strings=c(""," ","NA"))
+E4.debrief_labeled <- read.csv('../experiments/Hangman2/data/debrief.csv',na.strings=c(""," ","NA"))
 
@@
 
+### The letter frequency heuristic
+
+In this analysis, we asked whether pretenders are doing more than guessing frequent letters such as vowels.
+We therefore compared behaviour in pretend and non-pretend games to the behaviour of an agent that selects the most frequent letter among the ones that were not selected so far.
+By focusing our analysis on misses only (that is, selections of letters that are not in the target word), we ensured that any observed advantage in real players' behaviour is not simply due to making more correct guesses.
+
+```{r hangman-letter-freuqency}
+
+E4.misses_with_simulations <- E4.click_log_with_boards %>%
+  filter(!hit_bin)%>%
+  rbind(E4.click_log_with_boards_random %>%
+          filter(!hit_bin)) %>%
+  rbind(E4.click_log_with_boards_most_frequent %>%
+          filter(!hit_bin))
+
+E4.nic <- E4.misses_with_simulations%>%
+  group_by(subj_id,test_part, genuine_first) %>%
+  summarise(nic=sum(p_click>0)/n()*100)
+
+E4.nic %>%
+  group_by(test_part) %>%
+  mutate(test_part=factor(test_part,
+                          levels=c('random','nonpretend','pretend','most_frequent'),
+                          labels=c('random','non-pretend','pretend','most frequent'))) %>%
+  summarise(p_irrational=mean(nic),
+            sep_irrational=se(nic)) %>%
+  ggplot(aes(x=test_part,y=p_irrational, fill=test_part))+
+  geom_bar(stat="identity", color="black",
+           position=position_dodge()) +
+  geom_errorbar(aes(ymin=p_irrational-sep_irrational, ymax=p_irrational+sep_irrational), width=.2,
+                position=position_dodge(.9)) +
+  labs(x='',y='% errors that make sense') +
+  scale_fill_manual(values=c("#FDE725", "#69b3a2", "#404080", "#E88567")) +
+  theme_classic()+
+  theme(legend.position='na')
+
+ggsave('figures/forPresentation/hangman_rational.png',width=5,height=3,dpi=300)
+
+```
+
+Our dependent measure was the proportion of guesses that "make sense", that is, that are consistent with the board state and the word category.
+Pretend guesses made more sense according to this measure than what we would expect if people were only guessing the most frequent letters (difference in percentage of sensible errors between pretend and heuristic behaviour: `r apa_print(E4.nic %>% spread(test_part,nic) %>% mutate(diff=pretend- most_frequent)%>%pull(diff)%>%t.test())$full_result`).
+
 \newpage

### 2025-08-31 docs/paper_asAccepted.Rmd
Psych Science revisions

---
title             : "Pretending not to know reveals a capacity for model-based self-simulation"
shorttitle        : "Pretending not to know"

author: 
  - name          : "Matan Mazor"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "All Souls College, High Street, Oxford OX1 4AL"
    email         : "matan.mazor@all-souls.ox.ac.uk"
  #   role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
  #     - Conceptualization
  #     - Writing - Original Draft Preparation
  #     - Writing - Review & Editing

  - name          : "Chaz Firestone"
    affiliation   : "2"

  - name          : "Ian Phillips"
    affiliation   : "2"
    # role:
    #   - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "University of Oxford"
  - id            : "2"
    institution   : "Johns Hopkins University"

abstract: |
  Pretending not to know requires appreciating how one would behave without a given piece of knowledge, and acting accordingly. Here, two game-based experiments reveal a capacity to simulate decision-making under such counterfactual ignorance. 1001 English-speaking adults saw the solution to a game (ship locations in Battleship, the hidden word in Hangman) but attempted to play as though they never had this information. Pretenders accurately mimicked broad aspects of genuine play, including the number of guesses required to reach a solution, as well as subtle patterns, such as effects of decision uncertainty on decision time. While peers were unable to detect pretense, statistical analysis and computational modeling uncovered traces of ‘over-acting’ in pretenders’ decisions, suggesting a schematic simulation of their minds. Opening up a new approach to studying self-simulation, our results reveal intricate metacognitive knowledge about decision-making, drawn from a rich—but simplified—internal model of cognition.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "pretense; metacognition; theory of mind"
wordcount         : "7745"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word

bibliography: references.bib
---

```{r analysis-preferences, warning=F, message=F, echo=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning=F, message=F)

# Load packages with groundhog
library(groundhog)
groundhog.library(c(
  'png',
  'grid',
  'ggplot2',
  'svglite',
  'xtable',
  'papaja',
  'tidyverse',
  'broom',
  'cowplot',
  'MESS', # for AUCs
  'lsr', # for effect sizes
  'pwr', # for power calculations
  'brms', # for mixed effects modeling
  'BayesFactor', # for Bayesian t test
  'jsonlite', # parsing data from sort_trial
  'caret', #for cross validation
  'ggrepel' #for word scatterplots
), '2025-06-01')

# Load workspace (after running preregisteredMethodsAndResults.Rmd, 
# exploratoryResults.Rmd, and all the scripts that are linked to from these 
# documents). 
load('../.RData')

papaja::r_refs("references.bib")

#color scheme
colors <-  list()
colors$p <- "#404080"
colors$np <- "#69b3a2"
colors$random <- 'gray'
colors$greedy <- 'black'
```

# Research Transparency Statement

## General Disclosures

The authors declare no conflicts of interest. Funding: This study was supported by an NSF BCS grant #2021053 awarded to C.F. M.M is supported by a post-doctoral research fellowship at All Souls College. Artificial intelligence: No artificial intelligence assisted technologies were used in this research or the creation of this article. Ethics: The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.

## Experiment 1 (Battleship)

Preregistration: The hypotheses and methods were preregistered on 2021-12–21, prior to data collection, and following the analysis of an independent pilot sample. A detailed pre-registration can be accessed at <https://osf.io/v9zsb>. The pre-registration was time-locked using cryptographic randomisation-based time-locking (Mazor, Mazor, & Mukamel, 2019) (protocol sum: 60c270410375e8a192468fc1a0e9c93da60d5e203eb2760b621a8631a26f4c5c; [link to relevant lines in experimental code](https://github.com/self-model/pretendingNotToKnow/blob/44ef964c0d7d4a8f3b76ec52316a8e624c41a047/experiments/Battleships2/code/index.html#L865-L876), making experimental randomisation causally dependent on the content of the pre-registration and thus ensuring that all data were collected after pre-registration. All pre-registered analyses are publicly available, including the report-generating R script (<https://self-model.github.io/pretendingNotToKnow/docs/exp.-1-battleship.html>). Exploratory analyses are flagged as such. Materials: All study materials, including demos of analysis experiments, are publicly available (<https://github.com/self-model/pretendingNotToKnow>). Data: All primary data are publicly available (<https://osf.io/zma9b>). Analysis scripts: All analysis scripts are publicly available (<https://osf.io/zma9b>).

## Experiment 2 (Hangman)

Preregistration: The hypotheses and methods were preregistered on 2022-06–22, prior to data collection, and following the analysis of an independent pilot sample. A detailed pre-registration can be accessed at osf.io/3thry. The pre-registration was time-locked using cryptographic randomisation-based time-locking, making experimental randomisation causally dependent on the content of the pre-registration and thus ensuring that all data were collected after pre-registration [@mazor2019]. Due to an error in the experiment code, time-locking took effect only from player number 221 (batch 3) and on (protocol sum: c4929c7fe33df1b7b52f15c789d98eab30a9cee09a8121807a3c59e28e7430a4;[relevant lines in experimental code](https://github.com/self-model/pretendingNotToKnow/blob/f98903f3f2d0e21800497dc65f8d5f823c910c88/experiments/Hangman2/code/index.html#L772-L783). All pre-registered analyses are publicly available, including the report-generating R script (<https://self-model.github.io/pretendingNotToKnow/docs/exp.-2-hangman.html>). Exploratory analyses are flagged as such. Materials: All study materials, including demos of the experiments, are publicly available (<https://github.com/self-model/pretendingNotToKnow>). Data: All primary data are publicly available (<https://osf.io/zma9b>). Analysis scripts: All analysis scripts are publicly available (<https://osf.io/zma9b>).

# Introduction

<!-- The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory]. This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation]. This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own. -->

Pretense relies on an ability to simulate and mimic one's own behavior under a counterfactual belief state. For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("*I know that they are planning a surprise party, but they don't know that I know that...*") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("*Where would I look first had I not known? What would I say? How long would it take me to recover from the surprise?*"). Similar examples abound in higher-stakes contexts such as diplomacy, warcraft and law. This is not a trivial challenge: previous research on "hindsight biases" suggests that knowledge about the actual state of the world can interfere with our ability to correctly judge what we would have believed [@fischhoff1975; @fischhoff1977; @wood1978; @roese2012] or perceived [@harley2004; @bernstein2007; @bernstein2012] without this knowledge. Such biases remain potent even when instructing participants to overcome them [@harley2004; @pohl1996]. Moreover, even if pretenders can correctly determine what they would have believed, they must further accurately simulate how they would think and behave in this different belief state.

The reliance of this kind of epistemic pretense on self-simulation makes it a promising tool for revealing the structure and content of people's internal models of their own minds. When directly asked, participants are able to provide relatively accurate descriptions of their own decision-making [@morris] and perception [@levin2008; @mazor2023]. Pretending not to know opens a new window into the structure and content of this metacognitive knowledge, with two important advantages. First, by not relying on explicit reports, pretense has the potential to reveal implicit self-knowledge -- that is, structured knowledge about the self that is not reportable. And second, data obtained from pretense experiments can be analyzed and modeled using the same tools employed by cognitive scientists to study non-pretense behavior, affording a direct and finer-grained comparison between pretend and genuine decision-making.

Our research question is whether people can mentally simulate their actions under a counterfactual knowledge state of ignorance. To that end, we had participants “pretend not to know” in a game setting. Using an online version of the games Battleship and Hangman (in which players seek to uncover the locations of enemy ships or the identity of a word), participants played a ‘non-pretend’ (normal) version of the game, as well as a ‘pretend’ version where they were given complete information about the hidden ships / target word but were instructed to behave as if they didn’t have this information. Participants’ pretense behaviour mirrored broad patterns and subtle features of real players’ decisions and decision times. At the same time, epistemic pretense was characterized by over-acting, stereotypical behavior, and suboptimal incorporation of new information: all markers of model-based simulation. Together, we take these findings as evidence for a capacity to mentally simulate decisions and actions using a simplified and schematic self-model.

<!-- to what extent humans are capable of generating complex behavior that is aligned with a counterfactual belief state, and specifically with a counterfactual state of ignorance. -->

# Method

The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University. In two experiments, online participants played online versions of two information-seeking games. In Battleship (Exp. 1), `r E2.df$subj_id%>%unique()%>%length()` English-speaking players (recruited from Prolific.com) were presented with a 5x5 grid of yellow squares, and attempted to reveal one size-3 submarine and two size-2 patrol boats with as few guesses as possible. In our version of the game, ships could only touch corner-to-corner, but not side-to-side (this was explained to participants before playing), and participants were not notified once they had sunk a ship (only whether their guess was a hit or a miss). In Hangman, N=`r E4.df$subj_id%>%unique()%>%length()` English-speaking players attempted to reveal a hidden word, name or number (hereafter referred to broadly as "word') with as few letter-guesses as possible, based on word length and category (a famous person, number, fruit, US state, or body part). To ensure familiarity with US states, Hangman participants were all US-based.

Both games traditionally start in a state of ignorance, with a player’s goal being to reveal an unknown world state (ship locations in Battleship, a hidden word in Hangman) in as few steps (cell or letter selections) as possible. Critically, in addition to playing five standard games, players in our experiments also completed five ‘pretend’ games in which the solution to the game was known to them from the start and remained visible on the screen throughout the entire game (pretend-Battleship ship locations were marked with a cross, pretend-Hangman words were presented visually and had to be typed by players before the game, to ensure encoding; see Fig. \@ref(fig:design)). In these games, players’ task was to behave as if they were playing for real – i.e. to play as though they did not have this information.

Throughout the game, participants accrued points that were later converted to a monetary bonus. In non-pretend games, participants received points for revealing the ships, or the target word, with as few guesses as possible. In pretend games, participants were given different instructions:

> “In this round, we’re going to tell you where the ships are, but we want you to act like you don’t know this information. We’ve marked the ships’ locations with a cross, so you’ll know where they are the whole time; but your job is to play the game as if these hints aren’t there. To see how good you are at this, we’re going to compare your games to the games of people who actually had no hints, and see how similar they are. We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you’ll get bonus points. But if your games look different, you won’t get these bonus points. Your number of clicks in this part will not affect your bonus. Only your ability to play like you had no hints.”

And in Hangman:

> “In this round, we’re going to tell you the word in advance, but we want you to act like you don’t know this information. To see how good you are at this, we’re going to compare your games to the games of people who played normally, without knowing what the word was, and see how similar they are. We will measure which letters you click and the timing of your guesses; if your clicks look similar to people who played like normal (trying to reveal the word with as few guesses as possible, but without any hints), you’ll get bonus points. But if your games look different, you won’t get these bonus points. Your number of clicks in this part will not affect your bonus. Only your ability to play like you didn’t see the word in advance.”

We intentionally included no reference to an observer in these instructions, to have participants focusing on simulating their own behaviour rather than simulating how their behaviour would be perceived by another person. In reality, participants’ games were presented to other participants, and they received bonus points if they tricked these other participants into believing they did not have hints.

Players played pretend and standard games in separate blocks that were presented in random order after a first ‘practice’ game. In principle, participants could learn about their own behaviour from this practice game. To minimize such learning effects, we distinguished practice games from the main experimental blocks, using a smaller 4x4 grid with only two size-2 ships in Battleship, and a word category (animals) that was not used in the main experiment in Hangman. Each experimental block was followed by a half-game, where players were instructed to complete the game from a half-finished state. Finally, players were presented with replays of the games of previous players, and judged which were standard and which were pretend games. We measured players’ capacity to simulate a counterfactual state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games. Our full pre-registered results are available [online](https://self-model.github.io/pretendingNotToKnow/docs/exp.-1-battleship.html) together with the report-generating code. Unless otherwise specified, all reported findings similarly hold when analysing only the first condition performed by each subject in a between-subject analysis, thereby ensuring that findings are not driven by learning effects^[In both experiments, the order of pretend and non-pretend blocks was counterbalanced between participants. We observed no significant interaction between the strength of any of our effects (i.e., differences between pretend and non-pretend conditions) and part of the experiment (i.e., first versus second). For full details see Supplementary Materials.]. Readers are invited to try [demos of the experiments](https://self-model.github.io/pretendingNotToKnow/experiments/demos/pretend).

```{r design, echo=FALSE, fig.cap="Experimental Design in Exp. 1 (upper panel) and 2 (lower panel). In non-pretend games, players revealed ships by guessing cells in a grid (A) or revealed a word by guessing letters (D). In pretend games, we marked ship locations with a cross (B) and revealed the target word from the start (E), but asked players to play as if they didn’t have this information. Lastly, players watched replays of the games of previous players and guessed which were pretend games (C and F).", out.width = '100%'}
knitr::include_graphics("figures/methods2.png")
```

# Results

<!-- In two experiments, players played online versions of two information-seeking games: Battleship (N=`r E2.df$subj_id%>%unique()%>%length()` players) and Hangman (N=`r E4.df$subj_id%>%unique()%>%length()` players). These games traditionally start in a state of ignorance, with a player's goal being to reveal an unknown world state (ship locations in Battleship, a hidden word in Hangman) in as few steps (cell or letter selections) as possible. Critically, in addition to playing five standard games, players in our experiments also completed five 'pretend' games in which the solution to the game was known to them from the start, but where their task was to behave as if it was not -- i.e. to play as though they did not have this information (see Fig. @ref(fig:design)). We measured participants' capacity to simulate a counterfactual state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games. Our full pre-registered results are available [online](https://matanmazor.github.io/pretendingNotToKnow/docs/exp.-1-battleship.html) together with the report-generating code. Readers are invited to try [demos of the experiments](https://matanmazor.github.io/pretendingNotToKnow/experiments/demos/pretend). -->

## Measuring pretense quality

As a first measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games. Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()%>%printnum()`, SD = `r E2.median_clicks%>%pull(pretend)%>%sd()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()%>%printnum()`, SD = `r E2.median_clicks%>%pull(nonpretend)%>%sd()%>%printnum()`; `r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$statistic`; Cohen's $d$ = `r E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`; Fig. \@ref(fig:decisions)A). `r E2.lucky_pretenders%>%length` pretenders who immediately discovered all ships without making errors were excluded from all further analyses, in accordance with our pre-registered plan. With these subjects excluded, the number of cell selections remained very similar in pretend (mean = `r  E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()`; `r apa_print(E2.median_clicks_filtered%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$statistic`; Fig. \@ref(fig:decisions)A). In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()%>%printnum()` misses, SD = `r E4.median_misses%>%pull(pretend)%>%sd()%>%printnum()`; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()%>%printnum()` misses, , SD = `r E4.median_misses%>%pull(nonpretend)%>%sd()%>%printnum()`; `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$statistic`; Cohen's $d$ = `r E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`; Fig. \@ref(fig:decisions)B). Despite an overall bias in the number of guesses, pretend Hangman games showed a near-perfect item-specific alignment: pretenders were successful in making more incorrect letter guesses when attempting to reveal words that would have been harder to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`; Fig. \@ref(fig:decisions)B). This strong correlation provides evidence for a human capacity to act in accordance with a counterfactual knowledge state.

(ref:decisions-caption) Battleship and Hangman guesses in pretend and non-pretend games. A: median number of misses in Battleship and Hangman games, in non-pretend (green) and pretend (purple) games. For reference, the expected number of misses is indicated by a reference line for a fully random agent, and for a "greedy" agent that maximizes the probability of a hit in each step. B: The median number of misses in Hangman for pretend and non-pretend games, as a function of the target word. C: Spatial guess distributions for pretend and non-pretend half-games (where players continued the game from a half-finished state) alongside their corresponding hit probability maps. D: Cell and letter selections were ranked according to their relative hit probability given the players’ knowledge at the time of making the decision (dynamically updated after each guess). Plotting the median rank per subject in pretend and non-pretend games, with reference lines for the expected rank probability for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step. Note that the expected rank for a greedy agent is greater than 1 because there was not always a single optimal choice. E: same as panel D but discarding all guesses that resulted in a hit.

```{r decisions, echo=FALSE, message= F, warning=F, fig.cap="(ref:decisions-caption)", out.width = '100%'}

num_misses_df <- E2.click_df %>%
  dplyr::select(subj_id,test_part, num_clicks) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,test_part) %>%
  summarise(num_misses=median(num_clicks)-7) %>%
  mutate(game='Battleship') %>%
  rbind(
    E4.click_df %>%
  dplyr::select(subj_id,test_part, num_misses) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,test_part) %>%
  summarise(num_misses=median(num_misses)) %>%
    mutate(game='Hangman')
  )
load('../analysis/simulated_hangman_games.RData')

E2.random_mean_misses <- E2.random_click_df %>%
           group_by(subj_id) %>%
           summarise(num_misses=median(num_clicks)-7) %>%
           pull(num_misses) %>%
                  mean()

E2.greedy_mean_misses <- E2.optimal_click_df %>%
           group_by(subj_id) %>%
           summarise(num_misses=median(num_clicks)-7) %>%
           pull(num_misses) %>%
                  mean()

E4.random_mean_misses <- simulated_random_hangman_games %>% 
  group_by(subj_id) %>%
  summarise(num_misses=median(nmisses)) %>%
  pull(num_misses) %>%
  mean()

E4.greedy_mean_misses <- simulated_optimal_hangman_games %>% 
  pull(nmisses) %>%
  mean()

ref_lines <- data.frame(
  game = c('Battleship','Battleship','Hangman','Hangman'),
  xintercept = c(E2.random_mean_misses, E2.greedy_mean_misses, 
                 E4.random_mean_misses, E4.greedy_mean_misses),
  label = c('Random','Greedy','Random','Greedy'),
  y = c(75,75,75,115)
)

p <- num_misses_df %>%
  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
  labs(x='median number of misses', y='number of players')+
  theme(legend.position='none') +
  theme_classic() +
  geom_vline(data=ref_lines, aes(xintercept = xintercept), color='black', linetype=2) +
  geom_text(data=ref_lines, aes(x = xintercept, y=y, label=label), angle = 90, vjust = -1, size = 3, inherit.aes = FALSE)+
  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
  facet_wrap(~game, nrow=2) +
  theme(legend.position='na',strip.background = element_blank(), panel.border=element_blank())

ggsave('figures/num_misses_facet.png',p, width=4, height=3)

p<-  E2.click_df %>%
  dplyr::select(subj_id,test_part, num_clicks) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,test_part) %>%
  summarise(num_misses=median(num_clicks)-7) %>%
  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
  labs(x='median number of misses', y='number of players')+
  theme(legend.position='none') +
  theme_classic() +
  geom_vline(xintercept = E2.greedy_mean_misses, color='black', linetype=2) +
  annotate("text", x = E2.greedy_mean_misses, y = 75, label = "Greedy", angle = 90, vjust = -1, size = 5)+
  geom_vline(xintercept = E2.random_mean_misses, color='black', linetype=2) +
  annotate("text", x = E2.random_mean_misses, y = 75, label = "Random", angle = 90, vjust = -1, size = 5)+
  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
  theme(legend.pos='na')

ggsave('figures/num_misses_E2.png',p, width=4.5, height=2.5)

p <- E4.click_df %>%
  dplyr::select(subj_id,test_part, num_misses) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,test_part) %>%
  summarise(num_misses=median(num_misses)) %>%
  ggplot(aes(x=num_misses,fill=test_part, color=test_part))+
  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
  scale_x_continuous(limits=c(-1,19), breaks=seq(0,18)) +
  labs(x='median number of misses', y='number of players')+
  theme(legend.position='none') +
  theme_classic() +
  geom_vline(xintercept = E4.greedy_mean_misses, color='black', linetype=2) +
  annotate("text", x = E4.greedy_mean_misses, y = 95, label = "Greedy", angle = 90, vjust = -1, size = 5)+
  geom_vline(xintercept = E4.random_mean_misses, color='black', linetype=2) +
  annotate("text", x = E4.random_mean_misses, y = 95, label = "Random", angle = 90, vjust = -1, size = 5)+
  geom_bar(alpha=0.6, position = 'identity', color='black', size=0.5) +
  theme(legend.pos='na')

ggsave('figures/num_misses_E4.png', p, width=4.5, height=2.5)
knitr::include_graphics("figures/decisions.png")

```

Having established an alignment in the total number of guesses, we next turned to the content of pretend and non-pretend guesses. In order to directly compare pretend and non-pretend guesses for the same board state, Battleship players completed two half-games in which they were instructed to continue the game from a half-completed state. In standard games, players start in the same (blank) board state but quickly diverge as they make different guess sequences. Including half games allowed us access to hundreds of cell selections for the same board state from pretenders and non-pretenders. This way, we had sufficient statistical power to compare the two guess distributions. We find a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_A_pretend_nonpretend)$table$p.value`; board B: $r=$ `r apa_print(cor_B_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_B_pretend_nonpretend)$table$p.value`; Fig. \@ref(fig:decisions)C), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.

To further examine the decisional processes behind this strong alignment, we compared the degree to which pretend and non-pretend guesses made sense within the context of the game. When playing Battleship and Hangman, it makes sense to guess cells or letters for which the probability of hitting a ship or revealing a letter is high [this "greedy" behavior is not strictly optimal, but approximates optimal behavior in most cases, @audinot2014optimal]. To this end, we ranked cells based on the Bayesian probability of a hit given players’ knowledge at the time of making the decision. Critically, hit probability maps were dynamically updated after each guess. In Battleship, this model assumed that all legal board configurations are equally likely a priori, but board configurations were ruled out as the game progressed and the content of individual cells was revealed. In Hangman, we used the category information (e.g., ‘a fruit’), to obtain a probability-weighted list of category-compatible words (or names, in the case of famous people). We relied on prototypicality norms [@uyeda1980prototypicality] for words, and number of visits to Wikipedia entries for famous people. The full prior distributions for each category were included in the pre-registration (for details, see Supplementary Materials). Similar to Battleship, in deriving hit probability we assume access to the full list of options that is consistent with the game state (the number of hidden letters, the revealed letters and their positions, and the list of letters that do not appear in the game solution) at the time of making the decision.

In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`). Pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`, Cohen's $d$ = `r cohensD(-E4.mean_P_click_rank_wide%>%pull(p_vs_r))%>%printnum()`), but significantly less rational than non-pretend guesses (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(p_vs_np))%>%apa_print%>%'$'(statistic)`, Cohen's $d$ = `r cohensD(-E4.mean_P_click_rank_wide%>%pull(p_vs_np))%>%printnum()` Fig. \@ref(fig:decisions)D). Critically, pretend guesses were more rational than random guesses even when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r t.test(-E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`, Cohen's $d$ = `r cohensD(-E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r))%>%printnum()`; Fig. \@ref(fig:decisions)E): that is, even when incorrectly guessing a ship's location or a letter's identity, pretend guesses made sense given the limited information players pretended to have.

```{r pneighbour, echo=FALSE, message= F, warning=F}
source('../analysis/pneighbour.R')
```

A specific example of this effect in the game of Battleship can be observed in players’ behaviour immediately after hitting the last cell of a size-2 patrol boat (players attempted to reveal two size-2 patrol boats and one size-3 submarine). Among non-pretenders, the next cell selection was often directed at checking whether the two cells were part of the size-3 submarine, but this was only true if the size-3 submarine had not been sunk yet (`r round(E2.pneighbour_nonpretend$second%>%mean(na.rm=T)*100)`% of all cell selections), and not when it had been sunk (`r round(E2.pneighbour_nonpretend$secondlast%>%mean(na.rm=T)*100)`% of all cell selections, and significantly lower than `r round(E2.pneighbour_nonpretend$second%>%mean(na.rm=T)*100)`%: `r apa_print(t.test(E2.pneighbour_nonpretend$diff))$statistic`, Cohen’s $d$ = `r printnum(cohensD(E2.pneighbour_nonpretend$diff))`). Despite knowing with full certainty that the size-2 patrol boat was not a size-3 submarine, pretenders showed the same qualitative pattern: pretending to check if the revealed cells were part of a size-3 submarine only when they pretended not to know that it was fully sunk (`r round(E2.pneighbour_pretend$second%>%mean(na.rm=T)*100)`% of all cell selections), but not when the size-3 submarine had been sunk (`r round(E2.pneighbour_pretend$secondlast%>%mean(na.rm=T)*100)`% of all cell selections, `r apa_print(t.test(E2.pneighbour_pretend$diff))$statistic`, Cohen’s $d$ = `r printnum(cohensD(E2.pneighbour_pretend$diff))`). The tendency to check if the two cells were part of a bigger ship was weaker among pretenders (`r apa_print(t.test(E2.pneighbour_wide$diff_second))$statistic`, Cohen’s $d$ = `r printnum(cohensD(E2.pneighbour_wide$diff_second))`).

Good pretense is a function not only of the number and content of players' decisions, but also of their timing. Here too, pretend games showed the same qualitative patterns as non-pretend games. Like non-pretenders, pretenders were faster in their successful guesses (difference in decision time between hits and misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms) and slowed down immediately after a hit (difference in decision time between guesses that followed hits versus misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms; Fig. \@ref(fig:time)A). All effects are significant at the $0.001$ level with the pre-registered within-subject t-test, except for the post-hit slowing down in Battleship, which, due to outliers with extreme effects in the opposite direction (\>10 s), was only significant in a non-parametric Wilcoxon sign-rank test (`r apa_print(E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%wilcox.test())$statistic`). Effects remain significant at the $0.001$ level when statistically controlling for the serial position of guesses within the game.

We also examined the effect of decision uncertainty, quantified as the Shannon entropy of the posterior distribution over cell or letter options, on decision times. To this end we fitted subject-level linear models, predicting reaction times from the linear and quadratic expansions of decision entropy, and contrasted the coefficients against zero in a group-level t-test. In the non-pretend versions of the games, the quadratic coefficients were significantly negative, with the slowest responses associated with mid-range levels of entropy (see Fig. \@ref(fig:time)B; Battleship: `r apa_print(E2.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%t.test())$statistic`, Cohen's $d$ = `r E2.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%cohensD()%>%printnum()`; Hangman: `r apa_print(E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%t.test())$statistic`, Cohen's $d$ = `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%cohensD()%>%printnum()`. When restricting the analysis to those Battleship players who pretended after playing normally, this effect was significant only in a Wilcoxon rank-sum test, due to outliers in the sample: `r apa_print(E2.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared' & genuine_first)%>%pull(nonpretend)%>%wilcox.test())$statistic`). Critically, the quadratic coefficients were significantly negative also in pretend games (see Fig. \@ref(fig:time)B; Battleship: `r apa_print(E2.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(pretend)%>%t.test())$statistic`, Cohen's $d$ = `r E2.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(pretend)%>%cohensD()%>%printnum()`; Hangman: `r apa_print(E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(pretend)%>%t.test())$statistic`, Cohen's $d$ = `r E4.RT_by_entropy_coefs_wide %>% filter(term=='entropy_squared')%>%pull(pretend)%>%cohensD()%>%printnum()`). In other words, despite knowing the game's solution with full certainty, pretenders successfully feigned subtle qualitative effects of counterfactual uncertainty on their decision times.

```{r time, echo=FALSE, message= F, warning=F, fig.cap="Patterns of decision time in pretend and non-pretend games. A: median decision times for hits and misses, as well as the decisions following them. In both Battleship and Hangman, hits were faster on average than misses, but guesses following a hit were slower on average than those following a miss. This pattern was mimicked in pretend games. B: median decision times as a function of decision uncertainty, quantified as the entropy of the posterior over guess options. In both Hangman and Battleship, guesses were slowest for mid-range levels of entropy, and this pattern was mimicked in pretend games. Shaded areas represent the bootstrapped standard error of the median.", out.width = '100%'}
knitr::include_graphics("figures/time3.png")
```

```{r sd_misses, echo=FALSE, cache=TRUE}

sd_misses_df <- E2.click_df %>%
  dplyr::select(genuine_first, subj_id,test_part, num_clicks) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,genuine_first, test_part) %>%
  summarise(sd_misses=sd(num_clicks)) %>%
  mutate(game='Battleship') %>%
  rbind(
    E4.click_df %>%
  dplyr::select(genuine_first, subj_id,test_part, num_misses) %>%
  mutate(subj_id=as.character(subj_id)) %>%
  # rbind(E2.optimal_click_df) %>%
  # rbind(E2.random_click_df) %>%
  mutate(test_part = factor(test_part, levels=c(
    'nonpretend','pretend'), labels = c(
    'non-pretend','pretend'))) %>%
  group_by(subj_id,genuine_first, test_part) %>%
  summarise(sd_misses=sd(num_misses)) %>%
    mutate(game='Hangman')
  )

sd_misses_df_wide <- sd_misses_df %>%
  mutate(test_part=ifelse(test_part=='pretend','p','np'))%>%
  pivot_wider(values_from=sd_misses,
              names_from=c(test_part,game))

p <- sd_misses_df %>%
  mutate(sd_misses = round(sd_misses*2)/2)%>%
  ggplot(aes(x=sd_misses,fill=test_part, color=test_part))+
  scale_fill_manual(values=c(colors$np,colors$p, colors$random, colors$greedy)) +
  scale_x_continuous(limits=c(0,8), breaks=seq(0,10)) +
  labs(x='median number of misses', y='number of players')+
  theme(legend.position='none') +
  theme_classic() +
  # geom_vline(data=ref_lines, aes(xintercept = xintercept), color='black', linetype=2) +
  # geom_text(data=ref_lines, aes(x = xintercept, y=y, label=label), angle = 90, vjust = -1, size = 3, inherit.aes = FALSE)+
  geom_bar(alpha=0.6, position = 'identity', colour='black', size=0.5) +
  facet_wrap(~game, nrow=2) +
  theme(legend.position='na',strip.background = element_blank(), panel.border=element_blank())+
  labs(y='number of players',x="\u03C3(number of misses)")

ggsave('figures/sd_misses_facet.png',p, width=2.5, height=3)

```

## Stereotypical, imperfect self-simulation

Though impressive, the capacity for simulating a state of ignorance was not perfect. Importantly, the limitations and biases we observe are consistent with the simulation of a stereotypical, “cartoon” model of decision-making, rather than leakage of concealed information into the decision-making process as would be expected if pretenders’ success was due to efficient, but imperfect, suppression of their knowledge of the game solution. First, despite showing the same qualitative effects, decision time patterns in Battleship pretend games (but not Hangman pretend games) were systematically more pronounced relative to non-pretend games: a form of “over-acting”. Specifically, the difference in reaction times as a function of guess outcome (Fig.3A) was larger in pretend games, both when measured with respect to the current guess (`r E2.RT_by_hit_lags_long %>% filter(position==0)%>%spread(test_part,difference)%>%mutate(diff=-pretend+nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.RT_by_hit_lags_long %>% filter(position==0)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`), and with respect to the following guess (`r E2.RT_by_hit_lags_long %>% filter(position==1)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.RT_by_hit_lags_long %>% filter(position==1)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`). Similarly, the quadratic effect of decision entropy on decision times was stronger in pretend games (`r E2.RT_by_entropy_coefs%>%filter( term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=-pretend+nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.RT_by_entropy_coefs%>%filter( term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`).

Furthermore, pretend games followed stereotypical patterns, and as a result were more homogeneous than non-pretend games. Despite a highly similar average number of misses in pretend and non-pretend games (Fig. \@ref(fig:decisions)A), the number of unsuccessful guesses was overwhelmingly less variable in pretend relative to non-pretend games (Battleship: sd=`r sd_misses_df_wide$p_Battleship%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Battleship%>%mean(na.rm=T)%>%printnum()` in non-pretend games, `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Battleship-np_Battleship)%>%pull(diff)%>%t.test())$statistic`, Cohen's $d$ = `r printnum(sd_misses_df_wide%>%mutate(diff=p_Battleship-np_Battleship)%>%pull(diff)%>%cohensD())`; Hangman: sd=`r sd_misses_df_wide$p_Hangman%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Hangman%>%mean(na.rm=T)%>%printnum()` in non-pretend games, `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Hangman-np_Hangman)%>%pull(diff)%>%t.test())$statistic`, Cohen's $d$ = `r printnum(sd_misses_df_wide%>%mutate(diff=p_Hangman-np_Hangman)%>%pull(diff)%>%cohensD())`; Fig. \@ref(fig:rigidity)A). Moreover, while pretenders produced more letter misses for harder words (Fig. 2B), they underestimated the difficulty of the very hard “DALAI LAMA” and overestimated the difficulty of the easy number (“ELEVEN” and “NINETY SIX”) and state (“MONTANA” and “IOWA”) words. That is, pretenders consistently enacted what they saw as a ‘typical’ or a ‘representative’ game, one that is not unusual in the number of lucky or unlucky guesses. This is again consistent with shrinkage toward the mean of a generative self-model [@mazor2021; @jansen2021], with an attempt to avoid extreme outcomes to appear convincing to a hypothetical observer [@oey2023], and with representativeness skewing intuitions about randomness [@kahneman1972].

```{r entropy, echo=FALSE, message= F, warning=F}
source('../analysis/clickEntropyBattleship.R')
source('../analysis/clickEntropyHangman.R')
```

Next, we examined variability not in the number of guesses, but in their contents. We separately computed the Shannon entropy of the guess distribution across different games for each player, condition (pretend or non-pretend), and serial guess number. High entropy then corresponds to pronounced variability in the guess sequences of different games, and low entropy to a tendency to repeat the same sequence of guesses in different games. For example, if a player always starts their games by clicking in the top left corner, their guess entropy for the first click will be H([1,1,1,1,1])=0. Unsurprisingly, the within-participant sequential guess entropy increased as a function of guess number, consistent with players adjusting their behaviour in light of the outcomes of previous guesses, making individual games increasingly more varied (Fig. \@ref(fig:rigidity)B). If pretend games were a similar but noisier version of standard games, their associated guess entropy would be higher, reflecting the additional noise in the decision-making process, or the game-specific biases that are associated with the suppression of specific words or game states. Critically, however, entropy was systematically reduced in pretend games (p\<0.001 for a within-subject t-test of guess entropy in guesses number 1-4 in both Battleship and Hangman, see Supplementary Materials for guess-specific statistics). Thus, when pretending, participants produced similar guess sequences across different games. In Hangman, for example, this meant that non-pretenders more flexibly adjusted their first letter guess to the word category and number of letters, compared to pretenders, who tended to open the game with the same letter guess irrespective of the specific game state. This seems consistent with an attempt to enact what they saw as typical, representative, or average behaviour. In contrast, a reduction in the guess sequence entropy is inconsistent with leakage of suppressed knowledge into the decision-making process, as would be expected if differences between pretend and non-pretend games reflected the imperfect suppression of the game’s solution.

One possible account of the reduced decision entropy in pretend games is that it reflects pretenders’ miscalibrated intuitions about randomness, conforming to a prototype of randomness that is itself too ordered. If the same prototype of randomness is used by pretenders to determine the number of unsuccessful guesses per game, the two measures should be correlated across participants. Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$full_result`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$full_result`). This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$full_result`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$full_result`). We interpret this as evidence that the reduction in variance reflects miscalibrated beliefs not only about randomness, but also about participants’ behaviour under a counterfactual knowledge state. Those players who thought they would strictly follow a particular sequence of guesses (low entropy), ended up producing games of more variable lengths, as their success depended more on luck. Other players adjusted their decision strategy more flexibly, perhaps attempting to produce games that are not too long or short, in line with their intuitions about randomness.

Finally, Hangman pretenders were more likely to guess letters that appear frequently in English words (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig. \@ref(fig:rigidity)C). This suggests that in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result (but see Supplementary Materials for evidence that heuristic use alone cannot fully explain pretenders' behaviour).

This limitation on incorporating evidence into the (simulated) decision-making process was especially evident in Hangman half-games, where players completed the game from a half-completed state. When asked to reveal the hidden fruit "\_A_A_A", `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('banana', 'papaya'))%>%pull(bananahand)%>%mean())`% of the non-pretenders guessed one of the letters 'B' or 'N' (Fig. \@ref(fig:rigidity)D, yellow bars in left column). Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word== 'banana')%>%pull(bananahand)%>%mean())`% (this drop was significant in a binomial test: $p=$`r apa_print(binom.test(E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='banana')%>%pull(bananahand)%>%sum(),E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='banana')%>%nrow(),E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word =='banana')%>%pull(bananahand)%>%mean(na.rm=T)))$table$p.value`). Importantly, half of the pretenders were given different information: they were told that the hidden word was the less prototypical fruit PAPAYA. Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', with many guessing letters that are not consistent with either PAPAYA or BANANA (gray bars in Fig. 4D), revealing that many pretenders had the knowledge that PAPAYA would be a hard fruit to guess, therefore avoiding the letter ‘P’, but were still unable to predict that BANANA would have immediately come to their minds (Fig. \@ref(fig:rigidity)D, yellow bars in right column). A similar pattern was observed for the prototypical body part word HA(ND) and its surprising counterpart HA(IR): when playing normally, `r round(100*E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word %in% c('hand', 'hair'))%>%pull(bananahand)%>%mean())`% of the players selected letters that are consistent with the prototypical option HAND. This figure was `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hand')%>%pull(bananahand)%>%mean())`% among pretenders for whom the target word was HAND, in contrast to only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='hair')%>%pull(bananahand)%>%mean())`% among pretenders for whom the target word was HAIR (Fig. \@ref(fig:rigidity)D, blue bars).

```{r rigidity, echo=FALSE, message= F, warning=F, fig.cap="Limitations on flexible decision-making when pretending. A: variability in the number of misses (extracted individually for each player and then averaged) was lower in pretend games. B: sequential guess entropy, a measure of the (inverse) predictability of individual players’ guesses as a function of click number and guess number. In both Battleship and Hangman, sequential guess entropy increased with click number, and was overall lower in pretend games. Shaded areas represent the mean +- one standard error. C: letter frequency of Hangman guesses: the median rank frequency per participant in pretend and non-pretend games, with reference lines for the expected rank frequency for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step. D: first letter guesses in Hangman half-games, as a function of pretense condition and target word. Letters appear in alphabetical order. letters that appear in the target word are marked in yellow (BANANA), red (PAPAYA), blue (HAND) and magenta (HAIR). For reference, the overall distribution of letter guesses across all games and conditions is given below.", out.width = '100%'}
knitr::include_graphics("figures/stereotypical2.png")
```

## Failure to detect pretending

These systematic deviations of pretend games from behaviour under true ignorance made pretend games discernible from non-pretend games. To illustrate, a support vector machine (SVM) algorithm reached an accuracy level of 72% (Battleship, 2AFC) and 60% (Hangman, yes/no) in linearly classifying condition (pretend / non-pretend) based on just three basic summary features: median decision latency, game optimality score, and number of irrational guesses per game. Despite this, human observers were entirely fooled by players’ pretending. After performing pretend and non-pretend games, participants were presented with game replays of previous players and took on the new role of being a 'judge' who had to determine who was pretending and who was playing for real. In Battleship, judges were presented with five pairs of game replays from different players. Within each pair, one player was a pretender, and judges’ task was to identify the pretender (or, for a random half of the judges, identify the real player). In Hangman, judges were presented with five game replays, each from a different player attempting to reveal a different word. For each individual game replay, the probability of it being a pretend game was 0.5, such that the number of pretend game replays was anywhere between 0 and 5 following a binomial distribution. Judges’ task was to decide, for each game, whether it was played as a pretend or a non-pretend game. In both games, pretense detection was at chance (Battleship: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%, `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=mean(correct))%>%pull(accuracy)%>%t.test(mu=0.5))$statistic`; Hangman: `r round(E4.judge_summary_df$accuracy%>%mean()*100)`%, `r apa_print(E4.judge_df%>%group_by(subj_id)%>%summarise(accuracy=mean(correct))%>%pull(accuracy)%>%t.test(mu=0.5))$statistic`). This is in line with previous findings of near-chance accuracy in lie detection [@bond2006]. Moreover, we find no sign of a correlation between pretense quality (measured as players' ability to trick judges into thinking they were not pretending) and pretense detection ability (measured as proportion correct; Battleship: `r apa_print(cor.test(E2.performance_df$judge_rate,E2.performance_df$pretend_rate, method='spearman'))$estimate`, Hangman: `r apa_print(cor.test(E4.performance_df$pretend_rate,E4.performance_df$judge_rate,method='spearman'))$estimate`), indicating that pretense and pretense detection rely on at least partly different cognitive processes.

# Discussion

In two experiments, we examined participants' ability to mimic a state of ignorance in a game setting, building on the recent recognition of games as a powerful tool for studying decision making [@allen2024]. We find that pretenders were able to successfully emulate decisions taken under a true state of ignorance. By extracting the same statistical and model-derived measures from pretend and non-pretend behaviour, we were able to directly compare how people truly solve a puzzle with how they believe they would solve the puzzle had they not known the solution. This approach revealed that people are capable of reproducing both broad patterns and subtle effects of guess accuracy and decision uncertainty on decision time. We also identify reliable signatures of pretend-ignorance on players’ decisions, including a cost to decision rationality and an increased tendency to follow heuristics and rules, even though these signatures went undetected by ‘judges’ asked to discriminate real from pretend games. Collectively, our findings are most consistent with epistemic pretense involving model-based self-simulation, based on a simplified model of participants’ own cognition

Previous research has identified limitations in our capacity to prevent knowledge from influencing our decisions and behavior [@fischhoff1975; @fischhoff1977; @wood1978; @roese2012; @harley2004]. In some cases, attempts to suppress thoughts even give rise to the paradoxical enhancement of suppressed representations [@wegner1987paradoxical; @earp2013; @giuliano2010]. Our findings reveal that notwithstanding these limitations, humans are capable of approximating their hypothetical behavior had they not known what they in fact do know. This capacity goes beyond making similar decisions to the ones they would have made had they not known; pretenders were also able to generate decision times that reproduce subtle qualitative patterns observed under a true state of ignorance.

<!-- This alignment between pretend and non-pretend decision times may indicate that pretenders were capable of controlling their decision times at will, enacting their intuitive beliefs of how decision time should be affected by knowledge and uncertainty. In a recent developmental study, children aged 5-10 were capable of making inferences about knowledge from the observed response time of other agents (e.g., inferring that decision-makers who take longer to decide make higher-quality decisions @richardson2022). Our findings here may show that people are also capable of using this sort of intuitive knowledge to guide their own pretense behavior. Alternatively, this mimicry of decision time patterns may reflect a similarity in the computations leading to pretend and non-pretend decision making. According to this interpretation, pretenders may have paused for longer when making some decisions because the simulated computation that led to these decisions took longer. -->

Internal simulations of decision-making processes are often studied (for example, in research on Bayesian Theory of Mind) by measuring participants' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational; @richardson2022], or implicitly [@onishi2005fifteen; @liu2017ten]. Here we have proposed a complementary approach: Asking participants to generate behavior based on a counterfactual mental state---In this case, a counterfactual knowledge state in which a known piece of information is unknown. Instead of relying on model inversion (e.g., "*Which belief states would give rise to this behavior?*"), we ask participants to run the model forward, taking counterfactual beliefs and desires as input and producing behavior as output.

Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory]. As such, our findings support a simulation model of epistemic pretense, and perhaps of mentalizing more generally. Critically, however, unlike classic “self-simulation” accounts of mindreading and theory of mind [@gallese1998; @gordon1986; @perner1996], which, in their purest form, entail that simulating ignorance should require effectively deleting or hiding mental representations from one’s self [@gordon2007], here the simulation is not of one’s actual cognitive machinery, but of a simplified, “cartoon” model of it that depicts its most salient surface-level aspects while ignoring details [@graziano2015]. A simulation of a schematic model explains both participants’ ability to mimic subtle patterns of true ignorance in an online fashion, as well as their consistent biases and limitations relative to behaviour when in a true state of ignorance [@saxe2005].

<!-- An alternative interpretation of our results is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored the revealed game state such that their entire cognitive machinery was available to play the game. This would not require self-simulation, only a capacity to intentionally 'unsee', or forget, relevant evidence. While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense, for at least five reasons. First, we tried to make such suppression as hard as possible, by presenting the game solution on top of the game board for the entire duration of pretend games, and by having participants type the target word before pretend Hangman games. Second, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical]. Third, when asked how they had performed the task in a debrief question, the responses of a significant majority of participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://matanmazor.github.io/pretendingNotToKnow/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)). Fourth, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig. @ref(fig:time)A): a finding that is consistent with the simulation of a simplified self-model, but is hard to explain if participants were merely suppressing their current knowledge. Finally, pretend games were more similar to each other than were non-pretend games to themselves. This is again consistent with an attempt to simulate typical behavior. -->

We interpret participants’ success in emulating a state of ignorance as revealing a non-trivial capacity for model-based counterfactual simulation, over and beyond any ability to suppress or ignore information (here, the game’s solution). This interpretation is supported by our finding, observed in both experiments, that pretend games were more similar to each other than were non-pretend games to each other, consistent with an attraction to the mean of a prior distribution [@mazor2021], or with an attempt to simulate representative behavior [@kahneman1972]. Such a tendency to avoid extreme events has been observed in the way people lie to an opponent [@oey2023], and in the generation of pseudorandom sequences of coin flips [@bar-hillel1991; @falk1997; @nickerson2002]. A similar effect is observed in Generative Adversarial Networks (GANs) where the distribution of generated samples is often narrower than the distribution of training data [an effect known as “mode collapse”, @kossale2022]. This underestimation of variability in game length cannot be explained by suppression alone. Additional support for a model-based simulation interpretation comes from the exaggerated, over-acted response-time profiles in pretend Battleship.

An alternative interpretation of our results is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored the revealed game state such that their entire cognitive machinery was available to play the game. This would not require self-simulation, only a capacity to intentionally ‘unsee’, or forget, relevant evidence. While we cannot fully rule out this interpretation, we think it is unlikely to explain players’ successful pretense, for at least three reasons over and above the tendency to produce representative behaviour described above. First, we tried to make such suppression as hard as possible, by presenting the game solution for the entire duration of pretend games, and by having participants type the target word before pretend Hangman games. Second, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical; @earp2013; @giuliano2010]. Third, when asked how they had performed the task in a debrief question, the responses of a significant majority of participants were aligned with self-simulation or rule-following, and our main findings hold when excluding the 32 Battleship and 10 Hangman players who mentioned suppression in response to this question (see exploratory analysis).

```{r num_suppressors, echo=FALSE, cache=TRUE}

E2.n_suppressors <- 10;
E4.n_suppressors <- 32;

suppression_table <- matrix(c(E2.n_suppressors, 
                                      E2.df$subj_id%>%unique()%>%length()-E2.n_suppressors, 
                                      E4.n_suppressors, 
                                      E4.df$subj_id%>%unique()%>%length()-E4.n_suppressors), nrow = 2, byrow = TRUE)

```

Findings from Battleship and Hangman mostly aligned: For both environments, the median number of guesses was similar in pretend and non-pretend games, guesses (correct and incorrect) made sense within the context of the game, and reaction times were similarly sensitive to guess accuracy and uncertainty. We also observed a similar tendency to produce representative and stereotypical behaviour in both experiments. At the same time, some differences are worth noting. First, fewer participants reported suppression as a strategy in pretend-Hangman games (`r round(100*E2.n_suppressors/E2.df$subj_id%>%unique()%>%length())`% of all pretenders) compared to pretend-Battleship games (`r round(100*E4.n_suppressors/E2.df$subj_id%>%unique()%>%length())`% of all pretenders; $p$ `r printp(chisq.test(suppression_table)$p.value,add_equals = TRUE)` in a Chi-square test of independence). This may be related to the fact that only in Hangman were players required to type in the target word before pretending, making suppression much harder. A second notable difference is the failure of many participants to predict their behaviour in Hangman half-games — most notably, their inability to appreciate that a high frequency word (e.g., BANANA) would immediately come to mind — when knowing that the solution is a low-frequency word (e.g., PAPAYA). This failure may have to do with an important difference between the two games: in Battleship, success in the game depends on players’ ability to weigh the relative likelihood of a relatively constrained set of hypotheses (grid configurations), which are fully specified by the rules of the game. In Hangman, in contrast, even though the set of hypotheses may be tightly constrained, these hypotheses are not evident from the rules of the game themselves. As a result, success in Hangman depends also on specific hypotheses coming to mind: a process that is largely masked from awareness [@bear2020]. It is possible that, having conscious access to the process of deliberation between existing hypotheses but not to the process of generating new hypotheses, participants can successfully simulate the first but not the second. An additional, not mutually exclusive explanation is that successful pretense requires suppressing available representations as a precondition for the model-based simulation process, and that words are harder to suppress than grid configurations. Either way, identifying the limiting conditions on epistemic pretense would be an important next step for understanding the underlying cognitive mechanisms, and in identifying the scope and content of people’s models of their own minds

Our findings speak not only to people’s ability to simulate counterfactual mental states, but also to their ability to pretend, deceive and lie more broadly. Previous research has mostly focused on the simulation of counterfactual world states, with theoretical models that suggest a key role for model-based simulations in pretense behaviour [@nichols2000; @weisberg2013], a role for pretense in the development of reasoning about causation [@walker2013], and hard constraints on the capacity to deceive [@depaulo2003; @verschuere2023; @walczyk2003]. Others have focused on the interaction between liars and recipients, modelling the effect of liars’ models of recipients’ mental states [@oey2023] and showing consistently poor ability of observers to detect lies or pretense in others[@bond2006]. In contrast, our focus here is on a special kind of pretense, one involving simulations of a counterfactual internal belief state rather than a counterfactual state of the external world, and with no reference to a specific recipient. Such simulations are required not only in adversarial settings such as pretense and deceit, but also in teaching and explaining (“*Would I have understood my explanation if I was not familiar with the subject matter?*”), fairness judgments (“*Would I have been so impressed with this candidate if I didn't know they went to Harvard*”), intelligence attribution based on observed behaviour (“*They solved the puzzle faster than it would have taken me to solve it had I not known the solution*”) and legal settings (“*Please ignore this witness’s testimony in your decision, as they were found unreliable*”). As such, while our findings should be considered within the broader context of people’s ability to behave in accordance with an imaginary world state, we focus not on the dependence of deceit on models of the world or of other agents, but on its reliance on a model of the self. We suggest that this novel perspective may open entirely new avenues for research about self-models and metacognitive knowledge.

Together, our findings reveal a non-trivial capacity for pretending not to know. Complementing previous work on cognitive and perceptual hindsight biases, which traditionally focus on people’s inability to emulate ignorance, we show that people are in fact capable of accurately simulating diverse aspects of their decision-making processes, although they exhibit systematic shortcomings. We speculate that these shortcomings are consistent with the simulation of a simplified model of cognition, over and above any suppression of knowledge or sensory input. In revealing this powerful capacity, our findings raise many new theoretical questions to which we don’t yet have answers. Are there specific aspects of our knowledge, beliefs, or inferences that are harder than others to simulate, and is this related to a lack of metacognitive understanding of these aspects? Does pretending not to know rely on explicit, reportable self-knowledge, or on an implicit self-model? Is the ability to overcome the curse of knowledge in the context of pretending predictive of the ability to overcome it in communicating information to a naive audience? Further research into these and similar limitations may continue to reveal the simplifications, abstractions, and biases in people’s models of their own minds.

<!-- including patterns of active information seeking and knowledge update in the face of new evidence. -->

<!-- We consider two possible explanations to  -->

<!-- Our subjects rose to the challenge. Their pretense behavior, although not perfect, closely resembled that of non-pretenders \-\-- not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency. Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: "What would I have done if I didn't have this piece of knowledge?". -->

<!-- An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game. This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something. While we cannot fully rule out this interpretation, we think it is unlikely to explain participants' pretense ability in our data, for two indirect reasons. First, suppressing thoughts on demand is notoriously difficult [\@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible. Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://self-model.github.io/pretendingNotToKnow/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)). -->

<!-- In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects' models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense. While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation. -->

# References

::: {#refs custom-style="Bibliography"}
:::

### 2025-08-31 analysis/clickEntropyBattleship.R

@@ -1,0 +1,72 @@
+
+#compute Shannon entropy, but only until the first timepoint at which some gamed ended
+entropy <- function(target) {
+  if (length(target)<5) {
+    return(NaN)
+  } else {
+    freq <- table(target)/length(target)
+    # vectorize
+    vec <- as.data.frame(freq)[,2]
+    #drop 0 to avoid NaN resulting from log2
+    vec<-vec[vec>0]
+    #compute entropy
+    return(-sum(vec * log2(vec)))
+  }
+}
+
+E2.click_entropy_by_serial_position <- E2.click_log_with_boards %>%
+  rbind(E2.click_log_with_boards_optimal)%>%
+  rbind(E2.click_log_with_boards_random)%>%
+  # filter(click_number<15) %>%
+  group_by(subj_id,test_part,click_number)%>%
+  summarise(entropy=entropy(flat_position))
+
+E2.click_entropy_by_serial_position_wide <- E2.click_entropy_by_serial_position %>%
+  spread(test_part,entropy) %>% 
+  mutate(diff=pretend-nonpretend)
+
+E2.click_entropy_by_serial_position_summary <- E2.click_entropy_by_serial_position %>%
+  group_by(test_part,click_number) %>%
+  summarise(mean_entropy=mean(entropy, na.rm=T),
+            se_entropy=se(entropy, na.rm=T)) %>%
+  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));
+
+p <- E2.click_entropy_by_serial_position_summary %>%
+  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
+  geom_line() +
+  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
+  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
+  labs(x='click number', y='entropy') +
+  theme_classic()
+
+ggsave('../docs/figures/E2_entropy_by_click_nnumber.png',p,width=7,height=5,dpi=300)
+
+E2.entropy_summary <- E2.click_entropy_by_serial_position %>%
+  filter(click_number<6) %>%
+  group_by(subj_id,test_part) %>%
[…]

### 2025-08-31 analysis/clickEntropyHangman.R

@@ -1,0 +1,73 @@
+
+#compute Shannon entropy, but only until the first timepoint at which some gamed ended
+entropy <- function(target) {
+  if (length(target)<5) {
+    return(NaN)
+  } else {
+    freq <- table(target)/length(target)
+    # vectorize
+    vec <- as.data.frame(freq)[,2]
+    #drop 0 to avoid NaN resulting from log2
+    vec<-vec[vec>0]
+    #compute entropy
+    return(-sum(vec * log2(vec)))
+  }
+}
+
+E4.click_entropy_by_serial_position <- E4.click_log_with_boards %>%
+  mutate(flat_position = which(LETTERS==letter)) %>%
+  # filter(click_number<15) %>%
+  group_by(subj_id,test_part,click_number)%>%
+  summarise(entropy=entropy(flat_position))
+
+E4.click_entropy_by_serial_position_wide <- E4.click_entropy_by_serial_position %>%
+  spread(test_part,entropy) %>% 
+  mutate(diff=pretend-nonpretend)
+
+
+E4.click_entropy_by_serial_position_summary <- E4.click_entropy_by_serial_position %>%
+  group_by(test_part,click_number) %>%
+  summarise(mean_entropy=mean(entropy, na.rm=T),
+            se_entropy=se(entropy, na.rm=T)) %>%
+  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));
+
+p <- E4.click_entropy_by_serial_position_summary %>%
+  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
+  geom_line() +
+  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
+  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
+  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
+  labs(x='click number', y='entropy') +
+  theme_classic()
+
+ggsave('../docs/figures/e4_entropy_by_click_nnumber.png',p,width=7,height=5,dpi=300)
+
+
+E4.entropy_summary <- E4.click_entropy_by_serial_position %>%
+  filter(click_number<6) %>%
+  group_by(subj_id,test_part) %>%
+  summarise(entropy=mean(entropy,na.rm=T))
+
[…]

### 2025-08-31 analysis/pneighbour.R

@@ -1,0 +1,194 @@
+# Define the list of 5x5 matrices (9 grids)
+grids <- list(
+  matrix(c(
+    '0','0','0','0','0',
+    '0','0','0','0','A',
+    'b','B','B','b','A',
+    '0','0','0','0','A',
+    'C','C','c','0','0'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    'C','C','c','0','0',
+    '0','0','0','0','0',
+    'B','B','b','A','0',
+    '0','0','0','A','0',
+    '0','0','0','A','0'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    'B','B','b','0','0',
+    '0','0','0','0','0',
+    'A','A','A','0','0',
+    '0','0','0','0','0',
+    '0','0','c','C','C'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    '0','0','0','0','B',
+    '0','0','0','0','B',
+    '0','A','A','A','b',
+    '0','0','0','0','0',
+    '0','c','C','C','c'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    '0','0','0','0','B',
+    'A','A','A','c','B',
+    '0','0','0','C','b',
+    '0','0','0','C','0',
+    '0','0','0','c','0'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    '0','0','0','0','0',
+    'c','0','0','0','0',
+    'C','0','b','0','A',
+    'C','0','B','0','A',
+    'c','0','B','0','A'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    '0','0','0','B','0',
+    '0','0','0','B','0',
+    'C','C','c','b','A',
+    '0','0','0','0','A',
+    '0','0','0','0','A'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    '0','0','0','C','0',
+    '0','0','0','C','0',
+    '0','0','0','c','0',
+    'A','A','A','0','0',
+    '0','0','b','B','B'
+  ), nrow=5, byrow=TRUE),
+  
+  matrix(c(
+    '0','A','0','0','0',
+    '0','A','b','B','B',
+    '0','A','c','0','0',
+    '0','0','C','0','0',
+    '0','0','C','0','0'
+  ), nrow=5, byrow=TRUE)
+)
+
[…]

### 2025-10-09 docs/ExploratoryResults.Rmd
model fits merged into one file, saving intermediate data automatically, and beautifying some parts of the paper

@@
 classoption       : "man"
-output            : word_document
+output            : bookdown::gitbook
 
@@
                       message=F, sanitize = T, 
-                      fig.width = 10, fig.height = 5)
+                      fig.width = 10, fig.height = 5,
+                      dev = "ragg_png", dpi = 300, fig.width = 6, fig.height = 4)
 ```
@@
 ```{r, libraries, echo=FALSE, warning=FALSE, cache=F}
+
+# Run on R 4.2.2 ("Innocent and Trusting")
 
@@
   'caret', #for cross validation
-  'ggrepel' #for word scatterplots
-), '2025-06-01')
+  'ggrepel', #for word scatterplots
+  'ragg' # for png
+), '2022-12-01')
 
@@
 ```{r hangman-load-coefficients, echo=FALSE}
-
-#get list of subjects
-E4.data_for_turing <- read.csv('../analysis/dataForTuring/E4norm.csv')
-E4.subjects <- E4.data_for_turing$subj_id%>%unique()
-
-#initialize
-E4.model_coefficients_raw <- data.frame()
-
-#fill the array
-for (i_subj in 1:length(E4.subjects)) {
-  E4.model_coefficients_raw <- E4.model_coefficients_raw %>%
-    rbind(read.csv(sprintf('../analysis/model_fits/E4/subj%d_normalized.csv',i_subj)) %>%
-            mutate(subj_id=E4.subjects[i_subj]) %>%
-            dplyr::select(subj_id,term,estimate))
-}
-
-# Organize in a more convenient form
-# non-pretend dataframe
-
-E4.np_coefficients <- E4.model_coefficients_raw %>% 
-  group_by(subj_id) %>% 
-  summarize(test_part = 'nonpretend',
-            prior = estimate[term == 'coef_prior'],
-            post = estimate[term == 'coef_post'],
-            eig = estimate[term == 'coef_eig'])
-
-E4.p_coefficients <- E4.model_coefficients_raw %>% 
-  group_by(subj_id) %>% 
-  summarize(test_part='pretend',
-            prior = estimate[term == 'coef_prior'] +
-              estimate[term == 'coef_prior_delta'],
-            post = estimate[term == 'coef_post'] +
-              estimate[term == 'coef_post_delta'],
-            eig = estimate[term == 'coef_eig'] +
-              estimate[term == 'coef_eig_delta'])
-
-E4.model_coefficients <- E4.np_coefficients %>%
-  rbind(E4.p_coefficients)
+E4.model_coefficients <- read.csv('../analysis/model_fits/model_coefficients.csv')
 ```
@@
 E4.model_coefficients %>%
+  filter(type=='basic') %>%
   gather('term','coefficient',prior,post,eig) %>%
@@
 All three terms had a positive effect in both conditions.
-Prior probability (base rate) had a stronger effect in pretend games (`r apa_print(E4.model_coefficients %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
-Hit posterior probability showed no such difference, and had a similar effect on players' decisions in both conditions (`r apa_print(E4.model_coefficients %>% dplyr::select(subj_id,test_part,post)%>%spread(test_part,post)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
-Finally, expected information gain had a weaker effect in pretend games (`r apa_print(E4.model_coefficients %>% dplyr::select(subj_id,test_part,eig)%>%spread(test_part,eig)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
-
-```{r hangman-load-coefficients-orth, echo=FALSE}
-
-#get list of subjects
-
-#initialize
-E4.model_coefficients_orth_raw <- data.frame()
-
-#fill the array
-for (i_subj in 1:length(E4.subjects)) {
-  E4.model_coefficients_orth_raw <- E4.model_coefficients_orth_raw %>%
-    rbind(read.csv(sprintf('../analysis/model_fits/E4/subj%d_orth.csv',i_subj)) %>%
-            mutate(subj_id=E4.subjects[i_subj]) %>%
-            dplyr::select(subj_id,term,estimate))
-}
-
-# Organize in a more convenient form
-# non-pretend dataframe
-
-E4.np_orth_coefficients <- E4.model_coefficients_orth_raw %>% 
-  group_by(subj_id) %>% 
-  summarize(test_part = 'nonpretend',
-            prior = estimate[term == 'coef_prior'],
-            post = estimate[term == 'coef_post'],
-            eig = estimate[term == 'coef_eig'])
-
-E4.p_orth_coefficients <- E4.model_coefficients_orth_raw %>% 
-  group_by(subj_id) %>% 
-  summarize(test_part='pretend',
-            prior = estimate[term == 'coef_prior'] +
-              estimate[term == 'coef_prior_delta'],
-            post = estimate[term == 'coef_post'] +
-              estimate[term == 'coef_post_delta'],
-            eig = estimate[term == 'coef_eig'] +
-              estimate[term == 'coef_eig_delta'])
-
-E4.model_coefficients_orth <- E4.np_orth_coefficients %>%
-  rbind(E4.p_orth_coefficients)
-```
+Prior probability (base rate) had a stronger effect in pretend games (`r apa_print(E4.model_coefficients %>% filter(type=='basic') %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+Hit posterior probability showed no such difference, and had a similar effect on players' decisions in both conditions (`r apa_print(E4.model_coefficients %>% filter(type=='basic') %>% dplyr::select(subj_id,test_part,post)%>%spread(test_part,post)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+Finally, expected information gain had a weaker effect in pretend games (`r apa_print(E4.model_coefficients %>% filter(type=='basic') %>% dplyr::select(subj_id,test_part,eig)%>%spread(test_part,eig)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
 
@@
 To isolate unique contributions of prior probability and expected information gain, we repeated the same analysis, but this time serially orthogonalizing the features (in the order [posterior, prior, expected information gain], and using the Gram Schmidt algorithm) prior to fitting the multinomial regression model.
-This time, expected information gain had no significant effect on decisions in non-pretend (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='nonpretend')%>%pull(eig)%>%t.test())$statistic`) and pretend games (`r  apa_print(E4.model_coefficients_orth%>%filter(test_part=='pretend')%>%pull(eig)%>%t.test())$statistic`), rendering it possible that the previously observed effects were confounded with prior and posterior probability.
-In contrast, both prior and posterior probability had a significant positive effect, with prior probability having a stronger effect in pretend games (`r apa_print(E4.model_coefficients_orth %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
+This time, expected information gain had no significant effect on decisions in non-pretend (`r  apa_print(E4.model_coefficients %>% filter(type=='orth') %>%filter(test_part=='nonpretend')%>%pull(eig)%>%t.test())$statistic`) and pretend games (`r  apa_print(E4.model_coefficients%>% filter(type=='orth') %>%filter(test_part=='pretend')%>%pull(eig)%>%t.test())$statistic`), rendering it possible that the previously observed effects were confounded with prior and posterior probability.
+In contrast, both prior and posterior probability had a significant positive effect, with prior probability having a stronger effect in pretend games (`r apa_print(E4.model_coefficients %>% filter(type=='orth') %>% dplyr::select(subj_id,test_part,prior)%>%spread(test_part,prior)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$full_result`).
 
@@
 
-E4.model_coefficients_orth %>%
+E4.model_coefficients %>%
+  filter(type=='orth') %>%
   gather('term','coefficient',prior,post,eig) %>%
@@
 
+```{r save data, echo-FALSE}
+
+save.image(file = file.path(getwd(), "explor_results.RData"))
+
+```
 \newpage

### 2025-10-09 docs/paper.Rmd
model fits merged into one file, saving intermediate data automatically, and beautifying some parts of the paper

--- 2024-01-17 docs/paper.Rmd
@@
 ---
-title             : "Pretending not to know reveals a powerful capacity for self-simulation"
+title             : "Pretending not to know reveals a capacity for model-based self-simulation"
 shorttitle        : "Pretending not to know"
@@
     affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
+    corresponding : yes    
     address       : "All Souls College, High Street, Oxford OX1 4AL"
     email         : "matan.mazor@all-souls.ox.ac.uk"
-  #   role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-  #     - Conceptualization
-  #     - Writing - Original Draft Preparation
-  #     - Writing - Review & Editing
 
@@
     affiliation   : "2"
-    # role:
-    #   - Writing - Review & Editing
 
@@
 abstract: |
-  What we do depends on what we know.
-But sometimes we try to decouple our behavior from our knowledge so as to appear not to know what we really do know.
-Such pretense behavior requires understanding how we would behave with different knowledge, and so provides a window into counterfactual self-simulation.
-However, little research has characterized and evaluated pretense relative to non-pretense behavior.
-Here, in two large-scale, pre-registered experiments, we tested people’s ability to conceal knowledge in the context of information-seeking games (Battleship and Hangman).
-Participants played normal games (in which they began with no knowledge of the ships’ locations or the word’s identity), as well as ‘pretend’ games in which they were shown the game's full solution but were instructed to play *as though* they were ignorant.
-We find a remarkable capacity for pretense, with pretenders successfully reproducing broad patterns and subtle features of real players’ decision-making and timing — so much so that 'judge' participants failed to distinguish real players from pretenders.
-Nonetheless, pretense was characterized by pervasive differences, undetected by judges but revealed by follow-up analyses: pretenders played less rationally than real players, and had an increased tendency to rely on heuristics and rules.
-Our findings thus reveal a powerful but imperfect ability to simulate decisions and behavior based on a counterfactual belief state, and open up a new approach to studying self-simulation.
-
-  <!-- https://tinyurl.com/ybremelq -->
+  Pretending not to know requires appreciating how one would behave without a given piece of knowledge, and acting accordingly.
+Here, two game-based experiments reveal a capacity to simulate decision-making under such counterfactual ignorance.
+1001 English-speaking adults saw the solution to a game (ship locations in Battleship, the hidden word in Hangman) but attempted to play as though they never had this information.
+Pretenders accurately mimicked broad aspects of genuine play, including the number of guesses required to reach a solution, as well as subtle patterns, such as effects of decision uncertainty on decision time.
+While peers were unable to detect pretense, statistical analysis and computational modeling uncovered traces of ‘over-acting’ in pretenders’ decisions, suggesting a schematic simulation of their minds.
+Opening up a new approach to studying self-simulation, our results reveal intricate metacognitive knowledge about decision-making, drawn from a rich—but simplified—internal model of cognition.
 
 keywords          : "pretense; metacognition; theory of mind"
-wordcount         : "5986"
+wordcount         : "7745"
 
@@
 output            : papaja::apa6_word
+
 bibliography: references.bib
@@
 
-```{r analysis-preferences, warning=F, message=F}
+```{r analysis-preferences, warning=F, message=F, echo=FALSE}
 # Seed for random number generation
@@
 # documents). 
-load('../.RData')
+load('prereg_results.RData')
+load('explor_results.RData')
 
-r_refs("r-references.bib")
+papaja::r_refs("references.bib")
 
@@
 
+# Research Transparency Statement
+
+## General Disclosures
+
+The authors declare no conflicts of interest.
+Funding: This study was supported by an NSF BCS grant #2021053 awarded to C.F.
+M.M is supported by a post-doctoral research fellowship at All Souls College.
+Artificial intelligence: No artificial intelligence assisted technologies were used in this research or the creation of this article.
+Ethics: The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.
+
+## Experiment 1 (Battleship)
+
+Preregistration: The hypotheses and methods were preregistered on 2021-12–21, prior to data collection, and following the analysis of an independent pilot sample.
+A detailed pre-registration can be accessed at <https://osf.io/v9zsb>.
+The pre-registration was time-locked using cryptographic randomisation-based time-locking (Mazor, Mazor, & Mukamel, 2019) (protocol sum: 60c270410375e8a192468fc1a0e9c93da60d5e203eb2760b621a8631a26f4c5c; [link to relevant lines in experimental code](https://github.com/self-model/pretendingNotToKnow/blob/44ef964c0d7d4a8f3b76ec52316a8e624c41a047/experiments/Battleships2/code/index.html#L865-L876), making experimental randomisation causally dependent on the content of the pre-registration and thus ensuring that all data were collected after pre-registration.
+All pre-registered analyses are publicly available, including the report-generating R script (<https://self-model.github.io/pretendingNotToKnow/docs/exp.-1-battleship.html>).
+Exploratory analyses are flagged as such.
+Materials: All study materials, including demos of analysis experiments, are publicly available (<https://github.com/self-model/pretendingNotToKnow>).
+Data: All primary data are publicly available (<https://osf.io/zma9b>).
+Analysis scripts: All analysis scripts are publicly available (<https://osf.io/zma9b>).
+
+## Experiment 2 (Hangman)
+
+Preregistration: The hypotheses and methods were preregistered on 2022-06–22, prior to data collection, and following the analysis of an independent pilot sample.
+A detailed pre-registration can be accessed at osf.io/3thry.
+The pre-registration was time-locked using cryptographic randomisation-based time-locking, making experimental randomisation causally dependent on the content of the pre-registration and thus ensuring that all data were collected after pre-registration [@mazor2019].
+Due to an error in the experiment code, time-locking took effect only from player number 221 (batch 3) and on (protocol sum: c4929c7fe33df1b7b52f15c789d98eab30a9cee09a8121807a3c59e28e7430a4;[relevant lines in experimental code](https://github.com/self-model/pretendingNotToKnow/blob/f98903f3f2d0e21800497dc65f8d5f823c910c88/experiments/Hangman2/code/index.html#L772-L783).
+All pre-registered analyses are publicly available, including the report-generating R script (<https://self-model.github.io/pretendingNotToKnow/docs/exp.-2-hangman.html>).
+Exploratory analyses are flagged as such.
+Materials: All study materials, including demos of the experiments, are publicly available (<https://github.com/self-model/pretendingNotToKnow>).
+Data: All primary data are publicly available (<https://osf.io/zma9b>).
+Analysis scripts: All analysis scripts are publicly available (<https://osf.io/zma9b>).
+
 # Introduction
 
-<!-- The ability to intentionally deceive others relies on a capacity to reason about mental states [@frith2005theory].
-This is evident in a similar developmental trajectory for the acquisition of theory of mind and the ability to deceive and detect deception [@wimmer1983beliefs; @shultz1981development; @sodian1991early], and a similar distribution of deception and theory of mind in the animal kingdom [e.g., @emery2004mentality; @hall2017cooperation].
-This link makes conceptual sense: to deceive others, one needs to understand that others can have different knowledge and beliefs than one's own.
--->
-
 Pretense relies on an ability to simulate and mimic one's own behavior under a counterfactual belief state.
-For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("I know that they are planning a surprise party, but they don't know that I know that...") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("Where would I look first had I not known?
+For example, in order to successfully deceive your friends into thinking that you were surprised by the birthday party they threw for you, it is not sufficient that you are able to reason about their mental states ("*I know that they are planning a surprise party, but they don't know that I know that...*") --- you also need to convincingly simulate and mimic your hypothetical behavior had you not known about the party ("*Where would I look first had I not known?
 What would I say?
-How long would it take me to recover from the surprise?").
+How long would it take me to recover from the surprise?*").
+Similar examples abound in higher-stakes contexts such as diplomacy, warcraft and law.
 This is not a trivial challenge: previous research on "hindsight biases" suggests that knowledge about the actual state of the world can interfere with our ability to correctly judge what we would have believed [@fischhoff1975; @fischhoff1977; @wood1978; @roese2012] or perceived [@harley2004; @bernstein2007; @bernstein2012] without this knowledge.
@@
 
-<!-- to what extent humans are capable of generating complex behavior that is aligned with a counterfactual belief state, and specifically with a counterfactual state of ignorance. -->
+Our research question is whether people can mentally simulate their actions under a counterfactual knowledge state of ignorance.
+To that end, we had participants “pretend not to know” in a game setting.
+Using an online version of the games Battleship and Hangman (in which players seek to uncover the locations of enemy ships or the identity of a word), participants played a ‘non-pretend’ (normal) version of the game, as well as a ‘pretend’ version where they were given complete information about the hidden ships / target word but were instructed to behave as if they didn’t have this information.
+Participants’ pretense behaviour mirrored broad patterns and subtle features of real players’ decisions and decision times.
+At the same time, epistemic pretense was characterized by over-acting, stereotypical behavior, and suboptimal incorporation of new information: all markers of model-based simulation.
+Together, we take these findings as evidence for a capacity to mentally simulate decisions and actions using a simplified and schematic self-model.
 
-To this end, we examine pretense in a game setting.
-Using an online version of the games Battleship and Hangman (in which players seek to uncover the locations of enemy ships or the identity of a word), participants played a 'non-pretend' (normal) version of the game, as well as a 'pretend' version where they were given complete information about the hidden ships / target word but were instructed to behave as if they didn't have this information.
-Overall, we find a highly impressive capacity for pretending not to know, with pretenders mirroring broad patterns and subtle features of real players' decisions and decision times.
-However, we also find that pretense was characterized by over-acting, stereotypical behavior, and suboptimal incorporation of new information, although these were undetected by human observers.
+# Method
 
-# Results
+The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.
+In two experiments, online participants played online versions of two information-seeking games.
+In Battleship (Exp.
+1), `r E2.df$subj_id%>%unique()%>%length()` English-speaking players (recruited from Prolific.com) were presented with a 5x5 grid of yellow squares, and attempted to reveal one size-3 submarine and two size-2 patrol boats with as few guesses as possible.
+In our version of the game, ships could only touch corner-to-corner, but not side-to-side (this was explained to participants before playing), and participants were not notified once they had sunk a ship (only whether their guess was a hit or a miss).
+In Hangman, N=`r E4.df$subj_id%>%unique()%>%length()` English-speaking players attempted to reveal a hidden word, name or number (hereafter referred to broadly as "word') with as few letter-guesses as possible, based on word length and category (a famous person, number, fruit, US state, or body part).
+To ensure familiarity with US states, Hangman participants were all US-based.
 
-In two experiments, players played online versions of two information-seeking games: Battleship (N=`r E2.df$subj_id%>%unique()%>%length()` players) and Hangman (N=`r E4.df$subj_id%>%unique()%>%length()` players).
-These games traditionally start in a state of ignorance, with a player's goal being to reveal an unknown world state (ship locations in Battleship, a hidden word in Hangman) in as few steps (cell or letter selections) as possible.
-Critically, in addition to playing five standard games, players in our experiments also completed five 'pretend' games in which the solution to the game was known to them from the start, but where their task was to behave as if it was not -- i.e.
-to play as though they did not have this information (see Fig.
+Both games traditionally start in a state of ignorance, with a player’s goal being to reveal an unknown world state (ship locations in Battleship, a hidden word in Hangman) in as few steps (cell or letter selections) as possible.
+Critically, in addition to playing five standard games, players in our experiments also completed five ‘pretend’ games in which the solution to the game was known to them from the start and remained visible on the screen throughout the entire game (pretend-Battleship ship locations were marked with a cross, pretend-Hangman words were presented visually and had to be typed by players before the game, to ensure encoding; see Fig.
 \@ref(fig:design)).
-We measured participants' capacity to simulate a counterfactual state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games.
-Our full pre-registered results are available [online](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html) together with the report-generating code.
-Readers are invited to try [demos of the experiments](https://matanmazor.github.io/ignorance/experiments/demos/pretend).
+In these games, players’ task was to behave as if they were playing for real – i.e.
+to play as though they did not have this information.
+
+Throughout the game, participants accrued points that were later converted to a monetary bonus.
+In non-pretend games, participants received points for revealing the ships, or the target word, with as few guesses as possible.
+In pretend games, participants were given different instructions:
+
+> “In this round, we’re going to tell you where the ships are, but we want you to act like you don’t know this information.
+We’ve marked the ships’ locations with a cross, so you’ll know where they are the whole time; but your job is to play the game as if these hints aren’t there.
+To see how good you are at this, we’re going to compare your games to the games of people who actually had no hints, and see how similar they are.
+We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you’ll get bonus points.
+But if your games look different, you won’t get these bonus points.
+Your number of clicks in this part will not affect your bonus.
+Only your ability to play like you had no hints.”
+
+And in Hangman:
+
+> “In this round, we’re going to tell you the word in advance, but we want you to act like you don’t know this information.
+To see how good you are at this, we’re going to compare your games to the games of people who played normally, without knowing what the word was, and see how similar they are.
+We will measure which letters you click and the timing of your guesses; if your clicks look similar to people who played like normal (trying to reveal the word with as few guesses as possible, but without any hints), you’ll get bonus points.
+But if your games look different, you won’t get these bonus points.
+Your number of clicks in this part will not affect your bonus.
+Only your ability to play like you didn’t see the word in advance.”
+
+We intentionally included no reference to an observer in these instructions, to have participants focusing on simulating their own behaviour rather than simulating how their behaviour would be perceived by another person.
+In reality, participants’ games were presented to other participants, and they received bonus points if they tricked these other participants into believing they did not have hints.
+
+Players played pretend and standard games in separate blocks that were presented in random order after a first ‘practice’ game.
+In principle, participants could learn about their own behaviour from this practice game.
+To minimize such learning effects, we distinguished practice games from the main experimental blocks, using a smaller 4x4 grid with only two size-2 ships in Battleship, and a word category (animals) that was not used in the main experiment in Hangman.
+Each experimental block was followed by a half-game, where players were instructed to complete the game from a half-finished state.
+Finally, players were presented with replays of the games of previous players, and judged which were standard and which were pretend games.
+We measured players’ capacity to simulate a counterfactual state of ignorance by comparing patterns of decisions and decision times in pretend and non-pretend games.
+Our full pre-registered results are available [online](https://self-model.github.io/pretendingNotToKnow/docs/exp.-1-battleship.html) together with the report-generating code.
+Unless otherwise specified, all reported findings similarly hold when analysing only the first condition performed by each subject in a between-subject analysis, thereby ensuring that findings are not driven by learning effects^[In both experiments, the order of pretend and non-pretend blocks was counterbalanced between participants.
+We observed no significant interaction between the strength of any of our effects (i.e., differences between pretend and non-pretend conditions) and part of the experiment (i.e., first versus second).
+For full details see Supplementary Materials.].
+Readers are invited to try [demos of the experiments](https://self-model.github.io/pretendingNotToKnow/experiments/demos/pretend).
 
@@
 
+# Results
+
 ## Measuring pretense quality
@@
 As a first measure of pretense quality, we compared the total number of guesses in pretend and non-pretend games.
-Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()%>%printnum()`; $p=$ `r apa_print(E2.median_clicks$diff%>%t.test())$table$p.value`; Fig.
+Among Battleship players, the number of cell selections was similar in pretend (mean = `r E2.median_clicks%>%pull(pretend)%>%mean()%>%printnum()`, SD = `r E2.median_clicks%>%pull(pretend)%>%sd()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks%>%pull(nonpretend)%>%mean()%>%printnum()`, SD = `r E2.median_clicks%>%pull(nonpretend)%>%sd()%>%printnum()`; `r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$statistic`; Cohen's $d$ = `r E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`; Fig.
 \@ref(fig:decisions)A).
 `r E2.lucky_pretenders%>%length` pretenders who immediately discovered all ships without making errors were excluded from all further analyses, in accordance with our pre-registered plan.
-With these subjects excluded, the number of cell selections remained very similar in pretend (mean = `r  E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()`; $p=$`r apa_print(E2.median_clicks%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
-\@ref(fig:decisions)A)).
-In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()%>%printnum()` misses; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()%>%printnum()` misses; $p$ `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+With these subjects excluded, the number of cell selections remained very similar in pretend (mean = `r  E2.median_clicks_filtered%>%pull(pretend)%>%mean()%>%printnum()`) and non-pretend games (mean = `r E2.median_clicks_filtered%>%pull(nonpretend)%>%mean()%>%printnum()`; `r apa_print(E2.median_clicks_filtered%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$statistic`; Fig.
+\@ref(fig:decisions)A).
+In Hangman, pretenders tended to make about one additional letter guess on average than did non-pretenders, controlling for word length (pretend: `r E4.median_misses%>%pull(pretend)%>%mean()%>%printnum()` misses, SD = `r E4.median_misses%>%pull(pretend)%>%sd()%>%printnum()`; non-pretend: `r E4.median_misses%>%pull(nonpretend)%>%mean()%>%printnum()` misses, , SD = `r E4.median_misses%>%pull(nonpretend)%>%sd()%>%printnum()`; `r apa_print(E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test())$statistic`; Cohen's $d$ = `r E4.median_misses%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`; Fig.
 \@ref(fig:decisions)B).
-Despite an overall bias in the number of guesses, pretend Hangman games showed an impressive, item-specific alignment: pretenders were successful in making more letter guesses when attempting to reveal words that would have been harder to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`; Fig.
+Despite an overall bias in the number of guesses, pretend Hangman games showed a near-perfect item-specific alignment: pretenders were successful in making more incorrect letter guesses when attempting to reveal words that would have been harder to guess had they been playing for real (`r cor.test(E4.misses_per_word$pretend, E4.misses_per_word$nonpretend,method='spearman')%>%apa_print()%>%'$'(estimate)`; Fig.
 \@ref(fig:decisions)B).
@@
 C: Spatial guess distributions for pretend and non-pretend half-games (where players continued the game from a half-finished state) alongside their corresponding hit probability maps.
-D: Cell and letter selections were ranked according to their relative hit probability.
-Plotting the median rank per subject in pretend and non-pretend games, with reference lines for the expected rank probability for a random agent, and for a "greedy" agent that maximizes the probability of a hit in each step.
+D: Cell and letter selections were ranked according to their relative hit probability given the players’ knowledge at the time of making the decision (dynamically updated after each guess).
+Plotting the median rank per subject in pretend and non-pretend games, with reference lines for the expected rank probability for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step.
+Note that the expected rank for a greedy agent is greater than 1 because there was not always a single optimal choice.
 E: same as panel D but discarding all guesses that resulted in a hit.
@@
   mutate(subj_id=as.character(subj_id)) %>%
-  # rbind(E2.optimal_click_df) %>%
-  # rbind(E2.random_click_df) %>%
   mutate(test_part = factor(test_part, levels=c(
@@
 In order to directly compare pretend and non-pretend guesses for the same board state, Battleship players completed two half-games in which they were instructed to continue the game from a half-completed state.
-Hundreds of cell selections for the same board state revealed a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_pretend_nonpretend)$table$estimate`, $p=$ `r apa_print(cor_A_pretend_nonpretend)$table$p.value`; board B: $r=$ `r apa_print(cor_B_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_B_pretend_nonpretend)$table$p.value`; Fig.
+In standard games, players start in the same (blank) board state but quickly diverge as they make different guess sequences.
+Including half games allowed us access to hundreds of cell selections for the same board state from pretenders and non-pretenders.
+This way, we had sufficient statistical power to compare the two guess distributions.
+We find a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_A_pretend_nonpretend)$table$p.value`; board B: $r=$ `r apa_print(cor_B_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_B_pretend_nonpretend)$table$p.value`; Fig.
 \@ref(fig:decisions)C), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.
@@
 To further examine the decisional processes behind this strong alignment, we compared the degree to which pretend and non-pretend guesses made sense within the context of the game.
-When playing Battleship and Hangman, it makes sense to guess cells or letters for which the probability of hitting a ship or revealing a letter is high [this "greedy" behavior is not strictly optimal, but approximates optimal behavior in most cases, @audinot2014optimal, Section 3.3].
-In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).
-Despite being less rational than non-pretend guesses, pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(np_vs_r))%>%apa_print%>%'$'(statistic)`; Fig.
+When playing Battleship and Hangman, it makes sense to guess cells or letters for which the probability of hitting a ship or revealing a letter is high [this "greedy" behavior is not strictly optimal, but approximates optimal behavior in most cases, @audinot2014optimal].
+To this end, we ranked cells based on the Bayesian probability of a hit given players’ knowledge at the time of making the decision.
+Critically, hit probability maps were dynamically updated after each guess.
+In Battleship, this model assumed that all legal board configurations are equally likely a priori, but board configurations were ruled out as the game progressed and the content of individual cells was revealed.
+In Hangman, we used the category information (e.g., ‘a fruit’), to obtain a probability-weighted list of category-compatible words (or names, in the case of famous people).
+We relied on prototypicality norms [@uyeda1980prototypicality] for words, and number of visits to Wikipedia entries for famous people.
+The full prior distributions for each category were included in the pre-registration (for details, see Supplementary Materials).
+Similar to Battleship, in deriving hit probability we assume access to the full list of options that is consistent with the game state (the number of hidden letters, the revealed letters and their positions, and the list of letters that do not appear in the game solution) at the time of making the decision.
+
+In the non-pretend versions of both games, guesses were more rational according to this measure than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E4.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-nonpretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`).
+Pretend guesses were also more rational than expected by chance (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`, Cohen's $d$ = `r cohensD(-E4.mean_P_click_rank_wide%>%pull(p_vs_r))%>%printnum()`), but significantly less rational than non-pretend guesses (Battleship: `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r t.test(-E4.mean_P_click_rank_wide%>%pull(p_vs_np))%>%apa_print%>%'$'(statistic)`, Cohen's $d$ = `r cohensD(-E4.mean_P_click_rank_wide%>%pull(p_vs_np))%>%printnum()` Fig.
 \@ref(fig:decisions)D).
-Critically, the same was true when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses_only%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Hangman: `r t.test(-E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`; Fig.
+Critically, pretend guesses were more rational than random guesses even when restricting the analysis to unsuccessful guesses (Battleship: `r E2.mean_P_click_rank_misses%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.mean_P_click_rank_misses%>%spread(test_part,p_click_rank)%>%mutate(diff=-pretend+random)%>%pull(diff)%>%cohensD()%>%printnum()`; Hangman: `r t.test(-E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r))%>%apa_print%>%'$'(statistic)`, Cohen's $d$ = `r cohensD(-E4.mean_P_click_rank_misses_wide%>%pull(p_vs_r))%>%printnum()`; Fig.
 \@ref(fig:decisions)E): that is, even when incorrectly guessing a ship's location or a letter's identity, pretend guesses made sense given the limited information players pretended to have.
+
+```{r pneighbour, echo=FALSE, message= F, warning=F}
+source('../analysis/pneighbour.R')
+```
+
+A specific example of this effect in the game of Battleship can be observed in players’ behaviour immediately after hitting the last cell of a size-2 patrol boat (players attempted to reveal two size-2 patrol boats and one size-3 submarine).
+Among non-pretenders, the next cell selection was often directed at checking whether the two cells were part of the size-3 submarine, but this was only true if the size-3 submarine had not been sunk yet (`r round(E2.pneighbour_nonpretend$second%>%mean(na.rm=T)*100)`% of all cell selections), and not when it had been sunk (`r round(E2.pneighbour_nonpretend$secondlast%>%mean(na.rm=T)*100)`% of all cell selections, and significantly lower than `r round(E2.pneighbour_nonpretend$second%>%mean(na.rm=T)*100)`%: `r apa_print(t.test(E2.pneighbour_nonpretend$diff))$statistic`, Cohen’s $d$ = `r printnum(cohensD(E2.pneighbour_nonpretend$diff))`).
+Despite knowing with full certainty that the size-2 patrol boat was not a size-3 submarine, pretenders showed the same qualitative pattern: pretending to check if the revealed cells were part of a size-3 submarine only when they pretended not to know that it was fully sunk (`r round(E2.pneighbour_pretend$second%>%mean(na.rm=T)*100)`% of all cell selections), but not when the size-3 submarine had been sunk (`r round(E2.pneighbour_pretend$secondlast%>%mean(na.rm=T)*100)`% of all cell selections, `r apa_print(t.test(E2.pneighbour_pretend$diff))$statistic`, Cohen’s $d$ = `r printnum(cohensD(E2.pneighbour_pretend$diff))`).
+The tendency to check if the two cells were part of a bigger ship was weaker among pretenders (`r apa_print(t.test(E2.pneighbour_wide$diff_second))$statistic`, Cohen’s $d$ = `r printnum(cohensD(E2.pneighbour_wide$diff_second))`).
 
@@
 Here too, pretend games showed the same qualitative patterns as non-pretend games.
-Like non-pretenders, pretenders were faster in their successful guesses (Difference in decision time between hits and misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms) and slowed down immediately after a hit (Difference in decision time between guesses that followed hits versus misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms; Fig.
+Like non-pretenders, pretenders were faster in their successful guesses (difference in decision time between hits and misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(hit_diff)%>%median(na.rm=T)%>%round()` ms) and slowed down immediately after a hit (difference in decision time between guesses that followed hits versus misses in Battleship: $\Delta_{non-pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median()%>%round()` ms, $\Delta_{pretend}=$ `r E2.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median()%>%round()` ms; Hangman: $\Delta_{non-pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms, $\Delta_{pretend}=$ `r E4.RT_by_hit_lags %>% filter(test_part=='pretend')%>%pull(lag1_diff)%>%median(na.rm=T)%>%round()` ms; Fig.
 \@ref(fig:time)A).
-Furthermore, decision uncertainty, quantified as the Shannon entropy of the posterior distribution over cell or letter options, had a similar quadratic effect on decision time in both non-pretend and pretend versions of the games, with the slowest responses associated with mid-range levels of entropy (see Fig.
-\@ref(fig:time)B).
+All effects are significant at the $0.001$ level with the pre-registered within-subject t-test, except for the post-hit slowing down in Battleship, which, due to outliers with extreme effects in the opposite direction (\>10 s), was only significant in a non-parametric Wilcoxon sign-rank test (`r apa_print(E4.RT_by_hit_lags %>% filter(test_part=='nonpretend')%>%pull(lag1_diff)%>%wilcox.test())$statistic`).
+Effects remain significant at the $0.001$ level when statistically controlling for the serial position of guesses within the game.
+
+We also examined the effect of decision uncertainty, quantified as the Shannon entropy of the posterior distribution over cell or letter options, on decision times.
+To this end we fitted subject-level linear models, predicting reaction times from the linear and quadratic expansions of decision entropy, and contrasted the coefficients against zero in a group-level t-test.
+In the non-pretend versions of the games, the quadratic coefficients were significantly negative, with the slowest responses associated with mid-range levels of entropy (see Fig.
+\@ref(fig:time)B; Battleship: `r apa_print(E2.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%t.test())$statistic`, Cohen's $d$ = `r E2.RT_by_entropy_coefs  %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%cohensD()%>%printnum()`; Hangman: `r apa_print(E4.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%t.test())$statistic`, Cohen's $d$ = `r E4.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(nonpretend)%>%cohensD()%>%printnum()`.
+When restricting the analysis to those Battleship players who pretended after playing normally, this effect was significant only in a Wilcoxon rank-sum test, due to outliers in the sample: `r apa_print(E2.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared' & genuine_first)%>%pull(nonpretend)%>%wilcox.test())$statistic`).
+Critically, the quadratic coefficients were significantly negative also in pretend games (see Fig.
+\@ref(fig:time)B; Battleship: `r apa_print(E2.RT_by_entropy_coefs  %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(pretend)%>%t.test())$statistic`, Cohen's $d$ = `r E2.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(pretend)%>%cohensD()%>%printnum()`; Hangman: `r apa_print(E4.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(pretend)%>%t.test())$statistic`, Cohen's $d$ = `r E4.RT_by_entropy_coefs %>% spread(test_part,estimate) %>% filter(term=='entropy_squared')%>%pull(pretend)%>%cohensD()%>%printnum()`).
 In other words, despite knowing the game's solution with full certainty, pretenders successfully feigned subtle qualitative effects of counterfactual uncertainty on their decision times.
@@
 B: median decision times as a function of decision uncertainty, quantified as the entropy of the posterior over guess options.
-In both Hangman and Battleship, guesses were slowest for mid-range levels of entropy, and this pattern was mimicked in pretend games.", out.width = '100%'}
-knitr::include_graphics("figures/time.png")
+In both Hangman and Battleship, guesses were slowest for mid-range levels of entropy, and this pattern was mimicked in pretend games.
+Shaded areas represent the bootstrapped standard error of the median.", out.width = '100%'}
+knitr::include_graphics("figures/time3.png")
 ```
@@
 sd_misses_df <- E2.click_df %>%
-  dplyr::select(subj_id,test_part, num_clicks) %>%
+  dplyr::select(genuine_first, subj_id,test_part, num_clicks) %>%
   mutate(subj_id=as.character(subj_id)) %>%
-  # rbind(E2.optimal_click_df) %>%
-  # rbind(E2.random_click_df) %>%
   mutate(test_part = factor(test_part, levels=c(
@@
     'non-pretend','pretend'))) %>%
-  group_by(subj_id,test_part) %>%
+  group_by(subj_id,genuine_first, test_part) %>%
   summarise(sd_misses=sd(num_clicks)) %>%
@@
     E4.click_df %>%
-  dplyr::select(subj_id,test_part, num_misses) %>%
+  dplyr::select(genuine_first, subj_id,test_part, num_misses) %>%
   mutate(subj_id=as.character(subj_id)) %>%
-  # rbind(E2.optimal_click_df) %>%
-  # rbind(E2.random_click_df) %>%
   mutate(test_part = factor(test_part, levels=c(
@@
     'non-pretend','pretend'))) %>%
-  group_by(subj_id,test_part) %>%
+  group_by(subj_id,genuine_first, test_part) %>%
   summarise(sd_misses=sd(num_misses)) %>%
@@
   theme_classic() +
-  # geom_vline(data=ref_lines, aes(xintercept = xintercept), color='black', linetype=2) +
-  # geom_text(data=ref_lines, aes(x = xintercept, y=y, label=label), angle = 90, vjust = -1, size = 3, inherit.aes = FALSE)+
   geom_bar(alpha=0.6, position = 'identity', colour='black', size=0.5) +
@@
 Though impressive, the capacity for simulating a state of ignorance was not perfect.
-Importantly, the limitations and biases we observe are consistent with the simulation of a stereotypical, “cartoon” model of decision-making, rather than leakage of concealed information into the decision-making process.
+Importantly, the limitations and biases we observe are consistent with the simulation of a stereotypical, “cartoon” model of decision-making, rather than leakage of concealed information into the decision-making process as would be expected if pretenders’ success was due to efficient, but imperfect, suppression of their knowledge of the game solution.
 First, despite showing the same qualitative effects, decision time patterns in Battleship pretend games (but not Hangman pretend games) were systematically more pronounced relative to non-pretend games: a form of “over-acting”.
+Specifically, the difference in reaction times as a function of guess outcome (Fig.3A) was larger in pretend games, both when measured with respect to the current guess (`r E2.RT_by_hit_lags_long %>% filter(position==0)%>%spread(test_part,difference)%>%mutate(diff=-pretend+nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.RT_by_hit_lags_long %>% filter(position==0)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`), and with respect to the following guess (`r E2.RT_by_hit_lags_long %>% filter(position==1)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.RT_by_hit_lags_long %>% filter(position==1)%>%spread(test_part,difference)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`).
+Similarly, the quadratic effect of decision entropy on decision times was stronger in pretend games (`r E2.RT_by_entropy_coefs%>%filter( term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=-pretend+nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`, Cohen's $d$ = `r E2.RT_by_entropy_coefs%>%filter( term=='entropy_squared')%>%spread(test_part,estimate)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%cohensD()%>%printnum()`).
+
 Furthermore, pretend games followed stereotypical patterns, and as a result were more homogeneous than non-pretend games.
 Despite a highly similar average number of misses in pretend and non-pretend games (Fig.
-\@ref(fig:decisions)A), the number of unsuccessful guesses was overwhelmingly less variable in pretend relative to non-pretend games (Battleship: sd=`r sd_misses_df_wide$p_Battleship%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Battleship%>%mean(na.rm=T)%>%printnum()` in non-pretend games, $p$ `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Battleship-np_Battleship)%>%pull(diff)%>%t.test())$table$p.value`; Hangman: sd=`r sd_misses_df_wide$p_Hangman%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Hangman%>%mean(na.rm=T)%>%printnum()` in non-pretend games, $p$ `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Hangman-np_Hangman)%>%pull(diff)%>%t.test())$table$p.value`; Fig.
+\@ref(fig:decisions)A), the number of unsuccessful guesses was overwhelmingly less variable in pretend relative to non-pretend games (Battleship: sd=`r sd_misses_df_wide$p_Battleship%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Battleship%>%mean(na.rm=T)%>%printnum()` in non-pretend games, `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Battleship-np_Battleship)%>%pull(diff)%>%t.test())$statistic`, Cohen's $d$ = `r printnum(sd_misses_df_wide%>%mutate(diff=p_Battleship-np_Battleship)%>%pull(diff)%>%cohensD())`; Hangman: sd=`r sd_misses_df_wide$p_Hangman%>%mean(na.rm=T)%>%printnum()` in pretend versus `r sd_misses_df_wide$np_Hangman%>%mean(na.rm=T)%>%printnum()` in non-pretend games, `r apa_print(sd_misses_df_wide%>%mutate(diff=p_Hangman-np_Hangman)%>%pull(diff)%>%t.test())$statistic`, Cohen's $d$ = `r printnum(sd_misses_df_wide%>%mutate(diff=p_Hangman-np_Hangman)%>%pull(diff)%>%cohensD())`; Fig.
 \@ref(fig:rigidity)A).
-Moreover, the within-participant sequential guess entropy, a measure of the inverse predictability of decision sequences across different games of the same player, was significantly reduced in pretend games (Fig.
-\@ref(fig:rigidity)B; see [online supplementary materials](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#within-participant-click-entropy)), consistent with an attempt to enact typical, or average, behavior in a state of ignorance.
+Moreover, while pretenders produced more letter misses for harder words (Fig.
+2B), they underestimated the difficulty of the very hard “DALAI LAMA” and overestimated the difficulty of the easy number (“ELEVEN” and “NINETY SIX”) and state (“MONTANA” and “IOWA”) words.
+That is, pretenders consistently enacted what they saw as a ‘typical’ or a ‘representative’ game, one that is not unusual in the number of lucky or unlucky guesses.
+This is again consistent with shrinkage toward the mean of a generative self-model [@mazor2021; @jansen2021], with an attempt to avoid extreme outcomes to appear convincing to a hypothetical observer [@oey2023], and with representativeness skewing intuitions about randomness [@kahneman1972].
+
+```{r entropy, echo=FALSE, message= F, warning=F}
+source('../analysis/clickEntropyBattleship.R')
+source('../analysis/clickEntropyHangman.R')
+```
+
+Next, we examined variability not in the number of guesses, but in their contents.
+We separately computed the Shannon entropy of the guess distribution across different games for each player, condition (pretend or non-pretend), and serial guess number.
+High entropy then corresponds to pronounced variability in the guess sequences of different games, and low entropy to a tendency to repeat the same sequence of guesses in different games.
+For example, if a player always starts their games by clicking in the top left corner, their guess entropy for the first click will be H([1,1,1,1,1])=0.
+Unsurprisingly, the within-participant sequential guess entropy increased as a function of guess number, consistent with players adjusting their behaviour in light of the outcomes of previous guesses, making individual games increasingly more varied (Fig.
+\@ref(fig:rigidity)B).
+If pretend games were a similar but noisier version of standard games, their associated guess entropy would be higher, reflecting the additional noise in the decision-making process, or the game-specific biases that are associated with the suppression of specific words or game states.
+Critically, however, entropy was systematically reduced in pretend games (p\<0.001 for a within-subject t-test of guess entropy in guesses number 1-4 in both Battleship and Hangman, see Supplementary Materials for guess-specific statistics).
+Thus, when pretending, participants produced similar guess sequences across different games.
+In Hangman, for example, this meant that non-pretenders more flexibly adjusted their first letter guess to the word category and number of letters, compared to pretenders, who tended to open the game with the same letter guess irrespective of the specific game state.
+This seems consistent with an attempt to enact what they saw as typical, representative, or average behaviour.
+In contrast, a reduction in the guess sequence entropy is inconsistent with leakage of suppressed knowledge into the decision-making process, as would be expected if differences between pretend and non-pretend games reflected the imperfect suppression of the game’s solution.
+
+One possible account of the reduced decision entropy in pretend games is that it reflects pretenders’ miscalibrated intuitions about randomness, conforming to a prototype of randomness that is itself too ordered.
+If the same prototype of randomness is used by pretenders to determine the number of unsuccessful guesses per game, the two measures should be correlated across participants.
+Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$full_result`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$full_result`).
+This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$full_result`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$full_result`).
+We interpret this as evidence that the reduction in variance reflects miscalibrated beliefs not only about randomness, but also about participants’ behaviour under a counterfactual knowledge state.
+Those players who thought they would strictly follow a particular sequence of guesses (low entropy), ended up producing games of more variable lengths, as their success depended more on luck.
+Other players adjusted their decision strategy more flexibly, perhaps attempting to produce games that are not too long or short, in line with their intuitions about randomness.
+
 Finally, Hangman pretenders were more likely to guess letters that appear frequently in English words (E, T, A, etc.) irrespective of the game state, compared to genuine players (Fig.
 \@ref(fig:rigidity)C).
-This suggests that in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result.
+This suggests that in their attempt to behave as if they didn't know the true state of the game, pretenders had an increased tendency to follow rigid heuristics and rules, ignoring useful information as a result (but see Supplementary Materials for evidence that heuristic use alone cannot fully explain pretenders' behaviour).
 
@@
 \@ref(fig:rigidity)D, yellow bars in left column).
-Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word== 'banana')%>%pull(bananahand)%>%mean())`%.
+Among pretenders who knew that the hidden word was BANANA, this preference was reduced to `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word== 'banana')%>%pull(bananahand)%>%mean())`% (this drop was significant in a binomial test: $p=$`r apa_print(binom.test(E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='banana')%>%pull(bananahand)%>%sum(),E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='banana')%>%nrow(),E4.first_click_half_games %>% filter(test_part=='nonpretend_half_game' & word =='banana')%>%pull(bananahand)%>%mean(na.rm=T)))$table$p.value`).
 Importantly, half of the pretenders were given different information: they were told that the hidden word was the less prototypical fruit PAPAYA.
-Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', revealing that many pretenders were unable to predict that the first fruit that would have come to their minds was BANANA, not PAPAYA (Fig.
+Although good pretenders should simulate their behavior had they not known this information, only `r round(100*E4.first_click_half_games %>% filter(test_part=='pretend_half_game' & word =='papaya')%>%pull(bananahand)%>%mean())`% selected the letters 'B' or 'N', with many guessing letters that are not consistent with either PAPAYA or BANANA (gray bars in Fig.
+4D), revealing that many pretenders had the knowledge that PAPAYA would be a hard fruit to guess, therefore avoiding the letter ‘P’, but were still unable to predict that BANANA would have immediately come to their minds (Fig.
 \@ref(fig:rigidity)D, yellow bars in right column).
@@
 In both Battleship and Hangman, sequential guess entropy increased with click number, and was overall lower in pretend games.
+Shaded areas represent the mean +- one standard error.
 C: letter frequency of Hangman guesses: the median rank frequency per participant in pretend and non-pretend games, with reference lines for the expected rank frequency for a random agent, and for a “greedy” agent that maximizes the probability of a hit in each step.
@@
 
-Despite these limitations, pretenders' behavior resembled real players' behavior closely enough that they were able to fool other participants into believing they were playing for real.
+These systematic deviations of pretend games from behaviour under true ignorance made pretend games discernible from non-pretend games.
+To illustrate, a support vector machine (SVM) algorithm reached an accuracy level of 72% (Battleship, 2AFC) and 60% (Hangman, yes/no) in linearly classifying condition (pretend / non-pretend) based on just three basic summary features: median decision latency, game optimality score, and number of irrational guesses per game.
+Despite this, human observers were entirely fooled by players’ pretending.
 After performing pretend and non-pretend games, participants were presented with game replays of previous players and took on the new role of being a 'judge' who had to determine who was pretending and who was playing for real.
-In both games, pretense detection was at chance (Battleship: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%, Hangman: `r round(E4.judge_summary_df$accuracy%>%mean()*100)`%).
+In Battleship, judges were presented with five pairs of game replays from different players.
+Within each pair, one player was a pretender, and judges’ task was to identify the pretender (or, for a random half of the judges, identify the real player).
+In Hangman, judges were presented with five game replays, each from a different player attempting to reveal a different word.
+For each individual game replay, the probability of it being a pretend game was 0.5, such that the number of pretend game replays was anywhere between 0 and 5 following a binomial distribution.
+Judges’ task was to decide, for each game, whether it was played as a pretend or a non-pretend game.
+In both games, pretense detection was at chance (Battleship: `r E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=sum(correct)*20)%>%pull(accuracy)%>%mean()%>%round()`%, `r apa_print(E2.judge_df%>%group_by(subj_id)%>%summarise(accuracy=mean(correct))%>%pull(accuracy)%>%t.test(mu=0.5))$statistic`; Hangman: `r round(E4.judge_summary_df$accuracy%>%mean()*100)`%, `r apa_print(E4.judge_df%>%group_by(subj_id)%>%summarise(accuracy=mean(correct))%>%pull(accuracy)%>%t.test(mu=0.5))$statistic`).
 This is in line with previous findings of near-chance accuracy in lie detection [@bond2006].
@@
 In two experiments, we examined participants' ability to mimic a state of ignorance in a game setting, building on the recent recognition of games as a powerful tool for studying decision making [@allen2024].
-We find that pretenders were able to successfully emulate decisions taken under a true state of ignorance, including reproducing both broad patterns and subtle effects of guess accuracy and decision uncertainty on decision time.
-We also identify reliable signatures of pretend-ignorance on players' decisions, including a cost to decision rationality and an increased tendency to follow heuristics and rules, even though these signatures went undetected by 'judges' asked to discriminate real from pretend games.
-Collectively, our findings reveal a rich and impressive, but ultimately limited, capacity for online simulation of counterfactual belief states.
+We find that pretenders were able to successfully emulate decisions taken under a true state of ignorance.
+By extracting the same statistical and model-derived measures from pretend and non-pretend behaviour, we were able to directly compare how people truly solve a puzzle with how they believe they would solve the puzzle had they not known the solution.
+This approach revealed that people are capable of reproducing both broad patterns and subtle effects of guess accuracy and decision uncertainty on decision time.
+We also identify reliable signatures of pretend-ignorance on players’ decisions, including a cost to decision rationality and an increased tendency to follow heuristics and rules, even though these signatures went undetected by ‘judges’ asked to discriminate real from pretend games.
+Collectively, our findings are most consistent with epistemic pretense involving model-based self-simulation, based on a simplified model of participants’ own cognition
 
 Previous research has identified limitations in our capacity to prevent knowledge from influencing our decisions and behavior [@fischhoff1975; @fischhoff1977; @wood1978; @roese2012; @harley2004].
+In some cases, attempts to suppress thoughts even give rise to the paradoxical enhancement of suppressed representations [@wegner1987paradoxical; @earp2013; @giuliano2010].
 Our findings reveal that notwithstanding these limitations, humans are capable of approximating their hypothetical behavior had they not known what they in fact do know.
@@
 
-This alignment between pretend and non-pretend decision times may indicate that pretenders were capable of controlling their decision times at will, enacting their intuitive beliefs of how decision time should be affected by knowledge and uncertainty.
-In a recent developmental study, children aged 5-10 were capable of making inferences about knowledge from the observed response time of other agents [e.g., inferring that decision-makers who take longer to decide make higher-quality decisions, @richardson2022].
-Our findings here may show that people are also capable of using this sort of intuitive knowledge to guide their own pretense behavior.
-Alternatively, this mimicry of decision time patterns may reflect a similarity in the computations leading to pretend and non-pretend decision making.
-According to this interpretation, pretenders may have paused for longer when making some decisions because the simulated computation that led to these decisions took longer.
+Internal simulations of decision-making processes are often studied (for example, in research on Bayesian Theory of Mind) by measuring participants' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational; @richardson2022], or implicitly [@onishi2005fifteen; @liu2017ten].
+Here we have proposed a complementary approach: Asking participants to generate behavior based on a counterfactual mental state---In this case, a counterfactual knowledge state in which a known piece of information is unknown.
+Instead of relying on model inversion (e.g., "*Which belief states would give rise to this behavior?*"), we ask participants to run the model forward, taking counterfactual beliefs and desires as input and producing behavior as output.
 
-Research on Bayesian Theory of Mind provides some support for such internal simulations of decision-making processes.
-These are often studied by measuring participants' ability to infer beliefs and desires from observed behavior, either explicitly [@baker2009action; @baker2017rational; @richardson2022], or implicitly [@onishi2005fifteen; @liu2017ten].
-Here we have proposed a complementary approach: Asking participants to generate behavior based on a counterfactual mental state---In this case, a counterfactual knowledge state in which a known piece of information is unknown.
-Instead of relying on model inversion (e.g., "Which belief states would give rise to this behavior?"), we ask participants to run the model forward, taking counterfactual beliefs and desires as input and producing behavior as output.
 Due to the unconstrained space of possible behaviors in our task (cell selections x decision latencies), successfully pretending not to know demands a rich model of cognition, and is much harder to achieve based on a quasi-scientific theory of mental states [@gopnik1994theory].
+As such, our findings support a simulation model of epistemic pretense, and perhaps of mentalizing more generally.
+Critically, however, unlike classic “self-simulation” accounts of mindreading and theory of mind [@gallese1998; @gordon1986; @perner1996], which, in their purest form, entail that simulating ignorance should require effectively deleting or hiding mental representations from one’s self [@gordon2007], here the simulation is not of one’s actual cognitive machinery, but of a simplified, “cartoon” model of it that depicts its most salient surface-level aspects while ignoring details [@graziano2015].
+A simulation of a schematic model explains both participants’ ability to mimic subtle patterns of true ignorance in an online fashion, as well as their consistent biases and limitations relative to behaviour when in a true state of ignorance [@saxe2005].
+
+We interpret participants’ success in emulating a state of ignorance as revealing a non-trivial capacity for model-based counterfactual simulation, over and beyond any ability to suppress or ignore information (here, the game’s solution).
+This interpretation is supported by our finding, observed in both experiments, that pretend games were more similar to each other than were non-pretend games to each other, consistent with an attraction to the mean of a prior distribution [@mazor2021], or with an attempt to simulate representative behavior [@kahneman1972].
+Such a tendency to avoid extreme events has been observed in the way people lie to an opponent [@oey2023], and in the generation of pseudorandom sequences of coin flips [@bar-hillel1991; @falk1997; @nickerson2002].
+A similar effect is observed in Generative Adversarial Networks (GANs) where the distribution of generated samples is often narrower than the distribution of training data [an effect known as “mode collapse”, @kossale2022].
+This underestimation of variability in game length cannot be explained by suppression alone.
+Additional support for a model-based simulation interpretation comes from the exaggerated, over-acted response-time profiles in pretend Battleship.
 
 An alternative interpretation of our results is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored the revealed game state such that their entire cognitive machinery was available to play the game.
-This would not require self-simulation, only a capacity to intentionally 'unsee', or forget, relevant evidence.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain our players' successful pretense, for at least five reasons.
-First, we tried to make such suppression as hard as possible, by presenting the game solution on top of the game board for the entire duration of pretend games, and by having participants type the target word before pretend Hangman games.
-Second, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical].
-Third, when asked how they had performed the task in a debrief question, the responses of a significant majority of participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).
-Fourth, response time patterns were exaggerated in pretend Battleship relative to non-pretend Battleship games (see Fig.
-\@ref(fig:time)A): a finding that is consistent with the simulation of a simplified self-model, but is hard to explain if participants were merely suppressing their current knowledge.
-Finally, pretend games were more similar to each other than were non-pretend games to themselves.
-This is again consistent with an attempt to simulate typical behavior.
+This would not require self-simulation, only a capacity to intentionally ‘unsee’, or forget, relevant evidence.
+While we cannot fully rule out this interpretation, we think it is unlikely to explain players’ successful pretense, for at least three reasons over and above the tendency to produce representative behaviour described above.
+First, we tried to make such suppression as hard as possible, by presenting the game solution for the entire duration of pretend games, and by having participants type the target word before pretend Hangman games.
+Second, suppressing thoughts on demand is notoriously difficult, and often has an opposite, positive effect on the suppressed content [@wegner1987paradoxical; @earp2013; @giuliano2010].
+Third, when asked how they had performed the task in a debrief question, the responses of a significant majority of participants were aligned with self-simulation or rule-following, and our main findings hold when excluding the 32 Battleship and 10 Hangman players who mentioned suppression in response to this question (see exploratory analysis).
+
+```{r num_suppressors, echo=FALSE, cache=TRUE}
+
+E2.n_suppressors <- 10;
+E4.n_suppressors <- 32;
+
+suppression_table <- matrix(c(E2.n_suppressors, 
+                                      E2.df$subj_id%>%unique()%>%length()-E2.n_suppressors, 
+                                      E4.n_suppressors, 
+                                      E4.df$subj_id%>%unique()%>%length()-E4.n_suppressors), nrow = 2, byrow = TRUE)
+
+```
+
+Findings from Battleship and Hangman mostly aligned: For both environments, the median number of guesses was similar in pretend and non-pretend games, guesses (correct and incorrect) made sense within the context of the game, and reaction times were similarly sensitive to guess accuracy and uncertainty.
+We also observed a similar tendency to produce representative and stereotypical behaviour in both experiments.
+At the same time, some differences are worth noting.
+First, fewer participants reported suppression as a strategy in pretend-Hangman games (`r round(100*E2.n_suppressors/E2.df$subj_id%>%unique()%>%length())`% of all pretenders) compared to pretend-Battleship games (`r round(100*E4.n_suppressors/E2.df$subj_id%>%unique()%>%length())`% of all pretenders; $p$ `r printp(chisq.test(suppression_table)$p.value,add_equals = TRUE)` in a Chi-square test of independence).
+This may be related to the fact that only in Hangman were players required to type in the target word before pretending, making suppression much harder.
+A second notable difference is the failure of many participants to predict their behaviour in Hangman half-games — most notably, their inability to appreciate that a high frequency word (e.g., BANANA) would immediately come to mind — when knowing that the solution is a low-frequency word (e.g., PAPAYA).
+This failure may have to do with an important difference between the two games: in Battleship, success in the game depends on players’ ability to weigh the relative likelihood of a relatively constrained set of hypotheses (grid configurations), which are fully specified by the rules of the game.
+In Hangman, in contrast, even though the set of hypotheses may be tightly constrained, these hypotheses are not evident from the rules of the game themselves.
+As a result, success in Hangman depends also on specific hypotheses coming to mind: a process that is largely masked from awareness [@bear2020].
+It is possible that, having conscious access to the process of deliberation between existing hypotheses but not to the process of generating new hypotheses, participants can successfully simulate the first but not the second.
+An additional, not mutually exclusive explanation is that successful pretense requires suppressing available representations as a precondition for the model-based simulation process, and that words are harder to suppress than grid configurations.
+Either way, identifying the limiting conditions on epistemic pretense would be an important next step for understanding the underlying cognitive mechanisms, and in identifying the scope and content of people’s models of their own minds
+
+Our findings speak not only to people’s ability to simulate counterfactual mental states, but also to their ability to pretend, deceive and lie more broadly.
+Previous research has mostly focused on the simulation of counterfactual world states, with theoretical models that suggest a key role for model-based simulations in pretense behaviour [@nichols2000; @weisberg2013], a role for pretense in the development of reasoning about causation [@walker2013], and hard constraints on the capacity to deceive [@depaulo2003; @verschuere2023; @walczyk2003].
+Others have focused on the interaction between liars and recipients, modelling the effect of liars’ models of recipients’ mental states [@oey2023] and showing consistently poor ability of observers to detect lies or pretense in others[@bond2006].
+In contrast, our focus here is on a special kind of pretense, one involving simulations of a counterfactual internal belief state rather than a counterfactual state of the external world, and with no reference to a specific recipient.
+Such simulations are required not only in adversarial settings such as pretense and deceit, but also in teaching and explaining (“*Would I have understood my explanation if I was not familiar with the subject matter?*”), fairness judgments (“*Would I have been so impressed with this candidate if I didn't know they went to Harvard*”), intelligence attribution based on observed behaviour (“*They solved the puzzle faster than it would have taken me to solve it had I not known the solution*”) and legal settings (“*Please ignore this witness’s testimony in your decision, as they were found unreliable*”).
+As such, while our findings should be considered within the broader context of people’s ability to behave in accordance with an imaginary world state, we focus not on the dependence of deceit on models of the world or of other agents, but on its reliance on a model of the self.
+We suggest that this novel perspective may open entirely new avenues for research about self-models and metacognitive knowledge.
 
 Together, our findings reveal a non-trivial capacity for pretending not to know.
-Complementing previous work on cognitive and perceptual hindsight biases, we show that people are capable of accurately simulating diverse aspects of their decision-making processes, although they exhibit systematic shortcomings.
-Further research into these and similar limitations may continue to reveal the simplifications, abstractions, and biases in people's models of their own minds.
-
-<!-- including patterns of active information seeking and knowledge update in the face of new evidence. -->
-
-<!-- We consider two possible explanations to  -->
-
-<!-- Our subjects rose to the challenge.
-Their pretense behavior, although not perfect, closely resembled that of non-pretenders \-\-- not only in gross measures such as total number of clicks, but also in subtler patterns of cell selection and decision latency.
-Our interpretation of this finding is that participants engaged in rich and relatively accurate counterfactual self-simulation: "What would I have done if I didn't have this piece of knowledge?".
--->
-
-<!-- An alternative interpretation is that instead of simulating a counterfactual knowledge state, participants actively suppressed or ignored unwanted knowledge, such that their entire cognitive machinery was available to play the game.
-This does not require any self-modeling or self-simulation, beyond the knowledge that suppressing knowledge can be beneficial for pretending not to know something.
-While we cannot fully rule out this interpretation, we think it is unlikely to explain participants' pretense ability in our data, for two indirect reasons.
-First, suppressing thoughts on demand is notoriously difficult [\@wegner1987paradoxical], and our first-person experience of performing the task is that suppressing knowledge of ship locations is hardly possible.
-Second, when asked how they had performed the task in a debrief question, the responses of most participants were aligned with self-simulation or rule-following (see [exploratory analysis](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship-exploratory-results.html#suppression-or-simulation)).
--->
-
-<!-- In conclusion, we propose that pretense behavior provides a unique opportunity to directly test the richness and accuracy of subjects' models of their own cognitive processes, and find evidence for accurate pretending that surprisingly does not translate to an ability to detect pretense.
-While more work is needed to exclude alternative explanations, we interpret our findings as providing preliminary evidence for a remarkable capacity for counterfactual self-simulation.
--->
-
-# Methods
-
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-## Exp. 1: Battleship
-
-A detailed pre-registration can be accessed at [osf.io/v9zsb](https://osf.io/v9zsb).
-The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019] (protocol sum: 60c270410375e8a192468fc1a0e9c93da60d5e203eb2760b621a8631a26f4c5c; [link to relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/44ef964c0d7d4a8f3b76ec52316a8e624c41a047/experiments/Battleships2/code/index.html#L865-L876)).
-All pre-registered analyses are available [in this link](https://matanmazor.github.io/ignorance/docs/exp.-1-battleship.html).
-
-### Participants
-
-The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.
-`E2.df$subj_id%>%unique()%>%length()` Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
-They were selected based on their acceptance rate (\>95%) and for being native English speakers.
-The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment (`r E2.bonus$bonus%>%min()` - `r E2.bonus$bonus%>%max()` USD, mean = `r E2.bonus$bonus%>%mean()%>%printnum()`).
-
-### Procedure
-
-Participants were first instructed that the experiment, based on the game Battleship, had three parts, and that they could accumulate 'points' that would later translate to a monetary bonus payment.
-They were then presented with a leaderboard of previous players, and given the rules of the game:
-
-> *"In the game Battleship, your task is to sink all ships located in a grid with as few clicks as possible.
-What makes the game difficult is that you can't see the ships; all you can see is a grid of squares, and you have to guess where the ships are.
-To sink a ship, you need to click on all of the squares it is located in.
-If you hit part of a ship, the square will turn red.
-If there is no ship in the square, it will turn blue.*"
-
-We further explained that in this version of the game, ships can touch corners, but their sides can't touch.
-This explanation was accompanied by a visual presentation of legal and illegal ship configurations.
-
-After completing a comprehension question and a practice round, participants completed one 'pretend' and one 'non-pretend' block, each comprising five full games and one half-game (see below for details).
-The order of pretend and non-pretend blocks was counterbalanced between participants.
-The allocation of boards to conditions was randomized between participants such that exactly one board was played in both pretend and non-pretend conditions, and this common board was different for different participants.
-The order of boards within a block was fully randomized, with the exception that half-games were always played last.
-
-#### Non-pretend (normal) games
-
-In non-pretend games (Fig.
-\@ref(fig:design)A), participants aimed to sink two 2-square patrol boats and one 3-square submarine with as few clicks as possible.
-An online counter of the number of clicks was displayed on the screen.
-After each game, feedback was given about the number of clicks and resulting number of points obtained.
-
-#### Pretend games
-
-Participants in pretend games were given the same explanation of Battleship, and played a practice round. However, they were then given an additional instruction:
-
-> *"This time your goal is different.
-In this round, we're going to tell you where the ships are, but **we want you to act like you don't know this information**.
-We've marked the ships' locations with a cross, so you'll know where they are the whole time; but your job is to play the game as if these hints aren't there.
-To see how good you are at this, we're going to compare your games to the games of people who actually had no hints, and see how similar they are.
-We will measure where and when you clicked; if your clicks look similar to people who played like normal (trying to reveal all ships with as few clicks as possible, but without any hints), you'll get bonus points.
-But if your games look different, you won't get these bonus points.
-Your number of clicks in this part will not affect your bonus.
-Only your ability to play like you had no hints.*"
-
-We informed participants that both the location and timing of their cell clicks would be measured.
-After one practice round and one comprehension question, participants played five pretend games (Fig.
-\@ref(fig:design)B), followed by one pretend half-game.
-Each game was followed by a short message, reminding them that a game that looks similar to the games of participants who had no hints would be awarded 10 bonus points.
-
-#### Half-games
-
-In order to directly compare participants' pretend and non-pretend games for identical belief states (genuine or pretended ignorance about where the ships are hidden), participants completed one pretend and one non-pretend game given a partly finished board with the content of 7 cells already revealed (Fig.
-\@ref(fig:design)C).
-We designed our half-games to produce a strong expectation to find a ship in specific cells, but not in others.
-The assignment of half-completed boards to pretend and non-pretend conditions was randomized between participants.
-
-#### Judge trials
-
-```{r judge-game-proportion-Battleship, echo=FALSE, message= F, warning=F}
-E2.judge_np_game_proportion <- E2.click_log %>% 
-  filter(genuine_first & test_part=='nonpretend') %>%
-  group_by(subj_id,test_part,grid_number, genuine_first) %>% 
-  summarise(duration=sum(RT)/1000) %>% 
-  mutate(included = duration<=60) %>%
-  pull(included) %>%
-  mean()
-
-E2.judge_p_game_proportion <- E2.click_log %>% 
-  filter(test_part=='pretend') %>%
-  group_by(subj_id,test_part,grid_number, genuine_first) %>% 
-  summarise(duration=sum(RT)/1000) %>% 
-  mutate(included = duration<=60) %>%
-  pull(included) %>%
-  mean()
-```
-
-In the final part of the experiment, participants observed the games of previous players and tried to determine who were pretenders and who were non-pretenders.
-On each trial, two empty grids were presented side by side, with a smaller grid on top, displaying the hidden positions of ships on the grid (Fig.
-\@ref(fig:design)D).
-The two grids corresponded to the true games of two previous players who played a version of the top grid either as pretenders or as non-pretenders.
-For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
-For both pretend and non-pretend games, only games shorter than one minute (`r (E2.judge_np_game_proportion*100)%>%round()`% of included non-pretend games and `r (E2.judge_p_game_proportion*100)%>%round()`% of pretend games) were presented.
-Judge participants observed a real time replay of the two grids, showing not only where participants clicked, but also when.
-After making a decision, participants were informed whether they would receive the 10 bonus points, or alternatively, whether the pretender would receive them in the event the pretender managed to trick them.
-
-Readers are invited to try a [demo of the experiment](https://matanmazor.github.io/ignorance/experiments/demos/Battleship).
-
-## Hangman
-
-A detailed pre-registration can be accessed at [osf.io/3thry](https://osf.io/3thry).
-The pre-registration was time-locked using cryptographic randomization-based time-locking [@mazor2019].
-Due to an error in the experiment code, time-locking took effect only from player number 221 (batch 3) and on (protocol sum: c4929c7fe33df1b7b52f15c789d98eab30a9cee09a8121807a3c59e28e7430a4; [relevant lines in experimental code](https://github.com/matanmazor/ignorance/blob/f98903f3f2d0e21800497dc65f8d5f823c910c88/experiments/Hangman2/code/index.html#L772-L783)).
-
-### Participants
-
-The research complied with all relevant ethical regulations and was approved by the Institutional Review Board of Johns Hopkins University.
-`E4.df$subj_id%>%unique()%>%length()` Participants were recruited via Prolific (prolific.co) and gave their informed consent prior to their participation.
-They were selected based on their acceptance rate (\>95%) and for being native English speakers.
-The entire experiment took approximately 20 minutes to complete.
-Participants' pay was equivalent to an hourly wage of 9.50 USD, in addition to a bonus payment of 1 USD that was awarded to `r E4.bonus$bonus%>%sum(na.rm=T)` players who earned 100 points or more.
-
-### Procedure
-
-The first instructions screen informed participants that the experiment, based on the game Hangman, had three parts, and that the points they accumulate translate to a monetary bonus payment.
-They were then presented with a leaderboard of previous players.
-Then, the rules of the game were presented:
-
-> *"In the following game, your task is to reveal a hidden word or phrase by guessing letters.
-What makes the game difficult is that you can't see the word; all you can see is a row of squares - a square for each letter.
-Use your mouse to make letter guesses.
-We will have five types of words: body parts, numbers, US states, fruit, and famous people.
-You will start each game with 15 points and lose one point for every guess of a letter that is not in the word.*"
-
-We then explained that "the words in this game are the kind of words that will be familiar to most English-speaking fifth-graders. We didn't pick any strange or particularly difficult words".
-
-Once they responded correctly to a multiple-choice comprehension question ("the goal of the game is to...": "reveal the word with as few letter guesses as possible"), participants played a practice round, revealing the word PIGEON (see Fig.
-\@ref(fig:design)E).
-
-After the main instructions, comprehension question and practice round, participants completed one pretend and one non-pretend block, each followed by one half-game (see below for details).
-The order of pretend and non-pretend blocks was counterbalanced between participants.
-Each block comprised five games played with five out of ten different words, and one half-game.
-The allocation of words to conditions was randomized between participants, with the constraint that both pretend and non-pretend blocks included exactly one word from each category.
-The order of words within a block was randomized, except for the half-game, which was always delivered at the end.
-
-The ten words included two number words (ELEVEN, NINETY SIX), two famous people (DALAI LAMA, TAYLOR SWIFT), two fruits (STRAWBERRY, LEMON), two body parts (TOOTH, HEAD), and two US states (MONTANA, IOWA).
-
-#### Non-pretend games
-
-In non-pretend games, participants revealed a hidden word with as few letter guesses as possible.
-An online counter of the number of points was displayed on the screen, deducting one point for every guess of a letter that is not in the target word.
-After each game, feedback was given about the number of points obtained.
-
-After completing the five games, participants performed one half-game (see below for details).
-
-#### Pretend games
-
-Participants were given the following instructions:
-
-> *"*In the next part of the experiment, you'll play 6 games where you reveal a hidden word by guessing letters.
-
-> However, this time your goal is different.
-
-> In this round, we're going to tell you the word in advance, but **we want you to act like you don't know this information**.
-
-> To see how good you are at this, we're going to compare your games to the games of people who played normally, without knowing what the word was, and see how similar they are.
-We will measure which letters you click and the timing of your guesses; if your clicks look similar to people who played like normal (trying to reveal the word with as few guesses as possible, but without any hints), you'll get bonus points.
-But if your games look different, you won't get these bonus points.
-Your number of clicks in this part will not affect your bonus.
-Only your ability to play like you didn't see the word in advance.*"*
-
-After one practice round, pretending not to know that the hidden word is PIGEON, and one comprehension question ("In this part of the experiment my goal is to...": "play the game as if I don't know what the word is so that I look like someone who had no hints"), participants played five pretend games (Fig.
-\@ref(fig:design)F).
-Each game was preceded by a short message informing subjects about the identity of the target word.
-To start pretending, players were asked to type in the target word on their keyboard.
-The target word remained on the screen, in green letters, until the end of the game.
-After pretending, we reminded players that a game that looks similar to the games of participants who had no hints will be awarded 10 bonus points.
-
-After completing the five games, participants performed one half-game (see below for details).
-
-#### Half-games
-
-In order to directly compare participants' pretend and non-pretend games for identical belief states (true or pretended knowledge about the identity of the word), we asked participants to also complete one pretend and one non-pretend game, given a partly finished game with some letters already guessed (they were told that the computer made these guesses; Fig.
-\@ref(fig:design)G).
-The two half-game words were one fruit: PAPAYA or BANANA, with guessed letters [A, E, I, O, M, T], and one body part: HAND, or HAIR with guessed letters (A, E, O, M, T, H, P).
-The assignment of category (fruit or body part) to condition (pretend and non-pretend), as well as the identity of the target word within each category (e.g., PAPAYA or BANANA), was randomized between participants.
-
-Instructions for the non-pretend half-game were:
-
-> *"For the next game, the computer chose the first letters for you; you can take over from where it left off.
-Your challenge is to complete the game.
-Just like in the previous games, here also you will lose one point for each letter that you guess and is not in the word.*"
-
-Instructions for the pretend half-game were:
-
-> *"For the next game, the computer chose the first letters for you; you can take over from where it left off.
-Just like in the previous games, here also you will know what the word is, but your bonus points will depend on your ability to play as if you didn't know the word.*"
-
-#### Judge trials
-
-```{r judge-game-proportion-Hangman, echo=FALSE, message= F, warning=F}
-E4.judge_np_game_proportion <- E4.click_log %>% 
-  filter(genuine_first & test_part=='nonpretend') %>%
-  group_by(subj_id,test_part,word, genuine_first) %>% 
-  summarise(duration=sum(RT)/1000) %>% 
-  mutate(included = duration<=90) %>%
-  pull(included) %>%
-  mean()
-
-E4.judge_p_game_proportion <- E4.click_log %>% 
-  filter(test_part=='pretend') %>%
-  group_by(subj_id,test_part,word, genuine_first) %>% 
-  summarise(duration=sum(RT)/1000) %>% 
-  mutate(included = duration<=90) %>%
-  pull(included) %>%
-  mean()
-```
-
-In the final part of the experiment, participants observed five games of previous players and determined who had hints and who didn't. Instructions for this part were:
-
-> *"In this third and last part of the experiment, we ask you to be a judge for previous players, and see if you can tell which of the players were shown the word (but acted like they weren't).
-We will show you 5 replays of the games of previous players.
-Your task is to decide whether they played normally or pretended.
-For each game that you get right, you will receive 10 points.
-Good luck!*"
-
-Then, on each judge trial, one game of a previous player was replayed in real time, with the target word presented above.
-For non-pretend games, only games from the group of participants that pretended in the second block (and played normally in the first block) were chosen for presentation in this part.
-For both pretend and non-pretend games, only games shorter than 1.5 minutes (`r (E4.judge_np_game_proportion*100)%>%round()`% of included non-pretend games and `r (E4.judge_p_game_proportion*100)%>%round()`% of pretend games) were presented.
-Judge participants indicated their decision by pressing the P and N keys on their keyboard.
-After making a decision, participants were informed whether they received the 10 points.
-Whenever a pretend game was classified as a non-pretend game, they were informed that the pretender received these 10 points instead of them.
-
-Lastly, participants were asked the following debrief questions:
-
-> *"Did you have a strategy that you used for pretending you did not see the word?
-What was most difficult about pretending?
-How about telling between players who pretenders and who played for real - did you have a strategy for that?*"
-
-And:
-
-> *"We would appreciate it if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data.*"
-
-Readers are invited to try a [demo of the experiment](https://matanmazor.github.io/ignorance/experiments/demos/Hangman).
-
-\newpage
+Complementing previous work on cognitive and perceptual hindsight biases, which traditionally focus on people’s inability to emulate ignorance, we show that people are in fact capable of accurately simulating diverse aspects of their decision-making processes, although they exhibit systematic shortcomings.
+We speculate that these shortcomings are consistent with the simulation of a simplified model of cognition, over and above any suppression of knowledge or sensory input.
+In revealing this powerful capacity, our findings raise many new theoretical questions to which we don’t yet have answers.
+Are there specific aspects of our knowledge, beliefs, or inferences that are harder than others to simulate, and is this related to a lack of metacognitive understanding of these aspects?
+Does pretending not to know rely on explicit, reportable self-knowledge, or on an implicit self-model?
+Is the ability to overcome the curse of knowledge in the context of pretending predictive of the ability to overcome it in communicating information to a naive audience?
+Further research into these and similar limitations may continue to reveal the simplifications, abstractions, and biases in people’s models of their own minds.
 

### 2025-10-09 docs/preregisteredMethodsAndResults.Rmd
model fits merged into one file, saving intermediate data automatically, and beautifying some parts of the paper

@@
                       message=F, sanitize = T, 
-                      fig.width = 10, fig.height = 5)
+                      fig.width = 10, fig.height = 5,
+                      dev = "ragg_png", dpi = 300, fig.width = 6, fig.height = 4)
+
 ```
@@
   'caret', #for cross validation
-  'ggrepel' #for word scatterplots
+  'ggrepel', #for word scatterplots
+  'ragg' # for png
 ), '2022-12-01')
@@
 #                                               "genuine_first"))))
-
+# 
+# 
 # for (row in 1:nrow(E4.click_df)) {
@@
 # }
-
+# 
+# 
 # E4.click_log <- E4.click_log %>%
@@
 
-# save(..experiments/Hangman2/data/E4.click_log, file='E4.click_log.RData')
-
-load('..experiments/Hangman2/data/E4.click_log.RData')
+load('../experiments/Hangman2/data/E4.click_log.RData')
+
 ```
@@
 
+```{r save data, echo-FALSE}
+
+save.image(file = file.path(getwd(), "prereg_results.RData"))
+
+```
+
 \newpage

### 2025-10-21 docs/paper.Rmd
added file readmes, added gitbook version of paper

@@
 classoption       : "man"
-output            : papaja::apa6_word
+output            : bookdown::gitbook
 
@@
 This way, we had sufficient statistical power to compare the two guess distributions.
-We find a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_A_pretend_nonpretend)$table$p.value`; board B: $r=$ `r apa_print(cor_B_pretend_nonpretend)$table$estimate`, $p$ `r apa_print(cor_B_pretend_nonpretend)$table$p.value`; Fig.
+We find a strong correlation between the spatial distributions of pretend and non-pretend guesses (board A: $r=$ `r apa_print(cor_A_pretend_nonpretend)$table$estimate`, $p$ `r printp(cor_A_pretend_nonpretend$p.value, add_equals=T)`; board B: $r=$ `r apa_print(cor_B_pretend_nonpretend)$table$estimate`, $p$ `r printp(cor_B_pretend_nonpretend$p.value, add_equals=T)`; Fig.
 \@ref(fig:decisions)C), confirming that pretenders were sensitive not only to the number of guesses they would have made had they been playing for real, but also to their content.
@@
 If the same prototype of randomness is used by pretenders to determine the number of unsuccessful guesses per game, the two measures should be correlated across participants.
-Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$full_result`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$full_result`).
-This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$full_result`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$full_result`).
+Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`).
+This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`).
 We interpret this as evidence that the reduction in variance reflects miscalibrated beliefs not only about randomness, but also about participants’ behaviour under a counterfactual knowledge state.

### 2025-11-04 docs/paper.Rmd
fix wrong p value, knit to word

@@
 classoption       : "man"
-output            : bookdown::gitbook
+output            : papaja::apa6_word
 
@@
 Crucially, we find the exact opposite pattern: a negative correlation between variability in the number of unsuccessful guesses and game entropy (Battleship: `r apa_print(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$pretend_entropy,E2.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`, Hangman: `r apa_print(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E4.entropy_sd$pretend_entropy,E4.entropy_sd$pretend_sd, method='spearman')$p.value, add_equals=T)`).
-This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`).
+This negative correlation was not observed in non-pretend games (and was even positive for Battleship: `r apa_print(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E2.entropy_sd$nonpretend_entropy,E2.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`; Hangman: `r apa_print(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman'))$estimate`, $p$ `r printp(cor.test(E4.entropy_sd$nonpretend_entropy,E4.entropy_sd$nonpretend_sd, method='spearman')$p.value,add_equals=T)`).
 We interpret this as evidence that the reduction in variance reflects miscalibrated beliefs not only about randomness, but also about participants’ behaviour under a counterfactual knowledge state.
